### 2018-12-23 Metrics1.tex
Committed by TeXstudio

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Econometrics 1. Problem Set 1.
\end{center}
\end{document}

### 2019-01-11 micro3_1.tex
micro3_1

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Microeconomics 3. Problem Set 1.
\end{center}
\section*{Problem 1}
Graphically, argue that if the indifference curves of a consumer is thick (due to absence of local non-satiation), then the conclusion of the first fundamental theorem of welfare economics may fail. That is, there may exist an equilibrium that is not Pareto efficient.
(Hint. If the indifference curve of a consumer is thick, we can take some goods from her,
and give them to the other consumer, without hurting the former.)

\textbf{Solution}

Assume that for some $p$ the equilibrium allocation is $x$ (see Fig. 1). Moreover for the first consumer LNS property of preferences is violated, that is his indifference domain is depicted as a shaded area, while the second consumer has LNS and strictly convex preferences, which are represented by indifference curves. The depicted equilibrium point $x$ will not be a Pareto efficient because the first consumer is indifferent between points $x$ and $x'$, however the second strictly prefers $x'$ to $x$.
\newpage
\section*{Problem 2}
Normalize the price of good 1 by letting $p_1 = 1$. As a function of the parameter $\alpha \in (0, 1/2)$, compute the equilibrium value of $p_2$ and the equilibrium allocation for the
following utility functions and endowments:
\begin{align*}
u_1(x_{11}, x_{21}) = (x_{11})^{\alpha}(x_{21})^{1-\alpha}, \omega_1 = (2, 1)\\
u_2(x_{12}, x_{22}) = (x_{12})^{2\alpha}(x_{22})^{1-2\alpha}, \omega_2 = (2, 1)\\
\end{align*}

\textbf{Solution}

Assume $p_1 = 1$. To find the Walrasian demand functions we need to solve the following optimization problems:
\begin{align*}
\underset{x_{11} \ge 0, x_{21}\ge 0}{\max}\ (x_{11})^{\alpha}(x_{21})^{1-\alpha}\ \ s.t.\ \ x_{11} + p_2x_{21} \le 2 + p_2\\
\underset{x_{12} \ge 0, x_{22} \ge 0}{\max}\ (x_{12})^{2\alpha}(x_{22})^{1-2\alpha}\ \ s.t.\ \ x_{12} + p_2x_{22} \le 2 + p_2\\
\end{align*}
Since $\alpha \in (0, 1/2)$ it is a well-known \textit{Cobb-Douglas-utility-linear-budget-constraint-interrior-solution} optimization problem and its solutions are:
\begin{align*}
x_1 = \begin{pmatrix}
\alpha (2 + p_2)\\
\\
(1-\alpha)\frac{2+p_2}{p_2}
\end{pmatrix}\\
x_2 = \begin{pmatrix}
2\alpha (2 + p_2)\\
\\
(1-2\alpha)\frac{2+p_2}{p_2}
\end{pmatrix}\\
\end{align*}
Using market clearing condition $x_{11} + x_{12} = 4$ one can get
\begin{align*}
2\alpha + \alpha p_2 + 4\alpha + 2\alpha p_2 = 4\\
p_2 = \frac{4}{3\alpha} - 2
\end{align*}
Thus, the optimal allocation should be: 
\begin{align*}
x_1^* = \begin{pmatrix}
\frac{4}{3}\\
\\
\frac{4 - 4\alpha}{4 - 6\alpha}
\end{pmatrix}\\
x_2^* = \begin{pmatrix}
\frac{8}{3}\\
\\
\frac{4 - 8\alpha}{4 - 6\alpha}
\end{pmatrix}\\
\end{align*}
As it can be seen $\frac{4}{3} + \frac{8}{3} = 4 = \omega_{11} + \omega_{12}, \frac{4-4\alpha}{4 - 6\alpha} + \frac{4 - 8 \alpha}{4 - 6\alpha} = 2 = \omega_{21} + \omega_{22}$. That is, this allocation is non-wasteful.
\section*{Problem 3}
Normalize the price of good 1 by letting $p_1 = 1$. As a function of the parameter $a > 0$,
compute the equilibrium value of $p_2$ and the equilibrium allocation for the following utility
functions and endowments:
\begin{align*}
u_1(x_{11}, x_{21}) = x_{11} + x_{21},\ \omega_1 = (1, 1)\\
u_2(x_{12}, x_{22}) = \sqrt{x_{12} x_{22}},\ \omega_2 = (0, a)
\end{align*}

\textbf{Solution}

Assume $p_1 = 1$. To find the Walrasian demand functions we need to solve the following optimization problems:
\begin{align*}
\underset{x_{11} \ge 0, x_{21}\ge 0}{\max}\ x_{11} + x_{21}\ \ s.t.\ \ x_{11} + p_2x_{21} \le 1 + p_2\\
\underset{x_{12} \ge 0, x_{22} \ge 0}{\max}\ (x_{12})^{\frac{1}{2}}(x_{22})^{\frac{1}{2}}\ \ s.t.\ \ x_{12} + p_2x_{22} \le ap_2\\
\end{align*}
Obviously solutions are:
\begin{align*}
&x_1 = \begin{cases}
\begin{pmatrix}
0\\
\frac{1+p_2}{p_2}
\end{pmatrix}, p_2 < 1\\
\begin{pmatrix}
1+p_2\\
0
\end{pmatrix}, p_2 > 1\\
\left\{(x_{11}, x_{21}):x_{11} + x_{21} = 2 \right\}
\end{cases}\\
&x_2 = \begin{pmatrix}
\frac{p_2a}{2}\\
\\
\frac{a}{2}
\end{pmatrix}
\end{align*}
If $p_2 > 1, a > 0$ then $0 + \frac{a}{2} < 1 + a$ that is market clearing condition is violated, hence $p_2 \le 1$. In case $p_2 < 1$:
\begin{align*}
\frac{p_2a}{2} = 1 \to p_2 = \frac{2}{a}, a > 2
\end{align*}
and the allocation should be:
\begin{align*}
x_1^* = \begin{pmatrix}
0\\
1 + \frac{a}{2}
\end{pmatrix}\\
x_2^* = \begin{pmatrix}
1\\
\frac{a}{2}
\end{pmatrix}
\end{align*}
If $p_2 = 1$ then
\begin{align*}
x_1^* = \begin{pmatrix}
1 - \frac{a}{2}\\
1 + \frac{a}{2}
\end{pmatrix}\\
x_2^* = \begin{pmatrix}
\frac{a}{2}\\
\frac{a}{2}
\end{pmatrix}
\end{align*}
Of course, $1 - \frac{a}{2} \ge 0 \iff a \le 2$. Thus, the optimal allocation is:
\begin{align*}
(x_1^*, x_2^*) = \begin{cases}
\begin{cases}
\begin{pmatrix}
0\\
1 + \frac{a}{2}
\end{pmatrix}\\
\begin{pmatrix}
1\\
\frac{a}{2}
\end{pmatrix}
\end{cases}, p_2 < 1, a > 2\\
\begin{cases}
\begin{pmatrix}
1 - \frac{a}{2}\\
1 + \frac{a}{2}
\end{pmatrix}\\
\begin{pmatrix}
\frac{a}{2}\\
\frac{a}{2}
\end{pmatrix}
\end{cases}, p_2 = 1, 0 < a \le 2
\end{cases}
\end{align*}
\section*{Problem 4}
In a Robinson Crusoe economy, find the equilibrium allocation, relative prices, and the
firm profit for $u(x_1, x_2) = \ln x_1 + \ln x_2, f(z) = z^{1/4}$ and $L = 5$.

\textbf{Solution}

As the owner of the firm the consumer is solving the following optimization problem, taking $w, p$ as given:
\begin{align*}
\underset{z \ge 0}{\max}\ pz^{1/4} - wz\\
\text{FOC}:\ \frac{p}{4z^{3/4}} - w = 0\\
z^* = \left(\frac{p}{4w}\right)^{4/3}
\end{align*}
that means that the profit function is
\begin{align*}
\pi(p, w) = p \left(\frac{p}{4w}\right)^{1/3} - w\left(\frac{p}{4w}\right)^{4/3} = \left(\frac{1}{4^{1/3}} - \frac{1}{4^{4/3}}\right) \frac{p^{4/3}}{w^{1/3}} = \frac{3}{4^{4/3}}\frac{p^{4/3}}{w^{1/3}}
\end{align*}\\
Moreover, as a consumer the agent is solving the following optimization problem:
\begin{align*}
\underset{0 \le x_1 \le 5,\ x_2 \ge 0}{\max}\ \ln x_1 + \ln x_2\ s.t.\ \ wx_1 + px_2 \le 5w + \pi(p, w)
\end{align*}
Applying increasing transformation the problem can be reduced to a well known Cobb-Douglas optimization problem which has a solution:
\begin{align*}
x_1 = \frac{5w + \frac{3}{4^{4/3}}\frac{p^{4/3}}{w^{1/3}}}{2w}\\
x_2 = \frac{5w + \frac{3}{4^{4/3}}\frac{p^{4/3}}{w^{1/3}}}{2p}
\end{align*}
The market clearing condition should be satisfied, that is
\begin{align*}
x_2 = f(z^*)\\
x_1 = 5 - z^*
\end{align*}
Thus, 
\begin{align*}
\frac{5w + \frac{3}{4^{4/3}}\frac{p^{4/3}}{w^{1/3}}}{2p} &= \left(\frac{p}{4w}\right)^{1/3}\\
5w + \frac{3}{4^{4/3}}\frac{p^{4/3}}{w^{1/3}} &= 2 \frac{p^{4/3}}{4^{1/3}w^{1/3}}\\
\frac{w}{p} &= \frac{1}{4}
\end{align*}
\end{document}

### 2019-01-11 Metrics1.tex
Committed by TeXstudio
Metrics1

@@
 \end{center}
+\section*{Problem 1}
+ Suppose your calculate estimates for $\beta_0$ and $\beta_1$ by finding the solution to the following minimization problem:
+ \begin{align*}
+ \underset{b_0, b_1}{\min}\ \sum_{i=1}^n \text{exp} \left\{(y_i - b_0 - b_1x_i)^2\right\}
+ \end{align*}
+ Write down first-order conditions for the estimates.
+
+ \textbf{Solution}
+
+ \begin{align*}
+ \frac{\partial}{\partial b_0}: -2 \sum_{i=1}^n (y_i - b_0 - b_ix_i)e^{(y_i - b_0 - b_ix_i)^2} = 0\\
+ \frac{\partial}{\partial b_1}: -2 \sum_{i=1}^n x_i(y_i - b_0 - b_ix_i)e^{(y_i - b_0 - b_ix_i)^2} = 0
+ \end{align*}
+ \section*{Problem 2}
+ In the simple linear regression model $y = \beta_0 + \beta_1x + u$, suppose that $\expect (u) \neq 0$.
+Letting $a_0 = \expect(u)$, show that the model can always be rewritten with the same slope, but a new
+ intercept and error, where the new error has a zero expected value.
+
+ \textbf{Solution}
+ Assume $u' = a_0 - u$. Then $\expect(u') = a_0 - a_0 = 0$ and moreover
+ \begin{align*}
+ &y = \beta_0 + \beta_1 x + a_0 - u' = \beta_0' + \beta_1 x + u''\\
+ &\text{where}\ \beta_0' = \beta_0 + a_0\\
+ &\ \ \ u'' = -u'\\
+ &\text{and}\ \ \expect(u'') = -\expect(u') = 0
+ \end{align*}
+ Q.E.D.
+ \section*{Problem 3}
+ Consider the standard simple linear regression model $y = \beta_0 + \beta_1 x + u$
+ When $n = 3$, is it possible that the data point with maximal value of $y$ is located below the
+ OLS regression line? If answer is yes, provide an example, if no, provide a proof.
+
+ \textbf{Solution}
+
+ Yes, it is possible. Assume $X = (1, 2, 5)$ and corresponded $Y = (2, 3, 4)$ then
+ \begin{align*}
+ b_1 = \frac{12}{26}\\
+ b_0 = \frac{23}{13}
+ \end{align*}
+ and
+ \begin{align*}
+ \frac{12}{26} \cdot 5 + \frac{23}{13} = \frac{53}{13} \approx 4.08 > 4
+ \end{align*}
+ which means that the regression line is lying above the point (5, 4).
+
+ \section*{Problem 4}
+  Consider the following relation:
+\begin{align*}
+y = \sqrt{x} + u
+\end{align*}
+ where $x$ is uniformly distributed on segment $[0; 1]$ and error term $u$ has zero mean
+ conditional on $x$. A random sample of size $n$ is collected: $\left\{x_i, y_i\right\}_{i=1}^n$. Some econometrician
+ performs OLS estimation based on a simple linear model 
+ \begin{align*}
+	y = \beta_0 + \beta_1x + e
+\end{align*}
+ Based on her estimates $\hat{\beta_0}, \hat{\beta_1}$, the econometrician constructs the predictor for the conditional mean of the dependent variable, i.e.
+for $\expect(y|x)$, as $\widehat{\expect(y|x)} = \hat{\beta_0} + \hat{\beta_1} x$.
+ \begin{enumerate}[a.]
+ \item Calculate the expected value of y.
+ \item Find the expected values of $\hat{\beta_0}, \hat{\beta_1}$ conditional on realized values of $x$.
+ \end{enumerate}
+
+\textbf{Solution}
+\begin{enumerate}[a.]
+	\item 
+	\begin{align*}
+	\expect[\sqrt{x}] = \int_{0}^1 \sqrt{x}dx = \frac{2}{3}\\
+	\expect[u] = \expect[\expect[u|X]] = \expect[0] = 0\\
+	\expect[y] = \frac{2}{3} + 0 = \frac{2}{3}
+	\end{align*}
+	\item
+	\begin{align*}
+	\expect[\hat{\beta_1}|x_1, \dots, x_n] = \expect \left[\frac{\sum_{i=1}^n (x_i - \bar{x})(y_i - \bar{y})}{\sum_{i=1}^n (x_i - \bar{x})^2}\bigg|x_1, \dots, x_n\right] =\\
+	= \expect \left[\frac{\sum_{i=1}^n (x_i - \bar{x})(\sqrt{x_i} + u_i - \overline{\sqrt{x}} -  \bar{u})}{\sum_{i=1}^n (x_i - \bar{x})^2}\bigg|x_1, \dots, x_n\right]
+	\end{align*}
+	\end{enumerate}
 \end{document}

### 2019-01-12 Metrics1.tex
Metrics1

@@
 	\end{align*}
-	\item
+	\item Denote $\overline{\sqrt{x}} = \frac{1}{n}\sum_{i=1}^n \sqrt{x_i}$, then
 	\begin{align*}
@@
 	\end{align*}
+	Since
+	\begin{align*}
+	\expect \left[ \sum_{i=1}^n (x_i - \bar{x})(u_i - \bar{u})\bigg|x_1, \dots, x_n\right] = \expect \left[\sum_{i=1}^n (x_i - \bar{x})u_i\bigg|x_1, \dots, x_n\right]
+	\end{align*}
+	assuming independence of the sample, for each $i$
+	\begin{align*}
+	\expect[(x_i - \bar{x})u_i|x_1, \dots, x_n] = \expect[(x_i - \bar{x})u_i|x_i] = 0
+	\end{align*}
+	Thus, 
+	\begin{align*}
+	\expect[\hat{\beta_1}|x_1, \dots, x_n] = \frac{\sum_{i=1}^n (x_i - \bar{x})(\sqrt{x_i} - \overline{\sqrt{x}})}{\sum_{i=1}^n (x_i - \bar{x})^2}
+	\end{align*}
+	Similarly,
+	\begin{align*}
+	\expect[\hat{\beta_0}|x_1, \dots, x_n] = \expect[\bar{y} - \hat{\beta_1}\bar{x}|x_1, \dots, x_n] = \expect[\overline{\sqrt{x}} + \bar{u} - \hat{\beta_1}\bar{x}|x_1, \dots, x_n] = \\
+	=\overline{\sqrt{x}} + \expect[\bar{u}|x_1, \dots, x_n] - \bar{x}\expect[\hat{\beta_1}|x_1, \dots, x_n] = \overline{\sqrt{x}} - \bar{x} \cdot \frac{\sum_{i=1}^n (x_i - \bar{x})(\sqrt{x_i} - \overline{\sqrt{x}})}{\sum_{i=1}^n (x_i - \bar{x})^2}
+	\end{align*}
+	\end{enumerate}
+\section*{Problem 5}
+ Using R, compute the sample mean and standard deviation of \textit{ahe} (average hourly earnings),
+\textit{yrseduc} (years of education), and \textit{female}.
+
+\textbf{Solution}
+
+\begin{align*}
+\text{mean(ahe)} = 15.19042\\
+\text{sd(ahe)} = 8.027257\\
+\text{mean(yrseduc)} = 13.46549\\
+\text{sd(yrseduc)} = 2.479692\\
+\text{mean(female)} = 0.4385083\\
+\text{sd(female)} = 0.49627
+\end{align*}
+\section*{Problem 6}
+ Estimate a regression of \textit{ahe} on \textit{yrseduc}.
+ \begin{enumerate}
+ 	\item What is the coefficient on \textit{yrseduc}? Explain in words what it means. Is the numerical
+value of your estimate large or small in an economic (real-world) sense?
+\item Graph the data points and the estimated regression line.
+\item Is the slope coefficient statistically significantly different from zero? Show how you reach
+this conclusion. Use heteroskedasticity-robust standard errors to answer this question.
+\item Report the 95\% confidence interval for $\beta_1$, the slope of the population regression line.
+Use heteroskedasticity-robust standard errors to answer this question.
+\item  What is the $R^2$ of this regression? What does this mean?
+\item  Compute the correlation coefficient between ahe and yrseduc, and compare its square to
+the $R^2$. How are the correlation coefficient and the $R^2$
+related?
+\item What is the root mean squared error of the regression? What does this mean?
+\item Based on your graph from (b), does the error term appear to be homoskedastic or
+heteroskedastic?
+
+\end{enumerate}
+
+\textbf{Solution}
+\begin{enumerate}
+	\item Coefficient on \textit{yrseduc} is 1.32186. That means that on each additional years of education the average size of the per-hour salary increases by 1.32186 dollars/hour.
+	\item \begin{figure}[h]
+		\centering
+		\includegraphics[width=0.7\textwidth]{Rplot}
+		\caption{Regression line and data points}
+	\end{figure}
 \end{enumerate}

### 2019-01-13 Metrics1.R
Metrics1.R

@@ -1,0 +1,30 @@
+df <- read.csv("cps99_ps1.csv")
+#Problem 5
+
+ahe_mean <- mean(df$ahe)
+ahe_sd <- sd(df$ahe)
+yrseduc_mean <- mean(df$yrseduc)
+yrseduc_sd <- sd(df$yrseduc)
+female_mean <- mean(df$female)
+female_sd <- sd(df$female)
+
+#Problem 6
+#1
+fit <- lm(ahe ~ yrseduc, df)
+summary(fit)
+
+#2
+library(ggplot2)
+test <- function(x) {1.32186 * x -2.60905}
+
+ggplot(df, aes(yrseduc, ahe)) + geom_point() + geom_smooth(method='lm', se = F)
+
+#3
+library(sandwich)
+library(lmtest)
+coeftest(fit, vcov = vcovHC(fit, type = "HC1"))
+
+#4
+confint(fit, vcov = vcovHC(fit, type = "HC1"))
+
+

### 2019-01-13 Metrics1.tex
Metrics1

@@
 \usepackage{amssymb}
+\usepackage{float}
 \usepackage[section]{placeins}
@@
  Estimate a regression of \textit{ahe} on \textit{yrseduc}.
- \begin{enumerate}
+ \begin{enumerate}[a.]
  	\item What is the coefficient on \textit{yrseduc}? Explain in words what it means. Is the numerical
@@
 \textbf{Solution}
-\begin{enumerate}
+
+The results of regression are depicted below
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.7\textwidth]{lm}
+	\caption{Results of regression}
+\end{figure}
+\begin{enumerate}[a.]
 	\item Coefficient on \textit{yrseduc} is 1.32186. That means that on each additional years of education the average size of the per-hour salary increases by 1.32186 dollars/hour.
@@
 	\end{figure}
+\newpage
+\item Using coeftest one can get:
+ \begin{figure}[h]
+	\centering
+	\includegraphics[width=0.7\textwidth]{coeftest}
+	\caption{t-test of coefficients}
+\end{figure}
+
+that is, the slope significantly differs from zero.
+\item Using confint one can get:
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.6\textwidth]{confint}
+	\caption{Confidence intervals for coefficients}
+\end{figure}
+
+that is, the 95\% confidence interval for the slope is (1.227613, 1.416104).
+\item $R^2$ is equal to 0.1667.
+This number is rather close to zero that means that regression hardly explains the actual interaction between variables (this conclusion can be drawn just by the glance at the plot on the Fig.
+2).
+\item Correlation coefficient between \textit{ahe} and \textit{yrseduc} is equal to 0.4083341, and 
+\begin{align*}
+0.4083341^2 = 0.1667 = R^2
+\end{align*}
+For the regression with the single regressor the square of the correlation coefficient is exactly equal to $R^2$.
+\item 
+\begin{align*}
+\text{RMSE} = \sqrt{\frac{1}{n}\sum_{i=1}^n u_i^2} = 7.326572
+\end{align*}
+where $u_i = \hat{y_i} - y_i$.
+\item As for me the error term is in fact heteroskedastic.
+\end{enumerate}
+
+\section*{Problem 7}
+Estimate a regression of \textit{ahe} on \textit{female}.
+\begin{enumerate}[a.]
+\item What is the coefficient on \textit{female}? Explain in words what this means. Is the numerical
+value of your estimate large or small in an economic (real-world) sense?
+\item Test the hypothesis that average hourly earnings are the same for male and female
+workers, against the alternative that they differ, at the 5\% significance level. Make sure to use
+heteroskedasticity-robust standard errors to answer this question.
+\item Compute the sample average of \textit{ahe} for women, and the sample average of \textit{ahe} for men;
+from this compute an estimate of the “gender gap” in earnings; and construct the t-statistic
+testing the hypothesis that the gender gap is zero. Make sure to use heteroskedasticity-robust
+standard errors to answer this question. Compare your results to those in parts (a) and (b).
+
+\end{enumerate}
+
+\textbf{Solution}
+
+Results of regression are depicted below
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\textwidth]{lm_female}
+	\caption{Results of regression}
+\end{figure}
+
+\begin{enumerate}[a.]
+	\item Coefficient on \textit{female} is -3.2026. That means that average per-hour salary falls by 3.2026 dollars per hour for women compare with the one for men.
+	\item If average hourly earnings for women and men are equal then the slope coefficient should be equal to zero.
+Hence, let us test the hypothesis whether the slope coefficient is equal to zero.
+Using coeftest one can get:
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.6\textwidth]{coeftest_female}
+		\caption{Test of coefficients}
+	\end{figure}
+
+	that is, on the significance level of 5\% null is rejected that is, average hourly earnings are actually not the same for men and for women.
+	\item Means for women's $\bar{w}$ and men's $\bar{m}$ average hourly earnings are:
+	\begin{align*}
+	\bar{w} = 13.3922\\
+	\bar{m} = 16.59478\\
+	\end{align*}
+	and
+	\begin{align*}
+	s_w = 6.916653\\
+	s_m = 8.539754\\
+	\end{align*}
+	hence assuming normality of data
+	\begin{align*}
+	t = \frac{\bar{w} - \bar{m}}{\sqrt{\frac{s_m^2}{n_m} + \frac{s_w^2}{n_w}}}
+	\end{align*}
+	then using t.test() one can get the following results:
+		\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.6\textwidth]{t_test}
+		\caption{Results of the t-test for testing the equality of means in two groups}
+	\end{figure}
+that is we can reject the null hypothesis which means that there is in fact a "gender gap". This conclusion is fully consistent with those were gotten in (a) and (b).
 \end{enumerate}

### 2019-01-13 Metrics1.R
Metrics1

@@ -9,22 +9,40 @@ female_mean <- mean(df$female)
 female_sd <- sd(df$female)
 
 #Problem 6
-#1
+#a
 fit <- lm(ahe ~ yrseduc, df)
 summary(fit)
 
-#2
+#b
 library(ggplot2)
 test <- function(x) {1.32186 * x -2.60905}
 
 ggplot(df, aes(yrseduc, ahe)) + geom_point() + geom_smooth(method='lm', se = F)
 
-#3
+#c
 library(sandwich)
 library(lmtest)
 coeftest(fit, vcov = vcovHC(fit, type = "HC1"))
 
-#4
+#d
 confint(fit, vcov = vcovHC(fit, type = "HC1"))
 
+#f
+cor(df$ahe, df$yrseduc)
 
+#g
+sqrt(mean(fit$residuals^2))
+
+#Problem 7
+fit2 <- lm(ahe ~ female, df)
+summary(fit2)
+
+#b
+coeftest(fit2, vcov = vcovHC(fit2, type = "HC1"))
+
+#c
+male_mean <- mean(df$ahe[df$female == 0])
+female_mean <- mean(df$ahe[df$female == 1])
+sd_w = sd(df$ahe[df$female == 1])
+sd_m = sd(df$ahe[df$female == 0])
+t.test(df$ahe[df$female == 1], df$ahe[df$female == 0])

### 2019-01-14 Metrics1.tex
Metrics1

@@
 \begin{enumerate}[a.]
-	\item Coefficient on \textit{yrseduc} is 1.32186. That means that on each additional years of education the average size of the per-hour salary increases by 1.32186 dollars/hour.
+	\item Coefficient on \textit{yrseduc} is 1.32186.
+That means that on each additional years of education the average size of the per-hour salary increases by 1.32186 dollars/hour.
+Or put it another way, we predict that ceteris paribus each additional year of education increases average hours earnings by 1.32186 dollars per hour.
+
+	If a standard working week lasts 40-45 hours that means that in average each additional year of education gives about $4 \cdot 45 \cdot 1.32186 \approx \$238$ additional dollars per month which is a rather reasonable number, I suppose.
 	\item \begin{figure}[h]
@@
 \begin{enumerate}[a.]
-	\item Coefficient on \textit{female} is -3.2026. That means that average per-hour salary falls by 3.2026 dollars per hour for women compare with the one for men.
+	\item Coefficient on \textit{female} is -3.2026. That means that average per-hour salary falls by 3.2026 dollars per hour for women compare with the one for men. The number is also rather reasonable.
 	\item If average hourly earnings for women and men are equal then the slope coefficient should be equal to zero.

### 2019-01-14 Metrics1.R
Metrics1

@@ -15,8 +15,6 @@ summary(fit)
 
 #b
 library(ggplot2)
-test <- function(x) {1.32186 * x -2.60905}
-
 ggplot(df, aes(yrseduc, ahe)) + geom_point() + geom_smooth(method='lm', se = F)
 
 #c

### 2019-01-18 micro3_2.tex
Micro

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Microeconomics 3. Problem Set 2.
\end{center}
\section*{Problem 1}
Suppose that as a function of factor prices $w = (w_1, w_2) >> 0$, the firms' cost for producing
one unit of output are as follows:
\begin{align*}
c_1(w) = \left(\frac{w_1}{3/4}\right)^{3/4}\left(\frac{w_2}{1/4}\right)^{1/4}\\
c_1(w) = \left(\frac{w_1}{1/4}\right)^{1/4}\left(\frac{w_2}{3/4}\right)^{3/4}
\end{align*}
Let $p_1 = 4$ and $p_2 = \frac{4}{\sqrt{3}}$ be the unit price of the firms' outputs, respectively. Furthermore, assume that the total endowment of the two factors are given by $\bar{z_1} = 1 + \frac{1}{3\sqrt{3}}$ and $\bar{z_2} = 1+\sqrt{3}$. Find the factor prices and allocation in an \textit{interior} equilibrium.

\textbf{Solution}

Since the solution should be interior that means that $c_1(w) = p_1, c_2(w) = p_2$. Hence, 
\begin{align*}
\frac{c_1(w)}{c_2(w)} = \frac{p_1}{p_2}\\
\left(\frac{w_1}{w_2}\right)^{1/2} = \sqrt{3}\\
w_1 = 3w_2\\
\end{align*}
Plugging it into $c_1(w) = p_1$ one can obtain:
\begin{align*}
4w_2 = 4 \to w_2 = 1, w_1 = 3\\
\end{align*}
Now let us find demand vectors for factors:
\begin{align*}
a_1(w) = \begin{pmatrix}
\frac{\partial c_1(w)}{\partial w_1}\\
\frac{\partial c_1(w)}{\partial w_2}
\end{pmatrix} = \begin{pmatrix}
\left(\frac{3w_2}{w_1}\right)^{1/4}\\
\left(\frac{w_1}{3w_2}\right)^{3/4}
\end{pmatrix} = \begin{pmatrix}
1\\
1
\end{pmatrix}\\
a_2(w) = \begin{pmatrix}
\frac{\partial c_2(w)}{\partial w_1}\\
\frac{\partial c_2(w)}{\partial w_2}
\end{pmatrix} = \begin{pmatrix}
\left(\frac{w_2}{3w_1}\right)^{3/4}\\
\left(\frac{3w_1}{w_2}\right)^{1/4}
\end{pmatrix} = \begin{pmatrix}
\dfrac{1}{\sqrt{27}}\\
\\
\sqrt{3}
\end{pmatrix}\\
\end{align*}
Since
\begin{align*}
\frac{a_{11}}{a_{21}} = 1 > \frac{\bar{z_1}}{\bar{z_2}} = \frac{1+3\sqrt{3}}{3\sqrt{3}(1 + \sqrt{3})} = \frac{8 - 2\sqrt{3}}{6\sqrt{3}} > \frac{a_{12}}{a_{22}} = \frac{1}{9}
\end{align*}
that means that solution is indeed interior.
To find the optimal output level $(q_1^*, q_2^*)$ we need the market clearing condition to be satisfied, i.e. we need to solve the following linear system:
\begin{align*}
q_1 \begin{pmatrix}
1\\
1
\end{pmatrix} + q_2 \begin{pmatrix}
\frac{1}{\sqrt{27}}\\
\sqrt{3}
\end{pmatrix} = \begin{pmatrix}
1 + \frac{1}{3\sqrt{3}}\\
1 + \sqrt{3}
\end{pmatrix}
\end{align*}
and the solution is obviously $q_2^* = 1, q_1^* = 1$. Thus, prices for factors are $w = (3, 1)$ and the equilibrium factors' allocation is:
\begin{align*}
\begin{pmatrix}
1\\
1
\end{pmatrix}, \begin{pmatrix}
\frac{1}{3\sqrt{3}}\\
\sqrt{3}
\end{pmatrix}
\end{align*}
\end{document}

### 2019-01-18 micro3_2.tex
Micro

@@
 \end{align*}
+
+\section*{Problem 2}
+Suppose that as a function of factor prices $w = (w_1, w_2) >> 0$, the firms' cost for producing
+one unit of output are as follows:
+\begin{align*}
+c_1(w) = \left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3}\\
+c_2(w) = \left(\frac{w_1}{1/2}\right)^{1/2}\left(\frac{w_2}{1/2}\right)^{1/2}
+\end{align*}
+Let
+\begin{align*}
+p_1 = \frac{6}{2^{\frac{2}{3}}}\text{ and } p_2 = 3
+\end{align*}
+be the unit price of the firms' outputs, respectively. Furthermore, assume that the total
+endowment of the two factors are given by $\bar{z_1} = 2$ and $\bar{z_2} = 1$. Find the factor prices and
+allocation in an equilibrium in which firm 2 shuts down. Verify that the firms' decisions are
+truly optimal given the factor prices you determined.
+
+\textbf{Solution}
+
+As in previous exercise:
+\begin{align*}
+\frac{c_1(w)}{c_2(w)} = \frac{p_1}{p_2}\\ \frac{6}{2^{\frac{2}{3}}}\left(\frac{w_1}{w_2}\right)^{1/6} = \frac{6}{2^{\frac{2}{3}}3}\\
+w_2 = 3^6w_1
+\end{align*}
+Now let us find the demand for factors:
+\begin{align*}
+a_1(w) = \begin{pmatrix}
+\frac{\partial c_1(w)}{\partial w_1}\\
+\frac{\partial c_1(w)}{\partial w_2}
+\end{pmatrix} = \begin{pmatrix}
+\left(\frac{2w_2}{w_1}\right)^{1/3}\\
+\left(\frac{w_1}{2w_2}\right)^{2/3}
+\end{pmatrix} = \begin{pmatrix}
+2^{\frac{1}{3}}9\\
+\frac{1}{2^{\frac{2}{3}}81}
+\end{pmatrix}\\
+a_2(w) = \begin{pmatrix}
+\frac{\partial c_2(w)}{\partial w_1}\\
+\frac{\partial c_2(w)}{\partial w_2}
+\end{pmatrix} = \begin{pmatrix}
+\left(\frac{w_2}{w_1}\right)^{1/2}\\
+\left(\frac{w_1}{w_2}\right)^{1/2}
+\end{pmatrix} = \begin{pmatrix}
+27\\
+\\
+\frac{1}{27}
+\end{pmatrix}\\
+\frac{a_{11}}{a_{21}} = 2\cdot 3^6 > \frac{a_{12}}{a_{22}} = 3^6 > \frac{\bar{z_1}}{\bar{z_2}} = 2
+\end{align*}
+That means that in the equilibrium firm 1 should shut down.
+
+\section*{Problem 3}
+Suppose that for $w = (w_1, w_2) >> 0$, \begin{align*}
+a_{11} = \frac{2}{3w_1}\left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3}\\
+a_{21} = \frac{2}{3w_2}\left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3}\\
+a_{12} = 
+\end{align*}
 \end{document}

### 2019-01-18 micro3_2.tex
Micro

@@
 a_{11} = \frac{2}{3w_1}\left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3}\\
-a_{21} = \frac{2}{3w_2}\left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3}\\
-a_{12} = 
+a_{21} = \frac{1}{3w_2}\left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3}\\
+a_{12} = \frac{1}{3w_1}\left(\frac{w_1}{1/3}\right)^{1/3}\left(\frac{w_2}{2/3}\right)^{2/3}\\
+a_{22} = \frac{2}{3w_2}\left(\frac{w_1}{1/3}\right)^{1/3}\left(\frac{w_2}{2/3}\right)^{2/3}\\
+\end{align*}
+where $a_j(w) := (a_{1j}(w); a_{2j}(w))$ is the conditional factor demand of firm $j$ that corresponds
+to one unit of output. Assume $p_1 = p_2 = \frac{1}{2}$. Find the equilibrium factor allocation and
+prices as a function of factor endowments $\bar{z_1}; \bar{z_2} > 0$. (Be sure to cover boundary equilibria,
+if any.)
+
+\textbf{Solution}
+
+Assume $c_1(w), c_2(w)$ are cost functions of firms 1 and 2 for producing 1 unit of output. Then
+\begin{align*}
+\frac{\partial c_1(w)}{\partial w_1} = a_{11} = \frac{2}{3w_1}\left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3} \to c_1(w) =\left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3} + \varphi(w_2)\\
+\end{align*}
+where $\varphi(w_2)$ is an arbitrary function of $w_2$.
+\begin{align*}
+\frac{\partial c_1(w)}{\partial w_2} = \frac{1}{3w_2}\left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3} + \frac{d\varphi(w_2)}{dw_2} = a_{21} \to \varphi(w_2) = const
+\end{align*}
+if there are no fixed costs then of course $\varphi(w_2) = 0$. Thus, 
+\begin{align*}
+c_1(w) = \left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3}
+\end{align*}
+Similarly, 
+\begin{align*}
+c_2(w) = \left(\frac{w_1}{1/3}\right)^{1/3}\left(\frac{w_2}{2/3}\right)^{2/3}
+\end{align*}
+Suppose firstly that the equilibrium is interior. Hence
+\begin{align*}
+\frac{c_1(w)}{c_2(w)} &= \frac{p_1}{p_2}\\
+\left(\frac{w_1}{w_2}\right)^{1/3} = 1 &\to w_1 = w_2\\
+c_1(w) = p_1 &\to w_1^* = w_2^* = \frac{1}{2^{1/3}3}
+\end{align*}
+Now let us derive the inferior solution condition.
+\begin{align*}
+\frac{a_{11}}{a_{21}} &= 2\\
+\frac{a_{12}}{a_{22}} &= \frac{1}{2}\\
+\end{align*}
+\begin{align}\label{eq1}
+2 > \frac{\bar{z_1}}{\bar{z_2}} > \frac{1}{2} &\to \frac{\bar{z_2}}{2} < \bar{z_1} < 2\bar{z_2}
+\end{align}
+Market clearing conditions:
+\begin{align*}
+q_1 a_1(w) + q_2 a_2(w) = \bar{z}\\
+q_1 \begin{pmatrix}
+2^{1/3}\\
+\frac{1}{2^{2/3}}
+\end{pmatrix} + q_2 \begin{pmatrix}
+\frac{1}{2^{2/3}}\\
+2^{1/3}
+\end{pmatrix} = \begin{pmatrix}
+\bar{z_1}\\
+\bar{z_2}
+\end{pmatrix}\\
+q_2^* = \frac{2\bar{z_2} - \bar{z_1}}{4 - \frac{1}{2^{2/3}}}, q_1^* = \frac{\bar{z_1}}{2^{1/3}} - \frac{2\bar{z_2} - \bar{z_1}}{8 - 2^{1/3}}
+\end{align*}
+One can easily check that under \eqref{eq1}, $q_1^* > 0, q_2^* > 0$ i.e. solution is indeed interior. Thus, equilibrium allocation is:
+\begin{align*}
+\begin{pmatrix}
+\bar{z_1} - \frac{2\bar{z_2} - \bar{z_1}}{\frac{8}{2^{1/3}} - 1}\\
+\\
+\frac{\bar{z_1}}{2^{2/3}} - \frac{2\bar{z_2} - \bar{z_1}}{2^{2/3}8 - 2}
+\end{pmatrix}, \begin{pmatrix}
+\frac{2\bar{z_2} - \bar{z_1}}{\frac{8}{2^{1/3}} - 1}\\
+\\
+2^{1/3} \cdot \frac{2\bar{z_2} - \bar{z_1}}{4 - \frac{1}{2^{2/3}}}
+\end{pmatrix}
+\end{align*}
+Assume now that $\frac{\bar{z_1}}{\bar{z_2}} > 2$ that means that the second firm shuts down. Hence
+\begin{align*}
+\begin{cases}
+c_1(w) = p_1\\
+\frac{a_{11}}{a_{21}} = \frac{\bar{z_1}}{\bar{z_2}}
+\end{cases} \to \frac{2w_2}{w_1} = \frac{\bar{z_1}}{\bar{z_2}} \to w_1 = \frac{1}{3}\left(\frac{\bar{z_2}}{\bar{z_1}}\right)^{1/3}, w_2 = \frac{1}{6}\left(\frac{\bar{z_1}}{\bar{z_2}}\right)^{2/3}
+\end{align*}
+and the equilibrium allocation is
+\begin{align*}
+\begin{pmatrix}
+\bar{z_1}\\
+\bar{z_2}
+\end{pmatrix}, \begin{pmatrix}
+0\\
+0
+\end{pmatrix}
+\end{align*}
+Assume now that $\frac{\bar{z_1}}{\bar{z_2}} < \frac{1}{2}$ that means that the first firm shuts down. Hence
+\begin{align*}
+\begin{cases}
+c_2(w) = p_2\\
+\frac{a_{12}}{a_{22}} = \frac{\bar{z_1}}{\bar{z_2}}
+\end{cases} \to \frac{w_2}{2w_1} = \frac{\bar{z_1}}{\bar{z_2}} \to w_1 = \frac{1}{6}\left(\frac{\bar{z_2}}{\bar{z_1}}\right)^{2/3}, w_2 = \frac{1}{3}\left(\frac{\bar{z_1}}{\bar{z_2}}\right)^{1/3}
+\end{align*}
+and the equilibrium allocation is
+\begin{align*}
+\begin{pmatrix}
+0\\
+0
+\end{pmatrix},
+\begin{pmatrix}
+\bar{z_1}\\
+\bar{z_2}
+\end{pmatrix}
 \end{align*}

### 2019-01-18 micro3_2.tex
Micro

@@
 \end{align*}
+\section*{Problem 4}
+Consider the 2x2 production model with constant returns to scale. Suppose that as a
+function of factor prices $w = (w_1; w_2) >> 0$, the firms' cost of producing one unit of output
+are as follows:
+\begin{align*}
+c_1(w) = \left(\frac{w_1}{3/4}\right)^{3/4}\left(\frac{w_2}{1/4}\right)^{1/4}\\
+c_2(w) = \left(\frac{w_1}{1/2}\right)^{1/2}\left(\frac{w_2}{1/2}\right)^{1/2}
+\end{align*}
+The firms' output are sold in the international market at the following prices per unit,
+respectively:
+\begin{align*}
+p_1 = 4,\  p_2 = 2\sqrt{3}
+\end{align*}
+Furthermore, the total endowment of factors 1 and 2 are given as follows, respectively:
+\begin{align*}
+\bar{z_1} = 2 + \frac{1}{\sqrt{3}},\ \bar{z_2} = 2 + \sqrt{3}
+\end{align*}
+\begin{enumerate}[(i)]
+	\item Compute the factor prices, factor allocation, and firms' output in an interior equilibrium.
+	\item For which values of $\bar{z_1}$ and $\bar{z_2}$ would the firm 1 shut down in equilibrium? Briefly interpret
+	your answer to provide some intuition.
+\end{enumerate}
+
+\textbf{Solution}
+\begin{enumerate}[(i)]
+\item \begin{align*}
+\frac{c_1(w)}{c_2(w)} &= \frac{p_1}{p_2}\\
+\frac{2}{3^{3/4}}\left(\frac{w_1}{w_2}\right)^{1/4} &= \frac{4}{2\sqrt{3}}\\
+\frac{w_1}{w_2} = 3 &\to w_1 = 3w_2\\
+c_1(w) = p_1 \to 4w_2 = 4 &\to w_2 = 1, w_1 = 3
+\end{align*}
+\begin{align*}
+a_1(w) = \begin{pmatrix}
+\left(\frac{3w_2}{w_1}\right)^{1/4}\\
+\left(\frac{w_1}{3w_2}\right)^{3/4}\\
+\end{pmatrix} = \begin{pmatrix}
+1\\
+1
+\end{pmatrix}\\
+a_2(w) = \begin{pmatrix}
+\left(\frac{w_2}{w_1}\right)^{1/2}\\
+\left(\frac{w_1}{w_2}\right)^{1/2}\\
+\end{pmatrix} = \begin{pmatrix}
+\frac{1}{\sqrt{3}}\\
+\sqrt{3}
+\end{pmatrix}\\
+\end{align*}
+\begin{align*}
+q_1 \begin{pmatrix}
+1\\
+1
+\end{pmatrix} + q_2 \begin{pmatrix}
+\frac{1}{\sqrt{3}}\\
+\sqrt{3}
+\end{pmatrix} = \begin{pmatrix}
+2 + \frac{1}{\sqrt{3}}\\
+2 + \sqrt{3}
+\end{pmatrix}\\
+q_1^* = 2, q_2^* = 1\\
+\end{align*}
+Allocation:
+\begin{align*}
+\begin{pmatrix}
+2\\
+2
+\end{pmatrix}, \begin{pmatrix}
+\frac{1}{\sqrt{3}}\\
+\sqrt{3}
+\end{pmatrix}
+\end{align*}
+\item Firm 1 shuts down when:
+\begin{align*}
+\frac{a_{11}}{a_{21}} = 1 > \frac{a_{12}}{a_{22}} = \frac{1}{3} > \frac{\bar{z_1}}{\bar{z_2}}\\
+\text{i. e. } \bar{z_2} > 3 \bar{z_1}
+\end{align*}
+qualitatively this condition means that there exists tiny amount of factor $\bar{z_1}$ compare to the amount of factor $\bar{z_2}$, but the first firm disposes the first factor very wasteful: for each unit of factor 2 it needs a unit of factor 1.
+In the same time, the firm 2 is a prudent firm, it uses scarce resources more efficient, for each unit of factor 2 it needs only $\frac{1}{3}$ of factor 1.
+As a result in the equilibrium only the firm 2 produces.
+\end{enumerate}
 \end{document}

### 2019-01-18 micro3_2.tex
Micro

@@
 \begin{align*}
-\frac{c_1(w)}{c_2(w)} = \frac{p_1}{p_2}\\ \frac{6}{2^{\frac{2}{3}}}\left(\frac{w_1}{w_2}\right)^{1/6} = \frac{6}{2^{\frac{2}{3}}3}\\
-w_2 = 3^6w_1
+\frac{c_1(w)}{c_2(w)} = \frac{p_1}{p_2}\\ \frac{3}{2^{\frac{2}{3}}2}\left(\frac{w_1}{w_2}\right)^{1/6} = \frac{6}{2^{\frac{2}{3}}3}\\
+w_1 = \frac{4^6}{3^6}w_2
 \end{align*}
@@
 \end{pmatrix} = \begin{pmatrix}
-2^{\frac{1}{3}}9\\
-\frac{1}{2^{\frac{2}{3}}81}
+2^{\frac{1}{3}}\frac{9}{16}\\
+\\
+\frac{256}{2^{\frac{2}{3}}81}
 \end{pmatrix}\\
@@
 \end{pmatrix} = \begin{pmatrix}
-27\\
+\frac{27}{64}\\
 \\
-\frac{1}{27}
+\frac{64}{27}
 \end{pmatrix}\\
-\frac{a_{11}}{a_{21}} = 2\cdot 3^6 > \frac{a_{12}}{a_{22}} = 3^6 > \frac{\bar{z_1}}{\bar{z_2}} = 2
+\frac{\bar{z_1}}{\bar{z_2}}  = 2 > \frac{a_{11}}{a_{21}} = 2\cdot \frac{3^6}{4^6} > \frac{a_{12}}{a_{22}} = \frac{3^6}{4^6}
 \end{align*}
-That means that in the equilibrium firm 1 should shut down.
-
+That means that in the equilibrium firm 2 should indeed shut down. Thus, 
+\begin{align*}
+\begin{cases}
+c_1(w) = p_1\\
+\frac{a_{11}}{a_{21}} = \frac{\bar{z_1}}{\bar{z_2}}
+\end{cases}\ \to 2\frac{w_2}{w_1} = 2 \to w_1 = w_2 \to \frac{3w_1}{2^{2/3}} = \frac{6}{2^{2/3}} \to w_1 = 2 = w_2
+\end{align*}
+Allocation:
+\begin{align*}
+\begin{pmatrix}
+2\\
+1
+\end{pmatrix}, \begin{pmatrix}
+0\\
+0
+\end{pmatrix}
+\end{align*}
+If $w_1 = w_2 = 2$ then $c_2(2, 2) = 4$ that is to produce one unit firm 2 should spend 4 dollars, while the price of unit of its production is only 3, that is it is indeed optimal for the second firm not to produce.
+Firm's 1 decision is also optimal because with $w_1 = w_2 = 2$ its cost of producing 1 unit becomes equal to price of the output.
 \section*{Problem 3}

### 2019-01-19 metrics2.tex
Metrics

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Econometrics 1. Problem Set 2.
\end{center}
\section*{Problem 1}
Suppose that the population model determining $y$ is
\begin{align*}
y = \beta_0 + \beta_1 x_1 + \beta_2x_2 + \beta_3x_3 + u
\end{align*}
and this model satisfies the Gauss-Markov assumptions. However, we estimate the model that
omits $x_3$. Let $\tilde{\beta_1}, \tilde{\beta_2}$ be the OLS estimators from the regression of $y$ on $x_1$ and $x_2$. Show
that the expected value of $\tilde{\beta_1}$ (given the values of the independent variables in the sample) is
\begin{align*}
\expect \tilde{\beta_1} = \beta_1 + \beta_3 \frac{\sum_{i=1}^n \hat{r_{i1}x_{i3}}}{\sum_{i=1}^n \hat{r_{i1}}^2}
\end{align*}
where the $\hat{r_{i1}}$ are the OLS residuals from the regression of $x_1$ on $x_2$.

\textbf{Solution}
Plugging $y_i = \beta_0 + \beta_1x_{i1} + \beta_2x_{i2}+\beta_3x_{i3}$ into
\begin{align*}
\tilde{\beta_1} = \frac{\sum_{i=1}^n \hat{r_{i1}}y_i}{\sum_{i=1}^n \hat{r_{i1}}^2}
\end{align*}
one can obtain:
\begin{align} \label{eq1}
\tilde{\beta_1} = \frac{\sum_{i=1}^n \hat{r_{i1}}(\beta_0 + \beta_1x_{i1} + \beta_2x_{i2}+\beta_3x_{i3} + u_i)}{\sum_{i=1}^n \hat{r_{i1}}^2} = \nonumber \\
=\frac{\beta_0 \sum_{i=1}^n \hat{r_{i1}} + \beta_1 \sum_{i=1}^n \hat{r_{i1}}x_{i1} + \beta_2 \sum_{i=1}^n \hat{r_{i1}}x_{i2} + \beta_2 \sum_{i=1}^n \hat{r_{i1}}x_{i2} + \sum_{i=1}^n \hat{r_{i1}}u_i}{\sum_{i=1}^n \hat{r_{i1}}^2}
\end{align}
By properties of OLS estimators (corollary from FOC): $\sum_{i=1}^n \hat{r_{i1}} = 0, \sum_{i=1}^n \hat{r_{i1}}x_{i2} = 0$. Moreover,
\begin{align*}
\sum_{i=1}^n \hat{r_{i1}}x_{i1} = \sum_{i=1}^n \hat{r_{i1}} (\hat{\alpha_0} + \hat{\alpha_1}x_{i2} + \hat{r_{i1}}) = 0 + 0 + \sum_{i=1}^n \hat{r_{i1}}^2
\end{align*}
where $\hat{\alpha_0}, \hat{\alpha_1}$ are OLS estimates for coefficients of regression of $x_1$ on $x_2$. Plugging it into (\ref{eq1}):
\begin{align*}
\tilde{\beta_1} = \beta_1 + \beta_3\frac{\sum_{i=1}^n\hat{r_{i1}}x_{i3}}{\sum_{i=1}^n \hat{r_{i1}}^2} + \frac{\sum_{i=1}^n \hat{r_{i1}}u_i}{\sum_{i=1}^n \hat{r_{i1}}^2}\\
\expect \tilde{\beta_1} = \beta_1 + \beta_3\frac{\sum_{i=1}^n\hat{r_{i1}}x_{i3}}{\sum_{i=1}^n \hat{r_{i1}}^2} + \expect\left[ \frac{\sum_{i=1}^n \hat{r_{i1}}u_i}{\sum_{i=1}^n \hat{r_{i1}}^2} \right]
\end{align*}
\begin{align*}
\hat{r_{i1}} = x_{i1} - \hat{x_{i1}} = x_{i1} - \hat{\alpha_0} - \hat{\alpha_1}x_{i2}\ \to \expect[f(\hat{r_{i1}})|x_{i1}, x_{i2}] = f(\hat{r_{i1}}) \ \forall\ f(\cdot)
\end{align*}
Then,
\begin{align*}
\expect\left[\frac{\sum_{i=1}^n \hat{r_{i1}}u_i}{\sum_{i=1}^n \hat{r_{i1}}^2} \right] = \expect\left[\expect\left[\frac{\sum_{i=1}^n \hat{r_{i1}}u_i}{\sum_{i=1}^n \hat{r_{i1}}^2}\bigg|x_{1}, x_{2}\right] \right] \underset{\text{i.i.d.}}{\propto} \expect\left[\sum_{i=1}^n\hat{r_{i1}}\expect [u_i|x_{i1}, x_{i2}]\right] \underset{\text{G.-M. assump.}}{=} 0 
\end{align*}
Thus, 
\begin{align*}
\expect \tilde{\beta_1} = \beta_1 + \beta_3\frac{\sum_{i=1}^n\hat{r_{i1}}x_{i3}}{\sum_{i=1}^n \hat{r_{i1}}^2}
\end{align*}
\end{document}

### 2019-01-19 metrics2.tex
Metrics

@@
 \newcommand{\expect}{\mathbb{E}}
+\newcommand{\summa}{\sum_{i=1}^n}
 \usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
@@
 \end{align*}
+\section*{Problem 2}
+\begin{enumerate}[(i)]
+	\item Consider the simple regression model $y = \beta_0 +\beta_1x+u$ under the first four Gauss-Markov
+	assumptions. For some function $g(x)$, for example $g(x) = x^2$ or $g(x) = \ln (1 + x^2)$, define
+	$z_i = g(x_i)$. Define a slope estimator as
+	\begin{align*}
+	\tilde{\beta_1} = \frac{\sum_{i=1}^n (z_i - \bar{z})y_i}{\sum_{i=1}^n (z_i - \bar{z})x_i}
+	\end{align*}
+	Show that $\tilde{\beta_1}$ is linear and unbiased.
+	\item Add the homoskedasticity assumption, MLR.5. Show that
+	\begin{align*}
+	Var \tilde{\beta_1} = \sigma^2 \frac{\sum_{i=1}^n (z_i -  \bar{z})^2}{\left(\sum_{i=1}^n (z_i - \bar{z})x_i\right)^2}
+	\end{align*}
+	\item Show directly that, under the Gauss-Markov assumptions, $Var\hat{\beta_1} \le Var\tilde{\beta_1}$, where $\hat{\beta_1}$ is the OLS estimator.
+\end{enumerate}
+
+\textbf{Solution}
+
+\begin{enumerate}[(i)]
+	\item Since 
+	\begin{align*}
+	\tilde{\beta_1} = a_1 y_1 + \dots + a_ny_n
+	\end{align*}
+	is a linear function of $y_1, \dots, y_n$ with some coefficients, thus, it is a linear estimator. Now let us show unbiasedness  
+	\begin{align*}
+	\tilde{\beta_1} = \frac{\sum_{i=1}^n (z_i - \bar{z})(\beta_0 + \beta_1 x_i + u_i)}{\sum_{i=1}^n(z_i - \bar{z})x_i} = \beta_1 + \frac{\sum_{i=1}^n(z_i - \bar{z})u_i}{\sum_{i=1}^n(z_i - \bar{z})x_i}\\
+	\expect \tilde{\beta_1} = \beta_1 + \expect \left[\frac{\sum_{i=1}^n(z_i - \bar{z})u_i}{\sum_{i=1}^n(z_i - \bar{z})x_i}\right]
+	\end{align*}
+	\begin{align*}
+	\expect[(z_i - \bar{z})u_i] \underset{\expect u_i = 0}{ = } \expect[z_iu_i] = \expect[\expect[z_iu_i|x_1, \dots, x_n]] \underset{\text{i.i.d}}{ = } \expect[z_i\expect[u_i|x_i]] \underset{\text{G.-M.
+assump.}}{ = } 0
+	\end{align*}
+	Hence 
+	\begin{align*}
+	\expect \tilde{\beta_1} = \beta_1
+	\end{align*}
+	i.e. the estimator is unbiased.
+	\item Assume $Var(u|x) = \sigma^2$ then
+	\begin{align*}
+	Var \tilde{\beta_1} &= Var\left[\frac{\sum_{i=1}^n (z_i - \bar{z})y_i}{\sum_{i=1}^n (z_i - \bar{z})x_i}\right] = \frac{1}{\left(\sum_{i=1}^n (z_i - \bar{z})x_i\right)^2} Var\left[\sum_{i=1}^n (z_i - \bar{z})y_i\right] \underset{\text{i.i.d}}{=}\\
+	&\underset{\text{i.i.d}}{=}\frac{1}{\left(\sum_{i=1}^n (z_i - \bar{z})x_i\right)^2} \sum_{i=1}^n Var((z_i - \bar{z})y_i)\\
+	&Var((z_i - \bar{z})y_i) = \expect[Var[((z_i - \bar{z})y_i)|x_1, \dots, x_n]] + Var(\expect[((z_i - \bar{z})y_i)|x_1, \dots, x_n]) = \\
+	&=\expect[(z_i - \bar{z})^2 Var(y_i|x_1, \dots, x_n)] + Var((z_i - \bar{z})\expect[y_i|x_1, \dots, x_n]) \underset{\text{i.i.d}}{=}\\
+	&\underset{\text{i.i.d}}{=} \expect[(z_i-\bar{z})^2 Var(y_i|x_i)] + Var((z_i - \bar{z})\expect[y_i|x_i])
+	\end{align*}
+	\begin{align*}
+	Var(y_i|x_i) = Var(\beta_0 + \beta_1x_i + u_i|x_i) = Var(u_i|x_i) = \sigma^2\\
+	\expect[y_i|x_i] = \expect[\beta_0 + \beta_1x_i + u_i|x_i] = \beta_0 + \beta_1x_i
+	\end{align*}
+	Hence
+	\begin{align*}
+	Var ((z_i - \bar{z})y_i) = (z_i - \bar{z})^2\sigma^2 + Var(const) = (z_i - \bar{z})^2\sigma^2
+	\end{align*}
+	Thus, 
+	\begin{align*}
+	Var \tilde{\beta_1} = \sigma^2 \frac{\sum_{i=1}^n (z_i -  \bar{z})^2}{\left(\sum_{i=1}^n (z_i - \bar{z})x_i\right)^2}
+	\end{align*}
+	\item 
+	\begin{align*}
+	\hat{\beta_1} = \frac{\sum_{i=1}^n(x_i - \bar{x})y_i}{\sum_{i=1}^n(x_i - \bar{x})^2}
+	\end{align*}
+	Similarly to (ii):
+	\begin{align*}
+	Var \hat{\beta_1} = \frac{ \sigma^2}{\sum_{i=1}^n(x_i - \bar{x})^2}
+	\end{align*}
+	By the Cauchy - Schwartz inequality:
+	\begin{align*}
+	\left(\sum_{i=1}^n (z_i - \bar{z})(x_i - \bar{x})\right)^2 = \left(\sum_{i=1}^n (z_i - \bar{z})x_i\right)^2 \le \left(\sum_{i=1}^n (x_i - \bar{x})^2\right) \left(\sum_{i=1}^n (z_i - \bar{z})^2\right)
+	\end{align*} 
+	hence
+	\begin{align*}
+	Var \tilde{\beta_1} \ge \sigma^2 \frac{\summa (z_i - \bar{z})^2}{\left(\sum_{i=1}^n (x_i - \bar{x})^2\right) \left(\sum_{i=1}^n (z_i - \bar{z})^2\right)} = \frac{ \sigma^2}{\sum_{i=1}^n(x_i - \bar{x})^2} = Var \hat{\beta_1}
+	\end{align*}
+\end{enumerate}
 \end{document}

### 2019-01-19 metrics2.tex
Metrics

@@
 \end{enumerate}
+\section*{Problem 3}
+Consider the standard linear multivariate regression model
+\begin{align*}
+y = \beta_0 + \beta_1x_1 + \beta_2x_2 + u
+\end{align*}
+under the Gauss-Markov assumptions.
+\begin{enumerate}[(a)]
+\item Derive the OLS estimate $\hat{\beta_2}$ in terms of sample variances and sample covariances of
+random variables $x_1$, $x_2$ and $y$.
+\item How would you get the OLS estimates $\hat{\beta_1}$ and $\hat{\beta_2}$ by computing a series of OLS estimates
+from simple regressions only?
+\item Suppose you have decided to get OLS estimates for $\gamma_0$, $\gamma_1$ and $\gamma_2$ from
+\begin{align*}
+x_2 = \gamma_0 + \gamma_1x_1 + \gamma_2 y + \nu.
+\end{align*}
+Is $\tilde{\beta_2} = \frac{1}{\hat{\gamma_2}}$ an unbiased estimate of $\beta_2$?
+\end{enumerate}
+
+\textbf{Solution}
+
+\begin{enumerate}[(a)]
+	\item Denote sample variances as $s^2_{ab}, a, b \in \left\{y, x_1, x_2\right\}$. To compute the OLS estimate we need to solve the following optimization problem:
+	\begin{align*}
+	\underset{b_0, b_1, b_2}{\min} \summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})^2\\
+	\text{FOCs:} \begin{cases}
+	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i}) = 0\\
+	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})x_{1i} = 0\\
+	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})x_{2i} = 0
+	\end{cases}\\
+	b_0 = \bar{y} - b_1\bar{x_1} - b_2 \bar{x_2}\\
+	b_1\left(\summa x_{1i}^2\right) = \summa (y_i - \bar{y} + b_1\bar{x_1} + b_2\bar{x_2} - b_2x_{2i})x_{1i}
+	\end{align*}
+	\begin{align*}
+	b_1 = \frac{\summa (y_i - \bar{y} - b_2(x_{2i} - \bar{x}))x_{1i}}{\summa (x_{1i} - \bar{x})^2}&=\frac{\summa (y_i - \bar{y})(x_{1i} - \bar{x_1}) - b_2\summa(x_{2i} - \bar{x_2})(x_{1i} - \bar{x_1})}{\summa (x_{1i} - \bar{x})^2} =\\
+	&=\frac{s^2_{yx_1} - b_2s^2_{x_1x_2}}{s^2_{x_1x_1}}
+	\end{align*}
+	Similarly for $b_2$:
+	\begin{align}\label{eq2}
+	b_2 = \frac{s^2_{yx_2} - b_1s^2_{x_1x_2}}{s^2_{x_2x_2}} &= \frac{s^2_{yx_2} - s^2_{yx_1} + b_2s^2_{x_1x_2}}{s^2_{x_2x_2}} \nonumber\\
+	b_2 &= \frac{s^2_{yx_2} - s^2_{yx_1}}{s^2_{x_2x_2} - s^2_{x_1x_2}}
+	\end{align}
+	\item Performing simple linear regression of $y$ on $x_1$, $y$ on $x_2$, $x_1$ on $x_2$ and $x_2$ on $x_1$ denote OLS estimates for slope coefficients as $\alpha_1, \alpha_2, \alpha_3, \alpha_4$ respectively.
+It is known that
+	\begin{align*}
+	\alpha_1 = \frac{s^2_{yx_1}}{s^2_{x_1x_1}},\ \alpha_2 = \frac{s^2_{yx_2}}{s^2_{x_2x_2}},\ \alpha_3 = \frac{s^2_{x_1x_2}}{s^2_{x_2x_2}},\ \alpha_4 = \frac{s^2_{x_1x_2}}{s^2_{x_1x_1}}
+	\end{align*}
+	Divide numerator and denumerator of (\ref{eq2}) by $s^2_{x_1x_1}$ (assuming that it is not equal to 0):
+	\begin{align*}
+	b_2 = \frac{\alpha_2\frac{\alpha_4}{\alpha_3} - \alpha_1}{\frac{\alpha_4}{\alpha_3} - \alpha_4}
+	\end{align*}
+	Similarly for $b_1$:
+	\begin{align*}
+	b_1 = \frac{\alpha_1\frac{\alpha_3}{\alpha_4} - \alpha_2}{\frac{\alpha_3}{\alpha_4} - \alpha_3}
+	\end{align*}
+	\item From (\ref{eq2}):
+	\begin{align*}
+	\hat{\gamma_2} = \frac{s^2_{yx_2} - s^2_{x_1x_2}}{s^2_{yy} - s^2_{yx_1}} \to \frac{1}{\hat{\gamma_2}} = \frac{s^2_{yy} - s^2_{yx_1}}{s^2_{yx_2} - s^2_{x_1x_2}}
+	\end{align*}
+\end{enumerate}
 \end{document}

### 2019-01-19 metrics2.tex
metrics

@@
 \end{enumerate}
+\section*{Problem 4}
+Consider a simple regression model through the origin under the Gauss-Markov assumptions
+\begin{align*}
+y = \gamma x + \varepsilon
+\end{align*}
+and let
+\begin{align*}
+\tilde{\gamma} = \frac{\summa y_ix_i^3}{\summa x_i^4}
+\end{align*}
+be an estimator of the slope parameter $\gamma$.
+\begin{enumerate}[(a)]
+\item Under what condition this estimator is well defined?
+\item Show conditional unbiasness of $\tilde{\gamma}$.
+\item Derive conditional variance of the estimator.
+\end{enumerate}
+
+\textbf{Solution}
+
+\begin{enumerate}[(a)]
+	\item First of all $x_i \neq 0\ \forall i$ then the estimator is well-defined.
+	\item \begin{align*}
+	\expect[\tilde{\gamma}|x_1, \dots, x_n] = \expect\left[\frac{\summa (\gamma x_i + \varepsilon_i)x_i^3}{\summa x_i^4}\bigg|x_1, \dots, x_n\right] &= \gamma + \frac{1}{\summa x_i^4} \summa x_i^3\expect[\varepsilon_i|x_1, \dots, x_n] =\\
+	 &\underset{\text{G.-M. assump}}{=} \gamma + 0
+	\end{align*}
+	\item \begin{align*}
+	Var(\tilde{\gamma}|x_1, \dots, x_n) = Var\left[\frac{\summa (\gamma x_i + \varepsilon_i)x_i^3}{\summa x_i^4}\bigg|x_1, \dots, x_n\right] \underset{\text{i.i.d}}{=} \frac{\summa x_i^6 Var[\varepsilon_i|x_i]}{\left(\summa x_i^4\right)^2}
+	\end{align*}
+	If $Var(\varepsilon|x) = const = \sigma^2$ then
+	\begin{align*}
+	Var(\tilde{\gamma}|x_1, \dots, x_2) = \sigma^2 \frac{\summa x_i^6}{\left(\summa x_i^4\right)^2} 
+	\end{align*}
+\end{enumerate}
 \end{document}

### 2019-01-21 macro3_1.tex
Macro

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Macroeconomics 3. Problem Set 1.
\end{center}
\section*{Problem 1}
Consider the following model
\begin{align}\label{eq1}
\underset{\left\{k(t+1), c(t)\right\}_{t=0}^{\infty}}{\max}\ \sum_{t=0}^{\infty} \beta^t u(c(t))\nonumber\\
s.t.\ k(t+1) = k(t) - c(t) \\
c(t) \ge 0, k(t) \ge 0, k(0) > 0\ \text{is given}\nonumber
\end{align}
\begin{enumerate}
	\item Rewrite it recursively. Replace $\beta$ with $\frac{1}{1+\rho}$
	and derive the asset pricing equations for the whole
	cake and for the marginal piece of cake. Explain what are the “dividends” and the “capital gains”
	in each case.
	\item Does the value of the marginal piece of cake increase or decrease over time? Use the asset pricing
	equation and explain the intuition.
	\item Does the value of the whole cake increase or decrease over time? Explain the intuition.
	\item At which rate does the “current value” of the marginal piece of cake change over time? What
	about the “present value”? Use the optimality conditions and explain the intuition.
	\item Assume CRRA utility function,
	\begin{align*}
	u(c(t)) = \frac{c(t)^{1 - \sigma}}{1 - \sigma}, \sigma > 0
	\end{align*}
	Derive the Euler equation. Explain the intuition of how the consumption path depends on the
	intertemporal elasticity of substitution.

	\item Now assume that the cake goes bad as the time goes on, that is the constraint \eqref{eq1} changes to
	\begin{align*}
	k (t + 1) = (1 - \delta) k (t) - c (t)
	\end{align*}
	where $0 < \delta < 1$. Explain how does it change your answers to the previous questions.
	\item Assume $\sigma = 1$ and $\delta = 0$. Find the analytical solutions for the value and policy functions.
\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item Let $V(k)$ be the value of the agent who acts optimally and possesses $k$ units of capital now. Then Bellman equation is:
	\begin{align*}
	V(k) = \underset{0 \le c \le k}{\max}\left\{u(c) + \beta V(k - c)\right\} \\
	V(k) = \underset{0 \le c \le k}{\max}\left\{u(c) + \frac{1}{1+\rho} V(k - c)\right\} 
	\end{align*}
	Assume $c^*$ is the optimizer, then
	\begin{align*}
	V(k) = u(c^*) + \frac{1}{1+\rho}V(k-c^*)\\
	\rho = \frac{V(k - c^*) - V(k) + u(c^*)}{V(k) - u(c^*)}
	\end{align*}
	here $V(k-c^*) - V(k)$ are "capital gains" and $u(c^*)$ are "dividends". By the envelope theorem
	\begin{align*}
	V'(k) = \frac{1}{1+\rho}V'(k-c^*)\\
	\rho = \frac{V'(k-c^*) - V'(k)}{V'(k)}
	\end{align*}
	\item 
	\item 
	\item 
	\item Assume 
	\begin{align*}
	u(c(t)) = \frac{c(t)^{1 - \sigma}}{1 - \sigma}, \sigma > 0
\end{align*}
then
\begin{align*}
&L = \sum_{t=0}^{\infty} \beta^t(u(c(t)) + \lambda(t)(k(t) - c(t) - k(t+1)))\\
&\begin{cases}\frac{\partial}{\partial c(t)}: u'(c(t)) - \lambda(t) = 0\\
\frac{\partial}{\partial k(t+1)}: -\beta^t \lambda(t) + \beta^{t+1}\lambda(t+1) = 0\\
\end{cases}\\
&\lambda(t) = \beta \lambda(t+1)\\
&u'(c(t+1)) = \lambda(t+1)\\
&\frac{u'(c(t))}{u'(c(t+1))} = \beta\\
\end{align*}
Thus, the Euler's equation is:
\begin{align*}
\left(\frac{c(t+1)}{c(t)}\right)^{\sigma} = \beta\\
\frac{c(t+1)}{c(t)} = \beta^{\frac{1}{\sigma}}
\end{align*}
intertemporal elasticity of substitution is $\frac{1}{\sigma}$. If it is equal to 0 then any changes of interest rate have no effect on consumption growth. If the intertemporal elasticity becomes larger than 0, then the ratio $\frac{c(t+1)}{c(t)}$ becomes low, that is the price of consumption 
\item 
\item Assume $\sigma = 1, \delta = 0$ then
from Euler's equation:
\begin{align*}
c(t+1) = \beta c(t)
\end{align*}
it remains to find $c(0)$,
\begin{align*}
&k(t) = k(t+1) + c(t) = k(t+2) + c(t+1) + c(t) = k(t+2) + \beta c(t) + c(t) = \\
&=k(t+3) + c(t+2) + \beta c(t) + c(t) = k(t+3) + \beta^2 c(t) + \beta c(t) + c(t) = \dots =\\
&=k(t + T) + c(t) \sum_{\tau = 0}^{T-1} \beta^{\tau} \\
&\text{since the amount of cake cannot be negative hence for optimal consumption path } k\\
&\text{ should tend to 0 at infinity, }\\
&k(t + T) + c(t) \sum_{\tau = 0}^{T-1} \beta^{\tau} \underset{T \to \infty}{\to} \frac{c(t)}{1-\beta} \to c(0) = (1-\beta)k(0)
\end{align*}
Thus, optimal consumption path is:
\begin{align*}
c(t+1) = \beta c(t), c(0) = (1-\beta)k(0)
\end{align*}
capital:
\begin{align*}
k(t+1) = k(t) - c(t)
\end{align*}
and the value function is:
\begin{align*}
\sum_{t=0}^{\infty} \beta^t \ln c(t) = \sum_{t=0}^{\infty} \beta^{t} \ln (\beta^t c(0)) = \sum_{t=0}^{\infty} \beta^t t \ln \beta + \sum_{t=0}^{\infty} \beta^t \ln c(0) = \frac{\beta \ln \beta}{(1-\beta)^2} + \frac{\ln c(0)}{1 - \beta } = \\
=\frac{\beta \ln \beta}{(1-\beta)^2} + \frac{\ln((1-\beta)k(0))}{1-\beta}
\end{align*}
\end{enumerate}

\section*{Problem 2}
Problem:
\begin{align*}
\underset{\left\{k(t+1), c(t)\right\}_{t=1}^{\infty}}{\max}\ \sum_{t=1}^{\infty} 0.96^t \ln c(t)\\
\text{s.t.\ } k(t+1) = Ak(t)^{1/3} + (1-0.02)k(t) - c(t)\\
c(t) \ge 0, k(t) \ge 0, k(1) = 0.1
\end{align*}
Euler's equation is:
\begin{align*}
\frac{c(t+1)}{\beta c(t)} = \frac{A}{3k(t+1)^{2/3}} + 1 - \delta
\end{align*}
In the steady state $c(t+1) = c(t)$ hence:
\begin{align*}
\frac{1}{\beta} = \frac{A}{3(k^*)^{2/3}} + 1 - \delta\\
\text{if } k^* = 1 \to A = 3(\frac{1}{\beta} + \delta - 1) = 0.185
\end{align*}
Our initial guess for $k(2)$ should lie within the interval $ [k(1), f(k(1)) + (1 - \delta)k(1)]$ because even if we assume $c(1) = 0$ we will not be able to increase capital $k(2)$ more than $f(k(1)) + (1 - \delta)k(1)$ and moreover since $k^* = 1$ while $k(1) = 0.1$ by the theorem optimal capital growth $s(k(t))$ is non-decreasing in $k(t)$ hence $k(2)$ cannot be less than $k(1)$. The global result which follows is that $k(t)$ cannot be greater than $k^*$ and less than $k(1)$.

Applying shouting algoritm the following value of $k(2)$ has been obtained:
\begin{align*}
k(2) = 0.1368665097187844
\end{align*}
and corresponding capital and consumption paths are depicted on the Fig. \ref{fig1} , \ref{fig2} respectively.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{k(t)}
	\caption{}\label{fig1}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{c(t)}
	\caption{}\label{fig2}
\end{figure}
\end{document}

### 2019-01-21 macro3_1.tex
Macro

@@
 	\end{align*}
-	here $V(k-c^*) - V(k)$ are "capital gains" and $u(c^*)$ are "dividends". By the envelope theorem
+	or alternatively:
+	\begin{align*}
+	\rho = \frac{u(c^*(t)) + V(k^*(t+1)) - V(k^*(t))}{V(k^*(t)) - u(c^*(t))}
+	\end{align*}
+	(* means optimal trajectory).
+	Here $\frac{V(k(t+1)) - V(k(t))}{V(k(t))}$ are "capital gains" and $u(c^*)$ are "dividends".
+
+	By the envelope theorem
 	\begin{align*}
@@
 	\end{align*}
-	\item 
-	\item 
-	\item 
+	or alternatively:
+	\begin{align*}
+	\rho = \frac{V'(k(t+1)) - V'(k(t))}{V'(k(t))}
+	\end{align*}
+	as one can observe there are no marginal dividends (because there are no production) only $\frac{V'(k(t+1)) - V'(k(t))}{V'(k(t))}$ which describes percentage change of shadow price of capital i.e.
+marginal "capital gains".
+	\item Of course in general the answer to this question depends on the particular type of function $u()$, however if we assume "good" increasing and concave utility function then price of marginal piece of cake increases over time.
+From the Euler's equation (which is obtained in 5.) $\frac{u'(c(t))}{u'(c(t+1))} = \frac{\lambda(t)}{\lambda(t+1)} = \beta$.
+Since $\lambda(t) = V'(k(t))$ then $\frac{V'(k(t))}{V'(k(t+1))} = \beta < 1$.
+That means that $V'(k(t))$ is increasing over time.
+Since $\rho$ is a constant (its meaning is the interest rate) then to hold it constant with increasing denumerator, the numerator $V'(k(t+1)) - V'(k(t))$ should also increase but $V'(k(t+1)) - V'(k(t))$ is exactly the price of marginal piece of cake.
+Intuitivelly it can be understood as follows: the amount of capital is shrinking hence over time the agent values each marginal unit of capital more and more.
+	\item Using the same equations as in the 2. $\frac{V'(k(t))}{V'(k(t+1))} = \beta$, that means that the present value of the whole cake stay constant: \begin{align*}
+	V'(k(t)) = \frac{1}{\beta}V'(k(t-1)) &= \frac{1}{\beta^2}V'(k(t-2)) = \dots = \frac{1}{\beta^t}V'(k(0))\\
+	\end{align*}
+	But the present value of $V'(k(t))$ is $\beta^t V'(k(t))$ hence 
+	\begin{align*}
+	\forall\ t\ \beta^t V'(k(t)) = V'(k(0))
+	\end{align*}
+	In the same time the current value is decreasing, which can be inferred from $\frac{V'(k(t))}{V'(k(t+1))} = \beta < 1$.
+	\item Present value is constant, current is increasing with rate $\frac{1}{\beta}$.
 	\item Assume 
@@
 \end{figure}
+
+It should be borne in mind that aforementioned $k(2)$ is the $\textit{approximate}$ value of true $k(2)$ that means that with this particular (in general not optimal) $k(2)$ for large $t$ the algorithm will diverge from the steady state.
+Hence those plots describe true capital and consumption paths only approximately.
 \end{document}

### 2019-01-21 macro3_1.tex
Macro

@@
 	\end{align*}
-	In the same time the current value is decreasing, which can be inferred from $\frac{V'(k(t))}{V'(k(t+1))} = \beta < 1$.
+	In the same time the current value is increasing, which can be inferred from $\frac{V'(k(t))}{V'(k(t+1))} = \beta < 1$.
 	\item Present value is constant, current is increasing with rate $\frac{1}{\beta}$.
@@
 If the intertemporal elasticity becomes larger than 0, then the ratio $\frac{c(t+1)}{c(t)}$ becomes low, that is the price of consumption
-\item 
+\item Now FOCs are:
+\begin{align*}
+\begin{cases}\frac{\partial}{\partial c(t)}: u'(c(t)) - \lambda(t) = 0\\
+\frac{\partial}{\partial k(t+1)}: -\beta^t \lambda(t) + \beta^{t+1}(1 - \delta)\lambda(t+1) = 0\\
+\end{cases}\\
+\frac{\lambda(t+1)}{\lambda(t)} = \frac{1}{\beta(1 - \delta)} > \frac{1}{\beta}
+\end{align*}
+that means that in this case even present value will increase over time:
+\begin{align*}
+V'(k(t)) = \frac{1}{(1 - \delta)^t}V'(k(0))
+\end{align*}
+and the rate of increasing of current value $\frac{1}{\beta(1-\delta)}$ goes up compare to the case $\delta = 0$.
 \item Assume $\sigma = 1, \delta = 0$ then

### 2019-01-21 macro3_1.tex
Macro

@@
 	\item Of course in general the answer to this question depends on the particular type of function $u()$, however if we assume "good" increasing and concave utility function then price of marginal piece of cake increases over time.
+It can be inferred both from the asset pricing equation and from the Euler's equation.
+From the asset pricing equation it is obviously that LHS is nonnegative (its meaning is the interest rate) hence the RHS should be also nonnegative.
+The denominator of the LHS is nonnegative because economically it is a shadow price of capital or matimatically it is equal to corresponding lagrange multiplier which cannot be negative by the Kuhn and Tucker theorem.
+Thus, numerator should be also nonnegative, so $V'(k(t+1)) \ge V'(k(t))$ i.e.
+price of marginal piece of cake increases over time.
 From the Euler's equation (which is obtained in 5.) $\frac{u'(c(t))}{u'(c(t+1))} = \frac{\lambda(t)}{\lambda(t+1)} = \beta$.
@@
 That means that $V'(k(t))$ is increasing over time.
-Since $\rho$ is a constant (its meaning is the interest rate) then to hold it constant with increasing denumerator, the numerator $V'(k(t+1)) - V'(k(t))$ should also increase but $V'(k(t+1)) - V'(k(t))$ is exactly the price of marginal piece of cake.
 Intuitivelly it can be understood as follows: the amount of capital is shrinking hence over time the agent values each marginal unit of capital more and more.
-	\item Using the same equations as in the 2. $\frac{V'(k(t))}{V'(k(t+1))} = \beta$, that means that the present value of the whole cake stay constant: \begin{align*}
+	\item Under some assumptions on the utility function it can be proved that $V(k)$ is strictly increasing.
+In our problem $k$ is decreasing over time that means that $V(k)$ should also decrease over time, i.e.
+the value of the whole cake decreases.
+	\item Using the same equations as in 2. $\frac{V'(k(t))}{V'(k(t+1))} = \beta$, that means that the present value of the marginal piece of cake remains constant: \begin{align*}
 	V'(k(t)) = \frac{1}{\beta}V'(k(t-1)) &= \frac{1}{\beta^2}V'(k(t-2)) = \dots = \frac{1}{\beta^t}V'(k(0))\\
@@
 	\end{align*}
-	In the same time the current value is increasing, which can be inferred from $\frac{V'(k(t))}{V'(k(t+1))} = \beta < 1$.
-	\item Present value is constant, current is increasing with rate $\frac{1}{\beta}$.
+	At the same time the current value is increasing, which can be inferred from $\frac{V'(k(t))}{V'(k(t+1))} = \beta < 1$.
 	\item Assume 
@@
 \end{align*}
-that means that in this case even present value will increase over time:
+that means that in this case even present value of marginal piece of cake will increase over time:
 \begin{align*}
@@
 and corresponding capital and consumption paths are depicted on the Fig. \ref{fig1} , \ref{fig2} respectively.
-\begin{figure}[h]
+\begin{figure}[H]
 	\centering

### 2019-01-22 macro3_1.tex
Macro

@@
 	(* means optimal trajectory).
-	Here $\frac{V(k(t+1)) - V(k(t))}{V(k(t))}$ are "capital gains" and $u(c^*)$ are "dividends".
+	Here $V(k(t+1)) - V(k(t))$ are gross "capital gains" and $u(c^*)$ are gross "dividends".
 
@@
 From the asset pricing equation it is obviously that LHS is nonnegative (its meaning is the interest rate) hence the RHS should be also nonnegative.
-The denominator of the LHS is nonnegative because economically it is a shadow price of capital or matimatically it is equal to corresponding lagrange multiplier which cannot be negative by the Kuhn and Tucker theorem.
+The denominator of the RHS is nonnegative because economically it is a shadow price of capital or matimatically it is equal to corresponding lagrange multiplier which cannot be negative by the Kuhn and Tucker theorem.
 Thus, numerator should be also nonnegative, so $V'(k(t+1)) \ge V'(k(t))$ i.e.
@@
 the value of the whole cake decreases.
+The amount of cake is decreasing over time, hence less cake $\to$ less it costs.
 	\item Using the same equations as in 2. $\frac{V'(k(t))}{V'(k(t+1))} = \beta$, that means that the present value of the marginal piece of cake remains constant: \begin{align*}
@@
 intertemporal elasticity of substitution is $\frac{1}{\sigma}$.
-If it is equal to 0 then any changes of interest rate have no effect on consumption growth.
-If the intertemporal elasticity becomes larger than 0, then the ratio $\frac{c(t+1)}{c(t)}$ becomes low, that is the price of consumption
+It shows the willingness of agents to substitute consumption over time i.e.
+to refuse today's consumption in order to increase it tomorrow.
+If $\frac{1}{\sigma} > 1$ then consumption path is steep: for the sake of increase consumtion today the agent is able to cut his consumption tomorrow, otherwise the consumption path is flat: the agent is not willing to cut tomorrow's consumption hence he should accept lower level of today's consumption (compare to the case $\frac{1}{\sigma} > 1$).
+See Fig.
+\ref{fig3}.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{plotdraft}
+	\caption{}\label{fig3}
+\end{figure}
 \item Now FOCs are:

### 2019-01-24 metrics2.tex
metrics

@@
 \newcommand{\summa}{\sum_{i=1}^n}
+\newcommand{\yrseduc}{\textit{yrseduc}}
 \usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
@@
 	\begin{align}\label{eq2}
-	b_2 = \frac{s^2_{yx_2} - b_1s^2_{x_1x_2}}{s^2_{x_2x_2}} &= \frac{s^2_{yx_2} - s^2_{yx_1} + b_2s^2_{x_1x_2}}{s^2_{x_2x_2}} \nonumber\\
-	b_2 &= \frac{s^2_{yx_2} - s^2_{yx_1}}{s^2_{x_2x_2} - s^2_{x_1x_2}}
+	b_2 &= \frac{s^2_{yx_2} - b_1s^2_{x_1x_2}}{s^2_{x_2x_2}}\nonumber\\
+	\text{hence }\nonumber\\
+	b_2 &= \frac{s^2_{yx_2}s^2_{x_1x_1} - s^2_{yx_1}s^2_{x_2x_1}}{s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2}
 	\end{align}
@@
 	\end{align*}
-	Divide numerator and denumerator of (\ref{eq2}) by $s^2_{x_1x_1}$ (assuming that it is not equal to 0):
+	Divide numerator and denumerator of (\ref{eq2}) by $(s^2_{x_1x_1})^2$ (assuming that it is not equal to 0):
 	\begin{align*}
-	b_2 = \frac{\alpha_2\frac{\alpha_4}{\alpha_3} - \alpha_1}{\frac{\alpha_4}{\alpha_3} - \alpha_4}
+	b_2 = \frac{\alpha_2\frac{\alpha_4}{\alpha_3} - \alpha_1\alpha_4}{\frac{\alpha_4}{\alpha_3} - \alpha_4^2} = \frac{\frac{\alpha_2}{\alpha_3} - \alpha_1}{\frac{1}{\alpha_3} - \alpha_4}
 	\end{align*}
@@
 	\begin{align*}
-	b_1 = \frac{\alpha_1\frac{\alpha_3}{\alpha_4} - \alpha_2}{\frac{\alpha_3}{\alpha_4} - \alpha_3}
+	b_1 = \frac{\alpha_1 - \alpha_2 \alpha_4}{1 - \alpha_3}
 	\end{align*}
@@
 	\begin{align*}
-	\hat{\gamma_2} = \frac{s^2_{yx_2} - s^2_{x_1x_2}}{s^2_{yy} - s^2_{yx_1}} \to \frac{1}{\hat{\gamma_2}} = \frac{s^2_{yy} - s^2_{yx_1}}{s^2_{yx_2} - s^2_{x_1x_2}}
+	\hat{\gamma_2} = \frac{s^2_{yx_2}s^2_{x_1x_1} - s^2_{x_2x_1}s^2_{yx_1}}{s^2_{x_1x_1}s^2_{yy} - (s^2_{x_1y})^2} \to \frac{1}{\hat{\gamma_2}} = \frac{s^2_{x_1x_1}s^2_{yy} - (s^2_{x_1y})^2}{s^2_{yx_2}s^2_{x_1x_1} - s^2_{x_2x_1}s^2_{yx_1}}\\
+	\end{align*}
+	To find $\expect[\frac{1}{\hat{\gamma_2}}]$ firstly observe the following facts:
+	\begin{align*}
+	s^2_{x_1x_1}s^2_{yy} = s^2_{x_1x_1}\cdot(\summa (y_i - \bar{y})(\beta_1x_{1i} + \beta_2x_{2i} + u_i - \beta_1\bar{x_1} - \beta_2 \bar{x_2} - \bar{u})) = \\
+	=s^2_{x_1x_1} (\beta_1s^2_{yx_1} + \beta_2s^2_{yx_2} + s^2_{uy})\\
+	(s_{x_1y}^2)^2 = s_{x_1y}^2 \cdot (\summa (x_{1i} - \bar{x})(\beta_1x_{1i} + \beta_2x_{2i} + u_i - \beta_1\bar{x_1} - \beta_2\bar{x_2} - \bar{u})) =\\ =s_{x_1y}^2 (\beta_1 s^2_{x_1x_1} + \beta_2s^2_{x_1x_2} + s^2_{x_1u})
+	\end{align*}
+	hence
+	\begin{align*}
+	s^2_{x_1x_1}s^2_{yy} - (s_{x_1y}^2)^2 = \beta_2(s^2_{x_1x_1}s^2_{yx_2} - s^2_{x_1y}s^2_{x_1x_2}) + s^2_{x_1x_1}s^2_{uy} - s^2_{x_1y}s^2_{x_1u}
+	\end{align*}
+	Thus,
+	\begin{align*}
+	\frac{1}{\hat{\gamma_2}} = \beta_2 + \frac{s^2_{x_1x_1}s^2_{uy} - s^2_{x_1y}s^2_{x_1u}}{s^2_{x_1x_1}s^2_{yx_2} - s^2_{x_1y}s^2_{x_1x_2}}
 	\end{align*}
@@
 \end{enumerate}
+\section*{Problem 5}
+In this problem set you continue to explore the returns to education and the gender gap
+in earnings. The data set is the same as for problem set 1 (cps99 ps1.csv; see problem set 1
+for a description of the variables). Consider the regression of average hourly earnings (\textit{ahe}) on
+years of education (\textit{yrseduc}). Consider the following three variables that have been omitted
+from the regression:
+\begin{enumerate}[(a)]
+\item gender
+\item a binary variable that = 1 if the worker’s last name falls in the first half of the alphabet,
+and = 0 if it falls in the second half.
+\item the worker’s native ability (for example IQ or some better measure)
+\end{enumerate}
+For each: Will this omission arguably lead to omitted variable bias? Why or why not?
+If your answer is that it will, state the sign of
+
+\textbf{Solution}
+
+Omitted variable bias appears when omitted variable is correlated with explanatory variables and affects dependent variable.
+\begin{enumerate}[(a)]
+	\item First of all, let us compute a sample correlation between gender and \textit{yrseduc}.
+The correlation is insignificantly small, hence we have reason to believe that gender is actualy not correlated with \textit{yrseduc}.
+Indeed, in most of countries there is no an arguable relationship between gender of the person and number of his or her years of education.
+However, in Russia, for example, and maybe in other countries with a conscript army system (except Israel, I think) I think men usually tend to spend more years being enrolled in some educational institutions in order not to go to the army.
+So, in general omission of gender variable can indeed lead to the omitted variable bias, but in our particular case, I suppose it is not a problem.
+	\item I think that such a variable both does not affect the response variable and is not correlated with years of education, so there is no omitted variable bias.
+	\item Yes, the omission of the worker's naitive ability can lead to the omitted variable bias.
+Because I suppose it both affects the average salary and is correlated with number of years of education.
+In general, I think the bias should be positive i.e.
+obtained in the misspecified model OLS estimate of the slope coefficient sould be on average greater than the true value of slope coefficient for \textit{yrseduc}.
+Since $\tilde{\beta_1} = \hat{\beta_1} + \hat{\beta_2} \delta$ where $\tilde{\beta_1}$ is the OLS estimate of the coefficient of misspecified model, $\hat{\beta_1}, \hat{\beta_2}$ are OLS estimates for the miltiple regression with both variables \textit{yrseduc} and innate abilities, and $\delta$ is the OLS estimate in regression innate abilities on \textit{yrseduc}.
+It can be reasonable to assume that the more skillful pearson is the more higher his salary is, hence $\hat{\beta_2} > 0$.
+Not as much reasonable would it assume that the more skillful pearson is the more years he spend on education, hence $\delta > 0$.
+However it can be the case that more skillful people on the contrary spend less years on education (e.g.
+Bill Gates, Steave Jobs) because of high opportunity costs of it, in this situation $\delta < 0$ and bias is negative.
+\end{enumerate}
+\section*{Problem 6}
+Consider an experimental approach to estimation of the effect on earnings of education.
+\begin{enumerate}[(a)]
+\item Describe an experimental protocol for an idealized randomized controlled experiment
+that would permit unbiased estimation of the effect on earnings of an additional year of education at a typical US educational institution. (Ignore practical and ethical issues.)
+\item Explain why, precisely, your experimental protocol would eliminate omitted variable
+bias arising from omission of the variables in question 1 (or any other variables).
+\end{enumerate}
+
+\textbf{Solution}
+\begin{enumerate}[(a)]
+	\item An idealized randomized controlled experiment should be as follows: we should absolutely randomly choose two groups of people.
+Absolutely randomly in the sense that each person of the USA could equally likely be chosen for this experiment.
+People of the first group should be given an additional year of education, while people from the second group should be abstained from the education over the year.
+Then after this year
+\end{enumerate}
+\section*{Problem 7}
+Estimate two regressions: (i) \textit{ahe} on \textit{yrseduc} and (ii) \textit{ahe} on \textit{yrseduc} and \textit{female}.
+\begin{enumerate}[(a)]
+\item In regression (ii), what is the coefficient on \yrseduc? Explain what this means.
+\item In regression (ii), test the hypothesis that the population coefficient on female in
+specification (ii) is zero, against the hypothesis that it is nonzero, at the 5\% significance level.
+In everyday words (not statistical terms), what precisely is the hypothesis is that you are
+testing?
+\item Does the coefficient on \textit{yrseduc} change from regressions (i) to (ii) in a substantively
+important way, that is, is the difference between the two estimates large in a real-world sense?
+What is the reason that including \textit{female} in the regression results in a large (or only a small)
+change in the coefficient on \yrseduc? (Hint: what is the correlation between \yrseduc\ and
+\textit{female}?)
+
+\end{enumerate}
+
+\textbf{Solution}
+
+\begin{enumerate}[(a)]
+	\item Coefficient on \yrseduc\ is equal to 1.34147.
+That means that holding a gender of the person fixed each additional year of education on average gives 1.34147 dollars of increase of the per hour earnings.
+	\item Using coeftest() it can be obtained that the null hypothesis should be rejected on the 5\% significance level.
+The hypothesis was that holding a number of years of education fixed both males and females get on average equal salary per hour.
+	\item Coefficient on \yrseduc\ did not significantly change: from 1.32 to 1.34.
+The reason is the follwoing that as it has been already indicated in the problem 5 (a) the sample correlation between \yrseduc\ and \textit{female} is very low, as a result, using formula from the 5 (c) $\tilde{\beta_1} = \hat{\beta_1} + \hat{\beta_2} \delta$ one can conclude that extremely small $\delta$ results in $\tilde{\beta_1} \approx \hat{\beta_1}$ even if $\hat{\beta_2} \neq 0$.
+\end{enumerate}
 \end{document}

### 2019-01-24 Metrics2.R

@@ -1,0 +1,20 @@
+df <- read.csv('cps99_ps1.csv')
+
+#Problem 5
+cor(df$yrseduc, df$female)
+
+#Problem 7
+#a
+fit1 <- lm(df$ahe ~ df$yrseduc, data = df)
+fit2 <- lm(df$ahe ~ df$yrseduc + df$female, data = df)
+
+summary(fit2)
+
+#b
+library(lmtest)
+coeftest(fit2, vcov = vcovHC(fit2, type = "HC1"))
+
+#c
+summary(fit1)
+
+#Problem 8

### 2019-01-25 metrics2.tex
metrics

@@
 	\end{align*}
-	\item From (\ref{eq2}):
+	\item 
 	\begin{align*}
-	\hat{\gamma_2} = \frac{s^2_{yx_2}s^2_{x_1x_1} - s^2_{x_2x_1}s^2_{yx_1}}{s^2_{x_1x_1}s^2_{yy} - (s^2_{x_1y})^2} \to \frac{1}{\hat{\gamma_2}} = \frac{s^2_{x_1x_1}s^2_{yy} - (s^2_{x_1y})^2}{s^2_{yx_2}s^2_{x_1x_1} - s^2_{x_2x_1}s^2_{yx_1}}\\
+	\beta_2 = \frac{1}{\gamma_2}\\
+	\expect[\hat{\beta_2}] = \beta_2\\
+	\expect[\hat{\gamma_2}] = \gamma_2\\
+	\beta_2 = \frac{1}{\expect[\hat{\gamma_2}]}
 	\end{align*}
-	To find $\expect[\frac{1}{\hat{\gamma_2}}]$ firstly observe the following facts:
+	Since $\frac{1}{x}$ is a strictly concave function for $x < 0$ and strictly convex for $x > 0$ and moreover $\hat{\gamma_2}$ is not almost sure constant (we assume it) then by Jensen's inequality:
 	\begin{align*}
-	s^2_{x_1x_1}s^2_{yy} = s^2_{x_1x_1}\cdot(\summa (y_i - \bar{y})(\beta_1x_{1i} + \beta_2x_{2i} + u_i - \beta_1\bar{x_1} - \beta_2 \bar{x_2} - \bar{u})) = \\
-	=s^2_{x_1x_1} (\beta_1s^2_{yx_1} + \beta_2s^2_{yx_2} + s^2_{uy})\\
-	(s_{x_1y}^2)^2 = s_{x_1y}^2 \cdot (\summa (x_{1i} - \bar{x})(\beta_1x_{1i} + \beta_2x_{2i} + u_i - \beta_1\bar{x_1} - \beta_2\bar{x_2} - \bar{u})) =\\ =s_{x_1y}^2 (\beta_1 s^2_{x_1x_1} + \beta_2s^2_{x_1x_2} + s^2_{x_1u})
-	\end{align*}
-	hence
-	\begin{align*}
-	s^2_{x_1x_1}s^2_{yy} - (s_{x_1y}^2)^2 = \beta_2(s^2_{x_1x_1}s^2_{yx_2} - s^2_{x_1y}s^2_{x_1x_2}) + s^2_{x_1x_1}s^2_{uy} - s^2_{x_1y}s^2_{x_1u}
+	\expect\left[\frac{1}{\hat{\gamma_2}}\right] < \frac{1}{\expect[\hat{\gamma_2}]}, \hat{\gamma_2} < 0\\
+	\expect\left[\frac{1}{\hat{\gamma_2}}\right] > \frac{1}{\expect[\hat{\gamma_2}]}, \hat{\gamma_2} > 0\\
 	\end{align*}
@@
 	\begin{align*}
-	\frac{1}{\hat{\gamma_2}} = \beta_2 + \frac{s^2_{x_1x_1}s^2_{uy} - s^2_{x_1y}s^2_{x_1u}}{s^2_{x_1x_1}s^2_{yx_2} - s^2_{x_1y}s^2_{x_1x_2}}
+	\expect\left[\frac{1}{\hat{\gamma_2}}\right] < \beta_2, \hat{\gamma_2} < 0\\
+	\expect\left[\frac{1}{\hat{\gamma_2}}\right] > \beta_2, \hat{\gamma_2} > 0\\
 	\end{align*}
+	hence it is biased.
 \end{enumerate}
@@
 \end{enumerate}
+\section*{Problem 8}
+Use multiple linear regression of \textit{ahe} on \yrseduc\ and \textit{ba} (a binary variable equaling one for
+those who attained a bachelor’s degree) to estimate the marginal value of a bachelor’s degree,
+more specifically, the value of going to school for 16 years and receiving a bachelor’s degree,
+relative to going to school for 16 years but not receiving a bachelor’s degree. Construct a 95\%
+confidence interval for the marginal value of a bachelor’s degree.
 \end{document}

### 2019-01-25 micro3_3.tex
Micro

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\newcommand{\yrseduc}{\textit{yrseduc}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Microeconomics 3. Problem Set 3.
\end{center}
\section*{Problem 1}
Consider an Edgeworth box exchange economy with two commodities. Consumers'
endowments are given by 
\begin{align*}
\omega_1 = (\frac{\sqrt{2}}{2}, 4),\ \omega_2 = (\frac{\sqrt{2}}{2}, 5)
\end{align*}. 
Assume further that the utility
functions of the two consumers are as follows:
\begin{align*}
u_1(x_1) := x_{11} + \sqrt{x_{21}},\ \ 
u_2(x_2) := \sqrt{8}x_{12} + \sqrt{x_{22}}
\end{align*}
\begin{enumerate}[(i)]
\item Find the set of Pareto efficient allocations parametrically, as a function of the utility level
of consumer 2.
\item For each interior Pareto efficient allocation, find a corresponding equilibrium with transfers.

\end{enumerate}

\textbf{Solution}
\begin{enumerate}[(i)]
	\item To find the set of Pareto efficient allocations one need to solve the following optimization problem
	\begin{align*}
	\underset{x_{11}, x_{21} \ge 0}{\max\ }\ x_{11} + \sqrt{x_{21}}\\
	s.t.\ \sqrt{8}x_{12} + \sqrt{x_{22}} \ge \bar{u}\\
	x_{11} + x_{12} = \sqrt{2}\\
	x_{21} + x_{22} = 9
	\end{align*}
	Or alternatively
	\begin{align*}
	\underset{x_{11}, x_{21} \ge 0}{\max}\ x_{11} + \sqrt{x_{21}}\\
	s.t.\ 4 - \sqrt{8}x_{11} + \sqrt{9 - x_{21}} \ge \bar{u}\\
	x_{11} \le \sqrt{2},\ x_{21} \le 9
	\end{align*}
	constraint is obviously binding because otherwise we can slightly increase $x_{11}$ such that the constraint still will hold but the objective function will increase.
	(It should be borne in mind that $\bar{u}$ cannot exceed 7 because $u_2(\sqrt{2}, 9) = 7$). For the interior solution:
	\begin{align*}
	\text{FOCs}: &\begin{cases}
	1 - \sqrt{8}\lambda = 0\\
	\frac{1}{\sqrt{x_{21}}} - \frac{\lambda}{\sqrt{9 - x_{21}}} = 0
	\end{cases}\\
	\lambda = \frac{1}{\sqrt{8}}\ \to\ 
	\frac{1}{\sqrt{x_{21}}} &- \frac{1}{\sqrt{72 - 8 x_{21}}} = 0 \to\ \begin{cases} x_{21} = 8\\
	x_{11} = \frac{5 - \bar{u}}{\sqrt{8}}
	\end{cases}
	\end{align*}
	Thus the set of Pareto efficient allocation is:
	\begin{align}\label{eq1}
	\begin{cases}
	x^*_1 = 
	\begin{pmatrix}
	\sqrt{2}\\
	9 - \bar{u}^2
	\end{pmatrix},\ x^*_2 = \begin{pmatrix}
	0\\
	\bar{u}^2
	\end{pmatrix}, 0 \le \bar{u} < 1\\
	x^*_1 = 
	\begin{pmatrix}
	\frac{5 - \bar{u}}{\sqrt{8}}\\
	8
	\end{pmatrix},\ x^*_2 = \begin{pmatrix}
	 \frac{\bar{u}- 1}{\sqrt{8}}\\
	 1
	\end{pmatrix}, 1 \le \bar{u} < 5\\
	x^*_1 = 
	\begin{pmatrix}
	0\\
	9 - (\bar{u} - 4)^2\\
	\end{pmatrix},\ x^*_2 = \begin{pmatrix}
	\sqrt{2}\\
	(\bar{u} - 4)^2
	\end{pmatrix}, 5 < \bar{u} \le 7\\
	\end{cases}
	\end{align}
	\item To find euilibria with transfers we need to find wealth levels and prices corresponding to each interior Pareto efficient allocation. That means that we need to find such $w_1, w_2$ and such a price vector $p = (p_1\ p_2)$ that $x^*_1, x^*_2$ from the \eqref{eq1} maximize consumers' utility functions over the following budget sets:
	\begin{align*}
	\left\{(x_{11}, x_{21}) \in \mathbb{R}^2_{+}: p_1x_{11} + p_2x_{21} \le w_1\right\}\\	\left\{(x_{12}, x_{22}) \in \mathbb{R}^2_{+}: p_1x_{12} + p_2x_{22} \le w_2\right\}\\
	\end{align*}
	\begin{align*}
	\underset{x_{11}, x_{21} \ge 0}{\max}\ x_{11} + \sqrt{x_{21}}\\
	 s.t.\ p_1x_{11} + p_2x_{21} \le w_1
	\end{align*}
	In the interior allocations MRS = MRT hence
	\begin{align*}
	2\sqrt{x_{21}} = \frac{p_1}{p_2} \to\ p_1 = 4\sqrt{2}p_2
	\end{align*}
	We can assume $p_2 = 1$ hence $p_1 = 4\sqrt{2}$
	\begin{align*}
	\begin{cases}
	4\sqrt{2} \cdot \frac{5 - \bar{u}}{\sqrt{8}} + 8 = w_1\\
	4\sqrt{2} \cdot \frac{\bar{u} - 1}{\sqrt{8}} + 1 = w_2
	\end{cases}\\
	\begin{cases}
	w_1 = 18 - 2\bar{u}\\
	w_2 = 2\bar{u} - 1
	\end{cases}, 
	\end{align*}
\end{enumerate}
\end{document}

### 2019-01-26 micro3_3.tex
Micro

@@
 	\bar{u}^2
-	\end{pmatrix}, 0 \le \bar{u} < 1\\
+	\end{pmatrix}, 0 \le \bar{u} \le 1\\
 	x^*_1 = 
@@
 	 1
-	\end{pmatrix}, 1 \le \bar{u} < 5\\
+	\end{pmatrix}, 1 < \bar{u} < 5\\
 	x^*_1 = 
@@
 	(\bar{u} - 4)^2
-	\end{pmatrix}, 5 < \bar{u} \le 7\\
+	\end{pmatrix}, 5 \le \bar{u} \le 7\\
 	\end{cases}
@@
 	\begin{align*}
-	\begin{cases}
+	&\begin{cases}
 	4\sqrt{2} \cdot \frac{5 - \bar{u}}{\sqrt{8}} + 8 = w_1\\
@@
 	\end{cases}\\
-	\begin{cases}
+	&\begin{cases}
 	w_1 = 18 - 2\bar{u}\\
 	w_2 = 2\bar{u} - 1
-	\end{cases}, 
+	\end{cases}, 1 < u < 5
 	\end{align*}
 \end{enumerate}
+\section*{Problem 2}
+Let $\bar{\omega_1}$ and $\bar{\omega_2}$ denote the total endowments of the two goods in an Edgeworth box
+exchange economy. Assume further that the utility functions of the two consumers are as
+follows:
+\begin{align*}
+u_1(x_1) := (x_{11})^{2/3}(x_{21})^{1/3}\ \ 
+u_2(x_2) := (x_{12})^{1/3}(x_{22})^{2/3}
+\end{align*}
+Consider the following welfare maximization problem, as a function of $\lambda \in (0; 1)$:
+\begin{align} \label{eq2}
+\underset{(x_1, x_2)}{\max}\ \lambda u_1(x_1) + (1 - \lambda)u_2(x_2)\ s.t.\ x_{l1} + x_{l2} = \bar{\omega} \text{ and } x_{li} > 0 \text{ for } l, i \in \left\{1, 2\right\}
+\end{align}
+Let $(x_1(\lambda); x_2(\lambda))$ denote the solution, where $x_i(\lambda) = (x_{1i}(\lambda); x_{2i}(\lambda))$ for $i = 1; 2.$
+\begin{enumerate}[(i)]
+\item Assuming an interior solution with $x_i(\lambda) >> 0$ for $i = 1; 2$, use the first order conditions
+of problem \eqref{eq2} to compute the ratios $\frac{x_{11}(\lambda)}{
+x_{21}(\lambda)}$
+and $\frac{x_{12}(\lambda)}{
+x_{22}(\lambda)}$
+in terms of $\lambda$.
+\item  Note that $\frac{x_{11}(\lambda)}{x_{21}(\lambda)} >\frac{x_{12}(\lambda)}{x_{22}(\lambda)}$
+and both ratios are decreasing in $\lambda$. What additional conditions
+on $\bar{\omega_1}; \bar{\omega_2}$ and $\lambda$ are required for problem \eqref{eq2} to have an interior solution as you assumed?
+Graphically illustrate the set of interior efficient allocations, and clarify the role of $\lambda$ in
+the graph. (Hints. Feel free to do the algebra to characterize the interiority condition,
+but it is not really needed. Instead, let us think of a utility function $u_i$ as a production
+function $f_i$ that produces \textit{utils} using consumption goods as inputs. Then, in problem \eqref{eq2},
+$\lambda$ and $1-\lambda$ can be viewed as the prices of two outputs, namely the utility levels of agents 1
+and 2, respectively. This makes problem \eqref{eq2} equivalent to the revenue maximization problem
+(15.D.5) in Chapter 15 of the textbook. In turn, the factor allocation that solves the problem
+(15.D.5) is the same as the equilibrium factor allocation of the corresponding small open
+economy. Conclusion: Aside from the differences in optimization problems of individual
+agents/firms (utility maximization subject to a budget constraint vs profit maximization),
+the Edgeworth exchange model and the $2 \times 2$ open economy are quite similar to each other,
+at least from a mathematical point of view. The similarity is especially useful in efficiency
+issues.)
+\end{enumerate}
+
+\textbf{Solution}
+
+\begin{enumerate}[(i)] 
+	\item If we assume the interior solution then FOCs are
+	\begin{align*}
+	\begin{cases}
+	\frac{2 \lambda }{3} \left(\frac{x_{21}}{x_{11}}\right)^{1/3} = \mu_1\\
+	\frac{\lambda}{3} \left(\frac{x_{11}}{x_{21}}\right)^{2/3} = \mu_2\\
+	\frac{1 - \lambda}{3} \left(\frac{x_{22}}{x_{12}}\right)^{2/3} = \mu_1\\
+	\frac{2(1-\lambda)}{3} \left(\frac{x_{12}}{x_{22}}\right)^{1/3} = \mu_2\\
+	\end{cases}
+	\end{align*}
+	where $\mu_l$ are Lagrange multipliers.
+	\begin{align*}
+	\frac{x_{11}}{x_{21}} = \frac{2\mu_2}{\mu_1}\\
+	\frac{x_{12}}{x_{22}} = \frac{\mu_2}{2\mu_1}
+	\end{align*}
+	hence
+	\begin{align*}
+	\begin{cases}
+	\frac{2\lambda}{3} \left(\frac{\mu_1}{2 \mu_2}\right)^{1/3} = \mu_1\\
+	\frac{1-\lambda}{3} \left(\frac{2\mu_1}{\mu_2}\right)^{2/3} = \mu_1
+	\end{cases} \to \frac{\mu_2}{\mu_1} = \left(\frac{1-\lambda}{\lambda}\right)^{3} \to \begin{cases}
+	\frac{x_{11}}{x_{21}} = 2 \left(\frac{1-\lambda}{\lambda}\right)^{3}\\
+	\frac{x_{12}}{x_{22}} = \frac{1}{2}\left(\frac{1-\lambda}{\lambda}\right)^{3}
+	\end{cases}
+	\end{align*}
+	\item Following the hint and using logic from $2 \times 2$ production model we can infer that the solution will indeed interior iff
+	\begin{align*}
+	\frac{x_{11}}{x_{21}} >& \frac{\bar{\omega_1}}{\bar{\omega_2}} > \frac{x_{12}}{x_{22}}\\
+	2\left(\frac{1 - \lambda}{\lambda}\right)^3 >& \frac{\bar{\omega_1}}{\bar{\omega_2}} >\frac{1}{2} \left(\frac{1 - \lambda}{\lambda}\right)^3
+	\end{align*}
+	The set of interior efficient allocations is depicted below, on the Fig \ref{fig1}.
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.8\textwidth]{plotdraft}
+		\caption{}\label{fig1}
+	\end{figure}
+\end{enumerate}
+\section*{Problem 3}
+In the $2\times 2$ open economy model, an allocation $z = (z_1; z_2)$ is feasible if $z_1+z_2 = \bar{z}$, where
+$\bar{z}$ is the total endowment vector $(\bar{z_1}; \bar{z_2})$. In turn, a feasible allocation is said to be efficient
+if there is no other feasible allocation $\hat{z} = (\hat{z_1}; \hat{z_2})$ such that $f_j(\hat{z_j}) \ge f_j(z_j)$ for both firms $j$,
+with strict inequality for some $j$. Using our discussion of Pareto efficiency from Chapter 16,
+suggest two different optimization techniques to characterize efficient factor allocations in a
+$2 \times 2$ open economy with concave production functions. (No need to go into details, just
+state the optimization problems.)
+
+\textbf{Solution}
+
+There can be two approaches, namely:
+A feasible allocation $(z_1^*, z_2^*)$ is efficient $\iff\ \exists\ \bar{f} \ge 0$ such that $(z_1^*, z_2^*)$ solves the following optimization problem 
+\begin{align*}
+\underset{z_{11} z_{21} \ge 0}{\max}\ f_1(z_{11}, z_{21})\\
+s.t.\ f_2(z_{12}, z_{22}) \ge \bar{f}\\
+z_{11} + z_{12} = \bar{z_1}\\
+z_{21} + z_{22} = \bar{z_2}
+\end{align*}
+or:
+A feasible allocation $(z_1^*, z_2^*)$ is efficient $\iff\ \exists\ \lambda_1 \ge 0, \lambda_2 \ge 0: \lambda_1^2 + \lambda_2^2 > 0$ such that $(z_1^*, z_2^*)$ solves the following optimization problem:
+\begin{align*}
+\underset{z_{11}, z_{21}, z_{12}, z_{22} \ge 0}{\max}\ \lambda f_1(z_{11}, z_{21}) + (1 - \lambda)f_2(z_{12}, z_{22})\\
+s.t. z_{11} + z_{12} = \bar{z_1}\\
+	z_{12} + z_{22} = \bar{z_2}
+\end{align*}
+\section*{Problem 4}
+Assuming interior solutions, show that the first order conditions of problems (16.E.1)
+and (16.F.1) in the textbook are equivalent to each other. Under what conditions on utility
+functions will the two problems lead to the same solutions? Explain your answer.
+
+\textbf{Solution}
+
+Problem 16.E.1.:
+\begin{align*}
+\underset{(x, y)}{\max}\ \lambda_1u_1(x_1) + \dots + \lambda_Iu_I(x_I)\\
+s.t.\ \sum_{i=1}^I x_{li} = \bar{\omega_l} + \sum_{j=1}^Jy_{lj}, \forall\ l = 1, \dots, L\\
+x_{li} \ge 0, \forall\ i = 1, \dots, I, l = 1, \dots, L\\
+y_{i} \in Y_i,\ \forall\ i = 1, \dots, I
+\end{align*}
+Assuming interior solution FOCs will be:
+\begin{align*}
+\begin{cases}
+\lambda_1\frac{\partial u_1(x_1)}{\partial x_{l1}} - \mu_l = 0,\ \ \  l = 1, \dots, L\\
+\dots\ \dots\ \dots\\
+\lambda_i \frac{\partial u_i(x_i)}{\partial x_{li}} - \mu_l = 0,\ \ \  l = 1, \dots, L\\
+\dots\ \dots\ \dots\\
+\lambda_I\frac{\partial u_I(x_I)}{\partial x_{lI}} - \mu_l = 0,\ \ \  l = 1, \dots, L
+\end{cases}
+\end{align*}
+where $\mu_l$ are Lagrange multipliers which correspond to $L$ market-clearing constraints.
+
+Problem 16.F.1:
+\begin{align*}
+&\underset{(x, y)}{\max}\ u_1(x_1)\\
+&s.t.\ u_i(x_i) \ge \bar{u_i},\ \ \ i=2, \dots, L\\
+&\sum_{i=1}^Ix_{li} \le \bar{\omega_l} + \sum_{j=1}^Jy_{lj},\ \ \ l = 1, \dots, L\\
+&F_j(y_j) \le 0\ \ \ j = 1, \dots, J\\
+&x_{li} \ge 0, \ \forall i, l
+\end{align*}
+Of course market clearing constraint and constraint on utility levels are binding because in latter case for example otherwise we can decrease the amount of good for some agent and give this amount to the 1st agent, increasing the oblective function.
+Assuming interior solution FOCs will be:
+\begin{align*}
+\begin{cases}
+\frac{\partial u_1(x_1)}{\partial x_{l1}} - \nu_l = 0, \ \ \ l = 1, \dots, L\\
+\dots\ \dots\ \dots\\
+\eta_i\frac{\partial u_i(x_i)}{\partial x_{li}} - \nu_l = 0,\ \ \ l = 1, \dots, L\\
+\dots\ \dots\ \dots\\
+\eta_I\frac{\partial u_I(x_I)}{\partial x_{lI}} - \nu_l = 0, \ \ \ l = 1, \dots, L
+\end{cases}
+\end{align*}
+Where $\eta_i$ are Lagrange multiplier which correspond to $I-1$ utility levels' constraints and $\nu_l$ are Lagrange multiplier which correspond to $L$ market clearing constraints.
+As we can observe, if $\lambda_1 \neq 0$ in the problem 16.E.1 then these FOCs are equivalent to each other.
+
+If utility functions are concave then the utility possibility set is a convex set.
+Obviously for each particular $(\bar{u_2}, \dots, \bar{U_I})$ the solution to the problem 16.F.1.
+should lie on the frontier (see Fig.), but when the UPS is a convex set by the separating hyperplane theorem there should exists such a non-zero $\lambda = (\lambda_1\ \dots\ \lambda_I)$ such that is is a solution to the problem 16.E.1.
+	\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{plotdraft}
+	\caption{}\label{fig2}
+\end{figure}
+In that sense both problems will lead to the same solution.
+\section*{Problem 5}
+Consider the exchange economy depicted in Figure 15.B.10(a). Show that the initial
+endowment allocation can be supported as a part of a quasi-equilibrium with transfers.
+Describe the corresponding price vector and wealth distribution. Is this a full equilibrium
+with transfers?
+
+\textbf{Solution}
+
+A feasible allocation $(x^*, y^*)$, a price vector $p$ and wealth levels $w = (w_1\ \dots\ w_I)$ are said to be a quasi-equilibrium with transfers if:
+\begin{enumerate}[(i)]
+	\item (Profit maximization condition) $\forall\ y_j\ \in\ Y_j\ \to py_j^* \ge py_j\ \forall\ j = 1, \dots, J$
+	\item (Market clearing condition) $\sum_{i=1}^I x^*_i = \bar{\omega} + \sum_{j=1}^Jy^*_j$
+	\item $\forall\ x_i\ \in\ X_i: x_i \succ_i x_i^*\ \to px_i \ge w_i\ \forall\ i$ (and $px_i^* \le w_i\ \forall\ i$)
+	\item (Balanced budget condition) $\sum_{i=1}^Iw_i = p\bar{\omega} + \sum_{j=1}^J py_j$
+\end{enumerate}
+In the exchanged economy there is no any production hence we should not care about the profit maximization condition.
+Assume the economy 15.B.10(a) (see Fig ) and let us prove that the initial endowment allocation $x^*$ can indeed be supported as the quasi-equilibrium.
+Assume $p = (1\ 0)$ and $w_1 = px_1^* = 0, w_2 = px^*_2 = \bar{\omega_1}$.
+Balanced budget condition holds, since $w_1 + w_2 = \bar{\omega_1} = (1\ 0) \cdot (\bar{\omega_1}\ \bar{\omega_2})^T = \bar{\omega_1}$.
+As well as market clearing condition.
+It remains to show that (iii) condition holds either.
+To do so, observe that the consumer 1 strictly prefers each point with positive amount of good 1 to $x_1^*$ hence $\forall\ x_1\ \in X_1: x_1 \succ_1 x_1^*\ \to px_1 = x_{11} \ge w_1 = 0$.
+Similarly, the second consumer strictly prefers each point with greater amount of good 1 (than $\bar{\omega_1}$) to $x_2^*$ which means that $\forall\ x_2 \succ_2 x_2^*\ \to\ px_2 = x_{12} \ge \bar{\omega_1}$.
+Thus, all conditions hold and the point is indeed quasiequilibrium with transfers.
+
+However it is not a full equilibrium with transfers. Because in the definition of full equilibrium with transfers instead of (iii) condition is the following:
+\begin{align*}
+x_i^*\ \text{maximizes } \succsim_i\ \text{ over }B(p) = \left\{x_i \in X_i: px_i \le w_i \right\}\ \forall\ i = 1, \dots, I
+\end{align*}
+For the 1st consumer his budget set with such a price vector is:
+\begin{align*}
+\left\{(x_{11}, x_{21})\in X_1:\ (1\ 0)\cdot(x_{11}, x_{21})^T = x_{11} \le w_1 = 0\right\} \equiv \left\{(x_{11}, x_{21})\in X_1:\ x_{11} = 0 \right\}
+\end{align*} 
+Of course $x_1^*$ does not maximizes hit utility on the aforementioned set: increase of $x_{21}$ could make him strictly better off still remainig in the budget set.
+\section*{Problem 6}
+\begin{enumerate}[(i)]
+	\item Assume by contradiction $\exists\ $ good $l$: $p_l < 0$.
+By strict monotonicity of preferences $\forall\ \varepsilon >0\ \ \hat{x_i} = x_i^* + \mathbb{I}_l\varepsilon \succ_i x_i^*$ (where $\mathbb{I}_l$ is a vector of zeros with 1 as a $l$-th entry).
+However
+	\begin{align*}
+	p\hat{x_i} = px_i^* + p_l\varepsilon < px_i^* \le w_i
+	\end{align*}
+	As a result 
+	\begin{align*}
+	\exists\ \hat{x_i}: \hat{x_i} \succ x_i^*: p\hat{x_i} \le w_i
+	\end{align*}
+	which violates the definition of quasi equilibrium that means that all entried of p-vector are nonnegative.
+	\item By the balanced budget condition:
+	\begin{align*}
+	\sum_{i=1}^I w_i = p \bar{\omega} + \sum_{j=1}^J py_j^*\\
+	\sum_{i=1}^I w_i = p \left(\bar{\omega} + \sum_{j=1}^J y_j\right) > 0
+	\end{align*}
+	Since $p \ge 0$ and $\left(\bar{\omega} + \sum_{j=1}^J y_j\right) >> 0$. As a result $\sum_{i=1}^I w_i > 0$ hence there exists at least one agent whose wealth is positive.
+	\item A buget set of the consumer $i_0$ is:
+	\begin{align*}
+	B(p, w_{i_0}) = \left\{x_{i_0} \in X_{i_0}: px_{i_0} \le w_{i_0} \right\}
+	\end{align*}
+	Assume that $x_{i_0} \succ_{i_0} x^*_{i_0}$ and $x_{i_0} \in B(p, w_{i_0})$.
+By the strictly monotonicity of preferences the only case which is possible is $px_{i_0} = w_{i_0} > 0$.
+By continuity of preferences both sets:
+	\begin{align*}
+	U = \left\{x_i \in X_i: x_i \succsim_i x_i^* \right\}\\
+	L = \left\{x_i \in X_i: x_i \precsim_i x_i^* \right\}
+	\end{align*}
+	are closed
+	\item Assume by contradiction that there exists some good $l$ such that $p_l \le 0$.
+Consider $\varepsilon > 0$ and $\hat{x_{i_0}} = x_{i_0}^* + \mathbb{I}_l\varepsilon$.
+By the strict monotonicity $\hat{x_{i_0}} \succ x_{i_0}^*$ but $p\hat{x_{i_0}} = px_{i_0}^* + p_l\varepsilon \le px_{i_0}^* \le w_{i_0}$ which contradict to the fact that $x_{i_0}^*$ is the preferences maximizer over the budget set.
+Thus, $p_l > 0\ \forall\ l = 1, \dots, L$.
+	\item $w_i \ge 0, \forall\ i$ because otherwise there cannot exist such an allocation $x_i$ that for $p >> 0$, $px_i \le w_i$ hence there cannot exist any quasi-equilibria which contradicts to the set up of the problem.
+Then, if $w_i = 0$ then for $p >> 0$ the budget set contains only $x_i^* = 0$ and this point maximizes preferences over such a budget set.
+In case $w_i > 0$ see (iii).
+AS a result, a quasi-equilibrium is in fact a full equilibrium.
+	\item Without strong monotonicity we would not be able to conclude that $p >> 0$ (actually solution to the problem 5 demonstrate the case when p (as a vector) is not strictly positive (i.e.
+contains zero elements)).
+Hence strong monotonicity indeed plays crucial role for the equivalence between notions of quasi- and full equilibria.
+\end{enumerate}
 \end{document}

### 2019-01-26 micro3_3.tex
Micro

@@
 By the strictly monotonicity of preferences the only case which is possible is $px_{i_0} = w_{i_0} > 0$.
-By continuity of preferences both sets:
-	\begin{align*}
-	U = \left\{x_i \in X_i: x_i \succsim_i x_i^* \right\}\\
-	L = \left\{x_i \in X_i: x_i \precsim_i x_i^* \right\}
-	\end{align*}
-	are closed
+Obviously (by strictly monotonicity of preferences) for each positive $\varepsilon$, $x_{\varepsilon} = (1-\varepsilon)x_{i_0} \prec_i x_{i_0}$ and $px_{\varepsilon} < w_{i_0}$.
+Let us prove that there exists such a small $\varepsilon$ that $x_{i_0}^* \prec_{i_0} x_{\varepsilon} \prec_{i_0} x_{i_0}$.
+Indeed, assume by contradiction that $\forall\ \varepsilon > 0\ x_{i_0}^* \succsim x_{\varepsilon}$.
+By continuity of preferences $\lim_{\varepsilon \to 0} x_{\varepsilon} = x_{i_0} \precsim_{i_0} \lim_{\varepsilon \to 0} x_{i_0}^* = x_{i_0}^*$ which contradicts to the fact that $x_{i_0} \succ_{i_0} x_{i_0}^*$.
+Thus, $\exists\  x_{\varepsilon}: x_{\varepsilon} \succ_{i_0} x_{i_0}^*\text{ and } px_{\varepsilon} < w_{i_0}$.
+The contradiction to the definition of quasi-equilibrium demonstrates that $x_{i_0}^*$ indeed maximizes preferences over the budget set.
 	\item Assume by contradiction that there exists some good $l$ such that $p_l \le 0$.
@@
 In case $w_i > 0$ see (iii).
-AS a result, a quasi-equilibrium is in fact a full equilibrium.
-	\item Without strong monotonicity we would not be able to conclude that $p >> 0$ (actually solution to the problem 5 demonstrate the case when p (as a vector) is not strictly positive (i.e.
+As a result, a quasi-equilibrium is in fact a full equilibrium.
+	\item Without strong monotonicity we would not be able to conclude that $p >> 0$ (actually solution to the problem 5 demonstrates the case when $p$ (as a vector) is not strictly positive (i.e.
 contains zero elements)).

### 2019-01-26 gameth1.tex
Gametheory

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Game theory. Problem Set 1.
\end{center}
\section*{Problem 1}
John and Peter are bidding for an indivisible object. The rules of the
auction are as follows. The individual who places the highest bid wins the
object. If John wins he pays the amount of money Peter bade. If Peter
wins he pays the amount of money he bade. Both players place their bids
simultaneously. For each individual $i \in \left\{J, P\right\}$ let $v_i \ge 0$ denote their
valuation for the object and let $b_i \ge 0$ denote the bid they placed. Their
preferences are captured by the following payoff functions for John and
Peter respectively:
\begin{align*}
u_J(b_J, b_P) = \begin{cases}
v_J - b_P, &b_J > b_P\\
\frac{1}{2}(v_J - b_P), &b_J = b_P\\
0, &b_J < b_P
\end{cases}\\
u_P(b_J, b_P) = \begin{cases}
v_P - b_P, &b_P > b_J\\
\frac{1}{2}(v_P - b_P), &b_P = b_J\\
0, &b_P < b_J
\end{cases}
\end{align*}
The information above (including valuations) is common knowledge. Does
either player have a weakly dominant strategy to bid his valuation? Can
you find values for the parameters $(v_J , v_P ) \in \mathbb{R}^2_+$ such that the pair of
strategies $(b_J , b_P ) = (v_J , v_P)$ constitutes a Nash Equilibrium? For each
answer you give provide the relevant demonstration.

\textbf{Solution}

Suppose that Peter's strategy is to bid his valuation $v_P$. Then
\begin{align*}
u_P(b_J, v_P) = 0 < u_P\left(b_J, \frac{b_J + v_P}{2}\right) = \frac{v_P - b_J}{2}, \text{for }b_J < v_P\\
\end{align*}
hence such a strategy cannot be (weakly) dominant for Peter.
Suppose that John's strategy is to bid his valuation $v_J$. Then
\begin{align*}
\begin{cases}
u_J(v_J, b_P) = 0 \ge u_J(b_J, v_P)\ \forall\ b_J , \ \ \ &b_P \ge v_{J}\\
u_J(v_J, b_P) = v_J - b_P \ge u_J(b_J, v_P)\ \forall\ b_J, \ \ \ &b_P < v_J
\end{cases}
\end{align*}
Thus, for John, the strategy to bid his valuation is indeed weakly dominant strategy.
Assume $v_J \ge v_P$ then the pair of strategies $(b_J, b_P) = (v_J, v_P)$ indeed constitutes a Nash equilibrium. Let us prove it. Denote $r_J(b_P), r_P(b_J)$ best responses on the opponent's bid of John and Peter respectively. Obviously:
\begin{align*}
r_J(b_P) &= \begin{cases}
[0, b_P), b_P > v_J\\
[0, b_P], b_p = v_J\\
(b_P, +\infty), b_P < v_J
\end{cases}\\
r_P(b_J) &= \begin{cases}
[0, b_J), b_J > v_P\\
[0, b_J], b_j = v_P\\
b_J + \varepsilon, b_J < v_P
\end{cases}
\end{align*}
Then 
\begin{align*}
\begin{cases}
r_J(v_P) = (v_P, +\infty) \ni v_J \text{ and } r_P(v_J) = [0, v_J) \ni v_P, \ \ \ &v_J > v_P\\
r_J(v_P) = [0, v_P] \ni v_J \text{ and } r_P(v_J) = [0, v_J] \ni v_P, \ \ \ &v_J = v_P
\end{cases}
\end{align*} 
Thus, for each player his strategy belongs to his best response on rival's playing this strategy, so, it is a Nash equilibrium.
If $v_J < v_P$ then $r_J(v_P) = (0, v_P) \ni v_J$ however $r_P(v_J) = v_J + \varepsilon < v_P$ (we assume $\varepsilon$ is sufficiently small). Hence for $v_J < v_P$, $(b_J, b_P) = (v_J, v_P)$ is not a Nash equilibrium.
\section*{Problem 2}
Prove the following statement
In a finite simultaneous move game representable in normal form,
\begin{align*}
\Gamma = (I, \times_{i \in I}S_i, (u_i)_{i \in I})
\end{align*}
if some pure strategy $s_i \in S_i$
is strictly dominated by some $\sigma_i \in \Delta(S_i)$, then
for any Nash Equilibrium of the game $\sigma^* \in \times_{i \in I}\Delta(A_i)$, we have $\sigma^*_i(s_i) = 0$
(that is, the equilibrium strategy of player $i$ assigns zero probability to pure
strategy $s_i$).
Hint: Proceed by contradiction. As a first step, show that $\sigma^*_i
(a_i) \neq 1$.

\textbf{Solution}

Assume $\sigma^* \in \times_{i \in I}\Delta(S_i)$ is a Nash equilibrium. And assume that for a player $i$, $\sigma_i^*$ assigns a positive probability to the pure strategy $s_j$ which is strictly dominated by the strategy $\sigma_i$. Then 
\begin{align*}
&\exists\ \alpha_1, \dots, \alpha_{|S_i|}: \alpha_k \in [0, 1]\ \forall\ k, \sum_{k=1}^{|S_i|}\alpha_k = 1\text{ and }\\
 u_i(\sigma^*_i, \sigma^*_{-i}) = &\alpha_1 u_1(s_1, \sigma^*_{-i}) + \dots + \alpha_j u_i(s_j, \sigma^*_{-i}) + \dots + \alpha_{|S_i|}u_i(s_{|S_i|}, \sigma^*_{-i}),\ \  \alpha_j > 0
\end{align*}
Since $\sigma_i$ strictly dominates $s_j$ then
\begin{align*}
&\exists\ \beta_1, \dots, \beta_{|S_i|}:\beta_k \in [0, 1]\ \forall\ k,\ \sum_{k=1}^{|S_i|}\beta_k = 1\text{ and }
\forall\ \sigma_{-i} \in \times_{l \in I\setminus \left\{i\right\}}\Delta(S_l)\\
u(s_j, \sigma_{-i}) < &u(\sigma_i, \sigma_{-i}) = \beta_1 u_i(s_1, \sigma_{-i}) + \dots + \beta_j u_i(s_j, \sigma_{-i}) + \dots + \beta_{|S_i|}u_i(s_{|S_i|}, \sigma_{-i})
\end{align*}
But then
\begin{align*}
&u_i(\sigma_i^*, \sigma^*_{-i}) < \alpha_1u_1(s_1, \sigma^*_{-i}) + \dots + \alpha_j \left(\sum_{k=1}^{|S_i|}\beta_k u_i(s_k, \sigma^*_{-i})\right) + \dots + \alpha_{|S_i|}u_i(s_{|S_i|}, \sigma^*_{-i}) = \\
&= (\alpha_1 + \alpha_j\beta_1)u_i(s_1, \sigma^*_{-i}) + \dots + (\alpha_j + \alpha_j\beta_j)u_i(s_1, \sigma^*_{-i}) + \dots + (\alpha_{|S_i|} + \alpha_{|S_i|}\beta_{|S_i|})u_i(s_{|S_i|}, \sigma^*_{-i}) = \\
&=u(\hat{\sigma_i}, \sigma^*_{-i})
\end{align*}
Since $\exists\ t \in 1, \dots, |S_i|: \beta_{t} > 0\ \to \alpha_t + \alpha_j \beta_t \neq \alpha_t$ hence there exists another strategy $\hat{\sigma_i} \neq \sigma^*_i$ which gives strictly greater payoff than $\sigma^*_i$ hence $\sigma^*$ is not a Nash equilibrium, contradiction. This contradiction shows that indeed $\alpha_j = 0$ i.e. the euilibrium strategy must assign $0$ probability to the strictly dominated strategy.
\end{document}

### 2019-01-27 metrics2.tex
Metrcis

@@
 \begin{align*}
-\expect \tilde{\beta_1} = \beta_1 + \beta_3 \frac{\sum_{i=1}^n \hat{r_{i1}x_{i3}}}{\sum_{i=1}^n \hat{r_{i1}}^2}
+\expect \tilde{\beta_1} = \beta_1 + \beta_3 \frac{\sum_{i=1}^n \hat{r_{i1}}x_{i3}}{\sum_{i=1}^n \hat{r_{i1}}^2}
 \end{align*}
@@
 \textbf{Solution}
-Plugging $y_i = \beta_0 + \beta_1x_{i1} + \beta_2x_{i2}+\beta_3x_{i3}$ into
+
+Following hint, plugging $y_i = \beta_0 + \beta_1x_{i1} + \beta_2x_{i2}+\beta_3x_{i3}$ into
 \begin{align*}
@@
 \begin{enumerate}[(a)]
-	\item First of all $x_i \neq 0\ \forall i$ then the estimator is well-defined.
+	\item If $\exists\ i: x_i \neq 0$ then the estimator is well-defined.
 	\item \begin{align*}
@@
 \begin{enumerate}[(a)]
-	\item An idealized randomized controlled experiment should be as follows: we should absolutely randomly choose two groups of people.
-Absolutely randomly in the sense that each person of the USA could equally likely be chosen for this experiment.
-People of the first group should be given an additional year of education, while people from the second group should be abstained from the education over the year.
-Then after this year
+	\item Since the education is a sequential process i.e.
+to get an additional year of university one should primarily finish a school.
+That means that in order to estimate the effect of $i$-th additional year at school we should randomly choose two groups of children randomly in the sense that each child in the USA should be equally likelly chosen in these groups.
+Then these children should be randomly distributed among schools in the USA and study there $i-1$ years.
+Then children from the first group should be expelled from school and finish their education.
+Children from the second group should be allowed to study one more additional year and after that should also be expelled from the school.
+Then at some age of those people we should measure average salaries in both groups and compare these means using some statistical tests.
+	\item This protocol actually eliminates omitted variable bias because it eliminates correlation between additional year of education and any other factors which can influence earnings.
+To be more precise, let assume that earnings are described as follows $y = \beta_0 + \beta_1 \text{\textit{yrseduc}} + \beta_2 \text{\textit{intel.ability}} + u$.
+But in our controlled randomized experiment intelectual abilities of people and years of education are independent variables.
+Hence if we somehow measured intelectual abilities and regress them on \textit{yrseduc} in our experiment we would get something like on the Fig \ref{fig2}.
+		\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.8\textwidth]{iqyrs}
+		\caption{}\label{fig2}
+	\end{figure}
+i.e.
+estimated coefficient $\delta$ would be close to zero.
+But estimated coefficient from regression $y$ on \textit{yrseduc} is equal to $\hat{\beta_1} + \delta \hat{\beta_2}$ where $\hat{\beta_1}, \hat{\beta_2}$ are OLS estimates on hypothetical regression $y$ on \textit{yrseduc} and intelectual abilities.
+THe same thing with any other facctors which can in principle affect earnings.
+Thus, estimated value of additional school year becomes unbiased.
+Of course, such kind of experiment is not possible in reality because of ethical issues.
 \end{enumerate}
@@
 confidence interval for the marginal value of a bachelor’s degree.
+
+\textbf{Solution}
+
+The summary of regression is depicted below.
+	\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{ba}
+	\caption{}\label{fig1}
+\end{figure}
+
+Using confint() one can get the following 95\% confidence interval for the marginal value of a bachelor degree:
+\begin{align*}
+(-0.1332442; 1.705304)
+\end{align*}
+This interval include 0, hence we cannot infer that for people whose number of years of education are equal, obtainment of bachelor degree on average leads to a higher salary.
 \end{document}

### 2019-01-27 Metrics2.R
Metrcis

@@ -18,3 +18,6 @@ coeftest(fit2, vcov = vcovHC(fit2, type = "HC1"))
 summary(fit1)
 
 #Problem 8
+fit3 <- lm(df$ahe ~ df$yrseduc + df$ba, data = df)
+summary(fit3)
+confint(fit3, vcov = vcovHC(fit3, type = "HC1"))

### 2019-01-27 gameth1.tex
Gametheory

@@
 the euilibrium strategy must assign $0$ probability to the strictly dominated strategy.
+\section*{Problem 3}
+There are $n < +\infty$ firms in an industry. Each can try to convince Congress
+to give the industry a subsidy. Let $h_i$ denote the hours of effort put by
+some industry $i \in \left\{1,\dots , n\right\}$. Let $c(h_i) = w_ih_i^2$, with $w_i > 0$, be the cost of
+this effort to firm $i$. For each profile of efforts $(h_1, \dots , h_n) \in \mathbb{R}^n_+$, the value
+of the subsidy that gets approved equals
+
+\begin{align*}
+\alpha \sum_{i \in \left\{1, \dots, n\right\}} h_i + \beta \left(\prod_{i \in \left\{1, \dots, n\right\}} h_i\right)
+\end{align*}
+where $\alpha, \beta \ge 0$. The approved subsidy is shared equally among the firms in
+the industry. Suppose that all firms choose the effort level simultaneously.
+Show that each firm has a strictly dominant strategy if and only if $\beta = 0$.
+What is firm $i$’s strictly dominant strategy when this is so?
+
+\textbf{Solution}
+
+Each firm's payoff function is:
+\begin{align*}
+u_i(h_1, \dots, h_i, \dots, h_n) = \alpha \cdot \frac{h_1 + \dots h_i + \dots + h_n}{n} + \beta h_1\times\dots\times h_i \times \dots\times h_n - w_ih_i^2
+\end{align*}
+
+hence to find a strictly dominant strategy one need to find $h_i^*$:
+\begin{align*}
+u_i(h_i^*, h_{-i}) > u_i(h_i, h_{-i})\ \forall\ h_{-i}, \forall\ h_i
+\end{align*}
+Suppose $\beta = 0$ then 
+\begin{align*}
+u_i(h_i, h_{-i}) = \alpha\frac{h_i}{n} + \alpha\frac{h_1 + \dots+h_{i-1}+h_{i+1}+\dots+h_n}{n} - w_ih_i^2
+\end{align*}
+it is a quandratic function in $h_i$ and regardless of what $h_j\ \forall\ j \in\left\{1, \dots, n\right\}\setminus \left\{i\right\}$ is it attains maximum at $h_i = \frac{\alpha}{2nw_i}$.
+Thus, strategy $h_i = \frac{\alpha}{2nw_i}$ is a strictly dominant strategy for each firm in case $\beta = 0$.
+Now assume conversly that each firm has a strictly dominant strategy $h_i^*$.
+Then
+\begin{align*}
+u_i(h_i^*, h_{-i}) > u_i(h_i, h_{-i})\ \  \forall\ h_i, h_{-i}
+\end{align*}
+Suppose by contradiction that $\beta > 0$.
+Let us denote $\prod_{k \in \left\{1, \dots, n\right\}\setminus\left\{i\right\}}h_k$ as $H_{-i}$ then if for some $h_{-i}$ - profile of strategies of other players
+\begin{align*}
+u_i(h_i^*, h_{-i}) > u_i(h_i, h_{-i})\ \forall\ h_i \neq h_{i}^*
+\end{align*}
+then $h_{i}^*$ is the extremum of the payoff function, given $H_{-i}$, i.e.
+\begin{align*}
+h_i^* = \frac{\alpha}{2nw_i} + \frac{\beta}{2nw_i}H_{-i}
+\end{align*}
+but now suppose $\forall j = 1, \dots, i-1, i+1, \dots, n$, $\hat{h_j} = \varepsilon^{\frac{1}{n-1}}h_j$ for some positive $\varepsilon$. Then
+\begin{align*}
+\exists\ \hat{h_i} = \frac{\alpha}{2nw_i} + \frac{\varepsilon \beta}{2nw_i}H_{-i} \neq h^*_i: u_i(\hat{h_i}, \hat{h_{-i}}) \ge u_i(h_i^*, \hat{h_{-i}})
+\end{align*}
+Because the value of the function at its maximum is always greater or equal than in any other point.
+That means that $h_i^*$ is not a strictly dominant strategy.
+This contradiction shows that $\beta = 0$.
+Q.E.D.
+\section*{Problem 4}
+Prove the following statement
+In a finite simultaneous move game representable in normal form,
+\begin{align*}
+\Gamma =
+(I, \times_{i \in I}S_i,(u_i)_{i\in I})
+\end{align*}
+if a single profile of pure strategies survives the process of iterated deletion
+of strictly dominated strategies, then this profile must be the unique Nash
+equilibrium of the game
+
+\textbf{Solution}
+
 \end{document}

### 2019-01-27 gameth1.tex
gametheory

@@
 Prove the following statement
+
 In a finite simultaneous move game representable in normal form,
@@
 
+First of all let us prove that the survived strategy is indeed a Nash equilibrium, after that let us prove that none of Nash equilibria can be deleted over the iterated deletion process.
+
+Assume by contradiction that the profile is not a Nash equilibrium, i.e.
+for some player $i$, $s_i^*$ is not a best response on $s^*_{-i}$.
+That means that the best response $\hat{s_i}$ has been deleted through the process of iterated deletion.
+Let us consider the step when it occured.
+To be deleted there must exist a strategy $\tilde{s_i}$ such that
+\begin{align*}
+\forall\ s_{-i} \text{ which has survived until this step and for } s^*_{-i} \text{ in particular }\to u_i(\tilde{s_i}, s_{-i}) > u_i(\hat{s_i}, s_{-i})
+\end{align*}
+but that means that $\hat{s_i}$ is not a best response on $s^*_{-i}$. The contradiction shows that $s^*_{i}$ is indeed a best response for each player $i$, hence the profile is a Nash equilibrium.
+
+It remains to show that none of Nash equilibria can be deleted through the process of iterated deletion.
+Assume by contradiction that some Nash equilibrium $(s_1^*, \dots, s_I^*)$ has been deleted, wlog that means that for some player $i$, $s_i^*$ has been deleted as a strictly dominated strategy (while $s_{-i}^*$ has still not be deleted).
+That means that
+\begin{align*}
+\exists\ \hat{s_i} \in S_i:\ \forall\ s_{-i} \text{ and for } s_{-i}^* \text{ in particular }\to u_i(\hat{s_i}, s_{-i}) > u_i(s_i^*, s_{-i})
+\end{align*}
+but that means that $s_{i}^*$ cannot be a best response on $s_{-i}^*$ hence it cannot be part of the Nash equilibium. This contradiction shows that none of Nash equilibria can be deleted. Q.E.D.
 \end{document}

### 2019-01-29 micro3_3.tex
Micro

@@
 \begin{align*}
-\underset{z_{11}, z_{21}, z_{12}, z_{22} \ge 0}{\max}\ \lambda f_1(z_{11}, z_{21}) + (1 - \lambda)f_2(z_{12}, z_{22})\\
+\underset{z_{11}, z_{21}, z_{12}, z_{22} \ge 0}{\max}\ \lambda_1 f_1(z_{11}, z_{21}) + \lambda_2f_2(z_{12}, z_{22})\\
 s.t. z_{11} + z_{12} = \bar{z_1}\\
@@
 
+In both problems technological constraints are the same hence FOCs wrt $y_j$ will be similar in both problems either.
+
 Problem 16.F.1:
@@
 If utility functions are concave then the utility possibility set is a convex set.
-Obviously for each particular $(\bar{u_2}, \dots, \bar{U_I})$ the solution to the problem 16.F.1.
-should lie on the frontier (see Fig.), but when the UPS is a convex set by the separating hyperplane theorem there should exists such a non-zero $\lambda = (\lambda_1\ \dots\ \lambda_I)$ such that is is a solution to the problem 16.E.1.
+Obviously for each particular $(\bar{u_2}, \dots, \bar{u_I})$ the solution to the problem 16.F.1.
+should lie on the frontier (see Fig.
+\ref{fig2}), but when the UPS is a convex set by the separating hyperplane theorem there should exists such a non-zero $\lambda = (\lambda_1\ \dots\ \lambda_I)$ such that it is a solution to the problem 16.E.1.
 	\begin{figure}[h]
@@
 In that sense both problems will lead to the same solution.
+Put it another way, both problems have identical FOCs, but only under assumption of concavity these FOCs become \textit{sufficient} conditions for extremum.
 \section*{Problem 5}

### 2019-01-30 metrics3.tex
Metrics

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\newcommand{\yrseduc}{\textit{yrseduc}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Econometrics 1. Problem Set 3.
\end{center}
\section*{Problem 1}

\textbf{Solution}

\begin{enumerate}[(a)]
	\item 
	\begin{align*}
	t = \frac{-0.0023}{0.0022} \approx -1
	\end{align*}
	p-value is about $0.2$, consequently we have no reasons for rejecting the hypothesis that \textit{profitmarg} does not affect salary (i.e. slope coefficient on this variable is equal to zero) on the 5\% significance level.
	\item Similarly,
	\begin{align*}
	t_{2\text{nd col}} = 2.24,\ t_{3\text{nd col}} = 2.04
	\end{align*}
	in both cases the hypothesis that corresponding coeeficient is equal to zero should be rejected on the 5\% significance level. Since
	\begin{align*}
	\frac{\Delta \text{salary}}{\text{salary}} = 0.1 \cdot \frac{\Delta \text{mktval}}{\text{mktval}}\\
	100 \cdot \frac{\Delta \text{salary}}{\text{salary}} = 0.1 \frac{\Delta \text{mktval}}{\text{mktval}} \cdot 100
	\end{align*}
	that means that once \textit{mktval} changes by 1\% holding other thing equal, the average salary will change by 0.1\%.
	\item 
	\begin{align*}
	t_{\text{ceoten}} = \frac{0.0171}{0.0055} = 3.11\\
	\end{align*}
	hence on the 5\% significance level the coefficient is not equal to 0.
	Since
	\begin{align*}
		\frac{\Delta \text{salary}}{\text{salary}} = 0.0171 \Delta \text{ceoten}\\
		100 \cdot \frac{\Delta \text{salary}}{\text{salary}} = 1.71 \Delta \text{ceoten}
	\end{align*}
	that is one more year as a CEO with the current company increases the average salary by 1.71\% other thing equal.
	\begin{align*}
	t_{\text{comten}} = -\frac{0.0092}{0.0033} = 2.78\\
	\end{align*}
	hence on the 5\% significance level the coefficient is not equal to 0.
	Since
	\begin{align*}
	\frac{\Delta \text{salary}}{\text{salary}} = -0.0092 \Delta \text{ceoten}\\
	100 \cdot \frac{\Delta \text{salary}}{\text{salary}} = 0.92 \Delta \text{ceoten}
	\end{align*}
	that is one more year with the current company decreases the average salary by 0.92\% other thing equal.
\end{enumerate}

\section*{Problem 2}
\begin{enumerate}[(a)]
	\item \begin{align*}
	&y_i = \beta_0 + \beta_1 x_{i} + u\\
	&y_i = \gamma_0 + \gamma_1(x_i - \bar{x}) + v = \gamma_0 + \gamma_1 \tilde{x_i} + v\\
	&t_{\beta_1} = \frac{\hat{\beta_1}}{SE(\hat{\beta_1})},\  t_{\gamma_1} = \frac{\hat{\gamma_1}}{SE(\hat{\gamma_1})}
	\end{align*}
	\begin{align*}
	\hat{\beta_1} = \frac{\summa (x_i - \bar{x})y_i}{\summa (x_i - \bar{x})^2} = \frac{\summa \tilde{x_i} y_i}{\summa \tilde{x_i}^2} = \hat{\gamma_1}
	\end{align*}
	\begin{align*}
	SE(\hat{\beta_1}) = \frac{\sqrt{\summa \hat{u_i}^2}}{\sqrt{\summa (x_i - \bar{x})^2}}\\
	SE(\hat{\gamma_1}) = \frac{\sqrt{\summa \hat{v_i}^2}}{\sqrt{\summa (\tilde{x_i})^2}}\\
	\end{align*}
	Thus, the only difference in both standard errors is the difference in sum of squares of residuals. To that end in order to answer the question one need to compare these sums.
	\begin{align*}
	&\hat{u_i} = y_i - \hat{\beta_0} - \hat{\beta_1}x_{i}\\
	&\hat{v_i} = y_i - \hat{\gamma_0} - \hat{\gamma_1}\tilde{x_i} = y_i - \hat{\gamma_0} - \hat{\gamma_1}(x_i - \bar{x})
	\end{align*}
	Since $\hat{\beta_1} = \hat{\gamma_1}$, $\hat{u_i}$ and $\hat{v_i}$ differs by the term $\hat{\gamma_0} - \hat{\gamma_1}\bar{x} - \hat{\beta_0}$.
	\begin{align*}
	&\hat{\gamma_0} = \bar{y} - \hat{\gamma_1}\bar{\tilde{x}} = \bar{y}\\
	&\hat{\beta_0} = \bar{y} - \hat{\beta_1}\bar{x}\\
	&\hat{u_i} - \hat{v_i} = \hat{\beta_1}\bar{x} - \hat{\gamma_1}\bar{x} = 0
	\end{align*}
	Thus, t-statistics are equal.
	\item
	\begin{align*}
	Var \begin{pmatrix}
	\hat{\beta_0}\\
	\hat{\beta_1}
	\end{pmatrix} = \begin{pmatrix}
	2 & 1\\
	1 & 3
	\end{pmatrix}
	\end{align*}
	Since
	\begin{align*}
	\hat{\beta_1} = \beta_1 + \frac{\summa (x_i - \bar{x})u_i}{\summa (x_i - \bar{x})^2}\\
	\end{align*}
	\begin{align*}
	Var (\hat{\beta_1}|x_1, \dots, x_n) \underset{\text{i.i.d}}{=} \frac{\sigma^2_u}{\summa (x_i - \bar{x})^2}
	\end{align*}
	\begin{align*}
	&Var(\hat{\beta_0}|x_1, \dots, x_n) = Var(\bar{y} - \hat{\beta_1}\bar{x}|x_1, \dots, x_n) = Var(\bar{y}|x_1, \dots, x_n) + \bar{x}^2Var(\hat{\beta_1}|x_1, \dots, x_n) -\\
	&- 2\bar{x}cov(\bar{y}, \hat{\beta_1}|x_1, \dots, x_n)
	\end{align*}
	\begin{align*}
	&cov(\bar{y}, \hat{\beta_1}|x_1, \dots, x_n) = \frac{1}{n\summa (x_i - \bar{x})^2} cov \left(\summa y_i, \sum_{j=1}^n (x_j - \bar{x})y_j\bigg|x_1, \dots, x_n\right) = \\
	&=\frac{1}{n\summa (x_i - \bar{x})^2} \sum_{j=1}^n(x_j - \bar{x})\summa cov(y_j, y_i) = \frac{\sigma^2_u \summa (x_i - \bar{x})}{n\summa(x_i - \bar{x})} = 0
	\end{align*}
	Thus, 
	\begin{align*}
	Var(\hat{\beta_0}|x_1, \dots, x_n) = \frac{\sigma^2_u}{n} + \frac{\bar{x}^2\sigma^2_u}{\summa(x_i - \bar{x})^2} = \sigma^2_u\frac{n\bar{x}^2 + \summa(x_i - \bar{x})^2}{n\summa(x_i - \bar{x})^2} = \frac{\summa x_i^2}{n\summa(x_i - \bar{x})^2}\sigma_u^2
	\end{align*}
	\begin{align*}
	cov(\hat{\beta_0}, \hat{\beta_1}|x_1, \dots, x_n) = cov(\bar{y} - \hat{\beta_1}\bar{x}, \hat{\beta_1}|x_1, \dots, x_n) = 0 - \bar{x}Var(\hat{\beta_1}|x_1, \dots, x_n) =\\ -\frac{\bar{x} \sigma^2_u}{\summa (x_i - \bar{x})^2}
	\end{align*}
	To obtain the estimates of variances one should plug $\hat{\sigma_u}^2$ instead of $\sigma^2_u$. That means that:
	\begin{align*}
	\begin{cases}
	\frac{\hat{\sigma^2_u}}{\summa (x_i - \bar{x})^2} = 2\\
	\frac{\summa x_i^2}{n\summa(x_i - \bar{x})^2}\hat{\sigma_u^2} = 3\\
	-\frac{\bar{x} \sigma^2_u}{\summa (x_i - \bar{x})^2} = 1
	\end{cases}\\
	\end{align*}
	Using the fact $\summa (x_i - \bar{x})^2 = \summa x_i^2 - n\bar{x}^2$ we get a system of three equations with three unknows, which can be easily solved, as a result:
	\begin{align*}
	\hat{\sigma^2_u} = 255\\
	\summa x_i^2 = 153\\
	\bar{x} = -\frac{1}{2}
	\end{align*}
	\item Assuming all variances, condition on $x_1, \dots, x_n$
	\begin{align*}
	Var(\hat{\beta_0} + \alpha \hat{\beta_1}) = Var\hat{\beta_0} + \alpha Var \hat{\beta_1} -2\alpha \text{cov}(\hat{\beta_0}, \hat{\beta_1}) =\\
	=\frac{153}{13005}\hat{\sigma_u^2} + \frac{\alpha^2 \hat{\sigma^2_u}}{127.5} + 2\alpha \frac{ \sigma^2_u}{255} \to \min
	\end{align*}
	\begin{align*}
	FOC: \frac{4}{255} \alpha^* + \frac{2}{255} = 0\\
	\alpha^* = -\frac{1}{2}
	\end{align*}
\end{enumerate}
\section*{Problem 3}
The result of regression is depicted below on the Fig. \ref{fig1}.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{lnahe}
	\caption{Summary of the regression}\label{fig1}
\end{figure}

\begin{enumerate}[(a)]
	\item First of all, all three coefficients are significant (on 5\% significance level). Of course it would be more correct to use coeftest() in order to get heteroskedasticity-robust standard errors, but p-value is such a small that this procedure do not change our conclusion about significance (only standard errors and t-statistics change slightly). Then, using logic similar to the Problem 1, we can infer that, other things equal each additional year of education increases the average per hour salary by 8\%, holding other things equal each additional year of life increases per hour earnings by 0.6\%, and finally, holding other things equal women get on average 22\% less per hour salary than men.
	\item RMSE is approximately equal to $0.465$.
	\item Confidence interval for the gender gap in earnings is
	\begin{align*}
	(-0.254035195; -0.194271848)
	\end{align*}
	i.e. with probability 0.95 the interval $(19\%; 25\%)$ cover the value which indicates by how much percent the average per hour salary for women is less than the one for men.
	\item Based on the summary of regression, the marginal value of the high school diploma is 8.3, i.e. holding gender and age fixed, receiving a high school diploma after 12 years of studying against 12 years without diploma increases average per hour earnings by 8.3\%. Using coeftest() we can conclude that on the 5\% significance level we reject the hypothesis that the marginal value of high school diploma is zero. 8.3\% is not a huge value, in fact, I suppose average the salary of people with high school diploma much more higher than those without diploma.
	\item 
	Estimating the following regression model
	\begin{align*}
	\ln(\text{ahe}) = \beta_0 + \beta_1 \text{hsdipl} + \beta_2 \text{yrseduc} + \beta_3 \text{age} + \beta_4\text{female} + u
	\end{align*}
	one can get:
	\begin{align*}
	\ln \text{ahe}_{12} - \ln \text{ahe}_{10} &= \beta_1 + 2 \beta_2\\
	\frac{\text{ahe}_{12} - \text{ahe}_{10}}{\text{ahe}_{10}} &= V = e^{\beta_1 + 2\beta_2} - 1 \\
	\hat{V} = e^{\hat{\beta_1} + 2\hat{\beta_2}} - 1 &\approx 0.285
\end{align*}
I think it is a rather reasonable estimate.
\item 
\begin{align*}
\prob (\alpha_1 < e^{\hat{\beta_1} + 2 \hat{\beta_2}} - 1 < \alpha_{2}) = 0.95\\
\prob (\ln(\alpha_1 + 1) < \hat{\beta_1} + 2\hat{\beta_2} < \ln(\alpha_2 +1)) = 0.95
\end{align*}
Thus, we need to compute a confidence interval for $\hat{\beta_1} + 2 \hat{\beta_2}$. To do so let us modify our model and estimate regression $\ln( \text{ahe})$ on $1$, \textit{hsdipl } $+ 1$, \textit{yrseduc } $+2$, \textit{age} and \textit{female}.
\end{enumerate}
\end{document}

### 2019-01-31 metrics3.tex
Metrics

@@
 	\includegraphics[width=0.8\textwidth]{lnahe}
-	\caption{Summary of the regression}\label{fig1}
+	\caption{Summary of regression}\label{fig1}
 \end{figure}
@@
 Thus, we need to compute a confidence interval for $\hat{\beta_1} + 2 \hat{\beta_2}$.
-To do so let us modify our model and estimate regression $\ln( \text{ahe})$ on $1$, \textit{hsdipl } $+ 1$, \textit{yrseduc } $+2$, \textit{age} and \textit{female}.
+To do so let us modify our model and estimate regression $\ln( \text{ahe})$ on \textit{hsdipl } $- 1$, \textit{yrseduc } $-2$, \textit{age} and \textit{female}.
+It is easy to observe that in that case the intercept will be $\beta_0 + \beta_1 + 2\beta_2$.
+Using confint() for this new intercept we can get the following:
+\begin{align*}
+\prob (1.375500454 &< \widehat{\beta_0 + \beta_1 + 2\beta_2} < 1.585969536) = 0.95\\
+\prob(0.1458 &< \hat{\beta_1} + 2\hat{\beta_2} < 0.3563) = 0.95
+\end{align*}
+Thus, $\alpha_1 = 0.157$, $\alpha_2 = 0.428$ and
+\begin{align*}
+\prob(0.157 < \hat{V} < 0.428) = 0.95
+\end{align*}
+\end{enumerate}
+\section*{Problem 4}
+The summary of regression is depicted below (see Fig. \ref{fig2})
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{square}
+	\caption{Summary of the regression}\label{fig2}
+\end{figure}
+
+\begin{enumerate}[(a)]
+	\item Sign of the coefficient on \textit{age} is +, while on \textit{age}$^2$ is -. Denote $y = \ln \text{ahe}, x_1 = \text{ age}, x_3 = \text{ female}, x_4 = \text{hsdipl}$. Since 
+	\begin{align*}
+	\ln y = \beta_0 + \beta_1x_1 + \beta_2x_1^2 + \beta_3x_3 + \beta_4x_4 +u
+	\end{align*}
+	that means that
+	\begin{align*}
+	\frac{d \ln y}{d x_1} = \beta_1 + 2\beta_2x_1\\
+	\frac{dy}{y} = \beta_1dx_1 + 2\beta_2 x_1 dx_1 \\
+	\frac{\Delta y}{y} \approx \beta_1\Delta x_1 + 2\beta_2 x_1 \Delta x_1
+	\end{align*}
+	Since $\beta_1 > 0$ while $\beta_2 < 0$ that means that: the older the person becomes - the less (on average) increase in salary each additional year brings.
+For example, if $x_1 = 20, \Delta x_1 = 1$ then $\frac{\Delta y}{y} \cdot 100 \approx 2.96 \%$, but if $x_1 = 40, \Delta x_1 = 1$ then $\frac{\Delta y}{y} \cdot 100 \approx 0.86 \%$.
+	\item The relation is linear $\iff$ $\beta_2 = 0$. Using coeftest() we reject this hypothesis on the 5\% significance level, i.e. the relation is quadratic.
+	\item The estimated gender gap has changed from -0.224 to -0.225, which is not a significant change in a real-world sense.
+Econometric reason is that first of all $\text{age}^2$ and $\text{female}$ are almost uncorrelated (sample correaltion coeeficient is about 0.007) and moreover slope coefficient on $\text{age}^2$ is also very small, as a result including $\text{age}^2$ barely changes the estimated gender gap.
+	\item Let us estimate the following regression
+	\begin{align*}
+	\ln y = \beta_0 + \beta_1x_1 + \beta_2x_1^2 + \beta_{23}x_1^3+ \beta_3x_3 + \beta_4x_4 +u
+	\end{align*}
+	the relation is cubic $\iff \beta_{23} = 0$. Using coeftest() we can conclude that on the 5\% significance level we cannot reject this hypothesis, i.e. the relation is quadratic.
+	\item Here we need to test the hypothesis $\beta_2 = 0, \beta_{23} = 0$. The output of linearHypothesis() is depicted below (see Fig. \ref{fig3}).
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.8\textwidth]{hyptest}
+		\caption{Hypothesis test}\label{fig3}
+	\end{figure}
+
+	Thus, the hypothesis can be rejected on the 5\% significance level. That is the relation is more likely a polynomial of up to cubic degree. I recomend the specification
+	\begin{align*}
+	\ln y = \beta_0 + \beta_1x_1 + \beta_2x_1^2 + \beta_3x_3 + \beta_4x_4 +u
+	\end{align*}
 \end{enumerate}

### 2019-01-31 metrics3.tex
Metrics3

@@
 	\end{align*}
-	the relation is cubic $\iff \beta_{23} = 0$. Using coeftest() we can conclude that on the 5\% significance level we cannot reject this hypothesis, i.e. the relation is quadratic.
+	the relation is quadratic $\iff \beta_{23} = 0$. Using coeftest() we can conclude that on the 5\% significance level we cannot reject this hypothesis, i.e. the relation is quadratic.
 	\item Here we need to test the hypothesis $\beta_2 = 0, \beta_{23} = 0$. The output of linearHypothesis() is depicted below (see Fig. \ref{fig3}).
@@
 	\end{align*}
+	first of all because this model is more interpretable than the model with cubic relationship, and secondly, we have significant statistic evidence that the relation is not cubic and not linear (based on (d) and (e)).
+	\end{enumerate}
+
+\section*{Problem 5}
+\begin{enumerate}[(a)]
+	\item 
+\begin{table}[h]
+	\begin{tabular}{cccc}
+		Specification & Age 30 to 31  & Age 45 to 46 & Age 60 to 61 \\
+		\hline
+		Linear in \textit{age}&  &  &  \\
+		Quadratic in \textit{age}  &  &  &  \\
+		Cubic in \textit{age} &  &  &\\
+		\hline 
+	\end{tabular}
+\end{table}
+
+For linear model the starting point does not matter hence all entries should be equal to each other and from the Problem 3 (a):
+\begin{align*}
+\frac{\Delta y}{y} \cdot 100 \% \approx 0.6 \Delta x_1
+\end{align*}
+For quadratic model, as it has been already mentioned above:
+\begin{align*}
+\frac{\Delta y}{y} &\approx \beta_1 \Delta x_1 + 2 \beta_2x_1\Delta x_1\\
+\frac{\Delta y}{y} \cdot 100 \% &\approx 5.07 \Delta x_1 -0.105 x_1 \Delta x_1 = \begin{cases}
+1.92, x_1 = 30\\
+0.34, x_1 = 45\\
+-1.23, x_1 = 60
+\end{cases}
+\end{align*}
+Similarly for cubic model:
+\begin{align*}
+\frac{\Delta y}{y} \approx \beta_1 \Delta x_1 + 2 \beta_2 x_1 \Delta x_1 + 3\beta_{23}x_1^2 \Delta x_1
+\end{align*}
+Using summary from the cubic regression we can obtain:
+\begin{align*}
+\frac{\Delta y}{y} \cdot 100 \% \approx 7.754 \Delta x_1 - 0.234 x_1 \Delta x_1 + 0.0015 x_1^2 \Delta x_1 = \begin{cases}
+2.08, x_1 = 30\\
+0.26, x_1 = 45\\
+-0.886, x_1 = 60
+\end{cases}
+\end{align*}
+
+Thus, the table looks like:
+
+\begin{table}[H]
+	\begin{tabular}{cccc}
+		Specification & Age 30 to 31  & Age 45 to 46 & Age 60 to 61 \\
+		\hline
+		Linear in \textit{age}& 0.6  & 0.6 & 0.6 \\
+		Quadratic in \textit{age}  &1.92  &0.34  &-1.23  \\
+		Cubic in \textit{age} &2.08  &0.26  &-0.886\\
+		\hline 
+	\end{tabular}
+\end{table}
+\item From the linear to quadratic specification this values significantly differs, quadratic specification reflects the truth of life that the average salary firstly increases with age, then it attains maximum at middle age and then starts falling down.
+
+AS for comparison between quadratic and cubic models, both models predict approximately equal values in real world sense.
+\item As I have mentioned somewhere above, the second row makes more sense for me, because it is interpretable (formula for percent changes in $y$ does not include $\text{age}^2$ which is a very strange indicator in real-world terms), in contrast to the cubic model and moreover, estimates gotten from this model seem rather plausible.
 \end{enumerate}

### 2019-01-31 macro3_2.tex
Macro

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Macroeconomics 3. Problem Set 2.
\end{center}
\section*{Problem 1}
Consider a neoclassical growth model in continuous time with technological progress. Production function
is given by $F (K, AL)$, $A_t = Ze^{gt}, 0 < g < \rho$, where $\rho$ is the discount factor. Utility function is CRRA, $u (c) = c^{\gamma}, 0 < \gamma < 1$.
\begin{enumerate}[1.]
\item Write down the planner’s problem in terms of original variables $K_t, L_t, C_t$.
\item Denote $k_t \equiv \frac{K_t}{e^{gt}L_t}$ and $c_t =\frac{C_t}{e^{gt}L_t}$. Rewrite the planner’s problem in terms of these variables.
\item Write down all optimality conditions.
\item Plot the phase diagram in coordinates $(k, c)$. Show the optimal trajectory.
\item Suppose that at moment $t = 0$ the economy is in the steady state. Use the phase diagram to
describe what happens after an unexpected permanent change in labor productivity $Z$.
\item Suppose that at moment $t = 0$ the economy is in the steady state. Use the phase diagram to
describe what happens after an unexpected temporary change in labor productivity $Z$: at $t = 0$ $Z$
jumps to its new value, and it is expected that at $t = T > 0$ $Z$ drops back to its original value.
\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item The planner's problem is:
	\begin{align*}
	&\underset{C(t)}{\max}\ \int_{0}^{\infty} L(t) \left(\frac{C(t)}{L(t)}\right)^{\gamma}e^{-\rho t}dt\\
	s.t.\ &\dot{K}(t) = F(K(t), Ze^{gt}L(t)) - \delta K(t) - C(t)\\
	&C(t) \ge 0, K(t) \ge 0, L(t) \ge 0\\
	&K(0) \ge 0, L(0) \ge 0 \text{ are given }
	\end{align*}
	\item $k_t = \frac{K_t}{e^{gt}L_t}$ assuming $L_t = L(0)\ \forall\ t$, we obtain
	\begin{align*}
	\dot{K}(t) = \frac{d}{dt}(k_te^{gt}L_t) = \dot{k}_te^{gt}L_t + gk_te^{gt}L_t
	\end{align*}
	hence resources constraint can be rewritten as follows:
	\begin{align*}
	\dot{k_t} + gk_t = \frac{F(K_t, e^{gt}ZL_t)}{e^{gt}L_t} - \delta\frac{K(t)}{e^{gt}L_t} - \frac{C_t}{e^{gt}L_t}
	\end{align*}
	 assuming CRS production function, finally the planner's problem should be rewritten in the following manner:
	\begin{align*}
	&\underset{c_t}{\max}\ \int_{0}^{\infty} c_t^{\gamma}e^{-(\rho - g \gamma)t}dt\\
	s.t.\ &\dot{k}(t) = F(k_t, Z) - (\delta+g) k_t - c_t\\
	&c_t \ge 0, k_t \ge 0\\
	&k_0 \ge 0 \text{ is given}
	\end{align*}
	\item 
	\begin{align*}
	H = c_t^{\gamma}e^{-(\rho - g\gamma)t} + \lambda_t (F(k_t, Z) - (\delta +g) k_t - c_t)\\
	\begin{cases}
	\frac{\partial H}{\partial c_t} = \gamma c_t^{\gamma - 1}e^{-(\rho - g\gamma)t} - \lambda_t = 0\\
	\dot{\lambda}_t = -\frac{\partial H}{\partial k_t} = -\lambda_t(F'(k_t, Z) - \delta - g)\\
	\lim_{t \to \infty} \lambda_tk_t = 0
	\end{cases}
	\end{align*}
	\item \begin{align*}
	&\frac{d}{dt} \left(\frac{\gamma e^{-(\rho - g\gamma)t}}{c_t^{1 - \gamma}}\right) = \dot{\lambda}_t\\
	-\frac{\gamma(1 - \gamma)\dot{c}_t}{c_t^{2 - \gamma}}e^{-(\rho - g\gamma)t} - &\frac{\gamma(\rho - g\gamma)}{c_t^{1 - \gamma}}e^{-(\rho - g\gamma)t} = \dot{\lambda}_t = -\lambda_t(F'(k_t, Z) - \delta - g)\\
	\frac{\gamma(1 - \gamma)\dot{c}_t}{c_t^{2 - \gamma}}e^{-(\rho - \gamma g)t} + &\frac{\gamma(\rho - g\gamma)}{c_t^{1 - \gamma}}e^{-(\rho - g\gamma)t} = \frac{\gamma}{c_t^{1 - \gamma}}e^{-(\rho - g\gamma)t}(F'(k_t, Z) - \delta - g)\\
	&\frac{(1 - \gamma)\dot{c}_t}{c_t} + \rho - g\gamma = F'(k_t, Z) - \delta - g
	\end{align*}
	If $\dot{c}_t = 0$ then either $F'(k_t, Z) = \rho +(1-\gamma)g + \delta$ or $c_t = 0$. Denote the solution to this equation as $\tilde{k}$. If $\dot{k}_t = 0$ then $c_t = F(k_t, Z) - (\delta + g)k_t$ assuming $F(\cdot)$ satisfies Inada conditions, the isoclines will be as it depicted on the Fig \ref{fig1}. 
	\begin{figure}[h]
		\centering
		\includegraphics[width=0.8\textwidth]{plotdraft}
		\caption{}\label{fig1}
	\end{figure}
\item If $Z$ increases than $F(k_t, Z)$ increases either as a result isocline $\dot{k}(t) = 0$ moves upward because for a given $k_t$, $c_t$ increases. Moreover, if marginal productivity of capital increases with increase of the other factor (which is a standard assumption about production functions) then in order for $F'(k_t, Z)$ remains unchanged and equal to $\delta + g + \rho - g\gamma$, $k_t$ should also increase (since $\frac{\partial F}{\partial k}$ decreases in $k$). Thus, the isocline,corresponded to $\dot{c} = 0$ moves to the left (see Fig. ). Of course, exact values of these shifts depend on parameters and a type of the production function, as a result in the new steady state consumption could be more or less than the old one. Let us consider the case, which is depicted on the Fig. \ref{fig1}. In this case at $t = 0$ consumption will be increased such that the future dynamics will be along the optimal saddle trajectory of the new steady state (see Fig. \ref{fig1}).
\item In this case the social planner should from $t = 0$ to $T = t$ moves along the trajectory of new, modified system but he should choose those trajectory of new system which exactly at $t = T$ brings the system to the saddle trajectory of old steady state. See Fig. \ref{fig3} for clarifications.

	\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{plotdraft}
	\caption{}\label{fig3}
\end{figure}

\end{enumerate}
\section*{Problem 2}
Problem:
\begin{enumerate}[1.]
	\item
\begin{align*}
\underset{\left\{k(t+1), c(t)\right\}_{t=1}^{\infty}}{\max}\ \sum_{t=1}^{\infty} 0.96^t \ln c(t)\\
\text{s.t.\ } k(t+1) = Ak(t)^{1/3} + (1-0.02)k(t) - c(t)\\
c(t) \ge 0, k(t) \ge 0, k(1) = 0.1
\end{align*}
Euler's equation is:
\begin{align*}
\frac{c(t+1)}{\beta c(t)} = \frac{A}{3k(t+1)^{2/3}} + 1 - \delta
\end{align*}
In the steady state $c(t+1) = c(t)$ hence:
\begin{align*}
\frac{1}{\beta} = \frac{A}{3(k^*)^{2/3}} + 1 - \delta\\
\text{if } k^* = 1 \to A = 3(\frac{1}{\beta} + \delta - 1) = 0.185
\end{align*}
\item 
\begin{align*}
V(k) = \underset{c, k'}{\max}\ [u(c) + \beta V(k')]\\
s.t.\ k' = f(k) + (1 - \delta)k - c
\end{align*}
we don’t need values of $k$ below $0.1$ and
above $1$ because by theorem $k(t)$ monotonically increases from $k(1)$ towards the steady state $k^* = 1$.
\item Comparisons with the shooting algorithm for 100, 500 and 1000-points grids are depicted below on the Fig. \ref{fig4}, \ref{fig2} and \ref{fig5} respectively.
	\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{shooting100}
	\caption{Comparison with shooting algorithm for 100-points grid}\label{fig4}
\end{figure}
	\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{VFI}
	\caption{Comparison with shooting algorithm for 500-points grid}\label{fig2}
\end{figure}
	\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{shooting1000}
	\caption{Comparison with shooting algorithm for 1000-points grid}\label{fig5}
\end{figure}

For 1000 points-grid accuracy is indeed significantly higher comparing to 100-points. However even in case of 500-points the accuracy is high enough so, in fact there is no need in 1000-points grid which requires high computational power of the computer (it has taken about 10-15 minutes for my computer to execute the algorithm for 1000 points).
\end{enumerate}
\end{document}

### 2019-02-04 Metrics3.R
Metrics

@@ -1,0 +1,45 @@
+df <- read.csv('cps99_ps1.csv')
+
+#Problem 3
+#a
+fit1 <- lm(log(df$ahe) ~ df$yrseduc+age+female+hsdipl, data = df)
+reg <- summary(fit1)
+#b
+sqrt(mean(fit1$residuals^2))
+reg$sigma
+#c
+library(lmtest)
+library(sandwich)
+confint(fit1, vcov = vcovHC(fit1, type = "HC1"))
+#d
+coeftest(fit1, vcov = vcovHC(fit1, type = "HC1"))
+#e
+fit2 <- lm(log(ahe)~hsdipl+age+female, data = df)
+summary(fit2)
+
+#f
+df$hsdip_1 = df$hsdipl - 1
+df$yrseduc_2 = df$yrseduc - 2
+fit3 <- lm(log(ahe)~ id+hsdip_1+yrseduc_2+age+female, data=df)
+summary(fit3)
+confint(fit3, vcov = vcovHC(fit3, type = "HC1"))
+
+#Problem 4
+#a
+df$age_sq = df$age^2
+fit4 <- lm(log(ahe)~yrseduc+age+age_sq+female+hsdipl, data = df)
+summary(fit4)
+
+#b
+coeftest(fit4, vcov = vcovHC(fit4, type = "HC1"))
+#c
+cor(df$age^2, df$female)
+#d
+df$age_cub = df$age^3
+fit5 <- lm(log(ahe)~age+age_sq+age_cub+female+hsdipl+yrseduc, data = df)
+summary(fit5)
+coeftest(fit5, vcov = vcovHC(fit5, type = "HC1"))
+#e
+library(car)
+myH0 <- c("age_sq = 0", "age_cub = 0")
+linearHypothesis(fit5, myH0, vcov = vcovHC(fit5, type = "HC1"))

### 2019-02-05 metrics4.tex
metrics

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\newcommand{\yrseduc}{\textit{yrseduc}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Econometrics 1. Problem Set 4.
\end{center}
\section*{Problem 1}

\textbf{Solution}

Since $\hat{\beta_0}, \dots, \hat{\beta_k}$ are OLS estimates they satisfy the following system of equations:
\begin{align}\label{eq1}
\begin{cases}
\summa (y_i - \hat{\beta_0} - \hat{\beta_1}x_{1i} - \dots - \hat{\beta_k}x_{ki}) = 0\\
\summa x_{1i}(y_i - \hat{\beta_0} - \hat{\beta_1}x_{1i} - \dots - \hat{\beta_k}x_{ki}) = 0\\
\dots\\
\dots\\
\summa x_{ki}(y_i - \hat{\beta_0} - \hat{\beta_1}x_{1i} - \dots - \hat{\beta_k}x_{ki}) = 0
\end{cases}
\end{align}

This is also true for $\tilde{\beta_0}, \dots, \tilde{\beta_k}$, i.e.
\begin{align*}
\begin{cases}
\summa c_0(c_0y_i - \tilde{\beta_0} - \tilde{\beta_1}c_1x_{1i} - \dots - \tilde{\beta_k}c_kx_{ki}) = 0\\
\summa c_1x_{1i}(c_0y_i - \tilde{\beta_0} - \tilde{\beta_1}c_1x_{1i} - \dots - \tilde{\beta_k}c_kx_{ki}) = 0\\
\dots\\
\dots\\
\summa x_{ki}c_k(c_0y_i - \tilde{\beta_0} - \tilde{\beta_1}c_1x_{1i} - \dots - \tilde{\beta_k}c_kx_{ki}) = 0
\end{cases}
\end{align*}
Assuming that $c_0$ is not equal to 0 either, this system can be rewritten as follows:
\begin{align*}
\begin{cases}
\summa c_0^2(y_i - \frac{1}{c_0}\tilde{\beta_0} - \frac{c_1}{c_0}\tilde{\beta_1}x_{1i} - \dots - \frac{c_k}{c_0}\tilde{\beta_k}x_{ki}) = 0\\
\summa \frac{c_1}{c_0}x_{1i}(y_i - \frac{1}{c_0}\tilde{\beta_0} - \frac{c_1}{c_0}\tilde{\beta_1}x_{1i} - \dots - \frac{c_k}{c_0}\tilde{\beta_k}x_{ki}) = 0\\
\dots\\
\dots\\
\summa \frac{c_k}{c_0}x_{ki}(y_i - \frac{1}{c_0}\tilde{\beta_0} - \frac{c_1}{c_0}\tilde{\beta_1}x_{1i} - \dots - \frac{c_k}{c_0}\tilde{\beta_k}x_{ki}) = 0
\end{cases}
\end{align*}
or, finally
\begin{align*}
\begin{cases}
\summa (y_i - \frac{1}{c_0}\tilde{\beta_0} - \frac{c_1}{c_0}\tilde{\beta_1}x_{1i} - \dots - \frac{c_k}{c_0}\tilde{\beta_k}x_{ki}) = 0\\
\summa x_{1i}(y_i - \frac{1}{c_0}\tilde{\beta_0} - \frac{c_1}{c_0}\tilde{\beta_1}x_{1i} - \dots - \frac{c_k}{c_0}\tilde{\beta_k}x_{ki}) = 0\\
\dots\\
\dots\\
\summa x_{ki}(y_i - \frac{1}{c_0}\tilde{\beta_0} - \frac{c_1}{c_0}\tilde{\beta_1}x_{1i} - \dots - \frac{c_k}{c_0}\tilde{\beta_k}x_{ki}) = 0
\end{cases}
\end{align*}
Substituting 
\begin{align}\label{eq2}
\tilde{\beta_i} = \frac{c_0}{c_i}\tilde{\beta_i}', \forall\ i = 1, \dots, k\\
\tilde{\beta_0} = c_0 \tilde{\beta_0}' \nonumber
\end{align} 
we can get the system which is equivalent to the system \eqref{eq1}, as a result has the same solutions. Plugging these solutions $\hat{\beta_i}$ back to \eqref{eq2} we finally get
\begin{align*}
\tilde{\beta_i} = \frac{c_0}{c_i}\hat{\beta_i}\ \forall\ i=1, \dots, k\\
\tilde{\beta_0} = c_0 \hat{\beta_0}
\end{align*}
Q.E.D.

\section*{Problem 2}
\textbf{Solution}

\begin{align*}
\ln(wage) = \beta_0 + \beta_1 educ + \beta_2 educ \times pareduc + \beta_3 exper + \beta_4 tenure + u
\end{align*}

\begin{enumerate}[(i)]
	\item Obviously
	\begin{align*}
	\frac{d \ln (wage)}{d\ educ} = \beta_1 + \beta_2 pareduc
	\end{align*}
	hence
	\begin{align*}
	\frac{\Delta \ln (wage)}{\Delta educ} \approx \beta_1 + \beta_2 pareduc
	\end{align*}
	As for me it is difficult to say something about sign of $\beta_0$ apriori. If $\beta_2$ is positive then each additional year of parents' education yields greater average percentage change in salary with additional year of education. I cannot come up with any strong reasoning for $\beta_2$ be positive or negative.
	\item If $pareduc = 24$ then
	\begin{align*}
	100\% \cdot \frac{\Delta wage}{wage} = 4.7 \Delta educ + 0.078 \cdot 24 \Delta educ = 4.7 \Delta educ + 1.872 \Delta educ
	\end{align*}
	If $pareduc = 32$ then
	\begin{align*}
	100\% \cdot \frac{\Delta wage}{wage} = 4.7 \Delta educ + 0.078 \cdot 32 \Delta educ = 4.7 \Delta educ + 2.496 \Delta educ
	\end{align*}
	i.e. each additional year of parents' education increases by 0.078 percent points the value of percent change of salary in response to additional year of education.
	\item \begin{align*}
	t = \frac{0.0016}{0.0012} \approx 1.33
	\end{align*}
	p-value is about 0.09, hence we can reject the hypothesis on 5\% significance level. That is, the return to education does not depend on parents' education.
	\section*{Problem 3}

	\textbf{Solution}

	Since 
	\begin{align*}
	F \equiv \frac{(SSR_r - SSR_{ur}) / q}{SSR_{ur}/(n - k - 1)}
	\end{align*}
	where $SSR_r$ and $SSR_{ur}$ are sums of squares of residuals of restricted and unrestricted models respectively, $q = 2, k = 8$. 
	\begin{align*}
	R^2_{ur} = 1 - \frac{SSR_{ur}}{SST_{ur}}\\
	R^2_{r} = 1 - \frac{SSR_{r}}{SST_{r}}\\
	\end{align*}
	while
	\begin{align*}
	SST = \summa (y_i - \bar{y}) = SST_r = SST_{ur}
	\end{align*}
	hence
	\begin{align*}
	F = \frac{(R^2_{ur} - R^2_r)/q}{(1 - R^2_{ur})/(n-k-1)} = \frac{(0.232 - 0.229)/2}{(1-0.232)/(680 - 8 - 1)} \approx 1.31
	\end{align*}
	p-value is about 0.27, hence the hypothesis (that corresponding slope coefficients are 0) cannot be rejected at the 10\% significance level. That is, both these terms are jointly insignificant hence I would not include them in the model.
\end{enumerate}
\section*{Problem 4}

\end{document}

### 2019-02-07 metrics4.tex
Metrics

@@
 \section*{Problem 4}
+\begin{enumerate}[(a)]
+	\item From Problem Set 2, Problem 3:
+	\begin{align*}
+	\hat{\beta_1} = \frac{s^2_{yx_1}s^2_{x_2x_2} - s^2_{yx_2}s^2_{x_2x_1}}{s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2}\\
+	\hat{\beta_2} = \frac{s^2_{yx_2}s^2_{x_1x_1} - s^2_{yx_1}s^2_{x_2x_1}}{s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2}
+	\end{align*}
+	where $s^2_{xy}$ means a sample covariance between $x$ and $y$. That means that:
+	\begin{align}\label{eq4}
+	&Var(\hat{\beta_1}|x_1, x_2) = \left(\frac{s^2_{x_2x_2}}{s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2}\right)^2 Var(s^2_{yx_1}|x_1, x_2) + \nonumber\\
+	&+\left(\frac{s^2_{x_2x_1}}{s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2}\right)^2 Var(s^2_{yx_2}|x_1, x_2) - \frac{2s^2_{x_2x_2}s^2_{x_2x_1}}{(s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2)^2}cov(s^2_{yx_1}, s^2_{yx_2}|x_1, x_2) = \nonumber\\
+	&\underset{i.i.d}{=} \left(\frac{s^2_{x_2x_2}}{s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2}\right)^2 \summa (x_{i1} - \bar{x_1})^2Var(y_i|x_{i1}, x_{i2}) + \nonumber\\
+	&\left(\frac{s^2_{x_2x_1}}{s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2}\right)^2 \summa (x_{i2} - \bar{x_2})^2Var(y_i|x_{i1}, x_{i2}) -\nonumber\\
+	&-\frac{2s^2_{x_2x_2}s^2_{x_2x_1}}{(s^2_{x_1x_1}s^2_{x_2x_2} -(s^2_{x_1x_2})^2)^2} \summa \sum_{j=1}^n (x_{i1} - \bar{x_1})(x_{j2} -\bar{x_2})cov(y_i, y_j|x_{1}, x_2) = \nonumber\\
+	&=\left(\frac{s^2_{x_2x_2}}{s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2}\right)^2 s^2_{x_1x_1} \sigma^2 +
+	\left(\frac{s^2_{x_2x_1}}{s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2}\right)^2 s^2_{x_2x_2} \sigma^2 - \nonumber\\
+	&-\frac{2s^2_{x_2x_2}s^2_{x_2x_1}}{(s^2_{x_1x_1}s^2_{x_2x_2} -(s^2_{x_1x_2})^2)^2} s^2_{x_1x_2}\sigma^2 = \frac{s^2_{x_2x_2}}{s^2_{x_1x_1}s^2_{x_2x_2} -(s^2_{x_1x_2})^2}\sigma^2
+	\end{align}
+	Similarly for $\hat{\beta_2}$:
+	\begin{align}\label{eq3}
+	&Var(\hat{\beta_2}|x_1, x_2) = \frac{s^2_{x_1x_1}}{s^2_{x_1x_1}s^2_{x_2x_2} -(s^2_{x_1x_2})^2}\sigma^2
+	\end{align}
+	Then
+	\begin{align*}
+	Var(\hat{\beta_1}|x_1, x_2) - Var(\hat{\beta_2}|x_1, x_2) = \frac{\sigma^2}{s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2} (s^2_{x_2x_2} - s^2_{x_1x_1})
+	\end{align*}
+	and 
+	\begin{align*}
+	\widehat{Var(\hat{\beta_1}|x_1, x_2)} - \widehat{Var(\hat{\beta_2}|x_1, x_2)} =\frac{\hat{\sigma}^2}{s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2} (s^2_{x_2x_2} - s^2_{x_1x_1})
+	\end{align*}
+	Since $\widehat{Var(\hat{\beta_1}|x_1, x_2)} - \widehat{Var(\hat{\beta_2}|x_1, x_2)} > 0$ and the determinant of sample covariance matrix (which appears in the denominator) should be nonnegative (i.e.
+positive) then $s^2_{x_2x_2} - s^2_{x_1x_1}$ should be positive.
+Hence, once we have an invertible sample covariance matrix the sample variance of $x_1$ should be smaller than the sample variance of $x_2$.
+	\item 
+	\item Since
+	\begin{align*}
+	Var(\hat{\beta_1} + \hat{\beta_2}) = Var(\hat{\beta_1}) + Var(\hat{\beta_2}) + 2cov(\hat{\beta_1}, \hat{\beta_2})
+	\end{align*}
+	It remains to estimate the covariance of $\hat{\beta_1}$ with $\hat{\beta_2}$.
+	\begin{align*}
+	cov(\hat{\beta_1}, \hat{\beta_2}|x_1, x_2) = \frac{s^2_{x_2x_2}s^2_{x_1x_1}cov(s^2_{yx_1}, s^2_{yx_2}) - s^2_{x_2x_2}s^2_{x_2x_1}Var(s^2_{yx_1})}{(s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2)^2} - \\
+	-\frac{ s^2_{x_2x_1}s^2_{x_1x_1}Var(s^2_{yx_2}) - s^2_{x_2x_1}s^2_{x_2x_1}cov(s^2_{yx_2}, s^2_{yx_1})}{(s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2)^2} = \\
+	=\frac{-s^2_{x_1x_2}}{s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2}\sigma^2
+	\end{align*}
+	That means that
+	\begin{align*}
+	\widehat{Var(\hat{\beta_1} + \hat{\beta_2})} = 
+	\end{align*}
+	\item $\summa x_ix_i^T$ is a diagonal meaning that $s^2_{x_1x_2} = 0$ hence we need to find $s^2_{x_1x_1}$ and $s^2_{x_2x_2}$. From \eqref{eq3} and \eqref{eq4} it follows that:
+	\begin{align*}
+	2 = SE(\hat{\beta_1}) = \sqrt{\frac{s^2_{x_2x_2}3^2}{(n-1)s^2_{x_1x_1}s^2_{x_2x_2}}}\\
+	(n-1)s^2_{x_1x_1} = \frac{9}{4}
+	\end{align*}
+	Similarly for $\hat{\beta_2}$:
+	\begin{align*}
+	1 = \sqrt{\frac{s^2_{x_1x_1}3^2}{(n-1)s^2_{x_1x_1}s^2_{x_2x_2}}}\\
+	(n-1)s^2_{x_2x_2} = 3
+	\end{align*}
+	Thus, the diagonal of $\summa x_ix_i^T$ which is $n, \summa x_{1i}^2, \summa x_{2i}^2$ is equal to:
+	\begin{align*}
+	100,\ \frac{9}{4},\ 3
+	\end{align*}
+\end{enumerate}
+\section*{Problem 5}
+Summary of the regression is depicted on the Fig. \ref{fig1}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{lnahe_femyrs}
+	\caption{Summary of regression}\label{fig1}
+\end{figure}
+
+\begin{enumerate}[a.]
+	\item Denote coefficient on $yrseduc$ as $\beta_1$, on $female\times yrseduc$ as $\beta_2$ and on $female$ as $\beta_3$. For males ($female = 0$)
+	\begin{align*}
+	\frac{\Delta ahe}{ahe} \cdot 100\% = \beta_1 \cdot 100 \Delta yrseduc
+	\end{align*}
+	i.e. the coefficient on $yrseduc$ times 100 is the percent change of average per hour salary for men adding one additional year of education, holding other things fixed.
+
+	For women:
+	\begin{align*}
+	\frac{\Delta ahe}{ahe} \cdot 100\% = 100(\beta_1 + \beta_2)\Delta yrseduc
+	\end{align*}
+	i.e.
+the coefficient on $yrseduc$ plus the coefficient on $yrseduc \times female$ times 100 is the percent change of average per hour salary for women adding one additional year of education, holding other things fixed.
+	\item 
+	\item We should test whether $\beta_2 = 0$ or not.
+Using coeftest() we can conclude that the null hypothesis can be rejected at the 5\% significance level, i.e.
+the value of the aditional year of education for men and for women differs.
+	\item We need to test the hypothesis that both $\beta_2$ and $\beta_3$ are equal to 0.
+Using linearhypothesis() we can conclude that the hypothesis should be rejected, i.e.
+regression lines are different for men and for women.
+	\item For men additional year of education increases the average per hour salary by $8.26\%$ for women by $9.9\%$, in statistical terms this is a significant difference (since we reject the hypothesis that $\beta_2 = 0$) in real world sense I think not.
+\end{enumerate}
+\section*{Problem 6}
 

### 2019-02-08 metrics4.tex
metrics

@@
 \section*{Problem 6}
+Nothing changed. R did not include $male \times yrseduc$ in regression because $male \times yrseduc = yrseduc - female \times yrseduc$ i.e. regressors are perfectly linearly dependent.
+\section*{Problem 7}
+Summary for two models are depicted on the Fig. \ref{fig2} and \ref{fig3} respectively.
 
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{age_cub}
+	\caption{Summary of regression}\label{fig2}
+\end{figure}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{age_cub2}
+	\caption{Summary of regression}\label{fig3}
+\end{figure}
+\begin{enumerate}[a.]
+	\item Denote the coefficients of the models as follows:
+	\begin{align*}
+	\ln ahe = \beta_0 + \beta_1 female + \beta_2 yrseduc + \beta_3 age + \beta_4 age^2 + \beta_5 age^3 + \beta_6 female \times yrseduc + \\
+	+\beta_7 female \times age + \beta_8 female \times age^2 + \beta_9 female \times age^3
+	\end{align*}
+	For men :
+	\begin{align*}
+	\text{1st model: } 100 \beta_2 = 8.08\\
+	\text{2nd model: } 100 \beta_2 = 8.03\\
+	\end{align*}
+	i.e. interaction terms do not significantly affect the value of additional year of education for men, in a real-world sense.
+	For women:
+	\begin{align*}
+	\text{1st model: } 100(\beta_2 + \beta_6) = 10.13\\
+	\text{2nd model: } 100(\beta_2 + \beta_6) = 10.07
+	\end{align*}
+	i.e. interaction terms do not significantly affect the value of additional year of education for women, in a real-world sense, either.
+	\item \begin{align*}
+	\frac{\Delta ahe}{ahe} \cdot 100 = 100(\beta_1 + 16 \beta_6 + 25 \beta_7 + 25^2 \beta_8 + 25^3 \beta_9)\Delta female = -8.42 \Delta female
+	\end{align*}
+	i.e. on average 25-year old female workers with 16 years of education earn by 8.4\% less than their male counterparts.
+	\item \begin{align*}
+	\frac{\Delta ahe}{ahe} \cdot 100 = 100(\beta_1 + 16 \beta_6 + 55 \beta_7 + 55^2 \beta_8 + 55^3 \beta_9)\Delta female = -22.69 \Delta female
+	\end{align*}
+	i.e. on average 55-year old female workers with 16 years of education earn by 22.7\% less than their male counterparts.
+	\item We need to test the hypothesis that $\beta_7 = \beta_8 = \beta_9 = 0$.
+Using linearHypothesis() we can conclude that at the 5\% significance level the hypothesis should be rejected.
+That is, a gender gap depends on age.
+	\item Yes, there is evidence.
+Coefficients jointly significanly differ from zero and moreover based on (c) and (d) we can conclude that for 25 year-old female and for 55-year old the gender gap significantly differs in a real-world terms (8\% for the former and 22.7\% for the latter)
+\end{enumerate}
 \end{document}

### 2019-02-09 micro3_5.tex
Micro

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\newcommand{\yrseduc}{\textit{yrseduc}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Microeonomics 3. Problem Set 5.
\end{center}
\section*{Problem 1}
\begin{enumerate}[(a)]
	\item  Consider a modified version of Question 4 in Problem Set 4 in which trade takes
	place sequentially. Suppose there are two Arrow securities (denominated in terms of the
	physical good 1). Find a Radner equilibrium of this modified model. (You should determine
	the equilibrium values of all variables, i.e., consumption plans, asset portfolios, spot prices
	and asset prices.)
	\item Suppose now that in place of the Arrow securities, there are two assets defined by the
	return vectors $r_1 = (2, 1)$ and $r_2 = (1, 2)$, respectively. As usual, the $s$’th coordinate of $r_k$
	represents the amount of good 1 that the asset $k$ pays in state $s$ (for $k = 1, 2$ and $s = 1, 2$).
	Find a Radner equilibrium of the corresponding sequential exchange model.

\end{enumerate}

\textbf{Solution}

\begin{enumerate}[(a)]
	\item From the Problem 4 of Problem Set 4 it is known that the Arrow-Debreu equilibrium consists of prices:
	\begin{align*}
	p_{11} = p_{22} = 1,\ p_{12} = p_{21} = 3
	\end{align*}
	and allocation:
	\begin{align*}
	x_1 = \begin{pmatrix}
	3\\
	1\\
	3\\
	9
	\end{pmatrix},\ x_2 = \begin{pmatrix}
	9\\
	3\\
	1\\
	3
	\end{pmatrix}
	\end{align*}
From the theorem of equivalence between Arrow-Debreu equilibrium and Radner equilibrium it follows that prices of two Arrow securities are equal to:
\begin{align*}
q^* = \begin{pmatrix}
p_{11}\\
p_{12}
\end{pmatrix} = \begin{pmatrix}
1\\
3
\end{pmatrix}
\end{align*}
Equilibrium consumption plans do not change. It remains to find equilibrium portfolio for each consumer. 
\begin{align*}
z^*_{si} = \frac{1}{p_{1s}}p_s(x_{si} - \omega_{si})\\
z^*_{1} =  \begin{pmatrix}
-18\\
6
\end{pmatrix},\ z^*_{2} = \begin{pmatrix}
18\\
-6
\end{pmatrix}
\end{align*}
\item 
\begin{align*}
R = \begin{pmatrix}
2 & 1\\
1 & 2
\end{pmatrix}
\end{align*}
The return matrix has a full rank hence the consumption plan do not change and the asset prices will be:
\begin{align*}
\tilde{q}^* = R^Tq = \begin{pmatrix}
5\\
7
\end{pmatrix}
\end{align*}
Again, by the theorem of equivalence between Radner equilibrium with assets and Arrow-Debreu equilibrium, prices do not change. It remains to find assets portfolio for each consumer. To do so, for the first consumer we need to solve the following linear system:
\begin{align*}
\begin{pmatrix}
2 & 1\\
1 & 2
\end{pmatrix} \begin{pmatrix}
\tilde{z_{11}}^*\\
\tilde{z_{12}}^*
\end{pmatrix} = \begin{pmatrix}
-18\\
6
\end{pmatrix}
\end{align*}
and the solution is 
\begin{align*}
\tilde{z}_1^* = \begin{pmatrix}
-14\\
10
\end{pmatrix}
\end{align*}
For the second consumer, we need to clear the security market hence:
\begin{align*}
\tilde{z_2}^* = \begin{pmatrix}
14\\
-10
\end{pmatrix}
\end{align*}
(Also note that $\tilde{q}^*\tilde{z_i}^* = 0\ \forall\ i$ i.e. no-arbitrage condition holds)
\end{enumerate}
\section*{Problem 2}
\begin{enumerate}[(a)]
\item Consider a sequential version of Question 5 in Problem Set 4 with two Arrow
securities. Find a Radner equilibrium of this modified model.
\item  If we were to add one more asset to this economy defined by a return vector $r_3 \in \mathbb{R}^2$,
how would you compute the equilibrium price of this new asset?

\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item Again, from the Problem 5 of Problem Set 4 it is known that the Arrow-Debreu equilibrium consists of prices:
	\begin{align*}
	p_{11} = p_{21} = 1,\  p_{22} = p_{12} = 2 
	\end{align*}
	and the equilibrium allocation:
	\begin{align*}
	x_1 = \begin{pmatrix}
	4\\
	8\\
	2\\
	4
	\end{pmatrix},\ x_2 = \begin{pmatrix}
	8\\
	4\\
	4\\
	2
	\end{pmatrix}
	\end{align*}
	From the theorem of equivalence between Arrow-Debreu equilibrium and Radner equilibrium it follows that prices of two Arrow securities are equal to:
	\begin{align*}
	q^* = \begin{pmatrix}
	p_{11}\\
	p_{12}
	\end{pmatrix} = \begin{pmatrix}
	1\\
	2
	\end{pmatrix}
	\end{align*}
	Equilibrium consumption plans do not change. It remains to find equilibrium portfolio for each consumer. 
	\begin{align*}
	z^*_{si} = \frac{1}{p_{1s}}p_s(x_{si} - \omega_{si})\\
	z^*_{1} =  \begin{pmatrix}
	0\\
	0
	\end{pmatrix},\ z^*_{2} = \begin{pmatrix}
	0\\
	0
	\end{pmatrix}
	\end{align*}
	\item Since market with 2 Arrow securities was complete that means that we can express $r_3$ as a unique linear combination of columns of the identity matrix $2 \times 2$ i.e. $r_3 = r_{31} e_1 + r_{32} e_2$. Then:
	\begin{align*}
	q_3 = r_{31} q_1 + r_{32} q_2= r_{31} + 2r_{32}
	\end{align*}
\end{enumerate}
\section*{Problem 3}
Consider a sequential exchange model with two physical goods, two states, two consumers, and two assets. The return vectors of the assets are as follows:
\begin{align*}
r_1 = (1,\ 1)\ r_2 = (2,\ 1)
\end{align*}
As usual, the $s$’th entry of $r_k$ represents the return of asset $k$ in state $s$ in terms of the
physical good 1.
Consumers’ Bernoulli utility functions are state independent:
\begin{align*}
u_1(x_{s1}) = (x_{1s1})^{1/2}(x_{2s1})^{1/2},\ 
u_2(x_{s2}) = (x_{1s2})^{1/3}(x_{2s2})^{2/3}
\end{align*}
Here, $x_{si} = (x_{1si}, x_{2si})$ stands for the consumption bundle of consumer $i$ in state $s = 1, 2$.
Both consumers assign probability $1/2$ to each state.
Let $\omega_i = (\omega_{lsi})$, where $\omega_{lsi}$ is consumer $i$’s endowment of good $l$ in state $s$. Assume
further that:
\begin{align*}
\omega_1 = (\omega_{111}, \omega_{211}, \omega_{121}, \omega_{221}) = (5, 1, 0, 0),\ \ \omega_2 = (\omega_{112}, \omega_{212}, \omega_{122}, \omega_{222}) = (0, 0, 5, 1).
\end{align*}
Thus, consumer 1 has all the endowment in state 1 while consumer 2 has all the endowment
in state 2. Moreover, the total endowment vector equals $(5,\ 1)$ in both states.
Compute a Radner equilibrium of this model.

\textbf{Solution}

Let us compute firstly the Arrow-Debreu equilibrium. Consumers are solving the following problems:
\begin{align*}\
\underset{x_{111}, x_{211}, x_{121}, x_{221} \ge 0}{\max}\ \frac{1}{2} (x_{111})^{1/2} (x_{211})^{1/2} + \frac{1}{2} (x_{121})^{1/2}(x_{221})^{1/2}\\
s.t.\ p_{11}x_{111} + p_{21}x_{211} + p_{12}x_{121} + p_{22}x_{221} \le 5p_{11} + p_{21}\\
\underset{x_{112}, x_{212}, x_{122}, x_{222} \ge 0}{\max}\ \frac{1}{2} (x_{112})^{1/3} (x_{212})^{2/3} + \frac{1}{2} (x_{122})^{1/3}(x_{222})^{2/3}\\
s.t.\ p_{11}x_{111} + p_{21}x_{211} + p_{12}x_{121} + p_{22}x_{221} \le 5p_{12} + p_{22}\\
\end{align*}
And moreover market clearing conditions imply:
\begin{align*}
x_{111} + x_{112} = 5 \nonumber\\
x_{211} + x_{212} = 1\nonumber\\
x_{121} + x_{122} = 5\\
x_{221} + x_{222} = 1\nonumber
\end{align*}
Since derivatives of objective functions tend to infinity while some of $x$ tend to 0 hence the solution should lie at the interior. The solution is:
\begin{align*}
\begin{cases}
x_{111} = \frac{5p_{11} + p_{21}}{4p_{11}}\\
x_{211} = \frac{5p_{11} + p_{21}}{4p_{21}}\\
x_{121} = \frac{5p_{11} + p_{21}}{4p_{12}}\\
x_{221} = \frac{5p_{11} + p_{21}}{4p_{22}}\\
\end{cases}\\
\begin{cases}
x_{112} = \frac{5p_{12} + p_{22}}{6p_{11}}\\
x_{212} = \frac{5p_{12} + p_{22}}{3p_{21}}\\
x_{122} = \frac{5p_{12} + p_{22}}{6p_{12}}\\
x_{222} = \frac{5p_{12} + p_{22}}{3p_{22}}\\
\end{cases} 
\end{align*}
Market clearing conditions imply that:
\begin{align*}
\begin{cases}
-45p_{11} + 3p_{21} + 10p_{12} + 2p_{22} = 0\\
15p_{11} - 9p_{21} + 20p_{12} + 4p_{22} = 0\\
15p_{11} + 3p_{21} - 50p_{12} + 2p_{22} = 0\\
15p_{11} + 3p_{21} +20p_{12} - 8p_{22} = 0\\
\end{cases}
\end{align*}
Normalizing $p_{11} = 1$ we get the solution:
\begin{align*}
p_{11} = p_{12} = 1,\ p_{22} = p_{21} = 7
\end{align*}
and the equilibrium allocation is:
\begin{align*}
x_1^* = \begin{pmatrix}
3\\
\\
\frac{3}{7}\\
\\
3\\
\\
\frac{3}{7}
\end{pmatrix},\ x_2^* = \begin{pmatrix}
2\\
\\
\frac{4}{7}\\
\\
2\\
\\
\frac{4}{7}
\end{pmatrix}
\end{align*}
Then, in Radner equilibrium with 2 Arrow securities:
\begin{align*}
q^* = \begin{pmatrix}
1\\
1
\end{pmatrix}
\end{align*}
equilibrium portfolio for each consumer can be found as follows: 
\begin{align*}
z^*_{si} = \frac{1}{p_{1s}}p_s(x_{si} - \omega_{si})\\
z^*_{1} =  \begin{pmatrix}
-6\\
6
\end{pmatrix},\ z^*_{2} = \begin{pmatrix}
6\\
-6
\end{pmatrix}
\end{align*}
For
\begin{align*}
&R = \begin{pmatrix}
1 & 2\\
1 & 1
\end{pmatrix}\\
&\text{assets' price vector will be }\tilde{q}^* = R^Tq = \begin{pmatrix}
2\\
3
\end{pmatrix}
\end{align*}
It remains to find assets portfolio for each consumer. To do so, for the first consumer we need to solve the following linear system:
\begin{align*}
\begin{pmatrix}
1 & 2\\
1 & 1
\end{pmatrix} \begin{pmatrix}
\tilde{z_{11}}^*\\
\tilde{z_{12}}^*
\end{pmatrix} = \begin{pmatrix}
-6\\
6
\end{pmatrix}
\end{align*}
and the solution is 
\begin{align*}
\tilde{z}_1^* = \begin{pmatrix}
18\\
-12
\end{pmatrix}
\end{align*}
For the second consumer, we need to clear the security market hence:
\begin{align*}
\tilde{z_2}^* = \begin{pmatrix}
-18\\
12
\end{pmatrix}
\end{align*}
\section*{Problem 4}
Consider a sequential exchange model with one physical good, three states, three Arrow
securities and two consumers. Let $x_{si}$ stand for consumption of consumer $i = 1, 2$ in state
$s = 1, 2, 3$. Consumers’ Bernoulli utility functions are state independent, and defined as
follows:
\begin{align*}
u_1(x_{s1}) = \sqrt{x_{s1}},\  u_2(x_{s2}) = \ln x_{s2}.
\end{align*}
Both consumers assign probability $\pi_s$ to the state $s$, where 
\begin{align*}
\pi_1 =\frac{1}{6},\ \pi_2 =\frac{2}{6},\ \pi_3 =\frac{3}{6}
\end{align*}
Consumers’ endowment vectors are as follows:
$\omega_1 = (0,\ 3,\ 0), \omega_2 = (3,\ 0,\ 1)$.
(As usual, the $s$-th entry of $\omega_i$ stands for the endowment of consumer $i$ in state $s$.) Questions
(i)-(iv) below refer to a Radner equilibrium of this model.
\begin{enumerate}[(i)]
\item Let $q_s$ denote the price of the $s$-th Arrow security, which has a positive return in state $s$.
Find a relation between $q_1$ and $q_2$. Explain your answer. (Hint. Utility functions are state
independent and strictly concave, we have a common prior, and total endowments are the
same in states 1 and 2. This should remind you an earlier example.)
\item Normalize $q_3$ to $1$. Use your answer in part (i) to determine consumers’ consumption
plans and asset portfolios, as a function of $q_1$.
\item Compute the value of $q_1$.
\item Provide a formula for the price of any additional asset that we may introduce into this
economy
\end{enumerate}

\textbf{Solution}

\begin{enumerate}[(i)]
	\item At the interior equilibrium we should have:
	\begin{align*}
	\frac{\pi_1 u_1'(x_{1})}{\pi_2 u_1'(x_2)} = \frac{p_1}{p_2} = \frac{q_1}{q_2} = \frac{\pi_1u_2'(3 - x_1)}{\pi_2u_2'(3 - x_2)}
	\end{align*}
	If $x_1 > x_2$ then $3-x_1 < 3-x_2$ and by strict concavity of utility functions:
	\begin{align*}
	\frac{u_1'(x_1)}{u_1'(x_2)} < 1 < \frac{u_2'(3-x_1)}{u_2'(3-x_2)}
	\end{align*}
	it follows that $x_1 = x_2$ and as a result $q_1 = q_2$.
	\item $q_3 = 1$. Similarly for other states:
	\begin{align*}
	\frac{\pi_1 u_1'(x_{11})}{\pi_3 u_1'(x_{31})} = \frac{p_1}{p_3} = &q_1 = \frac{\pi_1u_2'(x_{12})}{\pi_3u_2'(x_{32})}\\
	\sqrt{\frac{x_{31}}{x_{11}}} = &q_1 = \frac{x_{32}}{x_{12}}
	\end{align*}
	incorporating market-clearing:
	\begin{align*}
	\sqrt{\frac{x_{31}}{x_{11}}} = q_1 = \frac{1- x_{31}}{3 - x_{11}}
	\end{align*}
	\begin{align*}
	x_{31} = 1 - (3-&x_{11})q_1\\
	\frac{1 - (3-x_{11})q_1}{x_{11}} &= q_1^2\\
	x_{11} = \frac{3q_1 - 1}{q_1(1 - q_1)} &= x_{21}\\
	x_{31} = \frac{q_1(3q_1 - 1)}{1 - q_1}\\
	x_1^* = \begin{pmatrix}
	\frac{3q_1 - 1}{q_1(1 - q_1)}\\
		\\
		\frac{3q_1 - 1}{q_1(1 - q_1)}\\
			\\
		\frac{q_1(3q_1 - 1)}{1 - q_1}	
	\end{pmatrix},\ &x_2^* = \begin{pmatrix}
	\frac{1 - 3q_1^2}{q_1(1 - q_1)}\\
	\\
	\frac{1 - 3q_1^2}{q_1(1 - q_1)}\\
	\\
	\frac{1 - 3q_1^2}{1 - q_1}\\
	\end{pmatrix}
	\end{align*}
	And the assets portfolios will be:
	\begin{align*}
	z_1^* = \begin{pmatrix}
	\frac{3q_1 - 1}{q_1(1 - q_1)}\\
	\\
	-\frac{1 - 3q_1^2}{q_1(1 - q_1)}\\
	\\
	\frac{q_1(3q_1 - 1)}{1 - q_1}\\
	\end{pmatrix},\ z_2^* = \begin{pmatrix}
	-\frac{3q_1 - 1}{q_1(1 - q_1)}\\
	\\
	-\frac{1 - 3q_1^2}{q_1(1 - q_1)}\\
	\\
	-\frac{q_1(3q_1 - 1)}{1 - q_1}	
	\end{pmatrix}
	\end{align*}
	\item To compute $q_1$ we need to use the budget constraint, i.e.
	\begin{align*}
	2\frac{3q_1 - 1}{1-q_1} + \frac{q_1(3q_1 - 1)}{1 - q_1} = 3q_1\\
	q_1 = \frac{\sqrt{13} - 1}{6}
	\end{align*}
	\item Since the market with 3 Arrow securities was complete alike in the Problem 2, price of new security can be computed as a linear combination of prices of basic securities, with weights being equal to returns of the new security. That is, if new asset pays $r_4 = (r_{14}\ r_{24}\ r_{32})$ then its price will be:
	\begin{align*}
	q_4 = r_{14}q_1 + r_{24}q_2+ r_{32}q_3
	\end{align*}
\end{enumerate}
\section*{Problem 5}
\textbf{Step 1}

The no arbitrage condition implies that the linear space $\Phi$ does not intersect the convex set $\mathbb{R}^s_+\setminus\left\{0\right\}$. Because if some $\tilde{z} \in \mathbb{R}^s_+\setminus\left\{0\right\}$ and at the same time $\tilde{z} \in \Phi$ that means that $\exists\ z\in \mathbb{R}^K$ such that $qz = 0$ and $Rz = \tilde{z} > 0$ which contradicts the no-arbitrage condition. Since $\mathbb{R}^s_+\setminus\left\{0\right\}$ is a convex set as well as a linear subspace $\Phi$, we can use a swparation hyperplane theorem. That is, $\exists\ \mu \in \mathbb{R}^s: \mu \neq 0$ and $a \in \mathbb{R}$ such that
\begin{align*}
\mu z \ge a \ge \mu \phi\, \forall\ z \in \mathbb{R}^s_+\setminus\left\{0\right\},\ \forall\ \phi \in \Phi
\end{align*}
Since $\left\{0\right\} \in \Phi$ then
\begin{align}\label{eq1}
\mu z \ge a \ge 0 \ge \mu \phi\, \forall\ z \in \mathbb{R}^s_+\setminus\left\{0\right\},\ \forall\ \phi \in \Phi
\end{align}
Moreover, since $\Phi$ is a linear space that means that $\forall\ \phi \in \Phi\ \exists\ -\phi \in \Phi$ hence:
\begin{align*}
\forall\ \phi\in\Phi\, \mu \phi \le 0\ \text{ and } -\mu \phi \le 0 \to \mu\phi = 0\
\end{align*}
It remains to prove that $\mu >0$. To do so, let us assume the opposite, i.e. $\exists\ \text{ a coordinate } s: \mu_s < 0$. Obviously $\forall\ \varphi\ \in \mathbb{R}^s_{+}\setminus\left\{0\right\}$ and $\forall\ \lambda > 0$, $\varphi + \mathbb{I}_{s}\lambda \in \mathbb{R}^s_{+}\setminus\left\{0\right\}$. But for all $\varphi$ we can pick such a large $\lambda$ that
\begin{align*}
\mu (\varphi + \mathbb{I}_s\lambda) = \mu \varphi + \mu_s\lambda < 0
\end{align*}
but by \eqref{eq1} it should be nonnegative for any $\lambda$. In conclusion, $\exists\ \mu >0:\ \mu \phi = 0\, \forall\ \phi \in \Phi$, as we sought.

\textbf{Step 2}

Given that $\mu$, the vector $\mu R$ belongs to the linear space $\left\{\alpha q : \alpha \in \mathbb{R}\right\}$. Indeed, otherwise, we can separate the vector $\mu R$ from the latter space, since both $\mu R$ as a point and $\left\{\alpha q : \alpha \in \mathbb{R}\right\}$ are convex sets. That is
\begin{align*}
\exists\ \text{ nonzero }\ v: \mu R v  \ge a \ge rv\ \forall\ r\ \in\left\{\alpha q : \alpha \in \mathbb{R}\right\}
\end{align*}
Since the latter space is a linear
space, just as in the previous step, $\forall\ r\ \in \left\{\alpha q : \alpha \in \mathbb{R}\right\},\ -r\ \in \left\{\alpha q : \alpha \in \mathbb{R}\right\} $ and as a result $rv = 0\ \forall\ r\in\left\{\alpha q : \alpha \in \mathbb{R}\right\}$. This leads to a separating normal vector $v \in \mathbb{R}^K$ such that
$qv = 0$ and $(\mu R)v = \mu (Rv) \ge 0$. Since $\mu >0$ and each column of $R$ also $>0$ hence $\mu R >0$ therefore if $z \neq 0$ and $\mu R z \ge 0$ then $\mu R v > 0$. Letting $\phi \equiv Rv \in \Phi$ yields a contradiction to the fact that $\mu \phi = 0\ \forall\ \phi$. The contradiction shows that $\mu R$ belongs to the linear space $\left\{\alpha q : \alpha \in \mathbb{R}\right\}$.

\textbf{Step 3}

As it has been concluded in step 2 $\mu R >0$. Moreover $q >> 0$ because otherwise no-arbitrage condition would be violated (assume there exists some coordinate $s$ such that $q_s \le 0$ then there exists $z = \mathbb{I}_s \lambda$ for some $\lambda > 0$ such that $qz \le 0$ however since there exists $s$ such that $r_{sk} >0$ hence $Rz > 0$).
Since $\mu R = \alpha q$ it follows that $\alpha >0$ but that means that
\begin{align*}
\exists\ \hat{\mu} = \frac{1}{\alpha}\mu >0: \hat{\mu}R = q
\end{align*}
and this completes the proof.
\end{document}

### 2019-02-11 gameth2.tex
gametheory

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Game theory. Problem Set 2.
\end{center}
\section*{Problem 1}
Whether candidate 1 or candidate 2 is elected depends on the votes of two
citizens. The world may be in one of two states, A or B. The citizens agree
that candidate 1 is best if the state is A and candidate 2 is best if the state
is B. Each citizen’s preferences are represented by the expected value of a
Bernoulli payoff function that assigns a payoff of 1 if the best candidate
for the state wins (obtains more votes than the other candidate), a payoff
of zero if the other candidate wins, and a payoff of $\frac{1}{2}$ if the candidates tie.
Citizen 1 is informed of the state of the world, whereas citizen 2 believes
it’s A with probability .9 and state B with probability .1. Each citizen may
vote for candidate 1, vote for candidate 2, or abstain (not vote). Find the
Bayes-Nash equilibrium (or equilibria) of this game involving each player
playing a pure strategy (if any) separately for each of the the following
assumptions:
\begin{enumerate}[(a)]
\item Citizens vote simultaneously.
\item Citizen 1 votes first and citizen 2 votes after observing citizen 1’s vote.
Does any of the equilibria you found involve a weakly dominated strategy?
\end{enumerate}

\textbf{Solution}

\begin{enumerate}[(a)]
	\item The tree of the game is depicted below:
	\begin{figure}[H]
		\centering
		\includegraphics[width=0.8\textwidth]{plotdraft}
		\caption{}\label{fig2}
	\end{figure}
	Obviously for player 1 the strategy $\left\{12\right\}$ (i.e. vote for candidate $1$ if A occurs and for candidate 2 if B occurs) is a weakly dominant strategy: regardless of what the player 2 does the player 1 gets greater than or equal payoff. I claim that the profile of strategies 
	\begin{align*}
	\left\{12; \text{abstain}\right\}
	\end{align*}
	is a Nash equilibria (and hence a SPE). Let us prove it. In playing the aforementioned profile, player's 2 expected payoff is
	\begin{align*}
	0.9 \cdot 1 + 0.1 \cdot 1 = 1
	\end{align*}
	If he deviates and plays $1$ or $2$ then his expected payoffs will be:
	\begin{align*}
	0.9 \cdot 1 + 0.1 \cdot 0.5 < 1\\
	0.9 \cdot 0.5 + 0.1 \cdot 1 < 1
	\end{align*}
	If the first player deviates and plays $\left\{\text{abstain}\ 2\right\}$, $\left\{\text{abstain}\ \text{abstain}\right\}$ or $\left\{1\ \text{abstain}\right\}$ (other strategies are striclty dominated and cannot be played in the equilibrium) then his expected payoffs will be:
	\begin{align*}
	\left\{\text{abstain}\ 2\right\}&: 0.9 \cdot 0.5 + 0.5 + 0.1 \cdot 1 < 1\\
	\left\{\text{abstain}\ \text{abstain}\right\}&: 0.9 \cdot 0.5 + 0.1 \cdot 0.5 = 0.5 < 1\\
	\left\{1\ \text{abstain}\right\}&: 0.9 \cdot 1 + 0.1 \cdot 0.5 < 1
	\end{align*}
	The profile of strategies $\left\{\text{abstain } 2; 1 \right\}$ is also a SPE because:
	\begin{align*}
	&u_2(\left\{\text{abstain } 2; 1 \right\}) = 0.95 > u_2(\left\{\text{abstain } 2; \text{ abstain } \right\}) = 0.55 > u_2(\left\{\text{abstain } 2; 2 \right\}) = 0.05\\
	&u_1(\left\{\text{abstain } 2; 1 \right\}) = 0.95 \ge u_1(\left\{1 2; 1 \right\}) = 0.95 > u_1(\left\{\text{abstain } \text{ abstain }; 1 \right\}) = 0.9 \ge \\
	&\ge u_1(\left\{1\text{ abstain }; 1 \right\}) = 0.9
	\end{align*}
	 To prove that there are no another equilibria below is a list of profitable unilateral deviations for another profiles of strategies (note that strictly dominated strategies are not taken into account)
	\begin{align*}
	u_2(\left\{\text{abstain } \text{ abstain }; 2 \right\}) &< u_2(\left\{\text{abstain } \text{ abstain }; 1 \right\})\\
	u_2(\left\{1 \text{ abstain }; 2 \right\}) &< u_2(\left\{1 \text{ abstain }; 1 \right\})
	\end{align*}
	Thus, there are two Bayes-Nash equilibria, namely:
	\begin{align*}
	\left\{12; \text{abstain}\right\}\\
	\left\{\text{abstain } 2; 1 \right\}
	\end{align*}
	The second equilibrium profile contains a weakly dominated strategy for the first player, it is easy to observe that the strategy $\left\{12\right\}$ gives the first player weakly greater payoff (than $\left\{\text{ abstain } 2\right\}$) regardless of what the second player does.

	\item If the first player votes for the 1st candidate than it is optimal for the second player to vote either for the 1st candidate or to abstain, both alternatives yield him equal expected payoffs $0.9$.

	If the first player abstains from voting then it is optimal for the second player to vote for the 1st candidate it yields him the greatest expected payoffs $0.9$.

	If the first player votes for the second candidate then it is optimal for the second player to vote for the 1st candidate it yields him the greates expected payoffs $0.5$.

	As a result the equilibria are (strategy $121$ means that if 1st player votes for 1st or for 2nd candidate - vote for the 1st candidate, if he abstained - vote for the second candidate):
	\begin{align*}
	&\left\{11\ ; 111\ \right\}\\
	&\left\{11\ ; \text{abstained }11\ \right\}\\
	&\left\{1 \text{ abstained }\ ; 111\ \right\}\\
	&\left\{1\text{ abstained }\ ; \text{ abstained }11\ \right\}\\
	&\left\{\text{ abstained } 1\ ; 111\ \right\}\\
	&\left\{\text{ abstained } 1\ ; \text{ abstained }11\ \right\}\\
	&\left\{\text{ abstained }\text{ abstained }\ ; 111\ \right\}\\
	&\left\{\text{ abstained }\text{ abstained }\ ; \text{ abstained }11\ \right\}\\	
	\end{align*}
	As in previous point, for player 1, to abstain in the first state is weakly dominated strategy by strategy of voting for the 1st candidate, because it yields him weakly smaller payoffs regardless of what player 2 does.
	\end{enumerate}

\section*{Problem 2}

\textbf{Solution}

I will denote a profile of strategies as $\left\{A, B, CD\right\}$ where $A$, $B$ are some actions of players 1 and 2 respectively, and $CD$ means that player 3 plays action $C$ once he is called upon to play in history $\left\{D\right\}$ and $D$ once he is called upon to play in history $\left\{Cd\right\}$

Obviously profiles of strategies
\begin{align*}
\left\{C, c, LR\right\}\\
\left\{C, c, LL\right\}\\
\end{align*}
are Nash equilibria because nobody has a profitable unilateral deviation: player 1 gets 1 instead of 3 if deviates, player 2 is not called upon to play hence his action cannot somehow affect his payoffs, and player 3 if deviates and plays $R$ instead of $L$ gets 0 instead of 2. The strategy
\begin{align*}
\left\{C, d, LR\right\}
\end{align*}
also constitutes Nash equilibrium because if player 1 decides to deviate he gets 0 instead of 3, and player 3 gets 0 instead of 2. The strategies
\begin{align*}
\left\{D, c, RR\right\}\\
\left\{D, d, RR\right\}
\end{align*}
are also Nash equilibria because both players 1 and 2 cannot get more by deviating, whereas player 3 gets strictly less, playing another strategy in the second node. Thus, all Nash equilibria in this game are:
\begin{align*}
\left\{C, c, LR\right\}\\
\left\{C, c, LL\right\}\\
\left\{C, d, LR\right\}\\
\left\{D, c, RR\right\}\\
\left\{D, d, RR\right\}
\end{align*}
It is easy to check that there are no other pure-strategies equilibria.

Among these Nash equilibria, strategies
\begin{align*}
\left\{C, d, LR\right\}\\
\left\{C, c, LR\right\}\\
\end{align*}
are SPE, because for player 3 it is optimal to play $L$ in the history $\left\{C\right\}$ and $R$ in the history $\left\{Dd\right\}$, player 2 is indifferent, and consequently for the 1st player it is optimal to play $C$. Thus, these profiles consits of strategies, optimal for each subgame.
\section*{Peroblem 3}
There is a seller who holds a good and can supply it at a cost $c > 0$. There
is also one buyer who values the good for $v > c$. They negotiate over the good and attempt to settle the deal. Their discount rate for the future is
$\delta \in (0, 1)$: if the good is sold at a price $p$ at period $t$, at period $s \le t$ the
buyer estimates his utility from the deal to be $\delta^{t-s}(v - p)$. Similarly, the
seller values the deal at $\delta^{t-s}(p-c)$. If the deal does not happen, both agents
receive zero utility. All the parameters are common knowledge. Assume
that the buyer always buys and the seller always sells when indifferent.
\begin{enumerate}[(a)]
\item Suppose the seller can make a take-it-or-leave-it offer $p$ at $t = 1$. Then,
the buyer either accepts or rejects it, and the game terminates. What
is the set of pure-strategy Nash equilibria in this game? What is the
Sub-game Perfect Equilibrium?
\item Assume the game does not terminate after the buyer answers the offer.
If he rejects at $t = 1$, at $t = 2$ he can make an offer in response,
proposing his own price. The seller then either accepts or rejects it, and
the game terminates. What is the subgame-perfect Nash equilibrium?
\item Now suppose the seller can make her offers at $t = 1$ and $t = 2$, and
the buyer is able to propose a price only at $t = 3$ (after which, again,
the game terminates). What is the Sub-game Perfect Equilibrium?
(Hint: if the buyer rejects the offer at $t = 1$, at $t = 2$ the game from
the previous point will begin, delivering the same payoffs in its SPE.)
\end{enumerate}

\textbf{Solution}

\begin{enumerate}[(a)]
	\item The stratefy of seller is a number $b \in \mathbb{R}$. His payoff is
	\begin{align*}
	u_s(b) = \begin{cases}
	b - c, &\text{the deal occurs}\\
	0, &\text{otherwise}
	\end{cases}
	\end{align*}
	taking $b$ as given strategy of buyer is a binary decision accept offer or reject. His payoff is:
	\begin{align*}
	u_b(b) = \begin{cases}
	v - b, &\text{if he accepts}\\
	0, &\text{otherwise}
	\end{cases}
	\end{align*}
	The best response of buyer is:
	\begin{align*}
	r_b(b) = \begin{cases}
	\text{accept}, &\text{ if } b \le v\\
	\text{reject}, &\text{ if } b > v
	\end{cases}
	\end{align*}
	Hence the optimal strategy of seller is $b = v$, and the equilibrium profile of strategies is
	\begin{align*}
	\left\{v, \text{accept}\right\}
	\end{align*}
	it is a SPE. However this game also possesses another equilibrium. The profile of strategies
	\begin{align*}
	\left\{b, \text{reject} \right\},\ b > v
	\end{align*}
	is also a Nash equilibrium because any unilateral deviations will not increase players' payoffs.
	\item Denote seller's bid as $b_s$, buyer's bid as $b_b$, and will denote profiles of strategies as
	\begin{align*}
	\left\{b_sR, Ab_b\right\}
	\end{align*}
	where $b_sR$ means that seller bids $b_s$ at $t = 1$ and rejects the offer at $t = 1$ and the same things for the buyer. The game tree is depicted below
	\begin{figure}[H]
		\centering
		\includegraphics[width=0.8\textwidth]{plotdraft}
		\caption{}\label{fig1}
	\end{figure}
	If at $t = 2$ the seller is called upon to play then his optimal strategy is:
	\begin{align*}
	\begin{cases}
	\text{accept}, &\text{ if } \delta(b_b - c) \ge 0\\
	\text{reject}, &\text{ if } \delta(b_b - c) < 0
	\end{cases}
	\end{align*}
	When buyer is called upon to play at his second node then his optimal strategy is $b_b = c$. When buyer is called upon to play at his first node and should decide whether to accept offer or not his optimal strategy is:
	\begin{align*}
	\begin{cases}
	\text{accept}, &\text{ if } b_s \le v - \delta(v - c)\\
	\text{reject}, &\text{ if } b_s > v - \delta(v - c)
	\end{cases}
	\end{align*}
	And finally at $t = 1$ optimal strategy for the seller will be $b_s = v - \delta(v - c)$. Thus, the subgame perfect equilibrium is:
	\begin{align*}
	\left\{v - \delta(v - c)A, Ac\right\}
	\end{align*}
	\item Since we have already analysed the subgame, starting from $t = 2$, at $t = 1$ when buyer is called upon to play his optimal strategy will be:
	\begin{align*}
	\begin{cases}
	\text{accept}, & \text{ if } v - b_s^{t=1} \ge \delta(v - v + \delta(v-c))\\
	\text{reject}, &\text{ if } v - b_s^{t=1} < \delta(v - v + \delta(v-c))
	\end{cases}
	\end{align*}
	hence optimal strategy of seller at $t = 1$ is:
	\begin{align*}
	b_s^{t=1} = v - \delta^2(v - c)
	\end{align*}
	Thus, the SPE profile of strategies is:
	\begin{align*}
	\left\{v - \delta^2(v - c),v - \delta(v - c),A\ ;\  A,A,c \right\}
	\end{align*}
\end{enumerate}
\section*{Problem 4}
Consider a game with two players, Player 1 and Player 2. Each player
$i \in \left\{1, 2\right\}$ can choose an action $a_i$ from a finite set of actions $A_i$. Player
$i$’s payoff at any action profile $(a_1, a_2)\in A_1 \times A_2$ is $f_i(a_1, a_2)$. Suppose
first that the two players move simultaneously. How many strategies does
each player have? Now suppose that player 1 moves first and that player 2
observes player 1’s move before choosing her move. How may strategies does
each player have now? Moreover, suppose that this latter game has multiple
sub-game perfect Nash equilibria. Show that if this is the case, then there
exist two pairs of actions $(a_1', a_2')$ and $(a_1'', a_2'')$ (with either $a_1' \neq a_2'$ or
$a_1'' \neq a_2''$ or both) such that either
\begin{align*}
f_1(a_1', a_2') &= f_1(a_1'', a_2'')\ \text{ or }\\
f_2(a_1', a_2') &= f_2(a_1'', a_2'')
\end{align*}

\textbf{Solution}

Players have infinitely many mixed strategies if $|A_i| > 1\ \forall\ i \in \left\{1, 2\right\}$. If both move simulteneously than each has $|A_i|$ pure strategies ($|\cdot|$ means cardinality of set). If player 1 moves first than he has still $|A_1|$ pure strategies, whereas player 2, who observes the player's 1 actions now have $|A_1| \times |A_2|$ pure strategies, because for each action of player 1 player 2 has $|A_2|$ actions to respond.

Obviously, if there exists a unique best response $r_2(a_1)$ of player 2 to each action $a_1$ of player 1, and moreover the set
\begin{align*}
\underset{a_i}{\text{argmax}}\ f_1(a_i, r_2(a_i))
\end{align*} 
is a singleton, then the profile of strategies \begin{align*}
\left\{\underset{a_i}{\text{argmax}}\ f_1(a_i, r_2(a_i)); r_2(\underset{a_i}{\text{argmax}}\ f_1(a_i, r_2(a_i)))\right\}
\end{align*} 
is a uniqiue SPE. Suppose now that there exists several SPE, it can be possible either if the set
\begin{align*}
\underset{a_i}{\text{argmax}}\ f_1(a_i, r_2(a_i))
\end{align*}
contains more than one element, or
\begin{align*}
\exists\ a_j\ \in A_1: |\underset{b_i \in A_2}{\text{argmax}}\ f_2(a_j, b_i)| > 1
\end{align*}
or both. If the set
\begin{align*}
\underset{a_i}{\text{argmax}}\ f_1(a_i, r_2(a_i))
\end{align*}
contains more than one element than
\begin{align*}
\exists\ \text{ different } a_1', a_2'' \in \underset{a_i}{\text{argmax}}\ f_1(a_i, r_2(a_i))\ \&\ \exists\ a_2' = r_2(a_1'), a_2'' = r_2(a_1'')\ : f_1(a_1', a_2') = f_1(a_1'', a_2'')
\end{align*}
If
\begin{align*}
\exists\ a_j\ \in A_1: |\underset{b_i \in A_2}{\text{argmax}}\ f_2(a_j, b_i)| > 1
\end{align*}
then
\begin{align*}
\exists\ a_1' = a_j, a_1'' = a_j\ \&\ \exists\ \text{different } a_2', a_2'' \in \underset{b_i \in A_2}{\text{argmax}}\ f_2(a_j, b_i):\ f_2(a_1', a_2') = f_2(a_1'', a_2'')
\end{align*}
The case $a_1' \neq a_1''\ \&\ a_2' \neq a_2''$ is analoguos.
Q.E.D.
\end{document}

### 2019-02-11 macro3_3.tex
Macro3

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Macroeconomics 3. Problem Set 3.
\end{center}
\section*{Problem 1}
\begin{enumerate}[1.]
	\item \begin{enumerate}[(a)]
		\item SOMCE are functions
		\begin{align*}
		\left\{c(t), l(t), i(t), b(t), w(t), r(t)\right\}
		\end{align*}
		such that
		\begin{itemize}
			\item Taking $w(t)$ as given households choose $\left\{c(t), l(t), b(t)\right\}$ as a solution to the following optimization problem:
			\begin{align*}
			&\max\ \int_{0}^{\infty} e^{-\rho t}u(c(t))\\
			&s.t.\ \dot{b}(t) = w(t)l(t) -c(t) + r(t)b(t),\ \forall\ t\\
			&\lim_{t \to \infty} b(t)e^{-\int_{0}^t r(\tau)d\tau} \ge 0, c(t) \ge 0, 0 \le l(t) \le 1, b(0) = 0
			\end{align*}
			\item Taking $w(t)$ as given, firms choose $\left\{l(t), i(t)\right\}$ as a solution to the following optimization problem:
			\begin{align*}
			\max\ \int_{0}^{\infty} &(F(k(t), l(t)) - w(t)l(t) - i(t))e^{-\rho t}dt\\
			&s.t.\ \dot{k}(t) = i(t) - \delta k(t)\\
			&k(t) \ge 0, k(0) > 0 \text{ is given}
			\end{align*}
			\item Markets are cleared, i.e.
			\begin{align*}
			c(t) + i(t) &= F(k(t), l(t)), \forall\ t\\
			b(t) &= 0, \forall t
			\end{align*}
		\end{itemize}
	\item $l(t) = 1,\ \forall t$ since there is no disutility from labour. Hamiltonian for households looks as follows
	\begin{align*}
	\mathscr{H}(c(t), b(t)) = e^{-\rho t}u(c(t)) + \lambda(t)(w(t)l(t) - c(t) + r(t)b(t))\\
	\end{align*}
	and optimality conditions are:
	\begin{align*}
	\begin{cases}
	e^{-\rho t}u'(c(t)) = \lambda(t)\\
	-\lambda(t)r(t) = \dot{\lambda}(t)\\
	\lim_{t \to \infty} \lambda(t)b(t) = 0
	\end{cases}\\
	\frac{\dot{c}(t)}{c(t)} = -\frac{u'(c(t))}{u''(c(t))c(t)}(r(t) - \rho)
	\end{align*}
	Firms have two control variables $l(t)$ and $i(t)$ and one state variable $k(t)$. Hamiltonian looks as follows:
	\begin{align*}
	\mathscr{H}(i(t), l(t), k(t)) = e^{-\rho t}(F(k(t), l(t)) - w(t)l(t) - i(t)) + \gamma(t)(i(t) - \delta k(t))
	\end{align*}
	then optimality conditions are:
	\begin{align*}
	\begin{cases}
	F'_l(k(t), l(t)) = w(t)\\
	\gamma(t) = e^{-\rho t}\\
	\dot{\gamma}(t) = -e^{-\rho t}F'_k(k(t), l(t)) + \gamma(t)\delta\\
	\lim_{t \to \infty} \gamma(t) k(t) = 0
	\end{cases} \to \begin{cases}
	F'_k(k(t), l(t)) = \rho + \delta\\
	F'_l(k(t), l(t)) = w(t)
	\end{cases}
	\end{align*}
	Then, the dynamic system could be:
	\begin{align*}
	\begin{cases}
	\frac{\dot{c}(t)}{c(t)} = -\frac{u'(c(t))}{u''(c(t))c(t)}(r(t) - \rho)\\
	F'_k(k(t), 1) = \rho + \delta\\
	\dot{k}(t) = F(k(t), 1) - c(t) - \delta k(t)
	\end{cases}
	\end{align*}
	No-arbitrage condition imply that at the equilibrium, firms should be indifferent between investments in capital and investments in consumers' bonds, i.e. gains from bonds and capital should be equal to each other, i.e. $r(t) = F'_k(k(t), l(t)) - \delta$. As a result, the dynamic system becomes equivalent to the social planner's one:
	\begin{align*}
	\begin{cases}
	\frac{\dot{c}(t)}{c(t)} = -\frac{u'(c(t))}{u''(c(t))c(t)}(F'_k(k(t), 1) - \rho - \delta)\\
	\dot{k}(t) = F(k(t), 1) - c(t) - \delta k(t)
	\end{cases}
	\end{align*}	
	Of course its solution coincides with the planner's solution.
	\item Consumers manage to save through firms i.e. if in some period consumers decide to consume less (to save) then market clearing condition implies that firms will have to invest more in capital, as a result more capital leads to more marginal product of labour, which means higher wages for consumers. Thus, by consuming less households automatically force firms to invest more. At equilibrium $r(t)$ is set in such a way as to eliminate any possibilities for arbitrage. With equilibrium interest rate firms are indifferent between investments in capital and investments in bonds, and consumers are indifferent to borrowing or lending.
	\end{enumerate}
	\item 
	\begin{align*}
	c(t) + b(t + 1) + p(t) a(t + 1) = w(t)l(t) + (1 + r (t)) b(t) + p(t) a(t) + d(t)a(t)
	\end{align*}
	\begin{enumerate}[(a)]
		\item 
		\begin{align*}
		&c(t)\Delta t + b(t + \Delta t) + p(t)a(t + \Delta t) = w(t)l(t)\Delta t + (1 + r(t)\Delta t)b(t) + p(t)a(t) + \\
		&+d(t)a(t)\Delta t\\
		&\dot{b}(t) + p(t)\dot{a}(t) = w(t)l(t) + r(t)b(t) + d(t)a(t) - c(t)
		\end{align*}
		Now, SOMCE are functions
		\begin{align*}
		\left\{c(t), l(t), a(t), i(t), b(t), w(t), r(t)\right\}
		\end{align*}
		such that
		\begin{itemize}
			\item Taking $w(t), p(t)$ as given households choose $\left\{c(t), l(t), b(t), a(t)\right\}$ as a solution to the following optimization problem:
			\begin{align*}
			&\max\ \int_{0}^{\infty} e^{-\rho t}u(c(t))\\
			&s.t.\ \dot{b}(t) + p(t)\dot{a}(t) = w(t)l(t) + r(t)b(t) + d(t)a(t) - c(t),\ \forall\ t\\
			&\lim_{t \to \infty} b(t)e^{-\int_{0}^t r(\tau)d\tau} \ge 0, c(t) \ge 0, 0 \le l(t) \le 1, b(0) = 0, a(t) \le 1\\
			\end{align*}
			\item Taking $w(t)$ as given, firms choose $\left\{l(t), i(t)\right\}$ as a solution to the following optimization problem:
			\begin{align*}
			\max\ \int_{0}^{\infty} &(F(k(t), l(t)) - w(t)l(t) - i(t) - d(t))e^{-\rho t}dt\\
			&s.t.\ \dot{k}(t) = i(t) - \delta k(t)\\
			&k(t) \ge 0, k(0) > 0 \text{ is given}
			\end{align*}
			\item Markets are cleared, i.e.
			\begin{align*}
			c(t) + i(t) + d(t)&= F(k(t), l(t)), \forall\ t\\
			b(t) &= 0,\ \forall\ t\\
			a(t) &= 1,\ \forall\ t
			\end{align*}
		\end{itemize}
	\item $x(t) = p(t)a(t)$. Then consumer's problem is:
	\begin{align*}
	&\max\ \int_{0}^{\infty} e^{-\rho t}u(c(t))\\
	&s.t.\ \dot{b}(t) + \dot{x}(t) = w(t)l(t) + r(t)b(t) + \frac{d(t) + \dot{p}(t)}{p(t)}x(t) - c(t),\ \forall\ t\\
	&\lim_{t \to \infty} b(t)e^{-\int_{0}^t r(\tau)d\tau} \ge 0, c(t) \ge 0, 0 \le l(t) \le 1, b(0) = 0, \frac{x(t)}{p(t)} \le 1\\
	\end{align*}
	No-arbitrage condition implies that at the equilibrium both assets should yield the same interest, that means that
	\begin{align*}
	r(t) = \frac{d(t) + \dot{p}(t)}{p(t)}
	\end{align*}
	it means that asset yields its owner a gain in a form of dividends $\frac{d(t)}{p(t)}$ and in a form of price change $\frac{\dot{p}(t)}{p(t)}$.
	\item Economic interpretation is the following: the price of asset is equal to discounted sum of all dividends it pays.
	\end{enumerate}
\end{enumerate}
\end{document}

### 2019-02-12 gameth2.tex
gt

@@
 
-	\item If the first player votes for the 1st candidate than it is optimal for the second player to vote either for the 1st candidate or to abstain, both alternatives yield him equal expected payoffs $0.9$.
+	\item Let us take a look at the game from the point of view of second player and use backward induction approach.
+Best responses of the second player on non-strictly dominated strategies of the 1st player are the following (here $a$ stands for "abstain" and as in previous part $121$-strategy means that the second player votes for 1 if 1st player votes for 1 or 2 and votes for 2 if 1st player abstains, $\forall$ - means that any action is possible):
+	\begin{itemize}
+	\item \begin{align*}
+	r_2(\left\{12\right\}) =\begin{cases} \left\{1\forall a \right\}\\
+	\left\{1\forall 2 \right\}\\
+	\left\{a\forall 2 \right\}\\
+	\left\{a\forall a \right\}\\
+	\end{cases}
+	\end{align*}
+	Of course, 
+	\begin{align*}
+	\left\{12\right\} \in r_1(r_2(\left\{12\right\}))
+	\end{align*}
+	hence all 12 profiles of strategies are equilibria. 
+	\item \begin{align*}
+	r_2(\left\{1a\right\}) = \begin{cases}
+	\left\{12\forall \right\}\\
+	\left\{a2\forall \right\}
+	\end{cases}
+	\end{align*}
+	Again
+	\begin{align*}
+	\left\{1a\right\} \in r_1(r_2(\left\{1a\right\}))
+	\end{align*}
+	hence 6 more profiles of strategies are equilibria.
 
-	If the first player abstains from voting then it is optimal for the second player to vote for the 1st candidate it yields him the greatest expected payoffs $0.9$.
-
-	If the first player votes for the second candidate then it is optimal for the second player to vote for the 1st candidate it yields him the greates expected payoffs $0.5$.
-
-	As a result the equilibria are (strategy $121$ means that if 1st player votes for 1st or for 2nd candidate - vote for the 1st candidate, if he abstained - vote for the second candidate):
+	\item \begin{align*}
+	r_2(\left\{a2\right\}) = \begin{cases}
+	\left\{\forall 1 2\right\}\\
+	\left\{\forall 1 a\right\}
+	\end{cases}
+	\end{align*}
+	$\left\{a2\right\} \in r_1(r_2(\left\{a2\right\}))$ hence these strategies are also equilibria.
+	\item \begin{align*}
+	r_2(\left\{aa\right\}) = \left\{\forall 1 \forall \right\}
+	\end{align*}
+	however 
 	\begin{align*}
-	&\left\{11\ ; 111\ \right\}\\
-	&\left\{11\ ; \text{abstained }11\ \right\}\\
-	&\left\{1 \text{ abstained }\ ; 111\ \right\}\\
-	&\left\{1\text{ abstained }\ ; \text{ abstained }11\ \right\}\\
-	&\left\{\text{ abstained } 1\ ; 111\ \right\}\\
-	&\left\{\text{ abstained } 1\ ; \text{ abstained }11\ \right\}\\
-	&\left\{\text{ abstained }\text{ abstained }\ ; 111\ \right\}\\
-	&\left\{\text{ abstained }\text{ abstained }\ ; \text{ abstained }11\ \right\}\\	
+	\left\{aa\right\} \notin r_1(r_2(\left\{aa\right\}))
 	\end{align*}
-	As in previous point, for player 1, to abstain in the first state is weakly dominated strategy by strategy of voting for the 1st candidate, because it yields him weakly smaller payoffs regardless of what player 2 does.
+	because once the state $b$ occurs it is more profitable for the 1st player to deviate and to vote for the second candidate. Hence, it is not an equilibrium profile.
+	\end{itemize}
+	As in previous point, for player 1, to abstain in both states is weakly dominated strategy by strategy of voting for the best candidate for the particular state, because it yields him weakly smaller payoffs regardless of what player 2 does.
 	\end{enumerate}

### 2019-02-12 macro3_3.tex
Macro

@@
 			\begin{align*}
-			&\max\ \int_{0}^{\infty} e^{-\rho t}u(c(t))\\
+			&\max\ \int_{0}^{\infty} e^{-\rho t}u(c(t))dt\\
 			&s.t.\ \dot{b}(t) = w(t)l(t) -c(t) + r(t)b(t),\ \forall\ t\\
@@
 			\begin{align*}
-			&\max\ \int_{0}^{\infty} e^{-\rho t}u(c(t))\\
+			&\max\ \int_{0}^{\infty} e^{-\rho t}u(c(t))dt\\
 			&s.t.\ \dot{b}(t) + p(t)\dot{a}(t) = w(t)l(t) + r(t)b(t) + d(t)a(t) - c(t),\ \forall\ t\\
@@
 	\begin{align*}
-	&\max\ \int_{0}^{\infty} e^{-\rho t}u(c(t))\\
+	&\max\ \int_{0}^{\infty} e^{-\rho t}u(c(t))dt\\
 	&s.t.\ \dot{b}(t) + \dot{x}(t) = w(t)l(t) + r(t)b(t) + \frac{d(t) + \dot{p}(t)}{p(t)}x(t) - c(t),\ \forall\ t\\
@@
 	\item Economic interpretation is the following: the price of asset is equal to discounted sum of all dividends it pays.
+	\item 
+	\item \begin{align*}
+	d(t) = r(t)k(t) - \dot{k}(t)
+	\end{align*}
+	\item
 	\end{enumerate}
 \end{enumerate}
+\section*{Problem 2}
+\begin{enumerate}
+	\item \begin{align}\label{eq1}
+	&\underset{c(t), l(t), a(t)}{\max}\ \int_{0}^{\infty} e^{-\rho t} u(c(t))dt\\
+	&s.t.\ \dot{a}(t) = w(t)l(t) + (r(t) - \tau - \delta)a(t) - c(t) - z(t)\nonumber\\
+	&\lim_{t \to \infty} a(t)e^{-\int_{0}^t r(\tau)d\tau} \ge 0, c(t) \ge 0, 0 \le l(t) \le 1, a(0) = k(0) > 0 \text{ is given }\nonumber
+	\end{align}
+	\item Taking $r(t), w(t)$ as given for each $t$ firms are solving the following optimization problem:
+	\begin{align}\label{eq2}
+	\underset{K(t), L^f(t)}{\max}\ F(K(t), L^f(t)) - r(t)K(t) - w(t)L^f(t)
+	\end{align}
+	\item The equilibrium is a set of functions $\left\{c(t), l(t), a(t), K(t), L^f(t), r(t), w(t), z(t)\right\}$ such that:
+	\begin{itemize}
+		\item Taking $w(t), z(t), r(t)$ as given, consumers choose $c(t), l(t), a(t)$ as a solution to optimization problem \eqref{eq1}
+		\item Taking $r(t), w(t)$ as given for each $t$ firms choose $K(t), L^f(t)$ as a solution to optimization problem \eqref{eq2}.
+		\item Markets are cleared and budget is balanced, i.e.
+		\begin{align*}
+		\forall\ t:\ &K(t) = A(t) = a(t) \cdot L(t)\\
+		&l(t) = L^f(t)\\
+		&c(t) + \dot{k}(t) + g = F(K(t), L^f(t)) - \delta k(t)\\
+		&g = \tau k(t) + z(t)
+		\end{align*}
+	\end{itemize}
+\item Assuming CRS production function, firms' problem can be rewritten in a following manner:
+\begin{align*}
+\max\ L^f(F(k_t, 1) - r(t)k(t) - w(t))
+\end{align*}
+Then FOCs imply:
+\begin{align*}
+&F'_k(k(t), 1) = f'(k(t)) = r(t) \to k(t)^*\\
+L^f = &\begin{cases}
+\infty, w(t) < F(k(t)^*, 1) - r(t)k^*(t)\\
+0, w(t) > F(k(t)^*, 1) - r(t)k^*(t)\\
+\forall, w(t) = F(k(t)^*, 1) - r(t)k^*(t)
+\end{cases}
+\end{align*}
+Incorporating market clearing:
+\begin{align*}
+L^f(t) &= l(t)\\
+w(t) &= F(k^*(t), 1) - r(t)k^*(t), \text{ where } k^*(t) = f'^{-1}(r(t))
+\end{align*}
+\item Firms have zero profit: the whole output is used to pay for labour and capital.
+\item Hamiltonian is:
+\begin{align*}
+\mathscr{H}(c(t), l(t), a(t)) = e^{-\rho t}u(c(t), l(t)) + \lambda(t)(w(t)l(t) + (r(t) - \tau)a(t) - c(t) - z(t))
+\end{align*}
+Since there are no disutility from labour, $l(t) = 1\ \forall\ t$.
+Optimality conditions:
+\begin{align*}
+\begin{cases}
+e^{-\rho t}u'(c(t)) - \lambda(t) = 0\\
+\dot{\lambda}(t) = -\lambda(t)(r(t) - \tau - \delta)\\
+\lim_{t \to \infty} a(t)\lambda(t) = 0
+\end{cases}\\
+\end{align*}
+and the system is:
+\begin{align*}
+\begin{cases}
+\frac{\dot{c}(t)}{c(t)} = -\frac{u'(c(t))}{u''(c(t))c(t)}(\rho - r(t) + \tau + \delta)\\
+\dot{k}(t) = F(k(t), 1) - \delta k(t) - c(t) - g
+\end{cases}
+\end{align*}
+\item The picture is provided below:
+	\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{plotdraft}
+	\caption{}\label{fig1}
+\end{figure}
+
+Yes, if a subsidy $-\tau = \rho$ then $f'(\bar{k}) = \delta$ which corresponds to a golden rule.
+Since this steady state does not coincide with the social planner's one, which as we know is the first best, it means that actually consumers are not benefited from translating into the golden-rule steady state.
+\item If the goverment increases $\tau$ then the isocline $\dot{c}(t) = 0$ moves to the left, that means that new steady state will correspond to lower consumption and lower capital (see Fig \ref{fig1}).
+Hence in order to move to a new saddle path consumption should be increased firstly after that it is decreasing gradually.
+Intuition is straightforward: high tax rate on capital - consumers do not want to own as much capital - they start consume more, invest less, as a consequences amount of capital decreases.
+\item In that case consumption should be increased at $t=0$ to attain the trajectory of the initial system which exactly at $t=T$ leads to the saddle path of the modified system.
+\item If at time $t=T$ government decides not to increase $\tau$, the economy should decrease consumption in order to attain the saddle path of initial system (see Fig \ref{fig2}).
+Also, the straightforward intuition: it turns out that tax rate is not increased, but we already have too less capital, that means it's time to "tighten belts" and start to amass capital.
+	\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{plotdraft}
+	\caption{}\label{fig2}
+\end{figure}
+\item In that case at $t = 0$ consumption should be increased in order to move to the trajectory of new system which exactly at $t = T$ leads to a saddle path of the initial system.
+
+	\end{enumerate}
 \end{document}

### 2019-02-14 matrics5.tex
metrics

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\newcommand{\yrseduc}{\textit{yrseduc}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Econometrics 1. Problem Set 5.
\end{center}
\section*{Problem 1}
Consider a simple model to estimate the effect of personal computer (PC) ownership on
college grade point average for graduating seniors at a large public university:
\begin{align*}
GPA = \beta_0 + \beta_1 PC + u
\end{align*}
where $PC$ is a binary variable indicating $PC$ ownership.
\begin{enumerate}[(i)]
\item Why might PC ownership be correlated with $u$?
\item Explain why $PC$ is likely to be related to parents’ annual income. Does this mean
parental income is a good IV for $PC$? Why or why not?
\item Suppose that, four years ago, the university gave grants to buy computers to roughly
one-half of the incoming students, and the students who received grants were randomly chosen.
Carefully explain how you would use this information to construct an instrumental variable for
$PC$.
\end{enumerate}

\textbf{Solution}

\begin{enumerate}[(i)]
	\item Variety of reasons are possible. For example, people who own PCs can spend a lot of time on playing computer games, as a result they pay less attention to studying, hence their GPA can be affected. That is, error term contains for example a number of hours a person spend on computer games, which of course correlated with PC ownership.
	\item PC is not very cheap purchase, so it can in principle be related to parents' income: smaller income - smaller probability that parents can afford purchasing a PC. Parental income is good IV if it is valid, i.e. is not correlated with error term. As for me, it is not a good IV because I suppose that children from not very rich families can tend to have a higher motivation for studying than their rich counterparts, as a result parental income can be negatively correlated with $GPA$.
	\item We can use as IV a dummy variable, whether or not a particular student have been chosen as a grant taker. Obviously it is not correlated with $GPA$ because students were chosen randomly. Moreover this variable is correlated with PC ownership, because once the person received a grant for purchasing PC, he should own a PC. Consequently, this variable is relevant, i.e. correlated with PC ownership, and exogenous, i.e. not correlated with $GPA$.
\end{enumerate}
\section*{Problem 2}
Consider a simple linear regression model
\begin{align*}
y = \beta_0 + \beta_1x + u
\end{align*}
The regressor $x$ is endogenous and $Cov(x, u) \neq 0$. There are two instruments $z_1$ and $z_2$, which
are valid and relevant and non-constant. The error term $u$ is homoscedastic and has zero
unconditional mean. The random variables $x$, $z_1$ and $z_2$ also have zero unconditional mean.
\begin{enumerate}[(a)]
\item Suppose that $Cov(x, u) > 0$. What is the sign of an asymptotic bias of the OLS estimate
of the intercept?
\item Consider the vector of random variables $\xi = (x, z_1, z_2)^T$. What is a minimal possible
rank of the following matrix of expectations: $\expect (\xi\xi^T)$?
\item Consider two IV estimates of the intercept. The first estimate is based on the following
two instruments: constant and $z_1$. The second is based on $z_1$ and $z_2$. Is it possible to get
a smaller asymptotic variance for the intercept estimator in the second case than in the first
case? If yes, provide an example.
\item Suppose someone wants to perform IV estimation with the following instruments: constant and $z_1$. The estimator does not observe $z_1$, instead (s)he observes this variable with a
measurement error which is i.i.d. and independent of all other random variables. Investigate
and describe the effect of the measurement error on efficiency of estimation.

\end{enumerate}

\textbf{Solution}

\begin{enumerate}[(a)]
	\item If $Cov(x, u) > 0$ then
	\begin{align*}
	\frac{\summa (x_i - \bar{x})y_i}{\summa (x_i - \bar{x})^2} = \hat{\beta}_1 \overset{p}{\to} \frac{cov(x, y)}{Var(y)} = \beta_1 + \frac{cov(x, u)}{Var(x)}
	\end{align*}
	\begin{align*}
	\hat{\beta}_0 = \bar{y} - \hat{\beta}_1 \bar{x} = \beta_0 + (\beta_1 - \hat{\beta_1})\bar{x} + \bar{u} \overset{p}{\to} \beta_0 + \frac{cov(x, u)}{Var(x)}\expect x + \expect u = \beta_0
	\end{align*}
	hence the intercept is still asymptotically unbiased.
	\item $\expect \xi \xi^T$ is in fact a covariance matrix of the random vector $\xi = (x\ z_1\ z_2)^T$, i.e. $Var(\xi)$. Assume firstly that $z_1$ and $z_2$ are perfectly linearly dependent i.e. $z_1 = -\alpha z_2, \alpha \neq 0$. That means that
	\begin{align*}
	\exists\ v = (0\ 1\ -\alpha)^T \neq 0: Var(v^T\xi) = v^T Var(\xi) v = Var(z_1 -\alpha z_2) = 0
	\end{align*}
	that means that matrix $Var(\xi) = \expect (\xi \xi^T)$ is singular, i.e. $rg(\expect(\xi\xi^T)) < 3$. Note that assumption of perfect linear dependency between $z_1$ and $z_2$ does not violate the premise of the problem, both $z_1$ and $z_2$ are still valid, relevant and non-constant. In conclusion, rank can be equal to 2.

	Now assume that $x = \alpha_1 z_1 + \alpha_2 z_2,\  \alpha_1\alpha_2 \neq 0$ then
	\begin{align*}
	\expect x^2 = \expect[\alpha_1^2z_1^2 + 2\alpha_1\alpha_2z_1z_2 + \alpha_2^2z_2^2] = \alpha_1\expect[\alpha_1z_1^2 + \alpha_2z_1z_2] + \alpha_2\expect[\alpha_1z_1z_1 + \alpha_2z_2^2] =\\ =\alpha_1\expect[xz_1] + \alpha_2 \expect[xz_2]
	\end{align*}
	As a result:
	\begin{align*}
	Varx = \alpha_1 cov(x, z_1) + \alpha_2 cov(x, z_2)\\
	cov(x, z_1) = \alpha_1 Varz_1 + \alpha_2 cov(z_1, z_2)\\
	cov(x, z_2) = \alpha_1 cov(z_1, z_2) + \alpha_2Varz_2
	\end{align*}
	Since
	\begin{align*}
	\expect \xi\xi^T = \begin{pmatrix}
	Varx & cov(x, z_1) & cov(x, z_2)\\
	cov(x, z_1) & Var z_1 & cov(z_1, z_2)\\
	cov(x, z_2) & cov(z_1, z_2) & Varz_2
	\end{pmatrix}
	\end{align*}
	we can conclude that the first row is a linear combination of two other rows, i.e. matrix has rank 1.

	However, in this case $z_1$ and $z_2$ become non-valid instruments, because
	\begin{align*}
	cov(u, z_1) = \frac{1}{\alpha_1}cov(x, u) - \frac{\alpha_2}{\alpha_1}cov(z_2, u) = \frac{1}{\alpha_1}cov(x, u) \neq 0\\
	cov(u, z_2) = \frac{1}{\alpha_2}cov(x, u) - \frac{\alpha_1}{\alpha_2}cov(z_1, u)= \frac{1}{\alpha_2}cov(x, u) \neq 0\\
	\end{align*}
	In conclusion, minimal rank is 2.
	\item Assume $x_i = (1\ x_i)^T, X = (x_1\ \dots\ x_n)^T, z_{1i} = (1\ z_{1i})^T, Z_1 = (z_{11}\ \dots\ z_{1n})^T, \beta = (\beta_0\ \beta_1)^T$.
	Then 
	\begin{align*}
	y &= X\beta + u\\
	X &= Z_1\gamma + v\\
	y &= Z_1\gamma\beta + u\\
	\hat{\beta} &= ((Z_1\hat{\gamma})^T(Z_1\hat{\gamma}))^{-1}(Z_1\hat{\gamma})^Ty\\
	\hat{\gamma} &= (Z^TZ_1)^{-1}Z_1^TX\\
	\hat{\beta}&=(X^TZ_1(Z_1^TZ_1)^{-1}Z_1^TX)^{-1}X^TZ_1(Z_1^TZ_1)^{-1}Z_1^Ty = (Z_1^TX)^{-1}Z_1^Ty\\
	Var(\hat{\beta}) &= (Z_1^TX)^{-1}Z_1^TZ_1(X^TZ_1)^{-1}\sigma^2\\
	\end{align*}
	\begin{align*}
	\frac{1}{n}(Z_1^TX) = \begin{pmatrix}
	1 & \bar{x}\\
	\bar{z_1} & \frac{1}{n}\summa z_{1i}x_i
	\end{pmatrix} &\overset{p}{\to} \begin{pmatrix}
	1 & 0\\
	0 & cov(z_1, x)
	\end{pmatrix}\\
	\frac{1}{n}(Z^TZ) = \begin{pmatrix}
	1 & \bar{z}\\
	\bar{z} & \frac{1}{n}\summa z_{1i}^2
	\end{pmatrix} &\overset{p}{\to} \begin{pmatrix}
	1 & 0\\
	0 & Var(z_1)
	\end{pmatrix}\\
	\frac{1}{n} (X^TZ_1) = \begin{pmatrix}
	1 & \bar{z_1}\\
	\bar{x} & \frac{1}{n}\summa z_{1i}x_i
	\end{pmatrix} & \overset{p}{\to} \begin{pmatrix}
	1 & 0\\
	0 & cov(z_1, x)
	\end{pmatrix}
	\end{align*}
	That means that $n Var_{as}(\hat{\beta_0}) = \sigma^2$

	Now, suppose $z_{2i} = (z_{1i}\ z_{2i})^T$ and $Z_2 = (z_1\ \dots\ z_n)^T$. Following the same logic:
	\begin{align*}
	Var(\hat{\beta}) = (Z_2^TX)^{-1}Z_2^TZ_2(X^TZ_2)^{-1}\sigma^2\\
	\end{align*}
	\begin{align*}
	\frac{1}{n}(Z_2^TX) = \begin{pmatrix}
	\bar{z_1} & \frac{1}{n}\summa z_{1i}x_i\\
	\\
	\bar{z_2} & \frac{1}{n}\summa z_{2i}x_i
	\end{pmatrix} &\overset{p}{\to} \begin{pmatrix}
	0 & cov(z_1, x)\\
	0 & cov(z_2, x)
	\end{pmatrix}\\
	\frac{1}{n}(Z^TZ) = \begin{pmatrix}
	\frac{1}{n}\summa z_{1i}^2 & \frac{1}{n}\summa z_{1i}z_{2i}\\
	\\
	\frac{1}{n}\summa z_{2i}x_i & \frac{1}{n}\summa z_{2i}^2
	\end{pmatrix} &\overset{p}{\to} \begin{pmatrix}
	Var(z_1) & cov(z_1, z_2)\\
	cov(z_1, z_2) & Var(z_2)
	\end{pmatrix}\\
	\frac{1}{n} (X^TZ_1) = \begin{pmatrix}
	\bar{z_1} & \bar{z_2}\\
	\frac{1}{n}\summa x_iz_{1i} & \frac{1}{n}\summa z_{2i}x_i
	\end{pmatrix} & \overset{p}{\to} \begin{pmatrix}
	0 & 0\\
	cov(x, z_1) & cov(z_2, x)
	\end{pmatrix}
	\end{align*}
	As we can see, asymptotic matrix is singular, i.e. it diverges to infinity whereas in previous case asymptotic variance $\times n$ is finite. We have to conclude that, it is not possible to get
	a smaller asymptotic variance for the intercept estimator in the second case than in the first
	case.
	\item Assume $\tilde{z_{1i}} = z_{1i} + \varepsilon_i$. Then
	\begin{align*}
	\hat{\beta_1}_{IV} = \frac{\summa (y_i - \bar{y})(z_{1i} + \varepsilon_i)}{\summa (x_i - \bar{x})(z_{1i} + \varepsilon_i)}
	\end{align*}
\end{enumerate}
\end{document}

### 2019-02-14 macro3_3.tex
Macro

@@
 			&s.t.\ \dot{b}(t) + p(t)\dot{a}(t) = w(t)l(t) + r(t)b(t) + d(t)a(t) - c(t),\ \forall\ t\\
-			&\lim_{t \to \infty} b(t)e^{-\int_{0}^t r(\tau)d\tau} \ge 0, c(t) \ge 0, 0 \le l(t) \le 1, b(0) = 0, a(t) \le 1\\
+			&\lim_{t \to \infty} b(t)e^{-\int_{0}^t r(\tau)d\tau} \ge 0, c(t) \ge 0, 0 \le l(t) \le 1, b(0) = 0, 0 \le a(t) \le 1\\
 			\end{align*}
@@
 	&s.t.\ \dot{b}(t) + \dot{x}(t) = w(t)l(t) + r(t)b(t) + \frac{d(t) + \dot{p}(t)}{p(t)}x(t) - c(t),\ \forall\ t\\
-	&\lim_{t \to \infty} b(t)e^{-\int_{0}^t r(\tau)d\tau} \ge 0, c(t) \ge 0, 0 \le l(t) \le 1, b(0) = 0, \frac{x(t)}{p(t)} \le 1\\
+	&\lim_{t \to \infty} b(t)e^{-\int_{0}^t r(\tau)d\tau} \ge 0, c(t) \ge 0, 0 \le l(t) \le 1, b(0) = 0,0 \le \frac{x(t)}{p(t)} \le 1\\
 	\end{align*}
@@
 	\item Economic interpretation is the following: the price of asset is equal to discounted sum of all dividends it pays.
-	\item 
+	\item Using capital motion equation, market clearing can be rewritten as follows:
+	\begin{align*}
+	c(t) + \dot{k}(t) + \delta k(t) + d(t) = F'_k(k(t), l(t))k(t) + F'_l(k(t), l(t))l(t)\\
+	c(t) + \dot{k}(t) + \delta k(t) + d(t) = F'_k(k(t), l(t))k(t) + w(t)l(t)\\
+	\end{align*}
+	Since at the equilibrium firms have zero profit, market clearing condition implies that 
+	\begin{align*}
+	c(t) = w(t)l(t)
+	\end{align*}
+	that means that
+	\begin{align*}
+	\dot{k}(t) + \delta k(t) + d(t) &= (r(t) + \delta)k(t)\\
+	r(t) &= \frac{\dot{k}(t) + d(t)}{k(t)}
+	\end{align*}
+	$k(t) = p(t)$ i.e. price of asset is equal to price of capital, it is a reasonable result. At equilibrium people should be indifferent between purchasing capital and purchasing assets.
 	\item \begin{align*}
@@
 	\end{align*}
-	\item
+	 Since $a(t) = 1$ then $x(t) = p(t) = k(t)$. Thus,
+	\begin{align*}
+	\dot{b}(t) + \dot{k}(t) = w(t)l(t) + r(t)b(t) + r(t)k(t) - c(t)
+	\end{align*}
+	So, we obtained the budget constraint in a model when consumers own capital.
+	\item Again, consumers save through firms: market clearing implies that the whole output should be either consumed by consumers or invested in future capital (or payed consumers in a form of dividends).
+As a result, consumers are able to reduce consumption and thereby increase investments in future capital which increases their future wages and dividends.
+$r(t)$ is also set such that consumers and firms become indifferent between different forms of saving (investment).
 	\end{enumerate}

### 2019-02-15 matrics5.tex
metrics

@@
 \usepackage{float}
+\usepackage{makecell}
 \usepackage[section]{placeins}
@@
 	\begin{align*}
-	\hat{\beta_1}_{IV} = \frac{\summa (y_i - \bar{y})(z_{1i} + \varepsilon_i)}{\summa (x_i - \bar{x})(z_{1i} + \varepsilon_i)}
+	\hat{\beta_1}_{IV} &= \frac{\summa (y_i - \bar{y})(z_{1i} + \varepsilon_i)}{\summa (x_i - \bar{x})(z_{1i} + \varepsilon_i)}\\
+	Var(\hat{\beta_1}_{IV}|x, \tilde{z_1}) &= \frac{s^2_{z_1z_1} + 2s^2_{z_1\varepsilon} + s^2_{\varepsilon \varepsilon}}{(s^2_{xz_1} + s^2_{x \varepsilon})^2}\sigma^2 \overset{p}{\to} \frac{Var(z_1) + Var(\varepsilon)}{(cov(x, z_1))^2}\sigma^2
 	\end{align*}
+	As we can observe, the error term affects asymptotic effectiveness of estimator, variance becomes larger.
 \end{enumerate}
+\section*{Problem 3}
+\begin{enumerate}[(a)]
+	\item If we suppose that wealth level remains unchanged in both cases (the expenditure on purchase is insufficiently small comparing to the whole wealth) then correlation should be positive, because if $p_{i0}$ is larger that means that on average $w_i$ is large, which means that $p_{1i}$ is also on average large.
+	\item Of course $cov(p_{i0}, \varepsilon_{i}^1) = 0$ i.e.
+the instrumental variable is exogenous.
+It remains to show that it is relevant.
+Obviously, by the set-up of the problem $cov(p_{i0}, w_i) > 0$.
+Moreover, as it has been stated in (a):
+	\begin{align*}
+	cov(p_{i0}, \beta_2w_i) > 0\\
+	cov(p_{i0}, \beta_2w_i + \varepsilon_i^1) > 0\\
+	cov(p_{i0}, f(\beta_2w_i + \varepsilon_i^1)) > 0\\
+	cov(p_{i0}, p_{i1}) > 0
+	\end{align*}
+	Thus, the IV is exogenous and relevant i.e. valid. Assume $p_i = A \cdot (\beta_2 w_i + \varepsilon_i) + B$ then
+	\begin{align*}
+	\frac{1}{n} (Z^TX) \overset{p}{\to} \begin{pmatrix}
+	1 & A\beta_2 \expect [w] + B & \expect[w]\\
+	A\beta_2\expect[w] + B & A^2\beta_2^2\expect[w^2] + 2AB\beta_2\expect[w]& A\beta_2 \expect[w^2]+B\expect[w]\\
+	\expect[w]& A\beta_2 \expect[w^2]+B\expect[w] & \expect[w^2]
+	\end{pmatrix}
+	\end{align*}
+	it is easy to see that determinant of this matrix is equal to
+	\begin{align*}
+	B^2 (\expect[w])^2
+	\end{align*}
+	As a result, if $B = 0$ then the aforementioned matrix is singular that means that asymptotic variance of estimator diverges to infinity (not well-behaved)
+	\item To find OLS estimates we need to solve the following linear system:
+	\begin{align*}
+	(X^TX) \hat{\beta} = X^Ty\\
+	\end{align*}
+	Since the inverse operation is interchangeable with convergence in probability (continuous mapping theorem) hence, the limit in probability of the solution to the system must coincide with the solution of system, which is obtained in limit.
+That is,
+	\begin{align*}
+	\text{If } \beta_n \overset{p}{\to} \beta\ \&\ X_n \overset{p}{\to} X, y_n \overset{p}{\to} y\ \text{ then }\\
+	\text{If } \forall\ n\ X_n^TX_n\beta_n = X^T_ny_n\text{ then } X^TX \beta = X^Ty
+	\end{align*}
+	\begin{align*}
+	\frac{1}{n} X^T_nX_n \gamma_n &= \frac{1}{n}X_n^Ty_n\\
+	\begin{pmatrix}
+	1 & \expect[p_0] & \expect[w]\\
+	\expect[p_0] & \expect[p_0^2] & \expect[p_0w]\\
+	\expect[w] & \expect[p_0w] & \expect[w^2]
+	\end{pmatrix} \begin{pmatrix}
+	\gamma_0\\
+	\gamma_1\\
+	\gamma_2
+	\end{pmatrix} &= \begin{pmatrix}
+	\expect[q_1]\\\expect[p_0q_1]\\
+	\expect[wq_1]
+	\end{pmatrix}\\
+	\end{align*}
+	\begin{align*}
+	\begin{pmatrix}
+	1 & \expect[p_0] & \expect[w] \bigg| \expect[q_1]\\
+	\expect[p_0] & \expect[p_0^2] & \expect[p_0w]\bigg| \expect[p_0q_1]\\
+	\expect[w] & \expect[p_0w] & \expect[w^2]\ \bigg|\expect[wq_1]
+	\end{pmatrix} &\sim \begin{pmatrix}
+	1 & \expect[p_0] & \expect[w] \bigg| \expect[q_1]\\
+	\expect[p_0] & \expect[p_0^2] & \expect[p_0w]\bigg| \expect[p_0q_1]\\
+	0 & cov(p_0,w) & Var(w)\ \ \  \bigg|cov(w,q_1)
+	\end{pmatrix} \sim \\
+	&\sim \begin{pmatrix}
+	1 & \expect[p_0] & \expect[w] \bigg| \expect[q_1]\\
+	0 & Var(p_0) & cov(p_0, w)\bigg| cov(p_0q_1)\\
+	0 & cov(p_0, w) & Var(w)\ \ \ \bigg|cov(w, q_1)
+	\end{pmatrix} 
+	\end{align*}
+	\begin{align*}
+	&\gamma_2 = \frac{cov(w, q_1)Var(p_0) - cov(p_0, w)cov(p_0, q)}{Var(w)Var(p_0) - (cov(p_0, w))^2}\\
+	&\gamma_1 = \frac{cov(p_0, q_1)Var(w) - cov(p_0, w)cov(w, q)}{Var(w)Var(p_0) - (cov(p_0, w))^2}\\
+	&\gamma_0 = \expect[q_1] - \expect[p_0] \beta_1 - \expect[w]\beta_2
+	\end{align*}
+	\item 
+	\begin{align*}
+	&q_i^1 - q_i^0 = \beta_1(p_i^1 - p_i^0) + \varepsilon_i^1 - \varepsilon_i^0\\
+	&\hat{\gamma_1} = \frac{\summa (q^1_i - \bar{q^1} - q^0_i + \bar{q^0})(p^1_i - \bar{p^1} - p^0_i + \bar{p^0})}{\summa (p^1_i - \bar{p^1} - p^0_i + \bar{p^0})^2} = \\
+	&= \frac{\summa (\beta_1(p_i^1 - p_i^0) + \varepsilon_i^1 - \varepsilon^0_i - \beta_1(\bar{p^1} - \bar{p^0}) - \bar{\varepsilon^1} + \bar{\varepsilon^0})(p^1_i - \bar{p^1} - p^0_i + \bar{p^0})}{\summa (p^1_i - \bar{p^1} - p^0_i + \bar{p^0})^2}
+	\end{align*}
+	\begin{align}\label{eq1}
+	\expect[\hat{\gamma_1}] = \expect[\expect[\hat{\gamma_1}|p^1, p^0]] = \beta_1 +  \expect\left[\frac{\summa (p^1_i - \bar{p^1} - p^0_i + \bar{p^0})(f^{-1}(p^1_i) - f^{-1}(p^0_i))}{\summa (p^1_i - \bar{p^1} - p^0_i + \bar{p^0})^2}\right]
+	\end{align}
+	where $f(\cdot)$ is such a function that:
+	\begin{align*}
+	p^0_i = f(\beta_2 w_i + \varepsilon^0_i)\\
+	p^1_i = f(\beta_2 w_i + \varepsilon^1_i)\\
+	\end{align*}
+	There are no reasons for the second term in \eqref{eq1} to have a zero mean, that means that $\hat{\gamma_1}$ is biased.
+	\item 
+	\begin{align}\label{eq2}
+	\frac{1}{n}(Z^TX) \overset{p}{\to} \begin{pmatrix}
+	1 & \expect[p^1] & \expect[w]\\
+	\expect[p^0] & \expect[p^1p^0] & \expect[p^0w]\\
+	\expect[w] & \expect[p^1w] & \expect[w^2]
+	\end{pmatrix} \sim \begin{pmatrix}
+	1 & \expect[p^1] & \expect[w]\\
+	0 & cov(p^1, p^0) & cov(p^0, w)\\
+	0 & cov(p^1, w) & Var(w)
+	\end{pmatrix}\\
+	\det \begin{pmatrix}
+	1 & \expect[p^1] & \expect[w]\\
+	0 & cov(p^1, p^0) & cov(p^0, w)\\
+	0 & cov(p^1, w) & Var(w)
+	\end{pmatrix} = cov(p^1, p^0)Var(w) - cov(p^1, w)cov(p^0, w)\nonumber
+	\end{align}
+	By the Cauchy - Schwartz inequality
+	\begin{align*}
+	|cov(p^1, w)| \le \sqrt{Var(p^1)Var(w)}
+	\end{align*}
+	If $p^1 = [\beta_2 w + \varepsilon^1]^3, p^0 = [\beta_2 w + \varepsilon^0]^3$ then $cov(p_1, w) \ge 0$ hence:
+	\begin{align*}
+	cov(p^1, p^0)Var(w) - cov(p^1, w)cov(p^0, w) \ge cov(p^1, p^0)Var(w) - \\
+	-\sqrt{Var(p^1)Var(p^0)}Var(w) > Var(w) > 0
+	\end{align*}
+	Since $cor(p^1, p^0) < 1$ because $p^1, p^0$ are not linearly dependent. That means that determinant of the matrix \eqref{eq2} is not zero, i.e. the estimator is well behaved asymptotically.	
+\end{enumerate}
+\section*{Problem 4}
+Estimate the own-price elasticity of demand by using OLS to regress the log of the quantity
+of grain shipped on the log of the price of shipping grain and the full set of month binary
+indicators.
+\begin{enumerate}[(a)]
+\item What is the estimated demand elasticity and its standard error?
+\item Explain why the interaction of supply and demand plausibly makes this estimator of
+the elasticity biased.
+\end{enumerate}
+
+\textbf{Solution}
+
+\begin{enumerate}[(a)]
+	\item The output of coeftest() is depicted below
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.8\textwidth]{price_elasticity}
+		\caption{}\label{fig1}
+	\end{figure}
+	Elasticity is equal to -0.25 and its standard error is equal to 0.029.
+	\item The estimated model is
+	\begin{align*}
+	\ln (quantity) = \beta_0 + \beta_1 \ln (price) + \beta_2 seas1 + \cdot \beta_{13} seas12 + u
+	\end{align*}
+	Where $u$ is the error term, which includes all factors which affect quantity sold apart from price and dummy months variables.
+This term can include for example, prices of substitutes (the cheaper substitutes are the smaller quantity) or total income of people (the poorer people are the smaller quantity is).
+However, price is obviously correlated with aforementioned factors: the less substitutes cost the smaller price is, that means that assumption of OLS is violated, which leads to biased estimate of elasticity.
+\end{enumerate}
+\section*{Problem 5}
+Consider two possibilities for instrumental variables: whether the railroad cartel was experiencing one of its periodic collapses ($cartel$), and whether the Great Lakes are closed because
+of ice ($ice$).
+\begin{enumerate}[(a)]
+\item Use economic reasoning to argue whether $cartel$ plausibly satisfies the two conditions
+for a valid instrument.
+\item Use economic reasoning to argue whether $ice$ plausibly satisfies the two conditions for
+a valid instrument
+\end{enumerate}
+
+\textbf{Solution}
+
+\begin{enumerate}[(a)]
+	\item Without doubt $cartel$ is correlated with $quantity$ (sample covariance is about 0.6).
+As it has been described cartels try to raise prices.
+Moreover, the existence of cartel can hardly affect the demand on grain, neither preferences nor income or prices of substitutes can be affected by existence of cartel, hence it is plausibly a valid instrument.
+	\item As in previous case, whether or not lakes are covered with ice cannot affect the consumers' demand for grain, however sample correlation between $ice$ and $\ln (price)$ is rather small (0.3) i.e.
+$ice$ can be a weak instrument.
+\end{enumerate}
+\section*{Problem 6}
+\begin{center}
+\begin{tabular}{cccccc}
+	\hline
+	& (1) & (2) & (3) & (4) & (5)\\
+	\hline\\
+	Dependent variable: & $\ln(P)$ & $\ln(Q)$ & $\ln(Q)$ & $\ln(Q)$ & $\ln(Q)$\\
+	Regressors:& & & &\\
+	cartel & $\underset{(13.6534)}{0.356}$ & – & – & – & –\\
+	ln(P) & – & $\underset{(-8.7779)}{-0.662}$ & $\underset{(-6.9133)}{-0.93}$ & $\underset{(-1.7509)}{-4.74}$ &$\underset{(-6.7985)}{-0.921}$\\
+	\makecell{F-statistic testing\\
+	coefficients on monthly\\
+	indicators (p-value)} &5.298e-08 & 1.725e-14&2.2e-16 & 0.6864&2.2e-16\\
+	Estimation method & OLS & OLS &TSLS &TSLS &TSLS\\
+	Instrumental variables& n/a& n/a& cartel& ice& \makecell{ice\\
+	cartel}\\
+	First-stage F-statistic& n/a &n/a&186.42&2.3907&92.847\\
+	J-test of overidentifying &n/a &n/a&&&\\
+	restrictions & n/a &n/a &n/a& n/a
+&
+\end{tabular}
+\end{center}
+\section*{Problem 7}
+I think the most reasonable results were gotten in TSLS regression with $cartel$ as IV.
+As we can conclude from F-statistic, $ice$ is a weak instrument which is barely correlated with prices and I think only add noise to the model.
+On the other hand, $cartel$ is correlated with prices, and give reasonable significant statistics to conclude that there was indeed cartel pricing (elasticity is very close to -1).
+
+Here, the main threat is a weak correlation between instruments and regressors
 \end{document}

### 2019-02-16 gameth3.tex
gameth

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage{makecell}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\newcommand{\yrseduc}{\textit{yrseduc}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Game theory. Problem Set 3.
\end{center}
\section*{Problem 1}
The tree of the game is depicted below:
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{plotdraft}
	\caption{}\label{fig1}
\end{figure}
Obviously, profile of strategies
\begin{align*}
\left\{L, R\right\}
\end{align*}
is a Nash equilibrium: the best response on $R$ is $L$, the best response on $L$ is $R$. Moreover, if we introduce belief $\mu$ as
\begin{align*}
\prob (\text{2nd player at node} \left\{L\right\}|\text{2nd player is called upon to play}) = \mu
\end{align*}
then 
\begin{align*}
-2 \mu < 5\mu + 7 - 7 \mu = 7 - 2\mu\ \forall\ \mu \in (0, 1)
\end{align*}
that means that optimal strategy for player 2 is playing $R$. And aware of it, optimal strategy for player 1 is playing L. As a result the outcome $\left\{LR\right\}$ is both weak sequential equilibrium and Nash equilibrium.
\section*{Problem 2}
Obviously the weak sequential equilibrium (which is also a SPE) is the following profile of strategies:
\begin{align*}
\left\{In\ Acquiesce;\ A\right\}
\end{align*}
and belief for player 2:
\begin{align*}
(0, 1)
\end{align*}
i.e. player 2 knows for sure that he is at node $\left\{In\ Acquiesce\right\}$. This conclusion is derived because in unique proper subgame the strategy $\left\{Fight\right\}$ is strictly dominated by strategy $\left\{Acquisce\right\}$.

Suppose player 1 plays $\left\{Out, Fight\right\}$ for sure. Then the information set of the 2nd player is reached with 0 probability. Hence the concept of weak sequential equilibrium allows us to put arbitrary beliefs but to the extent that given those beliefs player 1 still does not have a profitable deviation. Assume $\mu_{IF} = \mu$ then expected payoffs of player 2 are:
\begin{align*}
F : u_2 &= -1\\
A: u_2 &=-2\mu + 1 - \mu = 1 - 3\mu
\end{align*}
If $\mu < \frac{2}{3}$ then player 2 plays $A$, if $\mu > \frac{2}{3}$ then he plays $F$. If $\mu = \frac{2}{3}$ then player 2 mixes. On the other hand, suppose that player 1 deviates and plays $\left\{In, Fight\right\}$ then his expected payoffs is equal to:
\begin{align*}
\begin{cases}
3, \mu < \frac{2}{3}\\
-2, \mu > \frac{2}{3}\\
3 - 5\alpha, \mu = \frac{2}{3}
\end{cases}
\end{align*}
where $\alpha$ is a probability which player 2 mixes the strategy $F$ with. If $\mu < \frac{2}{3}$ then it is profitable for player 1 to deviate, hence it is not an equilibrium, if $\mu > \frac{2}{3}$ then player 1 does not have a profitable deviation. If $\mu = \frac{2}{3}$ then if $\alpha < \frac{3}{5}$ then again, player 1 has a profitable deviation, if $\alpha > \frac{3}{5}$ then the first player does not have incentives to deviate. If $\alpha = \frac{3}{5}$ then player is indifferent, however mixing for player 1 cannot be consistent with beliefs because once he assigns positive probability for $\left\{In\right\}$, $\mu$ automatically becomes zero, because strategy $\left\{In, Fight\right\}$ is strictly dominated by $\left\{In, Acquisce\right\}$. As a result the profiles of strategies and beliefs for player 2:
\begin{align*}
\begin{cases}
\left\{Out, Fight;\ F \right\}, \prob(In, Fight|In) = \mu > \frac{2}{3},\ \prob(In, Acquisce|In) = 1 - \mu\\
\left\{Out, Fight;\ \alpha F + (1-\alpha)A \right\}, \prob(In, Fight|In) = \frac{2}{3},\ \prob(In, Acquisce|In) = \frac{1}{3}, \alpha > \frac{3}{5}\\
\end{cases}
\end{align*}
constitute sequential equilibrium. These profiles are not SPE.
\section*{Problem 3}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{game3}
	\caption{}\label{fig2}
\end{figure}

\textbf{Solution}

Suppose that, conditional on the fact that player 2 is called upon to play, he believes that he is at node $\{AR\}$ with probability $\mu$ and at node $\left\{BR\right\}$ with probability $1 - \mu$. Then playing $W$ yields him $5$ whereas playing $E$ yields 
\begin{align*}
10 - 8 \mu
\end{align*}
hence he should play
\begin{align*}
\begin{cases}
W, \mu > \frac{5}{8}\\
E, \mu < \frac{5}{8}\\
\alpha W + (1 - \alpha)E, \mu = \frac{5}{8}
\end{cases}
\end{align*}
for $\alpha \in (0, 1)$. For player 1, if he plays $R$ he gets:
\begin{align*}
\begin{cases}
0, \mu > \frac{5}{8}\\
5, \mu < \frac{5}{8}\\
5 - 5\alpha, \mu = \frac{5}{8}
\end{cases}
\end{align*}
So, we can conclude that the profile of strategies and belief for player 2:
\begin{align*}
\left\{L, W\right\}, (\mu, 1-\mu), \mu > \frac{5}{8}
\end{align*}
is a weak sequential equilibrium.

If $\mu < \frac{5}{8}$ then player 1 plays for sure $R$ as a result $\mu \neq \frac{1}{2}$ is inconsistent. If $\mu = \frac{5}{8}$ then for $\alpha > \frac{3}{5}$ player 1 plays $L$ and this is also a weak sequential equilibrium, if $\alpha < \frac{3}{5}$ then player 1 should play $R$ for sure, but it is inconsistent with $\mu$. The only remained case is $\alpha = \frac{3}{5}$. In this case, player 1 is indifferent, hence he will mix, however $\mu \neq \frac{1}{2}$ will always be inconsistent with each $\beta$.

If $\mu = \frac{1}{2}$ then player 2 plays $E$ and player 1 plays $R$. This is also a weak sequential equilibrium, since belief is consistent with the strategy and both players maximize their expected payoffs given that the other player plays the strategy.
Thus, weak sequential equilibria of this game are:
\begin{align*}
\begin{cases}
\{L, W\},\ (\mu, 1-\mu),\ \mu > \frac{5}{8}\\
\{L, \alpha W + (1 - \alpha) L\},\ (\frac{5}{8}, \frac{3}{8}),\ \alpha > \frac{3}{5}\\
\{R, E\},\ (\frac{1}{2}, \frac{1}{2})
\end{cases}
\end{align*}
(Belief for player 1 is obviously $(0.5, 0.5)$)

\section*{Problem 4}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\textwidth]{game4}
	\caption{}\label{fig3}
\end{figure}

\textbf{Solution}
\begin{enumerate}[(a)]
	\item Yes, for example $\alpha = \frac{3}{4}$ in this case the profile of strategies $\{D, RL\}$ is a Nash equilibrium. Let us prove it. Obviosly, $\{RL\}$ is a best response on $\{D\}$. For player 1 expected payoffs of playing:
	\begin{align*}
	&U\text{ is } \frac{3}{8}\\
	&D\text{ is } \frac{9}{16} > \frac{3}{8}
	\end{align*}
	i.e. $D$ is also a best response on $\{RL\}$, hence it is a Nash equilibrium.
	\item $\alpha = \frac{1}{2}$. One of sequential equilibria is a mixed behavioural strategy for player 1: $\frac{2}{3}U + \frac{1}{3}D$ and for player 2: $\{\frac{2}{3}L + \frac{1}{3}R, R\}$ and believs for player 2: $(\frac{2}{3}, \frac{1}{3})$ (equal for both information sets, where he is called upon to play). That is, player 2 plays $\frac{2}{3}L + \frac{1}{3}R$ once the nature plays $T$ and $R$ otherwise. Suppose "1st information set" stands for the information set where the 2nd player is called upon to play and nature plays T, and "2nd information set" once the nature plays $B$. Since $\frac{1}{6} < \frac{1}{3}$ then at the 2nd information set the best response of player 2 is indeed $R$ and moreover, belief $(\frac{2}{3}, \frac{1}{3})$ is consistent with the strategy:
	\begin{align*}
	\frac{2}{3} = \frac{\frac{2}{3}(1-\alpha)}{(1 - \alpha)}
	\end{align*}. At the 1st information set, player's 2 expected payoffs from playing:
	\begin{align*}
	L: \frac{1}{3}\\
	R: \frac{1}{3}
	\end{align*}
	hence the best response is any mixed behavioural strategy and $\frac{2}{3}L + \frac{1}{3}R$ in particular. Also note, that at 1st information set belief for player 2 is consistent. It remains to check whether the first player is sequentially rational for this strategy. His expected payoffs of playing:
	\begin{align*}
	U: \frac{1}{3}\\
	D: \frac{1}{3}
	\end{align*}
	that means that he can play any mixed strategy and $\frac{2}{3}U + \frac{1}{3}D$ in particular.

	Thus, the strategy and believes are indeed sequential equilibrium of the game.
\end{enumerate}
\end{document}

### 2019-02-17 Metrics5.R
Metrics

@@ -1,0 +1,66 @@
+df = read.csv("grain.csv")
+
+#Problem 4
+fit <- lm(log(price) ~ log(quantity)+seas1+seas2+seas3+seas4+seas5+seas6+seas7+seas8+seas9+seas10+seas11+seas12, data=df)
+summary(fit)
+library(lmtest)
+coeftest(fit, vcov = vcovHC(fit, type = "HC1"))
+
+#Problem 5
+#a
+cor(df$cartel, log(df$price))
+#b
+cor(df$ice, log(df$price))
+
+#Problem 6
+#1
+fit1 <- lm(log(price) ~ cartel+seas1+seas2+seas3+seas4+seas5+seas6+seas7+seas8+seas9+seas10+seas11+seas12, data=df)
+coeftest(fit1, vcov = vcovHC(fit1, type = "HC1"))
+library(car)
+myH0 <- c("seas1 = 0", "seas2 = 0", "seas3 = 0", "seas4 = 0", "seas5 = 0", "seas6 = 0", "seas7 = 0", "seas8 = 0", "seas9 = 0", "seas10 = 0", "seas11 = 0", "seas12 = 0")
+linearHypothesis(fit1, myH0, vcov = vcovHC(fit1, type = "HC1"))
+
+#2
+fit2 <- lm(log(quantity) ~ log(price)+seas1+seas2+seas3+seas4+seas5+seas6+seas7+seas8+seas9+seas10+seas11+seas12, data=df)
+coeftest(fit2, vcov = vcovHC(fit2, type = "HC1"))
+linearHypothesis(fit2, myH0, vcov = vcovHC(fit2, type = "HC1"))
+
+#3
+library(AER)
+fit3 <- ivreg(log(quantity) ~ log(price)+seas1+seas2+seas3+seas4+seas5+seas6+seas7+seas8+seas9+seas10+seas11+seas12|seas1+seas2+seas3+seas4+seas5+seas6+seas7+seas8+seas9+seas10+seas11+seas12+cartel, data=df)
+coeftest(fit3, vcov = vcovHC(fit3, type = "HC1"))
+linearHypothesis(fit3, myH0, vcov = vcovHC(fit3, type = "HC1"))
+#First stage regression:
+first_stage <- lm(log(price) ~ seas1+seas2+seas3+seas4+seas5+seas6+seas7+seas8+seas9+seas10+seas11+seas12 + cartel, data=df)
[…]

### 2019-02-22 macro3_4.tex
macro

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage{makecell}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\newcommand{\yrseduc}{\textit{yrseduc}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Macroeconomics 3. Problem Set 4.
\end{center}
\section*{Problem 1}
\textbf{Solution}

\begin{enumerate}
	\item A SOMCE is a set of sequences $\{c(t), C(t), w(t), r(t), l(t), a(t), k(t), a(t)\}_{t=0}^{\infty}$. Such that:
	\begin{itemize}
		\item Taking $w(t)$ as given, workers chooses $\{c(t), l(t)\}_{t=0}^{\infty}$ such that it solves the following optimization problem:
		\begin{align*}
		&\underset{c(t), l(t)}{\max}\ \sum_{t=0}^{\infty} \beta^t \ln c(t)\\
		&s.t.\ \  c(t) = w(t)l(t), c(t) \ge 0, 0 \le l(t) \le 1
		\end{align*}
		\item Taking $r(t)$ as given, capitalists choose $\{C(t), a(t+1)\}_{t=0}^{\infty}$ such that it solves the following optimization problem:
		\begin{align*}
		&\underset{C(t)}{\max}\ \sum_{t=0}^{\infty} \beta^t \ln C(t)\\
		&s.t.\ C(t) + a(t+1) = (1 + r(t))a(t), C(t) \ge 0, a(t) \ge 0, a(0) \text{ is given }
		\end{align*}
		\item Taking $w(t), r(t)$ as given, firms choose $\{k(t+1), l(t)\}_{t=0}^{\infty}$ such that it solves the following optimization problem:
		\begin{align*}
		&\underset{k(t), l(t)}{\max}\ \sum_{t=0}^{\infty} \prod_{s=0}^t\left(\frac{1}{1 + r(s)}\right)(F(k(t), l(t)) - w(t)l(t) - r(t)k(t) - k(t+1) + (1 - \delta)k(t))\\
		&s.t.\ k(t) \ge 0, k(0) \text{ is given }
		\end{align*}
		\item Markets are cleared, i.e.:
		\begin{align*}
		&k(t) = a(t),\ \forall\ t\\
		&c(t) + C(t) + k(t+1) - (1 - \delta)k(t) = F(k(t), l(t)),\ \forall\ t\\
		&l(t) = 1,\ \forall\ t
		\end{align*}
	\end{itemize}
\begin{enumerate}[(a)]
	\item 	Optimality conditions for firms imply that:
	\begin{align}\label{eq1}
	\begin{cases}
	-\prod_{s=0}^{t}\left(\frac{1}{1+r(s)}\right) + \prod_{s=0}^{t+1}\left(\frac{1}{1+r(s)}\right)(F'_k(k(t+1), l(t+1)) - r(t+1) + 1 - \delta) = 0\\
	w(t) = F(k(t), 1) - r(t)k(t) - k(t+1) + (1 - \delta)k(t)
	\end{cases}
	\end{align}
	hence 
	\begin{align*}
	F'_k(k(t+1), l(t+1)) - r(t+1) + 1 - \delta = 1 + r(t+1)\\
	F'_k(k(t+1), l(t+1)) = \delta + 2r(t+1)
	\end{align*}
	At steady state $k(t) = \bar{k}$ hence $F'_k(\bar{k}, 1) = const = \delta + 2r(t+1)$ hence $r(t+1) = \bar{r}$. From BC for capitalists and equilibrium condition $a(t) = k(t) = \bar{k}$ one can get:
	\begin{align*}
	C(t) = \bar{r}\bar{k} = \bar{C}
	\end{align*}
	Moreover, from FOC for capitalists:
	\begin{align}
	\frac{C(t+1)}{\beta C(t)} = 1 + r(t+1)
	\end{align}
	hence at steady state $\bar{r} = \frac{1}{\beta}-1$ and steady state capital level is:
	\begin{align*}
	f(k) = F(k, 1)\\
	\bar{k}^* = f'^{-1}\left(\delta + 2\left(\frac{1}{\beta}-1\right)\right) 
	\end{align*}
	In the standard neoclassical growth model
	\begin{align*}
	\bar{k}^* = f'^{-1}\left(\frac{1}{\beta} - 1 + \delta\right)
	\end{align*}
	hence now at steady state there is less capital.
	\item Obviously $c(t) = w(t)\ \forall\ t$. And for capitalists:
	\begin{align}\label{eq2}
	\frac{C(t+1)}{\beta C(t)} = 1 + r(t+1)
	\end{align}
	From \eqref{eq2} and BC for capitalists it follows that:
	\begin{align*}
	&a(t) = \frac{C(t)}{1 + r(t)} + \frac{a(t+1)}{1+r(t)} = \frac{C(t)}{1 + r(t)} + \frac{C(t+1)}{(1+r(t))(1+r(t+1))} + \\ &+\frac{a(t+2)}{(1+r(t))(1+r(t+1))} = \frac{1+\beta}{1+r(t)}C(t) + \frac{C(t+2)}{(1+r(t))(1+r(t+1))(1+r(t+2))} + \\
	&+\frac{a(t+3)}{(1+r(t))(1+r(t+1))(1+r(t+2))} = \frac{1+\beta + \beta^2}{1+r(t)}C(t) +\\ &+\frac{a(t+3)}{(1+r(t))(1+r(t+1))(1+r(t+2))} = ... = \frac{1 - \beta^{T+1}}{(1-\beta)(1+r(t))}C(t) +\\ &+\prod_{s=t}^{T}\left(\frac{1}{1+r(s)}\right) a(t+T+1)
	\end{align*}
	Transversality condition implies that the last term tends to 0 while $T \to \infty$ hence
	\begin{align}\label{eq3}
	C(t) = (1 - \beta)(1+r(t))a(t)
	\end{align}
	\item Since
	\begin{align*}
	\frac{C(t+1)}{\beta C(t)} = 1 + r(t+1)
	\end{align*}
	then if $r(t+1)$ becomes larger than $C(t+1)$ becomes relatively larger to $C(t)$. The intuition is straightforward: consumption at $t$ becomes more expensive (because $r(t+1)$ is alternative cost of consumption) hence it is substituted by $C(t+1)$. 
\end{enumerate}
\item \begin{enumerate}[(a)]
	\item A SOMCE is a set of sequences $\{c(t), C(t), w(t), r(t), l(t), a(t), k(t), g(t), \tau(t), T(t)\}_{t=0}^{\infty}$. Such that:
	\begin{itemize}
		\item Taking $w(t), T(t)$ as given, workers chooses $\{c(t), l(t)\}_{t=0}^{\infty}$ such that it solves the following optimization problem:
		\begin{align*}
		&\underset{c(t), l(t)}{\max}\ \sum_{t=0}^{\infty} \beta^t \ln c(t)\\
		&s.t.\ \  c(t) = w(t)l(t) - T(t), c(t) \ge 0, 0 \le l(t) \le 1
		\end{align*}
		\item Taking $r(t), \tau(t)$ as given, capitalists choose $\{C(t), a(t+1)\}_{t=0}^{\infty}$ such that it solves the following optimization problem:
		\begin{align*}
		&\underset{C(t), a(t)}{\max}\ \sum_{t=0}^{\infty} \beta^t \ln C(t)\\
		&s.t.\ C(t) + a(t+1) = (1 - \tau(t))(1 + r(t))a(t), C(t) \ge 0, a(t) \ge 0, a(0) \text{ is given }
		\end{align*}
		\item Taking $w(t), r(t)$ as given, firms choose $\{k(t+1), l(t)\}_{t=0}^{\infty}$ such that it solves the following optimization problem:
		\begin{align*}
		&\underset{k(t), l(t)}{\max}\ \sum_{t=0}^{\infty} \left(\frac{1}{1 + r(t)}\right)^t(F(k(t), l(t)) - w(t)l(t) - r(t)k(t) - k(t+1) + (1 - \delta)k(t))\\
		&s.t.\ k(t) \ge 0, k(0) \text{ is given }
		\end{align*}
		\item Markets are cleared, i.e.:
		\begin{align*}
		&k(t) = a(t),\ \forall\ t\\
		&c(t) + C(t) + g(t) + k(t+1) - (1 - \delta)k(t) = F(k(t), l(t)),\ \forall\ t\\
		&l(t) = 1,\ \forall\ t\\
		&g(t) = T(t) + \tau(t)(1 + r(t))a(t)
		\end{align*}
	\end{itemize} 
\item \begin{align*}
&\underset{\makecell{c(t), C(t), l(t), \\ r(t), a(t), w(t), \\k(t), \tau(t), T(t)}}{\max}\ \sum_{t=0}^{\infty} \beta^t \ln c(t)\\
&s.t.\ \begin{cases}
\frac{C(t+1)}{\beta C(t)} = (1 - \tau(t+1))(1 + r(t+1))\\
c(t) = w(t) - T(t) = F(k(t), 1) - r(t)k(t) - k(t+1) + (1 - \delta)k(t) - T(t)\\
F'_k(k(t+1), l(t+1)) = \delta +2r(t+1)\\
c(t) + C(t) + g(t) + k(t+1) - (1 - \delta)k(t) = F(k(t), l(t))\\
C(t) + a(t+1) = (1 - \tau(t))(1 + r(t))a(t)\\
g(t) = T(t) + \tau(t) a(t)\\
a(t) = k(t)\\
l(t) = 1\\
C(t) \ge 0, c(t) \ge 0, k(t) \ge 0, a(t) \ge 0, k(0) \text{ is given }\\
TVC
\end{cases}
\end{align*}
\item Since planner cares only about workers and they cannot save, effectively planner should maximize just $w(t) - T(t)$ at steady state. i.e.
\begin{align*}
w(t) - T(t) = f(\bar{k}) - (\bar{r} + \delta)\bar{k} - T(t)
\end{align*}
\begin{align*}
\begin{cases}
\frac{C(t+1)}{\beta C(t)} = (1 - \tau(t+1))(1 + r(t+1))\\
c(t) = w(t) - T(t) = F(k(t), 1) - r(t)k(t) - k(t+1) + (1 - \delta)k(t) - T(t)\\
F'_k(k(t+1), l(t+1)) = \delta +2r(t+1)\\
c(t) + C(t) + T(t) + \tau(t) \bar{k} + \delta\bar{k} = F(k(t), l(t))\\
C(t) + a(t+1) = (1 - \tau(t))(1 + r(t))a(t)\\
g(t) = T(t) + \tau(t) a(t)\\
\end{cases}
\end{align*}
\begin{align*}
\begin{cases}
\frac{C(t+1)}{\beta C(t)} = (1 - \tau(t+1))(1 + r(t+1))\\
c(t) = F(k(t), 1) - (\bar{r} +\delta)\bar{k} - T(t)\\
F'_k(k(t+1), l(t+1)) = \delta +2r(t+1)\\
c(t) + C(t) + T(t) + \tau(t) \bar{k} + \delta\bar{k} = F(k(t), l(t))\\
C(t) + a(t+1) = (1 - \tau(t))(1 + r(t))a(t)\\
\end{cases}
\end{align*}
\begin{align*}
\begin{cases}
\frac{C(t+1)}{\beta C(t)} = (1 - \tau(t+1))(1 + r(t+1))\\
c(t) = F(k(t), 1) - (\bar{r} +\delta)\bar{k} - T(t)\\
F'_k(k(t+1), l(t+1)) = \delta +2r(t+1)\\
c(t) + C(t) + T(t) + \tau(t) \bar{k} + \delta\bar{k} = F(k(t), l(t))\\
C(t) = (\bar{r} - \tau(t)(1 + \bar{r}))\bar{k}\\
\end{cases}
\end{align*}
\begin{align*}
\begin{cases}
\frac{C(t+1)}{\beta C(t)} = (1 - \tau(t+1))(1 + r(t+1))\\
c(t) = F(k(t), 1) - (\bar{r} +\delta)\bar{k} - T(t)\\
F'_k(k(t+1), l(t+1)) = \delta +2r(t+1)\\
c(t) + T(t) + (\delta + (1-\tau(t))\bar{r})\bar{k} = F(k(t), l(t))\\
C(t) = (\bar{r} - \tau(t)(1 + \bar{r}))\bar{k}\\
\end{cases}
\end{align*}
\begin{align*}
\begin{cases}
\frac{C(t+1)}{\beta C(t)} = (1 - \tau(t+1))(1 + r(t+1))\\
c(t) = F(k(t), 1) - (\bar{r} +\delta)\bar{k} - T(t)\\
F'_k(k(t+1), l(t+1)) = \delta +2r(t+1)\\
F(k(t), 1) - (\bar{r} +\delta)\bar{k} - T(t) + T(t) + (\delta + (1-\tau(t))\bar{r})\bar{k} = F(k(t), l(t))\\
\end{cases}
\end{align*}
\end{enumerate}
	\end{enumerate}
\section*{Problem 2}
\begin{enumerate}
	\item Firstly let us solve the model without pension.
	$\forall\ t > 0$ taking $\{w(t), r(t)\}$ as given households choose $\{c_1(t), c_2(t+1), s(t)\}$ such that is solves the following optimization problem:
	\begin{align*}
	&\max\ \ln c_1(t) + \beta \ln c_2(t+1)\\
	s.t.\ &c_1(t) +s(t) = w(t)\\
	& c_2(t+1) = s(t)(r(t+1) + 1 - \delta)\\
	&c_1(t) \ge 0, c_2(t+1) \ge 0
	\end{align*}
	$c_2(0)$ solves:
	\begin{align*}
	&\max\ \ln c_2(0)\\
	s.t.\ &c_2(0) = s(-1)(r(0) + 1 - \delta)\\
	&s(-1) \text{ is given }
	\end{align*}
	Firms' optimization conditions imply that:
	\begin{align*}
	r(t) &= A\alpha k(t)^{\alpha - 1}\\
	w(t) &= A(1 - \alpha) k(t)^{\alpha}
	\end{align*}
	And market clearing conditions are:
	\begin{align*}
	k(t+1)(1+n) &= Ak(t)^{\alpha} + (1-\delta)k(t) - c_1(t) - \frac{c_2(t)}{n+1}\\
	k(t+1) &= \frac{s(t)}{1+n}
	\end{align*}
	FOC:
	\begin{align*}
	\frac{c_2(t+1)}{\beta c_1(t)} &= A\alpha k(t+1)^{\alpha-1} + 1 - \delta\\
	\frac{k(t+1)(n+1)(A\alpha k(t+1)^{\alpha - 1} + 1 -\delta)}{A(1 - \alpha) k(t)^{\alpha} - k(t+1)(n+1)} &= \beta(A\alpha k(t+1)^{\alpha-1} + 1 - \delta)\\
	k(t+1) &= \frac{A\beta (1-\alpha)k(t)^{\alpha}}{(1+\beta)(1+n)}\\
	\bar{k}^* &= \left(\frac{A\beta (1-\alpha)}{(1+\beta)(1+n)}\right)^{\frac{1}{1 - \alpha}}
	\end{align*}
	Now let us introduce a pension. Then problems of agents become:

	$\forall\ t > 0$ taking $\{w(t), r(t)\}$ as given households choose $\{c_1(t), c_2(t+1), s(t)\}$ such that is solves the following optimization problem:
	\begin{align*}
	&\max\ \ln c_1(t) + \beta \ln c_2(t+1)\\
	s.t.\ &c_1(t) +s(t) = w(t) - T\\
	& c_2(t+1) = s(t)(r(t+1) + 1 - \delta) + (1+n)T\\
	&c_1(t) \ge 0, c_2(t+1) \ge 0
	\end{align*}
	$c_2(0)$ solves:
	\begin{align*}
	&\max\ \ln c_2(0)\\
	s.t.\ &c_2(0) = s(-1)(r(0) + 1 - \delta) + (1+n)T\\
	&s(-1) \text{ is given }
	\end{align*}
	Firms' optimization conditions imply that:
	\begin{align*}
	r(t) &= A\alpha k(t)^{\alpha - 1}\\
	w(t) &= A(1 - \alpha) k(t)^{\alpha}
	\end{align*}
	And market clearing conditions are:
	\begin{align*}
	k(t+1)(1+n) &= Ak(t)^{\alpha} + (1-\delta)k(t) - c_1(t) - \frac{c_2(t)}{n+1}\\
	k(t+1) &= \frac{s(t)}{1+n}
	\end{align*}
	Plugging expressions for $c_1(t)$ and $c_2(t+1)$ into  the objective function and taking derivatives with respect to $s(t)$ we get:
	\begin{align}\label{eq4}
	\frac{1}{w(t) - s(t) - T} = \frac{\beta(1+r(t+1) - \delta)}{s(t)(1 + r(t+1) - \delta) + T(1+n)}\nonumber\\
	s(t) = \frac{\beta (1 + r(t+1) - \delta)(w(t) - T) - T(1+n)}{(1+\beta)(1 + r(t+1) - \delta)} = (n+1)k(t+1)
	\end{align}
	\begin{enumerate}[(a)]
	\item If $T > 0$ then savings are falling down, that means that steady state level of capital declines. It is impossible to express $k(t+1)$ as a function of $k(t)$ since $r(t+1)$ in \eqref{eq4} is a function of $k(t+1)$, but for a given interest rate after introduction $T>0$, $k(t+1)$ becomes smaller, i.e. the equilibrium trajectory also declines.
	\item If the initial state was dynamically (Pareto) efficient then the increase of $T$ benefits living at the moment old generation, but if the equilibrium was Pareto efficient is should hurt future generations. If the initial state was inefficient that means that  people overaccumulate capital, the increase of $T$ invokes the decrease of $s(t)$ as a result the steady state level of capital declines and both living and future generations will be benefited.
	\end{enumerate}
	\item Now
	$\forall\ t > 0$ taking $\{w(t), r(t)\}$ as given households choose $\{c_1(t), c_2(t+1), s(t)\}$ such that is solves the following optimization problem (assuming $\delta = 0$):
	\begin{align*}
	&\max\ \ln c_1(t) + \beta \ln c_2(t+1)\\
	s.t.\ &c_1(t) +s(t) = w(t) - T\\
	& c_2(t+1) = (s(t)+T)(r(t+1) + 1)\\
	&c_1(t) \ge 0, c_2(t+1) \ge 0
	\end{align*}
	$c_2(0)$ solves:
	\begin{align*}
	&\max\ \ln c_2(0)\\
	s.t.\ &c_2(0) = (s(-1)+T)(r(0) + 1)\\
	&s(-1) \text{ is given }
	\end{align*}
	Firms' optimization conditions imply that:
	\begin{align*}
	r(t) &= A\alpha k(t)^{\alpha - 1}\\
	w(t) &= A(1 - \alpha) k(t)^{\alpha}
	\end{align*}
	And market clearing conditions are:
	\begin{align*}
	k(t+1)(1+n) &= Ak(t)^{\alpha} + (1-\delta)k(t) - c_1(t) - \frac{c_2(t)}{n+1}\\
	k(t+1) &= \frac{s(t) + T}{1+n}
	\end{align*}
	Plugging expressions for $c_1(t)$ and $c_2(t+1)$ into  the objective function and taking derivatives with respect to $s(t)$ we get:
	\begin{align*}
	\frac{1}{w(t) - s(t) - T} = \frac{\beta}{(s(t)+T)(1 + r(t+1))}\nonumber\\
	s(t) = \frac{\beta w(t) - \beta T - T}{1+\beta} = (n+1)k(t+1) - T\\
	s(t) = \frac{\beta w(t)}{1+\beta} = (n+1)k(t+1)
	\end{align*}
	Which coincides with saving without pension, i.e. $T$ affects neither steady state level nor the equilibrium trajectory.
	\end{enumerate}
\end{document}

### 2019-02-23 metrics6.tex
metrics

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage{makecell}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\newcommand{\yrseduc}{\textit{yrseduc}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Econometrics 1. Problem Set 6.
\end{center}
\section*{Problem 1}
Assume that problem 8.7 in the textbook means the problem 7 from the chapter 8.

\textbf{Solution}

\begin{enumerate}[(i)]
	\item \begin{align*}
	Var(u_{i, e}) = Var(f_i + \nu_{i, e}) = Var(f_i) + Var(\nu_{i, e}) + 2cov(f_i, \nu_{i, e}) = \sigma^2_f + \sigma^2_{\nu} \sigma^2
	\end{align*}
	\item \begin{align*}
	cov(u_{i, e}, u_{i, g}) = cov(f_i + \nu_{i, e}, f_i + \nu_{i, g}) = \sigma^2_f + 0 + 0 + 0 = \sigma^2_f
	\end{align*}
	\item \begin{align*}
	\bar{u}_i &= \frac{1}{m_i} \sum_{e=1}^{m_i} u_{i, e}\\
	Var(u_i) = Var\left(f_i + \frac{1}{m_i} \sum_{e=1}^{m_i} \nu_{i, e}\right) &= \sigma^2_f + \frac{1}{m_i}Var(\nu_{1, e}) = \sigma^2_f + \frac{1}{m_i}\sigma^2_{\nu}
	\end{align*}
	\item 
	\begin{align*}
	Var(m_i \bar{u}_i) = m_i^2\sigma^2_f + m_i \sigma^2_{\nu}
	\end{align*}
	as a result it still depends on firm ($i$), so, such weights cannot lead to efficient WLS.
\end{enumerate}

\section*{Problem 2}

\textbf{Solution}

\begin{enumerate}[(a)]
	\item \begin{align}
	y = \beta_0 + \beta_1 x^* + u\nonumber\\
	x^* = \tilde{x_1} - e_1\label{eq1}\\
	x^* = \tilde{x_2} - e_1 - e_2\label{eq2}
	\end{align}
	Firstly regress $y$ on $\tilde{x}_1$ and get residuals $\hat{r}$ and the slope coefficient $\hat{\alpha_1}$. Then repeat the procedure for regression $y$ on $\tilde{x_2}$ and get residuals $\hat{v}$. Then 
	\begin{align*}
	\frac{1}{n \hat{\alpha}_1^2} \summa (\hat{r_i}^2 - \hat{v_i}^2)
	\end{align*}
	will be a consistent estimator for $\sigma^2_e$. Let us prove it.
	\begin{align*}
	\hat{r_{i}} = y_i - \hat{\alpha_0} - \hat{\alpha_1}\tilde{x}_1\\
	\hat{v}_i = y_i - \hat{\gamma}_0 - \hat{\gamma}_1 \tilde{x}_{2i}
	\end{align*}
	Where $\hat{\alpha_0}, \hat{\alpha_1}, \hat{\gamma}_0, \hat{\gamma}_1$ are OLS estimates from the regression $y$ on $\tilde{x}_1$. From \eqref{eq1} and we can infer that:
	\begin{align*}
	y = \beta_0 + \beta_1 \tilde{x}_1 - \beta_1 e_1 + u\\
	\end{align*}
	that means that $\hat{\alpha_0} \overset{p}{\to} \beta_0$ and $\hat{\alpha_1} \overset{p}{\to} \beta_1$. Since $\bar{\hat{r}}_i = 0$ and observations are i.i.d., that means that
	\begin{align*}
	\frac{1}{n} \summa \hat{r}_i^2 \overset{p}{\to} Var(u - \beta_1 e_1) \underset{indep.}{=} \sigma^2 + \beta_1^2 \sigma^2_e
	\end{align*}
	Similarly,
	\begin{align*}
	\frac{1}{n} \summa \hat{v}_i^2 \overset{p}{\to} Var(u - \beta_1 e_1 - \beta_1 e_2) \underset{indep.}{=} \sigma^2 + 2\beta_1^2 \sigma^2_e
	\end{align*}
	Then
	\begin{align*}
	\frac{1}{n \hat{\alpha}_1^2}\summa(\hat{v}_i^2 - \hat{r}_i^2) \overset{p}{\to} \sigma^2_e
	\end{align*}
	\item 
	\begin{align*}
	\hat{\alpha_1} = \beta_1 + \frac{-\beta_1 \summa (\tilde{x}_{1i} - \bar{\tilde{x}}_1)e_{1i} + \summa(\tilde{x}_{1i} - \bar{\tilde{x}}_1)u_{i} }{\summa (\tilde{x}_{1i} - \bar{\tilde{x}}_1)^2}
	\end{align*}
	and asymptotic bias is:
	\begin{align*}
	b_1 = \frac{-\beta_1 cov(\tilde{x}_1, e_1) + cov(\tilde{x}_1, u)}{Var(\tilde{x}_1)} = \frac{-\beta_1 \sigma^2_e}{Var(x^*) + \sigma^2_e}
	\end{align*}
	Similarly for 
	\begin{align*}
	\hat{\gamma}_1 = \beta_1 + \frac{-\beta_1 \summa (\tilde{x}_{2i} - \bar{\tilde{x}}_2)e_{1i} -\beta_1 \summa (\tilde{x}_{2i} - \bar{\tilde{x}}_2)e_{2i} + \summa(\tilde{x}_{2i} - \bar{\tilde{x}}_2)u_{i} }{\summa (\tilde{x}_{2i} - \bar{\tilde{x}}_2)^2}
	\end{align*}
	and asymptotic bias is:
	\begin{align*}
	b_2 = \frac{-\beta_1 cov(\tilde{x}_2, e_1) -\beta_1 cov(\tilde{x}_2, e_2)+ cov(\tilde{x}_2, u)}{Var(\tilde{x}_2)} = \frac{-2\beta_1 \sigma^2_e}{Var(x^*) + 2\sigma^2_e}
\end{align*}
hence
\begin{align*}
b_1 - b_2 = \frac{\beta_1\sigma^2_eVar(x^*)}{(Var(x^*) + \sigma^2_e)(Var(x^*) + 2\sigma^2_e)}\begin{cases}
>0, \beta_1 > 0\\
\le 0, \beta_1 \le 0
\end{cases}
\end{align*}
\item $\tilde{x} = \alpha \tilde{x}_1 + (1 - \alpha)\tilde{x}_2$
\begin{align*}
x^* &= \tilde{x} - e_1 - (1 - \alpha)e_2\\
y &= \beta_0 + \beta_1 \tilde{x} - \beta_1 e_1 - \beta_1(1 - \alpha)e_2 + u
\end{align*}
Regress $y$ on $\tilde{x}$:
\begin{align*}
y = \delta_0 + \delta_1 \tilde{x} + \varepsilon
\end{align*}
\begin{align*}
\hat{\delta}_1 = \beta_1 + \frac{-\beta_1 \summa (\tilde{x}_{i} - \bar{\tilde{x}})e_{1i} -\beta_1(1 - \alpha) \summa (\tilde{x}_{i} - \bar{\tilde{x}})e_{2i} + \summa(\tilde{x}_{i} - \bar{\tilde{x}})u_{i} }{\summa (\tilde{x}_{i} - \bar{\tilde{x}})^2}
\end{align*}
and asymptotic bias is:
\begin{align*}
b_2 = \frac{-\beta_1 cov(\tilde{x}, e_1) -\beta_1(1 - \alpha) cov(\tilde{x}, e_2)+ cov(\tilde{x}, u)}{Var(\tilde{x})} = \frac{(\alpha - 2)\beta_1 \sigma^2_e}{Var(x^*) + (1 + (1 - \alpha)^2)\sigma^2_e}
\end{align*}
as we can see, if $\alpha = 2$ then asymptotic bias is equal to 0.
\item Yes, if we regress $y$ on $2\tilde{x}_1 - \tilde{x}_2$ then as it has been shown above, asymptotic bias of OLS estimate of slope will be 0.
\end{enumerate}
\section*{Problem 3}
Assume that Problem C8.10 means the problem C10 from the 8th chapter.

\textbf{Solution}

\begin{enumerate}[(i)]
	\item The results for both approaches are depicted below:
	\begin{figure}[H]
		\centering
		\includegraphics[width=0.6\textwidth]{nonrobust}
		\caption{Usual OLS standard errors}\label{fig1}
	\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{robust}
	\caption{Heteroskedasticity robust standard errors}\label{fig2}
\end{figure}
As we can see, the errors differ, however it does not affect significance of estimates, so differences are not important.
\item 
\begin{align*}
Var(y|x) = \expect[u^2|x] = \expect[y|x] - (\expect[y|x])^2
\end{align*}
hence, if
\begin{align*}
\hat{u}_i^2 = \gamma_0 + \gamma_1 \hat{y}_i + \gamma_2 \hat{y}_i^2 + \varepsilon_i
\end{align*}
then
\begin{align*}
\hat{\gamma}_1 \overset{p}{\to} 1\\
\hat{\gamma}_2 \overset{p}{\to} -1\\
\hat{\gamma}_0 \overset{p}{\to} 0
\end{align*}
\item The results are depicted below
\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{white}
	\caption{}\label{fig3}
\end{figure}
as we can conclude the obtained estimates correspond to theoretical results from (ii).
\item The results are depicted below
\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{weighted}
	\caption{}\label{fig4}
\end{figure}
Most coefficient differ slightly, only coefficient on $male$ differs dramatically, however this coefficient is not significant in both models.
\end{enumerate}
\section*{Problem 4}

\textbf{Solution}

\begin{enumerate}[(i)]
	\item Coeftest of regression is depicted below
	\begin{figure}[H]
		\centering
		\includegraphics[width=0.6\textwidth]{401}
		\caption{}\label{fig5}
	\end{figure}
That means that on average participating in 401(k) plan increases the probability of having individual retirement account by 0.0536, holding income and age fixed.
\item The problem is that error term can be correlated with $p401k$ because people who have personal retirement account can be more inclined for saving and as a result can tend to participate in 401(k) more likely than others.
\item  $e401k$ should be correlated with $p401k$ and be uncorrelated with the error term. Of course it is correlated, because in order to participate in 401(k), a worker should be eligible for participating.

I have no idea how American pension system is organized, and what people should do in order to become eligible for participating, and unfortunately have no time to sort it out. So, if in order to become eligible people should undertake some actions then, $e401k$ is not a good IV, because of course those who are more inclined for saving would on average more likely to become eligible, otherwise it can be a good IV.
\item 	
The results are depicted below
\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{p401}
	\caption{}\label{fig6}
\end{figure}
Coefficient on $e401k$ is statistically significant and rather large, so indeed $e401k$ has correlation with $p401k$.
\item Results of IV regression are depicted below
\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{iv401}
	\caption{}\label{fig7}
\end{figure}
This estimate is almost twice smaller than usual OLS estimate.
\item As we can see, coefficient on residuals significantly differ from zero, that means that we should reject the hypothesis, i.e. $p401k$ is in fact endogenous.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{endog}
	\caption{}\label{fig8}
\end{figure}
\end{enumerate}

\end{document}

### 2019-02-23 Metrics6.R
metrics

@@ -1,0 +1,34 @@
+df <- read.csv("401KSUBS.csv")
+#Problem C8.10
+#i
+df$inc_2 = df$inc^2
+df$age_2 = df$age^2
+fit1 <- lm(e401k ~ inc+inc_2+age+age_2+male, data = df)
+summary(fit1)
+library(lmtest)
+library(sandwich)
+coeftest(fit1, vcov = vcovHC(fit1, type = "HC1"))
+#iii
+df$u_2 = fit1$residuals^2
+df$y_hat = fit1$fitted.values
+df$y_hat_2 = df$y_hat^2
+fit2 <- lm(u_2 ~ y_hat+y_hat_2, data = df)
+coeftest(fit2, vcov = vcovHC(fit2, type = "HC1"))
+#iiii
+max(df$y_hat)
+min(df$y_hat)
+df$h = df$y_hat * (1 - df$y_hat)
+fit3 <- lm(e401k ~ inc+inc_2+age+age_2+male, weights = 1/h, data = df)
+coeftest(fit3, vcov = vcovHC(fit3, type = "HC1"))
+
+#Problem C15.8
+#i
+fit4 <- lm(pira ~ p401k + inc+inc_2+age+age_2, data = df)
+coeftest(fit4, vcov = vcovHC(fit4, type = "HC1"))
+#iv
+fit5 <- lm(p401k ~ e401k+inc+inc_2+age+age_2+male, data=df)
+coeftest(fit5, vcov = vcovHC(fit5, type = "HC1"))
+#v
+df$res1 = fit5$residuals
+fit6 <- lm(pira ~ p401k + inc+inc_2+age+age_2+res1, data = df)
+coeftest(fit6, vcov = vcovHC(fit6, type = "HC1"))

### 2019-03-04 micro4_1.tex
Micro4

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Microeconomics 4. Problem Set 1.
\end{center}
\section*{1\ Strategic entry deterrence}
A market is open for two periods, and demand is initially uncertain. With probability $p$,
demand is high $(H)$ in both periods, and with probability $1-p$, demand is low $(L)$ in both
periods. In period 1, a single firm (the incumbent) is alone in the market. In period 2,
the incumbent remains in the market; another firm (the entrant) decides whether or not
to enter and can, if it wishes, base its decision upon the incumbent's first-period profits.
The incumbent's profitability in each period is limited by the monopoly profit (which is
$M^H$ if demand is high and $M^L < M^H$ if demand is low); however, the incumbent may
decide to earn a smaller profit in period 1, by secretly slacking. If the entrant enters,
both firms' second-period duopoly profit is $D^H$ if demand is high and $D^L$
if demand
is low, where $D^L < 0 < D^H < M^L$. If entry does not occur, the entrant obtains 0
in the second period, while the incumbent earns $M^L$ or $M^H$, depending upon demand.
The incumbent's objective at the start of the game is to maximize the sum of first- and
second-period profit
\begin{enumerate}
	\item  Assume demand is publicly observable. Explain why, in equilibrium, the incumbent
	always chooses to earn the monopoly profit in the first period and the entrant enters
	if and only if demand is high.
	\item Assume instead, form now on, that the incumbent observes the level of demand at
	the start of period 1 but that the entrant does not observe it until after making
	the entry decision in period 2. Explain why, under this assumption of asymmetric information, there does not exist a perfect Bayesian equilibrium in which the
	incumbent uses the same strategy as in the equilibrium in question 1.
	\item Suppose $p D^H + (1-p)D^L < 0$. Describe the strategies and beliefs in one of the
	perfect Bayesian equilibria in this signalling model. State whether your equilibrium
	is pooling or separating
\end{enumerate}

\textbf{Solution}
\begin{enumerate}
	\item The tree of this game is depicted below.
		\begin{figure}[H]
		\centering
		\includegraphics[width=0.8\textwidth]{plotdraft}
		\caption{}\label{fig1}
	\end{figure}
Since $D^L < 0 < D^H$ the entrant in case of low demand will always choose not enter, and in case of high demand will always choose enter. The incumbent hence should solve the following optimization problems:
\begin{align*}
\underset{m \in [0, M^H]}{\max}\ m+D^H\ \text{ if demand is high}\\
\underset{m \in [0, M^L]}{\max}\ m+M^L\ \text{ if demand is low}\\
\end{align*}
That's why the optimal strategy for incumbent is to choose $M^H$ and $M^L$ in case of high and low demand respectively.
\item The tree of this game is depicted below
		\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{plotdraft}
	\caption{}\label{fig2}
\end{figure}
Denote the belief of the entrant that he is at the high demand case as $\omega$. If incumbent chooses $M^H$ as a profit for high demand case and $M^L$ as a profit for the low demand case, then $\omega = 1$, that means that the strategy of the entrant consistent with such a belief is choosing not enter if $m = M^H$ and enter if $m = M^L$. But since $D^H < M^L$ the incumbent has incentive to deviate and choose $M^L$ if demand is high in order to prevent the entrant from entry and obtain $M^L + M^H$ instead of $M^H + D^H$. Thus, the "honest" strategy is not an equilibrium one for the incumbent.
\item The tree of the game is depicted below
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{plotdraft}
	\caption{}\label{fig3}
\end{figure}
Denote $\omega$ is a condition belief of the entrant that the demand is high. Assume $\omega = p: pD^H + (1-p)D^L < 0$. For such a belief the optimal strategy for the entrant is enter once the incumbent's profit if $M^H$ and not enter otherwise (once it is $M^L$). The incumbent's strategy hence is to choose $M^L$ in both cases because
\begin{align*}
M^H + D^H < M^H + M^L
\end{align*}
Moreover $\omega = p$ is consistent with such a strategy of the incumbent. Thus, the equilibrium is:
\begin{align*}
&\text{Incumbent : } \left\{M^L, M^L\right\}\\
&\text{Entrant : } \left\{\text{enter if incumbent's profit is } M^H \text{ not enter if it is } M^L\right\}\\
&\text{Entrant's belief : } (p, 1-p)
\end{align*}
This is a pooling equilibrium.
\end{enumerate}
\section*{2\ Beer and Quiche}
Nick wakes up and feels like fighting. He goes to a local pub in a hope of finding a target
there. Indeed, there he sees a man (called Paul) ordering a breakfast. Nick has seen Paul
before, but never talked to him. Nick knows that Paul can be of two types: Wimp or
Surly. Surly type is always eager to fight, and Nick does not want to fight with a Surly
type. In turn, Wimp is weak and easy to fight with, so Nick would be eager to release
his anger on a Wimp type. However, Nick does not know Paul's type exactly, he only
knows that the probability of each type is 0.5. Further, Nick has read in a book called
"Real Men Don't Eat Quiche" by Bruce Feirstein that surly people typically have beer
for breakfast, while wimp guys prefer ordering a quiche. Nick observes what Paul has
for breakfast, and decides whether to fight or pass. The payoffs are summarized in the
matrices below, where the strategies of Nick are in the columns, and the ones of Paul are
in the rows

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{surlywimp}
	\label{fig4}
\end{figure}
where $c$ is an additional cost for a wimp type to drink beer in the morning.
\begin{enumerate}
\item What conditions on $c$ guarantee the existence of a separating PBE? Fully describe
such an equilibrium.
\item For what values of $c$ is there a pooling equilibrium in which both Surly and Wimp
Paul gets beer for breakfast? Fully describe such an equilibrium.
\item Does the equilibrium in question 2 satisfy the intuitive criterion?
\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item In the separating PBE the surly type chooses $beer$ and the wimp type chooses $quiche$. Since at the separating equilibrium Nick chooses to fight once Paul gets quiche for breakfast and to pass otherwise, getting a beer for wimp type should be such costly as even being aware of Nick's strategy Paul would not have a profitable deviation. That is, $1 - c < -1$ hence $c > 2$.

	If we denote belief for Paul of coming across the surly type condition on the fact that Paul gets the beer as $\omega$ and  belief for Paul of coming across the wimp type condition on the fact that Paul gets the quiche as $\lambda$, then the separating equilibrium is:
	\begin{align*}
	&\text{Paul : } \{beer, quiche\}\\
	&\text{Nick : } \{pass, fight\}\\
	& \omega = 1, \lambda = 0
	\end{align*}
	(first action of Paul corresponds to surly type and the second to wimp type, while first action of Nick corresponds to Paul orders beer and the second corresponds to Paul orders quiche)
	\item Assume $\omega = \frac{1}{2} > \frac{1}{3}$ in this case
	\begin{align*}
	2\omega + 1 - \omega < 0
	\end{align*}
	that means that Nich chooses $pass$ once Paul gets beer. If $\lambda > \frac{1}{3}$ then $-2\lambda + 1 - \lambda < 0$ which means that Nick chooses to pass even if Paul gets quiche. Since $c > 0$, $1 - c < 1$ and Paul should choose quiche for wimp type. In case $\lambda < \frac{1}{3}$ Nick chooses to fight once Paul gets quiche, as a result if $-1 \le 1 - c$ i.e. $c \le 2$ even wimp type will choose beer.
	Thus, the equilibrium is:
	\begin{align*}
	&\text{Paul : } \{beer, beer\}\\
	&\text{Nick : } \{pass, fight\}\\
	& \omega = \frac{1}{2}, \lambda < \frac{1}{3}
	\end{align*}
	(again, first action of Paul corresponds to surly type and the second to wimp type, while first action of Nick corresponds to Paul orders beer and the second corresponds to Paul orders quiche)
	\item It does not satisfy the intuitive criterion. It makes no sense for wimpish Paul to deviate, if $c \le 2$, whereas a surly Paul is indifferent between beer and quiche and in principle can deviate, that means that Nick should believe that if Paul gets quiche, he is surly and as a result he should avoid fighting, but being aware of that wimpish Paul chooses $quiche$, and the equilibrium is broken.
\end{enumerate}
\section*{3\ Education}

\end{document}

### 2019-03-05 micro4_1.tex
micro

@@
 \section*{3\ Education}
+Consider Spence's job-marker signalling model. There are two firms on the market, and
+two types of worker, $\theta_H$ and $\theta_L$; $\theta_H > \theta_L > 0$. The probability of type $\theta_H$ is $Prob(\theta =
+\theta_H) = \lambda$, and is common knowledge. The worker knows his productivity and chooses
+education $e \ge 0$. Education is productive: the productivity of a worker of type $\theta_i$ with
+education $e$ is given by:
+\begin{align*}
+\theta_i + e
+\end{align*}
+The firms observe $e$ and simultaneously make the worker a wage offer. Then the worker
+accepts the highest of the wages, $w$. The payoff of the worker of productivity $\theta_i, i = L, H,$ is:
+\begin{align*}
+w - c(\theta_i, e)
+\end{align*}
+where $c(\theta_i, e)$ is the cost of education $e$ for the worker of ability $\theta_i$,
+\begin{align*}
+c(\theta_i, e) = \frac{e^2}{2\theta_i}
+\end{align*}
+The payoff of the firm that hires the worker of type $\theta_i$ with education $e$ at the wage $w$ is
+\begin{align*}
+\theta_i + e - w
+\end{align*}
+Assume also that the firms have the same beliefs about the worker's productivity off equilibrium path. 
+\begin{enumerate}
+\item Consider the full information case, in which firms observe the type of the worker. What are the wages $w^*_L, w^*_H$, and the levels of education $e^*_L, e^*_H$ under this scenario?
 
+Now assume that the firms observe the educational choices of the workers but not the
+types.
+\item  Can this game have a PBE which replicates the full information outcome in terms of
+wages and educational levels? If yes, find the corresponding restrictions on the set
+of parameters of the problem $(\theta_H; \theta_L; \lambda)$, and suggest an intuitive explanation for
+your answer. If no, explain why not. Illustrate your answer graphically in $(w; e)$
+space (do not worry about the precision of your graph).
+\item Now consider pooling equilibria in this model:
+\begin{enumerate}[(a)]
+\item Are there situations in which $e_p = 0$ cannot be supported as a pooling equilibrium?
+If yes, describe the set of parameters $(\theta_H; \theta_L; \lambda)$ that corresponds to such a situation, and suggest an intuitive explanation for your answer.
+If no,
+explain why not. Illustrate your answer graphically in $(w; e)$ space .
+\item Can the pooling equilibria in this model be fully Pareto-ordered? Explain your
+answer.
+\item Formulate the dominance requirement
+for PBE refinement. Assume that
+the parameters are such that the set $\Omega$
+of pooling equilibria surviving the
+dominance requirement is non-empty. Does $\Omega$
+contain the pooling equilibrium
+which delivers the highest (pooling equilibrium) payoff to $\theta_H$?
+\end{enumerate}
+\end{enumerate}
+
+\textbf{Solution}
+
+\begin{enumerate}
+	\item Competition requires that firms have zero profit, that means that
+	\begin{align*}
+	w_L &= \theta_L + e_L\\
+	w_H &= \theta_H + e_H
+	\end{align*}
+	Thus, choosing the level of education workers of each type maximize:
+	\begin{align*}
+	\theta_L &+ e_L - \frac{e^2_L}{2\theta_L}, \text{ if } \theta = \theta_L\\
+	\theta_H &+ e_H - \frac{e^2_H}{2\theta_H}, \text{ if } \theta = \theta_H
+	\end{align*}
+	hence $e^*_H = \theta_H, e^*_L = \theta_L$ and $w^*_L = 2\theta_L, w^*_H = 2\theta_H$.
+	\item In this case, firms can perfectly distinguish the type of workers by their education level. Hence the beliefs should be $\mu(e_H) = 1, \mu(e_L) = 0$ where
+	\begin{align*}
+	\mu(e) = \prob(\text{the worker is of high type}|\text{the education level } e \text{ is observed})
+	\end{align*}
+	let us specify the beliefs off the equilibrium path as follows:
+	\begin{align*}
+	\mu(e) = \begin{cases}
+	0, e < e_H\\
+	1, e \ge e_H
+	\end{cases}
+	\end{align*}
+	Hence in order for the low type workers do not have a profitable deviation the following condition on parameters should hold:
+	\begin{align*}
+	2\theta_H - \frac{\theta_H^2}{2\theta_L} &\le 2\theta_L - \frac{\theta_L^2}{2\theta_L}\\
+	2(\theta_H - \theta_L) - \frac{1}{2\theta_L}(\theta_H^2 - \theta^2_L) &\le 0\\
+	\frac{(\theta_H - \theta_L)(3\theta_L - \theta_H)}{2\theta_L} &\le 0\\
+	\theta_L &\le \frac{\theta_H}{3}
+	\end{align*}
+	So, when the productivity of the low type is very low compare to that of high type, costs of education are very high for low type workers and it is not optimal for them to pretend to be a high type, which means that low type workers choose $e^*_L = \theta_L$ and high type workers choose $e^*_H = \theta_H$ and firms perfectly distinguish between types.
+	\item 
+	\begin{enumerate}[(a)]
+	\item At the pooling equilibrium both types of workers choose
+	\begin{align*}
+	e^*_H = e^*_L = e_p
+	\end{align*}
+	and $\mu(e_p) = \lambda$. The expected profit of the firm, given an education level $e$ is
+	\begin{align*}
+	\mu(e) \theta_H + (1 - \mu(e)) \theta_L + e
+	\end{align*}
+	that means that the equilibrium wages should be
+	\begin{align*}
+	w(e_p) =  \lambda \theta_H + (1 - \lambda) \theta_L + e_p
+	\end{align*} 
+	Let us define beliefs off the equilibrium path as follows:
+	\begin{align*}
+	\mu(e) = \begin{cases}
+	\lambda, e=e_p = 0\\
+	0, e \neq e_p
+	\end{cases}
+	\end{align*}
+	in this case
+	\begin{align*}
+	w(e) = \begin{cases}
+	\lambda \theta_L + (1 - \lambda) \theta_H, e = 0\\
+	\theta_L + e, e > 0
+	\end{cases}
+	\end{align*}
+	Then the high type is solving the following optimization problem:
+	\begin{align*}
+	\max\left\{ \underset{e > 0}{\max}\ \theta_L + e - \frac{e^2}{2\theta_H}, \lambda \theta_L + (1 - \lambda) \theta_H \right\} = \max\left\{\theta_L + \frac{\theta_H}{2}, \lambda \theta_L + (1 - \lambda) \theta_H \right\}
+	\end{align*}
+	If $\lambda \le \frac{1}{2}$ then $\theta_L + \frac{\theta_H}{2} > \lambda \theta_L + (1 - \lambda)\theta_H$ and that means that high type will deviate and choose positive $e^*_H = \theta_H$ level of education, i.e.
+$e_p = 0$ is not supportable.
+	\item No, they cannot be.
+Firstly, obviously that the greatest amount of efforts the workers can spend on education at any pooling equilibria cannot exceed $\bar{e}$ (see Fig.
+\ref{fig5}) because for any other education level $e > \bar{e}$ and respective equilibrium wage level $w(e) = \lambda \theta_H + (1 - \lambda)\theta_L + e$ the low type worker prefers $e = 0$.
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.8\textwidth]{plotdraft}
+		\caption{}\label{fig5}
+	\end{figure}
+Suppose now that the pooling equilibrium is attained at education level $e' < \bar{e}$ (see Fig.
+\ref{fig5}).
+The corresponding indifference curves and wage schedule is depicted on the Fig.
+\ref{fig5}.
+Both types are maximizing their utilities, given the wage schedule and equilibrium wages are consistent with beliefs.
+And this pooling equilibrium Pareto dominates the one with $e = \bar{e}$ (and also each equilibrium with $e' < e < \bar{e}$) because firms have the same zero expected profit, but both types attain higher indifference curves and consequently higher utility levels at $e'$.
+However, there always exists an education level $\underline{e} = \theta_H$ such that indifference curve of high type is tangent to the line $\lambda \theta_H + (1 - \lambda)\theta_L + e$ at this point (see Fig.
+\ref{fig5}).
+And each pooling  equilibrium level of education $e'' < \underline{e}$ will yield the high type worker strictly less utility.
+Since each point at line $\lambda \theta_H + (1 - \lambda) \theta_L + e$ for $e < \underline{e}$ corresponds to lower indifference curve for high type.
+Similarly for low type workers there exists an education level $\underline{\underline{e}} = \theta_L$ such that indifference curve of low type is tangent to the line $\lambda \theta_H + (1 - \lambda)\theta_L + e$ at this point (see Fig.
+\ref{fig5}).
+And each pooling  equilibrium level of education $e''' < \underline{\underline{e}}$ will yield the low type worker strictly less utility.
+Since each point at line $\lambda \theta_H + (1 - \lambda) \theta_L + e$ for $e < \underline{\underline{e}}$ corresponds to lower indifference curve for low type.
+As a result, there is a full Pareto order for pooling equilibrium for $e \in [\theta_H, \bar{e}]$ and for $e \in [0, \theta_L]$, for $e \in (\theta_L, \theta_H)$ there is no Pareto order, since some pooling equilibria are better for low type, but are worse for high type and vice-versa.
+\item Highest pooling equilibrium payoff for the high type is delivered at pooling equilibrium education level $e^* = \theta_H$
+	\end{enumerate}
+\end{enumerate}
 \end{document}

### 2019-03-06 micro4_1.tex
Micro

@@
 \ref{fig5}) because for any other education level $e > \bar{e}$ and respective equilibrium wage level $w(e) = \lambda \theta_H + (1 - \lambda)\theta_L + e$ the low type worker prefers $e = 0$.
+(here $\bar{e} = \theta_L\left(1 + \sqrt{1 + \frac{2\lambda(\theta_H - \theta_L)}{\theta_L}}\right) > \theta_L$)
 	\begin{figure}[H]
@@
 And this pooling equilibrium Pareto dominates the one with $e = \bar{e}$ (and also each equilibrium with $e' < e < \bar{e}$) because firms have the same zero expected profit, but both types attain higher indifference curves and consequently higher utility levels at $e'$.
-However, there always exists an education level $\underline{e} = \theta_H$ such that indifference curve of high type is tangent to the line $\lambda \theta_H + (1 - \lambda)\theta_L + e$ at this point (see Fig.
+However, there always exists an education level $\underline{e} = \theta_H$ (which is wlog $< \bar{e}$) such that indifference curve of high type is tangent to the line $\lambda \theta_H + (1 - \lambda)\theta_L + e$ at this point (see Fig.
 \ref{fig5}).
@@
 As a result, there is a full Pareto order for pooling equilibrium for $e \in [\theta_H, \bar{e}]$ and for $e \in [0, \theta_L]$, for $e \in (\theta_L, \theta_H)$ there is no Pareto order, since some pooling equilibria are better for low type, but are worse for high type and vice-versa.
-\item Highest pooling equilibrium payoff for the high type is delivered at pooling equilibrium education level $e^* = \theta_H$
+
+Case $\theta_H > \bar{e}$ is analogous, for $e \in (\theta_L, \theta_H)$ equilibria cannot be Pareto ranked.
+\item Dominance requirement requires firms to prescribe positive $\mu(e)$ only to such $e$ that is not a strictly dominated action for high type.
+In the sense that there is no action $e'$ which yields the high type strictly greater payoff regardless of what equilibrium wages are.
+Assumption that $\Omega$ in non-empty means that firstly $\theta_H < \bar{e}$ and on the top of that the indifference curve of high type, which goes through $(\bar{e}, \theta_H + \bar{e})$ lies below the indifference curve which is tangent to the line $\lambda \theta_H + (1 - \lambda) \theta_L$ at the point $\theta_H$ (see Fig.
+\ref{fig6}) because otherwise, the high type chooses the level $\max\ \{\theta_H, \bar{e}\}$ and low type chooses $0$ and $\Omega$, the set of pooling equilibria, is empty.
+	\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{plotdraft}
+	\caption{}\label{fig6}
+\end{figure} 
+The highest pooling equilibrium payoff for the high type is delivered at the education level $e = \theta_H$.
+In our case this equilibrium belongs to $\Omega$ because neither of types have $e = \theta_H$ as a strictly dominated action.
+On the Fig.
+\ref{fig6} there is a wage schedule for which $e_H = e_L = \theta_H$ is the best response for both types (i.e.
+there is no action which strictly dominates this action).
 	\end{enumerate}
 \end{enumerate}
+\section*{4\ Exams}
+Karen would like her friend Tim to pass his exams. Tim may choose to work at a
+private cost of $c$. If he slacks, he fails. If he works, he passes with probability $\theta$. Hence $\theta$
+is Tim's $ability$. If he passes, Karen receives a private benefit of $V > 0$ and Tim receives
+a private benefit of $U > 0$.
+Karen offers to pay a reward $b \in [0; V]$ to Tim if and only if he passes. Hence Karen
+receives a payoff $V-b$ if Tim passes, and zero otherwise. Tim receives a payoff of zero if
+he slacks, $-c$ if he works and fails, and $U + b-c$ if he works and passes. For simplicity,
+assume that Tim will choose to work whenever he is otherwise indifferent between working
+and slacking.
+Tim's ability is initially unknown; $\theta_H$ with probability $\alpha$ and $\theta_L$ otherwise where
+\begin{align*}
+1 > \theta_H > \theta_L > 0, [\alpha \theta_H + (1-\alpha)\theta_L] U < c < \theta_L(U + V ) \text{ and } c < \theta_H U
+\end{align*}
+We refer to Tim's beliefs about $\theta$ as his "confidence". Begin by assuming that neither
+player observes $\theta$. Hence Karen begins by choosing a bonus $b$. Tim observes this bonus,
+and then chooses whether to work
+\begin{enumerate}
+	\item What bonus will Karen choose to offer?
+
+	Karen now observes Tim's ability $\theta$, but Tim does not. Hence, Nature begins by
+	choosing $\theta \in \{\theta_H; \theta_L\}$. Next, Karen observes $\theta$ and offers a bonus $b_H$ (if $\theta = \theta_H$) or $b_L$
+	(if $\theta = \theta_L$). Finally, Tim observes $b$, but not $\theta$, and then chooses whether to work. We seek pure strategy perfect Bayesian Nash equilibria.
+	\item Show that there is no separating equilibrium, and hence $b^*_H = b^*_L$ in any equilibrium.
+	\item Does the standard sorting (single crossing) condition hold in part 2?
+
+	Suppose now that Karen offers Tim an unconditional gift $w \ge 0$ as well as the bonus
+	$b \ge 0$. Hence she pays $w+b$ if Tim passes, and $w$ if he fails. So Karen offers $(w_H; b_H)$ when
+	she observes $\theta_H$ and $(w_L; b_L)$ when she observes $\theta_L$. We seek a separating equilibrium.
+
+	\item Show that $w^*_L = 0$ and $b^*_L = \frac{c}{\theta_L} - U$ in any separating equilibrium. Show that
+	together with $w^*_H = c - \theta_LU$ and $b_H^* = 0$ this constitutes an equilibrium.
+	\item Briefly interpret this equilibrium
+\end{enumerate}
+
+\textbf{Solution}
+
+\begin{enumerate}
+	\item Since Karen gets 0 if Tim slacks and if he works she in the worst case gets 0, she should offer such a $b$ that Tim would be indifferent between work and slack, and as a result chooses to work.
+	\begin{align*}
+	\alpha(\theta_H(U + b - c) &- (1 - \theta_H)c) + (1 - \alpha)(\theta_L(U + b - c) - (1 - \theta_L)c) = 0\\
+	&(U+b)(\alpha\theta_H + (1 - \alpha)\theta_L) = c\\
+	&b = \frac{c}{\alpha\theta_H + (1 - \alpha)\theta_L} - U
+	\end{align*}
+	\item 
+	\item In separating equilibrium Tim can perfectly distinguish whether $\theta$ is high or low, hence in case of high $\theta$ he will work only if:
+	\begin{align*}
+	(U + b_H)\theta_H - c \ge 0\\
+	b_H \ge \frac{c}{\theta_H} - U < 0
+	\end{align*}
+	i.e. if $\theta = \theta_H$ Karen's optimal offer is $b_H = 0$. If $\theta = \theta_L$ then Tim will work only if
+	\begin{align*}
+	b_L \ge \frac{c}{\theta_L} - U
+	\end{align*}
+	and hence optimal bonus which Karen offers is:
+	\begin{align*}
+	b_L = \max\ \left\{\frac{c}{\theta_l} - U, 0\right\}
+	\end{align*}
+	but if $b_L = 0$ then equilibrium becomes pooling and Tim cannot distinguish between states.
+If $\frac{c}{\theta_l} - U > 0$ then Karen will want to deviate and offer $b_L = 0$ in order to convince Tim that he is of high type.
+Hence at each equilibrium $b^*_H = b^*_L$.
+	\item  As always at separating equilibrium Tim can perfectly distinguish between states. That means that at low state he will work only if
+	\begin{align*}
+	\theta_L(U + b_L) + w_L - c \ge w_L
+	\end{align*}
+	hence Karen should offer such $(b_L, w_L)$ that:
+	\begin{align*}
+	\theta_L b_L = c - \theta_LU
+	\end{align*}
+	In high state it is profitable for Tim to accept any positive offer $(b_H, w_H)$ and it would be optimal for Karen to offer $(0, 0)$ but it cannot be a separating equilibrium because at low state Karen will have a profitable deviation, namely $(0, 0)$ and equilibrium becomes pooling.
+So, at the separating equilibrium $(w_H, b_H)$ should be such that Karen does not want to deviate at low state.
+That is:
+	\begin{align*}
+	\theta_L(V - b_H - w_H) + (1 - \theta_L)(-w_H) &\le \theta_L(V - b_L - w_L) + (1 - \theta_L)(-w_L)\\
+	\theta_L V - \theta_Lb_H - w_H &\le \theta_L(V + U) - c - w_L
+	\end{align*}
+	hence
+	\begin{align}\label{eq1}
+	\theta_L b_H + w_H = c +w_L - \theta_L U
+	\end{align}
+	and finally Karen should choose such $(b_H, w_H), (b_L, w_L)$ which maximize her expected payoff. At low state she should maximize:
+	\begin{align*}
+	\theta_L (V - b_L) - w_L = \theta_L V - c + \theta_LU - w_L \to \underset{w_L}{\max}
+	\end{align*}
+	obviously the solution is $w_L = 0$. Similarly, using \eqref{eq1}, at high state she should maximize:
+	\begin{align*}
+	\theta_HV - \theta_H b_H - w_H = \theta_HV - \theta_Hb_H - c + \theta_LU + \theta_L b_H \to \underset{b_H}{\max}\\
+	\theta_H V + \theta_l U - c - (\theta_H - \theta_L)b_H \to \underset{b_H}{\max}
+	\end{align*}
+	the solution is $b_H = 0$. Thus, the equilibrium is
+	\begin{align*}
+	w^*_L = 0,\ b^*_L = \frac{c}{\theta_L} - U\\
+	w^*_H = c - \theta_LU,\ b^*_H = 0
+	\end{align*}
+	as we sought.
+	\item The interpretation is straightforward.
+Without unconditional gift Karen pays her bonus ex post, as a result she had incentives to deviate and underpay Tim, to cheat him that he is of high type.
+Whereas the introduction of the gift which Karen pays ex ante can be viewed as a commitment that Karen will not cheat Tim and will not make his expected payoff negative when Tim is of low type.
+\end{enumerate}
+\section*{5\ Beliefs}
+The definition of the world $arbitrary$ is the following:
+\begin{align*}
+\text{based on random choice or personal whim, rather than any reason or system}
+\end{align*}First of all such beliefs are not fully arbitrary, they should be consistent with strategies, players play.
+That is, given those beliefs and actions the player should undertake according to them, players do not have profitable unilateral deviations from the equilibrium path.
+But even if we actually can arbitrary (to the aforementioned extent) choose beliefs for the off equilibrium path info sets, conclusions are not arbitrary in the slightest.
+They are fully consistent with this particular concept of equilibrium.
+We do not conclude that tomorrow will be rain because some information set at the game we study is reached with zero probability at the equilibrium.
+It is the example of arbitrary conclusion.
+But those conclusions we usually draw are correct for the model and concept of equilibrium we chose.
+If we do not like these conclusions we can only choose another concept of equilibrium, use some refinements for example, but we cannot say that conclusions are arbitrary, they are valid, correct for this concept and not random.
+\section*{6\ Wage Inequality}
+
 \end{document}

### 2019-03-06 micro4_1.tex
Micro

@@
 \section*{5\ Beliefs}
-The definition of the world $arbitrary$ is the following:
+The definition of the term $arbitrary$ is the following:
 \begin{align*}
@@
 \section*{6\ Wage Inequality}
+Suppose access to a university education becomes less dependent on wealth or social class.
+Using a signalling model of education, what predictions would you make about how wage
+inequality within age cohorts would change?
+
+\textbf{Solution}
+
+There can be three types of predictions: wage inequality can increase, decrease or remains unchanged.
+And in principle all three types can be possible, depending on parameters of the model and concept of equilibrium we use (PBE or PBE+intuitive criterion refinement)
+
+Assume for simplicity that we have only two types of workers (of high productivity and of low productivity) and that education is useless, in the sense that it does not increase productivity (a plausible assumption for Russian educational system).
+
+In addition assume that we have a separating equilibrium i.e.
+the wage inequality is $\theta_H - \theta_L$ and moreover let us use the intuitive criterion refinement which eliminates all pooling equilibria.
+Since indifference curves of low type workers become flatter, it is now becomes profitable for them to deviate and to pretend to be workers of high productivity.
+In this case either firms become unable to distinguish between the types and offer a single wage for all types (pooling equilibrium) and as a result, inequality decreases to 0, or workers of high type just increase their level of education (starting to pursue PhD for example) and firm still are able to distinguish between types and offer different wages, hence the wage inequality remains unchanged.
 

### 2019-03-11 metrics2_1.tex
metrics

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage{makecell}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\newcommand{\yrseduc}{\textit{yrseduc}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Econometrics 2. Problem Set 1.
\end{center}
\section*{Problem 1}
 In Example 10.4, we saw that our estimates of the individual lag coefficients in a distributed lag model were very imprecise. One way to alleviate the multicollinearity problem is to assume that the $\delta_j$ follow a relatively simple pattern. For concreteness, consider a model with four lags:
 \begin{align*}
y_t = \alpha_0 + \delta_0z_t + \delta_1 z_{t-1} + \delta_2 z_{t-2} + \delta_3 z_{t-3} + \delta_4 z_{t-4}+ u_t.
\end{align*} 
Now, let us assume that the $\delta_j$ follow a quadratic in the lag, $j$: 
\begin{align*}
\delta_j = \gamma_0 + \gamma_1 j + \gamma_2 j^2,
\end{align*}
for parameters $\gamma_0, \gamma_1$, and $\gamma_2$. This is an example of a polynomial distributed lag (PDL) model. 
\begin{enumerate}[(i)]
	\item  Plug the formula for each $\delta_j$ into the distributed lag model and write the model in terms of the parameters $\gamma_h$, for $h = 0,1,2$. 
	\item Explain the regression you would run to estimate the $\gamma_h$. 
	\item The polynomial distributed lag model is a restricted version of the general model. How many restrictions are imposed? How would you test these? (Hint: Think F test.)
\end{enumerate}

\textbf{Solution}

\begin{enumerate}[(i)]
	\item 
	\begin{align*}
	y_t = \alpha_0 + \gamma_0 z_t + (\gamma_0 + \gamma_1 + \gamma_2)z_{t-1}+(\gamma_0 + 2\gamma_1 + 4\gamma_2) z_{t-2} + (\gamma_0 + 3\gamma_1 + 9\gamma_2)z_{t-3} + \\
	+(\gamma_0 + 4\gamma_1 + 16\gamma_2)z_{t-4} + u_t\\
	y_t = \alpha_0 + (z_t + z_{t-1} + z_{t-2} + z_{t-3}+z_{t-4})\gamma_0 + (z_{t-1} + 2z_{t-2}+3z_{t-3}+4z_{t-4})\gamma_1 +\\
	+(z_{t-1} + 4z_{t-2} + 9z_{t-3} + 16z_{t-4})\gamma_2 + u_t
	\end{align*}
	\item I would create new variables $\tilde{z}_{1t} = z_t + z_{t-1} + z_{t-2} + z_{t-3}+z_{t-4}$, $\tilde{z}_{2t} = z_{t-1} + 2z_{t-2}+3z_{t-3}+4z_{t-4}$ and $\tilde{z}_{3t} = z_{t-1} + 4z_{t-2} + 9z_{t-3} + 16z_{t-4}$ and run the regression $y_t$ on $1, \tilde{z}_{1t}, \tilde{z}_{2t}, \tilde{z}_{3t}$, using OLS then to estimate $\gamma_h$
	\item In general model we have 6 parameters whereas in restricted model only 4, that means that 2 restrictions were imposed. Using formula for F-statistic through the $R^2$ of restricted and unrestricted models, namely:
	\begin{align*}
	F = \frac{(R^2_{ur} - R^2_r)(n-6)}{2(1 - R^2_{ur})}
	\end{align*}
	we can test the hypothesis whether restrictions are true or not. Under null hypothesis F-statistics has asymptotically $\frac{1}{2} \chi^2(2)$ distribution. And under assumption of normality of errors we can use also $F(2, n-6)$ distribution.
\end{enumerate}
\section*{Problem 2}
 In Example 10.4, we wrote the model that explicitly contains the long-run propensity, $\theta_0$, as 
 \begin{align*} 
 gfr_t = \alpha_0 + \theta_0 pe_t + \delta_1 (pe_{t-1} - pe_t) + \delta_2 (pe_{t-2} - pe_t) + u_t,
 \end{align*}  
 where we omit the other explanatory variables for simplicity. As always with multiple regression analysis, $\theta_0$ should have a ceteris paribus interpretation. Namely, if $pe_t$  increases by one (dollar) holding ($pe_{t-1} - pe_t$) and ($pe_{t-2} - pe_t$) fixed, $gfr_t$ should change by $\theta_0$. 
 \begin{enumerate}[(i)]
\item  If ($pe_{t - 1} - pe_t$) and ($pe_{t-2} - pe_t$) are held fixed but $pe_t$ is increasing, what must be true about changes in $pe_{t-1}$ and $pe_{t-2}$?
\item How does your answer in part (i) help you to interpret $\theta_0$ in the above equation as the LRP? 
\end{enumerate}

\textbf{Solution}

\begin{enumerate}[(i)]
	\item If $pe_t$ increases while $pe_{t-1} - pe_t$ remains unchanged that means that $pe_{t-1}$ should also increase by the same as $pe_t$, as well as $pe_{t-2}$.
	\item As the LRP $\theta_0$ shows the increase of $gfr$ in the long-run (from $t+2$ to $n$) once $pe_t$ permanently changes by 1 starting from period $t$. The result in (i) demonstrates that $\theta_0$ shows the increase in $gfr$ once starting from $t-2$ the $pe$ changes permanently. Obviously the results are equivalent up to renumbering of observations' indices. 
\end{enumerate}
\end{document}

### 2019-03-11 micro4_2.tex
Micro

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Microeconomics 4. Problem Set 1.
\end{center}
\section*{1\ The Doctor and the Patient}
Consider the cheap talk signalling model a-la Crawford and Sobel (1982). Suppose that
there is a patient that does not feel well and goes to see a doctor. The doctor should
decide which treatment $a$ the patient needs. The type of treatment will depend on
the health state of the patient $\theta$, which is only privately observed by the patient. The
doctor's prior about the patient's health is $\theta \sim U[0; 1]$. The patient can send a costless,
non-binding and non-verifiable signal $s \in [0; 1]$ about his health state (i.e. explain how
he feels). The doctor cares about her reputation, so she wants to prescribe a appropriate
treatment for the patient's health: $U
D(a; \theta) = -(a-\theta)^2$. Instead, the patient, that is
very hypochondriac, wants to be overtreated: $U^P(a; \theta) = -(a - (\theta + b))^2$, where $b > 0$ is
the bias towards excessive medication.
\begin{enumerate}
	\item Assume that b is such that there exist a 2-partitional equilibrium $(p = 2)$. Characterize this equilibrium.
	\item Under which condition will such an equilibrium exist?
	\item Is this the only equilibrium that exists? If yes, argue why. If no, characterize the
	other(s) equilibrium (equilibria).
	\item Compute the expected payoff of the doctor and the patient. How do the payoffs
	change with $b$?
\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item Assume that for $\theta < \theta_1$ the patient sends $s_1$ e.g. $s_1 = 0$, while for $\theta \ge \theta_1$ the patient sends $s_2 > s_1$ e.g. $s_2 = \theta_1$. Then the doctor chooses treatment as a maximizer of:
	\begin{align*}
	\begin{cases}
	\underset{a}{\max}\ \int_{0}^{\theta_1} -(a - \theta)^2d\theta, s < \theta_1\\
	\underset{a}{\max}\ \int_{\theta_1}^{1} -(a - \theta)^2d\theta, s \ge \theta_1\\
	\end{cases}
	\end{align*}
	and the solution is obviously
	\begin{align*}
	a(s) = \begin{cases}
	\frac{\theta_1}{2}, s < \theta_1\\
	\frac{1 + \theta_1}{2}, s \ge \theta_1
	\end{cases}
	\end{align*}
	The patient should send:
	\begin{align*}
	\begin{cases}s_1, -\left(\frac{\theta_1}{2} - \theta - b\right)^2 > -\left(\frac{\theta_1 + 1}{2} - \theta - b\right)^2\\
	s_2, \text{ otherwise}
	\end{cases}
	\end{align*}
	That means that $\theta_1$ should be such that the patient becomes indifferent between sending $s_1$ and $s_2$ i.e.
	\begin{align*}
	\frac{\frac{\theta_1}{2} - b + \frac{\theta_1+1}{2} - b}{2} &= \theta_1\\
	\theta_1 &= \frac{1}{2} - 2b
	\end{align*}
	\item 
	\begin{align*}
	\frac{1}{2} - 2b \ge 0\\
	b \le \frac{1}{4}
	\end{align*}
	\item No, there always exists a "babbling" equilibrium. In which the patient always sends the same, uninformative signal and doctor just chooses action which maximize his expected payoff taking in account doctor's prior beliefs about type. Given that, patient indifferent between any signal (he cannot change anything) hence he has not profitable deviation.
	\item The expected payoff of the doctor is:
	\begin{align*}
	\pi^D = \int_{0}^{\frac{1}{2} - 2b} - \left(\frac{1}{4} - b - \theta\right)^2d\theta + \int_{\frac{1}{2} - 2b}^{1}- \left(\frac{3}{4} - b - \theta\right)^2d\theta = -2\frac{\left(\frac{1}{4} - b\right)^3}{3} -2\frac{\left(\frac{1}{4} + b\right)^3}{3}
	\end{align*}
	The expected payoff of the patient is:
	\begin{align*}
	\pi^P = \int_{0}^{\frac{1}{2} - 2b} - \left(\frac{1}{4} - 2b - \theta\right)^2d \theta + \int_{\frac{1}{2} - 2b}^1 - \left(\frac{3}{4} - 2b - \theta\right)^2d \theta = -\frac{1}{3 \cdot 2^5} + \\
	+\frac{1}{3} \left(\left(\frac{3}{4} - 2b\right)^3 + \left(\frac{1}{4} - 2b\right)\right)
	\end{align*}
	\begin{align*}
	&\frac{\partial \pi^D}{\partial b} = 2\left(\frac{1}{4} - b - \frac{1}{4} - b\right)\left(\frac{1}{4} - b + \frac{1}{4} + b\right) = -2b < 0\\
	&\frac{\partial \pi^P}{\partial b} = - 2 \left(\frac{3}{4} - 2b - \frac{1}{4} + 2b\right)\left(\frac{3}{4} - 2b + \frac{1}{4} - 2b\right) = -1 + 4b < 0
	\end{align*}
	That is payoffs goes down once $b$ increases.
\end{enumerate}
\section*{ 2 Labour Market}
There are two types of workers at the labor market, $i = 1, 2$. A worker of type $i$ has a utility function
\begin{align*}
u_i(w, x) = w - c_i(x)
\end{align*}
where $w$ is the wage paid to the worker, $x$ is the output he produces, and $c_i(x)$ is a convex
and increasing cost function, $c'_1(x) < c'_2(x)$, $c_i(0) = 0$, $c'_i(0) = 0$, $c'_i(\infty) = \infty$ for $i = 1, 2$.
The share of workers of the first type is given by $\lambda$.

\begin{enumerate}
	\item Assume that there is a single hiring firm at the market, and it does not observe the
	workers' types. The profit of the firm is given by the value of output less the wage
	payment
	\begin{align*}
	\pi = x - w
	\end{align*}
	Assume that the reservation wage of both workers is zero. Characterize the monopolistic screening contracts $(w_i, x_i)$ offered by the firm.
	\item Now assume that there are multiple firms in the market, but none of them observes
	the workers' types. The firms are competing in hiring workers. Characterize the
	competitive screening SPNE.
	\item How would your answers to (1) and (2) change if the firms can observe the types
	of the workers? Why can the first best be (or not be) achieved in each of these
	scenarios?
	\item Assume that $c_i(x) = \frac{i}{2}x^2,\ i = 1,\ 2$ and $\lambda = 1/2$. Derive the levels of output and the
	wages in your monopolistic and competitive screening problems with unobservable
	and observable worker types (i.e., in (1)-(3))
\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item The monopoly's problem is:
	\begin{align*}
	\underset{(x_1, w_1), (x_2, w_2)}{\max}\ \lambda(x_1 - w_1) + (1 - \lambda)(x_2 - w_2)
	\end{align*}
	Since outside options of workers are 0, the optimal contracts monopoly proposes to workers should be such that
	\begin{align*}
	w_1^* = c_1(x_1^*)\\
	w_2^* = c_2(x_2^*)
	\end{align*}
	Assuming interior solution FOC is:
	\begin{align*}
	\begin{cases}
	1 - c_1'(x_1^*) = 0\\
	1 - c_2'(x_2^*) = 0
	\end{cases}\\
	\end{align*}
	\begin{align*}
	c_1'(x_1^*) = c_2'(x_2^*) \to x_1^* > x_2^* \to c_1(x_1^*) > c_1(x_2^*) \to w_2^* - c_1(x_2^*) > w_1^* - c_1(x_1^*) = 0
	\end{align*}
	That means that at any interior solution high type has incentive to deviate and mimic to low type. Hence we should add incentive compatibility constraints. The IC constraints are:
	\begin{align*}
	\begin{cases}
	w_1^* - c_2(x_1^*) \le w_2^* - c_2(x_2^*)\\
	w_2^* - c_1(x_2^*) \le w_1^* - c_1(x_1^*)
	\end{cases}\\
	\begin{cases}
	c_1(x_1^*) - c_2(x_1^*) \le 0\\
	c_2(x_2^*) - c_1(x_2^*) \le 0
	\end{cases}
	\end{align*}
	Since $c_2(x) > c_1(x) \forall x > 0$ it follows that $x_2^* = 0$.
\end{enumerate}
\end{document}

### 2019-03-14 metrics2_2.tex
metrics

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\newcommand{\yrseduc}{\textit{yrseduc}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Econometrics 2. Problem Set 2.
\end{center}
\section*{Problem 1}
Let $\{y_t: t = 1, 2,\dots\}$ follow a random walk, as in (11.20), with $y_0 = 0$. Show that $Corr(y_t, y_{t+h}) =\sqrt{\frac{t}{t + h}}$   for $t \ge 1, h > 0$.

\textbf{Solution}

\begin{align*}
y_t = y_{t-1} + e_t = y_{t-2} + e_{t-1} + e_t = \dots = y_0 + e_1 + \dots + e_t\\
y_{t+h} = y_{t+h-1} + e_{t+h} = \dots = y_0 + e_1 + \dots + e_{t} + \dots + e_{t+h}
\end{align*}
\begin{align*}
&Cov(y_t, y_{t+h}) = (y_0 + e_1 + \dots + e_t, y_0 + e_1 + \dots + e_{t} + \dots + e_{t+h}) \underset{e_t \text{ i.i.d. and indep. of }y_0}{=} t\sigma^2_e\\
&Var(y_t) = Var(y_0 + e_1 + \dots + e_t)\underset{e_t \text{ i.i.d. and indep. of }y_0}{=} t\sigma^2_e\\
&Var(y_{t+h}) = Var(y_0 + e_1 + \dots + e_{t} + \dots + e_{t+h}) \underset{e_t \text{ i.i.d. and indep. of }y_0}{=} (t+h) \sigma^2_e\\
\end{align*}
\begin{align*}
Corr(y_t, y_{t+h}) = \frac{t\sigma^2_e}{\sqrt{t(t+h)}\sigma^2_e} = \sqrt{\frac{t}{t+h}}
\end{align*}
\section*{Problem 2}
A partial adjustment model is
\begin{align*}
y_t^* = \gamma_0 + \gamma_1 x_t + e_t\\
y_t - y_{t-1} = \lambda(y_t^* - y_{t-1}) + a_t, 
\end{align*}
where $y_t^*$ is the desired or optimal level of $y$, and $y_t$ is the actual (observed) level. For example, $y_t^*$ is the desired growth in firm inventories, and $x_t$ is growth in firm sales. The parameter $\gamma_1$ measures the effect of $x_t$ on $y_t^*$. The second equation describes how the actual $y$ adjusts depending on the relationship between the desired $y$ in time $t$ and the actual $y$ in time $t-1$. The parameter $\lambda$ measures the speed of adjustment and satisfies $0 < \lambda <1$.

\begin{enumerate}[(i)]
	\item  Plug the first equation for $y_t^*$ into the second equation and show that we can write
	\begin{align*}
	y_t = \beta_0 = \beta_1 y_{t-1} + \beta_2 x_{t}+u_t.
	\end{align*}
	In particular, find the $\beta_j$ in terms of the $\gamma_j$ and $\lambda$ and find $u_t$ in terms of $e_t$ and $a_t$. Therefore, the partial adjustment model leads to a model with a lagged dependent variable and a contemporaneous $x$. 
	\item  If $\expect(e_t|x_t, y_{t-1}, x_{t-1}, \dots) = \expect(a_t|x_t, y_{t-1}, x_{t-1},\dots) = 0$ and all series are weakly dependent, how would you estimate the $\beta_j$? 
	\item If $\hat{\beta}_1 = 0.7$ and $\hat{\beta}_2=0.2$, what are the estimates of $\gamma_1$ and $\lambda$?
\end{enumerate}

\textbf{Solution}

\begin{enumerate}[(i)]
	\item \begin{align*}
	y_t - y_{t-1} = \lambda(\gamma_0 + \gamma_1 x_t + e_t - y_{t-1}) + a_t\\
	y_t = \lambda \gamma_0 + (1 - \lambda)y_{t-1} + \lambda \gamma_1 x_t + \lambda e_t + a_t\\
	\beta_0 = \lambda \gamma_0,\ \beta_1 = 1 - \lambda,\ \beta_2 = \lambda \gamma_1,\ u_t = \lambda e_t + a_t
	\end{align*}
	\item Since
	\begin{align*}
	\expect[u_t|x_t, y_{t-1}, x_{t-1}, \dots] = 0
	\end{align*}
	and series are weakly dependent then we can consistently estimate $\beta_j$ by OLS regressing $y_t$ on $y_{t-1}, x_{t}$.
	\item \begin{align*}
	\hat{\lambda} = 1 - \hat{\beta}_1 = 0.3\\
	\hat{\gamma}_1 = \frac{\hat{\beta}_2}{\hat{\lambda}} \approx 0.67
	\end{align*}
\end{enumerate}
\end{document}

### 2019-03-14 metrics2_2.tex
metrics

@@
 \end{enumerate}
+
+\section*{Problem 3}
+Suppose that the equation 
+\begin{align*}
+y_t = \alpha + \delta _t + \beta_1 x_{t1}+\dots+\beta_k x_{tk}+u_t
+\end{align*} satisfies the sequential exogeneity assumption in equation (11.40). 
+\begin{enumerate}[(i)]
+\item Suppose you difference the equation to obtain 
+\begin{align}\label{eq1}
+\Delta y_t = \delta +\beta_1 \Delta x_{t1} + \dots+ \beta_k \Delta x_{tk} + \Delta u_t.
+\end{align}
+How come applying OLS on the differenced equation does not generally result in consistent estimators of the $\beta_j$?
+\item What assumption on the explanatory variables in the original equation would ensure that OLS on the differences consistently estimates the $\beta_j$?
+\item Let $z_{t1},\dots, z_{tk}$ be a set of explanatory variables dated contemporaneously with $y_t$.
+If we specify the static regression model $y_t = \beta_0 + \beta_1z_{t1} +\dots+ \beta_k z_{tk} + u_t$, describe what we need to assume for $x_t = z_t$ to be sequentially exogenous.
+Do you think the assumptions are likely to hold in economic applications?
+\end{enumerate}
+
+\textbf{Solution}
+
+\begin{enumerate}
+	\item First of all, since we do not have any assumptions about the model apart from sequential exogeneity
+	\begin{align*}
+	\expect[u_t|\vec{x}_t, \vec{x}_{t-1}, \dots] = \expect[u_t] = 0
+	\end{align*}
+	we can have a perfect multicollinearity among $\vec{x}_t$. Assume for example $k = 3$, and the model is:
+	\begin{align*}
+	y_t =\alpha + \delta t + \beta_1 x_t + \beta_2 z_t + \beta_3 w_t + u_t
+	\end{align*} 
+	where $w_t = x_t + z_t\ \forall\ t$. In this case obviously $\Delta w_t = \Delta x_t + \Delta z_t$ and the OLS estimate for \eqref{eq1} is inconsistent.
+	\item To ensure that OLS for \eqref{eq1} is consistent. Assumptions TS1'-TS3' should hold. The process \eqref{eq1} should be weakly dependent (and stationary)
+\end{enumerate}
 \end{document}

### 2019-03-14 metrics2_2.tex
metrics

@@
 	where $w_t = x_t + z_t\ \forall\ t$. In this case obviously $\Delta w_t = \Delta x_t + \Delta z_t$ and the OLS estimate for \eqref{eq1} is inconsistent.
-	\item To ensure that OLS for \eqref{eq1} is consistent. Assumptions TS1'-TS3' should hold. The process \eqref{eq1} should be weakly dependent (and stationary)
+	\item To ensure that OLS for \eqref{eq1} is consistent.
+Assumptions TS1'-TS3' should hold.
+The process \eqref{eq1} should be weakly dependent (and stationary).
+Of course, if $h_t = \beta_1 x_{t1} + \dots + \beta_k x_{tk} + u_t$ is weakly dependent and stationary then the process $\Delta h_t = \beta_1 \Delta x_{t1} + \dots + \beta_k \Delta x_{tk} + \Delta u_t$ (and hence the process $\Delta y_t$) will be weakly dependent and stationary (though this condition is not necessary, e.g.
+random walk process is not a weakly dependent while the differenced process is).
+Moreover, $x_{t1}, \dots, x_{tk}$ should not be linearly dependent (then $\Delta x_{t1}, \dots, \Delta x_{tk}$ will not linearly dependent too).
+In addition, contemporaneous exogeneity assumption should hold, i.e.
+	\begin{align*}
+	\expect[\Delta u_t| \Delta x_{t1}, \dots, \Delta x_{tk}] = 0
+	\end{align*} 
+	\begin{align*}
+	\expect[\Delta u_t|\Delta x_{t1}, \dots, \Delta x_{tk}] = \expect[u_{t}|\Delta x_{t1}, \dots, \Delta x_{tk}] - \expect[u_{t-1}|\Delta x_{t1}, \dots, \Delta x_{tk}] = \\
+	=\expect[\expect[u_t|\vec{x}_{t}, \vec{x}_{t-1}, \dots]|\Delta x_{t1}, \dots, \Delta x_{tk}] - \expect[u_{t-1}|\Delta x_{t1}, \dots, \Delta x_{tk}] = \\
+	=0 - \expect[u_{t-1}|\Delta x_{t1}, \dots, \Delta x_{tk}]
+	\end{align*}
+	hence to ensure that TS3' holds the following condition should hold:
+	\begin{align*}
+	\expect[u_{t-1}|x_{t,1} - x_{t-1, 1}, \dots, x_{t, k} - x_{t-1, k}] = 0
+	\end{align*}
+	\item If we assume that 
+	\begin{align*}
+	\expect[y_t|\vec{x}_t, \vec{x}_{t-1}, \dots] = \vec{x}_t^T\vec{\beta}
+	\end{align*}
+	then
+	\begin{align*}
+	\expect[u_t|\vec{x}_t, \vec{x}_{t-1}, \dots] = 0
+	\end{align*}
+	that is, if we assume that only contemporaneously dated (with $y_t$) variables "explain" $y_t$ (i.e.
+the average of $y_t$ can be predicted based only on contemporaneously dated variables) then the model is sequentially exogenous.
+This is not a plausible assumption for economic applications.
 \end{enumerate}

### 2019-03-15 metrics2_2.tex
metrics

@@
 \end{enumerate}
+	\section*{Problem 4}
+	\begin{enumerate}[(i)]
+		\item The result of regression $\Delta inf$ on $\Delta unem$ is shown in the table below
+
+		\begin{center}
+		\begin{tabular}{lc} \hline
+			& (1) \\
+			VARIABLES & cinf \\ \hline
+			&  \\
+			cunem & -0.842** \\
+			& (0.314) \\
+			Constant & -0.0782 \\
+			& (0.348) \\
+			&  \\
+			Observations & 48 \\
+			R-squared & 0.135 \\ \hline
+			\multicolumn{2}{c}{ Standard errors in parentheses} \\
+			\multicolumn{2}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
+		\end{tabular}
+	\end{center}
+	We have obtained statistically significant estimate of $\beta_1$.
+It has a negative sign, that is there is a short-run trade-off between unemployment and inflation: a 1 percent point raise in unemployment leads to on average about 0.8 fall in inflation.
+	\item $R^2$ of model in (i) is higher than that of model 11.19, hence it fits the data better.
+	\end{enumerate}
 \end{document}

### 2019-03-16 micro4_2.tex
Micro

@@
 	\end{align*}
-	Since $c_2(x) > c_1(x) \forall x > 0$ it follows that $x_2^* = 0$.
+	Since $c_2(x) > c_1(x)\ \forall\ x > 0$ ($g(x) = c_2(x) - c_1(x)$, $g(0) = 0, g'(x) = c'_2(x) - c_1'(x) > 0$ hence $g(x)$ is strictly increasing, since $g(0) = 0$ then $g(x)>0\ \forall\ x > 0$) it follows that $x_2^* = 0$.
+That means that monopoly should offer the following contracts:
+	\begin{align*}
+	&(0, 0)\ \text{ which is supposed to be for low type}\\
+	&(w_1 = c_1(c_1'^{-1}(1)), x_1 = c_1'^{-1}(1))\ \text{ which is supposed to be for high type}
+	\end{align*}
+	\item At the competitive equilibrium firms should earn 0 profit. That means that if firms can observe types they should offer the following contracts to both types of workers:
+	\begin{align*}
+	(x_1, x_1), (x_2, x_2)
+	\end{align*}
+	Since firms are competing for workers, at the equilibrium both firms should offer workers such contracts that would maximize their utility (otherwise, if any of firm offer another contract, the rival firm can deviate, offer maximizing utility contract, and hire all workers).
+Of course if firms offer maximizing utility contracts workers have no incentives to deviate and choose a contract which was supposed to be for another type.
+As a result the firms should propose contracts:
+	\begin{align*}
+	(c_1'^{-1}(1), c_1'^{-1}(1))\\
+	(c_2'^{-1}(1), c_2'^{-1}(1))
+	\end{align*}
+	\item If monopoly can distinguish between workers, it should propose each type a contract which would maximize its profit and leave each type of workers with zero utility.
+That is, the contracts should be:
+	\begin{align*}
+	(w_1 = c_1(c_1'^{-1}(1)), x_1 = c_1'^{-1}(1))\ \text{ high type}\\
+	(w_2 = c_2(c_2'^{-1}(1)), x_2 = c_2'^{-1}(1))\ \text{ low type}
+	\end{align*}
+	If competitive firms can observe types of workers they anyway have to propose such contracts which would maximize workers' utility because otherwise the rival firm hire all workers.
+That means that firms should propose the same contracts as in the case of unobservable types.
+That is, first best is achieved under competition and is not achieved under monopoly.
+	\item \begin{itemize}
+		\item Monopoly, observable types:
+		\begin{align*}
+		\underset{x_1, x_2}{\max}\ \frac{1}{2}\left(x_1 -\frac{x_1^2}{2}\right) + \frac{1}{2} \left(x_2 - x_2^2\right)\\
+		x_1^* = 1,\ x_2^* = \frac{1}{2}\\
+		w_1^* = \frac{1}{2},\ w_2^* = \frac{1}{4}
+		\end{align*}
+		\item Monopoly, unobservable types:
+		\begin{align*}
+		\underset{x_1, x_2}{\max}\ \frac{1}{2}\left(x_1 -\frac{x_1^2}{2}\right) + \frac{1}{2} \left(x_2 - x_2^2\right)\\
+		s.t.\ 	\begin{cases}
+		\frac{x_1^2}{2} - x_1^2 \le 0\\
+		x_2^2 - \frac{x_2^2}{2} \le 0
+		\end{cases}\\
+		x_2^* = 0, x_1^* = 1\\
+		w_2 = 0, w_1 = \frac{1}{2}
+		\end{align*}
+		\item Competition, observable types:
+		\begin{align*}
+		x_1^* = 1, w_1^* = 1\\
+		x_2^* = \frac{1}{2}, w_2^* = \frac{1}{2}
+		\end{align*}
+		\item Competition, unobservable types:
+		\begin{align*}
+		x_1^* = 1, w_1^* = 1\\
+		x_2^* = \frac{1}{2}, w_2^* = \frac{1}{2}
+		\end{align*}
+	\end{itemize}
+\end{enumerate}
+\section*{Insurance}
+Consider the competitive screening model of the insurance market with two types of
+consumers $H$ and $L$. All consumers have initial wealth $W > 0$ and may face a loss $d > 0$,
+$d < W$. The probability of facing the loss is $p_L$ for the $L$ type of the consumer, and
+$p_H$ for the $H$ type, where $p_H > p_L$. Multiple competitive insurance firms offer insurance
+contracts against the loss to either type of the consumer. The probabilities of loss and the
+utility functions for both types of consumers are common knowledge to all participants.
+For parts (1) and (2), assume that all consumers have the same strictly concave
+(Bernoulli) utility of wealth $u(W)$ with $u'(W) > 0$ and $u''(W) < 0$. That is, the two types
+of consumers are distinguished only by the difference between $p_L$ and $p_H$
+\begin{enumerate}
+	\item  Assume that the types of the consumers are publicly observable. Draw a graph
+	representing the equilibrium outcome of this model. Explain. Derive the utility levels offered to either type of the consumer in this equilibrium.
+	\item Now assume that the types of the consumers are only known to the consumers themselves.
+Draw a graph to represent the separating equilibrium of this model.
+Explain the reasoning behind your graph.
+Compare the outcome to (1) and comment on
+	the source of difference/similarity.
+	\item Now suppose the $H$ types are more risk averse - with utility function $v(W)$ instead
+	of $u(W)$. As above, probabilities of loss and the utility functions for both types of
+	consumers are common knowledge, but the types of the consumers are only known
+	to the consumers themselves. Draw a new graph to show the effect of this change in
+	utility function on the contracts in the separating equilibrium. Suggest an intuitive
+	explanation for why this change from $u(W)$ to $v(W)$ increases or reduces the utility
+	for good types in the separating equilibrium.
+\end{enumerate}
+
+\textbf{Solution}
+
+\begin{enumerate}
+	\item 
 \end{enumerate}

### 2019-03-16 micro4_1.tex
Micro

@@
 Assume $\omega = p: pD^H + (1-p)D^L < 0$.
-For such a belief the optimal strategy for the entrant is enter once the incumbent's profit if $M^H$ and not enter otherwise (once it is $M^L$).
+For such a belief the optimal strategy for the entrant is enter once the incumbent's profit is $M^H$ and not enter otherwise (once it is $M^L$).
 The incumbent's strategy hence is to choose $M^L$ in both cases because
@@
 	\end{align*}
-	that means that Nich chooses $pass$ once Paul gets beer.
+	that means that Nick chooses $pass$ once Paul gets beer.
 If $\lambda > \frac{1}{3}$ then $-2\lambda + 1 - \lambda < 0$ which means that Nick chooses to pass even if Paul gets quiche.
@@
 	So, when the productivity of the low type is very low compare to that of high type, costs of education are very high for low type workers and it is not optimal for them to pretend to be a high type, which means that low type workers choose $e^*_L = \theta_L$ and high type workers choose $e^*_H = \theta_H$ and firms perfectly distinguish between types.
+On the figure below two examples are provided where the PBE which replicates the full information outcome is possible and where not.
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.8\textwidth]{plotdraft}
+		\caption{Left: low type has profitable deviation; Right: low type does not have a profitable deviation}\label{fig7}
+	\end{figure}
 	\item 
@@
 	w(e) = \begin{cases}
-	\lambda \theta_L + (1 - \lambda) \theta_H, e = 0\\
+	\lambda \theta_H + (1 - \lambda) \theta_L, e = 0\\
 	\theta_L + e, e > 0
@@
 	\begin{align*}
-	\max\left\{ \underset{e > 0}{\max}\ \theta_L + e - \frac{e^2}{2\theta_H}, \lambda \theta_L + (1 - \lambda) \theta_H \right\} = \max\left\{\theta_L + \frac{\theta_H}{2}, \lambda \theta_L + (1 - \lambda) \theta_H \right\}
+	\max\left\{ \underset{e > 0}{\max}\ \theta_L + e - \frac{e^2}{2\theta_H}, \lambda \theta_H + (1 - \lambda) \theta_L \right\} = \max\left\{\theta_L + \frac{\theta_H}{2}, \lambda \theta_H + (1 - \lambda) \theta_L \right\}
 	\end{align*}
-	If $\lambda \le \frac{1}{2}$ then $\theta_L + \frac{\theta_H}{2} > \lambda \theta_L + (1 - \lambda)\theta_H$ and that means that high type will deviate and choose positive $e^*_H = \theta_H$ level of education, i.e.
+	If $0 \le \lambda \le \frac{1}{2}$ then $\theta_L + \frac{\theta_H}{2} > \lambda \theta_H + (1 - \lambda)\theta_L$ and that means that high type will deviate and choose positive $e^*_H = \theta_H$ level of education, i.e.
 $e_p = 0$ is not supportable.
-	\item No, they cannot be.
-Firstly, obviously that the greatest amount of efforts the workers can spend on education at any pooling equilibria cannot exceed $\bar{e}$ (see Fig.
-\ref{fig5}) because for any other education level $e > \bar{e}$ and respective equilibrium wage level $w(e) = \lambda \theta_H + (1 - \lambda)\theta_L + e$ the low type worker prefers $e = 0$.
-(here $\bar{e} = \theta_L\left(1 + \sqrt{1 + \frac{2\lambda(\theta_H - \theta_L)}{\theta_L}}\right) > \theta_L$)
+Figure below illustrates the case.
 	\begin{figure}[H]
@@
 		\includegraphics[width=0.8\textwidth]{plotdraft}
-		\caption{}\label{fig5}
+		\caption{Unsupportable $e_p = 0$}\label{fig8}
 	\end{figure}
-Suppose now that the pooling equilibrium is attained at education level $e' < \bar{e}$ (see Fig.
-\ref{fig5}).
-The corresponding indifference curves and wage schedule is depicted on the Fig.
-\ref{fig5}.
-Both types are maximizing their utilities, given the wage schedule and equilibrium wages are consistent with beliefs.
-And this pooling equilibrium Pareto dominates the one with $e = \bar{e}$ (and also each equilibrium with $e' < e < \bar{e}$) because firms have the same zero expected profit, but both types attain higher indifference curves and consequently higher utility levels at $e'$.
-However, there always exists an education level $\underline{e} = \theta_H$ (which is wlog $< \bar{e}$) such that indifference curve of high type is tangent to the line $\lambda \theta_H + (1 - \lambda)\theta_L + e$ at this point (see Fig.
-\ref{fig5}).
-And each pooling  equilibrium level of education $e'' < \underline{e}$ will yield the high type worker strictly less utility.
-Since each point at line $\lambda \theta_H + (1 - \lambda) \theta_L + e$ for $e < \underline{e}$ corresponds to lower indifference curve for high type.
-Similarly for low type workers there exists an education level $\underline{\underline{e}} = \theta_L$ such that indifference curve of low type is tangent to the line $\lambda \theta_H + (1 - \lambda)\theta_L + e$ at this point (see Fig.
-\ref{fig5}).
-And each pooling  equilibrium level of education $e''' < \underline{\underline{e}}$ will yield the low type worker strictly less utility.
-Since each point at line $\lambda \theta_H + (1 - \lambda) \theta_L + e$ for $e < \underline{\underline{e}}$ corresponds to lower indifference curve for low type.
-As a result, there is a full Pareto order for pooling equilibrium for $e \in [\theta_H, \bar{e}]$ and for $e \in [0, \theta_L]$, for $e \in (\theta_L, \theta_H)$ there is no Pareto order, since some pooling equilibria are better for low type, but are worse for high type and vice-versa.
-
-Case $\theta_H > \bar{e}$ is analogous, for $e \in (\theta_L, \theta_H)$ equilibria cannot be Pareto ranked.
+	\item No, they cannot be.
+The levels of education which maximize workers' utilities provided that firms pay a wage $\lambda \theta_H + (1 - \lambda) \theta_l + e$ are $e_H = \theta_H, e_L = \theta_L$.
+Since $\theta_H > \theta_L$ all pooling equilibria with education level above $\theta_H$ are Pareto ordered because decreasing of $e$ benefits both types of workers and firm still earn zero profit.
+However, equilibria with education level between $\theta_L$ and $\theta_H$ are not Pareto ordered because once $e^*_p \to \theta_L$ the high type becomes worse-off and the low type becomes better-off.
 \item Dominance requirement requires firms to prescribe positive $\mu(e)$ only to such $e$ that is not a strictly dominated action for high type.
 In the sense that there is no action $e'$ which yields the high type strictly greater payoff regardless of what equilibrium wages are.
-Assumption that $\Omega$ in non-empty means that firstly $\theta_H < \bar{e}$ and on the top of that the indifference curve of high type, which goes through $(\bar{e}, \theta_H + \bar{e})$ lies below the indifference curve which is tangent to the line $\lambda \theta_H + (1 - \lambda) \theta_L$ at the point $\theta_H$ (see Fig.
+Assumption that $\Omega$ in non-empty means that firstly $\theta_H < \bar{e}$ (where $\bar{e}$ is the highest possible amount of education the low type worker obtains in any pooling equilibria) and on the top of that the indifference curve of high type, which goes through $(\bar{e}, \theta_H + \bar{e})$ lies below the indifference curve which is tangent to the line $\lambda \theta_H + (1 - \lambda) \theta_L$ at the point $\theta_H$ (see Fig.
 \ref{fig6}) because otherwise, the high type chooses the level $\max\ \{\theta_H, \bar{e}\}$ and low type chooses $0$ and $\Omega$, the set of pooling equilibria, is empty.
@@
 	\end{align*}
-	\item 
 	\item In separating equilibrium Tim can perfectly distinguish whether $\theta$ is high or low, hence in case of high $\theta$ he will work only if:
@@
 Hence at each equilibrium $b^*_H = b^*_L$.
+	\item Single crossing condition means that once cost of work increases the low type Tim needs higher bonus compensation to preserve the same utility level than the high type Tim.
+Expected utility levels of both types are the following:
+	\begin{align*}
+	u_H = \begin{cases}
+	(U + b) \theta_H - c, \text{ if Tim works}\\
+	0, \text{if Tim slacks}
+	\end{cases}\\
+	u_L = \begin{cases}
+	(U + b) \theta_L - c, \text{ if Tim works}\\
+	0, \text{if Tim slacks}
+	\end{cases}\\
+	\end{align*}
+	Since $\frac{1}{\theta_H} < \frac{1}{\theta_L}$ then for high levels of utility single crossing condition holds.
+However, for example for utility level $0$ and for $c < U \theta_L$ (see Fig.
+\ref{fig9})as we can see utility curves coincide, i.e.
+the single crossing condition does not hold.
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.8\textwidth]{plotdraft}
+		\caption{}\label{fig9}
+	\end{figure}
 	\item  As always at separating equilibrium Tim can perfectly distinguish between states. That means that at low state he will work only if
@@
 In this case either firms become unable to distinguish between the types and offer a single wage for all types (pooling equilibrium) and as a result, inequality decreases to 0, or workers of high type just increase their level of education (starting to pursue PhD for example) and firm still are able to distinguish between types and offer different wages, hence the wage inequality remains unchanged.
+\section*{7 Equilibrium Refinements}
+	There are many different criteria that can be used to refine the set of equilibria in a given
+	game. One of them, which you have already discussed, is the intuitive criterion. Think about
+	the logic it tries to capture. Can you think of other ways to select reasonable out-of-equilibrium path beliefs? Suggest one such criterion and explain intuitively why it is reasonable. You may
+	refer to Beer and Quiche or Education signalling or your own set up to illustrate the logic behind
+	your criterion.
 
+	\textbf{Solution}
+
+	Intuitive criterion and other criteria try to refine unreasonable out-off equilibrium beliefs.
+For example, in $Beer-and-Quiche$ set-up (Problem 2).
+Intuitive criterion requires $\lambda$ (the conditional probability of coming across surly type, condition on the fact that Paul has ordered quiche) being equal to 1, because it is unreasonable for wimpish Paul to deviate from equilibrium.
+It is a good logic, but as for me it does not fully capture reality.
+In reality there is always a some probability of mistake, moreover Nick cannot for sure know the cost of drinking beer for wimp Paul.
+That's why I think it would be more reasonable to consider beliefs in unreachable information set as a random variable distributed on $[0, 1]$.
+It can be beta $B(a, b)$ distribution.
+Where parameters $a$ and $b$ are chosen taking into account possible payoffs of players.
+For instance $c$ is very close to 2 then Paul can not care a lot about his order and the probability of mistake becomes higher.
+However even if $c$ is small Nick cannot rule out the chance that apart from being wimpish Paul is also stupid and orders $quiche$.
+Thus, the distribution of belief $\lambda \sim B(a, b)$ where parameters are chosen in a special way (if this problem costs more than 5 point maybe I will derive some possible formulas for $a$ and $b$) capture the reality more precisely, in my opinion.
 \end{document}

### 2019-03-19 micro4_2.tex
Micro

@@
 \end{enumerate}
-\section*{Insurance}
+\section*{3 Insurance}
 Consider the competitive screening model of the insurance market with two types of
@@
 \begin{enumerate}
-	\item 
+	\item Both consumers can happen to be in two states of the world:
+	\begin{align*}
+	W_g &= W - t\\
+	W_b &= W - d + r - t
+	\end{align*}
+	And they derive the following expected utility function:
+	\begin{align*}
+	V(W_b, W_g) = p_i U(W_b) + (1 - p_i)U(W_g),\ i \in \{L, H\}
+	\end{align*}
+	Insurance companies' profit is equal to:
+	\begin{align*}
+	\pi = t - pr
+	\end{align*}
+	and should be equal to zero. That means that
+	\begin{align*}
+	W - W_g - p(W_b + d - W_g) = 0\\
+	(1 - p)W_g + pW_b = W - pd
+	\end{align*}
+	Hence break-even lines for high and low types are given as:
+	\begin{align*}
+	W_b = -\frac{1-p_H}{p_H}W_g + \frac{W}{p_H} -d\\
+	W_b = -\frac{1-p_L}{p_L}W_g + \frac{W}{p_L} -d\\	
+	\end{align*}
+	At the same time, slopes of indifferent curves of consumers are given by:
+	\begin{align*}
+	\frac{dW_b}{dW_g} = - \frac{\frac{dV}{dW_g}}{\frac{dV}{dW_b}} = -\frac{1-p_i}{p_i}\frac{U'(W_g)}{U'(W_b)}
+	\end{align*}
+	In complete information case insurance companies can observe the type, hence equilibrium $(W_g^H, W_b^H), (W_g^L, W_b^L)$ should lie on the break-even lines. In other words,
+	\begin{align*}
+	\frac{dW_b}{dW_g} = - \frac{1-p_i}{p_i}\frac{U'(W_g)}{U'(W_b)} = -\frac{1 - p_i}{p_i}
+	\end{align*}
+	hence by properties of the utility function $W_g = W_b$ for both types. Thus, both types get full insurance in complete information case. The figure below illustrates the case.
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.8\textwidth]{plotdraft}
+		\caption{}\label{fig1}
+	\end{figure}
+Since $W_g^H = W_b^H, W_g^L = W_b^L$ that means that both types at the equilibrium derive the following utility levels:
+\begin{align*}
+p_iU(W_b^i) + (1 - p_i)U(W_b^i) = U(W_b^i) = U(W - p_id),\ i \in \{H, L\}
+\end{align*}
+\item In case of incomplete information at the equilibrium among other things the incentive compatibility constraints should hold. That means that 
+\begin{align*}
+p_L U(W - d + r_L - t_L) + (1 - p_L)U(W - t_L) \ge p_L U(W - d + r_H - t_H) + \\
++(1 - p_L)U(W - t_H)\\
+p_H U(W - d + r_H - t_H) + (1 - p_H)U(W - t_H) \ge p_H U(W - d + r_L - t_L) + \\
++ (1 - p_H)U(W - t_L)\\
+\end{align*}
+If we assume that as in complete information case $W_g^H = W_b^H, W_g^L = W_b^L$ then IC imply:
+\begin{align*}
+U(W^L_g) \ge U(W^H_g)\\
+U(W^H_g) \ge U(W^L_g)\\
+\end{align*}
+but by properties of utility functions this implies that $W^L_g = W^H_g$ which is not a separating equilibrium.
+Thus, at any separating equilibrium the high type should be offered his complete information contract $(W_g^H, W_g^H)$ while the incentive compatibility constraints imply that:
+\begin{align*}
+p_HU(W_b^H) + (1 - p_H)U(W_g^H) = p_H U(W_b^L) + (1 - p_H)U(W_g^L)\\
+U(W - p_Hd) = p_HU(W_b^L) + (1 -p_H)U(W_g^L)
+\end{align*}
+This separating equilibrium is illustrated below
+	\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{plotdraft}
+	\caption{}\label{fig2}
+\end{figure}
+The high risk type obtains the same utility level as in complete information case, whereas the low risk type becomes strictly worse-off, since $W^g_L \neq W_b^L$.
+The source of this is the existence of high risk individual who wants to mimic to low risk type and make insurance companies to lose money.
+\item If high type is more risk averse, then for a given wealth level in good state he will demand more wealth in bad state comparing with utility $u(W)$ to maintain the same utility level.
+That means that new separating equilibrium looks like as follows:
+	\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{plotdraft}
+	\caption{}\label{fig3}
+\end{figure}
+As we can see, in new equilibrium low type obtain higher level of utility, however still less than in the complete information case.
+Intuitively it happens because once the high risk type becomes more risk averse, the amount of riskier (with not complete insurance) contracts which attract this type decreases.
+As a result, now the low type becomes happier, obtaining the contract which is closer to complete insurance.
 \end{enumerate}

### 2019-03-20 micro4_2.tex
Micro

@@
 \end{center}
-\section*{1\ The Doctor and the Patient}
+\section*{1 The Doctor and the Patient}
 Consider the cheap talk signalling model a-la Crawford and Sobel (1982). Suppose that
@@
 \end{enumerate}
+
+\section*{4 Productive Tasks}
+	Extend the job market screening model that we saw in Lecture 8 to the case in which
+	tasks are productive. Assume that a type $\theta$ worker produces $\theta(1 + \mu t)$ units of output
+	when her task level is $t$ where $\mu > 0$.
+	Identify the SPNE of this game.
+
+	\textbf{Solution}
+	SPNE is depicted below.
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.8\textwidth]{plotdraft}
+		\caption{}\label{fig4}
+	\end{figure}
+	As in usual screening model pooling equilibria are impossible.
+And separating equilibrium is on the picture.
+It is indeed equilibria because in this case it cannot be beaten a separating equilibrium as in unproductive task model, because in this equilibrium high type obtains higher level of utility than in any pooling equilibrium.
+And moreover since indifference curve of high type is tangent to the line $w = \theta_H(1 + \mu t)$ firms cannot simultaneously offer a slightly worse (with negative profit) contract for the low type and slightly better (with positive profit) for low type, because the high type will become strictly worse-off.
+
+	\section*{The Patient}
+		Suppose that the hypochondriac patient of question 1 is a professional runner. The
+		patient has been training for a very important race for the last two years and he is very
+		much looking forward to it. In order to run in this important race, first he should be
+		cleared that he is "fit for running" by the Organization of the race. Suppose that when
+		the patient meets his doctor, there is also a member of the Organization in the surgery.
+		What do you think will happen to the quality of the information transmitted by the
+		patient?
+
+		\textbf{Solution}
+
+		Since the patient is very much looking forward to taking part in the race, he does not want to be declared unfit for running.
+That means parameter $b$ becomes smaller.
+Because if patient complains about his health a lot, the member of the Organization will likely to declare him as unfit for running.
+And the sportsman understands it.
+Thus, $b$ becomes smaller and as we can see in Problem 1, expected profits increases, i.e.
+the quality of information transmitted goes up.
+
+		\section*{6 Absence of Equilibria}
+			In the job market screening model that we saw in class, we found that a pooling contract
+			was not an equilibrium because a firm could make money by skimming off the good types.
+			We also saw that a separating equilibrium is beaten by a pooling contract under some
+			conditions. Do you view this result as troubling? How might we be able to alter the
+			model to get a more reasonable prediction?
+
+			\textbf{Solution}
+
+			We should use some refinements i.e.
+slightly alter the concept of equilibrium.
+We can use anticipatory (Wilson, JET, 1977 ) and reactive (Riley, Econometrica, 1979 ) equilbrium concepts.
+The anticipatory equilibrium concept manages to deal with absence of pooling equilibria.
+Once the previous contracts are withdrawn under this concept, firm skim off not only good type but also bad type, but in this situation firms begin losing money, so the deviation is no longer profitable.
+That is, once contracts are withdrawn firms cannot skim off only a good type.
+
+			The concept of reactive equilibrium manages to deal with the second problem.
+This concept allows firms to add more contracts.
+Then, if we are in the case when separating equilibrium is beaten by pooling, adding another contract (see Fig.
+\eqref{fig5}, point 2) makes the initial deviation no longer profitable.
+Hence the separating equilibrium is not beaten by pooling under this concept.
+				\begin{figure}[H]
+				\centering
+				\includegraphics[width=0.8\textwidth]{plotdraft}
+				\caption{}\label{fig5}
+			\end{figure}
 \end{document}

### 2019-03-20 micro4_2.tex
Micro

@@
 			\end{figure}
+
+		\section*{7 Single-crossing}
+			Most screening models assume single-crossing condition. Think about what this condition
+			would require in real life. Do you think single-crossing is always a reasonable assumption?
+			Explain why it is reasonable in any application, or suggest a real-life example where this assumption might fail and explain why.
+
+			\textbf{Solution}
+
+			This condition requires that whatever effort people have already put, for putting the additional amount of effort the low type always requires more compensation than the high type.
+I think that it is a reasonable assumption in real-life terms but only starting from some point.
+Some tasks can require extremely low amount of effort, as a result, both groups of people will demand the same amount of compensation for completing these tasks.
+But starting from some point, tasks become more complicated and require more time and cognitive abilities, in this case people of high type will spend less time, put less effort to solve such a kind of tasks.
+As a result, it will be cheaper for the company.
+i.e.
+high type will require less compensation than the low type, single-crossing condition is satisfied.
 \end{document}

### 2019-03-20 metrics2_3.tex
Metrics

--- 2019-02-05 metrics4.tex
@@
 \begin{center}
-	Econometrics 1. Problem Set 4.
+	Econometrics 2. Problem Set 3.
 \end{center}
 \section*{Problem 1}
+In Example 12.8, we found evidence of heteroskedasticity in $u_t$ in equation (12.47).
+Thus, we compute the heteroskedasticity-robust standard errors along with the usual standard errors.
+What does using the heteroskedasticity-robust $t$ statistic do to the significance of $return_{t-1}$?
 
@@
 
-Since $\hat{\beta_0}, \dots, \hat{\beta_k}$ are OLS estimates they satisfy the following system of equations:
-\begin{align}\label{eq1}
-\begin{cases}
-\summa (y_i - \hat{\beta_0} - \hat{\beta_1}x_{1i} - \dots - \hat{\beta_k}x_{ki}) = 0\\
-\summa x_{1i}(y_i - \hat{\beta_0} - \hat{\beta_1}x_{1i} - \dots - \hat{\beta_k}x_{ki}) = 0\\
+\begin{align*}
+t_{old} = \frac{0.059}{0.038} = 1.55\\
+t_{new} = \frac{0.059}{0.069} = 0.85
+\end{align*}
+Significance of $return_{t-1}$ becomes smaller because for non-robust case p-value is $0.06$ while in robust case it is $0.19$ i.e.
+for example at the significant level of 10\% non-robust t-test reject hypothesis that $return_{t-1}$ does not affect $return_t$ while robust test does not.
+
+\section*{Problem 2}
+\begin{align*}
+\hat{\nu} = \sum_{t=1}^n \hat{a}_t^2 + 2 \sum_{h=1}^g \frac{g+1-h}{g+1} \left(\sum_{t = h+1}^n \hat{a}_{t}\hat{a}_{t-h}\right)
+\end{align*}
+To prove that $\hat{\nu} \ge 0$ firstly let us fixed some $0<g<n$. Denote 
+\begin{align*}
+\mu_h = \sum_{t = h+1}^n\hat{a}_t \hat{a}_{t-h}
+\end{align*}
+This implies
+\begin{align*}
+\hat{\nu} = \mu_0 + 2\sum_{h=1}^g \frac{g+1-h}{g+1}\mu_h
+\end{align*}
+Then let us assume that $a_t \equiv 0,\ \forall\ t < 1, t > n$. Denote
+\begin{align*}
+x_i &= \begin{pmatrix}
+a_{i+1}\\
+a_{i+2}\\
 \dots\\
 \dots\\
-\summa x_{ki}(y_i - \hat{\beta_0} - \hat{\beta_1}x_{1i} - \dots - \hat{\beta_k}x_{ki}) = 0
-\end{cases}
-\end{align}
-
-This is also true for $\tilde{\beta_0}, \dots, \tilde{\beta_k}$, i.e.
+a_{i+n}
+\end{pmatrix}\\
+\sum_{i = -n + 1}^{n-1} x_i x_i^T &= \begin{pmatrix}
+\mu_0 & \mu_1 & \mu_2 & \dots & \mu_{n-1}\\
+\mu_1 & \mu_0 & \mu_1 & \dots & \mu_{n-2}\\
+\dots & \dots & \dots & \dots & \dots \\
+\dots & \dots & \dots & \dots & \dots \\
+\mu_{n-1} & \mu_{n-2} & \mu_{n-3} & \dots & \mu_0
+\end{pmatrix} = M
+\end{align*} 
 \begin{align*}
-\begin{cases}
-\summa c_0(c_0y_i - \tilde{\beta_0} - \tilde{\beta_1}c_1x_{1i} - \dots - \tilde{\beta_k}c_kx_{ki}) = 0\\
-\summa c_1x_{1i}(c_0y_i - \tilde{\beta_0} - \tilde{\beta_1}c_1x_{1i} - \dots - \tilde{\beta_k}c_kx_{ki}) = 0\\
-\dots\\
-\dots\\
-\summa x_{ki}c_k(c_0y_i - \tilde{\beta_0} - \tilde{\beta_1}c_1x_{1i} - \dots - \tilde{\beta_k}c_kx_{ki}) = 0
-\end{cases}
+\forall\ v \in \mathbb{R}^{n}\ v^TMv = \sum_{i=-n+1}^{n-1} v^Tx_ix_i^Tv = \sum_{i=-n+1}^{n-1} (x_i^Tv)^T(x_i^Tv) \ge 0
 \end{align*}
-Assuming that $c_0$ is not equal to 0 either, this system can be rewritten as follows:
+That is matrix, $M$ is positive semi-definite. Denote vector $s \in \mathbb{R}^n$ such that first $g+1$ entries are 1 and other are 0. Then
 \begin{align*}
-\begin{cases}
-\summa c_0^2(y_i - \frac{1}{c_0}\tilde{\beta_0} - \frac{c_1}{c_0}\tilde{\beta_1}x_{1i} - \dots - \frac{c_k}{c_0}\tilde{\beta_k}x_{ki}) = 0\\
-\summa \frac{c_1}{c_0}x_{1i}(y_i - \frac{1}{c_0}\tilde{\beta_0} - \frac{c_1}{c_0}\tilde{\beta_1}x_{1i} - \dots - \frac{c_k}{c_0}\tilde{\beta_k}x_{ki}) = 0\\
-\dots\\
-\dots\\
-\summa \frac{c_k}{c_0}x_{ki}(y_i - \frac{1}{c_0}\tilde{\beta_0} - \frac{c_1}{c_0}\tilde{\beta_1}x_{1i} - \dots - \frac{c_k}{c_0}\tilde{\beta_k}x_{ki}) = 0
-\end{cases}
-\end{align*}
-or, finally
-\begin{align*}
-\begin{cases}
-\summa (y_i - \frac{1}{c_0}\tilde{\beta_0} - \frac{c_1}{c_0}\tilde{\beta_1}x_{1i} - \dots - \frac{c_k}{c_0}\tilde{\beta_k}x_{ki}) = 0\\
-\summa x_{1i}(y_i - \frac{1}{c_0}\tilde{\beta_0} - \frac{c_1}{c_0}\tilde{\beta_1}x_{1i} - \dots - \frac{c_k}{c_0}\tilde{\beta_k}x_{ki}) = 0\\
-\dots\\
-\dots\\
-\summa x_{ki}(y_i - \frac{1}{c_0}\tilde{\beta_0} - \frac{c_1}{c_0}\tilde{\beta_1}x_{1i} - \dots - \frac{c_k}{c_0}\tilde{\beta_k}x_{ki}) = 0
-\end{cases}
-\end{align*}
-Substituting 
-\begin{align}\label{eq2}
-\tilde{\beta_i} = \frac{c_0}{c_i}\tilde{\beta_i}', \forall\ i = 1, \dots, k\\
-\tilde{\beta_0} = c_0 \tilde{\beta_0}' \nonumber
-\end{align} 
-we can get the system which is equivalent to the system \eqref{eq1}, as a result has the same solutions. Plugging these solutions $\hat{\beta_i}$ back to \eqref{eq2} we finally get
-\begin{align*}
-\tilde{\beta_i} = \frac{c_0}{c_i}\hat{\beta_i}\ \forall\ i=1, \dots, k\\
-\tilde{\beta_0} = c_0 \hat{\beta_0}
+\frac{1}{g+1}s^TMs = \mu_0 + 2\frac{g}{g+1}\mu_1 + 2\frac{g-1}{g+1}\mu_2 + \dots + 2 \frac{1}{g+1}\mu_g = \hat{\nu} \ge 0
 \end{align*}
 Q.E.D.
-
-\section*{Problem 2}
-\textbf{Solution}
-
-\begin{align*}
-\ln(wage) = \beta_0 + \beta_1 educ + \beta_2 educ \times pareduc + \beta_3 exper + \beta_4 tenure + u
-\end{align*}
-
-\begin{enumerate}[(i)]
-	\item Obviously
-	\begin{align*}
-	\frac{d \ln (wage)}{d\ educ} = \beta_1 + \beta_2 pareduc
-	\end{align*}
-	hence
-	\begin{align*}
-	\frac{\Delta \ln (wage)}{\Delta educ} \approx \beta_1 + \beta_2 pareduc
-	\end{align*}
-	As for me it is difficult to say something about sign of $\beta_0$ apriori.
-If $\beta_2$ is positive then each additional year of parents' education yields greater average percentage change in salary with additional year of education.
-I cannot come up with any strong reasoning for $\beta_2$ be positive or negative.
-	\item If $pareduc = 24$ then
-	\begin{align*}
-	100\% \cdot \frac{\Delta wage}{wage} = 4.7 \Delta educ + 0.078 \cdot 24 \Delta educ = 4.7 \Delta educ + 1.872 \Delta educ
-	\end{align*}
-	If $pareduc = 32$ then
-	\begin{align*}
-	100\% \cdot \frac{\Delta wage}{wage} = 4.7 \Delta educ + 0.078 \cdot 32 \Delta educ = 4.7 \Delta educ + 2.496 \Delta educ
-	\end{align*}
-	i.e. each additional year of parents' education increases by 0.078 percent points the value of percent change of salary in response to additional year of education.
-	\item \begin{align*}
-	t = \frac{0.0016}{0.0012} \approx 1.33
-	\end{align*}
-	p-value is about 0.09, hence we can reject the hypothesis on 5\% significance level. That is, the return to education does not depend on parents' education.
-	\section*{Problem 3}
-
-	\textbf{Solution}
-
-	Since 
-	\begin{align*}
-	F \equiv \frac{(SSR_r - SSR_{ur}) / q}{SSR_{ur}/(n - k - 1)}
-	\end{align*}
-	where $SSR_r$ and $SSR_{ur}$ are sums of squares of residuals of restricted and unrestricted models respectively, $q = 2, k = 8$. 
-	\begin{align*}
-	R^2_{ur} = 1 - \frac{SSR_{ur}}{SST_{ur}}\\
-	R^2_{r} = 1 - \frac{SSR_{r}}{SST_{r}}\\
-	\end{align*}
-	while
-	\begin{align*}
-	SST = \summa (y_i - \bar{y}) = SST_r = SST_{ur}
-	\end{align*}
-	hence
-	\begin{align*}
-	F = \frac{(R^2_{ur} - R^2_r)/q}{(1 - R^2_{ur})/(n-k-1)} = \frac{(0.232 - 0.229)/2}{(1-0.232)/(680 - 8 - 1)} \approx 1.31
-	\end{align*}
-	p-value is about 0.27, hence the hypothesis (that corresponding slope coefficients are 0) cannot be rejected at the 10\% significance level.
-That is, both these terms are jointly insignificant hence I would not include them in the model.
-\end{enumerate}
-\section*{Problem 4}
-
 \end{document}

### 2019-03-21 metrics3_3.tex
Metrics

--- 2019-03-20 metrics2_3.tex
@@
 Q.E.D.
+
+\section*{Problem 3}
+\begin{enumerate}[(i)]
+	\item 12 $\hat{h}_t$ are negative.
+	\item No, there are no negative $\hat{h}_t$
+	\item Estimating WLS one can get
+	\begin{center}
+		\begin{tabular}{lc} \hline
+			& (1) \\
+			VARIABLES & return \\ \hline
+			&  \\
+			return\_1 & 0.0389 \\
+			& (0.0455) \\
+			Constant & 0.155** \\
+			& (0.0780) \\
+			&  \\
+			Observations & 689 \\
+			R-squared & 0.001 \\ \hline
+			\multicolumn{2}{c}{ Robust standard errors in parentheses} \\
+			\multicolumn{2}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
+		\end{tabular}
+	\end{center}
+while usual OLS regression yields:
+\begin{center}
+	\begin{tabular}{lc} \hline
+		& (1) \\
+		VARIABLES & return \\ \hline
+		&  \\
+		return\_1 & 0.0589 \\
+		& (0.0692) \\
+		Constant & 0.180** \\
+		& (0.0853) \\
+		&  \\
+		Observations & 689 \\
+		R-squared & 0.003 \\ \hline
+		\multicolumn{2}{c}{ Robust standard errors in parentheses} \\
+		\multicolumn{2}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
+	\end{tabular}
+\end{center}
+So as we can see, $\hat{\beta}_1$ becomes smaller but the significance remains unchanged. The hypothesis $H_0: \beta_1 = 0$ cannot be rejected even at 10\% significance level.
+\item The results are depicted below
+\begin{center}
+\begin{tabular}{lc} \hline
+	& (1) \\
+	VARIABLES & return \\ \hline
+	&  \\
+	return\_1 & 0.0239 \\
+	& (0.0466) \\
+	Constant & 0.159** \\
+	& (0.0783) \\
+	&  \\
+	Observations & 688 \\
+	R-squared & 0.000 \\ \hline
+	\multicolumn{2}{c}{ Robust standard errors in parentheses} \\
+	\multicolumn{2}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
+\end{tabular}
+\end{center}
+As we can see $\hat{\beta}_1$ becomes even smaller. It does not change our findings from (iii), the hypothesis that $\beta_1 = 0$ still cannot be rejected.
+\end{enumerate}
+
+\section*{Problem 4}
+\begin{enumerate}[(i)]
+	\item Running regression $\hat{u}_{t}$ on $\hat{u}_{t-1}$ one can get $\hat{\rho} \approx 0.282$ with t-statistics 3.25.
+Hence the hypothesis $\rho =0$ is rejected at 5\% significance level, i.e.
+we cannot reject the hypothesis of presence of serial correlation.
+	Obviously time trend, monthly dummies and $wkends$ are strictly exogenous, it is determined variables.
+Also unemployment rate seems to be strictly exogenous because it seems that today's unexpected changes in fatal accident rate cannot affect future unemployment.
+However, law-variables do not seem to be strictly exogenous because the law can be enacted due to extremely high fatal accident rate, hence today's $u_t$ can predict future laws' enactment.
+
+	\item Results for the usual OLS (column 1) and serial correlation- and heteroskedasticity robust (column 2) are depicted below
+	\begin{center}
+	\begin{tabular}{lcc} \hline
+		& (1) & (2) \\
+		VARIABLES & prcfat & prcfat \\ \hline
+		&  &  \\
+		t & -0.00224*** & -0.00224*** \\
+		& (0.000458) & (0.000559) \\
+		\dots \dots \dots\\
+		wkends & 0.000626 & 0.000626 \\
+		& (0.00585) & (0.00516) \\
+		unem & -0.0154** & -0.0154** \\
+		& (0.00604) & (0.00602) \\
+		spdlaw & 0.0671*** & 0.0671** \\
+		& (0.0198) & (0.0267) \\
+		beltlaw & -0.0295 & -0.0295 \\
+		& (0.0237) & (0.0331) \\
+		Constant & 1.030*** & 1.030*** \\
+		& (0.0959) & (0.0938) \\
+		&  &  \\
+		Observations & 108 & 108 \\
+		R-squared & 0.717 &  \\ \hline
+		\multicolumn{3}{c}{ Robust standard errors in parentheses} \\
+		\multicolumn{3}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
+	\end{tabular}
+	\end{center}
+As we can see the significance of coefficient on $spdlaw$ decreases but it remained significant at 5\%. $beltlaw$ remained not significant its t-statistics becomes lower.
+\item The regression output is depicted below
+\begin{center}
+	\begin{tabular}{lc} \hline
+		& (1) \\
+		VARIABLES & prcfat \\ \hline
+		&  \\
+		t & -0.00215*** \\
+		& (0.000548) \\
+		\dots \dots \\
+		wkends & 0.000617 \\
+		& (0.00500) \\
+		unem & -0.0132* \\
+		& (0.00711) \\
+		spdlaw & 0.0641** \\
+		& (0.0268) \\
+		beltlaw & -0.0248 \\
+		& (0.0301) \\
+		Constant & 1.009*** \\
+		& (0.102) \\
+		&  \\
+		Observations & 108 \\
+		R-squared & 0.641 \\ \hline
+		\multicolumn{2}{c}{ Standard errors in parentheses} \\
+		\multicolumn{2}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
+	\end{tabular}
+\end{center}
+There are no changes both in significance of coefficients and their values.
+\end{enumerate}
 \end{document}

### 2019-03-23 micro4_2.tex
Micro

--- 2019-03-20 micro4_2.tex
@@
 he feels). The doctor cares about her reputation, so she wants to prescribe a appropriate
-treatment for the patient's health: $U
-D(a; \theta) = -(a-\theta)^2$. Instead, the patient, that is
+treatment for the patient's health: $U^D(a; \theta) = -(a-\theta)^2$. Instead, the patient, that is
 very hypochondriac, wants to be overtreated: $U^P(a; \theta) = -(a - (\theta + b))^2$, where $b > 0$ is
@@
 	\end{align*}
-	Since outside options of workers are 0, the optimal contracts monopoly proposes to workers should be such that
+	Since outside options of workers are 0, the optimal contracts monopoly proposes to workers in case of complete information should be such that
 	\begin{align*}
@@
 	\end{align*}
-	That means that at any interior solution high type has incentive to deviate and mimic to low type. Hence we should add incentive compatibility constraints. The IC constraints are:
+	That means that at any interior solution high type has incentive to deviate and mimic to low type. Hence we should add incentive compatibility and participation constraints.
 	\begin{align*}
@@
 	w_2^* - c_1(x_2^*) \le w_1^* - c_1(x_1^*)
-	\end{cases}\\
+	\end{cases}
 	\begin{cases}
-	c_1(x_1^*) - c_2(x_1^*) \le 0\\
-	c_2(x_2^*) - c_1(x_2^*) \le 0
+	w_1^* - c_1(x_1^*) \ge 0\\
+	w_2^* - c_2(x_2^*) \ge 0
 	\end{cases}
 	\end{align*}
-	Since $c_2(x) > c_1(x)\ \forall\ x > 0$ ($g(x) = c_2(x) - c_1(x)$, $g(0) = 0, g'(x) = c'_2(x) - c_1'(x) > 0$ hence $g(x)$ is strictly increasing, since $g(0) = 0$ then $g(x)>0\ \forall\ x > 0$) it follows that $x_2^* = 0$.
-That means that monopoly should offer the following contracts:
+	Since $c_2(x) > c_1(x)\ \forall\ x > 0$ ($g(x) = c_2(x) - c_1(x)$, $g(0) = 0, g'(x) = c'_2(x) - c_1'(x) > 0$ hence $g(x)$ is strictly increasing, since $g(0) = 0$ then $g(x)>0\ \forall\ x > 0$) it follows that $w_1^* - c_1(x_1^*) \ge w_2^* - c_1(x_2^*) > w_2^* - c_2(x_2^*) \ge 0$.
+Thus, $w_1^*-c_1(x_1^*) > 0$ is implied by other constraints.
+Moreover, if $w_2^* - c_2(x_2^*) > 0$ then one can slightly decrease both $w_1^*, w_2^*$ such that all constraints hold but the objective function increases.
+Hence $w_2^* = c_2(x_2^*)$.
+On top of that, if $w_1^* - c_1(x_1^*) > w_2^* - c_1(x_2^*)$ then we can slightly decrease $w_1^*$ such that all constraints hold but the objective function increases.
+That means that $w_1^* - c_1(x_1^*) = w_2 - c_1(x_2^*)$.
+Finally, the monopoly optimization problem looks as follows:
 	\begin{align*}
-	&(0, 0)\ \text{ which is supposed to be for low type}\\
-	&(w_1 = c_1(c_1'^{-1}(1)), x_1 = c_1'^{-1}(1))\ \text{ which is supposed to be for high type}
+	\underset{(x_1, w_1), (x_2, w_2)}{\max}\ \lambda(x_1 - w_1) + (1 - \lambda) (x_2 - w_2)\\
+	s.t.\ \begin{cases}
+	w_1 - c_1(x_1) = c_2(x_2) - c_1(x_2)\\
+	w_1 - c_2(x_1) \le 0\\
+	c_2(x_2) = w_2
+	\end{cases}
 	\end{align*}
+	FOCs imply:
+	\begin{align*}
+	\begin{cases}
+	c'_1(x_1^*) = 1\\
+	\lambda c_1'(x_2^*) + 1 - \lambda - c_2'(x_2^*) = 0
+	\end{cases}
+	\end{align*}
+	If functions $c_1(\cdot), c_2(\cdot)$ are given it is easy to find optimal contracts from FOCs.
 	\item At the competitive equilibrium firms should earn 0 profit. That means that if firms can observe types they should offer the following contracts to both types of workers:
@@
 		\begin{align*}
-		\underset{x_1, x_2}{\max}\ \frac{1}{2}\left(x_1 -\frac{x_1^2}{2}\right) + \frac{1}{2} \left(x_2 - x_2^2\right)\\
+		\underset{x_1, w_1, x_2, w_2}{\max}\ \frac{1}{2}\left(x_1 -w_1\right) + \frac{1}{2} \left(x_2 - x_2^2\right)\\
 		s.t.\ 	\begin{cases}
-		\frac{x_1^2}{2} - x_1^2 \le 0\\
-		x_2^2 - \frac{x_2^2}{2} \le 0
+		w_1 - \frac{x_1^2}{2} = \frac{x_2^2}{2}\\
+		w_1 - x_1^2 \le 0\\
+		w_2 = x_2^2
 		\end{cases}\\
-		x_2^* = 0, x_1^* = 1\\
-		w_2 = 0, w_1 = \frac{1}{2}
+		x_2^* = \frac{1}{3}, x_1^* = 1\\
+		w_2 = \frac{1}{9}, w_1 = \frac{5}{9}
 		\end{align*}
@@
 \item If high type is more risk averse, then for a given wealth level in good state he will demand more wealth in bad state comparing with utility $u(W)$ to maintain the same utility level.
+Put it another way, if less risk averse consumer is indifferent between full insurance and some other point, then more risk averse consumer strictly prefers complete insurance.
 That means that new separating equilibrium looks like as follows:
@@
 
-	\section*{The Patient}
+	\section*{5 The Patient}
 		Suppose that the hypochondriac patient of question 1 is a professional runner. The
@@
 		Since the patient is very much looking forward to taking part in the race, he does not want to be declared unfit for running.
-That means parameter $b$ becomes smaller.
+That means parameter $b$ effectively becomes smaller (of course parameter $b$ is the innate trait of the agent and cannot change, but once the member of Organization is in the surgery, the patient becomes to obtain disutility from the overtreating and declaring unfit for the race, i.e.
+his utility becomes $U^P = - (a - (b + \theta - m))^2$ where $m > 0$.
+If we denote $b' = b - m$ then his effective bias $b'$ becomes smaller once the member of Organization is in surgery).
 Because if patient complains about his health a lot, the member of the Organization will likely to declare him as unfit for running.
 And the sportsman understands it.
-Thus, $b$ becomes smaller and as we can see in Problem 1, expected profits increases, i.e.
-the quality of information transmitted goes up.
+As is has been shown in Problem 1, once bias becomes smaller, expected profits increases.
+The same result was shown for a more general set-up by Crawford and Sobel (1982):
+		\begin{quote}
+			There may be a good case for presuming that direct communication is more likely to play an important role, the more closely related are agents' goals.
+		\end{quote}
+	That is, the quality of information transmission is likely to become higher.
 
@@
 
-			We should use some refinements i.e.
-slightly alter the concept of equilibrium.
-We can use anticipatory (Wilson, JET, 1977 ) and reactive (Riley, Econometrica, 1979 ) equilbrium concepts.
-The anticipatory equilibrium concept manages to deal with absence of pooling equilibria.
+			We should use some refinements i.e. slightly alter the concept of equilibrium. We can use anticipatory (or E2) (Wilson, JET, 1977 ) and reactive (Riley, Econometrica, 1979) equilibrium concepts.
+			Wilson (1977) defines the anticipatory equilibrium as follows:
+			\begin{quote}
+				Each policy in the equilibrium set must earn nonnegative profits and there must not be a new set of policies which will earn nonnegative profits for each policy and strictly positive profits for some new policy after the policies in the original set which have become unprofitable are withdrawn
+			\end{quote} 
+		As a result, this concept manages to deal with absence of pooling equilibria.
 Once the previous contracts are withdrawn under this concept, firm skim off not only good type but also bad type, but in this situation firms begin losing money, so the deviation is no longer profitable.
 That is, once contracts are withdrawn firms cannot skim off only a good type.
-
-			The concept of reactive equilibrium manages to deal with the second problem.
-This concept allows firms to add more contracts.
+			Riley (1979) defines the concept of reactive equilibrium as follows:
+			\begin{quote}
+				A set of offers is a reactive equilibrium if, for any additional offer which generates as expected gain to the agent making the offer, there is another which yields a gain to a second agent and losses to the first.
+Moreover, no further addition to the set of offers generates losses to the second agent.
+			\end{quote}
+			This concept manages to deal with the second problem.
+It allows firms to add more contracts.
 Then, if we are in the case when separating equilibrium is beaten by pooling, adding another contract (see Fig.
@@
 			\end{figure}
+
+		\section*{7 Single-crossing}
+			Most screening models assume single-crossing condition. Think about what this condition
+			would require in real life. Do you think single-crossing is always a reasonable assumption?
+			Explain why it is reasonable in any application, or suggest a real-life example where this assumption might fail and explain why.
+
+			\textbf{Solution}
+
+			This condition requires that whatever effort people have already put to signal of their type, for putting the additional amount of effort the low type always requires more compensation than the high type.
+I think that it is a reasonable assumption in real-life terms but only starting from some point.
+Cost of some signals can be very low, as a result, both groups of people will demand the same amount of compensation for completing these tasks.
+For example, almost everybody graduates from school and the vast majority then enter in some universities.
+The costs of this signals are very low, so people of both types of abilities bear the similar costs of signaling.
+But starting from some point (on effort line), tasks become more complicated and require more time and cognitive abilities, in this case people of high type will spend less time, put less effort to solve such a kind of tasks.
+For example just to graduate from school is a very easy task, however graduate from school with some kind of honors (gold medal or smth like that) is not so easy, as a result the majority of people who succeed in it are either of high type or of low type but who put a lot of effort (maybe these efforts transformed them to people of high type) and here the single crossing condition is satisfied.
+I think that true indifference curves should look something like this:
+			 \begin{figure}[H]
+			 	\centering
+			 	\includegraphics[width=0.8\textwidth]{plotdraft}
+			 	\caption{}\label{fig6}
+			 \end{figure}
 \end{document}

### 2019-03-25 micro4_3.tex
micro

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Microeconomics 4. Problem Set 3.
\end{center}
\section*{Adverse selection with indivisible good}
	Consider the model we studied in class but where the agent may produce only zero or
	one unit of the good. Principal's valuation of one unit is $S$. Agent's costs of producing
	one unit are $\theta$ that takes two values: $\underline{\theta}$ with probability $\nu$ and $\overline{\theta}$ with probability $1-\nu$,
	$\Delta \theta = \overline{\theta} - \underline{\theta} > 0$. When an agent with costs $\theta$ produces one unit, the principal gets $S-t$ and the agent gets $t - \theta$, $t$ being the transfer from the principal to the agent. No production entails no cost to the agent and no value to the principal. Production is always efficient:
	$S > \overline{\theta}$. The agent's reservation utility is zero.
	In the beginning of the period, the principal offers a menu $\{(\underline{p}, \underline{t}), (\overline{p}, \overline{t})\}$, where $\underline{p}$ and $\overline{p}$ are the probabilities of trade with the low and high cost type, respectively (note
	that the transfer is made independently of whether the trade occurs or not).
\begin{enumerate}
	\item Find the first best.
	\item Find the second best.
	\item Can the principal replicate the second-best solution by a simpler contract? Which
one?
\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item Assume that once the trade does not occur, the agent does not produce anything. That means that the agent's objective function is $U^A(t, p, \theta) = t - p \theta$. The principal's objective function is $U^P(\underline{p}, \underline{\theta}, \overline{p}, \overline{\theta}) = \nu(\underline{p}S - \underline{t}) + (1 - \nu)(\overline{p}S - \overline{t})$. Obviously under complete information $\underline{t} = \underline{p}\underline{\theta}$ and $\overline{t} = \overline{p}\overline{\theta}$. And $\underline{p} = \overline{p} = 1$. That is, the first best is the following:
	\begin{align*}
	\{(1, \overline{\theta}), (1, \underline{\theta})\}
	\end{align*}
	\item Under incomplete information, since $\overline{\theta} > \underline{\theta}$ then $\overline{t} - \overline{p}\underline{\theta} = \overline{p} \Delta \theta > \underline{t} - \underline{p}\underline{\theta} = 0$. That is, the type $\underline{\theta}$ will pretend to be the type $\overline{\theta}$. Hence the principal should propose the contracts which would be incentive compatible. Thus, the principal is solving the following optimization problem:
	\begin{align*}
	\underset{(\underline{p}, \underline{t}), (\overline{t}, \overline{p})}{\max}\ \nu(\underline{p}S - \underline{t}) + (1 - \nu)(\overline{p}S - \overline{t})\\
	s.t.\ \begin{cases}
	\overline{t} - \overline{p} \overline{\theta} \ge 0\ (PC1)\\
	\underline{t} - \underline{p} \underline{\theta} \ge 0\ (PC2)\\
	\overline{t} - \overline{p} \overline{\theta} \ge \underline{t} - \underline{p} \overline{\theta}\ (IC1)\\
	\underline{t} - \underline{p} \underline{\theta} \ge \overline{t} - \overline{p} \underline{\theta}\ (IC2)
	\end{cases}
	\end{align*}
	$\underline{t} - \underline{p}\underline{\theta} \ge \overline{t} - \overline{p}\underline{\theta} > \overline{t} - \overline{p}\overline{\theta} \ge 0$, hence, $PC2$ is not binding. While $PC1$ is binding because otherwise the principal can decrease $\underline{t}, \overline{t}$ by the same value, which leads to increase of objective function without violating of any constraints. Moreover, $IC2$ should be binding either, because otherwise the principal can decrease $\underline{t}$. Finally, the principal's problem should look like:
	\begin{align*}
	&\underset{\overline{p}, \underline{p}}{\max}\ \nu(\underline{p}S - \overline{p}\Delta \theta - \underline{p}\underline{\theta}) + (1-\nu) \overline{p}(S - \overline{\theta})\\
	&s.t.\ \overline{p} \le \underline{p}\ (IC1)
	\end{align*}
	Since $\nu (S - \underline{\theta}) > 0$ then $\underline{p} = 1$. If $(1 - \nu)(S - \overline{\theta}) > \nu \Delta \theta$ then $\overline{p} = 1$ and $\overline{t} = \overline{\theta}, \underline{t} = \overline{\theta}$. Effectively, it means that the principal proposes just one contract $(1, \overline{\theta})$. If $(1 - \nu)(S - \overline{\theta}) < \nu \Delta \theta$ then $\overline{p} = 0$, and $\overline{t} = 0, \underline{t} = \underline{\theta}$. If $(1 - \nu)(S - \overline{\theta}) = \nu \Delta \theta$ then $\overline{p} \in [0, 1]$, $\overline{t} = \overline{p} \overline{\theta}, \underline{t} = \overline{p}\Delta \theta + \underline{\theta}$. So, the second best is the following:
	\begin{align*}
	\begin{cases}
	(1, \overline{\theta}), (1, \overline{\theta}),\ \frac{S - \overline{\theta}}{S - \underline{\theta}} > \nu\\
	(1, \underline{\theta}), (0, 0),\ \frac{S - \overline{\theta}}{S - \underline{\theta}} < \nu\\
	(1, \overline{p}\Delta \theta + \underline{\theta}), (\overline{p}, \overline{p}\overline{\theta}),\ \frac{S - \overline{\theta}}{S - \underline{\theta}} = \nu, \overline{p} \in (0, 1)
	\end{cases}
	\end{align*}
	\item By the revelation principle, to replicate the second best we need contracts with number of options equal to cardinality of types' space. Here the cardinality is equal to 2, hence we cannot replicate the second best with less number of options.
\end{enumerate}
\end{document}

### 2019-03-26 micro4_3.tex
Micro

@@
 \newcommand{\expect}{\mathbb{E}}
+\newcommand{\ubar}{\overline}
+\newcommand{\lbar}{\underline}
 \newcommand{\summa}{\sum_{i=1}^n}
@@
 \end{enumerate}
+\section*{2 Adverse selection with a privately informed buyer}
+	Here we consider another standard adverse selection setting: a seller with known costs
+	sells a good to a buyer who is privately informed about his valuation.
+	The cost for the principal of $q$ units of good (or quality level $q$) is $c(q)$ and $c'(q) >
+	0, c''(q) > 0, c(0) = 0$. The utility of the principal is $V = t-c(q)$. The agent's valuation
+	for the good is $\theta u(q)$, where $u'(q) > 0, u''(q) < 0, u(0) = 0$ and $\theta = \underline{\theta}$ with probability
+	$\nu$ and $\theta = \overline{\theta}$ with probability $1-\nu$. Denote $\Delta \theta = \overline{\theta}- \underline{\theta}> 0$. The utility of the agent is $U = \theta u(q) - t$.
+	\begin{enumerate}
+	\item Find the first best.
+	\item Check the Spence-Mirlees condition.
+	\item Find the second best.
+	\item When it is optimal to "shut down" the low-valuation type?
+	\end{enumerate}
+
+\textbf{Solution}
+
+\begin{enumerate}
+	\item If the principle knows the agent's valuation then she should solve the following optimization problems:
+	\begin{align*}
+	\underset{(\overline{t}, \overline{q})}{\max}\ \overline{t} - c(\overline{q})\\
+	s.t.\ \overline{\theta} u(\overline{q}) - \overline{t} \ge 0\\
+	\underset{(\underline{t}, \underline{q})}{\max}\ \underline{t} - c(\underline{q})\\
+	s.t.\ \underline{\theta} u(\underline{q}) - \underline{t} \ge 0\\
+	\end{align*}
+	The solutions can be derived from the following FOCs
+	\begin{align*}
+	\underline{\theta} u'(\underline{q}) - c'(\underline{q}) = 0 \to \underline{q}^* \to \underline{t}^* = \underline{\theta}u(\underline{q}^*)\\
+	\overline{\theta} u'(\overline{q}) - c'(\overline{q}) = 0 \to \overline{q}^* \to \overline{t}^* = \overline{\theta}u(\overline{q}^*)
+	\end{align*}
+	\item The Spence-Mirrlees condition
+	\begin{align*}
+	\frac{\partial ^2}{\partial q \partial \theta} \theta u(q) > 0\\
+	\frac{\partial ^2}{\partial q \partial \theta} \theta u(q) = u'(\theta) > 0
+	\end{align*}
+	hence the condition is satisfied.
+	\item If at the first best $\underline{q}^* > 0$ then $\overline{\theta} u(\underline{q}^*) - \underline{\theta}u(\underline{q}^*) > 0$ i.e.
+under asymmetric information the $\ubar{\theta}$ type would mimic the $\lbar{\theta}$ type.
+Hence the second best should be incentive compatible.
+The principal should solve:
+	\begin{align*}
+	\underset{(\underline{t}, \underline{q}), (\overline{t}, \overline{q})}{\max}\ \nu(\underline{t} - c(\underline{q})) + (1 - \nu)(\overline{t} - c(\overline{q}))\\
+	s.t.\ \begin{cases}
+	\overline{\theta} u(\overline{q}) - \overline{t} \ge 0\ (PC1)\\
+	\underline{\theta} u(\underline{q}) - \underline{t} \ge 0\ (PC2)\\
+	\overline{\theta} u(\overline{q}) - \overline{t} \ge \overline{\theta} u(\underline{q}) - \underline{t}\ (IC1)\\
+	\underline{\theta} u(\underline{q}) - \underline{t} \ge \underline{\theta} u(\overline{q}) - \overline{t}\ (IC2)
+	\end{cases}
+	\end{align*}
+	$\ubar{\theta}u(\ubar{q}) - \ubar{t} \ge \ubar{\theta} u(\lbar{q}) - \lbar{t} > \lbar{\theta}u(\lbar{q}) - \lbar{t} > 0$ hence $PC1$ is not binding.
+While $PC2$ is binding because otherwise the principal can increase $\ubar{t}, \lbar{t}$ by the same amount which leads to increase of the objective function without the violation of any constraints.
+Moreover, the $IC1$ is binding because otherwise the principal can slightly increase $\ubar{t}$.
+Hence the optimization problem should like:
+	\begin{align*}
+	&\underset{(\ubar{q}, \lbar{q})}{\max}\ \nu(\lbar{\theta}u(\lbar{q}) - c(\lbar{q})) + (1 - \nu)(\ubar{\theta}u(\ubar{q}) - \Delta \theta u(\lbar{q}) - c(\ubar{q}))\\
+	&s.t.\ u(\lbar{q}) \le u(\ubar{q})
+	\end{align*}
+	FOCs:
+	\begin{align*}
+	\begin{cases}
+	\lbar{\theta}u'(\lbar{q}^*) - c'(\lbar{q}^*) - \frac{1 - \nu}{\nu} \Delta \theta u'(\lbar{q}^*) = 0\\
+	\ubar{\theta}u'(\ubar{q}^*) - c'(\ubar{q}) = 0
+	\end{cases}
+	\end{align*}
+	As we can see for $\ubar{\theta}$ second best $\ubar{q}^*$ coincides with the first best $\ubar{q}^*$.
+And $\ubar{t}^{*SB} = \ubar{\theta} u(\ubar{q}^*) - \Delta \theta u(\lbar{q}^*) < \lbar{t}^{*FB}$.
+However for $\lbar{\theta}$, $\lbar{q}^{*SB} < \lbar{q}^{*FB}$, if $\Delta \theta > 0$ (because $\frac{u'(\lbar{q})}{c'(\lbar{q})} = \frac{1}{\lbar{\theta} - \frac{1-\nu}{\nu}\Delta \theta}$ goes up once $\Delta \theta > 0$ which implies that $\lbar{q}$ declines).
+This also guarantees that $u(\lbar{q}) \le u(\ubar{q})$ because $\lbar{q}^{FB} < \ubar{q}^{FB}$.
+	\item It is optimal to shut down the low-valuation type if:
+	\begin{align*}
+	(1 - \nu) (\ubar{\theta}u(\ubar{q}^{FB}) - c(\ubar{q}^{FB})) > \nu(\lbar{\theta}u(\lbar{q}^{FB}) - c(\lbar{q}^{FB})) + (1 - \nu)(\ubar{\theta}u(\lbar{q}^{FB}) - c(\lbar{q}^{FB}))
+	\end{align*}
+\end{enumerate}
 \end{document}

### 2019-03-27 micro4_3.tex
Micro

@@
 	\begin{align*}
-	\frac{\partial ^2}{\partial q \partial \theta} \theta u(q) > 0\\
+	\frac{\partial ^2}{\partial q \partial \theta} (\theta u(q) - t) \neq 0\\
 	\frac{\partial ^2}{\partial q \partial \theta} \theta u(q) = u'(\theta) > 0
@@
 \end{enumerate}
+\section*{3 Lending with adverse selection}
+	There is a continuum of risk neutral borrowers with no personal wealth and limited
+	liability. A proportion $\nu$ of borrowers (called type 1) have sure projects with return $h$ for
+	an investment of $1$. A proportion $1-\nu$ of borrowers (called type 2) have (stochastically
+	independent) projects with return $h$ only with probability $\theta$ in $(0; 1)$ and return $0$ with
+	probability $1-\theta$, for an investment of $1$. If he does not apply for a loan, the borrower
+	has an outside opportunity utility level of $u$.
+	There is a single risk neutral bank available for loans which has a financing cost of
+	$r$. The bank offers contracts to maximize its expected profit. For simplicity, we assume
+	that all projects are socially valuable, i.e.,
+	\begin{align*}
+	\theta h > r + u
+	\end{align*}
+	\begin{enumerate}
+	\item Explain why there is no loss of generality in considering the menus of contracts
+	$\{(R_1, P_1),(R_2, P_2)\}$ where $P_i$
+	is the probability of obtaining a loan and $R_i$
+	is the
+	repayment to the bank when the investment succeeds if the borrower announces
+	that he is of type $i$.
+	\item Write the maximization program of the bank which chooses the menu $\{(R_1, P_1),(R_2, P_2)\}$
+	to maximize its expected profit under the borrower's participation and incentive
+	constraints (for simplicity assume that if a borrower applies for a loan he loses his
+	outside opportunity $u$).
+	\item Show that the optimal contract entails a non-random allocation of loans (i.e., $P_i$
+	is
+	either $0$ or $1$, $i = 1, 2$). Characterize the optimal contract. Discuss its properties.
+	Draw indifference curves (in the space $(R, P)$) and look at the Spence-Mirrlees
+	condition. To answer this question, you will need the more general definition of the
+	Spence-Mirrlees condition. Let U be the borrower's utility, then the single crossing
+	condition says that $\frac{\partial }{\partial \theta}\left(\frac{\frac{\partial U}{\partial P}}{\frac{\partial U}{\partial R}}\right)$ has a constant sign.
+	\end{enumerate}
+
+\textbf{Solution}
+
+\begin{enumerate}
+	\item There is no loss of generality because of revelation principle.
+Which implies that set of contracts with number of options equal to cardinality of the type space, is sufficient to replicate any arbitrary complicated contract structure.
+	\item The bank is maximizing:
+	\begin{align*}
+	\underset{(R_1, P_1), (R_2, P_2)}{\max}\ \nu P_1(R_1 - r) + (1 - \nu)P_2(\theta R_2 - r)\\
+	s.t.\ \begin{cases}
+	P_1 (h - R_1) \ge u\ (PC_1)\\
+	P_2\theta(h - R_2) \ge u\ (PC_2)\\
+	P_1(h - R_1) \ge P_2(h - R_2)\ (IC_1)\\
+	P_2\theta (h - R_2) \ge P_1 \theta (h - R_1)\ (IC_2)
+	\end{cases}
+	\end{align*}
+	\item As always, $(PC_1)$ is not binding because $P_1(h - R_1) \ge P_2(h - R_2) > P_2 \theta (h - R_2) \ge u$.
+$(PC_2)$ is binding because otherwise bank increases $P_1R_1, P_2R_2$ by the same value, which leads to increase in objective function without violation of any constraints.
+Moreover, $(IC_1)$ and $(IC_2)$ are also binding because $\theta$ is canceled out.
+The bank's problem looks as follows:
+	\begin{align*}
+	\underset{P_1, P_2}{\max}\ \nu P_1(h - \frac{u}{P_1\theta} - r) + (1 - \nu)P_2(\theta h - \frac{u}{P_2} - r)
+	\end{align*} 
+	The solution is:
+	\begin{align*}
+	\begin{cases}
+	P_1^* = 1, R_1^* = h - \frac{u}{\theta}, h > r\\
+	P_1^* = 0, \forall\ R_1, h < r\\
+	P_1^* \in (0, 1), R_1^* = h - \frac{u}{\theta P_1}, h = r
+	\end{cases}
+	\begin{cases}
+	P_2^* = 1, R_2^* = h - \frac{u}{\theta}, \theta h - r > 0\\
+	P_2^* = 0, \forall\ R_2, \theta h - r < 0\\
+	P_2^* \in (0, 1), R_2^* = h - \frac{u}{P_2 \theta}, \theta h - r = 0 
+	\end{cases}
+	\end{align*}
+	As we can see, the optimal contract entails not only non-random allocation, for some values of parameters, $P_1$ or $P_2$ can be $\in (0, 1)$.
+Since I was asked to show that contracts are non-random, I will further assume that parameters are such that contracts are non-random.
+Depending on parameters three outcomes are possible: $P_1^* = P_2^* = 0$, if $h < r$, $P_1^* = P_2^* = 1$ if $\theta h < r < h$, or $P_1^* = 1, P_2^* = 0$ if $r < \theta h$.
+As we can see, if cost of financing are higher than possible returns then obviously bank does not finance anybody, if cost of financing are lower than expected return of risky project than bank finances both groups proposes the same contracts for both groups.
+If cost are higher than expected gains from risky project then bank stops dealing with risky investors.
+Since utility functions of agents are
+	\begin{align*}
+	U_1 &= P_1(h - R_1) - u\\
+	U_2 &= P_2\theta (h - R_2) - u 
+	\end{align*}
+	the indifference curves look as follows.
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.8\textwidth]{plotdraft}
+		\caption{}\label{fig1}
+	\end{figure}
+Single-crossing condition:
+\begin{align*}
+\frac{\partial}{\partial \theta} \left(\frac{R-h}{P}\right) = 0
+\end{align*}
+has constant sign, however indifference curves obviously intersect only in infinity.
+\end{enumerate}	
+\section*{4 The bribing game}
+	We consider an administration which is supposed to deliver with some fixed delay
+	a service to the citizens (passport, permits,...). With the normal functioning of the
+	administration, citizens derive a benefit $u_0$ which depends on their valuation of time.
+	With some additional effort the official can deliver the service with a shorter delay.
+	Let us call $q$ the decrease of delay that the official can provide at a cost $\frac{(q - Q)^2}{2}$
+	for him
+	where $Q$ is a constant.
+	We assume that there is a proportion $\nu$ (resp. $1-\nu$) of type 1 (resp. type 2) citizens
+	who derive a benefit from a decrease $q$ of delay equal to $\lbar{\theta}q$ ($\ubar{\theta}q$). Citizens are willing to
+	bribe the official to decrease delays.
+	Characterize the optimal bribing contract that the official will offer to the citizen
+
+	\textbf{Solution}
+	Assume that $\lbar{\theta} < \ubar{\theta}$. The official is solving:
+	\begin{align*}
+	\underset{(b_1, q_1), (b_2, q_2)}{\max}\ \nu\left(b_1 - \frac{(q_1 - Q)^2}{2}\right) + (1 - \nu) \left(b_2 - \frac{(q_2 - Q)^2}{2}\right)\\
+	s.t.\ \begin{cases}
+	\ubar{\theta}q_1 - b_1 \ge 0\ (PC_1)\\
+	\lbar{\theta}q_2 - b_2 \ge 0\ (PC_2)\\
+	\ubar{\theta}q_1 - b_1 \ge \ubar{\theta}q_2 - b_2\ (IC_1)\\
+	\lbar{\theta} q_2 - b_2 \ge \lbar{\theta}q_1 - b_1\ (IC_2)
+	\end{cases}
+	\end{align*}
+	Which is equivalent to:
+	\begin{align*}
+	&\underset{q_1, q_2}{\max}\ \nu \left(\ubar{\theta}q_1 - \Delta \theta q_2 - \frac{(q_1 - Q)^2}{2}\right) + (1 - \nu) \left(\lbar{\theta}q_2 -\frac{(q_2 - Q)^2}{2}\right) \\
+	&s.t.\ \lbar{\theta}q_1 - b_1 \le 0
+	\end{align*}
+	FOCs:
+	\begin{align*}
+	\begin{cases}
+	q_1^* = \ubar{\theta} + Q\\
+	q_2^* = \lbar{\theta} + Q - \frac{\nu}{1 - \nu} \Delta \theta	
+	\end{cases} \begin{cases}
+	b_1^* = \ubar{\theta}(\ubar{\theta} + Q) - \Delta \theta (\lbar{\theta} +Q - \frac{\nu}{1 - \nu} \Delta \theta)\\
+	b_2^* = \lbar{\theta}(\lbar{\theta} +Q - \frac{\nu}{1 - \nu} \Delta \theta)
+	\end{cases}
+	\end{align*}
+	\begin{align*}
+	b_1 = \ubar{\theta}(\ubar{\theta} + Q) - \Delta \theta (\lbar{\theta} +Q - \frac{\nu}{1 - \nu} \Delta \theta) > \ubar{\theta}(\ubar{\theta} + Q) - \Delta \theta (\ubar{\theta} + Q) = \lbar{\theta} q_1
+	\end{align*}
+
+	\section*{Flexibility}
+		Since flexibility is always good, an employer who does not know the worker's productivity
+		will benefit from adjusting the contract after the worker reveals his productivity. Discus
+
+		\section*{6 At your disservice}
+
+		I donít want to complain, but my favourite lunchtime haunt has the rudest barman I've
+		ever encountered - and I don't think it's an accident. The restaurant is famous for its
+		superb, sophisticated Italian cooking and it prices accordingly. A romantic meal for two
+		costs about \$150, plus the price of your selection from a wine list of biblical proportions.
+		Even if you pick the cheapest main course for lunch and sip water, this frugal meal for
+		one will set you back \$15.
+		Or, you could sit at the bar or on one of the tables in the bar area. The food is still
+		superb: you can fill up on rich, soft pork meatballs nestling on pillows of light polenta for
+		about \$8. The veal ragu is rich but you don't have to be, because this perfect spaghetti
+		is half the price of a pasta dish in the main restaurant.
+		It sounds too good to be true, and I'm afraid there is a catch. To get to the food,
+		you have to get past the barman who takes your orders, a man more Bond villain than
+		bon vivant. To walk into that bar is to laugh in the face of fear. The barman welcomes
+		me in with all the warmth of a Transylvanian butler in a B-movie. He drops menus on
+		my table with a sneer. He repeatedly ignores my attempts to order. As he walks past,
+		my companion whispers that he picked up his tattoos in a Russian prison. I think sheís
+		having me on, but I canít be sure.
+		Itís possible, of course, that this man is in the wrong job by accident, and when
+		the restaurant owner reads this column and works out that itís about his restaurant,
+		"Igor" will be fired. I'm not convinced. I think Igor is all part of the plan. One food
+		critic described a previous bartender as acting like sheíd just returned from a tax audit.
+		"Venomous" is obviously in the job description.
+		Why would a restaurant deliberately sabotage the dining experience?
 \end{document}

### 2019-03-27 micro4_3.tex
Micro

@@
 
-	\section*{Flexibility}
+	\section*{5 Flexibility}
 		Since flexibility is always good, an employer who does not know the worker's productivity
@@
 
+		\textbf{Solution}
+
+		In some sense employer indeed benefits from renegotiation of the labour contract, because it can implement the most efficient contract.
+But in the long run such an opportunity can harm the employer because of the following reason.
+The employer will renegotiate the contract only with the low productivity worker, because the initial contract with this type of workers was deliberately worse than it could be, in order to prevent the high productivity type from pretending to be low.
+But if ex-ante the high productivity worker had known that tomorrow contract with low type is renegotiated, he would not have reveled his type, and still would have chosen to pretend being of low type.
+That is, in short run the employer can benefit, but in the long run without commitment he will lose, because nobody of the high type wants to reveal their type.
+
 		\section*{6 At your disservice}
 
-		I donít want to complain, but my favourite lunchtime haunt has the rudest barman I've
+		I don't want to complain, but my favourite lunchtime haunt has the rudest barman I've
 		ever encountered - and I don't think it's an accident. The restaurant is famous for its
@@
 		my table with a sneer. He repeatedly ignores my attempts to order. As he walks past,
-		my companion whispers that he picked up his tattoos in a Russian prison. I think sheís
-		having me on, but I canít be sure.
-		Itís possible, of course, that this man is in the wrong job by accident, and when
-		the restaurant owner reads this column and works out that itís about his restaurant,
+		my companion whispers that he picked up his tattoos in a Russian prison. I think she's
+		having me on, but I can't be sure.
+		It's possible, of course, that this man is in the wrong job by accident, and when
+		the restaurant owner reads this column and works out that it's about his restaurant,
 		"Igor" will be fired. I'm not convinced. I think Igor is all part of the plan. One food
-		critic described a previous bartender as acting like sheíd just returned from a tax audit.
+		critic described a previous bartender as acting like she'd just returned from a tax audit.
 		"Venomous" is obviously in the job description.
 		Why would a restaurant deliberately sabotage the dining experience?
+
+		\textbf{Solution}
+
+		It is an example of the second degree price discrimination.
+Unless existence of Igor, everybody went to the bar, get the same, superb food, but paying less.
+But the firm wants to separate those who can pay more from those who cannot afford going to the restaurant and reserving tables.
+To do so, the contracts the restaurant proposes should be incentive compatible.
+That is, the rich guys should not have incentives to deviate and to go to the bar and "underpay".
+Hence the owner of the restaurant has to decrease the quality of serving in bar, in order to prevent rich guys from going there.
+
+		\section*{7 Revelation Principle}
+			The revelation principle is a very powerful tool that greatly simplifies the search for optimal
+			mechanisms. Think about the intuitive argument and the logic behind the revelation principle
+			and, importantly, the assumptions this argument implicitly makes. Suggest a real-life situation
+			where the revelation principle would fail. Explain why the logic of the revelation principle
+			cannot be applied to your example.
+
+			\textbf{Solution}
+
+			The principle is based on the idea that even if instead of reporting type 
 \end{document}

### 2019-03-29 micro4_3.tex
Micro

@@
 \end{center}
-\section*{Adverse selection with indivisible good}
+\section*{1 Adverse selection with indivisible good}
 	Consider the model we studied in class but where the agent may produce only zero or
@@
 	\end{align*}
-	\item By the revelation principle, to replicate the second best we need contracts with number of options equal to cardinality of types' space.
-Here the cardinality is equal to 2, hence we cannot replicate the second best with less number of options.
+	\item If $\frac{S - \ubar{\theta}}{S - \lbar{\theta}} \neq \nu$ then yes, the seller can merely either offer the price $\lbar{\theta}$ or $\ubar{\theta}$, depending on the sign of $\frac{S - \ubar{\theta}}{S - \lbar{\theta}} - \nu$.
+Otherwise, also yes, but however seller can pick any probability of trading, finally getting the same expected payoff.
 \end{enumerate}
@@
 That is, in short run the employer can benefit, but in the long run without commitment he will lose, because nobody of the high type wants to reveal their type.
+This results has been demonstrated by Dewatripoint (QJE, 1989):
+		\begin{quote}
+			"Once the parties cannot precommit any more not to make
+			Pareto-improving changes in the contract based on information
+			revealed in previous periods, the outcome will be worse than in
+			Grossman and Hart's model."
+		\end{quote}
+	\begin{quote}
+		"For example, if at time $t$ it becomes common
+		knowledge that the state of nature is $i$, then from $t + 1$ on
+		employment has to be set at the Walrasian level for state $i$. Any
+		other employment level would not be credible, since it would be
+		known to be Pareto dominated and would be changed through
+		renegotiation. In this sense, revealing information can be costly (ex
+		ante), so that full instantaneous revelation may not be optimal."
+	\end{quote}
 
@@
 
-			The principle is based on the idea that even if instead of reporting type 
+			The principle is based on the idea that even if instead of mere reporting his type, the agent reports some function of his type, which would be the best his option given the principal's policy, and then the principal uses some rule $\tilde{g}$ (see Fig.\ref{fig2}) to chose his action.
+It is equivalent to the agent just reporting the type.
+			\begin{figure}[H]
+				\centering
+				\includegraphics[width=0.8\textwidth]{DRP}
+				\caption{}\label{fig2}
+			\end{figure}
+			If one see the picture above the most obvious example when this principle can fail is if functions $m^*(\cdot)$ or $\tilde{g}(\cdot)$ are not one-to-one.
+The real-life situation which can be described by such function is the following.
+Assume that instead of one principal there are two principals (e.g.
+two suppliers) and one agent (buyer) who has a capacity constraint $q_1 + q_2 \le \bar{K}$.
+Where $q_1, q_2$ the amount the agent buys from the first and the second supplier respectively.
+In this case even if the agent can be of two types: of high valuation and of low valuation.
+Each supplier should take into account the mechanism which the other principal uses, because it affects the quantity the agent buys from this principal.
+As a result, the revelation principal in its traditional form fails here, however Peters (Econometrica, 2001) showed that
+			\begin{quote}
+				"In very general
+				common agency problems in which there is only a single agent, principals do not need
+				any more than menus of alternatives to respond to the mechanisms offered by the other
+				principals."
+			\end{quote}
 \end{document}

### 2019-04-01 micro4_4.tex
Micro

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\ubar}{\overline}
\newcommand{\lbar}{\underline}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Microeconomics 4. Problem Set 4.
\end{center}
\section*{1 Credit Markets}
	A risk-neutral entrepreneur has a project that requires capital $K$ to initiate. The project
	will either succeed, in which case it generates profit $R$, or it will fail, in which case
	it generates profit $0$. The probability of success is equal to the entrepreneur's effort
	level $e \in [0, 1]$, and his effort cost is $\frac{e^2}{2}$. He seeks outside funding from an investor. The
	investor receives a repayment $r$ if the project is successful and $0$ otherwise. Therefore, the
	entrepreneur's and the investor's expected profit is $\Pi _E = e(R-r) - \frac{e^2}{2}$ and $\Pi _I = er-K$,respectively. Assume that the entrepreneur makes a take-it-or-leave-it offer to an investor,
	whose outside option is $0$. (This is equivalent to assuming that the investor offers the
	contract and he operates in a perfectly competitive market.) Furthermore, assume that $R^2 > 4K$.
	\begin{enumerate}
	\item Assume that effort level is contractible. Derive the first-best outcome $e^{FB}$ and $r^{FB}$.
	\item  Now suppose that effort is not contractible. Find the optimal solution $e^*$ and $r^*$.
	\item  Assume that the entrepreneur has initial wealth $w \in [0, K]$. Therefore, he will invest
	his own wealth $w$ in the project, and borrow $K-w$ from an investor. Compute the entrepreneur's expected utility $V(w)$. Show that it is increasing and concave in $w$. Provide some intuition and interpretation.
	\end{enumerate}

\textbf{Solution}

Assume that $R > r$.
\begin{enumerate}
	\item If the entrepreneur makes take-it-or-leave offer than he is solving the following optimization problem:
	\begin{align*}
	\underset{e, r}{\max}\ e(R - r) - \frac{e^2}{2}\\
	s.t.\ er - K \ge 0
	\end{align*}
	The solution is obviously
	\begin{align*}
	e^{FB} = \min\{R, 1\},\ r^{FB} = \frac{K}{\min\{R, 1\}}
	\end{align*}
	\item If the effort level is not contractible then the entrepreneur will take $r$ as given and to choose an effort level $e$ he should solve:
	\begin{align*}
	\underset{e}{\max}\ e(R - r) - \frac{e^2}{2}
	\end{align*}
	the solution is $e^* = \min\{R - r, 1\}$. To choose $r$ such that the investor would be willing to accept take-it-or-leave offer, the entrepreneur should solve:
	\begin{align*}
	\underset{r}{\max}\ e^*(R-r) - \frac{e^{*2}}{2}\\
	s.t.\ e^*r - K \ge 0
	\end{align*}
	If $e^* = 1$ then obviously $r^* = \max\{K, R-1\}$. If $e^* = R - r$ then
	\begin{align*}
	\underset{r}{\max}\ \frac{(R - r)^2}{2}\\
	s.t.\ r(R - r) - K \ge 0\\
	r^2 - rR + K \le 0\\
	\frac{R - \sqrt{R^2 - 4K}}{2} \le r \le \frac{R + \sqrt{R^2 - 4K}}{2}
	\end{align*}
	Note that $R^2 - 4K > 0$ guarantees existence of $\sqrt{R^2 - 4K}$. Since we assumed that $R - r > 0$ then $r^* = \frac{R - \sqrt{R^2 - 4K}}{2}$. The solution is
	\begin{align*}
	e = \begin{cases}
	1, K \le R - 1\\
	\frac{R + \sqrt{R^2 - 4K}}{2}, K > R - 1
	\end{cases} r = \begin{cases}
	R - 1, K \le R - 1\\
	\frac{R - \sqrt{R^2 - 4K}}{2}, K > R - 1
	\end{cases}
	\end{align*}
	\item Now the entrepreneur should solve:
	\begin{align*}
	\underset{r}{\max}\ e^*(R - r) - \frac{e^{*2}}{2}\\
	s.t.\ e^*r - (K - w) \ge 0
	\end{align*}
	If we assume that $e = \frac{R + \sqrt{R^2 - 4(K - w)}}{2} \le 1$ then expected profit is
	\begin{align*}
	V(w) = \frac{\left(R + \sqrt{R^2 - 4(K - w)}\right)^2}{8}
	\end{align*}
	\begin{align*}
	V'(w) = \frac{R + \sqrt{R^2 - 4(K - w)}}{\sqrt{R^2 - 4(K - w)}} \ge 0\\
	V''(w) = \frac{1 - \frac{2R}{\sqrt{R^2 - 4(K - w)}}}{R^2 - 4(K - w)} \le 0
	\end{align*}
	Interpretation: if the entrepreneur has more wealth then his expected wealth will be more (surprisingly). And the less initial wealth of entrepreneur is, the more he is benefited from the loan.
\end{enumerate}

\section*{2 Raising Liability Rule}
	We consider a lender-borrower relationship under moral hazard. The risk-neutral borrower wants to borrow I from a lender to finance a project with safe return $V$ . The project may with probability $1-e$ harm a third-party. The amount of safety care $e$ costs $\psi(e)$ to the borrower with $(\psi' > 0, \psi'' > 0, \psi''' > 0)$. The harm has value $h$. A financial contract is a pair $(\ubar{t}, \lbar{t})$ where $\lbar{t}$ (resp. $\ubar{t}$) is the borrower's reimbursement to the bank if	there is no (resp. one) environmental damage.

	\begin{enumerate}
		\item Suppose that $e$ is observable. Compute the first-best level of safety care and assume
		that the project is socially valuable when the interest rate is $r$.
		\item Suppose now that $e$ is not observable. We suppose that the bank is competitive
		and that the borrower has sufficient liability. Show that the first-best is still implementable if the bank must reimburse $h$ to the third-party in case of an accident.
		\item Suppose that the bank must reimburse $c < h$ to the third-party. We denote by $w$
		the initial assets of the borrower. Show that as $w$ diminishes, the first-best level of
		effort can no longer be implemented.
		\item Compute the second-best optimal level of effort maximizing the borrower's expected payoff subject to the bank's zero profit constraint, the borrower's incentive
		constraint and his limited liability constraint.
		\item Show that raising the bank's liability c leads to a lower expected welfare.
		\item Show that this result no longer holds when the bank is a monopoly.
	\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item Obviously, if neither bank nor borrower cares about the third-party and safety care is costly, the optimal care is $e = 0$. Denote $r$ the marginal interest rate i.e. if the investor borrows $I$, he should return $(1+r)I$. In this case, the optimal contract is:
	\begin{align*}
	\lbar{t} = \ubar{t} &= V\\
	e &= 0
	\end{align*}
\end{enumerate}
\begin{enumerate}
	\item 
\end{enumerate}
\section*{3 Contracting under Coercion}
	A firm wishes to employ an agent. The agent has limited liability so wages are
	nonnegative, but the firm can punish the agent in a way that is unproductive (e.g. by
	humiliating him). The firm also can lower the worker's outside option by purchasing
	"guns" to harass the agent. More precisely, the timing is:
	\begin{enumerate}
	\item The firm chooses the level of guns, $g$, at cost $k(g)$.
	\item The firm offers a contract $\langle w(y), p(y)\rangle$ to the agent, where $y \in \{L, H\} = \{0, 1\}$ is the output, $w \ge 0$ is the wage, and $p \ge 0$ is the nonpecuniary punishment. If the
	agent rejects the contract, he receives $\ubar{u}-g$.
	\item If the agent accepts the contract, he chooses effort $a \in [0, 1]$ at cost $c(a)$, giving rise
	to output $\prob (y = H) = a$.
	\end{enumerate}

Payoffs for the firm and worker are:
\begin{align*}
\Pi = zy - w - k(g)\\
U = w - p -c(a)
\end{align*}
where $z$ is the price of output. For simplicity, suppose $k(\cdot)$ and $c(\cdot)$ are increasing,
convex and satisfy the appropriate Inada conditions (so we don't have to worry about
boundary problems).
\begin{enumerate}[(a)]
\item Write down the firm's problem subject to the $(PC)$ constraint and the $(IC)$ constraint.
\item Simplify the problem using the first-order approach. Is this approach valid here?
\item Argue that the optimal contract has $p_H = 0$ and $w_L = 0$, and that the $(PC)$ constraint binds.
\item Use the $(PC)$ and $(IC)$ constraints to write profits as follows:
\begin{align*}
\Pi = za - ac(a) + a(1-a)c'(a) - a\ubar{u} + ag-k(g)
\end{align*}
Henceforth, let us assume $\Pi$ is concave in $a$. For $x \in \mathscr{R}^n$, $t \in \mathscr{R}^n$, a function $f(x, t)$ is supermodular in $(x, t)$ if all the cross-partial derivatives are positive. Topkis proved that
if $f(x, t)$ is supermodular then the optimal solution
\begin{align*}
x(t) = argmax_xf(x, t)
\end{align*}
is increasing in the parameter $t$.
\item How does the optimal choice of $a$ and $g$, ($a^*(z, u), g^*(z, \ubar{u})$), vary in the price of
output and the outside option? Provide an intuition.
\end{enumerate}

\textbf{Solution}

\begin{enumerate}[(a)]
	\item 
\end{enumerate}

\section*{4 Moral Hazard in Teams}
	Consider Holmstrom's model of moral hazard in teams. There are $N > 1$ agents that
	work in a team. They exert efforts $e_i \ge 0, i = 1, \dots,N$ to jointly produce the output $x(e_1,\dots, e_N )$, where $x(e_1,\dots,  e_N )$ is increasing differentiable function of each effort $e_i$. For each agent,
	the cost of exerting an effort $e$ is given by an increasing, differentiable and convex function
	$c(e)$. Each agent observes her own effort and total output, but not the efforts of the other
	agents, so they cannot contract on the effort level.
	\begin{enumerate}
	\item Characterize the first-best (i.e., full information) levels of efforts $e_i, i = 1, \dots, N$
	\item Assume that the agents agreed on a differentiable balanced-budget sharing rule
	$\{t_i(x)\}_{i=1, \dots, N}$ . That is, agent $i$ receives $t_i(x)$, which yields her the payoff of
	\begin{align*}
	u_i = t_i(x(e_1, \dots, e_N )) - c(e_i),
	\end{align*}
	and the transfers to all agents exactly exhaust the budget
	\begin{align}
	\sum_{i=1}^N t_i(x) = x
	\end{align}
	Is there a differentiable balanced-budget sharing rule which allows to sustain the
	efficient production level $x^*= x(e_1,\dots,e_N )$? Support your answer analytically.
	Suggest an intuitive explanation.
	\item Assume now that there is an $N + 1$-th actor in this game, a principal, who does
	not contribute to the production but may be part of the sharing rule. Consider the
	following differentiable sharing rule
	\begin{align}
	t_i(x) = x - \frac{N-1}{N}x(e_1,\dots, e_N) \text{ for } i = 1, \dots, N\\
	t_{N+1}(x) = (N-1)(x(e_1, \dots, e_N) - x) 
	\end{align}
	\begin{enumerate}[i.]
	\item Show that this sharing rule is not balanced among agents $1, \dots, N$, but balanced
	among all $N + 1$ agents.
	\item  Does this rule allow to attain the efficient production level? Justify your answer analytically.
	\item  What is the role of the principal here? Does the principal receive any payoff
	\end{enumerate}
	in equilibrium?
	\item Assume that the sharing rule is as in (b), but now the principal can collude with
	one of the agents, say, agent $k$. That is, assume that the principal and agent $k$ can
	write a secret contract maximizing their joint welfare. More specifically, consider a
	contract that makes agent $k$ residual claimant on the joint surplus of her and the
	principal (i.e., she would pay the principal a fixed fee, and maximize their joint
	welfare). The other agents would not know about this collusion.
	Suppose that first, the principal and agent $k$ decide whether to collude, then all
	productive agents $1, \dots, N$ (including agent $k$) choose and exert efforts, and finally
	the transfers are made according to the rule (2)-(3).
	Is the sharing rule (2)-(3) susceptible to collusion (i.e., can the agents still attain
	the efficient production level?) Prove your answer and support it by an intuitive explanation.
	\end{enumerate}
\end{document}

### 2019-04-01 micro4_4.tex
Micro

@@
 \begin{enumerate}
-	\item Obviously, if neither bank nor borrower cares about the third-party and safety care is costly, the optimal care is $e = 0$.
-Denote $r$ the marginal interest rate i.e.
-if the investor borrows $I$, he should return $(1+r)I$.
-In this case, the optimal contract is:
+	\item First best level of safety care is determined by $(1 - e)h + \psi(e) \to \min$. Hence $\psi'(e^{FB}) = h$.
+	\item If the effort level is not observable, then the agent is solving:
 	\begin{align*}
-	\lbar{t} = \ubar{t} &= V\\
-	e &= 0
+	\underset{e}{\max}\ V - (1+r)I - e \lbar{t} - (1 - e)\ubar{t} - \psi(e)
 	\end{align*}
-\end{enumerate}
-\begin{enumerate}
-	\item 
+	and the solution is given by $\psi(e^*) = \ubar{t} - \lbar{t}$. If bank sets $\ubar{t} - \lbar{t} = h$ then $e^* = e^{FB}$. Bank's expected payoff in this case will not be negative:
+	\begin{align*}
+	e \lbar{t} + (1 - e) (\ubar{t} - h) = \lbar{t} \ge 0
+	\end{align*}
+	I assume that "sufficient liability" means that the borrower's utility does not become negative if $\lbar{t}, \ubar{t}$ are such as it is provided above.
+	\item If the borrower has $w$ initial wealth, his limited liability constraint implies that even in case of damaging he cannot pay to the bank more than $w$, i.e. $\ubar{t} \le w$. That means that
+	\begin{align*}
+	\lbar{t} = \ubar{t} - \psi(e) \le w - \psi'(e)
+	\end{align*}
+	Bank's profit is
+	\begin{align*}
+	e \lbar{t} + (1 - e) \ubar{t} - (1 + r)I - (1 - e)c \le w - e \psi'(e)- (1 + r)I - (1 - e)c
+	\end{align*}
+	Obviously, once $w$ diminishes the RHS of the expression above could become negative, that means that first best effort level could be no longer implementable.
+	\item \begin{align*}
+	&\underset{e}{\max}\ V - (1 + r)I - e \lbar{t} - (1 - e) \ubar{t} - \psi(e)\\
+	s.t.\ &e \lbar{t} + (1 - e) (\ubar{t} - c) - (1 + r)I = 0\ (\text{Bank's zero profit})\\
+	&\psi'(e) = \ubar{t} - \lbar{t}\ (IC)\\
+	&\ubar{t} \le w\ (LLC)
+	\end{align*}
+	The solution, hence, is determined by:
+	\begin{align*}
+	w - (1 - e)c - e \psi'(e) - (1 + r)I = 0
+	\end{align*}
+	\item Expected welfare decreases once $e$ decreases.
+	\begin{align*}
+	\frac{de}{dc} = \frac{1-e}{c - \psi'(e) - e\psi''(e)} < 0
+	\end{align*}
+	if $c < \psi'(e) + e\psi''(e)$.
+	\item If the bank is monopoly then it is solving:
+	\begin{align*}
+	&\underset{\lbar{t}, \ubar{t}}{\max}\ e \lbar{t} + (1 -e)(\ubar{t} - c) - (1 +r)I\\
+	s.t.\ &\ubar{t} - \lbar{t} = \psi'(e)\\
+	&\ubar{t} \le w\\
+	\end{align*}
+	the solution is determined by
+	\begin{align*}
+	w - e \psi''(e) - \psi'(e) - w + c = 0\\
+	c - e \psi''(e) - \psi'(e) = 0\\
+	\frac{de}{dc} = \frac{1}{2\psi''(e) + e\psi'''(e)} > 0
+	\end{align*}
 	\end{enumerate}

### 2019-04-03 micro4_4.tex
Micro

@@
 \begin{enumerate}[(a)]
+	\item \begin{align*}
+	&\underset{(w_L, p_L), (w_H, p_H), g, a}{\max}\ a(z - w_H) - (1 - a) w_L - k(g)\\
+	s.t.\ &a(w_H - p_H) + (1 - a)(w_L - p_L) - c(a) \ge \ubar{u} - g\ (PC)\\
+	&w_H - p_H - (w_L - p_L) = c'(a)\ (IC)
+	\end{align*}
 	\item 
+	\item $(PC)$ binds because otherwise the principal can decrease $w_H$ and $w_L$ by the same value which does not affect $(IC)$ but objective function will increase.
+Principal's objective function is increasing in $a$, decreasing in $w_L$ and does not directly depend on $p_H$.
+While $(IC)$ implies that
+	$w_H + p_L - p_H - w_L = c'(a)$.
+Since $c'(a)$ is increasing function, it is optimal for principal to make the LHS of the aforementioned expression as large as possible.
+Obviously $p_H = 0$, it does not directly affect the objective function and $w_L = 0$, the objective function is decreasing in $w_L$.
+	\item \begin{align*}
+	&\Pi = az - aw_H - k(g)\\
+	&(PC) \to aw_H - (1 - a)p_L - c(a) = \ubar{u} - g\\
+	&(IC) \to w_H + p_L = c'(a)\\
+	&ac'(a) - p_L - c(a) = \ubar{u} - g\\
+	&p_L = ac'(a) - c(a) - \ubar{u} + g\\
+	&w_H = c'(a) - ac'(a) + c(a) + \ubar{u} - g\\
+	&\Pi = az - ac'(a) + a^2c'(a) - ac(a) - a\ubar{u} + ag - k(g)\\
+	&\Pi = az -a(1 - a)c'(a) - ac(a) - a\ubar{u} + ag - k(g)
+	\end{align*}
+	\item The profit function above is supermodular in $(a, g, z, -\ubar{u})$. Indeed,
+	\begin{align*}
+	\frac{\partial^2 \Pi}{\partial a \partial g} = 1 \ge 0\ \frac{\partial^2 \Pi}{\partial a \partial z} = 1 \ge 0\  
+	\frac{\partial^2 \Pi}{\partial a \partial( -\ubar{u})} = 1 \ge 0\\
+	\frac{\partial^2 \Pi}{\partial g \partial z} = 0 \ge 0 
+	\frac{\partial^2 \Pi}{\partial g \partial (-\ubar{u})} = 0 \ge 0   \\
+	\frac{\partial^2 \Pi}{\partial z \partial(-\ubar{u})} = 0 \ge 0
+	\end{align*}
+	Hence the optimal choice of $a$ and $g$ increases in price of output $z$ and decreases in outside option $\ubar{u}$.
+The more the price of the output is the more firm values output hence the more it will spend on guns to incentivize the worker to work, similarly it should exert more efforts once the price of output increases.
+When the outside option increases firm has to pay worker more and punish him less, hence the level of guns decreases as well as the level of effort.
 \end{enumerate}

### 2019-05-07 taxes_evasion.tex
taxes

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage{makecell}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{xcolor,colortbl}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
\section{Enough undeclared money to pay the worker under the table}
This is the case once $e\theta - y' \ge \beta$.
\subsection{Enforceability problem}
\begin{align*}
&\underset{\beta', e, y'}{\max}\ \pi = y'(1 - t) - \beta'(1-t)(1 - \gamma) + \gamma[e\theta - y' - c(e)] - pk(e\theta - y')\\
s.t.\ &\frac{\delta \pi}{1 - \delta} \ge (1-t)\beta' + \gamma \beta\ (DE) \\
&e\theta - y' \ge \beta
\end{align*}
Using the (IC), (DE) becomes
\begin{align*}
\delta y'(1 - t - \gamma) + \delta \gamma e \theta - \delta p k(e \theta - y') \ge \gamma c(e) + \beta'(1-t)(1 - \gamma)
\end{align*}
Then the problem is
\begin{align*}
\underset{e, y', \beta'}{\max}\ &y'(1 - t) - \beta'(1-t)(1 - \gamma) + \gamma[e\theta - y' - c(e)] - pk(e\theta - y')\\
s.t.\ &\delta y'(1 - t - \gamma) + \delta \gamma e \theta - \delta p k(e \theta - y') \ge \gamma c(e) + \beta'(1-t)(1 - \gamma)\ \to \lambda \\
& e\theta - y' \ge c(e) - (1 - t)\beta'\ \to \mu \\
& \beta', y', e \ge 0\ \to \psi_1, \psi_2, \psi_3\ \\
\end{align*}

Then FOCs are 
\begin{align}
\frac{\partial \pi}{\partial e} &= \gamma \theta - \gamma c'(e)- \theta p k'(e \theta - y') + \lambda \delta \gamma \theta - \lambda \delta \theta pk'(e \theta - y') - \lambda \gamma c'(e) + \mu \theta - \mu c'(e) + \psi_3= 0\label{eq1}\\
\frac{\partial \pi}{\partial y'} &= (1 - t - \gamma)+pk'(e\theta - y') + \lambda \delta (1 - t - \gamma) +\lambda \delta pk'(e\theta - y') - \mu + \psi_2 = 0\label{eq2}\\
\frac{\partial \pi}{\partial \beta'} &= -(1-t)(1-\gamma) - \lambda (1-t)(1-\gamma) + \mu(1 - t) + \psi_1= 0\label{eq3}\\
&\lambda, \mu, \psi_1, \psi_2, \psi_3 \ge 0\nonumber\\
&+\text{ slackness complementary conditions}\nonumber
\end{align}

It is clear that if $e \theta - y' \ge c(e)$ then $\beta' = 0$. Indeed, assume the contrary, $\beta' > 0$ then the principal can slightly decrease $\beta'$. This cannot violate the (DE) constraint. Additionally, if $e\theta - y' \ge c(e)$ then it is for sure $\ge c(e) - (1 - t)\beta'$ for $\beta'>0$, thus, this constraint cannot be violated either. But the profit increases. Hence, $\beta'>0$ is not optimal. Intuitively: if there is enough undeclared money to fully compensate the worker's cost, it is optimal to pay this compensation under the table, since this income is tax free for the worker. Though, it could be not optimal for the firm to make the difference $e\theta - y'$ too large (e.g. more than $c(e)$) because the punishment for tax evasion depends positively upon it, but if it is optimal then the declared bonus is zero.

Let us consider this case. Moreover, suppose that $e \theta - y' > c(e)\ \to \mu = 0$, and that the (DE) constraint does not bind. This case is equivalent to the case without the enforceability problem, thus, the optimal level of effort and the amount of declared revenue are determined by
\begin{align*}
c'(e) &= \frac{(1 - t)\theta}{\gamma}\\
pk'(e\theta - y') &= - (1 - t - \gamma)
\end{align*}
Assuming that the (DE) binds, FOCs become
\begin{align*}
\frac{\partial \pi}{\partial e} &= (1 + \lambda \delta)(\gamma \theta - \theta p k'(e \theta - y')) -(1+ \lambda) \gamma c'(e)= 0\\
\frac{\partial \pi}{\partial y'} &= (1 + \lambda \delta)(1 - t - \gamma+pk'(e\theta - y')) = 0\\
&\text{if }1-t < \gamma \text{ then }c'(e) = \frac{(1+\lambda \delta)\theta(1-t)}{\gamma (1 + \lambda)} \underset{\delta < 1}{<} \frac{\theta(1-t)}{\gamma}\\
&pk'(e \theta - y') = -(1 - t - \gamma)
\end{align*}
(Note that $\lambda$ can be found from the DE constraint which is binding).

Assume that $\mu > 0$. That means that $e\theta - y' = c(e)$. If the (DE) constraint is not binding then FOCs become
\begin{align*}
\frac{\partial \pi}{\partial e} &= \gamma \theta- \gamma c'(e) - \theta p k'(e \theta - y') + \mu \theta - \mu c'(e) = 0\\
\frac{\partial \pi}{\partial y'} &= (1 - t - \gamma)+pk'(e\theta - y') - \mu  = 0\\
c'(e) &= \frac{\theta(1 - t)}{\mu + \gamma}\\
pk'(e\theta - y') &= \mu - (1 - t - \gamma)
\end{align*}
if the (DE) constraint is binding then 
\begin{align*}
\frac{\partial \pi}{\partial e} &= (1 + \lambda \delta)(\gamma \theta - \theta p k'(e \theta - y')) - (1+\lambda) \gamma c'(e)+ \mu \theta - \mu c'(e) = 0\\
\frac{\partial \pi}{\partial y'} &= (1+\lambda \delta)(1 - t - \gamma+pk'(e\theta - y')) - \mu = 0\\
c'(e) &= \frac{\theta(1-t)(1+\lambda \delta)}{(1 + \lambda)\gamma + \mu}\\
pk'(e\theta - y') &= \frac{\mu}{1+\lambda \delta} - (1 - t- \gamma)
\end{align*}

If $c(e) - (1-t)\beta' \le e \theta - y' < c(e)$ then $\beta'$ can (and should) be positive.
Assuming in addition $y' > 0, e > 0$ then $\psi_1 = 0, \psi_2 = 0, \psi_3= 0$ and hence from \eqref{eq3}:
\begin{align*}
(1 + \lambda)(1 - \gamma) = \mu > 0\ \to e \theta - y' = c(e) - (1 - t)\beta'
\end{align*}
Then if the (DE) constraint does not bind, $\lambda = 0, \mu = 1 - \gamma$ and FOCs imply:
\begin{align*}
\frac{\partial \pi}{\partial e} &= \theta - c'(e) - \theta p k'(e\theta - y') = 0\\
\frac{\partial \pi}{\partial y'} &= (1 - t - \gamma)+pk'(e\theta - y') - (1-\gamma) = 0
\end{align*}
\begin{align*}
c'(e) &= \theta(1 - t)\\
pk'(e\theta - y') &= t\\
\beta' &= \frac{c(e) - (e\theta - y')}{1-t}
\end{align*}
The price of money laundering $\gamma$ affects neither the optimal level of effort nor the amount of declared revenue because the whole undeclared revenue is paid as a black bonus to the worker.

Assume that (DE) does bind. That means that $\lambda > 0$.
\begin{align*}
\frac{\partial \pi}{\partial e} &= (1 + \lambda \delta)(\gamma \theta - \theta p k'(e \theta - y')) - (1+\lambda)c'(e) + (1+\lambda)(1-\gamma)\theta = 0\\
\frac{\partial \pi}{\partial y'} &= (1+\lambda \delta)(1 - t - \gamma+pk'(e\theta - y')) - (1+\lambda)(1-\gamma) = 0\\
c'(e) &= \frac{(1+\lambda \delta)\theta(\gamma - pk'(e \theta - y'))+ (1+\lambda)(1-\gamma)\theta}{1+\lambda} = \frac{\theta(1-t)(1 + \lambda \delta)}{1+\lambda}\underset{\delta < 1}{<} \theta(1-t)\\
pk'(e\theta - y') &= \frac{(1+\lambda)(1-\gamma)}{1+\lambda \delta} - (1 - t - \gamma)\\
\beta' &= \frac{\delta y'(1 - t - \gamma) + \delta \gamma e \theta - \delta p k(e \theta - y') - \gamma c(e)}{(1-t)(1-\gamma)} = \frac{c(e) - (e \theta - y')}{1-t}
\end{align*}
The optimal level of effort still does not depend on the cost of money laundering. While the amount of declared profit $y'$ does depend. That is because the (DE) constraint is binding and $\gamma$ affects this constraint.

The results of this section are summarized by the table below:
\begin{center}
	\begin{tabular}{ |c|c|c|c|c|c| } 
		\hline
		 & \makecell{$\mu = 0$\\$\lambda>0$}&\makecell{$\mu = 0$\\$\lambda=0$}&\makecell{$\mu > 0$\\$\lambda>0$}&\makecell{$\mu > 0$\\$\lambda=0$} \\
		\hline
		$\beta' = 0$ & \makecell{$c'(e) = \frac{(1+\lambda\delta)\theta(1-t)}{\gamma(1 + \lambda)}$\\\\$pk'(e\theta - y') =$\\$ -(1 - t - \gamma)$} & \makecell{$c'(e) = \frac{\theta(1-t)}{\gamma}$\\\\$pk'(e\theta - y') =$\\$ -(1 - t - \gamma)$}&\makecell{$c'(e) = \frac{\theta(1-t)(1+\lambda\delta)}{(1+\lambda)\gamma + \mu}$\\\\$pk'(e\theta - y') =$\\$ \frac{\mu}{1+\lambda\delta}-(1 - t - \gamma)$}&\makecell{$c'(e) = \frac{\theta(1-t)}{\mu +\gamma}$\\\\$pk'(e\theta - y') =$\\$\mu -(1 - t - \gamma)$} \\ 
		\hline
		$\beta' >0$ & \cellcolor[HTML]{B2BEB5}& \cellcolor[HTML]{B2BEB5}&\makecell{$c'(e) = \frac{\theta(1-t)(1+\lambda\delta)}{1+\lambda}$\\\\$pk'(e\theta - y') =$\\$\frac{(1+\lambda)(1-\gamma)}{1+\lambda \delta} -(1 - t - \gamma)$} & \makecell{$c'(e) = \theta(1-t)$\\\\$pk'(e\theta - y') =t$} \\ 
		\hline
	\end{tabular}
\end{center}
\section{Not enough undeclared money}
Consider the case $e\theta - y' < \beta$. Since it is less costly to pay the worker out of undeclared money, it is worth paying as much as possible out of undeclared money and then pay the remaining amount out of declared profit. That means that 
\begin{align*}
e\theta - y' = (1 - \alpha) \beta\\
\alpha = 1 - \frac{e \theta - y'}{\beta}
\end{align*}
The firm's profit becomes:
\begin{align*}
\pi = (y' - \beta')(1 - t) - \beta + e \theta - y' - pk(e \theta - y')
\end{align*}
Using (IC) $\beta = c(e) - \beta'(1-t)$ the profit function is
\begin{align*}
\pi = (y'-\beta')(1-t) - c(e) + \beta' (1-t) +e \theta -y' -pk(e \theta - y') = \\ =-y't - c(e) + e \theta -pk'(e \theta - y')
\end{align*}
\subsection{No credibility problems}
Without enforceability problems, FOCs are
\begin{align*}
\frac{\partial \pi }{\partial e} &= -c'(e) + \theta - \theta p k'(e\theta - y') = 0\\
\frac{\partial \pi}{\partial y'} &= -t + pk'(e \theta - y') = 0\\
c'(e) &= \theta(1-t)\\
pk'(e \theta -y') &= t
\end{align*}
\subsection{Enforceability problems}
The (DE) constraint is
\begin{align*}
\frac{\delta \pi}{1 - \delta} &\ge (1 - t)\beta' + \gamma \beta\\
-\delta y't - \delta c(e) + \delta e \theta -\delta pk'(e \theta - y') &\ge (1 - \delta)(1 - t)\beta' + (1 - \delta)\gamma c(e) - \gamma (1 - \delta )(1 - t)\beta'
\end{align*}
The problem becomes
\begin{align*}
\max\ &-y't - c(e) + e \theta -pk'(e \theta - y')\\
s.t.\ &-\delta y't + \delta e \theta -\delta pk'(e \theta - y') \ge (1 - \gamma)(1 - \delta)(1 - t)\beta' + \gamma c(e)\to \lambda \ (DE)\\
& e \theta - y' < c(e) - (1-t)\beta'\ \to \mu \\
& e\theta - y' \ge 0\ \to \eta\\
& \beta' \ge 0, y' \ge 0, e \ge 0\ \to \psi_1, \psi_2,\psi_3
\end{align*}
\begin{align}
\frac{\partial \pi}{\partial y'} &= -t + pk'(e \theta - y') - \lambda \delta t + \lambda \delta pk'(e \theta - y') + \mu - \eta + \psi_2 = 0 \label{eq4}\\
\frac{\partial \pi}{\partial e} &= -c'(e)+\theta - \theta pk'(e\theta - y')- \lambda \gamma c'(e) + \lambda \delta \theta - \lambda \delta \theta pk'(e \theta - y') - \mu \theta + \nonumber\\
&+\mu c'(e) + \eta \theta+ \psi_3 = 0\label{eq5}\\
\frac{\partial \pi}{\partial \beta'} &= - \lambda(1-\gamma)(1-\delta)(1-t) - \mu(1-t) + \psi_1 = 0\label{eq6}\\
& \lambda, \mu, \eta, \psi_1, \psi_2, \psi_3 \ge 0 \label{eq7}\\
&+\text{ slackness complementary conditions}\nonumber
\end{align}
Assume firstly that $\beta', y', e >0$, that means $\psi_1 = \psi_2=\psi_3 = 0$, consequently from \eqref{eq6} and \eqref{eq7}, $\lambda = \mu = 0$. If $e \theta - y' > 0$ then $\eta = 0$ and the solution is equivalent to the case without credibility problems.

Now assume that $\psi_1 > 0\ \to \beta' = 0$, thus, implying $\mu = 0$, $\lambda = \frac{\psi_1}{(1-t)(1-\delta)(1-\gamma)} > 0$ hence the (DE) constraint binds. Then from the FOCs
\begin{align*}
c'(e) = \frac{(1 + \lambda \delta)\theta(1-t)}{1+\lambda} \underset{\delta < 1}{<}\theta(1-t)\\
pk'(e\theta - y') = t
\end{align*}
and $\lambda$ can be found from the (DE) constraint.
\end{document}

### 2019-05-09 taxes_evasion.tex
taxes

@@
 
+\subsubsection{The declared bonus is zero}
 It is clear that if $e \theta - y' \ge c(e)$ then $\beta' = 0$.
@@
 \end{align*}
-(Note that $\lambda$ can be found from the DE constraint which is binding).
-
+(Note that $\lambda$ can be found from the DE constraint which is binding). Comparative statics are the following:
+\begin{itemize}
+	\item An increase in $\theta$ increases $e \theta$ directly and indirectly (through FOC), however, now the indirect influence is smaller than in case of the absence of the enforceability problem.
+This is because the increase in $e$ tightens the (DE) constraint, so it is harder for the firm to incentivize efforts.
+Also, $y'$ should increase to keep the FOC satisfied.
+	\item An increase in $\gamma$ (or $t$) decreases $e$ through FOC, however, now the decrease is smaller than in case of the absence of the enforceability problem.
+$y'$ should also decrease to keep FOC satisfied.
+	\item An increase in $\lambda$ also decreases $e$: once the "shadow" price of relaxing the (DE) constraint goes up - it becomes more costly to incentivize effort.
+Furthermore, comparing cases $\lambda = 0$ and $\lambda > 0$ we can conclude that if the latter case implies less effort, but in both cases $pk'(e\theta - y') = -(1 - t - \gamma)$ then, other things equal, $y'$ in the latter case should also be smaller.
+That is the presence of the (DE) constraint not only decreases the effort level but also decreases the declared revenue.
+\end{itemize}
 Assume that $\mu > 0$. That means that $e\theta - y' = c(e)$. If the (DE) constraint is not binding then FOCs become
@@
 
+Now it is unclear, whether the (DE) constraint causes an increase or a decrease in $e$. It depends on the relationship between $\gamma$ and $\frac{\delta}{1-\delta} \mu$: 
+\begin{align*}
+\text{if } \gamma > \frac{\delta}{1-\delta}\mu \text{  then  } \frac{\theta(1-t)}{\mu + \gamma} > \frac{\theta(1-t)(1 + \lambda \delta)}{(1 + \lambda)\gamma + \mu},
+\end{align*} hence, the presence of the (DE) constraint decreases the level of effort. If $\gamma \le \frac{\delta}{1-\delta}\mu$ then $e$ either increases or remains unchanged.
+The comparative statics are the following:
+\begin{itemize}
+	\item As always, increase in $\theta$ increases $e$ and $y'$.
+Now the increase in $e$ is even smaller (compare to the case $\mu = 0$) because of additional shadow cost of keeping the constraint $e\theta - y' = c(e)$ bound.
+	\item The effect of increasing $\gamma$ (or $t$) qualitatively is the same as in case of $\mu = 0$.
+	\item The effect on $e$ of increasing $\lambda$ is ambiguous and depends on the sign of $\gamma - \frac{\delta}{1-\delta}\mu$.
+\end{itemize}
+\subsubsection{The declared bonus is positive}
 If $c(e) - (1-t)\beta' \le e \theta - y' < c(e)$ then $\beta'$ can (and should) be positive.
@@
 \end{align*}
-The price of money laundering $\gamma$ affects neither the optimal level of effort nor the amount of declared revenue because the whole undeclared revenue is paid as a black bonus to the worker.
+The comparative statics are the following:
+\begin{itemize}
+\item An increase in $\theta$ increases $e$, 
+it also increases the declared revenue $y'$ to keep the FOC satisfied, consequently it increases the declared bonus $\beta'$ ($e \theta - y'$ remains unchanged and $c(e)$ goes up).
+\item An increase in $t$ decreases $e$ and $y'$, while its affect on the declared bonus $\beta'$ is ambiguous: it decreases both the numerator and the denominator, so the affect can be of both signs.
+\item The price of money laundering $\gamma$ affects neither the optimal level of effort nor the amount of declared revenue because the whole undeclared revenue is paid as a black bonus to the worker.
+\end{itemize}
 
@@
 \end{align*}
-The optimal level of effort still does not depend on the cost of money laundering.
-While the amount of declared profit $y'$ does depend.
-That is because the (DE) constraint is binding and $\gamma$ affects this constraint.
-
+The comparative statics are the following:
+\begin{itemize}
+\item An increase in $\theta$ increase $e$, but less than in case of the absence of the enforceability problem, because an increase in $e$ tightens the (DE) constraint.
+\item The optimal level of effort still does not depend on the cost of money laundering.
+However, now it positively affects $y'$ i.e.
+an increase in $\gamma$ causes the increase in $y'$, because an increase in $\gamma$ tightens the (DE) constraint and $y'$ should be increased to keep it satisfied.
+\end{itemize}
 The results of this section are summarized by the table below:
@@
 \section{Not enough undeclared money}
-Consider the case $e\theta - y' < \beta$.
-Since it is less costly to pay the worker out of undeclared money, it is worth paying as much as possible out of undeclared money and then pay the remaining amount out of declared profit.
-That means that
-\begin{align*}
-e\theta - y' = (1 - \alpha) \beta\\
-\alpha = 1 - \frac{e \theta - y'}{\beta}
-\end{align*}
-The firm's profit becomes:
-\begin{align*}
-\pi = (y' - \beta')(1 - t) - \beta + e \theta - y' - pk(e \theta - y')
-\end{align*}
-Using (IC) $\beta = c(e) - \beta'(1-t)$ the profit function is
-\begin{align*}
-\pi = (y'-\beta')(1-t) - c(e) + \beta' (1-t) +e \theta -y' -pk(e \theta - y') = \\ =-y't - c(e) + e \theta -pk'(e \theta - y')
-\end{align*}
-\subsection{No credibility problems}
-Without enforceability problems, FOCs are
-\begin{align*}
-\frac{\partial \pi }{\partial e} &= -c'(e) + \theta - \theta p k'(e\theta - y') = 0\\
-\frac{\partial \pi}{\partial y'} &= -t + pk'(e \theta - y') = 0\\
-c'(e) &= \theta(1-t)\\
-pk'(e \theta -y') &= t
-\end{align*}
-\subsection{Enforceability problems}
-The (DE) constraint is
-\begin{align*}
-\frac{\delta \pi}{1 - \delta} &\ge (1 - t)\beta' + \gamma \beta\\
--\delta y't - \delta c(e) + \delta e \theta -\delta pk'(e \theta - y') &\ge (1 - \delta)(1 - t)\beta' + (1 - \delta)\gamma c(e) - \gamma (1 - \delta )(1 - t)\beta'
-\end{align*}
-The problem becomes
-\begin{align*}
-\max\ &-y't - c(e) + e \theta -pk'(e \theta - y')\\
-s.t.\ &-\delta y't + \delta e \theta -\delta pk'(e \theta - y') \ge (1 - \gamma)(1 - \delta)(1 - t)\beta' + \gamma c(e)\to \lambda \ (DE)\\
-& e \theta - y' < c(e) - (1-t)\beta'\ \to \mu \\
-& e\theta - y' \ge 0\ \to \eta\\
-& \beta' \ge 0, y' \ge 0, e \ge 0\ \to \psi_1, \psi_2,\psi_3
-\end{align*}
-\begin{align}
-\frac{\partial \pi}{\partial y'} &= -t + pk'(e \theta - y') - \lambda \delta t + \lambda \delta pk'(e \theta - y') + \mu - \eta + \psi_2 = 0 \label{eq4}\\
-\frac{\partial \pi}{\partial e} &= -c'(e)+\theta - \theta pk'(e\theta - y')- \lambda \gamma c'(e) + \lambda \delta \theta - \lambda \delta \theta pk'(e \theta - y') - \mu \theta + \nonumber\\
-&+\mu c'(e) + \eta \theta+ \psi_3 = 0\label{eq5}\\
-\frac{\partial \pi}{\partial \beta'} &= - \lambda(1-\gamma)(1-\delta)(1-t) - \mu(1-t) + \psi_1 = 0\label{eq6}\\
-& \lambda, \mu, \eta, \psi_1, \psi_2, \psi_3 \ge 0 \label{eq7}\\
-&+\text{ slackness complementary conditions}\nonumber
-\end{align}
-Assume firstly that $\beta', y', e >0$, that means $\psi_1 = \psi_2=\psi_3 = 0$, consequently from \eqref{eq6} and \eqref{eq7}, $\lambda = \mu = 0$.
-If $e \theta - y' > 0$ then $\eta = 0$ and the solution is equivalent to the case without credibility problems.
 
-Now assume that $\psi_1 > 0\ \to \beta' = 0$, thus, implying $\mu = 0$, $\lambda = \frac{\psi_1}{(1-t)(1-\delta)(1-\gamma)} > 0$ hence the (DE) constraint binds. Then from the FOCs
-\begin{align*}
-c'(e) = \frac{(1 + \lambda \delta)\theta(1-t)}{1+\lambda} \underset{\delta < 1}{<}\theta(1-t)\\
-pk'(e\theta - y') = t
-\end{align*}
-and $\lambda$ can be found from the (DE) constraint.
 \end{document}

### 2019-05-13 taxes_evasion.tex
taxes

@@
 \usepackage{tikz} % 
-\def\myrad{2cm}% radius of the circle
-\def\myanga{45}% angle for the arc
-\def\myangb{195}
 \begin{document} % 
@@
 Intuitively: if there is enough undeclared money to fully compensate the worker's cost, it is optimal to pay this compensation under the table, since this income is tax free for the worker.
-Though, it could be not optimal for the firm to make the difference $e\theta - y'$ too large (e.g.
-more than $c(e)$) because the punishment for tax evasion depends positively upon it, but if it is optimal then the declared bonus is zero.
+The question is, what if it is not optimal for the firm to make the difference $e\theta - y'$ too large (e.g.
+more than $c(e)$) because the punishment for tax evasion depends positively upon it?
 
-Let us consider this case.
+Let us firstly consider the case $\beta'=0$.
 Moreover, suppose that $e \theta - y' > c(e)\ \to \mu = 0$, and that the (DE) constraint does not bind.
 This case is equivalent to the case without the enforceability problem, thus, the optimal level of effort and the amount of declared revenue are determined by
-\begin{align*}
-c'(e) &= \frac{(1 - t)\theta}{\gamma}\\
-pk'(e\theta - y') &= - (1 - t - \gamma)
-\end{align*}
+\begin{align}
+c'(e) &= \frac{(1 - t)\theta}{\gamma}\label{eq4}\\
+pk'(e\theta - y') &= - (1 - t - \gamma) \label{eq5}
+\end{align}
 Assuming that the (DE) binds, FOCs become
@@
 That is the presence of the (DE) constraint not only decreases the effort level but also decreases the declared revenue.
+	\item An increase in $p$ does not affect $e$, but increases $e \theta - y'$, which means that $y'$ increases.
 \end{itemize}
 Assume that $\mu > 0$. That means that $e\theta - y' = c(e)$. If the (DE) constraint is not binding then FOCs become
-\begin{align*}
-\frac{\partial \pi}{\partial e} &= \gamma \theta- \gamma c'(e) - \theta p k'(e \theta - y') + \mu \theta - \mu c'(e) = 0\\
-\frac{\partial \pi}{\partial y'} &= (1 - t - \gamma)+pk'(e\theta - y') - \mu  = 0\\
-c'(e) &= \frac{\theta(1 - t)}{\mu + \gamma}\\
-pk'(e\theta - y') &= \mu - (1 - t - \gamma)
-\end{align*}
+\begin{align}
+\frac{\partial \pi}{\partial e} &= \gamma \theta- \gamma c'(e) - \theta p k'(e \theta - y') + \mu \theta - \mu c'(e) = 0\nonumber\\
+\frac{\partial \pi}{\partial y'} &= (1 - t - \gamma)+pk'(e\theta - y') - \mu  = 0\nonumber\\
+c'(e) &= \frac{\theta(1 - t)}{\mu + \gamma}\label{eq6}\\
+pk'(e\theta - y') &= \mu - (1 - t - \gamma)\label{eq7}
+\end{align}
 if the (DE) constraint is binding then 
@@
 
-Now it is unclear, whether the (DE) constraint causes an increase or a decrease in $e$. It depends on the relationship between $\gamma$ and $\frac{\delta}{1-\delta} \mu$: 
+Denote $e_{\mu = 0}(e_{\mu > 0}), y'_{\mu=0}(y'_{\mu>0})$ the effort level and the declared revenue in case $\mu = 0\ (\mu > 0)$.
+Then from \eqref{eq4} and \eqref{eq6} it follows that $e_{\mu >0} \le e_{\mu = 0}$.
+However from \eqref{eq5} and \eqref{eq7}:
 \begin{align*}
-\text{if } \gamma > \frac{\delta}{1-\delta}\mu \text{  then  } \frac{\theta(1-t)}{\mu + \gamma} > \frac{\theta(1-t)(1 + \lambda \delta)}{(1 + \lambda)\gamma + \mu},
-\end{align*} hence, the presence of the (DE) constraint decreases the level of effort. If $\gamma \le \frac{\delta}{1-\delta}\mu$ then $e$ either increases or remains unchanged.
-The comparative statics are the following:
-\begin{itemize}
-	\item As always, increase in $\theta$ increases $e$ and $y'$.
-Now the increase in $e$ is even smaller (compare to the case $\mu = 0$) because of additional shadow cost of keeping the constraint $e\theta - y' = c(e)$ bound.
-	\item The effect of increasing $\gamma$ (or $t$) qualitatively is the same as in case of $\mu = 0$.
-	\item The effect on $e$ of increasing $\lambda$ is ambiguous and depends on the sign of $\gamma - \frac{\delta}{1-\delta}\mu$.
-\end{itemize}
+c(e_{\mu > 0}) = e_{\mu>0}\theta - y'_{\mu>0} \ge e_{\mu=0}\theta - y'_{\mu=0} \ge c(e_{\mu=0})
+\end{align*}
+which means that $e_{\mu>0} \ge e_{\mu = 0}$ which is impossible unless $\mu \neq 0$. Thus, $\underline{\mu = 0}$.
 \subsubsection{The declared bonus is positive}
@@
 \end{align*}
-The comparative statics are the following:
-\begin{itemize}
-\item An increase in $\theta$ increases $e$, 
-it also increases the declared revenue $y'$ to keep the FOC satisfied, consequently it increases the declared bonus $\beta'$ ($e \theta - y'$ remains unchanged and $c(e)$ goes up).
-\item An increase in $t$ decreases $e$ and $y'$, while its affect on the declared bonus $\beta'$ is ambiguous: it decreases both the numerator and the denominator, so the affect can be of both signs.
-\item The price of money laundering $\gamma$ affects neither the optimal level of effort nor the amount of declared revenue because the whole undeclared revenue is paid as a black bonus to the worker.
-\end{itemize}
+
+Denote this effort level and corresponding declared revenue as $e_{\beta' >0}$ and $y'_{\beta'>0}$ respectively.
+Also denote the effort level and corresponding revenue from \eqref{eq4} and \eqref{eq5} as $e_{\beta'=0}, y'_{\beta'=0}$.
+It follows that
+\begin{align*}
+e_{\beta'=0} &> e_{\beta'>0}\\
+e_{\beta'>0}\theta - y'_{\beta'>0} &> e_{\beta'=0}\theta - y'_{\beta'=0} \ge c(e_{\beta'=0})
+\end{align*}
+which is possible only if $e_{\beta'>0}\theta - y'_{\beta'>0} \ge c(e_{\beta'>0})$ but this case has been considered above and implies that $\beta' = 0$.
 
@@
 \frac{\partial \pi}{\partial y'} &= (1+\lambda \delta)(1 - t - \gamma+pk'(e\theta - y')) - (1+\lambda)(1-\gamma) = 0\\
-c'(e) &= \frac{(1+\lambda \delta)\theta(\gamma - pk'(e \theta - y'))+ (1+\lambda)(1-\gamma)\theta}{1+\lambda} = \frac{\theta(1-t)(1 + \lambda \delta)}{1+\lambda}\underset{\delta < 1}{<} \theta(1-t)\\
-pk'(e\theta - y') &= \frac{(1+\lambda)(1-\gamma)}{1+\lambda \delta} - (1 - t - \gamma)\\
-\beta' &= \frac{\delta y'(1 - t - \gamma) + \delta \gamma e \theta - \delta p k(e \theta - y') - \gamma c(e)}{(1-t)(1-\gamma)} = \frac{c(e) - (e \theta - y')}{1-t}
+c'(e) &= \frac{(1+\lambda \delta)\theta(\gamma - pk'(e \theta - y'))+ (1+\lambda)(1-\gamma)\theta}{1+\lambda} = \frac{\theta(1-t)(1 + \lambda \delta)}{1+\lambda}\\
+pk'(e\theta - y') &= \frac{(1+\lambda)(1-\gamma)}{1+\lambda \delta} - (1 - t - \gamma)
 \end{align*}
-The comparative statics are the following:
-\begin{itemize}
-\item An increase in $\theta$ increase $e$, but less than in case of the absence of the enforceability problem, because an increase in $e$ tightens the (DE) constraint.
-\item The optimal level of effort still does not depend on the cost of money laundering.
-However, now it positively affects $y'$ i.e.
-an increase in $\gamma$ causes the increase in $y'$, because an increase in $\gamma$ tightens the (DE) constraint and $y'$ should be increased to keep it satisfied.
-\end{itemize}
-The results of this section are summarized by the table below:
+
+Using the same logic, one can show that this case also implies $e \theta - y' \ge c(e)$ which, in turns, means that $\underline{\beta' = 0}$.
+
+In conclusion, if there is enough money to pay the worker under the table, the declared bonus is zero.
+
+The results of this section are summarized by the table below
 \begin{center} 
@@
 		\hline 
-		 & \makecell{$\mu = 0$\\$\lambda>0$}&\makecell{$\mu = 0$\\$\lambda=0$}&\makecell{$\mu > 0$\\$\lambda>0$}&\makecell{$\mu > 0$\\$\lambda=0$} \\
+		& \makecell{$\lambda>0$}&\makecell{$\lambda=0$} \\ 
 		\hline 
-		$\beta' = 0$ & \makecell{$c'(e) = \frac{(1+\lambda\delta)\theta(1-t)}{\gamma(1 + \lambda)}$\\\\$pk'(e\theta - y') =$\\$ -(1 - t - \gamma)$} & \makecell{$c'(e) = \frac{\theta(1-t)}{\gamma}$\\\\$pk'(e\theta - y') =$\\$ -(1 - t - \gamma)$}&\makecell{$c'(e) = \frac{\theta(1-t)(1+\lambda\delta)}{(1+\lambda)\gamma + \mu}$\\\\$pk'(e\theta - y') =$\\$ \frac{\mu}{1+\lambda\delta}-(1 - t - \gamma)$}&\makecell{$c'(e) = \frac{\theta(1-t)}{\mu +\gamma}$\\\\$pk'(e\theta - y') =$\\$\mu -(1 - t - \gamma)$} \\
-		\hline
-		$\beta' >0$ & \cellcolor[HTML]{B2BEB5}& \cellcolor[HTML]{B2BEB5}&\makecell{$c'(e) = \frac{\theta(1-t)(1+\lambda\delta)}{1+\lambda}$\\\\$pk'(e\theta - y') =$\\$\frac{(1+\lambda)(1-\gamma)}{1+\lambda \delta} -(1 - t - \gamma)$} & \makecell{$c'(e) = \theta(1-t)$\\\\$pk'(e\theta - y') =t$} \\
+		$\beta' = 0$ & \makecell{$c'(e) = \frac{(1+\lambda\delta)\theta(1-t)}{\gamma(1 + \lambda)}$\\\\$pk'(e\theta - y') =$\\$ -(1 - t - \gamma)$} & \makecell{$c'(e) = \frac{\theta(1-t)}{\gamma}$\\\\$pk'(e\theta - y') =$\\$ -(1 - t - \gamma)$}\\
 		\hline
@@
 \section{Not enough undeclared money}
+Consider the case $e\theta - y' < \beta = c(e) - (1-t)\beta'$.
+Since it is less costly to pay the worker out of undeclared money, it is worth paying as much as possible out of undeclared money and then pay the remaining amount out of declared profit.
+That means that
+\begin{align*}
+e\theta - y' = (1 - \alpha) \beta\\
+\alpha = \left(1 - \frac{e \theta - y'}{\beta}\right)\cdot\mathbb{I}_{\{e \theta - y' < \beta\}}
+\end{align*}
+where
+\begin{align*}
+\mathbb{I}_{\{e \theta - y' < \beta\}} = \begin{cases}
+1, e\theta - y' < \beta\\
+0, \text{ otherwise}
+\end{cases}
+\end{align*}
+The firm's profit becomes:
+\begin{align*}
+\pi = (y' - \beta')(1 - t) - \beta\mathbb{I}_{\{e \theta - y' < \beta\}} + (e \theta - y')\mathbb{I}_{\{e \theta - y' < \beta\}} + \gamma(e\theta - y' - \beta + \beta \mathbb{I}_{\{e \theta -y' < \beta\}} -\\ -\mathbb{I}_{\{e \theta - y' < \beta\}}(e\theta - y')) - pk(e \theta - y')
+\end{align*}
+Using (IC) $\beta = c(e) - \beta'(1-t)$ the profit function is
+\begin{align*}
+\pi = (y'-\beta')(1-t) +(\mathbb{I}_{\{e \theta - y' < \beta\}} + \gamma(1-\mathbb{I}_{\{e \theta - y' < \beta\}}))(e \theta -y' - c(e) + (1-t)\beta')-pk(e \theta - y')
+\end{align*}
+Denote $\mathbb{I}_{\{e \theta - y' < \beta\}} + \gamma(1-\mathbb{I}_{\{e \theta - y' < \beta\}})$ as $\xi$. Note that
+\begin{align*}
+\xi = \begin{cases}
+1, e\theta - y' < \beta\\
+\gamma, \text{ otherwise}
+\end{cases}
+\end{align*}
+\subsection{No credibility problems}
+Without enforceability problems, FOCs are
+\begin{align*}
+\frac{\partial \pi }{\partial e} &= -\xi c'(e) + \xi \theta - \theta p k'(e\theta - y') = 0\\
+\frac{\partial \pi}{\partial y'} &= 1-t - \xi + pk'(e \theta - y') = 0\\
+c'(e) &= \theta(1-t)\\
+pk'(e \theta -y') &= -(1 - t - \xi)
+\end{align*}
+\subsection{Enforceability problems}
+The (DE) constraint is
+\begin{align*}
+\frac{\delta \pi}{1 - \delta} \ge (1 - t)\beta' + \alpha \beta + \gamma(1 - \alpha) \beta
+\end{align*}
+\begin{align*}
+\delta (y'-\beta')(1-t) +\delta \xi(e \theta -y' - c(e) + (1-t)\beta')-\delta pk(e \theta - y') \ge (1 - \delta)(1 - t)\beta' +\\
++ (1-\delta)(\alpha + \gamma(1 - \alpha))\beta\\
+\delta y'(1-t) +\delta \xi(e \theta -y')-\delta pk(e \theta - y') \ge ((1-\delta)(\alpha + \gamma(1 - \alpha)) +\delta \xi)c(e) +\\ +(1 - \delta \xi + (1 - \delta)(\alpha + \gamma(1 - \alpha)))(1-t)\beta'
+\end{align*}
+The problem becomes
+\begin{align*}
+\max\ &(y'-\beta')(1-t) +\xi(e \theta -y' - c(e) + (1-t)\beta')-pk(e \theta - y')\\
+s.t.\ &\delta y'(1-t) +\delta \xi(e \theta -y')-\delta pk(e \theta - y') \ge ((1-\delta)(\alpha + \gamma(1 - \alpha)) +\delta \xi)c(e) +\\&+(1 - \delta \xi + (1 - \delta)(\alpha + \gamma(1 - \alpha)))(1-t)\beta'\to \lambda \ (DE)\\
+& e \theta - y' < c(e) - (1-t)\beta'\ \to \mu \\
+& e\theta - y' \ge 0\ \to \eta\\
+& \beta' \ge 0, y' \ge 0, e \ge 0\ \to \psi_1, \psi_2,\psi_3
+\end{align*}
+\begin{align}
+\frac{\partial \pi}{\partial y'} &= 1-t-\xi + pk'(e \theta - y') + \lambda \delta (1 - t - \xi) + \lambda \delta pk'(e \theta - y') + \mu - \eta + \psi_2 = 0 \label{eq8}\\
+\frac{\partial \pi}{\partial e} &= -\xi c'(e)+\delta \xi \theta - \theta pk'(e\theta - y')- \lambda (\delta\xi + (1-\delta)(\alpha + \gamma(1 - \alpha))) c'(e) + \lambda \delta \theta - \nonumber\\ 
+&-\lambda \delta \theta pk'(e \theta - y') - \mu \theta + \mu c'(e) + \eta \theta+ \psi_3 = 0\label{eq9}\\
+\frac{\partial \pi}{\partial \beta'} &= -(1-\xi)(1-t) - \lambda(1 - \delta \xi + (1 - \delta)(\alpha + \gamma(1 - \alpha)))(1-t) - \mu(1-t) + \psi_1 = 0\label{eq10}\\
+& \lambda, \mu, \eta, \psi_1, \psi_2, \psi_3 \ge 0 \label{eq11}\\
+&+\text{ slackness complementary conditions}\nonumber
+\end{align}
+Assume firstly that $\beta', y', e >0$, also suppose that $e\theta - y' < \beta \to \xi = 1$, that means $\psi_1 = \psi_2=\psi_3 = 0$, consequently from \eqref{eq10} and \eqref{eq11}, $\lambda = \mu = 0$.
+If $e \theta - y' > 0$ then $\eta = 0$ and the solution is equivalent to the case without credibility problems.
+Thus, the optimal level of effort and the amount of declared revenue are determined by
+\begin{align*}
+c'(e) &= \theta(1 - t)\\
+pk'(e\theta - y') &= t
+\end{align*}
 
+Now assume that $\psi_1 > 0\ \to \beta' = 0$, implying $\mu = 0$, $\lambda = \frac{\psi_1}{(1 - \delta \xi + (1 - \delta)(\alpha + \gamma(1 - \alpha)))(1-t)}$$> 0$ hence the (DE) constraint binds.
+Then from the FOCs
+\begin{align*}
+c'(e) &= \frac{(1 + \lambda \delta)\theta(1-t)}{\xi+\lambda(\delta\xi + (1 - \delta)(\alpha + \gamma(1 - \alpha)))}\\
+pk'(e\theta - y') &= -(1 - t - \xi)
+\end{align*}
+and $\lambda$ can be found from the (DE) constraint.
 \end{document}

### 2019-05-16 metrics3_1.tex
Metrics

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
	{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\newcommand{\yrseduc}{\textit{yrseduc}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
		\begin{tabular}{r}
			Danil Fedchenko, MAE 2020, group A \\
		\end{tabular}
	\end{flushright}

	\begin{center}
		Econometrics 3. Problem Set 1.
	\end{center}
	\section*{Problem 1}
	5.8
	\begin{enumerate}[(a)]
		\item $\expect^{**}[Y|X] = \gamma X$ where $\gamma \in \underset{c}{\text{ argmax }} \expect[(Y - cX)^2]$. FOC:
		\begin{align*}
		\frac{\partial }{\partial c}: -2 \expect[(Y - \gamma X)X] = 0\\
		\gamma = \frac{\expect[YX]}{\expect[X^2]}
		\end{align*}
		\item 
		\begin{align*}
		\expect[\expect^{**}[Y|X] - Y] = \expect\left[\frac{\expect[YX]}{\expect[X^2]}X - Y\right] = \frac{\expect[XY]\expect[X]}{\expect[X^2]} - \expect[Y]
		\end{align*}
		the expression above in general is not equal to 0, hence it is a biased predictor.
		\item \begin{align*}
		Cov(X, Y - \gamma X) = Cov(X, Y) - \gamma Var(X) = \expect[XY] - \expect[X]\expect[Y] - \expect[XY] +\\+ \frac{\expect[XY](\expect X)^2}{\expect[X^2]} = \frac{\expect[XY](\expect X)^2}{\expect[X^2]} - \expect[X]\expect[Y] \neq 0
		\end{align*}
		\item \begin{align*}
		\expect\left[(Y - \frac{\expect[YX]}{\expect[X^2]}X)^2\right] = \expect[Y^2] - \frac{(\expect[XY])^2}{\expect[X^2]}
		\end{align*}
		\item When the BLP is used the minimized value $\expect[U^2]$ cannot be higher, since $\alpha + \beta X$ is the special case of $\gamma X$ with the constraint $\alpha = 0$. Of course, the more constrained minimization problem results in greater or equal solution comparing to the less constrained problem. If $\expect[Y]$ is used then
		\begin{align*}
		\expect[U^2] = \expect[(Y - \expect[Y])^2] = Var(Y) = \expect[Y^2] - (\expect[Y])^2
		\end{align*}
		and the result is ambiguous because if for example $Y = const$ (and $X \neq const$) then of course the marginal expectation method results in smaller $\expect[U^2]$, on the other hand, if $Y = X \neq const$ then the best proportional predictor yields the smallest $\expect[U^2]$.  
	\end{enumerate}

6.7

\begin{enumerate}[(a)]
	\item \begin{align*}
	\expect[Y|X] = \expect[W|X] + X^2\expect[W|X] = 1 + X^2
	\end{align*}
	since $X$ and $W$ are independent and $\expect[W|X] = \expect[W] = 1$.
	\begin{align*}
	&\expect^*[Y|X] = \alpha + \beta X\\
	\text{where } &\beta = \frac{cov(X, Y)}{Var(X)},\ \alpha = \expect[Y] - \beta \expect[X]
	\end{align*}
	\begin{align*}
	&cov(X, Y) = cov(X, W + WX^2) = cov(X, WX^2) = \expect[W]\expect[X^3] - &\expect[X]\expect[W]\expect[X^2] = 0\\
	&\expect[Y] = 1 + 1 = 2\\
	&\expect^*[Y|X] = 2
	\end{align*}
	\item 
	\begin{align*}
	\expect[Y|X] = 1 + X^2
	\end{align*}\begin{align*}
	cov(X, Y) = 1,\ Var(X) = 1\ \to \beta &= 1\\
	\alpha &= 1\\
	\expect^*[Y|X] &= 1 + X
	\end{align*}
	\item CEF does not change because it does not depend on distribution of $X$, whereas $\expect^*[Y|X]$ does depend on the distribution of $X$ and it changes.
\end{enumerate}
	\section*{Problem 2}
	Let a random triple $(y,\ x_1,\ x_2)^T$ be distributed as
	\begin{align*}
	\norma \left( \begin{pmatrix}
	0\\
	0\\
	0
	\end{pmatrix}, \begin{pmatrix}
	1 & \rho_{y1} & \rho_{y2}\\
	\rho_{y1} & 1 & \rho_{12}\\
	\rho_{y2} & \rho_{12} & 1
	\end{pmatrix} \right)
	\end{align*}
	where $\rho_{y1}, \rho_{y2}, \rho_{12}$ are unknown, but it is known that they are not zero. A researcher claims
	that $\expect[y|x_1, x_2] = \alpha_1 x_1 +\alpha_2x_2$ is the true regression, while $\expect[y|x_1] =\alpha_1x_1$ and $\expect[y|x_2] =\alpha_2x_2$ are not because of the omitted variables bias. Please comment on this claim.

	\textbf{Solution}

	It is known that $y|x_1,x_2 \sim \norma(\mu, \Sigma)$ where
	\begin{align*}
	\mu = \begin{pmatrix}
	\rho_{y1} & \rho_{y2}
	\end{pmatrix} \begin{pmatrix}
	1 & \rho_{12}\\
	\rho_{12} & 1
	\end{pmatrix}^{-1} \begin{pmatrix}
	x_1\\
	x_2
	\end{pmatrix} = \frac{\rho_{y1} - \rho_{y2}\rho_{12}}{1 - \rho_{12}^2}x_1 + \frac{\rho_{y2} - \rho_{y1}\rho_{12}}{1 - \rho_{12}^2}x_2 = \alpha_1 x_1 + \alpha_2 x_2
	\end{align*}
	as the researcher claims.  
	\begin{align*}
	\expect[y|x_1] = \expect[\expect[y|x_1, x_2]|x_1] = \alpha_1 x_1 + \alpha_2\expect[x_2|x_1]
	\end{align*}
	Since $\expect[x_2|x_1] = \rho_{12}x_1 \neq 0$,  $\expect[y|x_1] \neq \alpha_1x_1$ as the researcher claims. The same arguments can be applied to show that $\expect[y|x_2] \neq \alpha_2 x_2$.

	\section*{Problem 3}
	A researcher draws a random sample of $n = 100$ from the population of high school students
	and their results of a certain test. For each student $i = 1, \dots, n$ the researcher observes: $x_i$ - 
	the type of school (0 if public, 1 if private), and $y_i$ - the test score (1, 2, or 3). The empirical
	distribution of $(x; y)$ is the following:

	\begin{center} 
		\begin{tabular}{ c|ccc|c } 
			\hline 
			& Test Score 1&Test Score 2&Test Score 3&Totals\\ 
			\hline 
			Public Schools & 0.20& 0.50& 0.10 &0.80\\
			Private Schools& 0.05& 0.10& 0.05 &0.20\\
			\hline
			Totals & 0.25 &0.60 &0.15 &1.00\\
			\hline
		\end{tabular} 
	\end{center}
\begin{enumerate}
	\item Provide an analog estimate the optimal predictor (under MSPE) of $y$ given $x$:
	\item Let $\mathbb{BLP}[x|y]$ denote the best linear predictor (under MSPE) of $x$ given $y$.
	Let $\widehat{\mathbb{BLP}}[x|y]$ denote the sample analog of $\mathbb{BLP}[x|y]$. Compute
	\begin{align*}
	\widehat{\mathbb{BLP}}[x|y = 3] - \widehat{\mathbb{BLP}}[x|y = 1]
	\end{align*}
	\item Discuss the statement: "The data indicate that enrolling in a private school
	rather than a public school substantially increases the chance that a student obtains the
	highest test score. The estimated effect of private school enrollment is to increase the
	probability of scoring 3 from $\frac{1}{8}$ to $\frac{1}{4}$"
\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item $\mathbb{BLP}[y|x] = \alpha + \beta x$. Where
	\begin{align*}
	\beta = \frac{cov(x, y)}{Varx}, \alpha = \expect[y] - \beta \expect[x]
	\end{align*}
	Due to analogy principle, $cov(x, y), Varx, \expect[y], \expect[x]$ can be estimated by using the empirical distribution. Namely,
	\begin{align*}
	\widehat{\expect[y]} &= \frac{1}{100} \left(25 + 2 \cdot 60 + 3 \cdot 15\right) = 1.9\\
	\widehat{\expect[x]} &= 0.2
	\end{align*}
	\begin{align*}
	&n^2\widehat{cov}(x, y) = 20(0 - 0.2)(1 - 1.9) + 50(0-0.2)(2-1.9) + 10(0-0.2)(3-1.9) + \\ &+5(1-0.2)(1-1.9) + 10(1-0.2)(2-1.9)+5(1-0.2)(3-1.9) = 2\\
	&n^2\widehat{Var}x = 80(0-0.2)^2 + 20(1-0.2)^2 = 16
	\end{align*}
	\begin{align*}
	\hat{\beta} = \frac{1}{8},\ \alpha = 1.875\\
	\mathbb{BLP}[y|x] = \frac{15}{8} + \frac{1}{8}x
	\end{align*}
	\item $\mathbb{BLP}[x|y] = \gamma + \delta y$ where
	\begin{align*}
	\delta = \frac{cov(x, y)}{Var(y)},\ \gamma = \expect[x] - \delta \expect[y]
	\end{align*}
	\begin{align*}
	\widehat{\mathbb{BLP}}[x|y = 3] - \widehat{\mathbb{BLP}}[x|y = 1] = 2 \hat{\delta}
	\end{align*}
	$n^2\widehat{Var}(y) = 25(1-1.9)^2+60(2-1.9)^2+15(3-1.9)^2 = 39$.
	\begin{align*}
	\widehat{\mathbb{BLP}}[x|y = 3] - \widehat{\mathbb{BLP}}[x|y = 1] = 2 \hat{\delta} = \frac{4}{39}
	\end{align*}
	\item In general the statement is false, because there can be a lot of threats for the internal validity which are not taken into account. For example, there can be omitted variables which bias the estimate, i.e. some other factors apart from type of school affect the students' performance. Moreover there can be an endogeneity problem which also biases the estimate and threats the internal validity of the model.
	\section*{Problem 4}
	See the solution in Ipython Notebook file.
\end{enumerate}
\end{document}

### 2019-09-02 metrics4_1.tex
metrics

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
	{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathcal{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\newcommand{\yrseduc}{\textit{yrseduc}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
		\begin{tabular}{r}
			Danil Fedchenko, MAE 2020\\
		\end{tabular}
	\end{flushright}

	\begin{center}
		Econometrics 4. Problem Set 1.
	\end{center}
	\section*{Problem 1}
	\textbf{Solution}

	\begin{enumerate}[a.]
		\item We need to show that $\theta_0$ minimizes $\expect \{(y - m(x, \theta))^2|x\}$ over $\Theta$ for any $x$. Using equation (12.4):
		\begin{align*}
		&\expect \{(y - m(x, \theta))^2|x\} = \expect \{(y - m(x, \theta_0) + m(x, \theta_0) - m(x, \theta))^2|x\} =\\
		&= \expect\{(y - m(x, \theta_0))^2|x\} + 2\expect\{(m(x, \theta_0) - m(x, \theta))u|x\} + \expect\{(m(x, \theta_0) - m(x, \theta))^2|x\} = \\
		&= \expect\{(y - m(x, \theta_0))^2|x\} + 2(m(x, \theta_0) - m(x, \theta))\expect\{u|x\} + (m(x, \theta_0) - m(x, \theta))^2
		\end{align*}
		Since $\expect[u|x] = 0$ the expression above is equal to
		\begin{align*}
		\expect\{(y - m(x, \theta_0))^2|x\} + (m(x, \theta_0) - m(x, \theta))^2
		\end{align*}
		which is apparently positive if $\theta \neq \theta_0$, while for $\theta = \theta_0$ it is equal to 0. That is why it attains minimum once $\theta = \theta_0$.
		\item The result in a. is stronger because it states that $\theta_0$ minimizes the average (over $y$) square of deviation between $y$ and $m(x, \theta)$ for each particular $x$, whereas the statement that $\theta_0$ minimizes $\expect[(y - m(x, \theta))^2]$ over $\Theta$ claims that $\theta_0$ minimizes the average square of deviation between $y$ and $m(x, \theta)$ and the average is taken both on $y$ and $x$.
	\end{enumerate}

\section*{Problem 2}
Suppose $y = \beta x + \varepsilon$ where all variables are scalars, $x$ and $\varepsilon$ are independent,
and the distribution of $\varepsilon$ is symmetric around $0$. For a random sample $\{x_i,y_i\}^n_{i=1}$ consider the
following extremum estimator of $\beta$:
\begin{align*}
\hat{\beta} = \underset{b}{argmax} -\sum_{i=1}^n exp(y_i-bx_i) + exp(-y_i + bx_i)
\end{align*}
\begin{enumerate}[a)]
\item Derive the asymptotic properties of $\hat{\beta}$, paying special attention to the identification
condition.
\item Compare this estimator with the OLS estimator in terms of asymptotic efficiency for the
case when $x$ and $\varepsilon$ are normally distributed.
\end{enumerate}

\textbf{Solution}

\begin{enumerate}[a)]
	\item If the identification condition
	\begin{align*}
	\expect[e^{\varepsilon} + e^{-\varepsilon}] < \expect[e^{\varepsilon}e^{x(\beta-b)} + e^{-\varepsilon}e^{-x(\beta - b)}], \text{ for all }b \neq \beta
	\end{align*}
	holds, then $\hat{\beta} \overset{p}{\to} \beta$. Since $\varepsilon$ and $x$ are independent and moreover $\varepsilon$ and $-\varepsilon$ have similar distributions due to symmetry, then the identification condition becomes
	\begin{align*}
	2 \expect[e^{\varepsilon}] < \expect[e^{\varepsilon}] \expect[e^{x(\beta - b)} + e^{-x(\beta - b)}], \text{ for all }b \neq \beta
	\end{align*}
	Since $e^{\varepsilon}$ is almost surely positive then so is its expectation. Hence the condition above becomes
	\begin{align*}
	\expect[e^{x(\beta - b)} + e^{-x(\beta - b)}] > 2, \text{ for all }b \neq \beta\\
	\expect\left[\frac{(e^{x(\beta-b)} - 1)^2}{e^{x(\beta - b)}}\right] > 0, \text{ for all }b \neq \beta
	\end{align*}
	Both numerator and denominator are almost surely positive hence so is their ratio. Thus, the identification condition holds which (coupled with some technical conditions for the validity of the uniform weak law of large numbers, which also hold for the exponential function) implies $\hat{\beta} \overset{p}{\to} \beta$.

	Additionally, the conditions for the asymptotic normality hold and that means that
	\begin{align*}
	\sqrt{n}(\hat{\beta} - \beta) \overset{d}{\to} \norma(0, A_0^{-1}B_0A_0^{-1})
	\end{align*}
	where
	\begin{align*}
	&A_0 = \expect[x^2(e^{\varepsilon} + e^{-\varepsilon})] = 2\expect[x^2] \expect[e^{\varepsilon}]\\
	&B_0 = \expect[x^2e^{2\varepsilon} + x^2e^{-2\varepsilon} - 2x^2] = \expect[x^2]\expect[(e^{\varepsilon} - e^{-\varepsilon})^2] = 2\expect[x^2](\expect[e^{2\varepsilon}] - 1)
	\end{align*}
	Thus,
	\begin{align*}
	\sqrt{n}(\hat{\beta} - \beta) \overset{d}{\to} \norma\left(0, \frac{\expect[e^{2 \varepsilon}]-1}{2\expect[x^2]\expect[e^{\varepsilon}]^2}\right)
	\end{align*}
	\item Suppose $x \sim \norma(\mu, \sigma_x^2)$ (wlog further assume that $\mu = 0$, otherwise we can introduce an intercept to the model) and $\varepsilon \sim \norma(0, \sigma^2_{\varepsilon})$ then
	\begin{align*}
	\sqrt{n}(\hat{\beta} - \beta) \overset{d}{\to} \norma\left(0, \frac{\sigma_{\varepsilon}^2}{\sigma^2_{x}}\right)
	\end{align*}
	Moreover, $2\varepsilon \sim \norma(0, 2\sigma^2_{\varepsilon})$ which means that the asymptotic variance for the estimator from a) is 
	\begin{align*}
	\frac{e^{\sigma^2_{\varepsilon}}-1}{2\sigma^2_xe^{\sigma^2_{\varepsilon}}}
	\end{align*}
	Since $\sigma^2_{\varepsilon} > 0$ then 
	\begin{align*}
	e^{-\sigma^2_{\varepsilon}} &> 1 - \sigma^2_{\varepsilon}\\
	\frac{1}{2} - \frac{1}{2e^{\sigma^2_{\varepsilon}}} &< \frac{\sigma^2_{\varepsilon}}{2} < \sigma^2_{\varepsilon}
	\end{align*}
	that means that the asymptotic variance of the OLS estimate is greater, in other words, the estimator from a) is asymptotically more efficient than the OLS estimator.
\end{enumerate}
\end{document}

### 2019-11-06 Draft.tex
Draft

\documentclass[a4paper]{article}
\usepackage[12pt]{extsizes} % 
\usepackage{setspace}
\doublespacing
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage{makecell}
\usepackage{amsthm}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{xcolor,colortbl}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathcal{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=15mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\begin{document} %

	\section{Question}
	Suppose we have a web-platform that is designed to implicitly match customers and service providers via the past-user-based recommendation (examples of such a kind of platforms include \textit{TripAdvisor, Yelp, etc}). The main goal of the platform is to render users satisfied with the quality of consumed products (in order to keep users using the platform). Users' reviews help the platform to infer about the qualities of different providers and based on this, to recommend users better options. However, once a new provider appears, the platform usually does not know the quality of the good provided, since there are no reviews yet. And the natural desire of the platform in this case is to figure out the quality of new alternative as quickly as possible in order to use (or not) it in recommendation for users. But how can the platform achieve this? Well, it seems that the platform can just start recommend the new alternative to users, collect reviews and then infer about the quality. But obviously the platform cannot do it too often and with too risky options : users will just stop using the platform that recommends them things that eventually turn out to be inferior. And here the first trade-off arises: how to induce exploration of new alternative without "sacrifice" customers too much? There are several papers on this topic that shed light on this trade-off and propose the optimal policy. The main idea in all those papers is that there is an optimal level of information obfuscation to create incentives for exploring new things.

	However, in all previous works the quality of alternatives was exogenous to the process of recommendation. But in real life, providers are also strategic agents (and in most cases they are way more sophisticated than customers) who can change quality of their products in response to the recommendation policy the platform implements, and taking this into account the recommendation policy itself (I think) should change! So, questions I would be interested to address are: \textbf{what would be the optimal information revelation (recommendation) policy in this case, and how it is affected by the presence of strategic providers? Whether the quality of products increases (compare to the case when alternatives are exogenous)?}

	With strategic alternatives the platform's and customers' preferences may become even more misaligned. Because now customers may be used by the platform not only for the sake of exploration (and to the benefit of subsequent agents) but also as a tool for encouraging the competition between alternatives.

	\section{Model}
	\subsection{In words}
	We have a platform that is designed for providing customers with some services, and two firms (providers) are operating on this platform (it can be restaurants, hotels etc.).

	 Customers arrive sequentially and each should decide on the product from which firm to choose. Although the qualities of the products are unknown \textit{ex ante}, customers have some priors and make their decision rationally taking into account all information available and updating their beliefs via the Bayes rule. After making a decision, each customer gets some reward, reports his experience to the platform (principal), and leaves the system.

	The platform accumulates agents' reviews and based on this can recommend a product from a particular firm to the subsequent agents. The goal of the platform is to maximize the (infinite horizon ?) expected discounted reward of all agents.

	Heterogeneous firms, in turn, are competing for agents and can invest to improve quality of their goods. The goal of each firm is to maximize its (infinite horizon ?) expected discounted profit taking into account the recommendation policy.

	At $t = 0$ each firm privately learns its type. Then, the timing of the $t$-th ($t=1, 2, \dots$) stage of the game is the following:
	\begin{enumerate}[1)]
		\item Each firm decides whether to stay in business or not (and incurs cost $f_i$ if decides to stay)
		\item Firms make private and costly investment in quality
		\item The unobserved quality is realized
		\item Customer arrives
		\item The principal makes a recommendation
		\item Customer decides which firm to choose
		\item Payoffs are realized and the principal observes customers' experience
	\end{enumerate}
	\subsection{Formalization}
	\subsubsection{Firms}
	We have two competing firms. Each firm has its own type that somehow is supposed to represent its "effectiveness" or "inclination to investing and sustaining the high quality of the good". The simplest example that comes in mind is that the type represents the firm's marginal cost of investment: the higher the type - the less effective the firm is. (types are either discrete $\theta^H, \theta^L$ or continuous $\theta \in (0, 1)$).

	\textbf{Question:} why do we need types?

	\textbf{Answer:} customers will learn and infer about the type of the firm

	Making investment $e \in \mathbb{R}_{+}$ (maybe make it discrete $e^L$ and $e^H$) in quality of its product, each firm incurs cost $c_i(e)$ that should depend on the firm's type (for example $c_i(e) = \theta_ie_i$ where $\theta_i$ is the type of the $i$-th firm). Moreover, if the firm decides to stay in business, it incurs fixed cost $f_i$ disregard whether the customer chooses the firm or not. $f_i$ is low enough to ensure that at least one firm will operate in any case (maybe fixed cost also should dependent on type).
	Each firm obtains a profit $1$ if the customer chooses its product. Denote for each firm $$d_{it} = \begin{cases}
	1, \text{if consumer }t\text{ chooses the firm }i\\
	0, \text{if consumer }t\text{ does not choose the firm }i\\
	\end{cases}$$
	and
	$$b_{it} = \begin{cases}
	1, \text{if the firm }i\text{ decided to stay in busimess at }t\\
	0, \text{if the firm }i\text{ decided not to stay in busimess at }t
	\end{cases}$$
	then each firm is maximizing $$\underset{\{e_{it}, b_{it}\}_{t=1}^{\infty} }{\max}\ \expect \left[\sum_{t=1}^{\infty} \delta^t b_{it}(d_{it} - c_i(e_{it})- f_i)\right]$$
	subject to the recommendation policy implemented by the principal.

	\textbf{Open questions:}

	can the firms observe the experience of users who chose the competitor? It seems reasonable if they could : the restaurant manager can send somebody to the nearby competitive restaurant to infer about the real quality.

	can the firm know that the competitor still in business?

	\underline{Maybe infinite-horizon model will be too hard, make it finite (start from two-periods)}

	\subsubsection{Products}
	Each firm provides a product which in time $t$ has quality $q_{it} \in (0, 1)$  which is endogenously formed from the firms' investment $e_{it}$ in this period.

	\textbf{Idea:} the quality should somehow represent consumers' satisfaction of the product, the higher quality - the higher (expected) satisfaction.

	\textbf{Realization:} Suppose that upon consuming the good from the firm $i$, the customer $t$ gets a reward $$r_{it} = \begin{cases}
	1, \text{ w.p. } q_{it}\\
	0, \text{ w.p } 1 - q_{it}
	\end{cases}$$
    The quality of the product from the firm $i$ at time $t$ is determined as $$q_{it} = \begin{cases} \frac{e_{it}}{e_{it}+e_{jt}},\ &j \neq i, e_{it}^2 + e_{jt}^2 \neq 0 \\
	0, &e_{it} = e_{jt} = 0
	\end{cases}
	$$

	\textbf{Question:} why is the quality of the alternative $i$ affected (negatively) by the investment made by the firm $j$?

	\textbf{One possible answer:} in real life there is no such a thing as an absolute quality. I think that good things are not good \textit{per se}, they are good \underline{in comparison} to other things. So, in this sense the assumption may seem to be reasonable.

	\textbf{Concern:} in this set-up if both firms make high and equal investment the probability of success for the customer is $\frac{1}{2}$, whereas if only one firm made high investment while the other made low, the probability of success for the first firm becomes higher than $\frac{1}{2}$. Does it represent the reality?

	\subsubsection{Principal}
	Principal observes the history of consumers' choices and their payoffs, $h^t$ (denote $H^t$ the whole set of such histories) and can commit to the recommendation policy, a function $R: H^t \to M^t$, where $M^t$ is the message space feasible for the customer $t$. The principal's goal is to design a revelation policy $R(\cdot)$ in order to maximize agents' expected discounted payoff:
	$$\pi = \underset{R(\cdot)}{\max}\ \expect\left[\sum_{t=1}^{\infty} \delta^t r_{t} \right]$$

	where $r_t$ is an expected reward of the customer $t$.

	\subsubsection{Customers}

	Customers (agents) arrive sequentially one by one, they do not observe the experience of other agents, but can receive information revealed by the principal. For each principal's message $m_t \in M^t$ the agent's strategy is a function $\sigma^t: M^t \to A$, where $A = \{a_1, a_2\}$ describes the set of possible alternatives (products from the firm 1 or 2). In other words, the agent's strategy prescribes for each principal's message an action, the product from which firm to choose. Choosing the strategy, each agent $t$ maximizes
	$$\underset{\{a_1, a_2\}}{\max}\ \expect_t \left[q_{it}|m_t \right]$$

	\textbf{Questions}: what are priors? which alternative do the customers choose when indifferent between two?

\section{Analysis}
\subsection{Straightforward observations}
First of all, it seems that in this model it will never be optimal (starting from some point) to start recommending only one firm. This is because in this case the firm which is not recommended leaves the business (because of fixed cost) and this eliminates any incentive for the incumbent to invest. Hence, the quality becomes $0$ and all subsequent customers will get 0. So, the platform should internalize its role in sustaining a competition.
\subsection{No disclosure at all}
Suppose that the platfotm employs the non-disclosure policy. That is, agents do not possess any information regarding the previous users' experience. In this case, agents make decisions based on priors. If customers are homogeneous and have identical priors about the quality, then the firm which the customers are $a\ priori$ inclined towards stays in business with zero investment while the second firm exits the market. In this equilibrium all customers get zero, so the platform's payoff is zero.

As we can see, now with strategic agents the non-disclosure policy becomes much worse. One reason is that it completely eliminates any incentive to increase the quality of alternatives.

\subsection{Full disclosure}
\subsubsection{1 period, observable qualities}
Suppose we have only 1 period, at the beginning of the period, firms privately draw its types $\theta$ according to some full-support continuous distribution with the p.d.f. $f(\theta)$. Then both firms simultaneously perform investment. After that the customer arrives and can observe realized qualities. The customer chooses the firm with higher quality (ties do not happen almost surely).
In this case the following Lemma is true
\begin{lemma}
	Suppose $\theta \sim U(0, 1)$ then
	in the one period simultaneous-move game with observable qualities there exist a symmetric equilibria in which both firms invest $e(\theta) = -\ln(\theta)$.
\end{lemma}
\begin{proof}
	See Appendix
\end{proof}

\subsubsection{Many periods, observable qualities}

If there are more than 1 period then in all periods after the first one the firm with higher type will learn that its type is high (because the customer did not choose it in the first period) and that is why the aforementioned equilibrium breaks down: less effective firm is better-off if it leaves the business, but if one of the firm leaves then the other has no incentive to invest: it becomes a monopoly. So let us note this result
\begin{lemma}
	Suppose $\theta \sim U(0, 1)$ then in many-periods simultaneous-move game with observable qualities there exists an equilibrium in which:
	\begin{enumerate}
		\item In the first period both firms play $e(\theta) = -\ln(\theta)$
		\item In all subsequent periods:
		\begin{itemize}
			\item the more efficient firm (with smaller $\theta$) plays $e(\theta) = -\ln(\theta)$ if and only if there is a competitor in business, and $e(\theta) = 0$ otherwise
			\item the less efficient firm (with higher $\theta$) leaves the business forever.
		\end{itemize}
	\end{enumerate}
\end{lemma}
\begin{proof}
	See Appendix.
\end{proof}
As we can see, if qualities are observable in many periods game, unravelling may happen, and the principal's expected payoff becomes equal to the expected reward of the first agent $$\pi_0 = r_1 = \expect\left[\frac{\ln \theta_i}{\ln \theta_i + \ln \theta_j}\bigg| \theta_i < \theta_j\right] = \frac{3}{4},\ i \neq j$$

\subsection{Not full disclosure}
\subsubsection{Qualities are observable by the principal}
Can the principal do better? Let us take the simplest case, when the principal can observe the qualities.
Suppose the principal commits to recommend high quality firm with probability $q$ and low probability good with probability $1-q$. In this case it is easy to see that the symmetric equilibrium of one-period game is an equilibrium in many-periods game
\begin{lemma}
	Suppose $\theta \sim U(0, 1)$, $f \le 0.5$, the platform can observe the realized qualities and commits to recommend to the customer better option with probability $q \in (0.5, 1)$ and worse option with probability $1-q$. Then there exists a symmetric equilibium in which both firms are playing $e(\theta) = -(2q - 1)\ln \theta$.

	Moreover, in this equilibrium the platform's expected payoff is strictly greater than the platform's expected payoff $\pi_0$ in case of unravelling (Lemma 2) 
\end{lemma}
\begin{proof}
	See Appendix.
\end{proof}

As we can see, the platform can encourage competition.

\section{Extensions}
\begin{enumerate}
	\item Agents do not know they place in line.
	\item Heterogeneous customers
\end{enumerate}
\subsection*{Appendix}

\begin{proof}[Proof of Lemma 1]
	Let us find the equilibrium strategy for firms in a form $\varepsilon(\theta)$ where $\varepsilon(\theta):(0, 1) \to \mathbb{R}$ is a decreasing function. Then each firm is maximizing $$\underset{e}{\max}\ \prob(\theta > \varepsilon^{-1}(e))(1 - \theta e - f) - \prob(\theta \le \varepsilon^{-1}(e))(\theta e + f)$$
	or using the fact that $$\prob(\theta \le x) = \begin{cases}
	1, x > 1\\
	x, x \in [0, 1]\\
	0, x < 0
	\end{cases}$$
	the firm is maximizing
	$$\underset{e}{\max}\ 1 - \varepsilon^{-1}(e) - \theta e - f$$
	F.O.C.:
	$$-\frac{1}{\varepsilon'(\varepsilon^{-1}(e))} = \theta$$
	Since the equilibrium we are interested in is symmetric, $\varepsilon(\theta) = e$ and hence
	$$\varepsilon'(\theta) = -\frac{1}{\theta}$$
	Assuming $\varepsilon(1) = 0$ we get $$\varepsilon(\theta) = -\ln\theta$$ 
\end{proof}

\begin{proof}[Proof of Lemma 3]

	If both firms stay in business in the particular period, then let us find the equilibrium strategy for firms in a form $\varepsilon(\theta)$ where $\varepsilon(\theta):(0, 1) \to \mathbb{R}$ is a decreasing function. Then each firm is maximizing 
	\begin{align*}\underset{e}{\max}\ q\{\prob(\theta > \varepsilon^{-1}(e))(1 - \theta e - f) + \prob(\theta \le \varepsilon^{-1}(e))(-\theta e - f)\} +\\+ (1-q)\{\prob(\theta > \varepsilon^{-1}(e))(-\theta e - f) + \prob(\theta \le \varepsilon^{-1}(e))(1 - \theta e - f)\}
	\end{align*}
	or using the fact that $$\prob(\theta \le x) = \begin{cases}
	1, x > 1\\
	x, x \in [0, 1]\\
	0, x < 0
	\end{cases}$$
	the firm is maximizing 
	$$\underset{e}{\max}\ q(1 - \varepsilon^{-1}(e)) + (1 - q)\varepsilon^{-1}(e) - \theta e - f$$
	F.O.C.: $$-\frac{q}{\varepsilon'(\varepsilon^{-1}(e))} + \frac{1-q}{\varepsilon'(\varepsilon^{-1}(e))} - \theta = 0$$
	Since the equilibrium we are interested in is symmetric, $\varepsilon(\theta) = e$ and hence
	$$\varepsilon'(\theta) = -\frac{2q - 1}{\theta}$$
	Assuming $\varepsilon(1) = 0$ we get $$\varepsilon(\theta) = -(2q - 1)\ln\theta$$.

	It remains to check that less effective firm stays in business. Expected profit from staying in business for less effective firm is equal to
	$$\expect[1-q + (2q-1) \theta \ln \theta - f] = \frac{5}{4} - \frac{3}{2}q - f \ge 0$$

	The expected payoff of the platform is equal to
	$$\pi = \frac{q}{1-\delta} \expect\left[\frac{\ln \theta_i}{\ln \theta_i + \ln \theta_j}\bigg| \theta_i < \theta_j\right] +\frac{1-q}{1-\delta}\expect\left[\frac{\ln \theta_j}{\ln \theta_i + \ln \theta_j}\bigg| \theta_j \ge \theta_i\right] = \frac{2q+1}{4(1 - \delta)} > \frac{3}{4} = \pi_0$$
\end{proof}
\end{document}

### 2019-12-24 Draft.tex
Draft

@@
 \usepackage{pdfpages}
+\usepackage{tikz}
 \usepackage{makecell}
@@
 \newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{proposition}[theorem]{Proposition}
+\newtheorem{assumption}[theorem]{Assumption}
+\title{Optimal recommender system with strategic alternatives \\ Market Design. Term Project}
+\date{}
+\author{Danil Fedchenko}
 \begin{document} % 
+	\maketitle
+	\section{Introduction}
+	Recommender systems became an integral part of our lives.
+The most widely known examples of such a kind of services include: \textit{TripAdvisor} which presents ranking on Hotels and Restaurants for customers based on previous users' experience; \textit{Yelp} which provides a forum with reviews and ranking for different businesses; \textit{Netflix} which recommends movies and series also based on information aggregated from different customers.
+All those services are very important since, on one hand, they provide people with information about options they otherwise may never be aware of, at no (or very low) cost.
+On the other hand, these services aggregate the information and may make rather precise recommendations to people which significantly reduce search cost for them.
+But here the interesting question arise: how to disclose the information, the platform aggregated, optimally?
+It turns out that the full disclosure may not be optimal from the social welfare point of view.
+This happens since users, who do not internalize the information externalities they impose to other agents, underexplore i.e.
+start to choose the safe option and do not try new things.
 
-	\section{Question}
-	Suppose we have a web-platform that is designed to implicitly match customers and service providers via the past-user-based recommendation (examples of such a kind of platforms include \textit{TripAdvisor, Yelp, etc}).
-The main goal of the platform is to render users satisfied with the quality of consumed products (in order to keep users using the platform).
-Users' reviews help the platform to infer about the qualities of different providers and based on this, to recommend users better options.
-However, once a new provider appears, the platform usually does not know the quality of the good provided, since there are no reviews yet.
-And the natural desire of the platform in this case is to figure out the quality of new alternative as quickly as possible in order to use (or not) it in recommendation for users.
-But how can the platform achieve this?
-Well, it seems that the platform can just start recommend the new alternative to users, collect reviews and then infer about the quality.
-But obviously the platform cannot do it too often and with too risky options : users will just stop using the platform that recommends them things that eventually turn out to be inferior.
-And here the first trade-off arises: how to induce exploration of new alternative without "sacrifice" customers too much?
-There are several papers on this topic that shed light on this trade-off and propose the optimal policy.
-The main idea in all those papers is that there is an optimal level of information obfuscation to create incentives for exploring new things.
+	Several papers studied this question in different frameworks (see Literature review section) and the main conclusion that all those paper draw is that there should be an optimal level of information obfuscation in order to induce short-living and myopic users to experiment.
+However, in the previous work the quality of alternatives was supposed to be exogenous to the process of recommendation and learning.
+So, the recommender policy the platform implements cannot affect the realized quality of alternatives that users choose.
+But in reality, these alternatives are also strategic agents (think about restaurants or hotels), and in most cases these agents are way more sophisticated than users.
+These strategic agents may adjust the quality of the good to the recommender policy the platform is employing.
+And the platform, in turn, should take the strategic nature of alternatives into account once developing the policy.
+To my best knowledge, there are no papers that study the optimal recommender policy with strategic alternative, so this work may be considered as the first small step in this direction.
 
-	However, in all previous works the quality of alternatives was exogenous to the process of recommendation.
-But in real life, providers are also strategic agents (and in most cases they are way more sophisticated than customers) who can change quality of their products in response to the recommendation policy the platform implements, and taking this into account the recommendation policy itself (I think) should change!
-So, questions I would be interested to address are: \textbf{what would be the optimal information revelation (recommendation) policy in this case, and how it is affected by the presence of strategic providers?
-Whether the quality of products increases (compare to the case when alternatives are exogenous)?}
+	I build a model that embraces three main counterparties: the platform, the firm and users.
+Users are choosing between the good of \textit{ex-ante} unknown quality provided by the firm and a safe outside option.
+The firm can exert costly private effort to increase the quality of the service and the platform design the optimal recommender policy taking this into account.
+I demonstrate that the optimal policy is different once the firm becomes strategic.
+Also, I derive the optimal recommender policy in the class of simple "1-round" policies.
+The rest of the paper is organized as follows: the next section presents the literature review, section 3 describes the model, section 4 contains the main results, and last three sections embrace a brief discussion, possible extensions and directions for the future research, and a short conclusion.
+	\section{Literature Review}
+	This work is related to the strand of literature on recommender systems.
+One of the first papers which studied the optimal design of recommender system is the work by Kremer et al \cite{kremer2014}.
+The authors build a simple model with two alternatives which generate deterministic but \textit{ex-ante} unknown payoff.
+One of the alternatives is \textit{a priori} better, in the sense that the expected reward it generates is higher.
+The finite number of short-living consumers arrive sequentially and each chooses between either of two alternatives.
+The platform accumulates review and commits to the recommendation policy it uses to make recommendations to the newcomers.
+The papers, firstly, neatly in simple terms demonstrates why the full revelation policy may be suboptimal.
+Secondly, the optimal recommender policy is derived which has a "threshold" nature.
+That is, \textit{a priori} better alternative is recommended if its quality turns out to be higher than the certain threshold.
+Otherwise the other alternative is recommended, and then once both alternatives were tried the best one is recommended until the end.
+The setup of this model is very similar to my, however, it has a clear distinction.
+In my model the quality of the alternative is endogenous to the process of recommendation.
+This suggests that the threshold-like type of policies cannot be optimal in the presence of strategic alternatives because once the firm knows that it will be recommended forever, it does not have any incentives to maintain the high quality of the service it provides.
 
-	With strategic alternatives the platform's and customers' preferences may become even more misaligned.
-Because now customers may be used by the platform not only for the sake of exploration (and to the benefit of subsequent agents) but also as a tool for encouraging the competition between alternatives.
+	Another relevant paper is that of Che and Horner \cite{che2015}.
+In the paper a continuous flow of customers arrives, and each should choose between the good or an outside option.
+The quality of the good is exogenously predetermined but unknown.
+Customers who decide to consume the good send noisy signal about its quality to the platform.
+The signal is noisy in the sense that even if the quality of the good is good the information about it reaches to the platform only with some probability (different from 1).
+The platform updates its beliefs about the quality and tailors recommender policy contingent on the beliefs.
+Authors demonstrate that the full revelation policy is suboptimal and find the optimal policy which again has a threshold nature.
+Namely, once the platform receives the signal that the quality is good, it will recommend the good until the end.
+Otherwise, if no news arrive it is optimal to experiment with the risky alternative (recommend it) until the beliefs that its quality is good fall below the certain threshold.
+After that the optimal policy never recommends this alternative.
+
+	Finally, Papanstasiou et al \cite{papanastasiou2017} extended ideas developed in \cite{kremer2014} and \cite{che2015} and build a dynamic, stochastic model.
+In their model, agents also arrive sequentially to the platform and should choose between two alternatives that generate stochastically exogenous payoff.
+The platform in the paper designs a recommendation policy.
+Although the optimal policy turns out to have a complicated form, authors demonstrate that in this case the full revelation policy is not the best one, and the optimal policy involves information obfuscation.
+
+	The common feature of all aforementioned models is some form of exogeneity of the quality of the alternative.
+In the next section I present a model with the strategic firm that can adjust the quality of the good to the learning process and recommendation policy.
 
 	\section{Model}
-	\subsection{In words}
-	We have a platform that is designed for providing customers with some services, and two firms (providers) are operating on this platform (it can be restaurants, hotels etc.).
+	In the model one firm is operating on the platform and provides customers with the service of ex-ante unknown quality.
+Unit mass of customers arrives sequentially, each customer decides between the risky good provided by the firm and the safe outside option (customers differ in their outside option).
+Customers base their decision on the recommendation that is sent by the recommender system.
+Those who consume the good report the quality to the platform and leave forever.
+The platform accumulates reviews and commits to the recommender policy that can be a function of past reviews.
 
-	 Customers arrive sequentially and each should decide on the product from which firm to choose.
-Although the qualities of the products are unknown \textit{ex ante}, customers have some priors and make their decision rationally taking into account all information available and updating their beliefs via the Bayes rule.
-After making a decision, each customer gets some reward, reports his experience to the platform (principal), and leaves the system.
+	Three subsequent subsections in details describe all parties of the model.
+	\subsection{Firm}
+	The quality of the good $\mathcal{Q}_t$ provided by the firm at period $t$ can be either good or bad, $\mathcal{Q}_t \in \{g, b\}$.
+I will refer to the quality of the good as "the state of the firm".
+Following this notation, the firm can be either in good state or in bad.
+The firm can exert costly effort which can increase the probability to remain in the good state (if the current state is good) or to move to the good state (if the current state is bad).
+Without efforts with probability 1 firms either remains in the bad state (if the current state is bad) or moves to the bad state (if the current state is good).
+At each state the firm chooses the level of effort $e\in\{0, 1\}$ which will be exerted by the firm before the customers arrive.
+For example, if the firms chooses effort only based on its current state, then $e = \{e^b, e^g\}$.
+Using the notation from the theory of Markov chains, the transitory matrix in this case looks as follows $$P = \begin{pmatrix}
+	1 - pe^{b} & pe^{b}\\
+	1-pe^{g} & pe^{g}
+	\end{pmatrix}$$
+	here the first row refers to the bad state and the second to the good state, $p$ is the exogenous parameter which reflects the probability that investment will be successful.
+For example, the element $(1, 2)$ of the matrix represents the probability that the firm will move to the state $g$ if its current state is $b$.
 
-	The platform accumulates agents' reviews and based on this can recommend a product from a particular firm to the subsequent agents.
-The goal of the platform is to maximize the (infinite horizon ?) expected discounted reward of all agents.
+	If at time $t$ the share of the customers $s_t \in [0, 1]$ decides to consume the good then the firm gets $$\pi_t = s_t - ce$$ where $c$ is the cost of effort and $e \in \{0, 1\}$.
+The firm's objective is to choose an investing strategy that would maximize $$\expect_0\left[\sum_{t=0}^{\infty} \delta^t \pi_t \right]$$ subject to the recommender policy implemented by the platform.
+	\subsection{Customers}
 
-	Heterogeneous firms, in turn, are competing for agents and can invest to improve quality of their goods.
-The goal of each firm is to maximize its (infinite horizon ?) expected discounted profit taking into account the recommendation policy.
+	The unit mass of heterogeneous customers arrives sequentially each period.
+Each customer has the outside option $r_i \in [0, 1]$.
+Upon consuming the good customers derive the utility $$v_t = \begin{cases}
+	1, \text{ if }\mathcal{Q}_t = g\\
+	q, \text{ if }\mathcal{Q}_t = b
+	\end{cases}$$
+	where $q \in (0, 1)$. The total utility of the customer who has an outside option $r_i$ and arrives at time $t$ is determined as follows $$u_{it} = \begin{cases}
+	v_t &,\text{if he chose the good}\\
+	r_{i} &,\text{if he chose an outside option}
+	\end{cases}$$
+	Choosing between outside option and the good, customers take into account a private signal received from the platform.
+This signal has a form of recommendation: the platform either recommends the customer trying the good or not.
+	\subsection{Platform}
+	At the beginning of the game the platform commits to the recommendation policy, that can be based on past reviews it collected.
+Then, each period the platform recommends the good to the share of customers $s_t$.
+Where $s_t$ is the mapping from the set of possible histories at time $t$ to $[0, 1]$.
+If at time $t$ the platform recommends the good to the share of customers $s_t$ and everybody follows the recommendation then the platform derives a utility $$\mathcal{U}_t = v_ts_t - \frac{s_t^2}{2}$$ which is basically an aggregate surplus obtained by customers who chose a good.
+The platform designs the recommendation policy that would maximize $$\expect_0\left[\sum_{t=0}^{\infty} \delta^t \mathcal{U}_t\right]$$ subject to the firm's strategy $\sigma(s_t(h^t))$ which, in turn, is induced by the policy.
+	\subsection{Timing}
+	Timing of each period of the game is the following.
+At the beginning of the period the firm decides on the level of effort it will exert.
+Then this level is exerted and, the firm privately learns its new state.
+After that the platform makes a recommendation, customers, who followed it, report reviews and the round ends.
+	\begin{center}
+	\begin{figure}[H]
+		\begin{tikzpicture}
+	\draw[ultra thick,->] (0,0) -- (11,0);
+	\draw[thick] (1,-0.2) -- (1,0.2);
+	\node[label=below:firm learns] at (1,-0.2) {};
+	\node[label=below:its state] at (1,-0.5) {};
+	\draw[thick] (3,-0.2) -- (3,0.2);
+	\node[label=above:firm makes] at (3,0.5) {};
+	\node[label=above:private investment] at (3,0.1) {};
+	\draw[thick] (5,-0.2) -- (5,0.2);
+	\node[label=below:customers] at (5,-0.2) {};
+	\node[label=below:arrive] at (5,-0.5) {};
+	\draw[thick] (7,-0.2) -- (7,0.2);
+	\node[label=above:a recommendation] at (7,0.2) {};
+	\node[label=above:platform makes] at (7,0.5) {};
+	\draw[thick] (9,-0.2) -- (9,0.2);
+	\node[label=below:platform receives] at (9,-0.2) {};
+	\node[label=below:reviews] at (9,-0.5) {};
+		\end{tikzpicture}
+	\caption{Timing}\label{fig1}
+	\end{figure}
+\end{center}
+ 	\subsection{Assumptions}
+ 	Before starting the analysis I would like to explicitly state assumptions the model is based on, and to provide some justifications of these assumptions.
+ 	\begin{assumption}
+ 		The platform knows customers' outside. options
+ 	\end{assumption}
+ 	This assumption at first glance may seem implausible, however, nowadays platforms have an enormous amount of personal data which can enable them to infer rather precisely about the potential outside option of each customer.
 
-	At $t = 0$ each firm privately learns its type. Then, the timing of the $t$-th ($t=1, 2, \dots$) stage of the game is the following:
-	\begin{enumerate}[1)]
-		\item Each firm decides whether to stay in business or not (and incurs cost $f_i$ if decides to stay)
-		\item Firms make private and costly investment in quality
-		\item The unobserved quality is realized
-		\item Customer arrives
-		\item The principal makes a recommendation
-		\item Customer decides which firm to choose
-		\item Payoffs are realized and the principal observes customers' experience
-	\end{enumerate}
-	\subsection{Formalization}
-	\subsubsection{Firms}
-	We have two competing firms.
-Each firm has its own type that somehow is supposed to represent its "effectiveness" or "inclination to investing and sustaining the high quality of the good".
-The simplest example that comes in mind is that the type represents the firm's marginal cost of investment: the higher the type - the less effective the firm is.
-(types are either discrete $\theta^H, \theta^L$ or continuous $\theta \in (0, 1)$).
+ 	\begin{assumption}
+ 		If the good is not recommended to the customer, then the customer does not consume it. While if the good is recommended the customer may or may not choose it.
+ 	\end{assumption}
+ 	This assumption implies that without the platform the customer is unable to find the good.
+Think about \textit{TripAdvisor}, if it does not show you a restaurant, you will not know about the existence of this restaurant.
+At the same time, if \textit{TripAdvisor} shows you an option it is up to you whether to go there or not.
+ 	\begin{assumption}
+ 		The utility customers obtain in the bad state, $q > 0$ and hence it is always profitable for the platform to recommend the good to at least a small share of customers.
+ 	\end{assumption}
+ This assumption ensures that the platform can get the information about the state of the firm each period.
+ \begin{assumption}
+ 	$p \ge 0.5$
+ \end{assumption}
+This assumption, on one hand, is pure technical, and simplifies an analysis of optimal policy, but on the other hand it seems reasonable that once the firm invests, the probability that investment will lead to an increase in quality is at least 50\%.
+\section{Analysis}
+Although the recommender policy the platform employs can be a complicated function of all previous states, I start the analysis from the simplest policy.
+Namely, I firstly consider the policy that takes into account only the previous realization of the quality of the good.
+\subsection{One-round recommendation policy}
+Suppose that the platform employs the recommender policy that is based only on the previous period realization of the quality of the good.
+In this case the policy consists of two shares $\{s^0, s^1\}$ such that $$s_t = \begin{cases}
+s^0, \text{ if }\mathcal{Q}_{t-1} = b\\
+s^1, \text{ if }\mathcal{Q}_{t-1} = g
+\end{cases}$$
+The platform should choose $s^0, s^1$ to maximize its objective taking into account incentives of the firm.
+But firstly as a benchmark consider the first-best case, when the firm always invest.
+In this case the recommender policy is characterized by the following lemma
 
-	\textbf{Question:} why do we need types?
-
-	\textbf{Answer:} customers will learn and infer about the type of the firm
-
-	Making investment $e \in \mathbb{R}_{+}$ (maybe make it discrete $e^L$ and $e^H$) in quality of its product, each firm incurs cost $c_i(e)$ that should depend on the firm's type (for example $c_i(e) = \theta_ie_i$ where $\theta_i$ is the type of the $i$-th firm).
-Moreover, if the firm decides to stay in business, it incurs fixed cost $f_i$ disregard whether the customer chooses the firm or not.
-$f_i$ is low enough to ensure that at least one firm will operate in any case (maybe fixed cost also should dependent on type).
-	Each firm obtains a profit $1$ if the customer chooses its product. Denote for each firm $$d_{it} = \begin{cases}
-	1, \text{if consumer }t\text{ chooses the firm }i\\
-	0, \text{if consumer }t\text{ does not choose the firm }i\\
-	\end{cases}$$
-	and
-	$$b_{it} = \begin{cases}
-	1, \text{if the firm }i\text{ decided to stay in busimess at }t\\
-	0, \text{if the firm }i\text{ decided not to stay in busimess at }t
-	\end{cases}$$
-	then each firm is maximizing $$\underset{\{e_{it}, b_{it}\}_{t=1}^{\infty} }{\max}\ \expect \left[\sum_{t=1}^{\infty} \delta^t b_{it}(d_{it} - c_i(e_{it})- f_i)\right]$$
-	subject to the recommendation policy implemented by the principal.
-
-	\textbf{Open questions:}
-
-	can the firms observe the experience of users who chose the competitor?
-It seems reasonable if they could : the restaurant manager can send somebody to the nearby competitive restaurant to infer about the real quality.
-
-	can the firm know that the competitor still in business?
-
-	\underline{Maybe infinite-horizon model will be too hard, make it finite (start from two-periods)}
-
-	\subsubsection{Products}
-	Each firm provides a product which in time $t$ has quality $q_{it} \in (0, 1)$  which is endogenously formed from the firms' investment $e_{it}$ in this period.
-
-	\textbf{Idea:} the quality should somehow represent consumers' satisfaction of the product, the higher quality - the higher (expected) satisfaction.
-
-	\textbf{Realization:} Suppose that upon consuming the good from the firm $i$, the customer $t$ gets a reward $$r_{it} = \begin{cases}
-	1, \text{ w.p. } q_{it}\\
-	0, \text{ w.p } 1 - q_{it}
-	\end{cases}$$
-    The quality of the product from the firm $i$ at time $t$ is determined as $$q_{it} = \begin{cases} \frac{e_{it}}{e_{it}+e_{jt}},\ &j \neq i, e_{it}^2 + e_{jt}^2 \neq 0 \\
-	0, &e_{it} = e_{jt} = 0
-	\end{cases}
-	$$
-
-	\textbf{Question:} why is the quality of the alternative $i$ affected (negatively) by the investment made by the firm $j$?
-
-	\textbf{One possible answer:} in real life there is no such a thing as an absolute quality.
-I think that good things are not good \textit{per se}, they are good \underline{in comparison} to other things.
-So, in this sense the assumption may seem to be reasonable.
-
-	\textbf{Concern:} in this set-up if both firms make high and equal investment the probability of success for the customer is $\frac{1}{2}$, whereas if only one firm made high investment while the other made low, the probability of success for the first firm becomes higher than $\frac{1}{2}$.
-Does it represent the reality?
-
-	\subsubsection{Principal}
-	Principal observes the history of consumers' choices and their payoffs, $h^t$ (denote $H^t$ the whole set of such histories) and can commit to the recommendation policy, a function $R: H^t \to M^t$, where $M^t$ is the message space feasible for the customer $t$.
-The principal's goal is to design a revelation policy $R(\cdot)$ in order to maximize agents' expected discounted payoff:
-	$$\pi = \underset{R(\cdot)}{\max}\ \expect\left[\sum_{t=1}^{\infty} \delta^t r_{t} \right]$$
-
-	where $r_t$ is an expected reward of the customer $t$.
-
-	\subsubsection{Customers}
-
-	Customers (agents) arrive sequentially one by one, they do not observe the experience of other agents, but can receive information revealed by the principal.
-For each principal's message $m_t \in M^t$ the agent's strategy is a function $\sigma^t: M^t \to A$, where $A = \{a_1, a_2\}$ describes the set of possible alternatives (products from the firm 1 or 2).
-In other words, the agent's strategy prescribes for each principal's message an action, the product from which firm to choose.
-Choosing the strategy, each agent $t$ maximizes
-	$$\underset{\{a_1, a_2\}}{\max}\ \expect_t \left[q_{it}|m_t \right]$$
-
-	\textbf{Questions}: what are priors? which alternative do the customers choose when indifferent between two?
-
-\section{Analysis}
-\subsection{Straightforward observations}
-First of all, it seems that in this model it will never be optimal (starting from some point) to start recommending only one firm.
-This is because in this case the firm which is not recommended leaves the business (because of fixed cost) and this eliminates any incentive for the incumbent to invest.
-Hence, the quality becomes $0$ and all subsequent customers will get 0.
-So, the platform should internalize its role in sustaining a competition.
-\subsection{No disclosure at all}
-Suppose that the platfotm employs the non-disclosure policy.
-That is, agents do not possess any information regarding the previous users' experience.
-In this case, agents make decisions based on priors.
-If customers are homogeneous and have identical priors about the quality, then the firm which the customers are $a\ priori$ inclined towards stays in business with zero investment while the second firm exits the market.
-In this equilibrium all customers get zero, so the platform's payoff is zero.
-
-As we can see, now with strategic agents the non-disclosure policy becomes much worse. One reason is that it completely eliminates any incentive to increase the quality of alternatives.
-
-\subsection{Full disclosure}
-\subsubsection{1 period, observable qualities}
-Suppose we have only 1 period, at the beginning of the period, firms privately draw its types $\theta$ according to some full-support continuous distribution with the p.d.f.
-$f(\theta)$.
-Then both firms simultaneously perform investment.
-After that the customer arrives and can observe realized qualities.
-The customer chooses the firm with higher quality (ties do not happen almost surely).
-In this case the following Lemma is true
-\begin{lemma}
-	Suppose $\theta \sim U(0, 1)$ then
-	in the one period simultaneous-move game with observable qualities there exist a symmetric equilibria in which both firms invest $e(\theta) = -\ln(\theta)$.
+\begin{lemma}(First-best)
+	If the firm always invests in quality then the optimal recommender policy implies $$s^1_{FB} = s^0_{FB} = p + (1-p)q$$
 \end{lemma}
 \begin{proof}
-	See Appendix
+	The proof is straightforward, if the firm invests then the expected quality of the service at each period is $p+(1-p)q$.
+The platform is maximizing $$\underset{s}{\max}\ (p+(1-p)q)s - \frac{s^2}{2}$$ and the solution is given by $$s = p + (1-p)q$$
 \end{proof}
+Clearly, if the platform implements this policy, the firm does not have incentives to invest in quality. 
+\begin{corollary}
+	If the firm is strategic then the first-best policy is not implementable i.e. if the platform employs the first-best recommender policy, the firm never invests.
+\end{corollary}
 
-\subsubsection{Many periods, observable qualities}
-
-If there are more than 1 period then in all periods after the first one the firm with higher type will learn that its type is high (because the customer did not choose it in the first period) and that is why the aforementioned equilibrium breaks down: less effective firm is better-off if it leaves the business, but if one of the firm leaves then the other has no incentive to invest: it becomes a monopoly.
-So let us note this result
+The lemma and the corollary above suggest that the optimal policy should induce the firm to invest in quality. The results below give a characterization of the optimal recommender policy.
 \begin{lemma}
-	Suppose $\theta \sim U(0, 1)$ then in many-periods simultaneous-move game with observable qualities there exists an equilibrium in which:
-	\begin{enumerate}
-		\item In the first period both firms play $e(\theta) = -\ln(\theta)$
-		\item In all subsequent periods:
-		\begin{itemize}
-			\item the more efficient firm (with smaller $\theta$) plays $e(\theta) = -\ln(\theta)$ if and only if there is a competitor in business, and $e(\theta) = 0$ otherwise
-			\item the less efficient firm (with higher $\theta$) leaves the business forever.
-		\end{itemize}
-	\end{enumerate}
+	Given the recommendation policy $\{s^0, s^1\}$, it is optimal for the firm to invest at the bad state if and only if it is optimal to invest at the good state.
 \end{lemma} 
 \begin{proof}
-	See Appendix.
+	Let $V_{ij}$ be the expected discounted value of the firm which is at state $i$ and in the previous period was at state $j$. Then this values are the solutions to the following system of equations:
+	\begin{align}\label{eq1}
+	\begin{cases}
+	V_{bb} = s^0 + \max\{\delta V_{bb}, \delta p V_{gb} + \delta (1-p)V_{bb} - c\}\\
+	V_{bg} = s^1 + \max\{\delta V_{bb}, \delta p V_{gb} + \delta (1-p)V_{bb} - c\}\\
+	V_{gb} = s^0 + \max\{\delta V_{bg}, \delta p V_{gg} + \delta (1-p)V_{bg} - c\}\\
+	V_{gg} = s^1 + \max\{\delta V_{bg}, \delta p V_{gg} + \delta (1-p)V_{bg} - c\}
+	\end{cases}
+	\end{align}
+	In each $\max$ the first component represents the case when the firm does not invest, while the second is the value the firm obtains if invests.
+
+	The firm exerts effort at bad state if and only if $$\delta V_{bb} \le \delta p V_{gb} + \delta (1-p)V_{bb} - c \iff V_{gb} - V_{bb} \ge \frac{c}{\delta p}$$
+	Similarly, the firm exerts effort at good state if and only if $$V_{gg} - V_{bg} \ge \frac{c}{\delta p}$$ Suppose that the firm exerts effort at bad state, by way of contradiction assume that the firm does not invest at good state.
+In this case the system of equation becomes
+	$$	\begin{cases}
+	V_{bb} = s^0 + \delta p V_{gb} + \delta (1-p)V_{bb} - c\\
+	V_{bg} = s^1 + \delta p V_{gb} + \delta (1-p)V_{bb} - c\\
+	V_{gb} = s^0 + \delta V_{bg}\\
+	V_{gg} = s^1 + \delta V_{bg}
+	\end{cases}$$ and the solution is $$\begin{cases}
+	V_{bb} = \frac{1+\delta p - \delta^2p}{(1-\delta)(1+\delta p)}s^0 + \frac{\delta^2 p}{(1-\delta)(1+\delta p)}s^1 - \frac{c}{(1-\delta)(1+\delta p)}\\
+	V_{bg} = \frac{\delta}{(1-\delta)(1+\delta p)}s^0 + \frac{1-\delta + \delta p}{(1-\delta)(1+\delta p)}s^1 - \frac{c}{(1-\delta)(1+\delta p)}\\
+	V_{gb} = \frac{(1+\delta p)(1-\delta)+\delta^2}{(1-\delta)(1+\delta p)}s^0 - \frac{\delta^2(1-p)}{(1-\delta)(1+\delta p)}s^1 - \frac{c \delta}{(1-\delta)(1+\delta p)}\\
+	V_{gg} = \frac{\delta^2}{(1-\delta)(1+\delta p)}s^0 + \frac{1 - \delta^2 + \delta p}{(1-\delta)(1+\delta p)}s^1 - \frac{c\delta}{(1-\delta)(1+\delta p)}
+	\end{cases}$$
+	the condition that the firm invest at bad state is
+	$$V_{gb} - V_{bb} = \frac{\delta}{1+\delta p}(s^1 - s^0) - \frac{c}{1+\delta p} \ge \frac{c}{\delta p}$$
+	But this implies
+	$$\frac{c}{\delta p} \le \frac{\delta}{1 + \delta p}(s^1 - s^0) - \frac{c}{1 + \delta p} = V_{gg} - V_{bg}$$ and hence it is also optimal to invest at good state.
+The contradiction demonstrates that if the firm invests at the bad state, it also invest at a good state.
+Using the similar argument it can be shown that investing at the good state also must imply investment at the bad state.
 \end{proof}
-As we can see, if qualities are observable in many periods game, unravelling may happen, and the principal's expected payoff becomes equal to the expected reward of the first agent $$\pi_0 = r_1 = \expect\left[\frac{\ln \theta_i}{\ln \theta_i + \ln \theta_j}\bigg| \theta_i < \theta_j\right] = \frac{3}{4},\ i \neq j$$
+The result means that either the firm invests at both states or does not invest at all.
+The next proposition characterize the policy that induce the firm to invest in quality.
+\begin{proposition}\label{prop1}
+	Given the recommender policy $\{s^0, s^1\}$, the firm exerts effort at both states if and only if $$s^1 - s^0 \ge \frac{c}{\delta^2 p}$$
+\end{proposition}
+\begin{proof}
+	Suppose the firm does not invest at all, in this case the system \eqref{eq1} becomes
+	\begin{align*}
+	\begin{cases}
+	V_{bb} = s^0 + \delta V_{bb}\\
+	V_{bg} = s^1 + \delta V_{bb}\\
+	V_{gb} = s^0 + \delta V_{bg}\\
+	V_{gg} = s^1 + \delta V_{bg}
+	\end{cases} \to \begin{cases}
+	V_{bb} = \frac{s^0}{1-\delta}\\
+	V_{bg} = s^1 +\frac{\delta}{1-\delta}s^0\\
+	V_{gb} = \delta s^1 + \frac{1-\delta + \delta^2}{1-\delta}s^0\\
+	V_{gg} = (1+\delta)s^1 + \frac{\delta^2}{1-\delta}s^0
+	\end{cases}
+	\end{align*}
+	The necessary condition that ensures that the firm does not invest is $	V_{gb} - V_{bb} < \frac{c}{\delta p}$.
+Here, $$	V_{gb} - V_{bb} = \delta(s^1 - s^0)$$ Hence, the firm will not invest if $s^1 - s^0 < \frac{c}{\delta^2 p}$.
 
-\subsection{Not full disclosure}
-\subsubsection{Qualities are observable by the principal}
-Can the principal do better? Let us take the simplest case, when the principal can observe the qualities.
-Suppose the principal commits to recommend high quality firm with probability $q$ and low probability good with probability $1-q$.
-In this case it is easy to see that the symmetric equilibrium of one-period game is an equilibrium in many-periods game
-\begin{lemma}
-	Suppose $\theta \sim U(0, 1)$, $f \le 0.5$, the platform can observe the realized qualities and commits to recommend to the customer better option with probability $q \in (0.5, 1)$ and worse option with probability $1-q$.
-Then there exists a symmetric equilibium in which both firms are playing $e(\theta) = -(2q - 1)\ln \theta$.
+	Now suppose, the firm invests in both states. In this case the system \eqref{eq1} becomes
+	\begin{align*}
+	\begin{cases}
+	V_{bb} = s^0 + \delta p V_{gb} + \delta(1-p)V_{bb} - c\\
+	V_{bg} = s^1 + \delta p V_{gb} + \delta(1-p)V_{bb} - c\\
+	V_{gb} = s^0 + \delta pV_{gg} + \delta(1-p)V_{bg} - c\\
+	V_{gg} = s^1 + \delta p V_{gg} + \delta (1 - p)V_{bg} - c
+	\end{cases}
+	\end{align*}
+	It is a system of linear equations, and its solution is given by:
+	$$\begin{cases}
+	V_{bb} = \frac{\delta^2p}{1-\delta}s^1 + \frac{1-\delta^2p}{1-\delta}s^0 - \frac{c}{1-\delta}\\
+	V_{bg} = \frac{1-\delta + \delta^2p}{1-\delta}s^1 + \frac{\delta(1-\delta p)}{1-\delta}s^0 - \frac{c}{1-\delta}\\
+	V_{gb} = \frac{\delta(1-\delta+\delta p)}{1-\delta}s^1 + \frac{1-\delta + \delta^2 - \delta^2 p}{1- \delta}s^0 - \frac{c}{1-\delta}\\
+	V_{gg} = \frac{1 - \delta^2 + \delta^2 p}{1 - \delta}s^1 + \frac{\delta^2(1-p)}{1- \delta}s^0 - \frac{c}{1 - \delta}
+	\end{cases}$$
+	and again to ensure that the firm indeed invests one needs to check that $V_{gb} - V_{bb} \ge \frac{c}{\delta p}$.
+In this case $$V_{gb} - V_{bb} = \delta(s^1 - s^0) \ge \frac{c}{\delta p} \iff s^1 - s^0 \ge \frac{c}{\delta^2 p}$$
+\end{proof}
+The result means that the difference in demand of customers that the platform provides the firm with should be high enough to make it optimal for the firm to invest in quality.
+Also, the straightforward corollary can be obtained
+\begin{corollary}
+	It becomes harder to encourage firm to invest if either cost of investment $c$ is high, or the probability that investment results in quality increase $p$ is low, or the discount factor $\delta$ is low.
+\end{corollary}
 
-	Moreover, in this equilibrium the platform's expected payoff is strictly greater than the platform's expected payoff $\pi_0$ in case of unravelling (Lemma 2) 
-\end{lemma}
+The next step is to choose the policy $\{s^0, s^1\}$ to maximize the platform's utility subject to the constraint imposed by the Proposition \ref{prop1}.
+The following theorem characterize the optimal one-round recommender policy.
+\begin{theorem}
+	If it is optimal to encourage the firm to invest, then the optimal one-round recommender policy is characterized by $$s^1 = \min \left\{ p + q(1-p) + \frac{c}{\delta^2 p} - \frac{c}{\delta}, 1\right\}, s^0 = \max\left\{ p + q(1-p) - \frac{c}{\delta}, 0\right\}$$
+\end{theorem} 
 \begin{proof}
-	See Appendix.
+	Let $U_i$ be the expected discounted value of the platform if the state of the firm in the previous period was $i$. Then these values are solutions to the following system of equations. 
+	\begin{align*}
+	\begin{cases}
+	U_b = p\left(s^0 - \frac{(s^0)^2}{2} + \delta U_g\right) + (1-p)\left(qs^0 - \frac{(s^0)^2}{2} + \delta U_b \right)\\
+	U_g = p\left(s^1 - \frac{(s^1)^2}{2} + \delta U_g\right) + (1-p)\left(qs^1 - \frac{(s^1)^2}{2} + \delta U_b \right)
+	\end{cases} \to \begin{cases}
+	U_b = \frac{(p + (1-p)q)}{1 - \delta(1-p)}\left(s^0 - \frac{(s^0)^2}{2} \right) + \frac{p\delta}{1 - \delta(1-p)}U_g\\
+	U_g = \frac{(p + (1-p)q)}{1 - \delta(1-p)}\left(s^1 - \frac{(s^1)^2}{2} \right) + \frac{p\delta}{1 - \delta(1-p)}U_b
+	\end{cases}
+	\end{align*}
+	the solution is given by
+	$$\begin{cases}
+	U_b = \frac{1-\delta p}{1-\delta}\left((p + q(1-p))s^0 - \frac{(s^0)^2}{2}\right) + \frac{\delta p}{1-\delta}\left((p + q(1-p))s^1 - \frac{(s^1)^2}{2}\right)\\
+	U_g = \frac{\delta (1-p)}{1-\delta}\left((p+q(1-p))s^0 - \frac{(s^0)^2}{2}\right) +\frac{1 - \delta (1-p)}{1-\delta}\left((p+q(1-p))s^1 - \frac{(s^1)^2}{2}\right)
+	\end{cases}$$
+	without loss of generality suppose that at $t = 0$ the firm is at bad state and the platform knows it. Then the platform should solve the following problem
+	\begin{align*}
+	\underset{s^1, s^0}{\max}&\ \frac{1-\delta p}{1-\delta}\left((p + q(1-p))s^0 - \frac{(s^0)^2}{2}\right) + \frac{\delta p}{1-\delta}\left((p + q(1-p))s^1 - \frac{(s^1)^2}{2}\right)\\
+	&\text{subject to } 0 \le s^0 \le 1,\ 0 \le s^1 \le 1,\ s^1 - s^0 \ge \frac{c}{\delta^2 p}
+	\end{align*}
+	Clearly $s^1 - s^0 \ge \frac{c}{\delta^2 p}$ should bind.
+Suppose that $s^1 < 1$, we can plug $s^0 = s^1 - \frac{c}{\delta^2 p}$ into the objective function.
+In this case the optimal solution is given by \begin{align}\label{eq2}
+	s^0 &= p + q(1-p) - \frac{c}{\delta} \nonumber\\
+	s^1 &= p+q(1-p) + \frac{c(1-\delta p)}{\delta^2 p}
+	\end{align}
+	We should consider the case $s^1 = 1$, however if $s^1$ obtained in \eqref{eq2} is less than 1, then in case $p \ge 0.5$ it can be shown that the platform's profit is greater for this $s^1$, comparing to $s^1 = 1$.
 \end{proof}
-
-As we can see, the platform can encourage competition.
+\section{Discussion}
+As we can see, the optimal one-round recommender policy with the strategic firm differs from the first-best policy when the firm is not strategic.
+With strategic firm the platform have to "punish" and "reward" the firm by decreasing and increasing the recommended share comparing to the first-best case.
+Namely, $s^0_{FB} - s^0 = \frac{c}{\delta}$, so the amount of punishment decreases once $\delta \to 1$ and increases once $c \to 1$.
+Note that if $q > 1 - \frac{c}{p\delta}$ then $s^0 < q$ which means that the share of customers $q - s^0$ who for sure could derive positive utility from using the good are not recommended doing so.
+This may suggest that it is not always optimal to induce the firm to invest in quality.
 
 \section{Extensions}
-\begin{enumerate}
-	\item Agents do not know they place in line.
-	\item Heterogeneous customers
-\end{enumerate}
-\subsection*{Appendix}
+As for possible extensions, although in the model I consider the simplest class of recommender policies, it would be interesting to consider more general policies ($K$-round policy, $K > 1$).
+In this case the platform gets more tools to induce the firm to invest, hence, I think the expected social welfare losses from providing the firm with incentives for investing could be lower.
 
-\begin{proof}[Proof of Lemma 1]
-	Let us find the equilibrium strategy for firms in a form $\varepsilon(\theta)$ where $\varepsilon(\theta):(0, 1) \to \mathbb{R}$ is a decreasing function.
-Then each firm is maximizing $$\underset{e}{\max}\ \prob(\theta > \varepsilon^{-1}(e))(1 - \theta e - f) - \prob(\theta \le \varepsilon^{-1}(e))(\theta e + f)$$
-	or using the fact that $$\prob(\theta \le x) = \begin{cases}
-	1, x > 1\\
-	x, x \in [0, 1]\\
-	0, x < 0
-	\end{cases}$$
-	the firm is maximizing
-	$$\underset{e}{\max}\ 1 - \varepsilon^{-1}(e) - \theta e - f$$
-	F.O.C.:
-	$$-\frac{1}{\varepsilon'(\varepsilon^{-1}(e))} = \theta$$
-	Since the equilibrium we are interested in is symmetric, $\varepsilon(\theta) = e$ and hence
-	$$\varepsilon'(\theta) = -\frac{1}{\theta}$$
-	Assuming $\varepsilon(1) = 0$ we get $$\varepsilon(\theta) = -\ln\theta$$ 
-\end{proof}
+Also, I would be intrigued to see how the optimal policy changes if the firm could charge customers for consuming the good.
+In this case the firm can use the price as a signal of the quality of the good.
 
-\begin{proof}[Proof of Lemma 3]
+Finally, it is worth mentioning that the results are based on the assumption that customers are behavioral and always follow the recommendation.
+I would be interested to see how the policy changes if customers may be strategic.
+In this case, it would be harder for the platform to induce exploration since it should take into account incentives of customers.
+\section{Conclusion}
+Recommender systems became very popular for last years.
+Several papers have studied the question of optimal design of these systems.
+While the optimality was mainly understood in terms of inducing exploration of new alternatives for myopic agents.
+This work is a small step towards another direction.
+I built a model that was supposed to present a different perspective on the role of the recommender system.
+Namely, I focused on firm's incentives to maintain the quality of the alternative, and how it can be affected by the policy implemented by the platform.
+Results demonstrate that the optimal policy should take into account the strategic nature of the firm, and the optimal one-round policy derived in this paper sheds light on how exactly the platform should do it.
 
-	If both firms stay in business in the particular period, then let us find the equilibrium strategy for firms in a form $\varepsilon(\theta)$ where $\varepsilon(\theta):(0, 1) \to \mathbb{R}$ is a decreasing function.
-Then each firm is maximizing
-	\begin{align*}\underset{e}{\max}\ q\{\prob(\theta > \varepsilon^{-1}(e))(1 - \theta e - f) + \prob(\theta \le \varepsilon^{-1}(e))(-\theta e - f)\} +\\+ (1-q)\{\prob(\theta > \varepsilon^{-1}(e))(-\theta e - f) + \prob(\theta \le \varepsilon^{-1}(e))(1 - \theta e - f)\}
-	\end{align*}
-	or using the fact that $$\prob(\theta \le x) = \begin{cases}
-	1, x > 1\\
-	x, x \in [0, 1]\\
-	0, x < 0
-	\end{cases}$$
-	the firm is maximizing 
-	$$\underset{e}{\max}\ q(1 - \varepsilon^{-1}(e)) + (1 - q)\varepsilon^{-1}(e) - \theta e - f$$
-	F.O.C.: $$-\frac{q}{\varepsilon'(\varepsilon^{-1}(e))} + \frac{1-q}{\varepsilon'(\varepsilon^{-1}(e))} - \theta = 0$$
-	Since the equilibrium we are interested in is symmetric, $\varepsilon(\theta) = e$ and hence
-	$$\varepsilon'(\theta) = -\frac{2q - 1}{\theta}$$
-	Assuming $\varepsilon(1) = 0$ we get $$\varepsilon(\theta) = -(2q - 1)\ln\theta$$.
+Being very simple, the model I presented can be extended in a variety of directions, that was discussed in the extension section and which I am planing to explore during the future research.
 
-	It remains to check that less effective firm stays in business. Expected profit from staying in business for less effective firm is equal to
-	$$\expect[1-q + (2q-1) \theta \ln \theta - f] = \frac{5}{4} - \frac{3}{2}q - f \ge 0$$
-
-	The expected payoff of the platform is equal to
-	$$\pi = \frac{q}{1-\delta} \expect\left[\frac{\ln \theta_i}{\ln \theta_i + \ln \theta_j}\bigg| \theta_i < \theta_j\right] +\frac{1-q}{1-\delta}\expect\left[\frac{\ln \theta_j}{\ln \theta_i + \ln \theta_j}\bigg| \theta_j \ge \theta_i\right] = \frac{2q+1}{4(1 - \delta)} > \frac{3}{4} = \pi_0$$
-\end{proof}
+\bibliography{lib}{}
+\bibliographystyle{unsrt}
 \end{document}

### 2020-01-11 Draft.tex
draft

@@
 \newtheorem{assumption}[theorem]{Assumption}
+\newtheorem{definition}[theorem]{Definition}
 \title{Optimal recommender system with strategic alternatives \\ Market Design. Term Project}
@@
 	\section{Introduction}
+	Suppose you would like to find a restaurant to go out tonight what would you do? 
 	Recommender systems became an integral part of our lives.
@@
 On the other hand, these services aggregate the information and may make rather precise recommendations to people which significantly reduce search cost for them.
-But here the interesting question arise: how to disclose the information, the platform aggregated, optimally?
+But here the interesting question arises: how to disclose the information, the platform aggregated, optimally?
 It turns out that the full disclosure may not be optimal from the social welfare point of view.
@@
 	Several papers studied this question in different frameworks (see Literature review section) and the main conclusion that all those paper draw is that there should be an optimal level of information obfuscation in order to induce short-living and myopic users to experiment.
-However, in the previous work the quality of alternatives was supposed to be exogenous to the process of recommendation and learning.
+However, in the previous works the quality of alternatives was supposed to be exogenous to the process of recommendation and learning.
 So, the recommender policy the platform implements cannot affect the realized quality of alternatives that users choose.
@@
 
-	I build a model that embraces three main counterparties: the platform, the firm and users.
+	I build a model that embraces three main parties: the platform, the firm and users.
 Users are choosing between the good of \textit{ex-ante} unknown quality provided by the firm and a safe outside option.
-The firm can exert costly private effort to increase the quality of the service and the platform design the optimal recommender policy taking this into account.
+The firm can exert costly private effort to increase the quality of the service and the platform designs the optimal recommender policy taking this into account.
 I demonstrate that the optimal policy is different once the firm becomes strategic.
@@
 	\section{Model}
-	In the model one firm is operating on the platform and provides customers with the service of ex-ante unknown quality.
+	In the model one firm is operating on the platform and provides customers with the service of ex-ante unknown quality at the exogenous price.
 Unit mass of customers arrives sequentially, each customer decides between the risky good provided by the firm and the safe outside option (customers differ in their outside option).
@@
 For example, the element $(1, 2)$ of the matrix represents the probability that the firm will move to the state $g$ if its current state is $b$.
+Firm is price taker, i.e.
+it charges all customers the exogenous price $\mathcal{P}$.
 
-	If at time $t$ the share of the customers $s_t \in [0, 1]$ decides to consume the good then the firm gets $$\pi_t = s_t - ce$$ where $c$ is the cost of effort and $e \in \{0, 1\}$.
+	If at time $t$ the share of the customers $s_t \in [0, 1]$ decides to consume the good then the firm gets $$\pi_t = \mathcal{P} s_t - ce$$ where $c$ is the cost of effort and $e \in \{0, 1\}$.
 The firm's objective is to choose an investing strategy that would maximize $$\expect_0\left[\sum_{t=0}^{\infty} \delta^t \pi_t \right]$$ subject to the recommender policy implemented by the platform.
@@
 	\end{cases}$$
-	where $q \in (0, 1)$. The total utility of the customer who has an outside option $r_i$ and arrives at time $t$ is determined as follows $$u_{it} = \begin{cases}
-	v_t &,\text{if he chose the good}\\
+	where $q \in (0, 1)$. The total utility of the customer who has an outside option $r_i$ and arrives at time $t$ is supposed to be quasilinear and is determined as follows $$u_{it} = \begin{cases}
+	v_t - \mathcal{P}&,\text{if he chose the good}\\
 	r_{i} &,\text{if he chose an outside option}
@@
 Where $s_t$ is the mapping from the set of possible histories at time $t$ to $[0, 1]$.
-If at time $t$ the platform recommends the good to the share of customers $s_t$ and everybody follows the recommendation then the platform derives a utility $$\mathcal{U}_t = v_ts_t - \frac{s_t^2}{2}$$ which is basically an aggregate surplus obtained by customers who chose a good.
+If at time $t$ the platform recommends the good to the share of customers $s_t$ and everybody follows the recommendation then the platform derives a utility $$\mathcal{U}_t = \int_{0}^{s_t} (v_t - \mathcal{P} - r_i)dr_i =  (v_t - \mathcal{P})s_t - \frac{s_t^2}{2}$$ which is basically an aggregate surplus obtained by customers who chose a good.
 The platform designs the recommendation policy that would maximize $$\expect_0\left[\sum_{t=0}^{\infty} \delta^t \mathcal{U}_t\right]$$ subject to the firm's strategy $\sigma(s_t(h^t))$ which, in turn, is induced by the policy.
@@
  	\begin{assumption}
- 		The platform knows customers' outside. options
+ 		The platform knows customers' outside options.
  	\end{assumption}
@@
 \begin{lemma}(First-best)
-	If the firm always invests in quality then the optimal recommender policy implies $$s^1_{FB} = s^0_{FB} = p + (1-p)q$$
+	If the firm always invests in quality then the optimal recommender policy implies $$s^1_{FB} = s^0_{FB} = p + (1-p)q - \mathcal{P} = \expect[v] - \mathcal{P}$$
 \end{lemma}
@@
 	The proof is straightforward, if the firm invests then the expected quality of the service at each period is $p+(1-p)q$.
-The platform is maximizing $$\underset{s}{\max}\ (p+(1-p)q)s - \frac{s^2}{2}$$ and the solution is given by $$s = p + (1-p)q$$
+The platform is maximizing $$\underset{s}{\max}\ (p+(1-p)q - \mathcal{P})s - \frac{s^2}{2}$$ and the solution is given by $$s = p + (1-p)q - \mathcal{P}$$
 \end{proof}
@@
 	\begin{cases}
-	V_{bb} = s^0 + \max\{\delta V_{bb}, \delta p V_{gb} + \delta (1-p)V_{bb} - c\}\\
-	V_{bg} = s^1 + \max\{\delta V_{bb}, \delta p V_{gb} + \delta (1-p)V_{bb} - c\}\\
-	V_{gb} = s^0 + \max\{\delta V_{bg}, \delta p V_{gg} + \delta (1-p)V_{bg} - c\}\\
-	V_{gg} = s^1 + \max\{\delta V_{bg}, \delta p V_{gg} + \delta (1-p)V_{bg} - c\}
+	V_{bb} = \mathcal{P}s^0 + \max\{\delta V_{bb}, \delta p V_{gb} + \delta (1-p)V_{bb} - c\}\\
+	V_{bg} = \mathcal{P}s^1 + \max\{\delta V_{bb}, \delta p V_{gb} + \delta (1-p)V_{bb} - c\}\\
+	V_{gb} = \mathcal{P}s^0 + \max\{\delta V_{bg}, \delta p V_{gg} + \delta (1-p)V_{bg} - c\}\\
+	V_{gg} = \mathcal{P}s^1 + \max\{\delta V_{bg}, \delta p V_{gg} + \delta (1-p)V_{bg} - c\}
 	\end{cases}
@@
 	$$	\begin{cases}
-	V_{bb} = s^0 + \delta p V_{gb} + \delta (1-p)V_{bb} - c\\
-	V_{bg} = s^1 + \delta p V_{gb} + \delta (1-p)V_{bb} - c\\
-	V_{gb} = s^0 + \delta V_{bg}\\
-	V_{gg} = s^1 + \delta V_{bg}
+	V_{bb} = \mathcal{P}s^0 + \delta p V_{gb} + \delta (1-p)V_{bb} - c\\
+	V_{bg} = \mathcal{P}s^1 + \delta p V_{gb} + \delta (1-p)V_{bb} - c\\
+	V_{gb} = \mathcal{P}s^0 + \delta V_{bg}\\
+	V_{gg} = \mathcal{P}s^1 + \delta V_{bg}
 	\end{cases}$$ and the solution is $$\begin{cases}
-	V_{bb} = \frac{1+\delta p - \delta^2p}{(1-\delta)(1+\delta p)}s^0 + \frac{\delta^2 p}{(1-\delta)(1+\delta p)}s^1 - \frac{c}{(1-\delta)(1+\delta p)}\\
-	V_{bg} = \frac{\delta}{(1-\delta)(1+\delta p)}s^0 + \frac{1-\delta + \delta p}{(1-\delta)(1+\delta p)}s^1 - \frac{c}{(1-\delta)(1+\delta p)}\\
-	V_{gb} = \frac{(1+\delta p)(1-\delta)+\delta^2}{(1-\delta)(1+\delta p)}s^0 - \frac{\delta^2(1-p)}{(1-\delta)(1+\delta p)}s^1 - \frac{c \delta}{(1-\delta)(1+\delta p)}\\
-	V_{gg} = \frac{\delta^2}{(1-\delta)(1+\delta p)}s^0 + \frac{1 - \delta^2 + \delta p}{(1-\delta)(1+\delta p)}s^1 - \frac{c\delta}{(1-\delta)(1+\delta p)}
+	V_{bb} = \frac{1+\delta p - \delta^2p}{(1-\delta)(1+\delta p)}\mathcal{P}s^0 + \frac{\delta^2 p}{(1-\delta)(1+\delta p)}\mathcal{P}s^1 - \frac{c}{(1-\delta)(1+\delta p)}\\
+	V_{bg} = \frac{\delta}{(1-\delta)(1+\delta p)}\mathcal{P}s^0 + \frac{1-\delta + \delta p}{(1-\delta)(1+\delta p)}\mathcal{P}s^1 - \frac{c}{(1-\delta)(1+\delta p)}\\
+	V_{gb} = \frac{(1+\delta p)(1-\delta)+\delta^2}{(1-\delta)(1+\delta p)}\mathcal{P}s^0 - \frac{\delta^2(1-p)}{(1-\delta)(1+\delta p)}\mathcal{P}s^1 - \frac{c \delta}{(1-\delta)(1+\delta p)}\\
+	V_{gg} = \frac{\delta^2}{(1-\delta)(1+\delta p)}\mathcal{P}s^0 + \frac{1 - \delta^2 + \delta p}{(1-\delta)(1+\delta p)}\mathcal{P}s^1 - \frac{c\delta}{(1-\delta)(1+\delta p)}
 	\end{cases}$$
 	the condition that the firm invest at bad state is
-	$$V_{gb} - V_{bb} = \frac{\delta}{1+\delta p}(s^1 - s^0) - \frac{c}{1+\delta p} \ge \frac{c}{\delta p}$$
+	$$V_{gb} - V_{bb} = \frac{\delta\mathcal{P}}{1+\delta p}(s^1 - s^0) + \frac{c}{1+\delta p} \ge \frac{c}{\delta p}$$
 	But this implies
-	$$\frac{c}{\delta p} \le \frac{\delta}{1 + \delta p}(s^1 - s^0) - \frac{c}{1 + \delta p} = V_{gg} - V_{bg}$$ and hence it is also optimal to invest at good state.
+	$$\frac{c}{\delta p} \le \frac{\delta\mathcal{P}}{1 + \delta p}(s^1 - s^0) + \frac{c}{1 + \delta p} = V_{gg} - V_{bg}$$ and hence it is also optimal to invest at good state.
 The contradiction demonstrates that if the firm invests at the bad state, it also invest at a good state.
@@
 	\begin{cases}
-	V_{bb} = s^0 + \delta V_{bb}\\
-	V_{bg} = s^1 + \delta V_{bb}\\
-	V_{gb} = s^0 + \delta V_{bg}\\
-	V_{gg} = s^1 + \delta V_{bg}
+	V_{bb} = \mathcal{P}s^0 + \delta V_{bb}\\
+	V_{bg} = \mathcal{P}s^1 + \delta V_{bb}\\
+	V_{gb} = \mathcal{P}s^0 + \delta V_{bg}\\
+	V_{gg} = \mathcal{P}s^1 + \delta V_{bg}
 	\end{cases} \to \begin{cases}
-	V_{bb} = \frac{s^0}{1-\delta}\\
-	V_{bg} = s^1 +\frac{\delta}{1-\delta}s^0\\
-	V_{gb} = \delta s^1 + \frac{1-\delta + \delta^2}{1-\delta}s^0\\
-	V_{gg} = (1+\delta)s^1 + \frac{\delta^2}{1-\delta}s^0
+	V_{bb} = \frac{\mathcal{P}s^0}{1-\delta}\\
+	V_{bg} = \mathcal{P}s^1 +\frac{\delta}{1-\delta}s^0\\
+	V_{gb} = \delta \mathcal{P}s^1 + \frac{1-\delta + \delta^2}{1-\delta}\mathcal{P}s^0\\
+	V_{gg} = (1+\delta)\mathcal{P}s^1 + \frac{\delta^2}{1-\delta}\mathcal{P}s^0
 	\end{cases}
@@
 	The necessary condition that ensures that the firm does not invest is $	V_{gb} - V_{bb} < \frac{c}{\delta p}$.
-Here, $$	V_{gb} - V_{bb} = \delta(s^1 - s^0)$$ Hence, the firm will not invest if $s^1 - s^0 < \frac{c}{\delta^2 p}$.
+Here, $$	V_{gb} - V_{bb} = \delta \mathcal{P}(s^1 - s^0)$$ Hence, the firm will not invest if $s^1 - s^0 < \frac{c}{\delta^2 p \mathcal{P}}$.
 
@@
 	\begin{cases}
-	V_{bb} = s^0 + \delta p V_{gb} + \delta(1-p)V_{bb} - c\\
-	V_{bg} = s^1 + \delta p V_{gb} + \delta(1-p)V_{bb} - c\\
-	V_{gb} = s^0 + \delta pV_{gg} + \delta(1-p)V_{bg} - c\\
-	V_{gg} = s^1 + \delta p V_{gg} + \delta (1 - p)V_{bg} - c
+	V_{bb} = \mathcal{P}s^0 + \delta p V_{gb} + \delta(1-p)V_{bb} - c\\
+	V_{bg} = \mathcal{P}s^1 + \delta p V_{gb} + \delta(1-p)V_{bb} - c\\
+	V_{gb} = \mathcal{P}s^0 + \delta pV_{gg} + \delta(1-p)V_{bg} - c\\
+	V_{gg} = \mathcal{P}s^1 + \delta p V_{gg} + \delta (1 - p)V_{bg} - c
 	\end{cases}
@@
 	$$\begin{cases}
-	V_{bb} = \frac{\delta^2p}{1-\delta}s^1 + \frac{1-\delta^2p}{1-\delta}s^0 - \frac{c}{1-\delta}\\
-	V_{bg} = \frac{1-\delta + \delta^2p}{1-\delta}s^1 + \frac{\delta(1-\delta p)}{1-\delta}s^0 - \frac{c}{1-\delta}\\
-	V_{gb} = \frac{\delta(1-\delta+\delta p)}{1-\delta}s^1 + \frac{1-\delta + \delta^2 - \delta^2 p}{1- \delta}s^0 - \frac{c}{1-\delta}\\
-	V_{gg} = \frac{1 - \delta^2 + \delta^2 p}{1 - \delta}s^1 + \frac{\delta^2(1-p)}{1- \delta}s^0 - \frac{c}{1 - \delta}
+	V_{bb} = \frac{\delta^2p}{1-\delta}\mathcal{P}s^1 + \frac{1-\delta^2p}{1-\delta}\mathcal{P}s^0 - \frac{c}{1-\delta}\\
+	V_{bg} = \frac{1-\delta + \delta^2p}{1-\delta}\mathcal{P}s^1 + \frac{\delta(1-\delta p)}{1-\delta}\mathcal{P}s^0 - \frac{c}{1-\delta}\\
+	V_{gb} = \frac{\delta(1-\delta+\delta p)}{1-\delta}\mathcal{P}s^1 + \frac{1-\delta + \delta^2 - \delta^2 p}{1- \delta}\mathcal{P}s^0 - \frac{c}{1-\delta}\\
+	V_{gg} = \frac{1 - \delta^2 + \delta^2 p}{1 - \delta}\mathcal{P}s^1 + \frac{\delta^2(1-p)}{1- \delta}\mathcal{P}s^0 - \frac{c}{1 - \delta}
 	\end{cases}$$
 	and again to ensure that the firm indeed invests one needs to check that $V_{gb} - V_{bb} \ge \frac{c}{\delta p}$.
-In this case $$V_{gb} - V_{bb} = \delta(s^1 - s^0) \ge \frac{c}{\delta p} \iff s^1 - s^0 \ge \frac{c}{\delta^2 p}$$
+In this case $$V_{gb} - V_{bb} = \delta\mathcal{P}(s^1 - s^0) \ge \frac{c}{\delta p} \iff s^1 - s^0 \ge \frac{c}{\delta^2 p \mathcal{P}}$$
 \end{proof}
@@
 \begin{corollary}
-	It becomes harder to encourage firm to invest if either cost of investment $c$ is high, or the probability that investment results in quality increase $p$ is low, or the discount factor $\delta$ is low.
+	It becomes harder to encourage firm to invest if either cost of investment $c$ is high, or the probability that investment results in quality increase $p$ is low, or the discount factor $\delta$ is low, ot the priced paid by customers $\mathcal{P}$ is low.
 \end{corollary}
@@
 \begin{theorem}
-	If it is optimal to encourage the firm to invest, then the optimal one-round recommender policy is characterized by $$s^1 = \min \left\{ p + q(1-p) + \frac{c}{\delta^2 p} - \frac{c}{\delta}, 1\right\}, s^0 = \max\left\{ p + q(1-p) - \frac{c}{\delta}, 0\right\}$$
+	If it is optimal to encourage the firm to invest, then the optimal one-round recommender policy is characterized by $$s^1 = \min \left\{ p + q(1-p) - \mathcal{P} + \frac{c}{\delta^2 p \mathcal{P}} - \frac{c}{\delta \mathcal{P}}, 1\right\}, s^0 = \max\left\{ p + q(1-p) - \mathcal{P} - \frac{c}{\delta \mathcal{P}}, 0\right\}$$
 \end{theorem} 
@@
 	\begin{align*}
+	&\begin{cases}
+	U_b = p\left((1-\mathcal{P})s^0 - \frac{(s^0)^2}{2} + \delta U_g\right) + (1-p)\left((q - \mathcal{P})s^0 - \frac{(s^0)^2}{2} + \delta U_b \right)\\
+	U_g = p\left((1- \mathcal{P})s^1 - \frac{(s^1)^2}{2} + \delta U_g\right) + (1-p)\left((q - \mathcal{P})s^1 - \frac{(s^1)^2}{2} + \delta U_b \right)
+	\end{cases} 
+	\end{align*}
+	which can be rewritten as
+	\begin{align*}
 	\begin{cases}
-	U_b = p\left(s^0 - \frac{(s^0)^2}{2} + \delta U_g\right) + (1-p)\left(qs^0 - \frac{(s^0)^2}{2} + \delta U_b \right)\\
-	U_g = p\left(s^1 - \frac{(s^1)^2}{2} + \delta U_g\right) + (1-p)\left(qs^1 - \frac{(s^1)^2}{2} + \delta U_b \right)
-	\end{cases} \to \begin{cases}
-	U_b = \frac{(p + (1-p)q)}{1 - \delta(1-p)}\left(s^0 - \frac{(s^0)^2}{2} \right) + \frac{p\delta}{1 - \delta(1-p)}U_g\\
-	U_g = \frac{(p + (1-p)q)}{1 - \delta(1-p)}\left(s^1 - \frac{(s^1)^2}{2} \right) + \frac{p\delta}{1 - \delta(1-p)}U_b
+	U_b = \frac{1}{1 - \delta(1-p)} \left( (p + (1-p)q - \mathcal{P})s^0 - \frac{(s^0)^2}{2} \right) + \frac{p\delta}{1 - \delta(1-p)}U_g\\
+	U_g = \frac{1}{1 - \delta p} \left( (p + (1-p)q - \mathcal{P})s^1 - \frac{(s^1)^2}{2} \right) + \frac{(1-p)\delta}{1 - \delta p}U_b
 	\end{cases}
@@
 	$$\begin{cases}
-	U_b = \frac{1-\delta p}{1-\delta}\left((p + q(1-p))s^0 - \frac{(s^0)^2}{2}\right) + \frac{\delta p}{1-\delta}\left((p + q(1-p))s^1 - \frac{(s^1)^2}{2}\right)\\
-	U_g = \frac{\delta (1-p)}{1-\delta}\left((p+q(1-p))s^0 - \frac{(s^0)^2}{2}\right) +\frac{1 - \delta (1-p)}{1-\delta}\left((p+q(1-p))s^1 - \frac{(s^1)^2}{2}\right)
+	U_b = \frac{1-\delta p}{1-\delta}\left((p + q(1-p) - \mathcal{P})s^0 - \frac{(s^0)^2}{2}\right) + \frac{\delta p}{1-\delta}\left((p + q(1-p) - \mathcal{P})s^1 - \frac{(s^1)^2}{2}\right)\\
+	U_g = \frac{\delta (1-p)}{1-\delta}\left((p + q(1-p) - \mathcal{P})s^0 - \frac{(s^0)^2}{2}\right) +\frac{1-\delta(1- p)}{1-\delta}\left((p + q(1-p) - \mathcal{P})s^1 - \frac{(s^1)^2}{2}\right)
 	\end{cases}$$
@@
 	\begin{align*}
-	\underset{s^1, s^0}{\max}&\ \frac{1-\delta p}{1-\delta}\left((p + q(1-p))s^0 - \frac{(s^0)^2}{2}\right) + \frac{\delta p}{1-\delta}\left((p + q(1-p))s^1 - \frac{(s^1)^2}{2}\right)\\
-	&\text{subject to } 0 \le s^0 \le 1,\ 0 \le s^1 \le 1,\ s^1 - s^0 \ge \frac{c}{\delta^2 p}
+	\underset{s^1, s^0}{\max}&\ \frac{1-\delta p}{1-\delta}\left((p + q(1-p) - \mathcal{P})s^0 - \frac{(s^0)^2}{2}\right) + \frac{\delta p}{1-\delta}\left((p + q(1-p) - \mathcal{P})s^1 - \frac{(s^1)^2}{2}\right)
 	\end{align*}
-	Clearly $s^1 - s^0 \ge \frac{c}{\delta^2 p}$ should bind.
-Suppose that $s^1 < 1$, we can plug $s^0 = s^1 - \frac{c}{\delta^2 p}$ into the objective function.
+	Clearly $s^1 - s^0 \ge \frac{c}{\delta^2 p \mathcal{P}}$ should bind.
+Suppose that $s^1 < 1$, we can plug $s^0 = s^1 - \frac{c}{\delta^2 p \mathcal{P}}$ into the objective function.
 In this case the optimal solution is given by \begin{align}\label{eq2}
-	s^0 &= p + q(1-p) - \frac{c}{\delta} \nonumber\\
-	s^1 &= p+q(1-p) + \frac{c(1-\delta p)}{\delta^2 p}
+	s^0 &= p + q(1-p) - \mathcal{P} - \frac{c}{\delta \mathcal{P}} \nonumber\\
+	s^1 &= p + q(1-p) - \mathcal{P} - \frac{c}{\delta \mathcal{P}} + \frac{c}{\delta^2 p \mathcal{P}}
 	\end{align}
@@
 \end{proof}
-\section{Discussion}
 As we can see, the optimal one-round recommender policy with the strategic firm differs from the first-best policy when the firm is not strategic.
-With strategic firm the platform have to "punish" and "reward" the firm by decreasing and increasing the recommended share comparing to the first-best case.
-Namely, $s^0_{FB} - s^0 = \frac{c}{\delta}$, so the amount of punishment decreases once $\delta \to 1$ and increases once $c \to 1$.
-Note that if $q > 1 - \frac{c}{p\delta}$ then $s^0 < q$ which means that the share of customers $q - s^0$ who for sure could derive positive utility from using the good are not recommended doing so.
+With strategic firm the platform has to "punish" and "reward" the firm by decreasing and increasing the recommended share comparing to the first-best case.
+Namely, $s^0_{FB} - s^0 = \frac{c}{\delta \mathcal{P}}$, so the amount of punishment decreases once $\delta \to 1$ or $\mathcal{P} \to 1$, and increases once $c \to 1$.
+Note that if $q > 1 - \frac{c}{\delta p \mathcal{P}} - \frac{\mathcal{P}}{p}$ then $s^0 < q$ which means that the share of customers $q - s^0$ who for sure could derive positive utility from using the good are not recommended doing so.
 This may suggest that it is not always optimal to induce the firm to invest in quality.
 
+In addition, the recommender policy above is clearly not incentive compatible: those customers whose outside option $r_i \in \left(p+q(1-p) - \mathcal{P}, s^1 \right)$ would rather choose the outside option than follow the recommendation.
+This is because by following recommendation, given that the firm is investing in quality, the expected utility of the customer is $p + q(1-p) - \mathcal{P}$ while the outside option yields more.
+But that means that the platform cannot induce the firm to invest in quality since the firm does not obtain enough profit.
+That is why, the optimal one-round recommender policy with strategic customers should be different from that obtained in this section.
+
+\subsection{Strategic customers}
+In this section, I relax the assumption that customers always follow the recommendation, prescribed by the platform.
+Now, they only do so if and only if the expected utility from choosing the good exceeds the outside option.
+Formally,
+\begin{definition}
+	Given $p, q, \mathcal{P}$, the customer with the outside option $r_i$ would follow the recommendation choosing the good if and only if $$p + (1-p)q - \mathcal{P} \ge r_i$$
+\end{definition}
+
+To make some customers willing to follow the recommendation the platform may introduce personal discounts.
+\begin{definition}
+	The platform employs the personal discount system $d(r_i): [0, 1] \to [0, 1]$ if the price paid by the customer with the outside option $r_i$ is $d(r_i) \cdot \mathcal{P}$
+\end{definition}
+In other words, by offering a personal discounts the platform makes the good more attractive for  customers.
+This eventually allows the platform to meet the incentive compatibility constraint of the customers and, as a result, to render the recommender policy incentive compatible.
+\begin{theorem}
+	If it is optimal to encourage the firm to invest in quality then the best one-round recommender policy embraces recommendations $$s^1 = \min \left\{ p + q(1-p) - \mathcal{P} + \frac{c}{\delta^2 p \mathcal{P}} - \frac{c}{\delta \mathcal{P}}, 1\right\}, s^0 = \max\left\{ p + q(1-p) - \mathcal{P} - \frac{c}{\delta \mathcal{P}}, 0\right\}$$ and the personal discount system $$d(r_i) = \begin{cases}
+	1, &r_i \le p + (1-p)q - \mathcal{P}\\
+	\frac{p+(1-p)q - r_i}{\mathcal{P}}, &r_i > p + (1-p)q - \mathcal{P}
+	\end{cases}$$
+\end{theorem}
+\begin{proof}
+	to be completed
+\end{proof}
 \section{Extensions}

### 2020-01-14 Draft.tex
draft

@@
 \usepackage[section]{placeins}
+\usepackage{breakcites}
 \usepackage[makeroom]{cancel}
@@
 	\section{Introduction}
-	Suppose you would like to find a restaurant to go out tonight what would you do? 
+	Suppose you came to New-York and would like to find a good restaurant to go out tonight.
+Or you are going to spend your upcoming vacation on the coast of Spain and you need to find a hotel.
+What would you do?
+Where would you search for?
+Before the World Wide Web, people in such cases relied upon guide-books and brochures distributed by traveling agents.
+Those guides usually quickly became out-of-dated and were criticized for containing a misleading information.
+
+	After the advent of the Internet, finding up-to-date information has become much easier.
+To make this process even easier there began to appear Internet platforms that aggregate customers' reviews on products and service providers.
+Some of that platforms simply collect the reviews and present a rating that is just an average measure of consumers' satisfaction (weighted or simple mean of rates).
+Examples embrace \textit{Yelp} for restaurants, home and car services, or \textit{SiteJabber} for online businesses.
+
+	 Other platforms came further and began to not only collect the reviews but also to make individual recommendations based on these reviews.
+Usually these recommendations have a form of ranking of several alternatives that are ranked according to the secret propriety algorithm that the particular platform employs.
+The most widely known examples of such a kind of platforms include: \textit{TripAdvisor} which presents ranking on Hotels and Restaurants based on previous users' experience; \textit{Trustpilot} which is one of the most popular review services for a large variety of businesses; \textit{Glassdoor} that uses reviews of employees to rank companies by workers' satisfaction; and many other examples.
+Also, the interesting example is \textit{Amazon} which before 2015 calculated ranking just as an average of customers' reviews but after 2015 it changed the algorithm to a more complex, secret algorithm.
+The reason why \textit{Amazon} may have done so will be highlighted later in the paper.
+
+	 In this work by recommender system I mean such a kind of platforms that aggregates reviews of customers and makes a recommendations to newcomers based on these reviews.
+Those recommendation may be direct or indirect (e.g.
+the platform can provide a ranking of alternatives, this is not the direct recommendation to choose the top-ranked alternative, however, users understand that the platform thinks that this alternative is the best fit for them, so the platform implicitly recommends this alternative).
+
 	Recommender systems became an integral part of our lives.
-The most widely known examples of such a kind of services include: \textit{TripAdvisor} which presents ranking on Hotels and Restaurants for customers based on previous users' experience; \textit{Yelp} which provides a forum with reviews and ranking for different businesses; \textit{Netflix} which recommends movies and series also based on information aggregated from different customers.
-All those services are very important since, on one hand, they provide people with information about options they otherwise may never be aware of, at no (or very low) cost.
-On the other hand, these services aggregate the information and may make rather precise recommendations to people which significantly reduce search cost for them.
+Those services are very popular since, on one hand, they provide people with information about options they otherwise may never be aware of, at no (or very low) cost.
+On the other hand, these services aggregate the information and may make rather accurate recommendations to people which significantly reduces search cost for them.
 But here the interesting question arises: how to disclose the information, the platform aggregated, optimally?
-It turns out that the full disclosure may not be optimal from the social welfare point of view.
+Is it the best option from the social welfare standpoint to disclose all the information?
+
+	Several papers studied this question in different frameworks (see Literature review section) and using different models.
+It turns out that the full disclosure may not be optimal.
 This happens since users, who do not internalize the information externalities they impose to other agents, underexplore i.e.
 start to choose the safe option and do not try new things.
-
-	Several papers studied this question in different frameworks (see Literature review section) and the main conclusion that all those paper draw is that there should be an optimal level of information obfuscation in order to induce short-living and myopic users to experiment.
+The main conclusion that all those paper draw is that there should be an optimal level of information obfuscation in order to induce users to experiment.
 However, in the previous works the quality of alternatives was supposed to be exogenous to the process of recommendation and learning.
 So, the recommender policy the platform implements cannot affect the realized quality of alternatives that users choose.
-But in reality, these alternatives are also strategic agents (think about restaurants or hotels), and in most cases these agents are way more sophisticated than users.
-These strategic agents may adjust the quality of the good to the recommender policy the platform is employing.
-And the platform, in turn, should take the strategic nature of alternatives into account once developing the policy.
-To my best knowledge, there are no papers that study the optimal recommender policy with strategic alternative, so this work may be considered as the first small step in this direction.
+But in reality, these alternatives may be also strategic agents (think about restaurants, hotels or any other businesses), and in most cases these agents are way more sophisticated than users who usually come to the platform choose a restaurant or a hotel and then leave the platform for a long time.
+These strategic alternatives may adjust the quality of the good to the recommender policy the platform is employing.
+And the platform, in turn, should take the strategic nature of alternatives into account once developing its policy.
+The main question I am interested to address here is: what should be the best recommender policy the platform may implement if it deals with strategic alternatives which can adjust the quality of provided service to the recommender policy.
+To my best knowledge, there are no papers that study the optimal recommender policy in this framework, so the model I present here may be considered as the first small step in this direction.
 
-	I build a model that embraces three main parties: the platform, the firm and users.
-Users are choosing between the good of \textit{ex-ante} unknown quality provided by the firm and a safe outside option.
-The firm can exert costly private effort to increase the quality of the service and the platform designs the optimal recommender policy taking this into account.
+	I build a model that embraces three main parties: the platform, the firm and users (customers).
+Users arrive each period and are choosing between the good which is supplied by the firm and the outside option.
+Customers differ in the value of the outside option, and the utility they derive upon choosing the outside option is known for each customer.
+At the same time, the value customers derive from consuming the good is \textit{ex-ante} unknown since the good can have different quality and customers are unaware of the quality before making a choice.
+The firm can exert costly private effort to increase the quality of the good and the platform designs the optimal recommender policy taking this into account.
 I demonstrate that the optimal policy is different once the firm becomes strategic.
 Also, I derive the optimal recommender policy in the class of simple "1-round" policies.
+Additionally, I analyze how this policy changes if the customers also are strategic and follow the recommendation the platform makes if and only if the expected utility they may derive exceeds the value of the outside option.
+The results presented here highlight a new perspective on the role of the recommender system: it can not only aggregate information and provides the consumer side with reliable reviews and recommendations but also to encourage the producer side to invest in the product quality and maintain the high level of it.
+
 	The rest of the paper is organized as follows: the next section presents the literature review, section 3 describes the model, section 4 contains the main results, and last three sections embrace a brief discussion, possible extensions and directions for the future research, and a short conclusion.
@@
 	This work is related to the strand of literature on recommender systems.
-One of the first papers which studied the optimal design of recommender system is the work by Kremer et al \cite{kremer2014}.
+The first prototype of the recommender system has been created by Xerox Palo Alto Research Center in the late 1980-s early 1990-s and was described by  \cite{goldberg1992using}.
+This system has been called \textit{Tapestry} and has been designed in order to filter out a skyrocketed amount of the incoming emails.
+The system was trying to make the incoming stream of emails interesting for the particular person based on whether these emails were interesting for people whose interests are close to the interests of this person.
+Since that time many papers has been written that mainly explore the recommender systems from the Computer Science standpoint (see \cite{Beel2015r} for the recent survey).
+
+	One of the first papers which highlights the issue from the point of view of Economics and studies the optimal design of recommender system is the work by \cite{kremer2014}.
 The authors build a simple model with two alternatives which generate deterministic but \textit{ex-ante} unknown payoff.
@@
 The platform accumulates review and commits to the recommendation policy it uses to make recommendations to the newcomers.
-The papers, firstly, neatly in simple terms demonstrates why the full revelation policy may be suboptimal.
+The papers, firstly, neatly in simple terms demonstrates why the full revelation policy may be suboptimal\footnote{Maybe the appearance of this paper and similar to that in 2014 was the reason why \textit{Amazon} decided to switch their recommendation policy from the fully revealing one.}.
 Secondly, the optimal recommender policy is derived which has a "threshold" nature.
@@
 Otherwise the other alternative is recommended, and then once both alternatives were tried the best one is recommended until the end.
-The setup of this model is very similar to my, however, it has a clear distinction.
-In my model the quality of the alternative is endogenous to the process of recommendation.
-This suggests that the threshold-like type of policies cannot be optimal in the presence of strategic alternatives because once the firm knows that it will be recommended forever, it does not have any incentives to maintain the high quality of the service it provides.
+The model I build and present in the next section has similar setup, however I allow the quality of the alternative to be exogenous to the process of learning.
+This suggests that the threshold-like type of policies similar to those obtained in \cite{kremer2014} may not be optimal in the presence of strategic alternatives because once the firm knows that it will be recommended forever, it does not have any incentives to maintain the high quality of the service it provides.
 
@@
 The quality of the good is exogenously predetermined but unknown.
-Customers who decide to consume the good send noisy signal about its quality to the platform.
+Customers who decide to consume the good send a noisy signal about its quality to the platform.
 The signal is noisy in the sense that even if the quality of the good is good the information about it reaches to the platform only with some probability (different from 1).
-The platform updates its beliefs about the quality and tailors recommender policy contingent on the beliefs.
+The platform updates its beliefs about the quality and tailors a recommender policy contingent on the beliefs.
 Authors demonstrate that the full revelation policy is suboptimal and find the optimal policy which again has a threshold nature.
@@
 Otherwise, if no news arrive it is optimal to experiment with the risky alternative (recommend it) until the beliefs that its quality is good fall below the certain threshold.
-After that the optimal policy never recommends this alternative.
+After that, the optimal policy prescribes to never recommend this alternative.
 
-	Finally, Papanstasiou et al \cite{papanastasiou2017} extended ideas developed in \cite{kremer2014} and \cite{che2015} and build a dynamic, stochastic model.
+	Finally, Papanstasiou et al.
+\cite{papanastasiou2017} extended ideas developed in \cite{kremer2014} and \cite{che2015}, and build a dynamic, stochastic model.
 In their model, agents also arrive sequentially to the platform and should choose between two alternatives that generate stochastically exogenous payoff.
@@
 
-	The common feature of all aforementioned models is some form of exogeneity of the quality of the alternative.
+	The model that will be considered lately in the paper includes the stream of agents who should take an action and the results of their action may affect the actions of subsequently arriving agents.
+This feature may links this work to the information cascading literature (seminal papers are those of \cite{banerjee1992simple} and \cite{bikhchandani1992theory}).
+The information cascading in my model may happen if the platform reveals the information of the previous users' experience to the newcomers.
+However the scope of this work is completely different: the platform designs a recommender policy that prevents the information cascading from happening and moreover this policy induces the firm to exert effort and maintain a high quality of the good.
+
+	The common feature of the models with recommender systems considered in the aforementioned papers is some form of exogeneity of the quality of the alternative: in \cite{kremer2014} and \cite{che2015} it is determined and fixed, while in \cite{papanastasiou2017} it realizes randomly according to the parametric family of distribution with unknown parameters.
 In the next section I present a model with the strategic firm that can adjust the quality of the good to the learning process and recommendation policy.
@@
 
+\newpage
 \bibliography{lib}{}
-\bibliographystyle{unsrt}
+%\bibliographystyle{unsrt}
+\bibliographystyle{apalike}
 \end{document}

### 2020-01-16 Draft.tex
Draft

@@
 \usepackage{wrapfig} % 
+%\newenvironment{abstract}[0]{\small\rm
+%	\begin{center}ABSTRACT
+%		\\ \vspace{8pt}
+%		\begin{minipage}{5.2in}\smalllineskip
+%			\hspace{1pc}}{\end{minipage}\end{center}\vspace{-1pt}}
 \newcommand{\prob}{\mathbb{P}}
@@
 \newtheorem{definition}[theorem]{Definition}
-\title{Optimal recommender system with strategic alternatives \\ Market Design. Term Project}
+\title{Optimal recommender system with strategic alternatives \\ Term Paper}
 \date{}
@@
 	\maketitle
+	\begin{abstract}
+		Recommender systems became an integral part of our lives.
+Many papers studied the design of an optimal system that can balance exploration and exploitation for the myopic short-living agents.
+In previous works, the quality of alternatives (the utility customers derive upon consuming) was supposed to be exogenous to the recommendation process.
+I would like to study the interaction between recommender policy and the quality of the provided service.
+I build a model in which the firm is operating on the platform and provides customers with the service of ex-ante unknown quality.
+The firm can make a private and costly investment to increase the quality.
+The platform observes agents’ experience and commits to recommendation policy.
+The model proposes a new perspective on the role of the recommender system.
+Apart from inducing exploration, here the system can also encourage the firm to invest in quality as well as can smooth negative shocks.
+	\end{abstract}
+	\newpage
 	\section{Introduction}
@@
 
-	Finally, Papanstasiou et al.
+	Papanstasiou et al.
 \cite{papanastasiou2017} extended ideas developed in \cite{kremer2014} and \cite{che2015}, and build a dynamic, stochastic model.
@@
 At the same time, if \textit{TripAdvisor} shows you an option it is up to you whether to go there or not.
- 	\begin{assumption}
- 		The utility customers obtain in the bad state, $q > 0$ and hence it is always profitable for the platform to recommend the good to at least a small share of customers.
- 	\end{assumption}
- This assumption ensures that the platform can get the information about the state of the firm each period.
- \begin{assumption}
- 	$p \ge 0.5$
- \end{assumption}
-This assumption, on one hand, is pure technical, and simplifies an analysis of optimal policy, but on the other hand it seems reasonable that once the firm invests, the probability that investment will lead to an increase in quality is at least 50\%.
 \section{Analysis}
@@
 Namely, I firstly consider the policy that takes into account only the previous realization of the quality of the good.
-\subsection{One-round recommendation policy}
+Moreover, in the following subsection I assume that the customers blindly follow the recommendation the platform makes.
+\subsection{Nonstrategic customers}
 Suppose that the platform employs the recommender policy that is based only on the previous period realization of the quality of the good.
@@
 \begin{proposition}\label{prop1}
-	Given the recommender policy $\{s^0, s^1\}$, the firm exerts effort at both states if and only if $$s^1 - s^0 \ge \frac{c}{\delta^2 p}$$
+	Given the recommender policy $\{s^0, s^1\}$, the firm exerts effort at both states if and only if $$s^1 - s^0 \ge \frac{c}{\delta^2 p \mathcal{P}}$$
 \end{proposition}
@@
 \begin{corollary}
+	It is possible to incentivize the firm to invest only if $c \le \delta^2 p \mathcal{P}$.
+\end{corollary}
+Thereafter I will assume that the condition above condition holds.
+\begin{corollary}
 	It becomes harder to encourage firm to invest if either cost of investment $c$ is high, or the probability that investment results in quality increase $p$ is low, or the discount factor $\delta$ is low, ot the priced paid by customers $\mathcal{P}$ is low.
@@
 The following theorem characterize the optimal one-round recommender policy.
-\begin{theorem}
-	If it is optimal to encourage the firm to invest, then the optimal one-round recommender policy is characterized by $$s^1 = \min \left\{ p + q(1-p) - \mathcal{P} + \frac{c}{\delta^2 p \mathcal{P}} - \frac{c}{\delta \mathcal{P}}, 1\right\}, s^0 = \max\left\{ p + q(1-p) - \mathcal{P} - \frac{c}{\delta \mathcal{P}}, 0\right\}$$
+\begin{theorem} \label{th1}
+	If it is optimal to encourage the firm to invest, then the optimal one-round recommender policy is characterized by $\{s^1, s^0\}$ such that:
+	\begin{itemize}
+		\item if $p+(1-p)q - \mathcal{P} - \frac{c}{\delta \mathcal{P}} < 0$ then $\left\{s^0 = 0, s^1 = \frac{c}{\delta^2 p \mathcal{P}} \right\}$
+		\item if $p+(1-p)q - \mathcal{P} - \frac{c}{\delta \mathcal{P}} \ge 0$ then
+		\begin{itemize}
+			\item if $p+(1-p)q - \mathcal{P} - \frac{c}{\delta \mathcal{P}} + \frac{c}{\delta^2 p \mathcal{P}} < 1$ then $$\begin{cases}s^0 = p+(1-p)q - \mathcal{P} - \frac{c}{\delta \mathcal{P}}\\ s^1 = p+(1-p)q - \mathcal{P} - \frac{c}{\delta \mathcal{P}} + \frac{c}{\delta^2 p \mathcal{P}} \end{cases}$$
+			\item if $p+(1-p)q - \mathcal{P} - \frac{c}{\delta \mathcal{P}} + \frac{c}{\delta^2 p \mathcal{P}} \ge 1$ then $\left\{s^0 = 1-\frac{c}{\delta^2 p \mathcal{P}}, s^1 =1 \right\}$
+		\end{itemize}
+	\end{itemize}
 \end{theorem} 
@@
 	\end{align}
-	We should consider the case $s^1 = 1$, however if $s^1$ obtained in \eqref{eq2} is less than 1, then in case $p \ge 0.5$ it can be shown that the platform's profit is greater for this $s^1$, comparing to $s^1 = 1$.
+	Keeping in mind that $s^0 = s^1 - \frac{c}{\delta^2 p \mathcal{P}}$ the objective function is a downward parabola, hence, if $p+q(1-p) - \mathcal{P} - \frac{c}{\delta \mathcal{P}} < 0$ (and using the assumption that $\frac{c}{\delta^2 p \mathcal{P}} < 1$) then $s^0 = 0$ and to keep the constraint that ensures that the firm exerts effort satisfied, $s^1 = \frac{c}{\delta^2 p \mathcal{P}}$.
+If $p+q(1-p) - \mathcal{P} - \frac{c}{\delta \mathcal{P}} \ge 0$ then two cases are possible: either $p+q(1-p) - \mathcal{P} - \frac{c}{\delta \mathcal{P}} + \frac{c}{\delta^2 p \mathcal{P}} < 1$ or $p+q(1-p) - \mathcal{P} - \frac{c}{\delta \mathcal{P}} + \frac{c}{\delta^2 p \mathcal{P}} \ge 1$.
+The former case implies an internal solution and $$\begin{cases}s^0 = p+(1-p)q - \mathcal{P} - \frac{c}{\delta \mathcal{P}}\\ s^1 = p+(1-p)q - \mathcal{P} - \frac{c}{\delta \mathcal{P}} + \frac{c}{\delta^2 p \mathcal{P}} \end{cases}$$
+	The latter case implies $s^1 = 1$ and consequently $s^0 = 1 - \frac{c}{\delta^2 p \mathcal{P}}$.
 \end{proof}
@@
 
-In addition, the recommender policy above is clearly not incentive compatible: those customers whose outside option $r_i \in \left(p+q(1-p) - \mathcal{P}, s^1 \right)$ would rather choose the outside option than follow the recommendation.
+In addition, the recommender policy above is clearly not incentive compatible in all cases: those customers whose outside option $r_i \in \left(p+q(1-p) - \mathcal{P}, s^1 \right)$ would rather choose the outside option than follow the recommendation.
 This is because by following recommendation, given that the firm is investing in quality, the expected utility of the customer is $p + q(1-p) - \mathcal{P}$ while the outside option yields more.
@@
 \begin{definition}
-	The platform employs the personal discount system $d(r_i): [0, 1] \to [0, 1]$ if the price paid by the customer with the outside option $r_i$ is $d(r_i) \cdot \mathcal{P}$
+	The platform employs the personal discount system $d(r_i): [0, 1] \to [0, 1]$ if the price paid by the customer with the outside option $r_i$ is $(1-d(r_i)) \cdot \mathcal{P}$
 \end{definition}
@@
 \begin{theorem}
-	If it is optimal to encourage the firm to invest in quality then the best one-round recommender policy embraces recommendations $$s^1 = \min \left\{ p + q(1-p) - \mathcal{P} + \frac{c}{\delta^2 p \mathcal{P}} - \frac{c}{\delta \mathcal{P}}, 1\right\}, s^0 = \max\left\{ p + q(1-p) - \mathcal{P} - \frac{c}{\delta \mathcal{P}}, 0\right\}$$ and the personal discount system $$d(r_i) = \begin{cases}
-	1, &r_i \le p + (1-p)q - \mathcal{P}\\
-	\frac{p+(1-p)q - r_i}{\mathcal{P}}, &r_i > p + (1-p)q - \mathcal{P}
+	If it is optimal to encourage the firm to invest in quality then the best one-round recommender policy embraces recommendations $\{s^1, s^0\}$ such that:
+	\begin{itemize}
+		\item if $p+(1-p)q - \mathcal{P} - \frac{c(1+\delta p)}{2\delta p \mathcal{P}} < 0$ then $\left\{s^0 = 0, s^1 = \frac{c}{\delta^2 p \mathcal{P}} \right\}$
+		\item if  $p+(1-p)q - \mathcal{P} - \frac{c(1+\delta p)}{2\delta p \mathcal{P}} \ge 0$ then
+		\begin{itemize}
+			\item if $p+(1-p)q - \mathcal{P} - \frac{c}{2\delta \mathcal{P}} + \frac{c}{2\delta^2 p \mathcal{P}} < 1$ then $$\begin{cases}s^0 = p+(1-p)q - \mathcal{P} - \frac{c(1+\delta p)}{2\delta p \mathcal{P}}\\ s^1 = p+(1-p)q - \mathcal{P} - \frac{c}{2\delta \mathcal{P}} + \frac{c}{2\delta^2 p \mathcal{P}} \end{cases}$$
+			\item if $p+(1-p)q - \mathcal{P} - \frac{c}{2\delta \mathcal{P}} + \frac{c}{2\delta^2 p \mathcal{P}} \ge 1$ then $\left\{s^0 = 1-\frac{c}{\delta^2 p \mathcal{P}}, s^1 =1 \right\}$
+		\end{itemize}
+	\end{itemize}
+
+	and the personal discount system $$d(r_i) = \begin{cases}
+	0, &r_i \le p + (1-p)q - \mathcal{P}, r_i > s^1\\
+	1-\frac{p+(1-p)q - r_i}{\mathcal{P}}, &p + (1-p)q - \mathcal{P} \le r_i \le s^1
 	\end{cases}$$
@@
 \begin{proof}
-	to be completed
+	If the firm invests in quality then each period the expected quality of the good is $$\expect[v] = p + (1-p)q$$
+	hence, those customers who are recommended to choose the good and have $r_i \le \expect[v]$ will follow the recommendation.
+The platform needs to incentivize only those whose $r_i \in (\expect[v], s^1)$.
+To do so, the platform offers those customers a personal price $\mathcal{P}_i  = p+(1-p)q - r_i = (1-d(r_i)) \mathcal{P}$, hence, for each agent the platform is losing $$\mathcal{P} - \mathcal{P}_i = \mathcal{P} -  p-(1-p)q + r_i$$ because of the personal discount
+	$$d(r_i) = 1 - \frac{p+(1-p)q - r_i}{\mathcal{P}},\ p+(1-p)q - \mathcal{P} \le r_i \le s^1$$
+	Again without loss of generality assuming that initially the firm is in the bad state, the platform is solving the following problem 
+	\begin{align*} \underset{s^1, s^0}{\max}\ &\frac{1-\delta p}{1-\delta}\left((p + q(1-p) - \mathcal{P})s^0 - \frac{(s^0)^2}{2}\right) + \frac{\delta p}{1-\delta}\left((p + q(1-p) - \mathcal{P})s^1 - \frac{(s^1)^2}{2}\right) - \\
+	&- \frac{1}{1-\delta} \begin{cases} \int_{p+(1-p)q - \mathcal{P}}^{s^1} (\mathcal{P} -  p-(1-p)q + r_i)dr_i, &s^1 \ge p+(1-p)q - \mathcal{P} \\ 0, &\text{ else }\end{cases} = \\
+	= \underset{s^1, s^0}{\max}\ &\frac{1-\delta p}{1-\delta}\left((p + q(1-p) - \mathcal{P})s^0 - \frac{(s^0)^2}{2}\right) + \frac{\delta p}{1-\delta}\left((p + q(1-p) - \mathcal{P})s^1 - \frac{(s^1)^2}{2}\right) - \\
+	&- \frac{1}{1-\delta} \begin{cases} \frac{(s^1 + \mathcal{P} - q (1 - p) - p)^2}{2}, &s^1 \ge p+(1-p)q - \mathcal{P} \\ 0, &\text{ else }\end{cases}\\
+	&\text{s.t. } 0\le s^0\le 1, 0 \le s^1 \le 1, s^1 - s^0 \ge \frac{c}{\delta^2 p \mathcal{P}}
+	\end{align*}
+	the last term of the objective function represents the platform cost on discounts for customers to encourage them following the recommendation.
+Clearly $s^1 - s^0 \ge \frac{c}{\delta^2 p \mathcal{P}}$ binds, using this, under assumption that the solution lies in the interior, by taking FOC one can obtain
+	\begin{align*} s^1 &= p + q(1-p) - \mathcal{P} + \frac{c}{2 \delta^2 p \mathcal{P}} - \frac{c}{2 \delta \mathcal{P}}\\
+	 s^0 &= p + q(1-p) - \mathcal{P} - \frac{c}{2 \delta \mathcal{P}} \frac{1+\delta p}{\delta p}
+	 \end{align*}
+	 Similar to the proof of the theorem \ref{th1} one can obtain the corner solutions
 \end{proof}
+	 As we can see, for the interior solution the share of customers which are recommended the good after the good state $s^1$ (and hence $s^0$) is lower than in case when customers are not strategic.
+This happens since now the platform has to induce the customers to follow the recommendation and make personal discounts for them.
+Note that discounts are offered for those customers whose outside option exceeds the average product quality.
+\section{Discussion}
+The results presented above demonstrate that first of all, the optimal policy differs once the firm becomes strategic.
+The platform has to incentivize the firm to maintain the high quality of the service.
+To do so, the platform uses the demand the firm faces as a tool to regulate its profit.
+With the strategic firm the platform increases the demand by recommending the firm to customers if the realized quality of the service the firm provides is good, and decreases the demand by not recommending the firm to users if in the previous period consumers were not satisfied with the quality.
+However, to ensure that rational customers follow the recommendations the platform has to introduce personal discounts.
+This partly may be the reason why \textit{TripAdvisor} proposes personal discounts for customers who book tables through the platform (see Fig.
+\ref{fig2}).
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\textwidth]{restaurant}
+	\caption{\textit{TripAdvisor} offers personal discounts for booking some restaurants via the platform}\label{fig2}
+\end{figure}
+The platform faces the following trade-off here: on one hand, it is costly to give discounts for customers, but on the other hand if they do not follow the recommendation it will not become optimal for the firm to exert efforts, and, as a result, the platform loses in the long-run.
+This trade-off makes the platform to reduce shares of customers that are recommended the good, comparing to the case when the strategic nature of customers is not an issue (they are blindly following the recommendation).
+
 \section{Extensions}
-As for possible extensions, although in the model I consider the simplest class of recommender policies, it would be interesting to consider more general policies ($K$-round policy, $K > 1$).
-In this case the platform gets more tools to induce the firm to invest, hence, I think the expected social welfare losses from providing the firm with incentives for investing could be lower.
+As for possible extensions, firstly, it would be interesting to see how the policy changes if there are more than 1 firm.
+In this case, the platform may use the competition among firms and to switch stream of customers from one firm to another based on the previous users' satisfaction.
+Also, in my model has been assumed the fact that the platform knows the initial state of the firm.
+However, I would be particularly interested in question: if the new firm comes to the platform and the platform is not aware of its state how will the optimal recommendation policy deal with this?
 
-Also, I would be intrigued to see how the optimal policy changes if the firm could charge customers for consuming the good.
-In this case the firm can use the price as a signal of the quality of the good.
+Secondly, in my model the effort the firm exerts affects only current quality of the product, it is interesting to see whether the quantitative results are robust to the case if the firm may make long-term investment, or the quality of the product may depreciate gradually.
 
-Finally, it is worth mentioning that the results are based on the assumption that customers are behavioral and always follow the recommendation.
-I would be interested to see how the policy changes if customers may be strategic.
-In this case, it would be harder for the platform to induce exploration since it should take into account incentives of customers.
+Thirdly, another direction for future research may include to make the quality not binary but continuous, or to make the firm's decision more complex.
+The reason I use the simplest state space and the simplest action space is to make the model as simple as possible but still being able to preserve the strategic nature of the firm.
+Nevertheless, it is, indeed, interesting to check the robustness of my results in these, more complex, frameworks.
+
+Finally, although in the model I consider the simplest class of recommender policies, it would be interesting to consider more general policies ($K$-round policy, $K > 1$).
+In this case the platform gets more tools to induce the firm to invest, however, the problem will become harder and results may become intractable.
 \section{Conclusion}
@@
 While the optimality was mainly understood in terms of inducing exploration of new alternatives for myopic agents.
-This work is a small step towards another direction.
+This work is a step towards another direction.
 I built a model that was supposed to present a different perspective on the role of the recommender system.
@@
 Results demonstrate that the optimal policy should take into account the strategic nature of the firm, and the optimal one-round policy derived in this paper sheds light on how exactly the platform should do it.
+Furthermore, the model suggests that it is optimal for the platform to introduce personal discounts for customers to ensure that they follow the recommendation.
+These discounts actually can be observed in reality, for example, \textit{TripAdvisor} may offer you a personal discount for a bill you would pay in the particular restaurant if you booked the table via \textit{TripAdvisor}.
 

### 2020-04-02 Thesis.tex
Add files via upload
Thesis

\documentclass[a4paper]{article}
\usepackage[12pt]{extsizes} % 
\usepackage{setspace}
\doublespacing
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
%\usepackage{harvard}
\usepackage[round]{natbib}
\usepackage{pdfpages}
\usepackage{tikz}
\usepackage{makecell}
\usepackage{amsthm}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{xcolor,colortbl}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage{breakcites}
\interfootnotelinepenalty=10000
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
%\newenvironment{abstract}[0]{\small\rm
%	\begin{center}ABSTRACT
%		\\ \vspace{8pt}
%		\begin{minipage}{5.2in}\smalllineskip
%			\hspace{1pc}}{\end{minipage}\end{center}\vspace{-1pt}}
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathcal{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=25mm, top=20mm, right=25mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{definition}[theorem]{Definition}
%\let\cite = \citeasnoun
\title{To run with the hare and hunt with the hounds: Optimal recommendation system with competing sellers. \\ Master Thesis}
\date{}
\author{Danil Fedchenko}
\begin{document} % 
	\maketitle
	\begin{abstract}
		Many e-commerce platforms that connect buyers and sellers employ recommender systems to help customers find products and services. These systems may have several potential effects on strategies of competing sellers: on one hand, more people become aware of products, so sellers may sell more goods and benefit from the presence of the recommender system. On the other hand, at the same time, more people become aware of the competitor seller. The latter effect may intensify the price competition and may hurt sellers but, of course, benefit customers. However, many platforms charge firms for sales and as a result have the stakes in firms profits. Thus, by designing the recommender policy the platform should balance the pro-competitive care about customers and anti-competitive incentives to keep high prices. The main goal of this work is to study this trade-off in details. I am solving for an optimal recommender policy and investigate factors that may affect it.
	\end{abstract}
\newpage
	\section{Introduction}
	\section{Literature review}
	\section{Model}
	The model I consider is similar to that considered by \cite{levin2009quality}, the difference is that I introduce a new player -- a platform which employs a recommender system (I will use words ``platform'' and ``recommender system'' interchangeably).

	There is a unit mass of heterogenous customers. Preferences of customer are characterized by the location of the customer on the unity interval. Two competing sellers (firms) which I will refer to $A$ and $B$, who locate at the extremes of the unit interval, are selling substitutable goods which they produce at equal marginal cost normalized to 0. Goods of each seller may have different quality which means that the customers with the same ``horizontal preference'' for the two sellers may derive different utility upon consuming goods of different sellers. Sellers are competing in prices which they set simultaneously. In addition, the sellers may costly disclose the information about the quality of their good prior to price setting. The recommender system observes prices and sellers' decisions regarding the quality disclosure, and makes recommendation to customers which seller to choose. In the baseline model I assume that the recommender system knows the exact location of the customer, lately this assumption is relaxed, and I consider the scenario when the recommender system receives a noisy signal regarding the customers' locations. Finally, the platform charges sellers a commission for sales according to a commission rate $\alpha$ which the platform can choose.

	The solution concept is perfect Bayesian equilibrium, and all parties are assumed to be risk-neutral.

	The next three subsections describes formally all parties of my model.

	\subsection{Customers}
	Customer who locates at $x \in [0, 1]$ derives a net utility $$V + q_A - p_A - x$$ if  he buys the good at a price $p_A$ from the seller $A$ who locates at a point $0$, and net utility $$V + q_B - p_B - (1-x)$$ if  he buys the good at a price $p_B$ from the seller $B$ who locates at a point $1$. $V$ is assumed to be higher than $\frac{3}{2}$ to ensure that the market is fully covered in equilibrium, $q_A, q_B \in [0, 1]$ are the qualities of the goods. The outside option for the customers is zero, so they follow the recommendation if the expected net utility from buying the good from the recommended seller is higher than 0.

	Customers do not know about the sellers unless the recommender system recommends them.

	\subsection{Sellers}

	Prior to price-setting sellers privately learns the quality of their good and should decide whether to disclose this information to the platform or not. Upon disclosure the sellers bear cost $c$. Moreover, the sellers pay the platform commission from sales. That is, the seller's $i$ profit is given by the following formula
	$$\pi_i = (1-\alpha)p_i D_i - c d_i$$
	where $$d_i = \begin{cases}
	1, \text{ if sellers }i\text{ discloses }\\
	0, \text{ otherwise }
	\end{cases}$$ and $D_i$ is the demand on good $i$ which may be a function of prices and decisions regarding the information disclosure, and is determined by the recommender policy the platform employs.

	\subsection{Platform}

	The platform commits to the commission rate $\alpha$ and the recommender policy. The recommender policy specifies functions $r_A(x, p_A, p_B, d_A, d_B, q_A, q_B)$ and $r_B(x, p_A, p_B, d_A, d_B, q_A, q_B)$ which represents probabilities that the customer with location $x$ will be recommended the seller $A$ and $B$ respectively. The objective function of the platform is given by \begin{align}\label{pl_profit}
	\pi_P = \alpha(p_A D_A + p_B D_B) + \gamma CS
	\end{align}

	where $CS$ is a consumer surplus, $\gamma$ represents the platform's degree of consumer-orientedness. I assume that the maximum rate of commission the platform can charge the sellers is $\bar{\alpha} \in (0, 1)$, and $\gamma > \bar{\alpha}$.

	The goal of the platform is to choose the commission rate $\alpha$ and the recommender policy in order to maximize \eqref{pl_profit}.

	\subsection{Timing}
	The timing of the game looks as follows:
	\begin{enumerate}
		\item The platform commits to the commission rate $\alpha$ and the recommender policy.
		\item The sellers privately observes $q_A, q_B$ and decide whether to disclose it to the platform.
		\item The sellers simultaneously set prices.
		\item The platform observes prices and whether the sellers disclosed quality, and recommends sellers to customers. After that the payoffs are realized and the game ends.
	\end{enumerate}

	\section{Results}
	Firstly, I consider the case with observable qualities and without the recommender system. I need this comparison in order to demonstrate how and via which mechanisms the presence of the recommender system affects equilibrium behavior of sellers.

	\subsection{Observable qualities. Without the platform}

	\begin{lemma}\label{without}
		If qualities $q_A$ and $q_B$ are observable then the demand for each good and equilibrium prices are defined as follows:
		\begin{align*}
		D_i = x_0 &= \frac{1}{2} - \frac{p_i - p_{-i}}{2} + \frac{q_i-q_{-i}}{2}\\
		p_i &= 1 + \frac{q_i - q_{-i}}{3}
		\end{align*}
	\end{lemma}
\begin{proof}
	The proof will be completed.
\end{proof}
	The lemma above demonstrates that without the platform the seller with the higher quality sets a higher price, and if seller $i$ decides to undercut the competitor and sets the price $p_{-i} - \varepsilon$ then the gain in additional demand will be $\frac{\varepsilon}{2}$.

	In the next subsection I introduce the platform.

	\subsection{Observable qualities. With the platform.}
	For simplicity, I assume firstly that $\alpha$ is fixed.
	\begin{proposition} \label{with}
		If qualities $q_A$ and $q_B$ are observable then the recommender system induces the following demands for each good and equilibrium prices:
		\begin{align*}
		D_A &= \hat{x} = \frac{1}{2} - \frac{\gamma-\alpha}{2 \gamma}(p_A - p_B) + \frac{q_A - q_B}{2}\\
		D_B &= 1 - D_A\\
		p_i &= \frac{\gamma}{\gamma - \alpha} + \frac{\gamma(q_{i} -q_{-i})}{3(\gamma-\alpha)}
		\end{align*}
		and the optimal recommender policy implies:
		\begin{itemize}
			\item recommending the seller $A$ for those customers whose location $x \le \hat{x}$
			\item recommending the seller $B$ otherwise
		\end{itemize}
	\end{proposition}
	\begin{proof}
		The proof will be completed.
	\end{proof}
	Comparing the results of the Proposition \ref{with} with Lemma \ref{without} several things are worth noting. First of all, with the recommender system the term $\frac{p_A-p_B}{2}$ in the expression for demand has an extra multiplier $\frac{\gamma-\alpha}{\gamma} \in (0, 1)$. The latter fact means that is the seller $A$ decides to undercut the competitor and sets a price $p_B - \varepsilon$, with the recommender system the resulting gain in demand will be $\frac{\gamma - \alpha}{\gamma} \frac{\varepsilon}{2}$ which is strictly below $\frac{\varepsilon}{2}$ the result we obtained for the case without the platform. That means that now the sellers have fewer incentives to undercut the competitor, i.e. the presence of recommender system which has stakes in sellers' profits softens competition. As a result, the equilibrium prices a higher. The picture below (Fig. \ref{fig1}) demonstrates graphically the impact of the recommender system.

	\begin{figure}[H]
		\centering
		\includegraphics[width=0.6\textwidth]{ImpactRS}
		\caption{Impact of RS $(p_A < p_B)$}\label{fig1}
	\end{figure}
	As we can see, there exists a share of customers $\Delta = \frac{\alpha}{2 \gamma} (p_B - p_A)$ such that they strictly prefer the firm $A$ to the firm $B$ while are recommended the firm $B$. This happens since from the platform's standpoint the losses in consumer surplus are lower than gains driven by high-price sales. Note, also, that $\frac{\partial \Delta}{\partial \alpha} > 0$, $\frac{\partial \Delta}{\partial \gamma} < 0$ and $\lim_{\gamma \to \infty} \Delta = 0$. That is, the impact of the platform eliminates completely if $\gamma$ becomes very high, and the more the platform is profit-oriented (the higher is $\alpha$) the higher is $\Delta$.
	\subsection{Qualities are not observable}
	Here I consider the case when qualities $q_A, q_B$ are not observable to the platform unless the sellers decides to disclose it. In this subsection I derive the optimal recommender system of the platform. Following \cite{levin2009quality} I consider symmetric PBE in which the sellers disclose the quality if and only if the quality falls above some threshold $q^*$ and do no not disclose otherwise, and if the seller does not disclose, the beliefs of the platform regarding the quality do not depend on prices.

	The optimal platform's policy can be found separately: firstly the platform may obtain the optimal recommendation (as a function of $\alpha$) treating the commission rate $\alpha$ as given, and then find an optimal commission rate. The next proposition characterize the optimal recommendation strategy given the commission rate $\alpha$.
	\begin{proposition}\label{precise}
		Given $\alpha$, the optimal recommender policy is given by $$\hat{x} = \frac{1}{2} - \frac{\gamma-\alpha}{2 \gamma}(p_A - p_B) + \frac{\tilde{q}_A - \tilde{q}_B}{2}$$
		Equilibrium prices are $$p_i = \frac{\gamma}{\gamma - \alpha} + \frac{\gamma(\tilde{q}_{i} -\tilde{q}_{-i})}{3(\gamma-\alpha)}$$ where $$\tilde{q}_i = \begin{cases}
		q_i, \text{ if firm }i\text{ discloses }\\
		\frac{q^*}{2}, \text{ otherwise}
		\end{cases}$$
		and
		\begin{align*}
		q^* = -\frac{5}{3} + \frac{1}{3 \gamma} \sqrt{25 \gamma^2 + \frac{216 c \gamma (\gamma - \alpha)}{1 - \alpha}}
		\end{align*}
	\end{proposition} 
\begin{proof}
	Will be completed
\end{proof}
Comparing this result to the optimal recommendation strategy in case of unobservable quality, we note that instead of real qualities, in all formulas are used ``perceived'' qualities i.e. the exact qualities if the firm discloses and the conditional expectation of the quality given the fact that the firm does not disclose. The threshold $q^*$ is found as a quality that renders the seller indifferent between costly disclosure and non-disclosure. The following step would be to find an optimal commission rate $\alpha$ but prior to that let me analyse the equilibrium information disclosure and how the recommender system affects it.
\begin{proposition}\label{thres}
	\begin{enumerate}[(i)]
		\item If $\gamma <1$ then $\frac{\partial q^*}{\partial \alpha} < 0$ i.e. sellers disclose more information if commissions increase
		\item If $\gamma > 1$ then $\frac{\partial q^*}{\partial \alpha} > 0$ i.e. sellers disclose less information if commissions increase
		\item If $\gamma = 1$ then $\frac{\partial q^*}{\partial \alpha} = 0$ i.e. amount of information disclosed does not depend on the commission rate
	\end{enumerate}
\end{proposition}
\begin{proof}
	Proof will be completed.
\end{proof}
The figure below (Fig. \ref{fig2}) demonstrate the threshold $q^*$ as a function of $\alpha$.
	\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{thres}
	\caption{$q^*$ as a function of $\alpha$}\label{fig2}
\end{figure}

Let me explain the intuition behind the Proposition \ref{thres}. A choice of $\alpha$ has several potential effects on firms' strategies. On one hand, an increase in commission rate, other things equal, decreases sellers' profits since they should pay more to the platform. This makes it less attractive to disclose the information. On the other hand, an increase in commission increases the platform's stakes in sellers' profit, hence prices and sellers' profits go up. As a result, the latter effect makes the disclosure more attractive. The total effect is determined by the platform's degree of consumer-orientedness $\gamma$. If $\gamma > 1$ the first effect prevails, i.e. an increase in prices is not enough for sellers to offset losses from higher commissions. If $\gamma < 1$ the second effect dominates, and sellers benefit more from higher prices than lose from higher commissions. If $\gamma = 1$ the two effects completely offset each other.

Finally, let me summarize the optimal recommender policy by finding an optimal commission rate $\alpha$.
\begin{proposition}\label{opt_alpha}
	\begin{enumerate}[(i)]
		\item If $\gamma < 1$ then $\frac{\partial \pi}{\partial \alpha} > 0$, so the optimal recommendation policy implies the highest commission $\alpha = \bar{\alpha}$
		\item If $\gamma > 1$ then $\frac{\partial \pi}{\partial \alpha} < 0$, so the optimal recommendation policy implies $\alpha = 0$
	\end{enumerate}
\end{proposition}
\begin{proof} Will be completed.
\end{proof}

The Proposition \ref{opt_alpha} demonstrates that it turns out that the highest possible commission may be not always in the platform's best interest. If $\gamma$ is high enough it is not optimal for the platform to set commission. This happens since as we see from the Proposition \ref{thres}, if the platform charges commission, sellers will disclose less information but this brings losses in $CS$ and since $\gamma$ is high, these losses do not outweight direct benefits from high commissions.

It turns out that the optimality of the recommendation policy described in Proposition \ref{opt_alpha} strongly relies upon the assumption that the platform perfectly observes the customers' preferences (location on the line). In the next subsection this assumption is relaxed. In this scenario, there arise another effects that change the optimal recommender policy implemented by the platform.

\subsection{Not precise recommendations}
In this section I relax the assumption that the platform perfectly observes the customers' locations. A common way (see e.g. \cite{lewis1994supplying} ) to model this is to assume that prior to making a recommendation the platform can receive a noisy signal regarding the customer's location. I believe that this assumption is not very far away from reality: usually platforms collect data and implement some data-analysis tools to estimate people's preferences. Clearly, in most cases these estimates are not 100\% precise, so sometimes the platform receives a wrong signal regarding the customer's preferences.

Formally, I assume that before making recommendation to a particular customer whose true location $x$ is equal to $x_0 \in [0, 1]$ the platform receives a signal $s$ such that $$\mathbb{P}(s=x_0|x=x_0) = \beta \in [0, 1]$$ and with probability $1-\beta$, $s \sim U(0, 1)$. I will refer to $\beta$ as the recommender system precision. In this case, as it is shown by \cite{li2018recommender} the conditional expectation regarding the true location is equal to $$\mathbb{E}[x|s=x_0] = \beta x_0 + \frac{1-\beta}{2}$$
The result above is quite intuitive: the higher is precision the more weight the platform assigns to the signal, while when the precision is low the platform more frequently infers that the signal is not correct and prescribes to the customer an average location $\frac{1}{2}$.

The main goal of this subsection is to demonstrate how this recommender system's inaccuracy may affect the platform's optimal recommendation strategy and optimal commission rate. Moreover, I find an optimal precision for the recommendation system. Interesting, it turns out that it is not always optimal to have $\beta = 1$.

\begin{proposition}\label{notprecise}
	Given the commission rate $\alpha$ and precision $\beta$ the optimal recommender policy implies:
	$$\hat{x} = \frac{1}{2} - \frac{\gamma - \alpha}{\gamma \beta} \frac{p_A-p_B}{2} + \frac{\tilde{q}_A - \tilde{q}_B}{2 \beta}$$
	Equilibrium prices are
	$$p_i = \frac{\gamma}{\gamma - \alpha} \beta + \frac{\gamma(\tilde{q}_i - \tilde{q}_{-i})}{3(\gamma - \alpha)}$$
	where
	$$\tilde{q}_i = \begin{cases}
	q_i, \text{ if firm }i\text{ discloses }\\
	\frac{q^*}{2}, \text{otherwise}
	\end{cases}$$
	and $$q^* = -\frac{6 \beta - 1}{3} + \frac{1}{3 \gamma} \sqrt{(6 \beta - 1)^2\gamma^2 + \frac{216 \beta c \gamma (\gamma - \alpha)}{1-\alpha}}$$
\end{proposition}
\begin{proof}
	Will be completed.
\end{proof}
Note that for $\beta = 1$ i.e. for the highest precision of the recommender system, we get the exact result of the Proposition \ref{precise}. Several observations are worth noting from the Proposition above. First of all, low precision makes the impact of the recommender system less prominent. We have a multiplier $\frac{\gamma - \alpha}{2 \gamma \beta}$ before the term $(p_A - p_B)$ in the threshold which determines the demand for sellers. The lower is $\beta$ the stronger becomes incentives for the sellers to undercut the competitor. This can be explained in the following way: now there is a chance that the platform mistakes regarding the location of the customer, in this case, the platform makes a recommendation assuming that it is an average customer (location $\frac{1}{2}$). The latter increases the share of customers in the middle whom the sellers have the highest incentives to compete for.

Interestingly, qualitatively the behavior of the threshold $q^*$ as a function of $\alpha$ does not depend on the fact that the recommender system is not precise. That is, the Proposition \ref{thres} is true even if $\beta < 1$.
	\section{Discussion}
	\section{Extensions}
	\section{Conclusion}
	\newpage
	\bibliography{lib}{}
	%\bibliographystyle{unsrt}
	%\bibliographystyle{apalike}
	\bibliographystyle{abbrvnat}
\end{document}

### 2020-04-03 Thesis.tex
Add files via upload
Thesis

@@
 \date{}
-\author{Danil Fedchenko}
+\author{Danil Fedchenko\thanks{I am very grateful to Sergei Izmalkov for his help during the work on this project}}
 \begin{document} % 
@@
 	\end{abstract}
+\newpage
+\tableofcontents
 \newpage
@@
 		\begin{align*}
-		D_i = x_0 &= \frac{1}{2} - \frac{p_i - p_{-i}}{2} + \frac{q_i-q_{-i}}{2}\\
+		D_A &= x_0 = \frac{1}{2} - \frac{p_A - p_{B}}{2} + \frac{q_A-q_{B}}{2}\\
+		D_B &= 1 - D_A\\
 		p_i &= 1 + \frac{q_i - q_{-i}}{3}
@@
 \begin{proof}
-	The proof will be completed.
+	The standard way to solve the Hotelling model is to define a marginal consumer $x_0$ i.e.
+a consumer who is indifferent between buying from sellers $A$ or $B$.
+Thus, $x_0$ is determined by the following equation: $$V+q_A-p_A - x_0 = V+q_B - p_B - (1-x_0)$$
+	hence $$x_0 = \frac{1}{2} - \frac{p_A - p_B}{2} + \frac{q_A - q_B}{2}$$
+	Sellers set prices $p_A, p_B$ simultaneously to solve the following problems:
+	$$p_A \cdot x_0 = p_A \left(\frac{1}{2} - \frac{p_A - p_B}{2} + \frac{q_A - q_B}{2}\right) \to \underset{p_A}{\max}$$
+	$$p_B \cdot (1-x_0) = p_B \left(\frac{1}{2} + \frac{p_A - p_B}{2} - \frac{q_A - q_B}{2}\right) \to \underset{p_B}{\max}$$
+	The system of first order conditions is given by:
+	\begin{align*}
+	\begin{cases}
+	p_A = \frac{1}{2} + \frac{p_B}{2} + \frac{q_A - q_B}{2}\\
+	p_B = \frac{1}{2} + \frac{p_A}{2} - \frac{q_A - q_B}{2}
+	\end{cases}
+	\end{align*}
+	Solving the system above yields $p_i = 1 + \frac{q_i - q_{-i}}{3}$
 \end{proof}
@@
 	\begin{proof}
-		The proof will be completed.
+	 The optimal recommender strategy consists of two functions $r_A(x, p_A, p_B, q_A, q_B)$, $r_B(x, p_A, p_B, q_A, q_B)$ (which for simplicity I will denote as simply $r_A(x), r_B(x)$) which are the probabilities for the customer with location $x$ to be recommended the seller $A$ and $B$ respectively.
+Clearly these two functions should satisfy $$0 \le r_i(x) \le 1,\ \forall\ i \in \{A, B\},\ \forall\ x \in [0, 1] $$ $$0 \le r_A(x) + r_B(x) \le 1,\ \forall\ x \in [0, 1]$$
+	 Firstly, let me show that it is not optimal to have $r_A(x) + r_B(x) < 1$ for any $x$.
+Suppose on the contrary that $\exists\ x \in [0, 1], r_A(x)+r_B(x) < 1$.
+If for that customer any of expressions $V+q_A-p_A-x$, $V+q_B-p_B-(1-x)$ are positive then the platform can increase the probability of recommendation the seller whose good yields the customer positive utility (if both yields positive the good that yields more).
+Since the customer derives positive utility, $CS$ increases, this, in turn, increases the platform's objective function, hence $r_A(x) +r_B(x) < 1$ was not optimal.
+If both expressions $V+q_A-p_A-x$, $V+q_B-p_B-(1-x)$ are negative customer will not follow the recommendation no matter what are $r_A(x), r_B(x)$, so, without loss of generality we can restrict attention only to cases $r_A(x) + r_B(x) = 1$ i.e.
+to find only $r_A(x)$.
+
+	 The objective of the recommender system is \begin{align*}
+	 \int_{0}^1 \alpha p_A r_A(x)dx + \int_{0}^1 \alpha p_B r_B(x)dx + \gamma \biggl( \int_{0}^1 (V+q_A-p_A-x)r_A(x)dx + \\+\int_{0}^1 (V+q_B-p_B-(1-x))r_B(x)dx \biggr) = \int_0^1 (\gamma V + \gamma q_A - (\gamma - \alpha)p_A - \gamma x) r_A(x)dx +\\+ \int_0^1 (\gamma V + \gamma q_B - (\gamma - \alpha)p_B - \gamma(1- x))(1- r_A(x))dx
+	 \end{align*}
+	 The platform maximizes the above function subject to the constraint $0 \le r_A(x) \le 1,\ \forall\ x \in [0, 1]$. Clearly, the objective function is maximized for $$r_A(x) = \begin{cases}
+	 1, \text{if the expression under the first integral is higher}\\
+	 0, \text{if the expression under the second integral is higher}
+	 \end{cases}$$
+	 That is, the seller $A$ is recommended to customers whose locations are to the left of $$\hat{x} = \frac{1}{2} - \frac{\gamma - \alpha}{\gamma} \frac{p_A - p_B}{2} + \frac{q_A - q_B}{2}$$
+	 and the seller $B$ is recommended otherwise.
+
+	 Given the demand, 	sellers set prices $p_A, p_B$ simultaneously to solve the following problems:
+	 $$(1-\alpha)\cdot p_A \cdot \hat{x} = p_A \left(\frac{1}{2} - \frac{\gamma - \alpha}{\gamma} \frac{p_A - p_B}{2} + \frac{q_A - q_B}{2}\right) \to \underset{p_A}{\max}$$
+	 $$(1-\alpha)\cdot p_B \cdot (1-\hat{x}) = p_B \left(\frac{1}{2} + \frac{\gamma - \alpha}{\gamma} \frac{p_A - p_B}{2} - \frac{q_A - q_B}{2}\right) \to \underset{p_B}{\max}$$
+	 The system of first order conditions is given by:
+	 \begin{align*}
+	 \begin{cases}
+	 p_A = \frac{\gamma}{2(\gamma - \alpha)} + \frac{p_B}{2} + \frac{\gamma(q_A - q_B)}{2(\gamma - \alpha)}\\
+	 p_B = \frac{\gamma}{2(\gamma - \alpha)} + \frac{p_A}{2} - \frac{\gamma(q_A - q_B)}{2(\gamma - \alpha)}
+	 \end{cases}
+	 \end{align*}
+	 Solving the system above yields $p_i = \frac{\gamma}{\gamma - \alpha} + \frac{\gamma(q_i - q_{-i})}{3(\gamma - \alpha)}$
 	\end{proof}
@@
 		and
-		\begin{align*}
+		\begin{align}\label{threshold}
 		q^* = -\frac{5}{3} + \frac{1}{3 \gamma} \sqrt{25 \gamma^2 + \frac{216 c \gamma (\gamma - \alpha)}{1 - \alpha}}
-		\end{align*}
+		\end{align}
 	\end{proposition} 
 \begin{proof}
-	Will be completed
+	From the Bayes rule it follows that if the seller does not disclose the information the platform forms beliefs that the quality is $\frac{q^*}{2}$.
+Given the platform's beliefs regarding the quality, the proof exactly repeats that of Proposition \ref{with}.
+It remains to find $q^*$.
+This threshold is found as the minimal quality that renders the seller indifferent between disclosing and not disclosing the information.
+That is, $q^*$ is determined as a solution to the following equation:
+	\begin{align*}
+	\mathbb{E} [\pi_i|q_i=q^*, \text{disclose}] - c = \mathbb{E} [\pi_i|q_i=q^*, \text{do not disclose}]
+	\end{align*}
+	\begin{align*}
+	&(1-\alpha)\int_{0}^{q^*} \left(\frac{\gamma}{\gamma - \alpha} + \frac{\gamma\left(q^* - \frac{q^*}{2}\right)}{3(\gamma - \alpha)} \right) \left(\frac{1}{2} - \frac{q^* - \frac{q^*}{2}}{6}\right)dq_B +\\ &+(1-\alpha)\int_{q^*}^{1} \left(\frac{\gamma}{\gamma - \alpha} + \frac{\gamma(q^* - q_B)}{3(\gamma - \alpha)} \right) \left(\frac{1}{2} - \frac{q^* - q_B}{6}\right)dq_B - c = \\
+	&=(1-\alpha)\int_{0}^{q^*} \left(\frac{\gamma}{\gamma - \alpha} + \frac{\gamma\left(\frac{q^*}{2} - \frac{q^*}{2}\right)}{3(\gamma - \alpha)} \right) \left(\frac{1}{2} - \frac{\frac{q^*}{2} - \frac{q^*}{2}}{6}\right)dq_B +\\
+	&+ (1-\alpha)\int_{q^*}^{1} \left(\frac{\gamma}{\gamma - \alpha} + \frac{\gamma(\frac{q^*}{2} - q_B)}{3(\gamma - \alpha)} \right) \left(\frac{1}{2} - \frac{\frac{q^*}{2} - q_B}{6}\right)dq_B
+	\end{align*}
+	Solving the equation above yields
+	$$q^* = -\frac{5}{3} + \frac{1}{3 \gamma} \sqrt{25 \gamma^2 + \frac{216 c \gamma (\gamma - \alpha)}{1 - \alpha}}$$
 \end{proof}
@@
 \begin{proof}
-	Proof will be completed.
+	The proof is straightforward, differentiating \eqref{threshold} with respect to $\alpha$, one can get:
+	\begin{align*}
+	\frac{\partial q^*}{\partial \alpha} = - \frac{36 c (1-\gamma)}{\sqrt{25 \gamma^2 + \frac{216 c \gamma (\gamma - \alpha)}{1 - \alpha}}} 
+	\end{align*}
+	which immediately implies the result.
 \end{proof}
@@
 \end{proposition}
-\begin{proof} Will be completed.
+\begin{proof} Given prices $p_A, p_B$ and qualities $q_A, q_B$, total benefit of the platform is equal to:
+	\begin{align*}
+	\alpha(p_A \hat{x} + p_B(1-\hat{x})) + \gamma \left(\int_{0}^{\hat{x}} (V + q_A - p_A - x)dx + \int_{\hat{x}}^1 (V + q_B - p_B -(1-x))dx\right) = \\
+	= 	\alpha(p_A \hat{x} + p_B(1-\hat{x})) + \gamma \left(V - \frac{1}{2}+(1-\hat{x})(q_B - p_B) + \hat{x}(q_A-p_A) - \hat{x}^2  \right)
+	\end{align*}
+	Where $\hat{x}$ is defined according to Proposition \ref{precise}. Hence, expected total benefit of the platform is
+	\begin{align*}
+	&\pi = \int_{0}^{q^*} \int_0^{q^*} \biggl( \frac{\alpha \gamma}{\gamma - \alpha} + \gamma \left(V - \frac{3}{4}+ \frac{q^*}{2} - \frac{\gamma}{\gamma - \alpha} \right)\biggr) dq_A dq_B +\\+
+	 &\int_0^{q^*} \int_{q^*}^1 \biggl[ \alpha \biggl( \left( \frac{\gamma}{\gamma - \alpha} + \frac{\gamma\left( \frac{q^*}{2} - q_B \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} + \frac{\frac{q^*}{2} - q_B}{6} \right) + \left( \frac{\gamma}{\gamma - \alpha} - \frac{\gamma\left( \frac{q^*}{2} - q_B \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} - \frac{\frac{q^*}{2} - q_B}{6} \right) \biggr)+\\
+	 &+\gamma \biggl(V - \frac{1}{2} + \left(\frac{1}{2} - \frac{\frac{q^*}{2} - q_B}{6} \right) \left( q_B - \frac{\gamma}{\gamma - \alpha} + \frac{\gamma \left( \frac{q^*}{2} - q_B \right)}{3 (\gamma - \alpha)} \right)  + \\
+	 &+ \left(\frac{1}{2} + \frac{\frac{q^*}{2} - q_B}{6} \right) \left(\frac{q^*}{2} - \frac{\gamma}{\gamma - \alpha} - \frac{\gamma \left( \frac{q^*}{2} - q_B \right)}{3(\gamma - \alpha)} \right)  - \left( \frac{1}{2} + \frac{\frac{q^*}{2} - q_B}{6} \right)^2 \biggr)  \biggr] dq_A dq_B +\\
+	 +
+	 &\int_{q^*}^{1} \int_{0}^{q^*} \biggl[ \alpha \biggl( \left( \frac{\gamma}{\gamma - \alpha} + \frac{\gamma\left(q_A - \frac{q^*}{2} \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} + \frac{q_A - \frac{q^*}{2}}{6} \right) + \left( \frac{\gamma}{\gamma - \alpha} - \frac{\gamma\left(q_A- \frac{q^*}{2} \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} - \frac{q_A -\frac{q^*}{2}}{6} \right) \biggr)+\\
+	 &+\gamma \biggl(V - \frac{1}{2} + \left(\frac{1}{2} - \frac{q_A-\frac{q^*}{2}}{6} \right) \left( \frac{q^*}{2} - \frac{\gamma}{\gamma - \alpha} + \frac{\gamma \left(q_A -  \frac{q^*}{2} \right)}{3 (\gamma - \alpha)} \right)  + \\
+	 &+ \left(\frac{1}{2} + \frac{q_A-\frac{q^*}{2}}{6} \right) \left(q_A - \frac{\gamma}{\gamma - \alpha} - \frac{\gamma \left(q_A- \frac{q^*}{2} \right)}{3(\gamma - \alpha)} \right)  - \left( \frac{1}{2} + \frac{q_A - \frac{q^*}{2}}{6} \right)^2 \biggr)  \biggr] dq_A dq_B + \\
+	 +
+	  &\int_{q^*}^{1} \int_{q^*}^{1} \biggl[ \alpha \biggl( \left( \frac{\gamma}{\gamma - \alpha} + \frac{\gamma\left(q_A - q_B \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} + \frac{q_A - q_B}{6} \right) + \left( \frac{\gamma}{\gamma - \alpha} - \frac{\gamma\left(q_A- q_B \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} - \frac{q_A -q_B}{6} \right) \biggr)+\\
+	 &+\gamma \biggl(V - \frac{1}{2} + \left(\frac{1}{2} - \frac{q_A-q_B}{6} \right) \left( q_B - \frac{\gamma}{\gamma - \alpha} + \frac{\gamma \left(q_A -  q_B \right)}{3 (\gamma - \alpha)} \right)  + \\
+	 &+ \left(\frac{1}{2} + \frac{q_A-q_B}{6} \right) \left(q_A - \frac{\gamma}{\gamma - \alpha} - \frac{\gamma \left(q_A- q_B \right)}{3(\gamma - \alpha)} \right)  - \left( \frac{1}{2} + \frac{q_A - q_B}{6} \right)^2 \biggr)  \biggr] dq_A dq_B = \\
+	 & = \frac{\gamma(216 V - q^{*3} - 161)}{216}
+	\end{align*}
+	Thus $$\frac{\partial \pi}{\partial \alpha} = -\frac{3 \gamma q^{*2}}{216} \frac{\partial q^*}{\partial \alpha}$$ and using the result of the Proposition \ref{thres} one may finally prove the result.
 \end{proof}
@@
 I will refer to $\beta$ as the recommender system precision.
-In this case, as it is shown by \cite{li2018recommender} the conditional expectation regarding the true location is equal to $$\mathbb{E}[x|s=x_0] = \beta x_0 + \frac{1-\beta}{2}$$
+In this case, as it is shown by \cite{li2018recommender} the conditional expectation regarding the true location is equal to \begin{align}\label{signal}
+\mathbb{E}[x|s=x_0] = \beta x_0 + \frac{1-\beta}{2}
+\end{align}
 The result above is quite intuitive: the higher is precision the more weight the platform assigns to the signal, while when the precision is low the platform more frequently infers that the signal is not correct and prescribes to the customer an average location $\frac{1}{2}$.
@@
 \begin{proof}
-	Will be completed.
+	Using \eqref{signal} and repeating the proof of the Proposition \ref{with} one can obtain the expression for $\hat{x}$ and equilibrium prices.
+$q^*$ is found in the same way as in the Proposition \ref{precise}.
 \end{proof}
@@
 That is, the Proposition \ref{thres} is true even if $\beta < 1$.
+
+However, it turns out that depending on cost of information disclosure, the amounf of information disclosed in equilibrium depends differently on the recommender system precision.
+
+\begin{proposition}\label{thres_beta}
+	\begin{enumerate}[(i)]
+		\item If $c \le \frac{\gamma(1-\alpha)}{9(\gamma - \alpha)}$ then $q^* \in [0, 1]$ and $\frac{\partial q^*}{\partial \beta} \le 0$ i.e.
+sellers disclose more information if the recommender system is more precise
+		\item If $\frac{\gamma(1-\alpha)}{9(\gamma - \alpha)} < c \le \frac{13}{8}\frac{\gamma(1-\alpha)}{9(\gamma - \alpha)}$ then $q^* \in [0, 1]$ and $\frac{\partial q^*}{\partial \beta} > 0$ i.e.
+sellers disclose less information if the recommender system is more precise
+		\item If $c > \frac{13}{8}\frac{\gamma(1-\alpha)}{9(\gamma - \alpha)}$ then $\frac{\partial q^*}{\partial \beta} > 0$ and $q^* \in [0, 1] \iff 0 \le \beta \le \frac{\gamma(1-\alpha)}{72 c (\gamma - \alpha) - 12 \gamma (1-\alpha)}$
+	\end{enumerate}
+\end{proposition}
+\begin{proof}
+	First of all, let us find a condition that guarantees that $q^* \le 1$. 
+	\begin{align*}
+	\sqrt{(6 \beta - 1)^2 \gamma^2 + \frac{216 \beta c \gamma (\gamma - \alpha)}{1-\alpha}} &\le 2\gamma (1+3 \beta)\\
+	-\gamma(1+12 \beta) + \frac{72 \beta c (\gamma - \alpha)}{1 - \alpha} &\le 0\\
+	\beta &\le \frac{\gamma(1-\alpha)}{72c (\gamma - \alpha) - 12 \gamma (1-\alpha)}
+	\end{align*}
+	Clearly, if $\frac{\gamma(1-\alpha)}{72c (\gamma - \alpha) - 12 \gamma (1-\alpha)} > 1$, $q^* \le 1$ for all $\beta$. Thus, 
+	\begin{align*}
+	\frac{\gamma(1-\alpha)}{72c (\gamma - \alpha) - 12 \gamma (1-\alpha)} &> 1\\
+	\frac{13}{72} \frac{\gamma(1-\alpha)}{\gamma - \alpha} &> c
+	\end{align*}
+	Differentiated the expression for $q^*$ with respect to $\beta$ yields:
+	\begin{align*}
+	\frac{\partial q^*}{\partial \beta} = -2 + \frac{2 \gamma (6 \beta - 1) + \frac{36 c (\gamma - \alpha)}{1-\alpha}}{ \sqrt{(6 \beta - 1)^2\gamma^2 + \frac{216 \beta c \gamma (\gamma - \alpha)}{1-\alpha}}}
+	\end{align*} 
+	\begin{align*}
+	-2 + \frac{2 \gamma (6 \beta - 1) + \frac{36 c (\gamma - \alpha)}{1-\alpha}}{ \sqrt{(6 \beta - 1)^2\gamma^2 + \frac{216 \beta c \gamma (\gamma - \alpha)}{1-\alpha}}} &\le 0 \iff\\
+	\iff \gamma(6 \beta - 1) + \frac{18c(\gamma - \alpha)}{1- \alpha} &\le \sqrt{(6 \beta - 1)^2\gamma^2 + \frac{216 \beta c \gamma (\gamma - \alpha)}{1-\alpha}} \iff\\
+	\iff 36 \gamma (6 \beta - 1)(1-\alpha) + 18^2 c(\gamma - \alpha) &\le 216 \beta \gamma (1-\alpha) \iff\\
+	\iff c &\le \frac{\gamma(1 - \alpha)}{9(\gamma - \alpha)}  
+	\end{align*} 
+	This completes the proof.
+\end{proof}
+
+What drives the result above?
+The recommender system's precision has several effects on sellers: on one hand, as we see from the Proposition \ref{notprecise} higher precision means softer competition and as a result higher prices, and hence more incentives for sellers to disclose the information.
+On the other hand, high $\beta$ renders the contribution of quality into demand lower (we have the term $\frac{\tilde{q}_i - \tilde{q}_{-i}}{2 \beta}$) in the demand for the seller's $i$ good.
+As a result, with high precision the impact of quality decreases, which makes sellers less willing to bear cost of disclosure.
+That is why, if cost of disclosure is high the latter effect dominates and improving the recommender system's precision adversely effects the amount of information disclosed in the equilibrium.
+If, on the other hand, cost of disclosure is small, higher precision increases the amount of disclosed information.
+
+This ambiguous influence of the precision on the amount of information disclosed, qualitatively changes an impact of commission rate on the platform's total benefit.
+Assuming that the recommendation precision $\beta$ is an exogenous constant, the next proposition describes the optimal policy of the platform regarding commissions it charges sellers.
+
+\begin{proposition}\label{profit_beta}
+	\leavevmode 
+	\begin{itemize}
+		\item If $\beta \le 0.5$ and  $\gamma > 1$ then $\frac{\partial \pi }{\partial \alpha} > 0$ and the optimal commission is $\alpha = \bar{\alpha}$
+		\item  If $\beta \le 0.5$ and $\gamma < 1$ then $\frac{\partial \pi }{\partial \alpha} < 0$ and the optimal commission is $\alpha = 0$	
+		\item If $\beta > 0.5$ and  $\gamma > 1$ then $\frac{\partial \pi }{\partial \alpha} < 0$ and the optimal commission is $\alpha = 0$
+		\item If $\beta > 0.5$ and $\gamma < 1$ then $\frac{\partial \pi }{\partial \alpha} > 0$ and the optimal commission is $\alpha = \bar{\alpha}$
+	\end{itemize}
+\end{proposition}
+\begin{proof}
+	content...
+\end{proof}
 	\section{Discussion}

### 2020-04-04 Thesis.tex
Add files via upload
Thesis

@@
 	\section{Literature review}
+
+		This work is related to the strand of literature on recommender systems.
+The first prototype of modern recommender system has been created by Xerox Palo Alto Research Center in the late 1980-s early 1990-s and was described by  \cite{goldberg1992using}.
+This system has been called \textit{Tapestry} and has been designed in order to filter out a skyrocketing amount of the incoming emails.
+The system was trying to make the incoming stream of emails interesting for the particular person based on whether these emails were interesting for people whose interests are close to the interests of this person.
+Since that time many papers have been written that mainly explore the recommender systems from the Computer Science standpoint (see \cite{Beel2015r} for the recent survey).
+
+	One of the first papers which highlights the issue from the point of view of Economics is work by \cite{avery1999market}.
+\cite{avery1999market} considered the influence of the recommender system on choice regarding the consumption of good for the sequentially arriving agents.
+
+	\cite{bergemann2006optimal} address the question of optimal pricing under the presence of recommender system.
+
+	In the two papers mentioned above the authors assumed the recommendation strategy to be given and mainly just consider the impact of the recommender system on other strategic agents.
+
+	One of the first works that focus on design of optimal recommendation policy is the work by \cite{kremer2014}.
+The authors build a simple model with two alternatives which generate deterministic but \textit{ex-ante} unknown payoff.
+One of the alternatives is \textit{a priori} better, in the sense that the expected reward it generates is higher.
+The finite number of consumers arrive sequentially and each chooses between either of two alternatives.
+After the choice is made the consumer leaves forever.
+The platform accumulates reviews and commits to the recommendation policy it uses to make recommendations to the newcomers.
+\cite{kremer2014}, firstly, neatly in simple terms demonstrate why the full revelation policy may be suboptimal\footnote{Maybe the appearance of this paper and similar to that in 2014 was the reason why \textit{Amazon} decided to switch their recommendation policy from the fully revealing one.}.
+Secondly, the optimal recommender policy is derived which has a threshold nature.
+That is, \textit{a priori} better alternative is recommended if its quality turns out to be higher than the certain threshold.
+Otherwise, the other alternative is recommended, and then once both alternatives were tried the best one is recommended until the end.
+
+	Another relevant paper is that of \cite{che2015}.
+In the paper, a continuous flow of customers arrives, and each should choose between the good or an outside option.
+The quality of the good is exogenously predetermined but unknown.
+Customers who decide to consume the good send a noisy signal about its quality to the platform.
+The signal is noisy in the sense that even if the quality of the good is good the information about it reaches to the platform only with some probability (different from 1).
+The platform updates its beliefs about the quality and tailors a recommender policy contingent on the beliefs.
+Authors demonstrate that the full revelation policy is suboptimal and find the optimal policy which again has a threshold nature.
+Namely, once the platform receives the signal that the quality is good, it will recommend the good until the end.
+Otherwise, if no news arrives it is optimal to experiment with the risky alternative (recommend it) until the beliefs that its quality is good fall below the certain threshold.
+After that, the optimal policy prescribes to never recommend this alternative.
+
+	\cite{papanastasiou2017} extend ideas developed by \cite{kremer2014} and \cite{che2015}, and build a dynamic, stochastic model.
+In their model, agents also arrive sequentially to the platform and should choose between two alternatives that generate stochastically exogenous payoff.
+The platform in the paper designs a recommendation policy.
+Although the optimal policy turns out to have a complicated form, authors demonstrate that in this case the full revelation policy is not the best one, and the optimal policy involves information obfuscation.
+
+	Three papers mentioned above although address the question on optimal recommender policy, do not take into account the strategic nature of recommended alternatives.
+	\cite{chen2016advertising} consider the informative role of recommender system on strategies of competing sellers and the equilibrium outcomes.
+They consider two sellers located at the extremes of Hotelling-like (\cite{harold1929stability}) linear city who are competing for the unit mass of customers, located in between.
+There is also a recommender system that receives a noisy information about each customer's location and can recommend one of the sellers to the customers.
+Although in this paper, the recommendation strategy is exogenous, the authors consider how the presence of the recommender system affects competition between sellers.
+In particular, they are looking at prices and advertising expenditure.
+The main finding of this paper is that the presence of the recommender system may fundamentally change the nature of competition.
+In particular, the authors demonstrate that with the recommender system prices and advertising expenditure switch from being strategic substitutes to strategic complements.
+This results in the fact that with recommender system sellers prefer to advertise less, relying more on advertising provided by the recommendation system (the authors called this effect \textit{advertising effect}).
+At the same time, prices with recommender system are lower (the authors called this effect \textit{competition effect}) which is the result of tougher competition.
+Interestingly, the total effect of recommender system on sellers profits becomes a result of an interplay between \textit{competition effect} and \textit{advertising effect}: for some values of parameters the sellers becomes worse-off with the presence of the recommender system, \textit{competition effect} dominates, while for other values the sellers benefit from ``free'' advertising provided by the platform.
+
+	Another paper, which is closely related to the one mentioned previously is the paper by \cite{li2018recommender} (actually there are two quite different versions of this paper).
+In the two papers mentioned above, the qualities of goods were supposed to be equal and known \textit{ex-ante} to all parties which makes the type of differentiation considered to be purely horizontal.
+
+	\cite{levin2009quality} add a vertical component, focusing on the sellers' decisions regarding the quality disclosure, although, in the framework without the recommender system.
+In the model which I consider, I firstly endogenise the recommendation strategy that the platform implements, and secondly, I am looking at the interplay between the recommendation policy and sellers' incentives to disclose the information about the goods' quality.
+
 	\section{Model}
@@
 The main goal of this subsection is to demonstrate how this recommender system's inaccuracy may affect the platform's optimal recommendation strategy and optimal commission rate.
-Moreover, I find an optimal precision for the recommendation system.
-Interesting, it turns out that it is not always optimal to have $\beta = 1$.
 
@@
 	$$\hat{x} = \frac{1}{2} - \frac{\gamma - \alpha}{\gamma \beta} \frac{p_A-p_B}{2} + \frac{\tilde{q}_A - \tilde{q}_B}{2 \beta}$$
+	i.e. those customers for whom signal regarding their location falls to the left of $\hat{x}$ are recommended the seller $A$, otherwise they are recommended the seller $B$.
 	Equilibrium prices are
@@
 The latter increases the share of customers in the middle whom the sellers have the highest incentives to compete for.
+As a result, other things equal, equilibrium prices are lower, the less precise is the recommender system implemented by the platform.
 
@@
 
-However, it turns out that depending on cost of information disclosure, the amounf of information disclosed in equilibrium depends differently on the recommender system precision.
+However, it turns out that depending on cost of information disclosure, the amount of information disclosed in equilibrium depends differently on the recommender system precision.
 
@@
 \begin{proof}
-	content...
+ First of all, we need to find the platform's expected profit. Given prices $p_A, p_B$ and qualities $q_A, q_B$, total benefit of the platform is equal to:
+ \begin{align}\label{profit}
+ \alpha(p_A \hat{x} + p_B(1-\hat{x})) + \gamma \biggl(\int_{0}^{\hat{x}} \biggl[ (\beta +(1-\beta)\hat{x})(V + q_A - p_A - x) + \nonumber \\+ (1-\beta)(1-\hat{x})(V + q_B - p_B - (1-x))\biggr]dx + \int_{\hat{x}}^1 \biggl[(1-\beta) \hat{x}(V+q_A-p_A-x) + \nonumber\\
+ +(\beta + (1-\beta)(1-\hat{x}))(V + q_B - p_B -(1-x)) \biggr] dx\biggr) = \\
+ = 	\alpha(p_A \hat{x} + p_B(1-\hat{x})) + \gamma \left(V - \frac{1}{2}+(1-\hat{x})(q_B - p_B) + \hat{x}(q_A-p_A) + \beta \hat{x} - \beta \hat{x}^2  \right)\nonumber
+ \end{align}
+ Where $\hat{x}$ is defined according to Proposition \ref{notprecise}.
+For those customers located to the left of $\hat{x}$ the probability of being recommended a seller $A$ is $\beta + (1-\beta) \hat{x}$.
+Since with probability $\beta$ the platform receives the true location of customer and with probability $(1-\beta) \hat{x}$ the platform receives a signal from the interval $[0, \hat{x}]$.
+In the following way, one can obtain probabilities of being recommended seller $B$ for those customers located to the left of $\hat{x}$ as well as probabilities of being recommended sellers $A$ and $B$ for the customers located to the right of $\hat{x}$.
+Finally, we obtain an expression in \eqref{profit}.
+Then, in the same way as in the Proposition \ref{opt_alpha} one can obtain the expression for the profit:
+ \begin{align*}
+ \pi = \frac{\gamma(54 \beta^2(1+4V) - (1-q^{*3})(1-2\beta) - 216 \beta^3)}{216 \beta^2}
+ \end{align*}
+ Thus, \begin{align*}
+ \frac{\partial \pi }{\partial \alpha} = \frac{\gamma q^{*2}(1-2\beta)}{72 \beta^2} \frac{\partial q^*}{\partial \alpha}
+ \end{align*}
+ Using the result of the Proposition \ref{thres} we finally obtain the result we were to prove.
 \end{proof}
+In contrast to the case with precise recommendation system (Proposition \ref{precise}), if the platform's precision is not very high (lower than 0.5), it may be optimal to charge maximal commission even in case when $\gamma > 1$.
+This happens since with very low precision the weight of quality for the recommendation becomes higher, as a result, sellers have incentives to reveal quality even with high commissions.
+In the similar fashion can be explained the fact that for low precision not very consumer-oriented recommender system (with $\gamma < 1$) prefers to set $\alpha = 0$: low precision adversely affects the consumer surplus, the situation deteriorates if in addition sellers decide not to disclose the information.
+Consequently, it is optimal for the platform to set $\alpha = 0$.
 	\section{Discussion}

### 2020-04-05 Thesis.tex
Add files via upload
Thesis

@@
 \cite{avery1999market} considered the influence of the recommender system on choice regarding the consumption of good for the sequentially arriving agents.
+They model they consider include sequentially arriving agents who can either consume the good and leave a review for the next agents or he can wait until somebody else will try the good and report the review.
+The main focus of the work is on finding a socially optimal mechanism, in particular, the authors propose a mechanism which includes monetary transfers to some agents to induce them to try the good.
 
 	\cite{bergemann2006optimal} address the question of optimal pricing under the presence of recommender system.
+They consider the seller with recommender system who is competing for unit mass of heterogeneous buyers with the competitive fringe of sellers who do not deploy the recommender system.
+The seller with recommender system uses reviews from past customers to make recommendations to the newcomers.
+The authors consider several type of equilibria concluding that the presence of the recommender system creates an additional value by reducing uncertainty for agents about the quality of the good.
 
@@
 
-	One of the first works that focus on design of optimal recommendation policy is the work by \cite{kremer2014}.
+	One of the first works that addresses the question of optimal design of the recommendation policy is the work by \cite{kremer2014}.
 The authors build a simple model with two alternatives which generate deterministic but \textit{ex-ante} unknown payoff.
@@
 	Another paper, which is closely related to the one mentioned previously is the paper by \cite{li2018recommender} (actually there are two quite different versions of this paper).
-In the two papers mentioned above, the qualities of goods were supposed to be equal and known \textit{ex-ante} to all parties which makes the type of differentiation considered to be purely horizontal.
+In the first one, considering the same set-up as \cite{chen2016advertising}, the authors do not allow sellers to advertise themselves, assuming the shares of informed and uninformed customer to be exogenous.
+The authors demonstrate that the presence of the recommender system on one hand increase the share of informed customers and hence increases demand on goods, but on the other hand it intensifies competition by making a demand more elastic.
+\cite{li2018recommender} refer to these two effects as \textit{demand effect} and \textit{substitution effect} respectively.
+In the similar way as in \cite{chen2016advertising}, the authors demonstrate that the total effect of recommender system on sellers becomes a result of interaction between these tow effects.
+Also, in this paper the authors make the first attempt to endogenise the recommender policy, showing that the recommender system may increase its profit by changing the recommendation strategy.
+In the second version of the paper the authors elaborate on this, formulating the recommender system's objective and finding an optimal recommender policy the recommender system should implement.
+In the two (in fact, three) papers mentioned above, the qualities of goods were supposed to be equal and known \textit{ex-ante} to all parties which makes the type of differentiation considered to be purely horizontal.
 
@@
 	\end{cases}$$
-	and $$q^* = -\frac{6 \beta - 1}{3} + \frac{1}{3 \gamma} \sqrt{(6 \beta - 1)^2\gamma^2 + \frac{216 \beta c \gamma (\gamma - \alpha)}{1-\alpha}}$$
+	and \begin{align}\label{thres_b}
+	q^* = -\frac{6 \beta - 1}{3} + \frac{1}{3 \gamma} \sqrt{(6 \beta - 1)^2\gamma^2 + \frac{216 \beta c \gamma (\gamma - \alpha)}{1-\alpha}}
+	\end{align}
 \end{proposition}
@@
 Then, in the same way as in the Proposition \ref{opt_alpha} one can obtain the expression for the profit:
- \begin{align*}
+ \begin{align}\label{profit_b}
  \pi = \frac{\gamma(54 \beta^2(1+4V) - (1-q^{*3})(1-2\beta) - 216 \beta^3)}{216 \beta^2}
- \end{align*}
+ \end{align}
  Thus, \begin{align*}
@@
 Consequently, it is optimal for the platform to set $\alpha = 0$.
+
+In the exposition above I treated the precision of the recommender system $\beta$ as given, since I think it may be a reasonable assumption for the platform in the short-run.
+In real life, this precision is usually determined by the amount of data the platform collected about customers as well as the complexity of algorithm the platform employs to process and use this data.
+And once the amount of data and the algorithm are chosen and  fixed, the precision cannot be improved in the short-run and, hence, may be supposed to be fixed.
+At the same time, the platform may invest resources into increasing the precision in the long-run.
+And the question that logically arises is: is it also beneficial for the platform to increase the precision of the recommender system?
+Interestingly, the model I consider suggests that it might not always to be the case.
+\begin{proposition}
+	Given $\alpha$, the perfect precision $\beta = 1$ is always worse for the platform than some intermediate $\beta \in (0, 1)$
+\end{proposition}
+\begin{proof}
+	Plugging \eqref{thres_b} into \eqref{profit_b} one can obtain
+	\begin{align*}
+	\pi = \frac{\gamma}{216 \beta ^2}  \biggl\{-2 \beta  \left(\frac{\left(\sqrt{\gamma  \left(\frac{216 \beta  c (\gamma-\alpha)}{1-\alpha}+(1-6
+				\beta )^2 \gamma \right)}-6 \beta  \gamma +\gamma \right)^3}{27 \gamma ^3}-1\right)+\\
+			+\frac{\left(\sqrt{\gamma 
+				\left(\frac{216 \beta  c (\gamma - \alpha)}{1-\alpha}+(1-6 \beta )^2 \gamma \right)}-6 \beta  \gamma +\gamma
+			\right)^3}{27 \gamma ^3}-216 \beta ^3+54 \beta ^2 (4 V+1)-1\biggr\}
+	\end{align*}
+	Then
+	\begin{align*}
+	\pi\bigg|_{\beta = 1} = \frac{\gamma }{216} \biggl\{\frac{\left(\sqrt{\gamma  \left(\frac{216 c (\gamma-\alpha)}{1-\alpha}+25 \gamma \right)}-5
+		\gamma \right)^3}{27 \gamma ^3}-2 \left(\frac{\left(\sqrt{\gamma  \left(\frac{216 c (\gamma-\alpha)}{1-\alpha}+25
+			\gamma \right)}-5 \gamma \right)^3}{27 \gamma ^3}-1\right)+\\
+		+54 (4 V+1)-217\biggr\}
+	\end{align*}
+	At the same time, 
+	\begin{align*}
+	\pi \bigg|_{\beta = \frac{1}{2}} = \gamma  \left(V-\frac{1}{4}\right)
+	\end{align*}
+	Consider the difference $\pi\bigg|_{\beta = 1} - \pi\bigg|_{\beta = \frac{1}{2}}$:
+	\begin{align*}
+	\pi\bigg|_{\beta = 1} - \pi\bigg|_{\beta = \frac{1}{2}} = -\frac{\gamma}{216} \biggl\{ 107+ \frac{\left(\sqrt{\gamma  \left(\frac{216 c (\gamma-\alpha)}{1-\alpha}+25 \gamma \right)}-5
+		\gamma \right)^3}{27 \gamma ^3} \biggr\} < 0
+	\end{align*}
+	Thus, setting $\beta = \frac{1}{2}$ does the platform strictly better-off comparing to the case $\beta = 1$.
+\end{proof}
 	\section{Discussion}

### 2020-04-07 Thesis.tex
Add files via upload
Thesis

@@
 \date{}
-\author{Danil Fedchenko\thanks{I am very grateful to Sergei Izmalkov for his help during the work on this project}}
+\author{Danil Fedchenko}%\thanks{I am very grateful to Sergei Izmalkov for his help during the work on this project}}
 \begin{document} % 
@@
 	\section{Introduction}
+
+		Suppose you came to New-York and would like to find a good restaurant to go out tonight.
+Or you are going to spend your upcoming vacation on the coast of Spain and you need to find a hotel.
+What would you do?
+What will be the first thing you interact with?
+Before the World Wide Web, people in such cases relied upon guide-books and brochures distributed by travel agents.
+Those guides usually quickly became out-of-date and were criticized for containing misleading information.
+
+	After the advent of the Internet, finding up-to-date information has become much easier.
+To make this process even easier Internet platforms began to appear.
+These platforms aggregated customers' reviews on products and service providers.
+Some of those platforms simply collect the reviews and present a rating that is just an average measure of consumers' satisfaction (weighted or simple mean of rates).
+Examples include \textit{Yelp} for restaurants, home and car services, or \textit{SiteJabber} for online businesses.
+
+Other platforms came further and began to not only collect the reviews but also to make individual recommendations based on these reviews.
+Usually, these recommendations take a form of ranking of several alternatives that are ranked according to the secret propriety algorithm that the particular platform employs.
+The most widely known examples of such a kind of platforms include: \textit{TripAdvisor} which presents a ranking on Hotels and Restaurants based on previous users' experience; \textit{Trustpilot} which is one of the most popular review services for a large variety of businesses; \textit{Glassdoor} that uses reviews of employees to rank companies by workers' satisfaction; and many other examples.
+
+In the rows of examples, the special attention should be paid to online retailers such as \textit{Amazon}, \textit{eBay}, \textit{AliExpress} and many others.
+These platforms stand at the forefront of online Economics, generating huge amounts of money and completely transforming the shopping experience for people all other the world.
+These platforms play roles of intermediaries between buyers and sellers, helping the buyers to find the goods they need.
+And this process clearly includes making recommendations which seller to display you when you are searching for something.
+
+	All these services discussed above use some variety of what is called \textit{recommender system} which may be defined as any system which uses information about past users' experience in order to make the recommendations to the newcomers.
+These recommendations may be direct or indirect (e.g.
+the platform can provide a ranking of alternatives, this is not the direct recommendation to choose the top-ranked alternative, however, users understand that the platform thinks that this alternative is the best fit for them, so the platform implicitly recommends this alternative).
+
+I argue, that nowadays the recommender systems are one of \textbf{the key elements} of online Economics.
+Any Internet or mobile service you can think of almost for sure uses some form of a recommender system.
+That is why it is crucially important to understand how to recommend optimally and what factors should be taken into account once the service is designing its recommender system.
+
+For example, consider a retailer (such as \textit{Amazon} or \textit{eBay}), these retailers charge sellers commissions for each sale.
+These commissions usually take the form of a percentage commission.
+That means that the retailer obtains a higher profit if the price of the deal is higher.
+So, one of the questions that could potentially arise then is: should, in this case, the recommender system which Amazon employs out of two sellers who sell (almost) substitutable goods recommend the one who has a higher price?
+Or suppose that \textit{eBay} is thinking on the commission rate it charges (now it is about 8\%), should the commission be set at the highest possible level that guarantees that sellers do not refrain from using the platform?
+What if, at the same time, the quality of goods may differ and the sellers may have an option to spend some resources on disclosing this quality (e.g.
+to make the detailed description or take high-resolution photos).
+Clearly the higher is the commission the less willing are the sellers to spend money on quality disclosure: cost does not outweigh benefits.
+But if sellers do not disclose the information this may adversely affect the customers' satisfaction since this may create a mismatch between buyers and sellers: people end up with the good which does not fit them well.
+Thus, thinking about commission rate the platform should take this aspect into account.
+Additionally, suppose that the recommendation system is not precise i.e.
+it sometimes makes mistakes on figuring out the customer's preferences.
+How does this inaccuracy affect the behavior (price setting) of competing sellers and how does it interact with their incentives to disclose the quality of their goods?
+What should be the optimal commissions and recommender policy in this case?
+Although in this work I am touching upon all questions mentioned above, the main objective of this paper is to understand what should be the optimal recommendation strategy and commission rate the platform charges the competitive sellers, with the particular emphasis on the interaction with sellers' incentives to disclose private information about their goods.
+
+	To address these questions I build a theoretical one-period model (please see the detailed description of the model in the corresponding section).
+I consider the unit mass of heterogeneous customers whose preferences are summarized by the point on this unit line.
+At the extremes of the unit interval, there are two sellers.
+Customers demand one good and derive utility upon consuming the good but bear ``transportation'' cost by purchasing.
+This cost, of course, has little to do with any transportation but rather represents the outside option, opportunity cost of the customer: the utility the customer should refrain from in order to consume the good.
+I suppose that the qualities of goods offered by sellers can be different and that these qualities are \textit{ex-ante} privately known only to the corresponding sellers who can costly disclose this information before setting prices.
+The main player in this system is the recommender system (the platform) who beforehand receives (possibly noisy) signal regarding each customer's location and then may recommend any of the two sellers to this customer.
+This feature of my model has clear real-life interpretation: location of the customer represents his or her tastes, the signal which reveals the location can be interpreted as the result of some fancy machine-learning algorithm the majority of platform nowadays use in order to assess, to estimate the true preferences, tastes of the customer.
+The platform also sets a commission rate (percentage) it charges the sellers for each sale.
+However, I explicitly assume that the platform cares not only about the profit (commissions) but also about customers i.e.
+consumer surplus directly enters the platform's objective function.
+I suppose that this assumption is not very far away from reality: clearly, if the platform cares only about the profit and this adversely affects customers' satisfaction and happiness from using the service, in the long-run people simply will refuse using it or switch to the competitive platforms.
+The latter suggests that for the long-run sustainability the platform should pay some attention to the consumer surplus.
+
+In the work, I find the optimal recommender policy and investigate the effect of the presence of the recommender system on equilibrium behavior of sellers.
+I demonstrate that the recommender system which has stakes in the sellers' profits softens competition.
+The latter effect leads to higher prices.
+The amount of information disclosed in equilibrium is determined in the interplay between two effects: on one hand, higher commission directly decreases profits and, hence, incentives of sellers to disclose the information; on the other hand, the higher commission increases platform's stakes in profits and softens competition which results in higher prices and higher profits.
+The total value of these two effects is determined by the degree of consumer-orientedness of the platform: if it is high enough, increase in prices does not outweigh direct losses, so the total effect is negative; in the opposite case total effect is positive and an increase in commission increases the amount of information disclosed.
+
+In the extension of the baseline model, I allow for the not perfectly precise recommender system.
+Namely, I assume that the signal, the platform receives reveals the true location of the customer only with some probability $\beta$, and with probability $1 - \beta$ the signal follows a uniform distribution.
+This has several consequences: first of all, now the platform adjusts its recommender strategy incorporating this inaccuracy.
+This adjustment manifests itself in the way how the platform treats and processes the signal regarding the location, now the platform takes into consideration that the signal could potentially be wrong.
+The latter effectively increases the share of customers in the middle, increasing incentives for sellers to undercut the competitor and to get access to this increased demand.
+Thus, the direct effect of the platform's inaccuracy intensifies the competition between sellers.
+This decreases the sellers' profits and as a result, decreases sellers' incentives to reveal the quality.
+At the same time, the recommender system's inaccuracy increases the attention the platform pays to the quality of the good when it makes recommendations.
+The latter effect increases the sellers' returns on quality disclosure making the latter more profitable.
+Eventually, the total effect also depends on parameters: if the cost of disclosure is high enough, the first effect dominates and improving the recommender system precision decreases the amount of information the sellers disclose, for the low cost of disclosure the second effect prevails and increasing the platform's precision increases the amount of information disclosed.
+
+However, it turns out that increasing the precision of the recommender system up to 1 is not the optimal thing to do.
+The last thing I show in this work is that there exists some optimal level of $\beta \in (0, 1)$.
+This can be explained as follows: making a recommender system a bit imprecise, intensifies competition and lowers prices, at the same time if precision is not too low, the direct negative effect on consumer surplus is not very high.
+As a result, I show that some noise in the platform's evaluation of the customers' preferences can be beneficial for the platform.
+
+The rest of the paper is organized as follows: the next section presents the literature review, Section 3 describes the model, Section 4 contains the main results, and last three Sections encompass a brief discussion, possible extensions and directions for the future research, and a short conclusion.
+
 	\section{Literature review}
@@
 
-	One of the first papers which highlights the issue from the point of view of Economics is work by \cite{avery1999market}.
-\cite{avery1999market} considered the influence of the recommender system on choice regarding the consumption of good for the sequentially arriving agents.
-They model they consider include sequentially arriving agents who can either consume the good and leave a review for the next agents or he can wait until somebody else will try the good and report the review.
+One of the first papers which highlight the issue from the point of view of Economics is work by \cite{avery1999market}.
+\cite{avery1999market} considered the influence of the recommender system on a choice regarding the consumption of good for the sequentially arriving agents.
+The model they consider include sequentially arriving agents who can either consume the good and leave a review for the next agents or he can wait until somebody else will try the good and report the review.
 The main focus of the work is on finding a socially optimal mechanism, in particular, the authors propose a mechanism which includes monetary transfers to some agents to induce them to try the good.
 
-	\cite{bergemann2006optimal} address the question of optimal pricing under the presence of recommender system.
-They consider the seller with recommender system who is competing for unit mass of heterogeneous buyers with the competitive fringe of sellers who do not deploy the recommender system.
-The seller with recommender system uses reviews from past customers to make recommendations to the newcomers.
-The authors consider several type of equilibria concluding that the presence of the recommender system creates an additional value by reducing uncertainty for agents about the quality of the good.
+\cite{bergemann2006optimal} address the question of optimal pricing under the presence of a recommender system.
+They consider the seller with a recommender system who is competing for a unit mass of heterogeneous buyers with the competitive fringe of sellers who do not deploy the recommender system.
+The seller with the recommender system uses reviews from past customers to make recommendations to the newcomers.
+The authors consider several types of equilibria concluding that the presence of the recommender system creates additional value by reducing uncertainty for agents about the quality of the good.
 
@@
 They consider two sellers located at the extremes of Hotelling-like (\cite{harold1929stability}) linear city who are competing for the unit mass of customers, located in between.
-There is also a recommender system that receives a noisy information about each customer's location and can recommend one of the sellers to the customers.
+There is also a recommender system that receives noisy information about each customer's location and can recommend one of the sellers to the customers.
 Although in this paper, the recommendation strategy is exogenous, the authors consider how the presence of the recommender system affects competition between sellers.
@@
 At the same time, prices with recommender system are lower (the authors called this effect \textit{competition effect}) which is the result of tougher competition.
-Interestingly, the total effect of recommender system on sellers profits becomes a result of an interplay between \textit{competition effect} and \textit{advertising effect}: for some values of parameters the sellers becomes worse-off with the presence of the recommender system, \textit{competition effect} dominates, while for other values the sellers benefit from ``free'' advertising provided by the platform.
+Interestingly, the total effect of recommender system on sellers profits becomes a result of an interplay between \textit{competition effect} and \textit{advertising effect}: for some values of parameters the sellers become worse-off with the presence of the recommender system, \textit{competition effect} dominates, while for other values the sellers benefit from ``free'' advertising provided by the platform.
 
@@
 In the first one, considering the same set-up as \cite{chen2016advertising}, the authors do not allow sellers to advertise themselves, assuming the shares of informed and uninformed customer to be exogenous.
-The authors demonstrate that the presence of the recommender system on one hand increase the share of informed customers and hence increases demand on goods, but on the other hand it intensifies competition by making a demand more elastic.
+The authors demonstrate that the presence of the recommender system, on one hand, increase the share of informed customers and hence increases demand on goods, but on the other hand, it intensifies competition by making demand more elastic.
 \cite{li2018recommender} refer to these two effects as \textit{demand effect} and \textit{substitution effect} respectively.
-In the similar way as in \cite{chen2016advertising}, the authors demonstrate that the total effect of recommender system on sellers becomes a result of interaction between these tow effects.
+In a similar way as in \cite{chen2016advertising}, the authors demonstrate that the total effect of the recommender system on sellers becomes a result of interaction between these two effects.
 Also, in this paper the authors make the first attempt to endogenise the recommender policy, showing that the recommender system may increase its profit by changing the recommendation strategy.
-In the second version of the paper the authors elaborate on this, formulating the recommender system's objective and finding an optimal recommender policy the recommender system should implement.
+In the second version of the paper, the authors elaborate on this, formulating the recommender system's objective and finding an optimal recommender policy the recommender system should implement.
 In the two (in fact, three) papers mentioned above, the qualities of goods were supposed to be equal and known \textit{ex-ante} to all parties which makes the type of differentiation considered to be purely horizontal.
@@
 In the model which I consider, I firstly endogenise the recommendation strategy that the platform implements, and secondly, I am looking at the interplay between the recommendation policy and sellers' incentives to disclose the information about the goods' quality.
-
 	\section{Model}
@@
 
-	There is a unit mass of heterogenous customers.
+There is a unit mass of heterogeneous customers.
 Preferences of customer are characterized by the location of the customer on the unity interval.
 Two competing sellers (firms) which I will refer to $A$ and $B$, who locate at the extremes of the unit interval, are selling substitutable goods which they produce at equal marginal cost normalized to 0.
-Goods of each seller may have different quality which means that the customers with the same ``horizontal preference'' for the two sellers may derive different utility upon consuming goods of different sellers.
+Goods of each seller may have a different quality which means that the customers with the same ``horizontal preference'' for the two sellers may derive different utility upon consuming goods of different sellers.
 Sellers are competing in prices which they set simultaneously.
 In addition, the sellers may costly disclose the information about the quality of their good prior to price setting.
-The recommender system observes prices and sellers' decisions regarding the quality disclosure, and makes recommendation to customers which seller to choose.
-In the baseline model I assume that the recommender system knows the exact location of the customer, lately this assumption is relaxed, and I consider the scenario when the recommender system receives a noisy signal regarding the customers' locations.
+The recommender system observes prices and sellers' decisions regarding the quality disclosure and makes a recommendation to customers which seller to choose.
+In the baseline model, I assume that the recommender system knows the exact location of the customer, lately, this assumption is relaxed, and I consider the scenario when the recommender system receives a noisy signal regarding the customers' locations.
 Finally, the platform charges sellers a commission for sales according to a commission rate $\alpha$ which the platform can choose.
 
-	The solution concept is perfect Bayesian equilibrium, and all parties are assumed to be risk-neutral.
+The solution concept is the perfect Bayesian equilibrium, and all parties are assumed to be risk-neutral.
 
-	The next three subsections describes formally all parties of my model.
+The next three subsections describe formally all parties of my model.
 
@@
 Prior to price-setting sellers privately learns the quality of their good and should decide whether to disclose this information to the platform or not.
-Upon disclosure the sellers bear cost $c$.
+Upon disclosure, the sellers bear cost $c$.
 Moreover, the sellers pay the platform commission from sales.
@@
 	\item The sellers simultaneously set prices.
-		\item The platform observes prices and whether the sellers disclosed quality, and recommends sellers to customers. After that the payoffs are realized and the game ends.
+	\item The platform observes prices and whether the sellers disclosed quality, and recommends sellers to customers. After that, the payoffs are realized and the game ends.
 \end{enumerate}
@@
 Firstly, I consider the case with observable qualities and without the recommender system.
-I need this comparison in order to demonstrate how and via which mechanisms the presence of the recommender system affects equilibrium behavior of sellers.
+I need this comparison in order to demonstrate how and via which mechanisms the presence of the recommender system affects the equilibrium behavior of sellers.
 
@@
 That means that now the sellers have fewer incentives to undercut the competitor, i.e.
-the presence of recommender system which has stakes in sellers' profits softens competition.
+the presence of a recommender system which has stakes in sellers' profits softens competition.
 As a result, the equilibrium prices a higher.
@@
 	\subsection{Qualities are not observable}
-	Here I consider the case when qualities $q_A, q_B$ are not observable to the platform unless the sellers decides to disclose it.
-In this subsection I derive the optimal recommender system of the platform.
-Following \cite{levin2009quality} I consider symmetric PBE in which the sellers disclose the quality if and only if the quality falls above some threshold $q^*$ and do no not disclose otherwise, and if the seller does not disclose, the beliefs of the platform regarding the quality do not depend on prices.
+Here I consider the case when qualities $q_A, q_B$ are not observable to the platform unless the sellers decide to disclose it.
+In this subsection, I derive the optimal recommender system of the platform.
+Following \cite{levin2009quality} I consider symmetric PBE in which the sellers disclose the quality if and only if the quality falls above some threshold $q^*$ and do not disclose otherwise, and if the seller does not disclose, the beliefs of the platform regarding the quality do not depend on prices.
 
 The optimal platform's policy can be found separately: firstly the platform may obtain the optimal recommendation (as a function of $\alpha$) treating the commission rate $\alpha$ as given, and then find an optimal commission rate.
-The next proposition characterize the optimal recommendation strategy given the commission rate $\alpha$.
+The next proposition characterizes the optimal recommendation strategy given the commission rate $\alpha$.
+
 	\begin{proposition}\label{precise}
@@
 \begin{proof}
-	From the Bayes rule it follows that if the seller does not disclose the information the platform forms beliefs that the quality is $\frac{q^*}{2}$.
+From the Bayes rule, it follows that if the seller does not disclose the information the platform forms beliefs that the quality is $\frac{q^*}{2}$.
 Given the platform's beliefs regarding the quality, the proof exactly repeats that of Proposition \ref{with}.
@@
 This threshold is found as the minimal quality that renders the seller indifferent between disclosing and not disclosing the information.
-That is, $q^*$ is determined as a solution to the following equation:
-	\begin{align*}
+That is, $q^*$ is determined as a solution to the following equation	\begin{align*}
 	\mathbb{E} [\pi_i|q_i=q^*, \text{disclose}] - c = \mathbb{E} [\pi_i|q_i=q^*, \text{do not disclose}]
@@
 If $\gamma$ is high enough it is not optimal for the platform to set commission.
-This happens since as we see from the Proposition \ref{thres}, if the platform charges commission, sellers will disclose less information but this brings losses in $CS$ and since $\gamma$ is high, these losses do not outweight direct benefits from high commissions.
+This happens since as we see from the Proposition \ref{thres}, if the platform charges commission, sellers will disclose less information but this brings losses in $CS$ and since $\gamma$ is high, these losses do not outweigh direct benefits from high commissions.
 
 It turns out that the optimality of the recommendation policy described in Proposition \ref{opt_alpha} strongly relies upon the assumption that the platform perfectly observes the customers' preferences (location on the line).
-In the next subsection this assumption is relaxed.
-In this scenario, there arise another effects that change the optimal recommender policy implemented by the platform.
+In the next subsection, this assumption is relaxed.
+In this scenario, there arise other effects that change the optimal recommender policy implemented by the platform.
 
@@
 We have a multiplier $\frac{\gamma - \alpha}{2 \gamma \beta}$ before the term $(p_A - p_B)$ in the threshold which determines the demand for sellers.
-The lower is $\beta$ the stronger becomes incentives for the sellers to undercut the competitor.
+The lower is $\beta$ the stronger become incentives for the sellers to undercut the competitor.
 This can be explained in the following way: now there is a chance that the platform mistakes regarding the location of the customer, in this case, the platform makes a recommendation assuming that it is an average customer (location $\frac{1}{2}$).
@@
 
-However, it turns out that depending on cost of information disclosure, the amount of information disclosed in equilibrium depends differently on the recommender system precision.
+However, it turns out that depending on the cost of information disclosure, the amount of information disclosed in equilibrium depends differently on the recommender system precision.
 
@@
 On the other hand, high $\beta$ renders the contribution of quality into demand lower (we have the term $\frac{\tilde{q}_i - \tilde{q}_{-i}}{2 \beta}$) in the demand for the seller's $i$ good.
-As a result, with high precision the impact of quality decreases, which makes sellers less willing to bear cost of disclosure.
-That is why, if cost of disclosure is high the latter effect dominates and improving the recommender system's precision adversely effects the amount of information disclosed in the equilibrium.
-If, on the other hand, cost of disclosure is small, higher precision increases the amount of disclosed information.
+As a result, with high precision the impact of quality decreases, which makes sellers less willing to bear the cost of disclosure.
+That is why, if the cost of disclosure is high the latter effect dominates and improving the recommender system's precision adversely affects the amount of information disclosed in the equilibrium.
+If, on the other hand, the cost of disclosure is small, higher precision increases the amount of disclosed information.
 
@@
 \end{proof}
-In contrast to the case with precise recommendation system (Proposition \ref{precise}), if the platform's precision is not very high (lower than 0.5), it may be optimal to charge maximal commission even in case when $\gamma > 1$.
+In contrast to the case with precise recommendation system (Proposition \ref{precise}), if the platform's precision is not very high (lower than 0.5), it may be optimal to charge maximal commission even in the case when $\gamma > 1$.
 This happens since with very low precision the weight of quality for the recommendation becomes higher, as a result, sellers have incentives to reveal quality even with high commissions.
-In the similar fashion can be explained the fact that for low precision not very consumer-oriented recommender system (with $\gamma < 1$) prefers to set $\alpha = 0$: low precision adversely affects the consumer surplus, the situation deteriorates if in addition sellers decide not to disclose the information.
+In the similar fashion can be explained the fact that for low precision not very consumer-oriented recommender system (with $\gamma < 1$) prefers to set $\alpha = 0$: low precision adversely affects the consumer surplus, the situation deteriorates if in addition, sellers decide not to disclose the information.
 Consequently, it is optimal for the platform to set $\alpha = 0$.
 
-In the exposition above I treated the precision of the recommender system $\beta$ as given, since I think it may be a reasonable assumption for the platform in the short-run.
-In real life, this precision is usually determined by the amount of data the platform collected about customers as well as the complexity of algorithm the platform employs to process and use this data.
+In the exposition above I treated the precision of the recommender system $\beta$ as given since I think it may be a reasonable assumption for the platform in the short-run.
+In real life, this precision is usually determined by the amount of data the platform collected about customers as well as the complexity of the algorithm the platform employs to process and use this data.
 And once the amount of data and the algorithm are chosen and fixed, the precision cannot be improved in the short-run and, hence, may be supposed to be fixed.
@@
 Interestingly, the model I consider suggests that it might not always to be the case.
+
 \begin{proposition}
@@
 	\end{align*}
-	Thus, setting $\beta = \frac{1}{2}$ does the platform strictly better-off comparing to the case $\beta = 1$.
+	Thus, setting $\beta = \frac{1}{2}$ makes the platform strictly better-off comparing to the case $\beta = 1$.
 \end{proof}
+The result above although seems strange at first glance, can be explained.
+Low precision definitely creates cases of a mismatch for a customer (i.e.
+some customers are not recommended the best alternatives, even from the platform's standpoint), as a result, the direct effect on consumer surplus is negative.
+But at the same time, the low precision intensifies competition, as well as renders the disclosure of information more attractive for sellers, since the lower is precision the more disclosed quality matters for the recommendation.
+Thus, the total effect of decreasing precision from $\beta = 1$ to some intermediate value is positive.
+
+Making the recommendation system too inaccurate is also not optimal for the platform.
+It can be easily shown that the platform's profit \eqref{profit_b} tends to minus infinity once $\beta \to 0$.
+In other words, there exists some optimal level of precision that lies strictly between 0 and 1.
 	\section{Discussion}

### 2020-04-19 Thesis.tex
Add files via upload
Thesis

@@
 \usepackage[utf8]{inputenc}
+\usepackage{pgf,tikz,pgfplots}
+\usetikzlibrary{patterns}
 \usepackage{setspace,amsmath}
@@
 \date{}
-\author{Danil Fedchenko}%\thanks{I am very grateful to Sergei Izmalkov for his help during the work on this project}}
-\begin{document} % 
+\author{Danil Fedchenko\thanks{I am very grateful to: Sergei Izmalkov for constant help and support during the work on this project; Marta Troya Martinez for many valuable comments and suggestions.
+\newline Any remaining errors are mine.}}
+\begin{document} 
 	\maketitle
@@
 Or you are going to spend your upcoming vacation on the coast of Spain and you need to find a hotel.
+Or you would like to buy something using your favorite online retailer.
 What would you do?
 What will be the first thing you interact with?
-Before the World Wide Web, people in such cases relied upon guide-books and brochures distributed by travel agents.
+Before the World Wide Web, to find hotels or restaurants people relied upon guide-books and brochures distributed by travel agents.
 Those guides usually quickly became out-of-date and were criticized for containing misleading information.
@@
 In the rows of examples, the special attention should be paid to online retailers such as \textit{Amazon}, \textit{eBay}, \textit{AliExpress} and many others.
-These platforms stand at the forefront of online Economics, generating huge amounts of money and completely transforming the shopping experience for people all other the world.
+These platforms stand at the forefront of online Economics, generating huge amounts of money\footnote{e.g.
+\textit{Amazon}'s gross merchandise value (total value of goods sold on the platform) in 2019 fiscal year account for \$335 billion.} and completely transforming the shopping experience for people all other the world.
 These platforms play roles of intermediaries between buyers and sellers, helping the buyers to find the goods they need.
@@
 That means that the retailer obtains a higher profit if the price of the deal is higher.
-So, one of the questions that could potentially arise then is: should, in this case, the recommender system which Amazon employs out of two sellers who sell (almost) substitutable goods recommend the one who has a higher price?
-Or suppose that \textit{eBay} is thinking on the commission rate it charges (now it is about 8\%), should the commission be set at the highest possible level that guarantees that sellers do not refrain from using the platform?
+So, one of the questions that could potentially arise then is: should, in this case, the recommender system which \textit{Amazon} employs out of two sellers who sell (almost) substitutable goods recommend the one who has a higher price?
+Or suppose that \textit{eBay} is thinking on the commission rate it charges, %(now it is about 8\%),
+ should the commission be set at the highest possible level that guarantees that sellers do not refrain from using the platform?
 What if, at the same time, the quality of goods may differ and the sellers may have an option to spend some resources on disclosing this quality (e.g.
@@
 
-	To address these questions I build a theoretical one-period model (please see the detailed description of the model in the corresponding section).
-I consider the unit mass of heterogeneous customers whose preferences are summarized by the point on this unit line.
+	To address these questions I build a theoretical one-period model.
+I consider a unit mass of heterogeneous customers whose preferences are summarized by the point on this unit line.
 At the extremes of the unit interval, there are two sellers.
 Customers demand one good and derive utility upon consuming the good but bear ``transportation'' cost by purchasing.
-This cost, of course, has little to do with any transportation but rather represents the outside option, opportunity cost of the customer: the utility the customer should refrain from in order to consume the good.
-I suppose that the qualities of goods offered by sellers can be different and that these qualities are \textit{ex-ante} privately known only to the corresponding sellers who can costly disclose this information before setting prices.
-The main player in this system is the recommender system (the platform) who beforehand receives (possibly noisy) signal regarding each customer's location and then may recommend any of the two sellers to this customer.
-This feature of my model has clear real-life interpretation: location of the customer represents his or her tastes, the signal which reveals the location can be interpreted as the result of some fancy machine-learning algorithm the majority of platform nowadays use in order to assess, to estimate the true preferences, tastes of the customer.
+This cost although has nothing to do with transportation can be interpreted in several possible ways.
+For example, it can be think of as cost of the outside option, opportunity cost of the customer: the utility the customer should refrain from in order to consume the good; or as the measure of quality of fit: how well the product fits the particular customer.
+The latter allows to treat the location of the customer as the customer's taste regarding the two goods offered by sellers: closer the customer is to the seller $A$ -- better the good of seller $A$ fits this customer, other things being equal.
+Moreover, I suppose that the qualities of goods offered by sellers can be different.
+In other words, all customers regardless of their location get the same additional value upon consuming the good with higher quality.
+I assume that these qualities are \textit{ex-ante} privately known only to the corresponding sellers who can costly disclose this information before setting prices.
+As I mentioned earlier, the cost of disclosure may include time and monetary cost of making a detailed description of the good, or taking many high-resolution photos of the item.
+The main player in this interaction is the recommender system (the platform) who beforehand receives (possibly noisy) signal regarding each customer's location and then may recommend any of the two sellers to this customer.
+This feature of my model has clear real-life interpretation: the signal which reveals the location can be interpreted as the result of some fancy machine-learning algorithm the majority of platform nowadays use in order to assess, to estimate the true preferences, tastes of the customer.
 The platform also sets a commission rate (percentage) it charges the sellers for each sale.
@@
 
-In the work, I find the optimal recommender policy and investigate the effect of the presence of the recommender system on equilibrium behavior of sellers.
+	Also, I argue that the model I consider, as a whole, fits well the research questions I am addressing.
+The recommender systems usually help when customers are horizontally differentiated: there can be 10 restaurants of top quality, but you may love only one of them simply because it fits your tastes better (e.g.
+you just like Indian food).
+In this case, a good recommender system should recommend you precisely this restaurant out of ten.
+If, on the contrary, everybody agreed that restaurant $A$ is in all respects worse than the restaurant $B$, nobody would need the recommender system.
+
+In the work, I find the optimal recommendation policy and investigate the effect of the presence of the recommender system on equilibrium behavior of sellers.
 I demonstrate that the recommender system which has stakes in the sellers' profits softens competition.
 The latter effect leads to higher prices.
-The amount of information disclosed in equilibrium is determined in the interplay between two effects: on one hand, higher commission directly decreases profits and, hence, incentives of sellers to disclose the information; on the other hand, the higher commission increases platform's stakes in profits and softens competition which results in higher prices and higher profits.
-The total value of these two effects is determined by the degree of consumer-orientedness of the platform: if it is high enough, increase in prices does not outweigh direct losses, so the total effect is negative; in the opposite case total effect is positive and an increase in commission increases the amount of information disclosed.
+The amount of information about the quality of goods, disclosed in equilibrium is determined in the interplay between two effects: on one hand, higher commission directly decreases profits and, hence, incentives of sellers to disclose the information; on the other hand, the higher commission increases platform's stakes in profits and softens competition which results in higher prices and higher profits.
+The total value of these two effects is determined by the degree of consumer-orientedness of the platform: if it is high enough, increase in prices does not outweigh direct losses, so the total effect on the amount of disclosed information is negative; in the opposite case total effect is positive and an increase in commission increases the amount of information disclosed.
 
@@
 This adjustment manifests itself in the way how the platform treats and processes the signal regarding the location, now the platform takes into consideration that the signal could potentially be wrong.
-The latter effectively increases the share of customers in the middle, increasing incentives for sellers to undercut the competitor and to get access to this increased demand.
+In this case, the best prediction of the customer's location the platform could made is simply average location.
+The latter effectively increases the share of customers in the middle of the unity interval, increasing incentives for sellers to undercut the competitor and to get access to this increased demand.
 Thus, the direct effect of the platform's inaccuracy intensifies the competition between sellers.
 This decreases the sellers' profits and as a result, decreases sellers' incentives to reveal the quality.
-At the same time, the recommender system's inaccuracy increases the attention the platform pays to the quality of the good when it makes recommendations.
+At the same time, the recommender system's inaccuracy increases the attention the platform pays to the quality of the good when it makes recommendations: since the platform assumes that horizontal component (location) may be measured imprecisely, making a recommendation it begins to put more weights on vertical component (quality) which, if disclosed, is precise.
 The latter effect increases the sellers' returns on quality disclosure making the latter more profitable.
@@
 There is also a recommender system that receives noisy information about each customer's location and can recommend one of the sellers to the customers.
-Although in this paper, the recommendation strategy is exogenous, the authors consider how the presence of the recommender system affects competition between sellers.
+Although in this paper, the recommendation strategy is exogenously given and fixed, the authors consider how the presence of the recommender system affects competition between sellers.
 In particular, they are looking at prices and advertising expenditure.
@@
 
-\cite{levin2009quality} add a vertical component, focusing on the sellers' decisions regarding the quality disclosure, although, in the framework without the recommender system.
+\cite{levin2009quality} add a vertical component to the standard Hotelling model, focusing on the sellers' decisions regarding the quality disclosure, although, in the framework without the recommender system.
 In the model which I consider, I firstly endogenise the recommendation strategy that the platform implements, and secondly, I am looking at the interplay between the recommendation policy and sellers' incentives to disclose the information about the goods' quality.
@@
 Sellers are competing in prices which they set simultaneously.
-In addition, the sellers may costly disclose the information about the quality of their good prior to price setting.
+In addition, the sellers may costly disclose the information about the quality of their good to the platform prior to price setting.
 The recommender system observes prices and sellers' decisions regarding the quality disclosure and makes a recommendation to customers which seller to choose.
@@
 Customer who locates at $x \in [0, 1]$ derives a net utility $$V + q_A - p_A - x$$ if  he buys the good at a price $p_A$ from the seller $A$ who locates at a point $0$, and net utility $$V + q_B - p_B - (1-x)$$ if  he buys the good at a price $p_B$ from the seller $B$ who locates at a point $1$.
-$V$ is assumed to be higher than $\frac{3}{2}$ to ensure that the market is fully covered in equilibrium, $q_A, q_B \in [0, 1]$ are the qualities of the goods.
+$V$ is assumed to be higher than $\frac{3}{2}$ to ensure that the market is fully covered in equilibrium, $q_A, q_B \in (0, 1)$ are the qualities of the goods.
+These qualities are \textit{ex-ante} known only to corresponding sellers, and are $i.i.d.\  U(0, 1)$.
 The outside option for the customers is zero, so they follow the recommendation if the expected net utility from buying the good from the recommended seller is higher than 0.
 
-Customers do not know about the sellers unless the recommender system recommends them.
+Customers do not know about the sellers (hence cannot sell) unless the recommender system recommends them.
 
@@
 The platform commits to the commission rate $\alpha$ and the recommender policy.
-The recommender policy specifies functions $r_A(x, p_A, p_B, d_A, d_B, q_A, q_B)$ and $r_B(x, p_A, p_B, d_A, d_B, q_A, q_B)$ which represents probabilities that the customer with location $x$ will be recommended the seller $A$ and $B$ respectively.
+The recommender policy specifies functions $r_A(x, p_A, p_B, d_A, d_B, \mathbb{E}q_A, \mathbb{E}q_B)$ and $r_B(x, p_A, p_B, d_A, d_B, \mathbb{E} q_A, \mathbb{E} q_B)$ which represents probabilities that the customer with location $x$ will be recommended the seller $A$ and $B$ respectively.
 The objective function of the platform is given by \begin{align}\label{pl_profit}
-\pi_P = \alpha(p_A D_A + p_B D_B) + \gamma CS
+\pi_P = \mathbb{E}\left[\alpha(p_A D_A + p_B D_B) + \gamma CS\right]
 \end{align}
 
-where $CS$ is a consumer surplus, $\gamma$ represents the platform's degree of consumer-orientedness.
+where expectation is taken over unknown qualities $q_A, q_B$; $CS$ is a consumer surplus, $\gamma$ represents the platform's degree of consumer-orientedness.
 I assume that the maximum rate of commission the platform can charge the sellers is $\bar{\alpha} \in (0, 1)$, and $\gamma > \bar{\alpha}$.
+Usually in real life there are several competing platforms, so the platform cannot increase the commission up to 100\% since in this case sellers will just switch to the competitive online retail platform.
+So, $\bar{\alpha}$ can be think of as a maximum commission rate the platform can charge that still guarantees that sellers are willing to sell their goods through the platform.
 
@@
 \section{Results}
-Firstly, I consider the case with observable qualities and without the recommender system.
+Firstly, I consider the case with observable to buyers qualities and without the recommender system.
 I need this comparison in order to demonstrate how and via which mechanisms the presence of the recommender system affects the equilibrium behavior of sellers.
@@
 	\centering
-		\includegraphics[width=0.6\textwidth]{ImpactRS}
+	\begin{tikzpicture}[scale=1.5]
+	\draw [fill = black] (0, 0) rectangle (6.5, 0.03);
+	\draw (6.5, -0.4) node[color=red]{$B$};
+	\draw (0, 0.34) node[]{$0$};
+	\draw (6.5, 0.34) node[]{$1$};
+	\draw (-0.05, -0.4) node[color=blue]{$A$};
+	\draw (3.25, -0.47) node[]{$\frac{1}{2}$};
+	\draw (3.65, -0.4) node[]{$\hat{x}$};
+	\draw (4.46, -0.4) node[]{ $x_0$};
+	\draw [fill = black] (0, -0.1) rectangle (0.02, 0.12);
+	\draw [fill = black] (3.24, -0.1) rectangle (3.26, 0.12);
+	\draw [fill = black] (6.48, -0.1) rectangle (6.5, 0.12);
+	\draw [fill = black] (3.64, -0.1) rectangle (3.66, 0.12);
+	\draw [fill = black] (4.45, -0.1) rectangle (4.47, 0.12);
+	\draw[pattern=north west lines, pattern color=blue] (0,0) rectangle (3.64,0.2);
+	\draw[pattern=north west lines, pattern color=red] (3.66,0) rectangle (6.48,0.2);
+	\draw[pattern=north west lines, pattern color=blue] (0,0) rectangle (4.45,-0.2);
+	\draw[pattern=north west lines, pattern color=red] (4.47,0) rectangle (6.48,-0.2);
+	\draw [dashed] (3.65,0) -- (3.65,0.7);
+	\draw [dashed] (4.46,0) -- (4.46,0.7);
+	\draw [<->] (3.65,0.7) -- (4.46,0.7);
+	\draw (4.05, 0.9) node[]{$\Delta$};
+	\end{tikzpicture}
 	\caption{Impact of RS $(p_A < p_B)$}\label{fig1}
@@
 I will refer to $\beta$ as the recommender system precision.
-In this case, as it is shown by \cite{li2018recommender} the conditional expectation regarding the true location is equal to \begin{align}\label{signal}
+In this case, it can be shown (see  \cite{li2018recommender} for example) that the conditional expectation regarding the true location is equal to \begin{align}\label{signal}
 \mathbb{E}[x|s=x_0] = \beta x_0 + \frac{1-\beta}{2}

### 2020-05-02 Thesis.tex
Thesis

@@
 \usepackage{pgfplots}
+\usepackage[toc,page]{appendix}
 \usepackage{titlesec}
@@
 \date{}
-\author{Danil Fedchenko\thanks{I am very grateful to: Sergei Izmalkov for constant help and support during the work on this project; Marta Troya Martinez for many valuable comments and suggestions.
+\author{Danil Fedchenko}%\thanks{I am very grateful to: Sergei Izmalkov for constant help and support during the work on this project; Marta Troya Martinez for many valuable comments and suggestions.
 \newline Any remaining errors are mine.}}
@@
 
-		Suppose you came to New-York and would like to find a good restaurant to go out tonight.
+		Suppose you came to New York and would like to find a good restaurant to go out tonight.
 Or you are going to spend your upcoming vacation on the coast of Spain and you need to find a hotel.
-Or you would like to buy something using your favorite online retailer.
-What would you do?
+Or you would like to buy something at your favorite online retailer.
+How would you proceed?
 What will be the first thing you interact with?
 Before the World Wide Web, to find hotels or restaurants people relied upon guide-books and brochures distributed by travel agents.
-Those guides usually quickly became out-of-date and were criticized for containing misleading information.
+Those guides, once printed, usually quickly became out-of-date in the constantly changing environment, and could be criticized for containing misleading information.
 
 	After the advent of the Internet, finding up-to-date information has become much easier.
-To make this process even easier Internet platforms began to appear.
-These platforms aggregated customers' reviews on products and service providers.
+To make this process even easier Internet platforms began appearing.
+These platforms aggregate customers' reviews on products and service providers.
 Some of those platforms simply collect the reviews and present a rating that is just an average measure of consumers' satisfaction (weighted or simple mean of rates).
@@
 Other platforms came further and began to not only collect the reviews but also to make individual recommendations based on these reviews.
-Usually, these recommendations take a form of ranking of several alternatives that are ranked according to the secret propriety algorithm that the particular platform employs.
+Usually, these recommendations take a form of ranking of several alternatives, ranked according to the secret propriety algorithm that the particular platform employs.
 The most widely known examples of such a kind of platforms include: \textit{TripAdvisor} which presents a ranking on Hotels and Restaurants based on previous users' experience; \textit{Trustpilot} which is one of the most popular review services for a large variety of businesses; \textit{Glassdoor} that uses reviews of employees to rank companies by workers' satisfaction; and many other examples.
 
-In the rows of examples, the special attention should be paid to online retailers such as \textit{Amazon}, \textit{eBay}, \textit{AliExpress} and many others.
+In the rows of examples, we should pay the special attention to online retailers such as \textit{Amazon}, \textit{eBay}, \textit{AliExpress} and many others.
 These platforms stand at the forefront of online Economics, generating huge amounts of money\footnote{e.g.
-\textit{Amazon}'s gross merchandise value (total value of goods sold on the platform) in 2019 fiscal year account for \$335 billion.} and completely transforming the shopping experience for people all other the world.
+\textit{Amazon}'s gross merchandise value (total value of goods sold on the platform) in 2019 fiscal year was \$335 billion.} and completely transforming the shopping experience for people all other the world.
 These platforms play roles of intermediaries between buyers and sellers, helping the buyers to find the goods they need.
-And this process clearly includes making recommendations which seller to display you when you are searching for something.
+And this process clearly includes making recommendations: which seller to display the buyer when he or her is searching for something.
 
-	All these services discussed above use some variety of what is called \textit{recommender system} which may be defined as any system which uses information about past users' experience in order to make the recommendations to the newcomers.
+	All these services discussed above use some variety of what is called \textit{recommender system} which may be defined as any system which uses information about past users' experience in order to make recommendations to newcomers.
 These recommendations may be direct or indirect (e.g.
@@
 I argue, that nowadays the recommender systems are one of \textbf{the key elements} of online Economics.
-Any Internet or mobile service you can think of almost for sure uses some form of a recommender system.
+Any Internet or mobile service you can think of, almost for sure uses some form of a recommender system.
 That is why it is crucially important to understand how to recommend optimally and what factors should be taken into account once the service is designing its recommender system.
@@
 For example, consider a retailer (such as \textit{Amazon} or \textit{eBay}), these retailers charge sellers commissions for each sale.
-These commissions usually take the form of a percentage commission.
+These commissions usually take the form of a percentage commission \footnote{e.g.
+\textit{eBay} charges 8\% for each sale}.
 That means that the retailer obtains a higher profit if the price of the deal is higher.
@@
 What should be the optimal commissions and recommender policy in this case?
-Although in this work I am touching upon all questions mentioned above, the main objective of this paper is to understand what should be the optimal recommendation strategy and commission rate the platform charges the competitive sellers, with the particular emphasis on the interaction with sellers' incentives to disclose private information about their goods.
+Thus, in this work I am focusing on the following questions:
+ \begin{itemize}
+ 	\item How does the presence of the recommender system which has stakes in sellers' profits affect competition and equilibrium behavior of sellers?
+ 	\item How does the rate of commission the platform charges sellers affect sellers' incentives to disclose information about their goods?
+What is the optimal rate of commission from the platform's standpoint?
+ 	\item How does the recommender system's precision affect competition and equilibrium behavior? And what should be the optimal precision?
+ \end{itemize}
 
@@
 Customers demand one good and derive utility upon consuming the good but bear ``transportation'' cost by purchasing.
-This cost although has nothing to do with transportation can be interpreted in several possible ways.
-For example, it can be think of as cost of the outside option, opportunity cost of the customer: the utility the customer should refrain from in order to consume the good; or as the measure of quality of fit: how well the product fits the particular customer.
+This cost can be interpreted as the measure of quality of fit: how well the product fits the particular customer.
 The latter allows to treat the location of the customer as the customer's taste regarding the two goods offered by sellers: closer the customer is to the seller $A$ -- better the good of seller $A$ fits this customer, other things being equal.
+
 	Moreover, I suppose that the qualities of goods offered by sellers can be different.
@@
 I assume that these qualities are \textit{ex-ante} privately known only to the corresponding sellers who can costly disclose this information before setting prices.
-As I mentioned earlier, the cost of disclosure may include time and monetary cost of making a detailed description of the good, or taking many high-resolution photos of the item.
+This cost of disclosure can be interpreted in several ways.
+First of all, as I mentioned earlier, the cost of disclosure may represent actual time and monetary cost of making a detailed description of the good, or taking many high-resolution photos of the item.
+For those readers who are not convinced that it can be costly for sellers to disclose the information, I can offer another interpretation.
+Following \cite{levin2009quality} and \cite{hirshleifer2002disclosure}, we can interpret this cost as the share of naive customers who do not update their beliefs about the quality of the good if a seller does not disclose it.
+In this case, even if physically the information disclosure is costless, the seller may not disclose the information if quality is low in order not to lower the perception of quality for these naive consumers.
+
 	The main player in this interaction is the recommender system (the platform) who beforehand receives (possibly noisy) signal regarding each customer's location and then may recommend any of the two sellers to this customer.
-This feature of my model has clear real-life interpretation: the signal which reveals the location can be interpreted as the result of some fancy machine-learning algorithm the majority of platform nowadays use in order to assess, to estimate the true preferences, tastes of the customer.
+I assume that prior to interaction with the platform, customers are unaware of sellers and become aware only if the platform recommends the particular seller to the particular customer.
+The signal which the platform receives regarding the location of customers has clear real-life interpretation: it can be interpreted as the result of some fancy machine-learning algorithm the majority of platform nowadays use in order to assess, to estimate the true preferences, tastes of the customer.
+
 	The platform also sets a commission rate (percentage) it charges the sellers for each sale.
@@
 The latter suggests that for the long-run sustainability the platform should pay some attention to the consumer surplus.
+Or, if the substantial part of the platform's profit comes from advertising, the platform is interested in high stream of visitors.
+That is why, the platform should take into account users' satisfaction.
 
@@
 you just like Indian food).
-In this case, a good recommender system should recommend you precisely this restaurant out of ten.
+In this case, the recommender system creates value if it recommends you this precise restaurant.
 If, on the contrary, everybody agreed that restaurant $A$ is in all respects worse than the restaurant $B$, nobody would need the recommender system.
@@
 In the work, I find the optimal recommendation policy and investigate the effect of the presence of the recommender system on equilibrium behavior of sellers.
-I demonstrate that the recommender system which has stakes in the sellers' profits softens competition.
+I demonstrate that the recommender system which has stakes in the sellers' profits softens competition i.e.
+sellers are less willing to undercut the competitor.
 The latter effect leads to higher prices.
 The amount of information about the quality of goods, disclosed in equilibrium is determined in the interplay between two effects: on one hand, higher commission directly decreases profits and, hence, incentives of sellers to disclose the information; on the other hand, the higher commission increases platform's stakes in profits and softens competition which results in higher prices and higher profits.
-The total value of these two effects is determined by the degree of consumer-orientedness of the platform: if it is high enough, increase in prices does not outweigh direct losses, so the total effect on the amount of disclosed information is negative; in the opposite case total effect is positive and an increase in commission increases the amount of information disclosed.
+The sign of the cumulative effect is determined by the degree of consumer-orientedness of the platform: if it is high enough, increase in prices does not outweigh direct losses, so the total effect on the amount of disclosed information is negative; in the opposite case total effect is positive and an increase in commission increases the amount of information disclosed.
 
-In the extension of the baseline model, I allow for the not perfectly precise recommender system.
+In the extension of the baseline model, I allow for the not perfectly precise recommender system, or in other words, I allow for some noise in evaluating customers' location.
 Namely, I assume that the signal, the platform receives reveals the true location of the customer only with some probability $\beta$, and with probability $1 - \beta$ the signal follows a uniform distribution.
-This has several consequences: first of all, now the platform adjusts its recommender strategy incorporating this inaccuracy.
+This has several consequences: first of all, clearly the platform should adjust its recommender strategy incorporating this inaccuracy.
 This adjustment manifests itself in the way how the platform treats and processes the signal regarding the location, now the platform takes into consideration that the signal could potentially be wrong.
-In this case, the best prediction of the customer's location the platform could made is simply average location.
-The latter effectively increases the share of customers in the middle of the unity interval, increasing incentives for sellers to undercut the competitor and to get access to this increased demand.
-Thus, the direct effect of the platform's inaccuracy intensifies the competition between sellers.
+In this case, the best prediction of the customer's location the platform could made is simply an average location (since I model customers' location as uniform $(0, 1)$ random variable, an average location is $\frac{1}{2}$) .
+The latter effectively increases the share of customers in the middle of the unit interval, increasing incentives for sellers to undercut the competitor and to get access to this increased demand.
+Thus, the direct effect of the signal's inaccuracy intensifies the competition between sellers.
 This decreases the sellers' profits and as a result, decreases sellers' incentives to reveal the quality.
@@
 The latter effect increases the sellers' returns on quality disclosure making the latter more profitable.
-Eventually, the total effect also depends on parameters: if the cost of disclosure is high enough, the first effect dominates and improving the recommender system precision decreases the amount of information the sellers disclose, for the low cost of disclosure the second effect prevails and increasing the platform's precision increases the amount of information disclosed.
+Eventually, the combined effect of recommender system's inaccuracy on the amount of information disclosed by sellers depends on the value of cost parameter: if the cost of disclosure is high enough, the direct negative effect caused by more intensive competition outweighs the positive effect brought about by the increased return on disclosure and, hence, improving the recommender system precision decreases the amount of information the sellers disclose; for the low cost of disclosure the opposite is true, i.e.
+increasing the platform's precision increases the amount of information disclosed.
 
-However, it turns out that increasing the precision of the recommender system up to 1 is not the optimal thing to do.
+However, it turns out that it is not optimal for the platform to increase the precision of the recommender system up to 1.
 The last thing I show in this work is that there exists some optimal level of $\beta \in (0, 1)$.
 This can be explained as follows: making a recommender system a bit imprecise, intensifies competition and lowers prices, at the same time if precision is not too low, the direct negative effect on consumer surplus is not very high.
-As a result, I show that some noise in the platform's evaluation of the customers' preferences can be beneficial for the platform.
+As a result, I show that the platform benefits from some noise in the platform's evaluation of the customers' preferences.
 
@@
 The first prototype of modern recommender system has been created by Xerox Palo Alto Research Center in the late 1980-s early 1990-s and was described by  \cite{goldberg1992using}.
-This system has been called \textit{Tapestry} and has been designed in order to filter out a skyrocketing amount of the incoming emails.
-The system was trying to make the incoming stream of emails interesting for the particular person based on whether these emails were interesting for people whose interests are close to the interests of this person.
+This system has been called \textit{Tapestry} and has been designed to filter out a skyrocketing amount of the incoming emails.
+The system was trying to make the incoming stream of emails interesting for a particular person based on whether these emails were interesting for people whose interests are close to the interests of this person.
 Since that time many papers have been written that mainly explore the recommender systems from the Computer Science standpoint (see \cite{Beel2015r} for the recent survey).
+Computer Science literature mainly focuses on how to use data in order to estimate users' preferences which would allow making more precise and tailored recommendations.
+At the same time, Economists usually assume that we have already learned how to estimate preferences, and try to explore how utilization of such systems may affect the behavior of economic agents, or how to use these systems more effectively.
 
@@
 \cite{avery1999market} considered the influence of the recommender system on a choice regarding the consumption of good for the sequentially arriving agents.
-The model they consider include sequentially arriving agents who can either consume the good and leave a review for the next agents or he can wait until somebody else will try the good and report the review.
+The model they consider include sequentially arriving agents each of whom can either consume the good and leave a review for the next agents or he can wait until somebody else will try the good and report the review.
 The main focus of the work is on finding a socially optimal mechanism, in particular, the authors propose a mechanism which includes monetary transfers to some agents to induce them to try the good.
@@
 The authors consider several types of equilibria concluding that the presence of the recommender system creates additional value by reducing uncertainty for agents about the quality of the good.
+In addition, they demonstrate that if the recommender system is effective in reducing uncertainty, then it is optimal for the seller to differentiate (otherwise symmetric) goods by segmenting the market into two categories: large group of buyers-experimenters who get the good at a lower price in exchange for reviews, and small groups of inflexible buyers who are paying a higher price and get the access to these reviews.
 
-In the two papers mentioned above the authors assumed the recommendation strategy to be given and mainly just consider the impact of the recommender system on other strategic agents.
-
+In the two papers ( \cite{avery1999market}, \cite{bergemann2006optimal}) mentioned above the authors assumed the recommendation strategy to be given and mainly just consider the impact of the recommender system on other strategic agents.
 One of the first works that addresses the question of optimal design of the recommendation policy is the work by \cite{kremer2014}.
@@
 Three papers mentioned above although address the question on optimal recommender policy, do not take into account the strategic nature of recommended alternatives.
+While the latter may be crucially important for implications since in the vast majority of real cases, recommended alternatives (e.g.
+restaurants/hotels/sellers) are strategic agents who can adjust their behavior to the recommendation policy.
+It follows that the platform should take into account the strategic nature of recommended alternatives designing its recommender policy.
 \cite{chen2016advertising} consider the informative role of recommender system on strategies of competing sellers and the equilibrium outcomes.
@@
 \cite{levin2009quality} add a vertical component to the standard Hotelling model, focusing on the sellers' decisions regarding the quality disclosure, although, in the framework without the recommender system.
-In the model which I consider, I firstly endogenise the recommendation strategy that the platform implements, and secondly, I am looking at the interplay between the recommendation policy and sellers' incentives to disclose the information about the goods' quality.
+The authors are focusing on information disclosure and the welfare analysis,  considering two modes of competition: duopoly and cartel.
+The main findings are that under cartel sellers disclose more information than under duopoly; comparing cartel regime to the duopoly one, social welfare may increase or decrease depending on the value of cost of disclosure; and finally, the authors demonstrate that equilibrium disclosure both under duopoly and cartel is socially excessive (i.e.
+sellers disclose more information than is socially optimal).
+
+As you could see, the economic literature on recommending systems was evolving by the following scenario: firstly people explore how the presence of (passive) recommender system affect the common economic environment and behavior of economic agents (e.g.
+\cite{avery1999market}, \cite{bergemann2006optimal}); then the question of optimal recommender policy arose and people explored it under assumption of non-strategic alternatives (e.g.
+\cite{kremer2014}, \cite{che2015}, \cite{papanastasiou2017}); eventually, people started exploring the question of optimal recommender policy with strategic alternatives (e.g.
+\cite{li2018recommender}, \cite{chen2016advertising}).
+The model which I consider is closer to the latest stage of development of the recommender system literature i.e.
+I am addressing the question of optimal recommender policy with strategic alternatives.
+Moreover, employing the theoretical framework introduced by \cite{levin2009quality}, I am looking at the interplay between the recommendation policy and sellers' incentives to disclose the information about the goods' quality.
 	\section{Model}
@@
 There is a unit mass of heterogeneous customers.
-Preferences of customer are characterized by the location of the customer on the unity interval.
-Two competing sellers (firms) which I will refer to $A$ and $B$, who locate at the extremes of the unit interval, are selling substitutable goods which they produce at equal marginal cost normalized to 0.
-Goods of each seller may have a different quality which means that the customers with the same ``horizontal preference'' for the two sellers may derive different utility upon consuming goods of different sellers.
+Preferences of customer are characterized by the location of the customer on the unit interval.
+Two competing sellers (firms) whom I will refer to as $A$ and $B$, who locate at the extremes of the unit interval, are selling substitutable goods which they produce at equal marginal cost normalized to 0.
+Goods of each seller may have a different quality which means that customers with the same ``horizontal preferences'' for the two sellers may derive different utility upon consuming goods of different sellers.
 Sellers are competing in prices which they set simultaneously.
-In addition, the sellers may costly disclose the information about the quality of their good to the platform prior to price setting.
+In addition, the sellers may costly disclose the information about the quality of their good to the platform and to the other seller prior to price setting.
 The recommender system observes prices and sellers' decisions regarding the quality disclosure and makes a recommendation to customers which seller to choose.
@@
 
+The timing of the game looks as follows:
+\begin{enumerate}
+	\item The platform commits to the commission rate $\alpha$.
+	\item The sellers privately observes qualities of their goods $q_A, q_B$ and simultaneously decide whether to disclose them.
+	\item The sellers simultaneously set prices (observing each other's decisions from the Step 2).
+	\item The platform observes prices and whether the sellers disclosed quality, and recommends sellers to customers. After that, the payoffs are realized and the game ends.
+\end{enumerate}
+
 The next three subsections describe formally all parties of my model.
 
+\subsection{Sellers}
+
+Qualities of goods are \textit{ex-ante} unknown and assumed to be $i.i.d.\ U(0, 1)$.
+Prior to price-setting sellers privately learn these qualities (each seller learns only the quality of his own good) and should decide whether to disclose this information to the platform (and to the other seller) or not.
+Upon disclosure, the sellers bear cost $c$.
+Moreover, the sellers pay the platform commission from sales.
+That is, the seller's $i$ profit is given by the following formula
+$$\pi_i = (1-\alpha)p_i D_i - c d_i$$
+where $$d_i = \begin{cases}
+1, \text{ if seller }i\text{ discloses }\\
+0, \text{ otherwise }
+\end{cases}$$ and $D_i$ is the demand on good $i$ which may be a function of prices and decisions regarding the information disclosure, and is determined by the recommender policy the platform employs.
+
+	\subsection{Platform}
+
+The platform announces the commission rate $\alpha$, and after observing sellers' prices and qualities (if disclosed), should design the recommender policy.
+The recommender policy specifies functions $r_A(x, p_A, p_B, d_A, d_B, \mathbb{E}[q_A|d_A], \mathbb{E}[q_B|d_B])$ and \newline
+$r_B(x, p_A, p_B, d_A, d_B, \mathbb{E}[q_A|d_A], \mathbb{E}[q_B|d_B])$ which represents probabilities that the customer with location $x$ will be recommended the seller $A$ and $B$ respectively.
+The objective function of the platform is given by \begin{align}\label{pl_profit}
+\pi_P = \mathbb{E}\left[\alpha(p_A D_A + p_B D_B) + \gamma CS\right],
+\end{align}
+where expectation is taken over unknown qualities $q_A, q_B$; $CS$ is a consumer surplus, $\gamma$ represents the platform's degree of consumer-orientedness.
+I assume that the maximum rate of commission the platform can charge the sellers is $\bar{\alpha} \in (0, 1)$, and $\gamma > \bar{\alpha}$.
+I need the latter assumption $\gamma > \bar{\alpha}$ since if it does not hold it might be optimal for the platform to completely disregard the consumer surplus and induce as high price as possible to extract the whole consumer surplus.
+Although the assumption of maximal commission is for convenience it can be justified by assuming that sellers have non-zero outside option, so the platform cannot extract the whole revenue.
+That is, $\bar{\alpha}$ can be thought of as a maximum commission rate the platform can charge that still guarantees that sellers are willing to sell their goods through the platform.
+
+The goal of the platform is to choose the commission rate $\alpha$ and the recommender policy in order to maximize \eqref{pl_profit}.
+
 	\subsection{Customers}
-Customer who locates at $x \in [0, 1]$ derives a net utility $$V + q_A - p_A - x$$ if  he buys the good at a price $p_A$ from the seller $A$ who locates at a point $0$, and net utility $$V + q_B - p_B - (1-x)$$ if  he buys the good at a price $p_B$ from the seller $B$ who locates at a point $1$.
-$V$ is assumed to be higher than $\frac{3}{2}$ to ensure that the market is fully covered in equilibrium, $q_A, q_B \in (0, 1)$ are the qualities of the goods.
+Customer who locates at $x \in [0, 1]$ derives a net utility $$U(x, A) =  V + q_A - p_A - x$$ if  he buys the good at a price $p_A$ from the seller $A$ who locates at a point $0$, and net utility $$U(x, B) = V + q_B - p_B - (1-x)$$ if  he buys the good at a price $p_B$ from the seller $B$ who locates at a point $1$.
+$V$ is common knowledge and is assumed to be higher than $\max \left\{\frac{7\gamma + 2\bar{\alpha}}{6(\gamma - \bar{\alpha})},  \frac{3 \gamma - \bar{\alpha}}{2(\gamma - \bar{\alpha})} \right\}$ to ensure that all customers are willing to buy in equilibrium, $q_A, q_B \in (0, 1)$ are the qualities of the goods.
 These qualities are \textit{ex-ante} known only to corresponding sellers, and are $i.i.d.\  U(0, 1)$.
@@
 
-	\subsection{Sellers}
-
-Prior to price-setting sellers privately learns the quality of their good and should decide whether to disclose this information to the platform or not.
-Upon disclosure, the sellers bear cost $c$.
-Moreover, the sellers pay the platform commission from sales.
-That is, the seller's $i$ profit is given by the following formula
-$$\pi_i = (1-\alpha)p_i D_i - c d_i$$
-where $$d_i = \begin{cases}
-1, \text{ if sellers }i\text{ discloses }\\
-0, \text{ otherwise }
-\end{cases}$$ and $D_i$ is the demand on good $i$ which may be a function of prices and decisions regarding the information disclosure, and is determined by the recommender policy the platform employs.
-
-	\subsection{Platform}
-
-The platform commits to the commission rate $\alpha$ and the recommender policy.
-The recommender policy specifies functions $r_A(x, p_A, p_B, d_A, d_B, \mathbb{E}q_A, \mathbb{E}q_B)$ and $r_B(x, p_A, p_B, d_A, d_B, \mathbb{E} q_A, \mathbb{E} q_B)$ which represents probabilities that the customer with location $x$ will be recommended the seller $A$ and $B$ respectively.
-The objective function of the platform is given by \begin{align}\label{pl_profit}
-\pi_P = \mathbb{E}\left[\alpha(p_A D_A + p_B D_B) + \gamma CS\right]
-\end{align}
-
-where expectation is taken over unknown qualities $q_A, q_B$; $CS$ is a consumer surplus, $\gamma$ represents the platform's degree of consumer-orientedness.
-I assume that the maximum rate of commission the platform can charge the sellers is $\bar{\alpha} \in (0, 1)$, and $\gamma > \bar{\alpha}$.
-Usually in real life there are several competing platforms, so the platform cannot increase the commission up to 100\% since in this case sellers will just switch to the competitive online retail platform.
-So, $\bar{\alpha}$ can be think of as a maximum commission rate the platform can charge that still guarantees that sellers are willing to sell their goods through the platform.
-
-The goal of the platform is to choose the commission rate $\alpha$ and the recommender policy in order to maximize \eqref{pl_profit}.
-
-	\subsection{Timing}
-The timing of the game looks as follows:
-\begin{enumerate}
-	\item The platform commits to the commission rate $\alpha$ and the recommender policy.
-	\item The sellers privately observes $q_A, q_B$ and decide whether to disclose it to the platform.
-	\item The sellers simultaneously set prices.
-	\item The platform observes prices and whether the sellers disclosed quality, and recommends sellers to customers. After that, the payoffs are realized and the game ends.
-\end{enumerate}
-
 \section{Results}
-Firstly, I consider the case with observable to buyers qualities and without the recommender system.
+In the next subsection I relax the assumption that customers do not know about sellers, and consider the case with observable to buyers qualities and without the recommender system.
+Then I will turn back to the baseline model in which only the recommender system knows about sellers.
 I need this comparison in order to demonstrate how and via which mechanisms the presence of the recommender system affects the equilibrium behavior of sellers.
+Although, one may argue that the information structures in these two scenarios are different and the comparison is not fully valid.
+However, since I will mainly focus on sellers' strategies, it does not matter from the seller perspective how the demand for his good has been induced.
+That is why, I argue that it is still worth looking at the comparison between the two aforementioned cases.
 
@@
 	\subsection{Observable qualities. With the platform.}
-	For simplicity, I assume firstly that $\alpha$ is fixed.
+	For simplicity, I assume firstly that $\alpha$ is fixed. Then I will solve for optimal $\alpha$ as well.
 	\begin{proposition} \label{with}
@@
 	 \end{align*}
-	 Solving the system above yields $p_i = \frac{\gamma}{\gamma - \alpha} + \frac{\gamma(q_i - q_{-i})}{3(\gamma - \alpha)}$
+	 Solving the system above yields $p_i = \frac{\gamma}{\gamma - \alpha} + \frac{\gamma(q_i - q_{-i})}{3(\gamma - \alpha)}$.
+
+	 Finally, let me demonstrate that given the assumption that $V \ge \max \left\{\frac{7\gamma + 2\bar{\alpha}}{6(\gamma - \bar{\alpha})},  \frac{3 \gamma - \bar{\alpha}}{2(\gamma - \bar{\alpha})} \right\}$ all customers are willing to follow the recommendation and buy the good in the equilibrium.
+Given the equilibrium prices $$\hat{x} = \frac{1}{2} + \frac{q_A-q_B}{6}$$ hence to ensure that the customer with location $\hat{x}$ the following condition should be satisfied: \begin{align*}
+	 V+q_A - \frac{1}{2} - \frac{q_A-q_B}{6} - \frac{\gamma}{\gamma - \alpha} - \frac{\gamma(q_A-q_B)}{2(\gamma- \alpha)} \ge 0 \iff\\
+	 \iff V - \frac{1}{2} -\frac{\gamma}{\gamma - \alpha} + \frac{2 \gamma - 5 \alpha}{6(\gamma - \alpha)}q_A + \frac{4 \gamma - \alpha}{6(\gamma - \alpha)}q_B \ge 0
+	 \end{align*}
+	 $\frac{4 \gamma - \alpha}{6(\gamma - \alpha)} >0$, moreover, if $\gamma > \frac{5}{2} \bar{\alpha}$ then $\frac{2 \gamma - 5  \alpha}{6(\gamma - \alpha)} > 0$ and hence
+	 \begin{align*}
+	 V - \frac{1}{2} -\frac{\gamma}{\gamma - \alpha} + \frac{2 \gamma - 5 \alpha}{6(\gamma - \alpha)}q_A + \frac{4 \gamma - \alpha}{6(\gamma - \alpha)}q_B \ge V - \frac{1}{2} - \frac{\gamma}{\gamma - \alpha} \ge 0 \Leftarrow V \ge \frac{3 \gamma - \bar{\alpha}}{2(\gamma - \bar{\alpha})}
+	 \end{align*}
+	 If, on the other hand, $\gamma < \frac{5}{2} \bar{\alpha}$ then $\frac{2 \gamma - 5  \alpha}{6(\gamma - \alpha)}$ can be negative which means that
+	 \begin{align*}
+	 V - \frac{1}{2} -\frac{\gamma}{\gamma - \alpha} + \frac{2 \gamma - 5 \alpha}{6(\gamma - \alpha)}q_A + \frac{4 \gamma - \alpha}{6(\gamma - \alpha)}q_B \ge V - \frac{1}{2}- \frac{\gamma}{\gamma - \alpha} - \frac{5 \alpha - 2 \gamma}{6(\gamma - \alpha)} \ge 0 \Leftarrow \\
+	 \Leftarrow V \ge \frac{7 \gamma + 2 \bar{\alpha}}{6(\gamma - \bar{\alpha})}
+	 \end{align*}
+	 Similarly, we need to check that 
+	 \begin{align*}
+	 V+q_B - \left(1-\frac{1}{2} - \frac{q_A - q_B}{6} \right) - \frac{\gamma}{\gamma - \alpha} - \frac{\gamma(q_A-q_B)}{3(\gamma - \alpha)} \ge 0 \iff\\
+	 \iff V - \frac{1}{2} - \frac{\gamma}{\gamma-\alpha} + \frac{3 \gamma - 5 \alpha}{6(\gamma - \alpha)}q_B + \frac{3 \gamma - \alpha}{6(\gamma - \alpha)}q_A \ge 0
+	 \end{align*}
+	 The last term is non-negative, moreover if $\gamma > \frac{5}{3} \bar{\alpha}$ the penultimate term is also non-negative, hence
+	 \begin{align*}
+	V - \frac{1}{2} - \frac{\gamma}{\gamma-\alpha} + \frac{3 \gamma - 5 \alpha}{6(\gamma - \alpha)}q_B + \frac{3 \gamma - \alpha}{6(\gamma - \alpha)}q_A \ge 0 \Leftarrow V \ge \frac{3 \gamma - \bar{\alpha}}{2(\gamma - \bar{\alpha})}
+	 \end{align*}
+	 If, on the other hand, $\gamma < \frac{5}{2} \bar{\alpha}$ then $\frac{3 \gamma - 5 \alpha}{6(\gamma - \alpha)}$ can be negative. In this case, 
+	 \begin{align*}
+	 	V - \frac{1}{2} - \frac{\gamma}{\gamma-\alpha} + \frac{3 \gamma - 5 \alpha}{6(\gamma - \alpha)}q_B + \frac{3 \gamma - \alpha}{6(\gamma - \alpha)}q_A \ge 0 \Leftarrow  V \ge \frac{3 \gamma + \bar{\alpha}}{3(\gamma-\bar{\alpha})}
+	 \end{align*}
+	 Since $$\frac{7 \gamma + 2 \bar{\alpha}}{6(\gamma - \bar{\alpha})} - \frac{3 \gamma + \bar{\alpha}}{3(\gamma - \bar{\alpha})} = \frac{\gamma}{6(\gamma - \bar{\alpha})} > 0$$ it follows that $V \ge \max \left\{\frac{7\gamma + 2\bar{\alpha}}{6(\gamma - \bar{\alpha})},  \frac{3 \gamma - \bar{\alpha}}{2(\gamma - \bar{\alpha})} \right\}$ is a sufficient condition to guarantee that all customers derive non-negative utility in equilibrium.
 	\end{proof}
@@
 The picture below (Fig.
-\ref{fig1}) demonstrates graphically the impact of the recommender system.
+\ref{fig1}) demonstrates graphically the impact of the recommender system (for simplicity I assumed that $q_A = q_B$).
+Clearly, if $q_A=q_B$ \textit{in equilibrium} $\hat{x} = x_0 = \frac{1}{2}$ but returns on deviation in these two cases are different because of the influence of the recommender system.
 
@@
 In this subsection, I derive the optimal recommender system of the platform.
-Following \cite{levin2009quality} I consider symmetric PBE in which the sellers disclose the quality if and only if the quality falls above some threshold $q^*$ and do not disclose otherwise, and if the seller does not disclose, the beliefs of the platform regarding the quality do not depend on prices.
+Following \cite{levin2009quality} I consider symmetric PBE in which the sellers disclose the quality if and only if the quality is above some threshold $q^*$ and do not disclose otherwise, and if the seller does not disclose, the beliefs of the platform regarding the quality do not depend on prices.
+I will also assume that if the seller which is recommended to some customer does not disclose the quality, the beliefs of the platform regarding this seller's product translate into beliefs of this customer.
+For example, the platform can present the quality of the good as the position in ranking (or number of stars), and customers are supposed to trust the platform's ranking.
+In other words, I assume that if \textit{Amazon} tells you that some seller has $4.5$ stars out of $5$, you trust this information and make your purchasing decision as if the quality is really $4.5/5$.
 
@@
 \end{proposition}
-\begin{proof} Given prices $p_A, p_B$ and qualities $q_A, q_B$, total benefit of the platform is equal to:
-	\begin{align*}
-	\alpha(p_A \hat{x} + p_B(1-\hat{x})) + \gamma \left(\int_{0}^{\hat{x}} (V + q_A - p_A - x)dx + \int_{\hat{x}}^1 (V + q_B - p_B -(1-x))dx\right) = \\
-	= 	\alpha(p_A \hat{x} + p_B(1-\hat{x})) + \gamma \left(V - \frac{1}{2}+(1-\hat{x})(q_B - p_B) + \hat{x}(q_A-p_A) - \hat{x}^2  \right)
-	\end{align*}
-	Where $\hat{x}$ is defined according to Proposition \ref{precise}. Hence, expected total benefit of the platform is
-	\begin{align*}
-	&\pi = \int_{0}^{q^*} \int_0^{q^*} \biggl( \frac{\alpha \gamma}{\gamma - \alpha} + \gamma \left(V - \frac{3}{4}+ \frac{q^*}{2} - \frac{\gamma}{\gamma - \alpha} \right)\biggr) dq_A dq_B +\\+
-	 &\int_0^{q^*} \int_{q^*}^1 \biggl[ \alpha \biggl( \left( \frac{\gamma}{\gamma - \alpha} + \frac{\gamma\left( \frac{q^*}{2} - q_B \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} + \frac{\frac{q^*}{2} - q_B}{6} \right) + \left( \frac{\gamma}{\gamma - \alpha} - \frac{\gamma\left( \frac{q^*}{2} - q_B \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} - \frac{\frac{q^*}{2} - q_B}{6} \right) \biggr)+\\
-	 &+\gamma \biggl(V - \frac{1}{2} + \left(\frac{1}{2} - \frac{\frac{q^*}{2} - q_B}{6} \right) \left( q_B - \frac{\gamma}{\gamma - \alpha} + \frac{\gamma \left( \frac{q^*}{2} - q_B \right)}{3 (\gamma - \alpha)} \right)  + \\
-	 &+ \left(\frac{1}{2} + \frac{\frac{q^*}{2} - q_B}{6} \right) \left(\frac{q^*}{2} - \frac{\gamma}{\gamma - \alpha} - \frac{\gamma \left( \frac{q^*}{2} - q_B \right)}{3(\gamma - \alpha)} \right)  - \left( \frac{1}{2} + \frac{\frac{q^*}{2} - q_B}{6} \right)^2 \biggr)  \biggr] dq_A dq_B +\\
-	 +
-	 &\int_{q^*}^{1} \int_{0}^{q^*} \biggl[ \alpha \biggl( \left( \frac{\gamma}{\gamma - \alpha} + \frac{\gamma\left(q_A - \frac{q^*}{2} \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} + \frac{q_A - \frac{q^*}{2}}{6} \right) + \left( \frac{\gamma}{\gamma - \alpha} - \frac{\gamma\left(q_A- \frac{q^*}{2} \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} - \frac{q_A -\frac{q^*}{2}}{6} \right) \biggr)+\\
-	 &+\gamma \biggl(V - \frac{1}{2} + \left(\frac{1}{2} - \frac{q_A-\frac{q^*}{2}}{6} \right) \left( \frac{q^*}{2} - \frac{\gamma}{\gamma - \alpha} + \frac{\gamma \left(q_A -  \frac{q^*}{2} \right)}{3 (\gamma - \alpha)} \right)  + \\
-	 &+ \left(\frac{1}{2} + \frac{q_A-\frac{q^*}{2}}{6} \right) \left(q_A - \frac{\gamma}{\gamma - \alpha} - \frac{\gamma \left(q_A- \frac{q^*}{2} \right)}{3(\gamma - \alpha)} \right)  - \left( \frac{1}{2} + \frac{q_A - \frac{q^*}{2}}{6} \right)^2 \biggr)  \biggr] dq_A dq_B + \\
-	 +
-	  &\int_{q^*}^{1} \int_{q^*}^{1} \biggl[ \alpha \biggl( \left( \frac{\gamma}{\gamma - \alpha} + \frac{\gamma\left(q_A - q_B \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} + \frac{q_A - q_B}{6} \right) + \left( \frac{\gamma}{\gamma - \alpha} - \frac{\gamma\left(q_A- q_B \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} - \frac{q_A -q_B}{6} \right) \biggr)+\\
-	 &+\gamma \biggl(V - \frac{1}{2} + \left(\frac{1}{2} - \frac{q_A-q_B}{6} \right) \left( q_B - \frac{\gamma}{\gamma - \alpha} + \frac{\gamma \left(q_A -  q_B \right)}{3 (\gamma - \alpha)} \right)  + \\
-	 &+ \left(\frac{1}{2} + \frac{q_A-q_B}{6} \right) \left(q_A - \frac{\gamma}{\gamma - \alpha} - \frac{\gamma \left(q_A- q_B \right)}{3(\gamma - \alpha)} \right)  - \left( \frac{1}{2} + \frac{q_A - q_B}{6} \right)^2 \biggr)  \biggr] dq_A dq_B = \\
-	 & = \frac{\gamma(216 V - q^{*3} - 161)}{216}
-	\end{align*}
-	Thus $$\frac{\partial \pi}{\partial \alpha} = -\frac{3 \gamma q^{*2}}{216} \frac{\partial q^*}{\partial \alpha}$$ and using the result of the Proposition \ref{thres} one may finally prove the result.
+\begin{proof} 
+	See Appendix.
 \end{proof}
@@
 Formally, I assume that before making recommendation to a particular customer whose true location $x$ is equal to $x_0 \in [0, 1]$ the platform receives a signal $s$ such that $$\mathbb{P}(s=x_0|x=x_0) = \beta \in [0, 1]$$ and with probability $1-\beta$, $s \sim U(0, 1)$.
-I will refer to $\beta$ as the recommender system precision.
+$\beta$ is common knowledge and I will refer to it as the recommender system precision.
 In this case, it can be shown (see  \cite{li2018recommender} for example) that the conditional expectation regarding the true location is equal to \begin{align}\label{signal}
@@
 	\newpage
+\appendix
+\section*{Appendix}
+	\textbf{Proof of Proposition \ref{opt_alpha}}
+	\begin{proof} Given prices $p_A, p_B$ and qualities $q_A, q_B$, total benefit of the platform is equal to:
+		\begin{align*}
+		\alpha(p_A \hat{x} + p_B(1-\hat{x})) + \gamma \left(\int_{0}^{\hat{x}} (V + q_A - p_A - x)dx + \int_{\hat{x}}^1 (V + q_B - p_B -(1-x))dx\right) = \\
+		= 	\alpha(p_A \hat{x} + p_B(1-\hat{x})) + \gamma \left(V - \frac{1}{2}+(1-\hat{x})(q_B - p_B) + \hat{x}(q_A-p_A) - \hat{x}^2  \right)
+		\end{align*}
+		Where $\hat{x}$ is defined according to Proposition \ref{precise}. Hence, expected total benefit of the platform is
+		\begin{align*}
+		&\pi = \int_{0}^{q^*} \int_0^{q^*} \biggl( \frac{\alpha \gamma}{\gamma - \alpha} + \gamma \left(V - \frac{3}{4}+ \frac{q^*}{2} - \frac{\gamma}{\gamma - \alpha} \right)\biggr) dq_A dq_B +\\+
+		&\int_0^{q^*} \int_{q^*}^1 \biggl[ \alpha \biggl( \left( \frac{\gamma}{\gamma - \alpha} + \frac{\gamma\left( \frac{q^*}{2} - q_B \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} + \frac{\frac{q^*}{2} - q_B}{6} \right) + \left( \frac{\gamma}{\gamma - \alpha} - \frac{\gamma\left( \frac{q^*}{2} - q_B \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} - \frac{\frac{q^*}{2} - q_B}{6} \right) \biggr)+\\
+		&+\gamma \biggl(V - \frac{1}{2} + \left(\frac{1}{2} - \frac{\frac{q^*}{2} - q_B}{6} \right) \left( q_B - \frac{\gamma}{\gamma - \alpha} + \frac{\gamma \left( \frac{q^*}{2} - q_B \right)}{3 (\gamma - \alpha)} \right)  + \\
+		&+ \left(\frac{1}{2} + \frac{\frac{q^*}{2} - q_B}{6} \right) \left(\frac{q^*}{2} - \frac{\gamma}{\gamma - \alpha} - \frac{\gamma \left( \frac{q^*}{2} - q_B \right)}{3(\gamma - \alpha)} \right)  - \left( \frac{1}{2} + \frac{\frac{q^*}{2} - q_B}{6} \right)^2 \biggr)  \biggr] dq_A dq_B +\\
+		+
+		&\int_{q^*}^{1} \int_{0}^{q^*} \biggl[ \alpha \biggl( \left( \frac{\gamma}{\gamma - \alpha} + \frac{\gamma\left(q_A - \frac{q^*}{2} \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} + \frac{q_A - \frac{q^*}{2}}{6} \right) + \left( \frac{\gamma}{\gamma - \alpha} - \frac{\gamma\left(q_A- \frac{q^*}{2} \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} - \frac{q_A -\frac{q^*}{2}}{6} \right) \biggr)+\\
+		&+\gamma \biggl(V - \frac{1}{2} + \left(\frac{1}{2} - \frac{q_A-\frac{q^*}{2}}{6} \right) \left( \frac{q^*}{2} - \frac{\gamma}{\gamma - \alpha} + \frac{\gamma \left(q_A -  \frac{q^*}{2} \right)}{3 (\gamma - \alpha)} \right)  + \\
+		&+ \left(\frac{1}{2} + \frac{q_A-\frac{q^*}{2}}{6} \right) \left(q_A - \frac{\gamma}{\gamma - \alpha} - \frac{\gamma \left(q_A- \frac{q^*}{2} \right)}{3(\gamma - \alpha)} \right)  - \left( \frac{1}{2} + \frac{q_A - \frac{q^*}{2}}{6} \right)^2 \biggr)  \biggr] dq_A dq_B + \\
+		+
+		&\int_{q^*}^{1} \int_{q^*}^{1} \biggl[ \alpha \biggl( \left( \frac{\gamma}{\gamma - \alpha} + \frac{\gamma\left(q_A - q_B \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} + \frac{q_A - q_B}{6} \right) + \left( \frac{\gamma}{\gamma - \alpha} - \frac{\gamma\left(q_A- q_B \right) }{3(\gamma - \alpha)} \right)\left(\frac{1}{2} - \frac{q_A -q_B}{6} \right) \biggr)+\\
+		&+\gamma \biggl(V - \frac{1}{2} + \left(\frac{1}{2} - \frac{q_A-q_B}{6} \right) \left( q_B - \frac{\gamma}{\gamma - \alpha} + \frac{\gamma \left(q_A -  q_B \right)}{3 (\gamma - \alpha)} \right)  + \\
+		&+ \left(\frac{1}{2} + \frac{q_A-q_B}{6} \right) \left(q_A - \frac{\gamma}{\gamma - \alpha} - \frac{\gamma \left(q_A- q_B \right)}{3(\gamma - \alpha)} \right)  - \left( \frac{1}{2} + \frac{q_A - q_B}{6} \right)^2 \biggr)  \biggr] dq_A dq_B = \\
+		& = \frac{\gamma(216 V - q^{*3} - 161)}{216}
+		\end{align*}
+		Thus $$\frac{\partial \pi}{\partial \alpha} = -\frac{3 \gamma q^{*2}}{216} \frac{\partial q^*}{\partial \alpha}$$ and using the result of the Proposition \ref{thres} one may finally prove the result.
+	\end{proof}
 	\bibliography{lib}{}

### 2020-05-21 Thesis.tex
Add files via upload
Thesis

@@
 \usepackage{mathtools}
+\usepackage{pdfpages}
 \usepackage{pgfplots}
@@
 \date{}
-\author{Danil Fedchenko}%\thanks{I am very grateful to: Sergei Izmalkov for constant help and support during the work on this project; Marta Troya Martinez for many valuable comments and suggestions.
-\newline Any remaining errors are mine.}}
+\author{Danil Fedchenko\thanks{I am very grateful to: Sergei Izmalkov for constant help and support during the work on this project; Marta Troya Martinez for many valuable comments and suggestions.
+\newline Any remaining errors and inaccuracies are mine.}}
 \begin{document} 
+	\includepdf[pages=-,pagecommand={},width=\paperwidth]{titul.pdf}
 	\maketitle
@@
 
+	\subsection{Motivation}
+
 		Suppose you came to New York and would like to find a good restaurant to go out tonight.
@@
 Before the World Wide Web, to find hotels or restaurants people relied upon guide-books and brochures distributed by travel agents.
-Those guides, once printed, usually quickly became out-of-date in the constantly changing environment, and could be criticized for containing misleading information.
+Those guides, once printed, could quickly become out-of-date in the constantly changing environment, and, as a result, could be criticized for containing misleading information.
 
@@
 These platforms play roles of intermediaries between buyers and sellers, helping the buyers to find the goods they need.
-And this process clearly includes making recommendations: which seller to display the buyer when he or her is searching for something.
+And this process clearly includes making recommendations: which seller to display the buyer when he or she is searching for something.
 
@@
 What should be the optimal commissions and recommender policy in this case?
-Thus, in this work I am focusing on the following questions:
+I find all those issues interesting both from pure theoretical and practitioner's standpoints and explore them in this work.
+ So, to summarize this subsection let me list the questions I will address.
  \begin{itemize}
@@
 
+ 	\subsection{Overview of the Model and Results}
 	To address these questions I build a theoretical one-period model.
@@
 At the same time, the recommender system's inaccuracy increases the attention the platform pays to the quality of the good when it makes recommendations: since the platform assumes that horizontal component (location) may be measured imprecisely, making a recommendation it begins to put more weights on vertical component (quality) which, if disclosed, is precise.
-The latter effect increases the sellers' returns on quality disclosure making the latter more profitable.
+The latter effect increases the sellers' returns on quality disclosure, making the latter more profitable.
 Eventually, the combined effect of recommender system's inaccuracy on the amount of information disclosed by sellers depends on the value of cost parameter: if the cost of disclosure is high enough, the direct negative effect caused by more intensive competition outweighs the positive effect brought about by the increased return on disclosure and, hence, improving the recommender system precision decreases the amount of information the sellers disclose; for the low cost of disclosure the opposite is true, i.e.
@@
 I assume that the maximum rate of commission the platform can charge the sellers is $\bar{\alpha} \in (0, 1)$, and $\gamma > \bar{\alpha}$.
-I need the latter assumption $\gamma > \bar{\alpha}$ since if it does not hold it might be optimal for the platform to completely disregard the consumer surplus and induce as high price as possible to extract the whole consumer surplus.
+I need the latter assumption $\gamma > \bar{\alpha}$ to rule out possible cases which lie beyond the focus of this work: if this assumption does not hold it might be optimal for the platform to completely disregard the consumer surplus and induce as high price as possible to extract the whole consumer surplus.
 Although the assumption of maximal commission is for convenience it can be justified by assuming that sellers have non-zero outside option, so the platform cannot extract the whole revenue.
@@
 In other words, there exists some optimal level of precision that lies strictly between 0 and 1.
-	\section{Discussion}
+To find this optimal precision may be an interesting problem, although to keep the paper to an appropriate length I leave this for the future research.
 	\section{Extensions}
+	Being rather simple, the model I considered in the previous sections allows for several extensions.
+In this section, I firstly discuss potential ways of extending the existing model.
+Then I suggest some ideas which might render the model more sophisticated.
+The latter, in principle, may give us new insights which cannot be seen in this work.
+
+	First of all, it is worth studying what happens if the assumption on the common component of agents' utilities, $V$ does not hold.
+In this case, markets may not be fully covered in the type of equilibrium I considered.
+This fact may affect the strategic behavior of sellers, as a result, outcomes of interaction can change substantially.
+
+	Moreover, in this work, I demonstrated that perfect precision is not optimal for the recommender system, although I did not explore what is the optimal precision.
+To find an optimal precision and investigate which factors affect it -- is another interesting possibility for future work.
+
+	Another appealing direction is to endogenise $\gamma$ by allowing the platform to charge buyers for receiving recommendations (subscription fee)\footnote{I thank Sergey Kovbasyuk for this suggestion}.
+In reality,  there are many platforms which indeed charge subscription fees providing customers with recommendations in return.
+In this case, the platform gets a tool which enables it to directly affect consumer surplus.
+So, it would be very interesting to see how, in this case, the optimal policy: recommendations, sellers commissions, buyers commissions would look like.
+
+	In my model sellers simultaneously undertake strategic actions, although in real life many sellers may, in the first place, observe actions of their competitors.
+Hence, it can be interesting to investigate how robust are theoretical results obtained in this work to the change of mode of competition between sellers.
+In particular, one potential extension in this direction may include considering Stackelberg-like leader-follower mode of competition \footnote{I thank Shlomo Weber for this suggestion}.
+Additionally, it may be interesting to follow \cite{levin2009quality} and consider the cartel model (i.e.
+allow sellers to collude and undertake strategic actions cooperatively).
+
+	Finally, it could be interesting to test and validate the kind of theoretical predictions, I made in this work, empirically.
+Although it can be difficult to obtain data on exact recommendation strategies different platform use, data on commissions can be found relatively easily.
 	\section{Conclusion}
+	Recommender systems became widespread in the last several decades.
+And although Computer and Data Scientists advanced a lot in our abilities to pin down users' preferences which allows us to make personal recommendations.
+It was clear that these abilities of platforms to make recommendations might substantially transform the information structure of the interaction between economic agents.
+The latter suggests that the platform should judiciously choose its recommendation policy (and maybe even not always recommend an alternative which is the best for a user from his or her own standpoint\footnote{In my model this indeed happens.
+Under some conditions, there may exist users who strictly prefer one seller to another while are recommended the latter one simply because this guarantees the platform higher profit from commissions.}).
+This work belongs to the strand of economic literature which explores and suggests how the optimal recommendation policy should look like and what factors may influence it.
+
+	Turning to the set of questions I listed in the introduction, in the work I derived the optimal recommendation policy.
+I demonstrate that the presence of a recommender system which has stakes in sellers' profit softens competition.
+Moreover, the effect is stronger the lower is the degree of platform's consumer orientedness (which I denoted as $\gamma$).
+This parameter plays a crucial role in determining the effect of commissions on the amount of information which sellers disclose.
+I showed that if $\gamma$ is high then high commission results in less information being disclosed by sellers in equilibrium.
+While for small $\gamma$'s the relation goes in another direction i.e.
+an increase in commission increases the amount of disclosed information.
+This result stems from the interplay between two counteracting forces: an increase in commission decreases the sellers' profits directly; at the same time, this increases the platform's stakes in profit and as a result softens competition.
+The implication of this result is that for large $\gamma$ the platform should charge sellers the smallest possible commission.
+
+As for precision, I demonstrated that low precision weakens the effect of the presence of the recommender system.
+In other words, other things being equal, lower precision means harsher competition.
+This happens since the imprecision effectively increases the share of ``average'' customers who are the most desirable target the competing sellers would like to compete for.
+The imprecision also affects the amount of information the sellers disclose in equilibrium.
+However, the direction of this effect depends on the cost of disclosure: for low cost making the recommender system more precise may increase the amount of information disclosed, while for high cost the result is the opposite.
+The reason is that imprecision makes it optimal for the platform to put more weight on disclosed information (since this information is precise).
+This motivates sellers to disclose the information, partially offsetting the direct negative effect of imprecision brought about by harsher competition and lower profits.
+The last thing I showed in this paper is that some imprecision, in fact, can be optimal for the platform.
+The intuition behind this result is the following: on one hand, the imprecision directly harms consumer surplus (since mismatch between seller and buyer is more probable); but at the same time, perfect precision makes the competition too soft, and this negative (for consumer surplus) effect outweighs benefits caused by the absence of mismatch.
+In other words, it is optimal sometimes to ``sacrifice'' some customers (and recommend them not the best alternative for them) in order to intensify competition between sellers and benefit everybody (except for sellers) by lower prices.
+
+	The theoretical model considered in this work is, of course, only a simplification of reality.
+It can be extended further to incorporate more real-life features of the type of interactions I considered.
+However, I believe the results I obtained can give practitioners insights about the thing they should pay attention once designing the optimal recommendation policy.
 	\newpage

### 2021-01-04 ZeroDerivative.tex
Add files via upload

\documentclass[a4paper]{article}
\usepackage[12pt]{extsizes} % 
\usepackage{setspace}
\doublespacing
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
%\usepackage{harvard}
\usepackage[round]{natbib}
\usepackage{pdfpages}
\usepackage{tikz}
\usepackage{makecell}
\usepackage{amsthm}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{xcolor,colortbl}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage{breakcites}
\interfootnotelinepenalty=10000
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
	{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
%\newenvironment{abstract}[0]{\small\rm
%	\begin{center}ABSTRACT
%		\\ \vspace{8pt}
%		\begin{minipage}{5.2in}\smalllineskip
%			\hspace{1pc}}{\end{minipage}\end{center}\vspace{-1pt}}
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathcal{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\eps}{\varepsilon}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=25mm, top=20mm, right=25mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{definition}[theorem]{Definition}
%\let\cite = \citeasnoun
\title{Zero derivative and local risk aversion}
\date{}
\author{Danil Fedchenko}
\begin{document} % 
	\maketitle
	\begin{abstract}
	Consider a random variable $\varepsilon$ such that $\expect[\varepsilon] = 0$. Denote $\pi(t)$ a risk premium, an expected-utility maximizer with Bernoulli utility $u(x)$ is willing to pay to avoid risk $t \varepsilon$ (i.e. $u(x-\pi(t)) = \int u(x+t\varepsilon)dF(\varepsilon)$). As \cite{segal1990first} stated, $\pi'(0) = 0$. This is true, however, only if $u'(x) \neq 0$. In this note I show that an attitude toward risk of an individual with $u'(x) = 0$ in some cases is similar to being what \cite{segal1990first} called of order 1. Which in particular implies that such an individual may find optimal to buy full insurance even for a small positive risk in the neighborhood of the points where $u'(x) = 0$.
	\end{abstract}
	\newpage
	\section{Introduction}
	Consider an individual with expected utility preferences represented by strictly increasing, continuous Bernoulli utility function $u(x)$. Suppose this individual for some $t \in \mathbb{R}$ faces a lottery $t \eps$ represented by a random variable $\varepsilon$ such that $\expect[\varepsilon] = 0$. Define a risk premium $\pi(t)$ implicitly as $x-\pi(t) \sim x+t\eps$ or equivalently $u(x - \pi(t)) = \int u(x + t \eps)\ dF(\eps)$ (from now on whenever I write $\int$, I assume the integral exists). It is easy to show that as long as $u'(x) \neq 0$, $\pi'(0) = 0$. Indeed, 
	\begin{align*}
	\pi(t) &= x - u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right),\ u^{-1} \text{ exists since }u(x)\text{ is strictly increasing and continuous}\\
	\frac{\partial \pi(t)}{\partial t} &= -\left(u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right) \right)' \int u'(x+t \varepsilon)\varepsilon dF(\varepsilon) = \\
	&=-\frac{\int u'(x+t \varepsilon)\varepsilon dF(\varepsilon)}{u'\left(u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right)\right)}\ ,\ \ \  \text{if }u'\left(u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right)\right) \neq 0\\
	\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} &= -\frac{u'(x)}{u'(x)} \expect[\eps] = -\expect[\eps] = 0, \text{ if } u'(x) \neq 0
	\end{align*}
This in particular implies that for small enough $t$ $$\pi(t) \approx -\frac{t^2 \expect[\eps^2]}{2}\frac{u''(x)}{u'(x)}$$ and hence it approaches zero faster than $t$ which, in turn, implies that for small $t$ an individual is almost risk neutral. This assertion, however, heavily relies upon the assumption $u'(x) \neq 0$ which is known sometimes to be violated even for strictly increasing functions ($u(x) = x^3$ at $x=0$ is probably the most famous example). 
In this paper I partially characterize $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0}$ at points where $u'(x) = 0$, and show that it is not equal to zero in general. Applying this to the optimal insurance problem, as \cite{segal1990first} show, means that an expected utility maximizer may find optimal to buy full insurance even for small risk and even if insurance includes positive loading. I illustrate this result by deriving a slope of indifference curves (provided that it is non zero and finite) of the individual near the certainty line on the state-contingent  \cite{yaari1978some}-type diagram. It turns out that the slope differs from the case where $u'(x) \neq 0$.
\section{Order of Risk Aversion}
\cite{segal1990first} define an attitude toward risk of individual being of order 1 if for \textit{every} non-degenerate random variable $\eps$ such that $\expect[\eps] = 0$, $\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0^+} \neq 0$, and of order 2 if $\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0^+} = 0$. 
In the next proposition I derive $\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0}$ (assuming that it exists, is non-zero and finite) at the points where for some $k\ge 1$ first $k-1$ derivatives of the Bernoulli utility function $u(x)$ exist and equal to zero.
	\begin{proposition}\label{derivative}
		If $\exists\ x^* \in \mathbb{R}$ and $k \ge 2$ such that $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, $u'(x) > 0$ for $x \in (x^*-\delta, x^*+\delta)\setminus \{x^*\}$ for some small $\delta>0$, and there exists $u^{(k)}(x^*) \neq 0$. Then if there exists finite and non-zero $\expect[\eps^k]$, and at $x^*$ there exists finite, non-zero $\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0}$ then :
		\begin{align}
			\left( \frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0} \right)^k = (-1)^k\ \expect[\varepsilon^k]
		\end{align}
	Provided that \begin{align}\label{limsup}
	\forall\ i = \{2,\dots, k-1\}:\ \bigg|\lim_{t\to0} \inf \frac{\partial^{(i)} \pi(t)}{\partial t^{(i)}}\bigg| < \infty, \bigg|\lim_{t\to0} \sup \frac{\partial^{(i)} \pi(t)}{\partial t^{(i)}}\bigg| < \infty
\end{align} for $k>2$.
	\end{proposition}
\begin{proof}
	\begin{align*}
		\pi(t) &= x^* - u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\\
		\frac{\partial \pi(t)}{\partial t} &= -\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right) \right)' \int u'(x^*+t \varepsilon)\varepsilon dF(\varepsilon) = \\
		&=-\frac{\int u'(x^*+t \varepsilon)\varepsilon dF(\varepsilon)}{u'\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right)}\ ,\ \ \  \text{if }u'\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right) \neq 0
	\end{align*}
As $t \to 0$ since $u'(x^*) = 0$, both numerator and denominator approach zero.
Using the L'Hospital rule we get:
\begin{align*}
		\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0} = \lim_{t \to 0} -\frac{\int u''(x^*+t \varepsilon) \varepsilon^2 dF(\varepsilon)}{u''\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right) \underbrace{\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right) \right)'}_{-\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0}}}
\end{align*}
If $u''(x^*) \neq 0$ (i.e. $k=2$) then $\left( \frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0} \right)^2 = \expect[\varepsilon^2]$. If $k>2$ then one should apply L'Hospital rule once again. That is,
\begin{align*}
	\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0} = \lim_{t \to 0} -\frac{\int u'''(x^*+t \varepsilon) \varepsilon^3 dF(\varepsilon)}{u'''\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right) \left( \frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0}\right)^2 - u''\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right) \frac{\partial^2\pi(t)}{\partial t^2} }
\end{align*}
Since $u''(x^*) = 0$ and \eqref{limsup} implies  $$\lim_{t \to 0} u''\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right) \frac{\partial^2\pi(t)}{\partial t^2} = 0,$$ if $k=3$ then we have $$\left( \frac{\partial \pi(t)}{\partial t} \bigg|_{t=0}\right)^3 = -\expect[\eps^3] $$ which is what we sought to prove. If $k>3$ we should proceed in the same way, i.e. by applying L'Hospital rule $k-3$ times we get what we wanted to prove.
\end{proof}
Although, the proposition above does not tell us whether $\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0} \neq 0$, so we cannot conclude that the individual's attitude toward risk is necessary of order 1 (according to \cite{segal1990first}). We will see in Section 4 that, in general neither $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} \neq 0$ nor $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} = 0$, so the individual's attitude toward risk is neither of order 1 nor of order 2. Nevertheless, if it happens to be $\expect[\eps^k] \neq 0$ and $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} \neq 0$, optimal behavior of such an individual regarding the choice of insurance is similar to that of the one whose attitude toward risk is of order 1.
\section{Optimal Insurance}
Discussing the optimal insurance problem consider the lottery $x^*+t \eps $ where $\expect[\eps] > 0$. Then, the following proposition is an analogue to the Proposition 1 of \cite{segal1990first}.
\begin{proposition}\label{LRA}
	If $\exists\ x^* \in \mathbb{R}$ and $k \ge 2$ such that $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, there exist non-zero left and right derivatives $u^{(k)}_{+}(x^*)$ and $u^{(k)}_{-}(x^*)$, and $u(x)$ is $k$ times differentiable on $(x^*-\delta, x^*+\delta)\setminus\{x^*\}$ for some $\delta>0$. If $\eps$ possesses all finite moments, and in particular, there exist $\expect[ \eps^k |\eps\ge0]$, $\expect[ \eps^k |\eps<0]$ and \begin{align}\label{condition}
		u^{(k)}_{+}(x^*) \cdot \expect[ \eps^k | \eps\ge0]\prob(\eps\ge0) + u^{(k)}_{-}(x^*) \cdot \expect[ \eps^k | \eps<0]\prob(\eps<0) < 0,
		\end{align}
	then for sufficiently small $t > 0$ \textbf{an individual strictly prefers a certain outcome $x^*$ to the lottery $x^*+t\eps$.}
\end{proposition}
\begin{proof}
	\begin{align*}
		\int u(x^*+t\eps)dF(\eps) = \int \left( u(x^*) +  \bigg[u^{(k)}_{+}(x^*)\mathbb{I}[\eps\ge0]  + u^{(k)}_{-}(x^*)\mathbb{I}[\eps<0] \bigg] \frac{t^k \eps^k }{k!} +o_p(t^k \eps^k)\right)dF(\eps) = \\
	= u(x^*) + \frac{(u^{(k)}_{+}(x^*) \expect[\eps^k|\eps\ge 0]\prob(\eps\ge0)+u^{(k)}_{-}(x^*) \expect[\eps^k|\eps< 0]\prob(\eps<0))t^k}{k!} + \expect[o_p(t^k \eps^k)] < u(x^*)
	\end{align*}
where the first equality follows from Taylor formula and the fact that $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, and the last inequality follows from \eqref{condition}. Note that for the $k$ times differentiable $u(x)$, \eqref{condition} becomes $u^{(k)}(x^*)\cdot\expect[\eps^k] < 0$
\end{proof}
The result above means that the expected-utility maximizer under some conditions exhibits local risk aversion at the vicinity of the points where his strictly increasing Bernoulli utility function fails to have positive derivatives up to the order $k$. In other words, this individual prefers a certain outcome to the lottery with small positive expected value. This result stands in contrast to the well-known result that the expected-utility maximizer with increasing Bernoulli utility function $u(x)$ is locally risk neutral as long as $u'(x) > 0$. That is, for $t$ small enough he would strictly prefer the lottery with positive mean to the certain outcome.

Let me illustrate the proposition \ref{LRA} by deriving slopes of the individual's indifference curves at a certainty line on the state-contingent diagram. 
Consider an individual whose expected utility preferences represented by strictly increasing and continuous Bernoulli utility function $u(x)$. Suppose there are two states of nature, the first happens with probability $p$ and the second with probability $1-p$. At state $i \in \{1, 2\}$ the individual gets $x_i$. Then the following proposition holds:
\begin{proposition}\label{slope}
	If $\exists\ x^* \in \mathbb{R}$ and $k \ge 2$ such that $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, $u'(x) > 0$ for $x \in (x^*-\delta, x^*+\delta)\setminus \{x^*\}$ for some small $\delta>0$. If there exist left and right derivatives $u_{+}^{(k)}(x^*) \neq 0$, $u_{-}^{(k)}(x^*) \neq 0$ such that \footnote{$$sign(x) = \begin{cases}
			-1, x <0\\
			0, x=0\\
			1, x>0
		\end{cases}$$}$\text{sign}(u_{+}^{(k)}(x^*) \cdot u_{-}^{(k)}(x^*)) = (-1)^{k+1}$

	Then consider an indifference curve $x_2(x_1)$ implicitly given by 
	\begin{align}\label{ind}
		\{ (x_1, x_2): pu(x_1)+(1-p)u(x_2) = u(x^*) \}
	\end{align}
then if there exist finite, non-zero $\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0}$,  $\frac{d x_2}{d x_1} \bigg|_{x_1=x^*-0}$:
	\begin{align}
		\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0} = -\sqrt[k]{\frac{p}{1-p} \bigg|  \frac{u^{(k)}_{+}(x^*)}{u^{(k)}_{-}(x^*)} \bigg|  }\\
		\frac{d x_2}{d x_1} \bigg|_{x_1=x^*-0} = -\sqrt[k]{\frac{p}{1-p} \bigg|  \frac{u^{(k)}_{-}(x^*)}{u^{(k)}_{+}(x^*)} \bigg|  }
	\end{align}
Provided that 
\begin{align}\label{der}
\forall\ i = \{1, \dots, k-1\}:\ \bigg| \lim_{x_1 \to x^* \pm 0} \sup \frac{d^{(i)}x_2(x_1)}{dx_1^{(i)}}\bigg| < \infty, \bigg| \lim_{x_1 \to x^* \pm 0} \inf \frac{d^{(i)}x_2(x_1)}{dx_1^{(i)}}\bigg| < \infty
\end{align}
\end{proposition}
\begin{proof}
	From \eqref{ind}, implicit function theorem and the fact that $u'(x) > 0$ for $x \in (x^*-\delta, x^*+\delta)\setminus \{x^*\}$ it follows that
	\begin{align*}
		\lim_{x_1 \to x^*+0} \frac{dx_2}{dx_1} = \lim_{x_1 \to x^*+0} -\frac{p}{1-p} \frac{u'(x_1)}{u'(x_2(x_1))}
	\end{align*}
Since $x_2(x^*) = x^*$ and $u'(x^*) = 0$ both numerator and denominator approach zero as $x_1 \to x^*+0$. Then if $	\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0}$ exists we can use L'Hospital rule, i.e.
\begin{align*}
		\lim_{x_1 \to x^*+0} \frac{dx_2}{dx_1} = -\frac{p}{1-p} \frac{u''_{+}(x_1)}{u''_{-}(x_2(x_1)) \left(\frac{dx_2}{dx_1}\right)\bigg|_{x^*+0}}
	\end{align*}
We have left second derivative in the denominator since from the fact that $u(x)$ is strictly increasing it follows that the indifference curve should be downward slopping. Meaning that if $x_1$ approaches $x^*$ from above, $x_2(x_1)$ approaches $x^*$ from below.

If $k = 2$ and $u''_{+}(x^*) \cdot u''_{-}(x^*) < 0$ then $$	\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0} = -\sqrt{\frac{p}{1-p} \bigg|  \frac{u''_{+}(x^*)}{u''_{-}(x^*)} \bigg|  }$$
Minus sign again follows from downward slopping indifference curve. Left derivative can be found in exactly the same way.

If $k > 2$ then apply the L'Hospital rule once again. We get 
\begin{align*}
	\lim_{x_1 \to x^*+0} \frac{dx_2}{dx_1} = -\frac{p}{1-p} \frac{u'''_{+}(x_1)}{u'''_{-}(x_2(x_1)) \left(\frac{dx_2}{dx_1}\right)^2\bigg|_{x^*+0} + u''_{-}(x_2(x_1))\left(\frac{d^2x_2}{dx_1^2}\right)\bigg|_{x^*+0} }
\end{align*}
From \eqref{der} and $u''(x^*) = 0$ it follows that $$\lim_{x_1 \to x^*+0} u''_{-}(x_2(x_1))\left(\frac{d^2x_2}{dx_1^2}\right)\bigg|_{x^*+0} = 0$$
If $k = 3$ and $u'''_{+}(x^*) \cdot u'''_{-}(x^*) > 0$ then $$	\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0} = -\sqrt[3]{\frac{p}{1-p} \bigg|  \frac{u'''_{+}(x^*)}{u'''_{-}(x^*)} \bigg|  }$$
which is what we sought to prove. If $k>3$ then one should apply the L'Hospital rule as many times as it requires to get rid of indeterminacies in numerator and denominator. Which eventually leads to what we wanted to prove.
\end{proof}
In contrast to the result above it is known that in case of differentiable Bernoulli utility, slope of an expected-utility maximizer's indifference curve at certainty points where $u'(x) \neq 0$ is equal to $-\frac{p}{1-p}$. If this individual is offered to buy insurance which would cost $\gamma$ for every insured unit and pay $\$\ 1$ in case of the bad event occurs, then the individual's budget constraint would have a slope $-\frac{\gamma}{1-\gamma}$. This in particular explains why such expected-utility maximizer never buys full insurance unless it is actuarially fair (i.e. $\gamma = p$). At the same time, as we see from the Proposition \ref{slope} an individual with $k$ times differentiable Bernoulli utility with first $k-1$ zero derivatives at some $x$ has slope of the indifference curve at $(x, x)$ given by $-\sqrt[k]{\frac{p}{1-p}}$. Which means that there exists a range of positive loading $\gamma > p$ such that the expected utility maximizer would find optimal to buy full insurance. In terms of Proposition \ref{LRA}, this range corresponds to the range of lotteries with positive expectation satisfying condition \eqref{condition}. The next section illustrates Propositions \ref{LRA} and \ref{slope} by concrete examples.
\section{Examples}
Without loss of generality, for simplicity restrict attention to $x^* = 0$. In all examples I assume that there are two states of the world: the first occurs with a given probability $p$ and the second with a probability $1-p$, and in state $i\in\{1, 2\}$ the individual receives $x_i$.
\subsection{Ex. 1}
Consider $u(x) = x^3$, this is what first comes to mind when we are talking about a strictly increasing function with zero first $k$ derivatives at some point. First of all, let us derive a risk premium $\pi(t)$ for the lottery $x+t\eps$ at $x = 0$:
\begin{align}
	-\pi(t)^3 = \int t^3 \eps^3 dF(\eps)\\
	\pi(t) = -t \sqrt[3]{\expect[\eps^3]}
\end{align}
Note that if $\expect[\eps^3] = 0$ then $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} = 0$ while if $\expect[\eps^3] \neq 0$, $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} \neq 0$ which demonstrates that in general, the individual's attitude toward risk is neither of order 1 nor of order 2.
So, if $\expect[\eps^3] \neq 0$, $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} = -\sqrt[3]{\expect[\eps^3]}$ exactly as Proposition \ref{derivative} claims.

The indifference curve corresponding to the utility level $u(0) = 0$ is given by:
\begin{align*}
	px_1^3 + (1-p)x_2^3 = 0\\
	x_2 = -\sqrt[3]{\frac{p}{1-p}} x_1
\end{align*}
As Proposition \ref{slope} states, $\frac{dx_2}{dx_1}\bigg|_{x_1=0} = -\sqrt[3]{\frac{p}{1-p}}$.
Consider the plot on the Fig \ref{fig1}. 
\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{pic1}
	\caption{}\label{fig1}
\end{figure}

A blue line represents the indifference curve. Suppose that the individual initially faces the lottery $(x_1, p; x_2, 1-p)$ but in addition is offered to buy $t \ge 0$ units of insurance each at a cost $\gamma$, which would pay him $\$ 1$ in (bad) state 1. A green line is the budget line which represents lotteries with insurance for all possible $t$. An orange line represents the line of constant, zero expectation, that is, each point on this line represents the lottery with zero expected payoff. Since the slope of this line is lower (in absolute value) than the slope of the green line, we have insurance with positive loading ($\gamma > p$). At the same, since the individual's utility increases as we are moving to the north-east, it is clear that the individual strictly prefers the lottery $(0, p; 0, 1-p)$ (that is, full insurance) to any other point on the green line. Clearly, the situation above is possible only if 
\begin{align}\label{negative moment}
	\frac{p}{1-p} < \frac{\gamma}{1-\gamma}< \sqrt[3]{\frac{p}{1-p}} \Rightarrow \begin{cases}
		\gamma(1-p) - (1-\gamma)p > 0\\
		-(1-\gamma)^3p+\gamma^3(1-p)<0
		\end{cases}
\end{align}
Note that the two conditions in \eqref{negative moment} are equivalent to 
\begin{align*}
	\begin{cases}
			\expect[\eps] >0\\
	\expect[\eps^3] < 0
	\end{cases}
\end{align*}
where $\eps = (-(1-\gamma), p; \gamma, (1-p))$.These two conditions are exactly what Proposition \ref{LRA} requires for the individual with positive $u'''(0)$ being locally risk averse!
\subsection{Ex. 2}
Consider $$u(x) = \begin{cases}
	x^2, &x \ge 0\\-x^2, &x < 0
\end{cases}$$
this example is interesting since $u(x)$ does not have a second derivative at $x^* = 0$. However, there exist finite left and right second derivatives, $u''_{+}(x^*) = 2, u''_{-}(x^*) = -2$. An indifference curve corresponding to the utility level 0 is given by:
\begin{align*}
	-px_1^2 + (1-p)x_2^2 = 0,\ \text{if } x_1 \le 0, x_2 \ge 0\\
	px_1^2-(1-p)x_2^2 = 0,\ \text{if } x_1 > 0, x_2 < 0\\
\end{align*}
$$	x_2 = -\sqrt{\frac{p}{1-p}}x_1$$
So, we have the same situation as in Ex.1. That is, $$\frac{p}{1-p}<\frac{\gamma}{1-\gamma}<\sqrt{\frac{p}{1-p}}$$ guarantees local risk aversion. This condition again implies that $$\begin{cases}
	\gamma(1-p)-p(1-\gamma)>0\\
	\gamma^2(1-p) - (1-\gamma)^2p < 0
\end{cases} \iff \begin{cases}
\expect[\eps]>0\\
\expect[ \eps^k | \eps\ge0]\prob(\eps\ge0) -\expect[ \eps^k | \eps<0]\prob(\eps<0) < 0
\end{cases}$$
where $\eps = (-(1-\gamma), p; \gamma, (1-p))$. Since $u''_{-}(0) = -2, u''_{+}(0)=2$, the second condition above is equivalent to \eqref{condition} requirement in Proposition \ref{LRA} for the individual being locally risk averse.
\section{Conclusion}
	\newpage
	\bibliography{lib}{}
	%\bibliographystyle{unsrt}
	%\bibliographystyle{apalike}
	\bibliographystyle{abbrvnat}
	\end{document}

### 2021-04-01 ZeroDerivative.tex
Add files via upload

@@
 \usepackage{float}
+\usepackage[affil-it]{authblk}
 \usepackage[section]{placeins}
@@
 \newcommand{\norma}{\mathcal{N}}
+\renewcommand*{\Affilfont}{\normalfont}
 \newcommand{\expect}{\mathbb{E}}
+\newcommand{\posder}{u^{(k)}_{+}}
+\newcommand{\negder}{u^{(k)}_{-}}
 \newcommand{\eps}{\varepsilon}
@@
 \newtheorem{definition}[theorem]{Definition}
+\newtheorem{observation}[theorem]{Observation}
 %\let\cite = \citeasnoun
 \title{Zero derivative and local risk aversion}
-\date{}
-\author{Danil Fedchenko}
+\author{Danil Fedchenko\thanks{I thank Eddie Dekel for help and suggestions. Any remaining errors are mine. }}
+\affil{Northwestern University}
+\date{\today}
 \begin{document} % 
@@
 	Consider a random variable $\varepsilon$ such that $\expect[\varepsilon] = 0$.
-Denote $\pi(t)$ a risk premium, an expected-utility maximizer with Bernoulli utility $u(x)$ is willing to pay to avoid risk $t \varepsilon$ (i.e.
-$u(x-\pi(t)) = \int u(x+t\varepsilon)dF(\varepsilon)$).
-As \cite{segal1990first} stated, $\pi'(0) = 0$.
-This is true, however, only if $u'(x) \neq 0$.
-In this note I show that an attitude toward risk of an individual with $u'(x) = 0$ in some cases is similar to being what \cite{segal1990first} called of order 1.
-Which in particular implies that such an individual may find optimal to buy full insurance even for a small positive risk in the neighborhood of the points where $u'(x) = 0$.
+Denote $\pi(t, x)$ a risk premium, an expected-utility maximizer with Bernoulli utility $u(x)$ is willing to pay to avoid risk $t \varepsilon$ (i.e.
+$u(x-\pi(t, x)) = \int u(x+t\varepsilon)dF(\varepsilon)$).
+As \cite{segal1990first} stated $\frac{\partial \pi(t, x)}{\partial t}\bigg|_{t=0} = 0$ for all $x$, provided that $u'(x) \neq 0$.
+In this note I show that an attitude toward risk of an individual with $u'(x) = 0$ typically is similar to what \cite{segal1990first} called of order 1.
+This in particular implies that such an individual may find optimal to buy full insurance even for a small positive risk in the neighborhood of the points where $u'(x) = 0$.
 	\end{abstract}
@@
 Suppose this individual for some $t \in \mathbb{R}$ faces a lottery $t \eps$ represented by a random variable $\varepsilon$ such that $\expect[\varepsilon] = 0$.
-Define a risk premium $\pi(t)$ implicitly as $x-\pi(t) \sim x+t\eps$ or equivalently $u(x - \pi(t)) = \int u(x + t \eps)\ dF(\eps)$ (from now on whenever I write $\int$, I assume the integral exists).
-It is easy to show that as long as $u'(x) \neq 0$, $\pi'(0) = 0$.
+Define a risk premium $\pi(t, x)$ implicitly as $x-\pi(t, x) \sim x+t\eps$ or equivalently $u(x - \pi(t, x)) = \int u(x + t \eps)\ dF(\eps)$ (from now on whenever I write $\int$, I assume the integral exists).
+It is easy to show that as long as $u'(x) \neq 0$, $\frac{\partial \pi(t, x)}{\partial t}\bigg|_{t=0} = 0$.
 Indeed,
 	\begin{align*}
-	\pi(t) &= x - u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right),\ u^{-1} \text{ exists since }u(x)\text{ is strictly increasing and continuous}\\
-	\frac{\partial \pi(t)}{\partial t} &= -\left(u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right) \right)' \int u'(x+t \varepsilon)\varepsilon dF(\varepsilon) = \\
+	\pi(t, x) &= x - u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right) \footnotemark\\
+	\frac{\partial \pi(t, x)}{\partial t} &= -\left(u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right) \right)' \int u'(x+t \varepsilon)\varepsilon dF(\varepsilon) = \\
 	&=-\frac{\int u'(x+t \varepsilon)\varepsilon dF(\varepsilon)}{u'\left(u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right)\right)}\ ,\ \ \  \text{if }u'\left(u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right)\right) \neq 0\\
-	\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} &= -\frac{u'(x)}{u'(x)} \expect[\eps] = -\expect[\eps] = 0, \text{ if } u'(x) \neq 0
+	\frac{\partial \pi(t, x)}{\partial t}\bigg|_{t=0} &= -\frac{u'(x)}{u'(x)} \expect[\eps] = -\expect[\eps] = 0, \text{ if } u'(x) \neq 0
 	\end{align*}
-This in particular implies that for small enough $t$ $$\pi(t) \approx -\frac{t^2 \expect[\eps^2]}{2}\frac{u''(x)}{u'(x)}$$ and hence it approaches zero faster than $t$ which, in turn, implies that for small $t$ an individual is almost risk neutral.
-This assertion, however, heavily relies upon the assumption $u'(x) \neq 0$ which is known sometimes to be violated even for strictly increasing functions ($u(x) = x^3$ at $x=0$ is probably the most famous example).
-In this paper I partially characterize $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0}$ at points where $u'(x) = 0$, and show that it is not equal to zero in general.
-Applying this to the optimal insurance problem, as \cite{segal1990first} show, means that an expected utility maximizer may find optimal to buy full insurance even for small risk and even if insurance includes positive loading.
-I illustrate this result by deriving a slope of indifference curves (provided that it is non zero and finite) of the individual near the certainty line on the state-contingent  \cite{yaari1978some}-type diagram.
-It turns out that the slope differs from the case where $u'(x) \neq 0$.
-\section{Order of Risk Aversion}
-\cite{segal1990first} define an attitude toward risk of individual being of order 1 if for \textit{every} non-degenerate random variable $\eps$ such that $\expect[\eps] = 0$, $\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0^+} \neq 0$, and of order 2 if $\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0^+} = 0$.
-In the next proposition I derive $\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0}$ (assuming that it exists, is non-zero and finite) at the points where for some $k\ge 1$ first $k-1$ derivatives of the Bernoulli utility function $u(x)$ exist and equal to zero.
+\footnotetext{$u^{-1} \text{ exists since }u(x)\text{ is strictly increasing and continuous}$}
+Using this and continuing similarly one can show that for small enough $t$ $$\pi(t, x) \approx -\frac{t^2 \expect[\eps^2]}{2}\frac{u''(x)}{u'(x)}$$ and hence it approaches zero faster than $t$ which, in turn, implies that for small $t$ an individual is almost risk neutral.
+This assertion, however, relies upon the assumption $u'(x) \neq 0$ which may not only be violated for strictly increasing functions at some point ($u(x) = x^3$ at $x=0$ is probably the most famous example) but even more so, there exist strictly increasing and continuous functions whose derivative equals zero almost everywhere.
+
+In this paper, in Section \ref{order},  I derive $\frac{\partial \pi(t, x)}{\partial t}\bigg|_{t=0}$ at points where $u'(x) = 0$, and show that it is not equal to zero in general\footnote{One might mistakenly conjecture  that if $\frac{\partial \pi(t, x_n)}{\partial t}\bigg|_{t=0}=0$ for some $x_n \underset{n \to \infty}{\to}x$, then $\frac{\partial \pi(t, x)}{\partial t}\bigg|_{t=0}=0$, but our results show such a conjecture is typically false.
+Which means that $\frac{\partial \pi(t, x)}{\partial t}$ typically is not continuous at $(0, 0)$}.
+In Section \ref{insurance}, I apply this to the optimal insurance problem and show that an expected utility maximizer whose Bernoulli utility has zero derivative at some point may exhibit local risk aversion.
+That is, such an individual may find optimal to buy full insurance even for small risk and even if insurance includes positive loading.
+I illustrate this result by deriving a slope of indifference curves of the individual near the certainty line on the state-contingent  \cite{yaari1978some}-type diagram where we see that the slope differs from the case where $u'(x) \neq 0$.
+In Section \ref{Bsingular}, I do the same exercise considering an individual whose Bernoulli utility function is strictly increasing, continuous but has zero derivative almost everywhere.
+Finally, in Section \ref{Examples}, I consider several concrete examples which illustrate general results derived in the previous sections.
+\section{Order of Risk Aversion}\label{order}
+\cite{segal1990first} define an attitude toward risk of individual being of order 1 if for every non-degenerate random variable $\eps$ such that $\expect[\eps] = 0$, $\frac{\partial \pi(t, x)}{\partial t} \bigg|_{t = 0^+} \neq 0$ for all $x$, and of order 2 if $\frac{\partial \pi(t, x)}{\partial t} \bigg|_{t = 0^+} = 0$ for all $x$.
+In the next proposition I derive $\frac{\partial \pi(t, x)}{\partial t} \bigg|_{t = 0}$ (assuming that it exists) at a point $x^*$ where for some $k\ge 1$ the first $k-1$ derivatives of the Bernoulli utility function $u(x)$ exist and equal zero.
+Hereafter I assume that $\pi(t, x)$ is continuous in $t$ at $x^*$.
 	\begin{proposition}\label{derivative}
-		If $\exists\ x^* \in \mathbb{R}$ and $k \ge 2$ such that $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, $u'(x) > 0$ for $x \in (x^*-\delta, x^*+\delta)\setminus \{x^*\}$ for some small $\delta>0$, and there exists $u^{(k)}(x^*) \neq 0$.
-Then if there exists finite and non-zero $\expect[\eps^k]$, and at $x^*$ there exists finite, non-zero $\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0}$ then :
+		Suppose $\exists\ x^* \in \mathbb{R}$ and $k \ge 2$ such that $u(x)$ is $k$ times differentiable on $(x^*-\delta, x^*+\delta)$ for some $\delta>0$, and $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, while $u^{(k)}(x^*) \neq 0$.
+Then if $\expect|\eps^k| < \infty$ and at $x^*$ there exists $\frac{\partial \pi(t, x)}{\partial t} \bigg|_{t = 0}$ :
 		\begin{align}
-			\left( \frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0} \right)^k = (-1)^k\ \expect[\varepsilon^k]
+			\left( \frac{\partial \pi(t, x)}{\partial t} \Bigg|_{\substack{t = 0\\x = x^*}} \right)^k = (-1)^k\ \expect[\varepsilon^k]
 		\end{align}
-	Provided that \begin{align}\label{limsup}
-	\forall\ i = \{2,\dots, k-1\}:\ \bigg|\lim_{t\to0} \inf \frac{\partial^{(i)} \pi(t)}{\partial t^{(i)}}\bigg| < \infty, \bigg|\lim_{t\to0} \sup \frac{\partial^{(i)} \pi(t)}{\partial t^{(i)}}\bigg| < \infty
-\end{align} for $k>2$.
 	\end{proposition}
 \begin{proof}
+	By definition of $\pi(t, x)$
 	\begin{align*}
-		\pi(t) &= x^* - u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\\
-		\frac{\partial \pi(t)}{\partial t} &= -\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right) \right)' \int u'(x^*+t \varepsilon)\varepsilon dF(\varepsilon) = \\
-		&=-\frac{\int u'(x^*+t \varepsilon)\varepsilon dF(\varepsilon)}{u'\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right)}\ ,\ \ \  \text{if }u'\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right) \neq 0
+		u(x-\pi(t, x)) &= \int u(x+t\eps)dF(\eps)
 	\end{align*}
-As $t \to 0$ since $u'(x^*) = 0$, both numerator and denominator approach zero.
-Using the L'Hospital rule we get:
+expanding this in the neighborhood of $t=0$ for $x=x^*$ and using $\pi(0, x^*) = 0, u^{(i)}(x^*) = 0\ \forall\ i\in\{1, \dots, k-1\}$, we get
+\begin{align}\label{deriva}
+	u(x^*) + (-1)^k \frac{u^{(k)}(x^*)}{k!}\pi(t, x^*)^k + o(\pi(t, x^*)^k) &= \int \left(u(x^*) + \frac{u^{(k)}(x^*)t^k\eps^k}{k!}+o(t^k\eps^k) \right)dF(\eps) \nonumber\\
+(-1)^k \frac{u^{(k)}(x^*)}{k!}\pi(t, x)^k + \pi(t, x^*)^ko(1) &= \int \left(\frac{u^{(k)}(x^*)t^k\eps^k}{k!}+t^k\eps^ko_{\eps}(1) \right)dF(\eps)
+\end{align}
+where all $o(\cdot)$ are defined as $t \to 0$.
+Since for $t$ small enough $|\eps^ko_{\eps}(1)| < |\eps^k|$ and $\expect|\eps^k|<\infty$, by the dominated convergence theorem $\lim_{t\to0}\expect[\eps^ko_{\eps}(1)] = \expect[\lim_{t\to0}\eps^ko_{\eps}(1)] = 0$, i.e.
+$\expect[\eps^ko_{\eps}(1)] = o(1), t\to0$.
+Then \eqref{deriva} implies
+\begin{align}\label{deriv}
+	\left(\frac{\pi(t, x^*)}{t}\right)^k = \frac{\left( \frac{u^{(k)}(x^*)\expect[\eps^k]}{k!} + o(1) \right)}{\left( (-1)^k \frac{u^{(k)}(x^*)}{k!} + o(1)) \right)}
+\end{align}
+Since
 \begin{align*}
-		\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0} = \lim_{t \to 0} -\frac{\int u''(x^*+t \varepsilon) \varepsilon^2 dF(\varepsilon)}{u''\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right) \underbrace{\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right) \right)'}_{-\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0}}}
+	\frac{\partial \pi(t, x^*)}{\partial t}\bigg|_{t = 0} = \lim_{t\to0}\frac{\pi(t, x^*)}{t}
 \end{align*}
-If $u''(x^*) \neq 0$ (i.e.
-$k=2$) then $\left( \frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0} \right)^2 = \expect[\varepsilon^2]$.
-If $k>2$ then one should apply L'Hospital rule once again.
-That is,
+from \eqref{deriv} it follows that
 \begin{align*}
-	\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0} = \lim_{t \to 0} -\frac{\int u'''(x^*+t \varepsilon) \varepsilon^3 dF(\varepsilon)}{u'''\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right) \left( \frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0}\right)^2 - u''\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right) \frac{\partial^2\pi(t)}{\partial t^2} }
+	\left( \frac{\partial \pi(t, x)}{\partial t} \bigg|_{\substack{t = 0\\x = x^*}} \right)^k = (-1)^k\ \expect[\varepsilon^k]
 \end{align*}
-Since $u''(x^*) = 0$ and \eqref{limsup} implies  $$\lim_{t \to 0} u''\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right) \frac{\partial^2\pi(t)}{\partial t^2} = 0,$$ if $k=3$ then we have $$\left( \frac{\partial \pi(t)}{\partial t} \bigg|_{t=0}\right)^3 = -\expect[\eps^3] $$ which is what we sought to prove.
-If $k>3$ we should proceed in the same way, i.e.
-by applying L'Hospital rule $k-3$ times we get what we wanted to prove.
 \end{proof}
-Although, the proposition above does not tell us whether $\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0} \neq 0$, so we cannot conclude that the individual's attitude toward risk is necessary of order 1 (according to \cite{segal1990first}).
-We will see in Section 4 that, in general neither $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} \neq 0$ nor $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} = 0$, so the individual's attitude toward risk is neither of order 1 nor of order 2.
-Nevertheless, if it happens to be $\expect[\eps^k] \neq 0$ and $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} \neq 0$, optimal behavior of such an individual regarding the choice of insurance is similar to that of the one whose attitude toward risk is of order 1.
-\section{Optimal Insurance}
-Discussing the optimal insurance problem consider the lottery $x^*+t \eps $ where $\expect[\eps] > 0$. Then, the following proposition is an analogue to the Proposition 1 of \cite{segal1990first}.
+Thus, if $u'(x) = 0$ the individual's attitude toward risk is in general neither of order 1 nor of order 2, according to \cite{segal1990first} (e.g.
+for $k=3$ it could be the case $\expect[\eps^3] = 0$) Nevertheless, if it happens to be $\expect[\eps^k] \neq 0$ (where $k$ is defined as in the proposition above) and $\frac{\partial \pi(t, x)}{\partial t}\bigg|_{\substack{t=0\\x = x^*}} \neq 0$, optimal behavior of such an individual regarding the choice of insurance in a neighborhood of $x^*$ is similar to that of the one whose attitude toward risk is of order 1.
+\section{Optimal Insurance}\label{insurance}
+Discussing the optimal insurance problem, for $t \ge 0$ consider the lottery $x^*+t \eps $ where \textbf{$\expect[\eps] > 0$}.
+Then, the following proposition is an analogue to Proposition 1 of \cite{segal1990first}.
 \begin{proposition}\label{LRA}
-	If $\exists\ x^* \in \mathbb{R}$ and $k \ge 2$ such that $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, there exist non-zero left and right derivatives $u^{(k)}_{+}(x^*)$ and $u^{(k)}_{-}(x^*)$, and $u(x)$ is $k$ times differentiable on $(x^*-\delta, x^*+\delta)\setminus\{x^*\}$ for some $\delta>0$.
-If $\eps$ possesses all finite moments, and in particular, there exist $\expect[ \eps^k |\eps\ge0]$, $\expect[ \eps^k |\eps<0]$ and \begin{align}\label{condition}
-		u^{(k)}_{+}(x^*) \cdot \expect[ \eps^k | \eps\ge0]\prob(\eps\ge0) + u^{(k)}_{-}(x^*) \cdot \expect[ \eps^k | \eps<0]\prob(\eps<0) < 0,
+	Suppose $\exists\ x^* \in \mathbb{R}$ and $k \ge 2$ such that $u(x)$ is $k$ times differentiable on $(x^*-\delta, x^*+\delta)\setminus\{x^*\}$ for some $\delta>0$, and $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$.
+In addition, there exist finite left and right derivatives $u^{(k)}_{+}(x^*)$ and $u^{(k)}_{-}(x^*)$ such that at least one of these derivatives is non-zero.
+If there exist finite $\expect| \eps^k |\eps\ge0]$, $\expect[ \eps^k |\eps<0]$ and \begin{align}\label{condition}
+		u^{(k)}_{+}(x^*) \cdot \expect[ \eps^k | \eps\ge0]\ \prob(\eps\ge0) + u^{(k)}_{-}(x^*) \cdot \expect[ \eps^k | \eps<0]\ \prob(\eps<0) < 0,
 		\end{align}
@@
 	\begin{align*}
-		\int u(x^*+t\eps)dF(\eps) = \int \left( u(x^*) +  \bigg[u^{(k)}_{+}(x^*)\mathbb{I}[\eps\ge0]  + u^{(k)}_{-}(x^*)\mathbb{I}[\eps<0] \bigg] \frac{t^k \eps^k }{k!} +o_p(t^k \eps^k)\right)dF(\eps) = \\
-	= u(x^*) + \frac{(u^{(k)}_{+}(x^*) \expect[\eps^k|\eps\ge 0]\prob(\eps\ge0)+u^{(k)}_{-}(x^*) \expect[\eps^k|\eps< 0]\prob(\eps<0))t^k}{k!} + \expect[o_p(t^k \eps^k)] < u(x^*)
+		\int u(x^*+t\eps)dF(\eps) = \int \left( u(x^*) +  \bigg[u^{(k)}_{+}(x^*)\mathbb{I}[\eps\ge0]  + u^{(k)}_{-}(x^*)\mathbb{I}[\eps<0] \bigg] \frac{t^k \eps^k }{k!} +o(t^k \eps^k)\right)dF(\eps) = \\
+	= u(x^*) + \frac{(u^{(k)}_{+}(x^*) \expect[\eps^k|\eps\ge 0]\prob(\eps\ge0)+u^{(k)}_{-}(x^*) \expect[\eps^k|\eps< 0]\prob(\eps<0))t^k}{k!} + t^k\expect[\eps^k o_{\eps}(1)] < u(x^*)
 	\end{align*}
-where the first equality follows from Taylor formula and the fact that $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, and the last inequality follows from \eqref{condition}.
-Note that for the $k$ times differentiable $u(x)$, \eqref{condition} becomes $u^{(k)}(x^*)\cdot\expect[\eps^k] < 0$
+where the first equality follows from Taylor formula and the fact that $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, and the last inequality follows from \eqref{condition} and that $\lim_{t\to0}\expect[\eps^ko_{\eps}(1)]=0$, as in the proof of Proposition \ref{derivative} (here $\expect|\eps^k| < \infty$, since there exist finite $\expect[ \eps^k |\eps\ge0]$, $\expect[ \eps^k |\eps<0]$).
 \end{proof}
-The result above means that the expected-utility maximizer under some conditions exhibits local risk aversion at the vicinity of the points where his strictly increasing Bernoulli utility function fails to have positive derivatives up to the order $k$.
-In other words, this individual prefers a certain outcome to the lottery with small positive expected value.
+Note that for a $k$ times differentiable $u(x)$, \eqref{condition} becomes $u^{(k)}(x^*)\cdot\expect[\eps^k] < 0$, meaning that if $k$ is the order of first non-zero derivative of $u(x)$ then its sign times the sign of the $k$-th moment determines if the person is locally risk loving or risk averse.
+In other words, the result above means that the expected-utility maximizer under some conditions exhibits local risk aversion (or risk loving) at the vicinity of the points where his strictly increasing Bernoulli utility function fails to have positive derivatives up to the order $k$.
+Consequently, this individual (in the case of risk-aversion) prefers a certain outcome to the lottery with a small positive expected value.
 This result stands in contrast to the well-known result that the expected-utility maximizer with increasing Bernoulli utility function $u(x)$ is locally risk neutral as long as $u'(x) > 0$.
-That is, for $t$ small enough he would strictly prefer the lottery with positive mean to the certain outcome.
+That is, for $t$ small enough he would strictly prefer the lottery with a positive mean to the certain outcome.
 
-Let me illustrate the proposition \ref{LRA} by deriving slopes of the individual's indifference curves at a certainty line on the state-contingent diagram. 
-Consider an individual whose expected utility preferences represented by strictly increasing and continuous Bernoulli utility function $u(x)$.
+I illustrate Proposition \ref{LRA} by deriving slopes of the individual's indifference curves at a certainty line on the state-contingent diagram. 
+Consider an individual whose expected utility preferences are represented by strictly increasing and continuous Bernoulli utility function $u(x)$.
 Suppose there are two states of nature, the first happens with probability $p$ and the second with probability $1-p$.
@@
 \begin{proposition}\label{slope}
-	If $\exists\ x^* \in \mathbb{R}$ and $k \ge 2$ such that $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, $u'(x) > 0$ for $x \in (x^*-\delta, x^*+\delta)\setminus \{x^*\}$ for some small $\delta>0$.
-If there exist left and right derivatives $u_{+}^{(k)}(x^*) \neq 0$, $u_{-}^{(k)}(x^*) \neq 0$ such that \footnote{$$sign(x) = \begin{cases}
-			-1, x <0\\
-			0, x=0\\
-			1, x>0
-		\end{cases}$$}$\text{sign}(u_{+}^{(k)}(x^*) \cdot u_{-}^{(k)}(x^*)) = (-1)^{k+1}$
+	Suppose $\exists\ x^* \in \mathbb{R}$ and $k \ge 2$ such that $u(x)$ is $k$ times differentiable on $(x^*-\delta, x^*+\delta)\setminus \{x^*\}$ for some $\delta>0$, and $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$.
+In addition, there exist left and right derivatives: $u_{-}^{(k)}(x^*)$, $u_{+}^{(k)}(x^*)$ such that  at least one of these derivatives is non-zero.
 
@@
 	\end{align}
-then if there exist finite, non-zero $\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0}$,  $\frac{d x_2}{d x_1} \bigg|_{x_1=x^*-0}$:
+then if there exist $\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0}$,  $\frac{d x_2}{d x_1} \bigg|_{x_1=x^*-0}$ then
 	\begin{align}
@@
 	\end{align}
-Provided that 
-\begin{align}\label{der}
-\forall\ i = \{1, \dots, k-1\}:\ \bigg| \lim_{x_1 \to x^* \pm 0} \sup \frac{d^{(i)}x_2(x_1)}{dx_1^{(i)}}\bigg| < \infty, \bigg| \lim_{x_1 \to x^* \pm 0} \inf \frac{d^{(i)}x_2(x_1)}{dx_1^{(i)}}\bigg| < \infty
-\end{align}
+\end{proposition}
+In the proof of this proposition we will need the following lemma
+\begin{lemma}\label{lemma}
+	Under conditions of Proposition \ref{slope}:
+	$$\footnote{$sign(x) = \begin{cases} -1, &x < 0\\ 0, &x = 0\\ 1, &x > 0 \end{cases}$}sign\left(\frac{\posder(x^*)}{\negder(x^*)} \right) \neq (-1)^k$$
+	where we assume that ($sign(-\infty) = -1, sign(+\infty) = 1$)
+\end{lemma}
+\begin{proof}
+	Suppose $k=2$, without loss of generality assume $x^* = 0$ and $u(x^*) = 0$. Then $u'(0) = 0$. Since $u$ is twicely differentiable to the left and to the right of zero, there should exist 
+	\begin{align*}
+		u''_{-}(0) = \lim_{\varepsilon \to 0^-} \frac{u'(\varepsilon)}{\varepsilon},\  u''_{+}(0) = \lim_{\varepsilon \to 0^+} \frac{u'(\varepsilon)}{\varepsilon}
+		\end{align*}
+	By assumption, $u$ is strictly increasing and differentiable in a neighborhood of $x^*$ which means that $u'(x) \ge 0$. Hence \begin{align*}
+		u''_{-}(0) = \lim_{\varepsilon \to 0^-} \frac{u'(\varepsilon)}{\varepsilon} \le 0\\
+		u''_{+}(0) = \lim_{\varepsilon \to 0^+} \frac{u'(\varepsilon)}{\varepsilon} \ge 0
+	\end{align*}
+and consequently $sign\left(\frac{u''_{+}(x^*)}{u''_{-}(x^*)} \right) \neq 1$.
+
+Suppose $k = 3$ then $u'(0) = u''(0) = 0$ and as we saw for the case of $k=1$, 
+ \begin{align*}
+	u''_{-}(0) \le 0,\ u''_{+}(0) \ge 0
+\end{align*}
+then \begin{align*}
+	u^{(3)}_{-}(0) = \lim_{\varepsilon \to 0^-} \frac{u''(\varepsilon)}{\varepsilon} \ge 0\\
+	u^{(3)}_{+}(0) = \lim_{\varepsilon \to 0^+} \frac{u''(\varepsilon)}{\varepsilon} \ge 0
+\end{align*}
+meaning that $sign\left(\frac{u^{(3)}_{+}(x^*)}{u^{(3)}_{-}(x^*)} \right) \neq -1$. Proceeding in the same way, we can prove the result for an arbitrary $k$.
+\end{proof}
+\begin{proof} of Proposition \ref{slope}
+
+	Consider small $\epsilon > 0$, then
+	\begin{align*}
+		u(x^* + \epsilon) = u(x^*) + \frac{u^{(k)}_{+}(x^*) \epsilon^k}{k!} + o(\epsilon^k)\\
+		u(x^* + \epsilon)= u(x^*)+\epsilon^k\left( \frac{u^{(k)}_{+}(x^*)}{k!} + o(1) \right)
+	\end{align*}
+where $o(\cdot)$ is defined as $\epsilon \to 0$.
+Similarly, for small $\omega <0$
+\begin{align*}
+	u(x^* + \omega) = u(x^*) + \frac{u^{(k)}_{-}(x^*) \omega^k}{k!} + o(\omega^k)\\
+	u(x^* + \omega) = u(x^*) + \omega^k\left( \frac{u^{(k)}_{-}(x^*)}{k!} + o(1) \right)
+\end{align*}
+where $o(\cdot)$ is defined as $\omega \to 0$.
+Consider $x_1 = x^*+\epsilon$, then since $u(x)$ is strictly increasing implies that $x_2 = x^* + \omega(\epsilon)$. Then,
+\begin{align*}
+	\frac{dx_2}{dx_1}\bigg|_{x_1=x^*+0} = \lim_{\epsilon \to 0^+} \frac{\omega(\epsilon)}{\epsilon}
+\end{align*}
+and \eqref{ind} implies:
+\begin{align*}
+		p\epsilon^k\left( \frac{u^{(k)}_{+}(x^*)}{k!} + o(1) \right) + (1-p)\omega^k(\epsilon)\left( \frac{u^{(k)}_{-}(x^*)}{k!} + o(1) \right) = 0\\
+		\left(\frac{\omega(\epsilon)}{\epsilon}\right)^k = -\frac{p}{1-p} \frac{\left( \frac{u^{(k)}_{+}(x^*)}{k!} + o(1) \right)}{\left( \frac{u^{(k)}_{-}(x^*)}{k!} + o(1) \right)}
+\end{align*}
+By Lemma \ref{lemma} $sign\left(\frac{\posder(x^*)}{\negder(x^*)} \right) \neq (-1)^k$ then 
+\begin{align*}
+		\frac{dx_2}{dx_1}\bigg|_{x_1=x^*+0} = \lim_{\epsilon \to 0^+}\frac{\omega(\epsilon)}{\epsilon} = -\sqrt[k]{\frac{p}{1-p} \bigg|\frac{\posder(x^*)}{\negder(x^*)} \bigg| }
+\end{align*}
+$\frac{dx_2}{dx_1}\bigg|_{x_1=x^*-0}$ can be found in exactly the same way.
+\end{proof}
+In contrast to the result above it is known that in case of differentiable Bernoulli utility, the slope of an expected-utility maximizer's indifference curve at certainty points where $u'(x) \neq 0$ is equal to $-\frac{p}{1-p}$.
+If this individual is offered to buy insurance which would cost $\gamma$ for every insured unit and pay $\$\ 1$ in case of the bad event occurs, then the individual's budget constraint would have a slope $-\frac{\gamma}{1-\gamma}$.
+This in particular explains why such an expected-utility maximizer never buys full insurance unless it is actuarially fair (i.e.
+$\gamma = p$).
+At the same time, as we see from Proposition \ref{slope} an individual with $k$ times differentiable Bernoulli utility with the first $k-1$ derivatives equal to zero at some $x$ has slope of the indifference curve at $(x, x)$ given by $-\sqrt[k]{\frac{p}{1-p}}$.
+So depending on $p$ there may exist a range of positive loading $\gamma > p$ for which the expected utility maximizer would find it optimal to buy full insurance.
+In terms of Proposition \ref{LRA}, this range corresponds to the range of lotteries with positive expectation satisfying condition \eqref{condition}.
+Section \ref{Examples} illustrates Propositions \ref{LRA} and \ref{slope} by concrete examples.
+
+But before going to examples, let's consider another class of Bernoulli utility functions that have zero derivative.
+In particular, consider strictly increasing and continuous utility functions which have first derivative equals zero not at a particular point but almost everywhere.
+
+\section{Strictly singular Bernoulli utility function}\label{Bsingular}
+\begin{definition}
+	A continuous, non-constant function $f:\mathbb{R} \to \mathbb{R}$ is called (strictly) \textbf{singular} if it is (strictly) monotone and has zero derivative almost everywhere i.e.
+$$f'(x) = 0 \text{ a.e.}$$.
+\end{definition}
+For examples of such functions see \cite{salem1943some} or \cite{berg2000rham}.
+From now on, unless otherwise stated I will consider increasing singular functions defined on the interval $[a, b]$ for some $a, b \in \mathbb{R}, a<b$.
+For the result below it is useful to have the following proposition.
+\begin{proposition}\label{inverse}
+	If $f$ is a strictly singular function, then the inverse $g = f^{-1}$ is 
+	also strictly singular.
 \end{proposition}
 \begin{proof}
-	From \eqref{ind}, implicit function theorem and the fact that $u'(x) > 0$ for $x \in (x^*-\delta, x^*+\delta)\setminus \{x^*\}$ it follows that
+	See \cite{berg2000rham}
+\end{proof}
+Suppose an individual has a strictly singular Bernoulli utility function $u: [a, b] \to [u(a), u(b)]$.
+Denote $$E = \{ y \in [u(a), u(b)]: \text{ either } \nexists\ (u^{-1})'(y) \text{ or }\exists\ (u^{-1})'(y) \neq 0 \}.$$ By Proposition \ref{inverse}, Lebesgue measure of $E$ equals zero.
+For the next propositions we will be interested in a set $D = u^{-1}(E) = \{x\in[a, b]: u(x) \in E\}$, the preimage of set $E$ under the mapping $u$.
+\begin{proposition}\label{pi}
+	Suppose $u(x)$ is strictly singular.
+Then $\forall\ x \in [a, b]\setminus D: \frac{\partial \pi (t, x)}{\partial t} \bigg|_{t=0} = 0$.
+In particular, if $|D| = 0$, $\frac{\partial \pi (t, x)}{\partial t} \bigg|_{t=0} = 0\ \text{a.e.}$
+\end{proposition}
+\begin{proof}
+	By definition
 	\begin{align*}
-		\lim_{x_1 \to x^*+0} \frac{dx_2}{dx_1} = \lim_{x_1 \to x^*+0} -\frac{p}{1-p} \frac{u'(x_1)}{u'(x_2(x_1))}
+		\pi(t, x) &= x - u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right)\\
+		\frac{\partial \pi(t, x)}{\partial t} &= -\left(u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right) \right)' \int u'(x+t \varepsilon)\varepsilon dF(\varepsilon)\\
+		\frac{\partial \pi(t, x)}{\partial t} \bigg|_{t=0} &= -\left(u^{-1}\left( u(x) \right) \right)' \int u'(x)\varepsilon dF(\varepsilon)
 	\end{align*}
-Since $x_2(x^*) = x^*$ and $u'(x^*) = 0$ both numerator and denominator approach zero as $x_1 \to x^*+0$. Then if $	\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0}$ exists we can use L'Hospital rule, i.e.
-\begin{align*}
-		\lim_{x_1 \to x^*+0} \frac{dx_2}{dx_1} = -\frac{p}{1-p} \frac{u''_{+}(x_1)}{u''_{-}(x_2(x_1)) \left(\frac{dx_2}{dx_1}\right)\bigg|_{x^*+0}}
+Since $\forall\ x \in [a, b]\setminus D,\ u(x) \notin E\ \to\ (u^{-1}(u(x))) = 0$.
+ Thus, $$\frac{\partial \pi(t, x)}{\partial t}\bigg|_{t = 0} = 0$$
+\end{proof}
+\begin{corollary}
+	If $u$ is strictly singular and a preimage of any null set has Lebesgue measure zero, then $$\frac{\partial \pi (t, x)}{\partial t} \bigg|_{t=0} = 0\ \text{a.e.}$$
+\end{corollary}
+\begin{proof}
+If a preimage of any null set has Lebesgue measure zero, then $|D| = 0$.
+\end{proof}
+As we can see, an individual with a strictly singular Bernoulli utility function under some conditions has an attitude toward risk of order 2.
+This stands in contrast to the case considered in Section \ref{order} where an individual has Bernoulli utility with zero derivatives up to some order at a particular point.
+At the same time, as we know, the same result holds in case an individual has a utility function with $u'(x) > 0$.
+However, this is more of a coincidence rather than something general.
+The next proposition shows that characteristics of indifference curves of an individual with strictly singular Bernoulli utility function differ from that of an individual with $u'(x) > 0$.
+In particular, the slope (derivative) of the indifference curve does not equal $-\frac{p}{1-p}$ almost everywhere.
+\begin{proposition}\label{singular}
+	Suppose $u(x)$ is strictly singular.
+Consider an indifference curve $x_2(x_1)$ implicitly given by $$\{ (x_1, x_2):pu(x_1)+(1-p)u(x_2) = u(x^*) \}$$ for some $x^*$ between $x_1$ and $x_2$.
+Denote $F(x^*) = \{ x_1 \in [a, b]: \frac{u(x^*) - p u(x_1)}{1-p} \in E\}$.
+Then $$\forall\ x_1 \in [a, b] \setminus F(x^*):  \frac{dx_2(x_1)}{dx_1} = 0$$ In particular, if $|F(x^*)| = 0$ then $$\frac{dx_2(x_1)}{dx_1} = 0 \text{ a.e.}$$
+\end{proposition}
+\begin{proof}
+	Since $u(x)$ is strictly increasing and continuous, there exists strictly increasing and continuous $u^{-1}$ which is also strictly singular by Proposition \ref{inverse}. \begin{align*}
+		x_2(x_1) = u^{-1} \left( \frac{u(x^*) - pu(x_1)}{1-p} \right)
 	\end{align*}
-We have left second derivative in the denominator since from the fact that $u(x)$ is strictly increasing it follows that the indifference curve should be downward slopping.
-Meaning that if $x_1$ approaches $x^*$ from above, $x_2(x_1)$ approaches $x^*$ from below.
+By the chain rule then \begin{align*}
+	\frac{dx_2(x_1)}{dx_1} = -(u^{-1})'\left( \frac{u(x^*) - pu(x_1)}{1-p} \right)\frac{p}{1-p}u'(x_1)
+\end{align*}
+Since $\forall\ x_1 \in [a, b] \setminus F(x^*):\ \frac{u(x^*) - pu(x_1)}{1-p} \notin E\to (u^{-1})'\left( \frac{u(x^*) - pu(x_1)}{1-p} \right) = 0$. Hence, $$\frac{dx_2(x_1)}{dx_1} = 0$$
+If Lebesgue measure of $F(x^*)$ is zero then $$\frac{dx_2(x_1)}{dx_1} = 0 \text{ a.e.}$$
+\end{proof}
+\begin{corollary}
+	If $u$ is strictly singular and a preimage of any null set has Lebesgue measure zero, then $$\frac{dx_2(x_1)}{dx_1} = 0 \text{ a.e.}$$
+\end{corollary}
+\begin{proof}
+If a preimage of any null set has Lebesgue measure zero, then $|D| = 0$ but since $\frac{u(x^*) - pu(x_1)}{1-p}$ is simply an affine transformation of $u(x_1)$, clearly $|F(x^*)| = 0$ as well.
+\end{proof}
+Talking about optimal insurance, it seems that from the proposition above it follows that similar to the case with zero derivatives at a particular point, an individual with a strictly singular Bernoulli utility function must admit local risk aversion and prefer full insurance.
+This, however, is not completely true.
+The reason is that for singular functions the derivative can hardly be interpreted in a usual manner (as a slope).
+Thus, even though mathematically the derivative of the function which defines indifference curves indeed may be equal to zero, for the typical singular function, indifference curves will be ``very irregular'', meaning that it is impossible to say in general which choice of insurance (even locally) would be optimal for such an individual.
 
-If $k = 2$ and $u''_{+}(x^*) \cdot u''_{-}(x^*) < 0$ then $$	\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0} = -\sqrt{\frac{p}{1-p} \bigg|  \frac{u''_{+}(x^*)}{u''_{-}(x^*)} \bigg|  }$$
-Minus sign again follows from downward slopping indifference curve. Left derivative can be found in exactly the same way.
-
-If $k > 2$ then apply the L'Hospital rule once again. We get 
-\begin{align*}
-	\lim_{x_1 \to x^*+0} \frac{dx_2}{dx_1} = -\frac{p}{1-p} \frac{u'''_{+}(x_1)}{u'''_{-}(x_2(x_1)) \left(\frac{dx_2}{dx_1}\right)^2\bigg|_{x^*+0} + u''_{-}(x_2(x_1))\left(\frac{d^2x_2}{dx_1^2}\right)\bigg|_{x^*+0} }
-\end{align*}
-From \eqref{der} and $u''(x^*) = 0$ it follows that $$\lim_{x_1 \to x^*+0} u''_{-}(x_2(x_1))\left(\frac{d^2x_2}{dx_1^2}\right)\bigg|_{x^*+0} = 0$$
-If $k = 3$ and $u'''_{+}(x^*) \cdot u'''_{-}(x^*) > 0$ then $$	\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0} = -\sqrt[3]{\frac{p}{1-p} \bigg|  \frac{u'''_{+}(x^*)}{u'''_{-}(x^*)} \bigg|  }$$
-which is what we sought to prove.
-If $k>3$ then one should apply the L'Hospital rule as many times as it requires to get rid of indeterminacies in numerator and denominator.
-Which eventually leads to what we wanted to prove.
-\end{proof}
-In contrast to the result above it is known that in case of differentiable Bernoulli utility, slope of an expected-utility maximizer's indifference curve at certainty points where $u'(x) \neq 0$ is equal to $-\frac{p}{1-p}$.
-If this individual is offered to buy insurance which would cost $\gamma$ for every insured unit and pay $\$\ 1$ in case of the bad event occurs, then the individual's budget constraint would have a slope $-\frac{\gamma}{1-\gamma}$.
-This in particular explains why such expected-utility maximizer never buys full insurance unless it is actuarially fair (i.e.
-$\gamma = p$).
-At the same time, as we see from the Proposition \ref{slope} an individual with $k$ times differentiable Bernoulli utility with first $k-1$ zero derivatives at some $x$ has slope of the indifference curve at $(x, x)$ given by $-\sqrt[k]{\frac{p}{1-p}}$.
-Which means that there exists a range of positive loading $\gamma > p$ such that the expected utility maximizer would find optimal to buy full insurance.
-In terms of Proposition \ref{LRA}, this range corresponds to the range of lotteries with positive expectation satisfying condition \eqref{condition}.
-The next section illustrates Propositions \ref{LRA} and \ref{slope} by concrete examples.
-\section{Examples}
+\section{Examples}\label{Examples}
 Without loss of generality, for simplicity restrict attention to $x^* = 0$.
 In all examples I assume that there are two states of the world: the first occurs with a given probability $p$ and the second with a probability $1-p$, and in state $i\in\{1, 2\}$ the individual receives $x_i$.
-\subsection{Ex. 1}
-Consider $u(x) = x^3$, this is what first comes to mind when we are talking about a strictly increasing function with zero first $k$ derivatives at some point.
+\subsection{Example 1}
+Consider $u(x) = x^3$, this is what first comes to mind when we are talking about a strictly increasing function with zero first derivative at some point.
 First of all, let us derive a risk premium $\pi(t)$ for the lottery $x+t\eps$ at $x = 0$:
 \begin{align}
-	-\pi(t)^3 = \int t^3 \eps^3 dF(\eps)\\
-	\pi(t) = -t \sqrt[3]{\expect[\eps^3]}
+	-\pi(t, 0)^3 = \int t^3 \eps^3 dF(\eps)\\
+	\pi(t, 0) = -t \sqrt[3]{\expect[\eps^3]}
 \end{align}
-Note that if $\expect[\eps^3] = 0$ then $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} = 0$ while if $\expect[\eps^3] \neq 0$, $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} \neq 0$ which demonstrates that in general, the individual's attitude toward risk is neither of order 1 nor of order 2.
-So, if $\expect[\eps^3] \neq 0$, $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} = -\sqrt[3]{\expect[\eps^3]}$ exactly as Proposition \ref{derivative} claims.
+Note that if $\expect[\eps^3] = 0$ then $\frac{\partial \pi(t, 0)}{\partial t}\bigg|_{t=0} = 0$ while if $\expect[\eps^3] \neq 0$, $\frac{\partial \pi(t, 0)}{\partial t}\bigg|_{t=0} \neq 0$ which demonstrates that in general, the individual's attitude toward risk is neither of order 1 nor of order 2.
+So, if $\expect[\eps^3] \neq 0$, $\frac{\partial \pi(t, 0)}{\partial t}\bigg|_{t=0} = -\sqrt[3]{\expect[\eps^3]}$ exactly as Proposition \ref{derivative} claims.
 
@@
 A blue line represents the indifference curve.
-Suppose that the individual initially faces the lottery $(x_1, p; x_2, 1-p)$ but in addition is offered to buy $t \ge 0$ units of insurance each at a cost $\gamma$, which would pay him $\$ 1$ in (bad) state 1.
-A green line is the budget line which represents lotteries with insurance for all possible $t$.
+Suppose that the individual initially faces the lottery $(x_1, p; x_2, 1-p)$ but in addition is offered to buy $0 \le c \le \frac{x_2}{\gamma}$ units of insurance each at a cost $\gamma$, which would pay him $\$ 1$ in (bad) state 1.
+A green line is the budget line which represents lotteries with insurance for all possible $c$.
 An orange line represents the line of constant, zero expectation, that is, each point on this line represents the lottery with zero expected payoff.
@@
 \end{align*}
-where $\eps = (-(1-\gamma), p; \gamma, (1-p))$.These two conditions are exactly what Proposition \ref{LRA} requires for the individual with positive $u'''(0)$ being locally risk averse!
-\subsection{Ex. 2}
+where $\eps = (-(1-\gamma), p; \gamma, (1-p))$.These two conditions are exactly what Proposition \ref{LRA} requires for the individual with positive $u'''(0)$ being locally risk averse.
+\subsection{Example 2}
 Consider $$u(x) = \begin{cases}
@@
 Since $u''_{-}(0) = -2, u''_{+}(0)=2$, the second condition above is equivalent to \eqref{condition} requirement in Proposition \ref{LRA} for the individual being locally risk averse.
+\subsection{Example 3}
+Consider $$u(x) = \begin{cases}
+	x^3, &x \ge 0\\-x^2, &x < 0
+\end{cases}$$
+In this example, as in example 2, $u(x)$ is not differentiable at $x^*=0$. What is more interesting, here only left second derivative is non-zero at $x^*=0$.
+An indifference curve corresponding to the utility level 0 is given by:
+\begin{align*}
+	-px_1^2 + (1-p)x_2^3 = 0,\ \text{if } x_1 \le 0, x_2 \ge 0\\
+	px_1^3-(1-p)x_2^2 = 0,\ \text{if } x_1 > 0, x_2 < 0\\
+\end{align*}
+\begin{align}	\label{ic}
+	x_2 = \begin{cases} 
+	\sqrt[3]{\frac{p}{1-p}}\ |x_1|^{\frac{2}{3}}, &x_1\le0\\
+	-\sqrt{\frac{p}{1-p}}\ x_1^{\frac{3}{2}}, &x_1>0
+	\end{cases}
+\end{align}
+Here, $k=2$, $u^{''}_{+}(0) = 2, u^{''}_{-}(0) = 0$. From \eqref{ic} it follows
+\begin{align*}
+	\frac{dx_2}{dx_1}\bigg|_{x_1=0^-} = +\infty,\  \frac{dx_2}{dx_1}\bigg|_{x_1=0^+} = 0
+\end{align*}
+which corresponds to what Proposition \ref{slope} claims.
+Fig.
+\ref{fig2} demonstrates the point.
+Here, as in example 1, an individual strictly prefers full insurance with positive loading to his initial lottery $(x_1, x_2)$.
+At the same time, in contrast to example 1, here it is the case not for all $(x_1, x_2)$ which lies on the insurance budget constraint (green line) but only down to the point $(\bar{x}_1, \bar{x_2})$.
+That is, in terms of Proposition \ref{LRA}, an individual prefers a certain outcome to the small positive risk only for $t$ small enough.
+Note also, that condition \eqref{condition} of the Proposition \ref{LRA} holds, since $u''_{-}(0) = -2,\ \expect[\eps^2|\eps\ge0] > 0$, and $u''_{+}(0) = 0$, where $\eps = (-(1-\gamma), p; \gamma, (1-p))$.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{pic2}
+	\caption{}\label{fig2}
+\end{figure}
+It is worth mentioning also that in this example, since the left slope of the indifference curve at zero is infinite, even for very large positive loading $\gamma >> p$, there exists small enough $t$ such that the individual would prefer full insurance to a lottery with small positive expectation.
+
+Finally, note that, if instead we considered $u(x) = \begin{cases}
+	x^2, & x\ge0\\
+	x^3, & x <0
+\end{cases}$, it would be $\frac{dx_2}{dx_1}\bigg|_{x_1=0^-} = 0$.
+That means that the indifference curve would lie below the insurance budget constraint in the left neighborhood of $x^*=0$.
+Thus, in this case, no matter how small $t$ was, the individual would strictly prefer small positive risk to full insurance.
+So, Proposition \ref{LRA} ``does not work'' here.
+The reason for that is a violation of condition \eqref{condition} ($u''_{+}(0)\  \expect[\eps^2|\eps\ge0]\ \prob(\eps\ge0)>0$).
 \section{Conclusion}
